### 2023-03-13 docs/Manuscript/Manuscript.Rmd
first commit - analysis scripts + manuscript bibliographies & text

---
title: "Clinical prediction using similarity-based cohorts: A scoping review"
author:
  - name: Adi Cohen
    email: 
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Patti McCall
    email: 
    affiliation: Some Institute of Technology
  - name: Jason Cory Brunson
    email: jason.brunson@medicine.ufl.edu
    affiliation: Laboratory for Systems Medicine, University of Florida
    correspondingauthor: true
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Laboratory for Systems Medicine, University of Florida
    address: P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225
footnote:
  - code: 1
    text: "This is an illustrative author footnote."
abstract: |
  This is the abstract.

  It consists of two paragraphs.
keywords: 
  - case-based reasoning
  - localized modeling
  - nearest neighbors
journal: "Artificial Intelligence In Medicine"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib, SR-bibliography,
  SR-0-Found, SR-2-Included, SR-3-Tracked
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    latex_engine: xelatex
    citation_package: natbib
---

# Introduction

Clinical decision support (CDS) systems have been a fixture of artificial intelligence (AI) research since inception. As machine learning (ML) emerged, clinical modeling tasks became some of the most common and celebrated applications. A technique core to clinical applications of both AI and ML is _similarity-matching_, which relies on a _patient similarity measure_[^psm] to extract a cohort of past or training cases used to inform the diagnosis, prognosis, or care of a new or test case. Similarity-matching powers the retrieval step of most case-based reasoning (CBR) systems and provides the neighborhoods from which nearest neighbor (NN) models derive predictions. Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), most CBR tools return the retrieved cases to the user to inform human decision-making.

[^psm]: Similarity measures between more granular units of analysis, such as encounters and decision points, are often still referred to as patient similarity measures.

We discuss in this review a hybrid computational approach we call _localized modeling_ that borrows elements from both clinical application areas and appears to have been introduced independently several times. The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching. The idea is to use patient similarity not to obtain an ideally homogeneous population matched to the index case, as in CBR and NN prediction, but to obtain a relevant yet heterogeneous population, each member of which deviates slightly from the index case. As a result, the variation of any single factor will be (1) centered near its value on the index case and (2) most different from the index case in cases that in other respects are more similar to it. This makes interpretative models like GLMs better able to identify and measure the effects of determining factors. Localized modeling shows promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.

One memorable aspiration of clinical informatics was the “green button” that would, in response to clinical queries, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014. Localized modeling has recently been put forth as one answer to this appeal. Studies using localized models have been limited by computational cost—a naïve implementation involves fitting, optimizing, and evaluating a model on a separate population for each training case—and recent studies have described algorithms and implementations designed to reduce this cost. As greater access to health data, informatical knowledge, algorithmic efficiency, and computational power continue to spur innovation in CDS, we may expect localized modeling to become more accessible and efficient, its advantages better understood, and its disadvantages ameliorated.

The motivation for this scoping review is to describe the settings in and problems with which localized models have been tasked to date, and to synthesize existing methods into a general framework. Along the way, we attempt to reconcile varying terminology, summarize evaluation techniques, propose needed follow-up work, and recommend best practices.

## Related work

Clinical CBR emerged among rule-based approaches and other AI tools in the development of expert systems [@Aamodt1994]. Early implementations realized a general workflow described as “the four REs”, later the “R4 cycle” [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system retrieves one or more past cases (solved problems) from a corpus, reuses these to generate an understanding of (solution to) the new case, revises this understanding to better fit the new case (sometimes called “adaptation”), and retains the new case and its eventual understanding in the corpus to be retrieved and reused in future. In contrast to rule-based systems, which are variable-based and generally interpretable as a single rule applied to all new cases, case-based systems provide case-specific interpretations in the form of a number of more fully understood reference cases. While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.

@Kolodner1992 distinguished two styles of CBR: problem-solving, which is more procedural and used when objectives are more clearly defined, and interpretative, which provides categorizations and justifications for possible solutions. A similar dichotomy is commonly used to distinguish the performance criterion for the usefulness of predictive models from the interpretability criterion for their usability. In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort can be understood as an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a ML workflow. Accordingly, localized models have been primarily used for and evaluated on their ability to more accurately predict classes or outcomes. Nevertheless, as some more recent applications have begun to show, they can be used to draw individualized inferences about the importance of different risk factors. While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user. Provided the model family fitted to the cohorts is interpretable, the localized model as a whole inherits this property. Thus, localized models may embody both styles of CBR.

We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998a; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018]. While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis / dimension reduction / cluster analysis, risk evaluation / outcome prediction, clinical decision support / software tools). For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive modeling.

## Objectives

Our goals in this review are (1) to describe applications to date of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models. Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:

* Pulls case-level data from a corpus of past cases with known response values\newline
(response may be outcome, diagnosis, subtype, etc.)
* Defines a numeric multivariate case similarity measure\newline
(include integer-valued measures)
* Uses the similarity measure to select cohorts for index cases from the corpus\newline
(for example, based on a training–testing partition)
* Fits statistical models to cohorts from which to draw inferences about index cases\newline
(outcome predictions count, survival estimates, risk factor effect estimates, model evaluation statistics, etc.)

We expected the general framework to boil down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family. Such a general method would be straightforward, if laborious, to implement, and specializations of it would account for the majority of real-world applications. A general implementation based on this method would allow researchers to expedite every step of the analysis process, including selection, optimization, and evaluation, and enable sensitivity, robustness, and multiverse analyses that help identify the most consequential choices along the way.

# Methods

We modeled our literature review on PRISMA guidelines (PRISMA Group, 2021). Because we focus on methodologies rather than approaches, we deviated in some ways from these guidelines.

## PRISMA-S checklist

The PRISMA-S checklist comprises 16 items categorized into information sources and methods, search strategies, peer review, and managing records. We use the checklist to report the steps of our systematic review. Some items of the PRISMA-S checklist are intended for meta analysis and, therefore, did not apply to this study. We followed the checklist throughout the search process and explain the deviations from it in [Supplemental Table 1](https://docs.google.com/spreadsheets/d/1lKI5WWGIzo_hE9olkWWVCbLjCY9To5fiKPiefXUqgoU/edit).

## Search

We derived the eligibility criteria itemized in the Introduction from a seed set of previously read studies. Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and Mathematical Reviews. We finalized the PubMed search first, then adapted it to the other platforms. Through each platform, we searched those databases included by default.[^databases]

[^databases]: This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts & Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.

The structure and terms of our search strings were inspired in part by previous reviews adjacent to our topic of interest [@Sharafoddini2017; @Parimbelli2018]. We organized the PubMed search in disjunctive normal form (an OR of ANDs). Following the solidification of an outline, we expanded each term to include synonyms that are similar enough to be applicable to our search. We then evaluated the expanded search string using the PubMed Advanced Search platform. We initially included each term and their synonyms separately to evaluate what resulted. We pruned several terms that returned no results (“phrases not found”) or to reduce the number of results. At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.

We conducted all searches over 2021 July 19–22. We organized the results into a Zotero collection with one subcollection for each database. We then imported the pooled results to Covidence, which merged some duplicate entries.

## Title/Abstract screen

Within Covidence, we screened titles and abstracts for relevance. We decided on four eligibility criteria to expedite the screening process. These include articles written in English for readability, original studies to strictly include primary sources with new information/conclusions, medical or clinical related settings due to the applicability of our data to this field, and alternative use of the key terms within the search string, meaning that it was clear from the title and abstract that the entry was included due to the use of one or more search terms with a different meaning than ours. Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, an entry would be included.

## Full text review

We set out 4 criteria for full-text review to restrict to studies that used some form of localized modeling on health data:

* Pulls case-level data from a corpus of past cases with known classes or outcomes
* Defines a numeric multivariate case similarity measure
* Uses the similarity measure to select cohorts for index cases from the corpus
* Fits statistical models to cohorts to make inferences about index cases

Note that classical CBR satisfies the first and third criteria by definition and in most cases will satisfy the second.

In Covidence, two authors (AC and JCB) independently evaluated each manuscript for the four eligibility criteria. The first criterion that a manuscript failed was designated the reason for exclusion. An antecedent criterion was used to exclude manuscripts that did not report the results of original studies involving real-world experiments or empirical data, for example surveys of prior work and proposals of frameworks. In cases of disagreement between the authors on the reason for exclusion, the first criterion was adopted. There were 60 disagreements over inclusion versus exclusion. The authors arrived at agreement on each manuscript through discussion.

Finally, one author (JCB) applied the same review process to the seed set that inspired the review. This obtained 6 additional studies that had not been retrieved by any of our database searches [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019].

## Taxonomy of included studies

We noticed during full text review that our inclusion criteria admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the responses of the cohort. Mean and plurality can be understood as statistical models, and other studies used equally simple though less conventional calculations. However, a review of all studies that use nearest neighbors prediction, even in the clinical domain, would be impractical. Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to analyze only those studies whose approach was not equivalent to nearest neighbors prediction using a conventional similarity measure. The unanalyzed studies comprised 25 of the 44 studies included from the search and 1 of those from the seed set, and our analysis focuses on the remaining 19 + 1 + 5 = 25 studies.

## Coding

Following the determination of 25 sources aligning with our systematic review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)). The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach). We used these data to detect and visualize trends amongst the included studies.

## Synthesis

Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize possibly independent research efforts that have converged on a common modeling technique. The goal will be to describe a unified framework for localized models that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.

# Results

## Selection

Figure \ref{fig:prisma} depicts our identification of studies via databases and registers. Following the completion and input of each search string, there were a total of 1,422 sources within all of the platforms used. De-duplication resulted in 1,309 entries, which were added to the title/abstract screening for review. Of these, 328 entries met the screening criteria and were assessed through full text review. Of these, 44 fit the criteria and were considered reports of included studies. Exclusion of NN models left 19 studies being included in the method synthesis.

```{r fig:prisma, fig.align='center', out.width="80%", fig.cap = "\\label{fig:prisma}PRISMA-S flow chart.", echo=FALSE}
knitr::include_graphics("../../fig/fig-prisma.pdf")
```

We then reviewed 43 studies that comprised a seed set that inspired this review. After removing duplicates and screening for eligibility, we were left with 6 additional studies, 1 of which was excluded from the synthesis for using nearest neighbors prediction. Finally, citations in reference lists from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010]. This left us with 51 studies reviewed and 25 included in the synthesis.

Inter-rater reliability was low, at 82% relative to a 72% probability of random agreement. This was due, in parts, to different interpretations of some eligibility criteria by the authors, inconsistent terminology across the sample, and incomplete reporting of resources and methods in the sample. Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning. We will discuss the different domains, terminologies, and reporting issues of the sample in later sections.

## Overview

The 25 included studies were analyzed based on 20 characteristics (Table \ref{tab:synthesis}), and we describe some observations here. The studies were published in a variety of journals, with some of greater frequency, though no single journal published exceptionally many. The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, and _Hindawi - Journal of Healthcare Engineering_ published 3. Another distinguishing factor was the years of publication of the included studies (Figure \ref{fig:year}). While they stretch back to the late 1990s, they have become more common in recent years, which suggests that this is still an active, though not rapidly expanding methodological approach. We note that, while CBR in health and medicine originated as early as 1990, these studies did emerge early on once CBR had established itself. So the idea has been "in the air" for as long as CBR has been in use.

```{r tab:synthesis, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
googlesheets4::read_sheet(
  ss = "1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo", sheet = 1L
) |>
  rename_with(snakecase::to_snake_case) |>
  arrange(date_of_publication) |> 
  transmute(
    Citation = paste0("@", zotero_key),
    Date = format(date_of_publication, "%Y %b %d"),
    Task = objective,
    Source = type_of_data,
    Type = range_of_data,
    `Cases` = number_of_cases_incidents,
    `Features` = number_of_predictors_features
  ) |>
  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) |> 
  mutate(across(
    c(`Cases`, `Features`),
    sapply, FUN = paste0, collapse = "; "
  )) |> 
  mutate(across(
    everything(),
    # ~ stringr::str_replace_all(., "<U+001B>", "")
    textclean::replace_non_ascii
  )) ->
  tab_synthesis
tab_synthesis |> 
  knitr::kable(
    align = "lllllrr",
    caption = paste0(
      "\\label{tab:synthesis}",
      "Studies included in the method synthesis.",
      " 'Date' is the earliest the study is known to have been public."
    )
  )
```

```{r fig:year, fig.align='center', out.width="80%", fig.cap = "\\label{fig:year}Number of publications in our sample each year.", echo=FALSE}
knitr::include_graphics("../../fig/fig-years.png")
```

Another dominant characteristic of included studies is their remit. We categorized studies as “science” or “quality improvement” based on whether their aim was to produce generalizable knowledge or to improve practice. We classified 7 of the 25 studies as “science”, making the majority of studies from a “quality improvement” standpoint. These studies aimed primarily to provide providers with tools to use in clinic or to improve outcomes for patients currently receiving care. Lastly, we observed a commonality among 6 of the 25 included studies in having evaluated methods using leave-one-out cross-validation (LOOCV). The purpose of this measure is to estimate the overall performance of certain factors when used to make predictions, particularly utilized on smaller data sets, where models benefit greatly from larger training sets and additional model fitting is less costly.

In the Discussion, we discuss trends in the literature we reviewed, identify common limitations and gaps in understanding, and attempt to synthesize their methods into a general framework.

# Discussion

## Trends

We draw several observations from our encoding of the included studies. We first focus on two driving questions: What challenges did investigators set out to solve? How did their approaches meet these challenges? We then comment on the connectedness of the included studies and the potential for recombination.
See Table \ref{tab:composite} for details on approaches taken by the synthesized studies.

(See [this sheet](https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/).)

```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
googlesheets4::read_sheet(
  ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 1L
) |>
  rename_with(snakecase::to_snake_case) |>
  transmute(
    Citation = paste0("@", zotero_key),
    Elements = str_remove_all(approach_type_s, "\\[|\\]"),
    Terminology = term_s#,
    # Outcomes = outcome
  ) |> 
  mutate(Elements = str_replace_all(Elements, ",", ";")) |> 
  left_join(select(tab_synthesis, Citation, Date), by = "Citation") |> 
  arrange(Date) |> 
  select(-Date) |> 
  mutate(across(
    everything(),
    # ~ str_replace_all(., "<U+001B>", "")
    textclean::replace_non_ascii
  )) |>
  knitr::kable(
    align = "lllllrr",
    caption = paste0(
      "\\label{tab:composite}",
      "Methodological elements of studies included in the synthesis. "
    )
  )
```

### Application domain

While all included studies were reported in scientific and medical journals, the vast majority were oriented toward clinical practice rather than medical research. For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006]. This was part of a much larger literature on CBR systems that was included here despite relying on NN prediction because it used a non-traditional voting scheme to generate recommendations. A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021a]. In both cases a prototype implementation was deployed in an experimental setting for evaluation.

The most common clinical motivations were individualized detection or diagnosis, prognosis or outcome prediction, and treatment or therapy recommendation. The plurality focused on prognosis or outcome prediction, often using time-to-event analysis: @Mariuzzi1997 used CBR to predict survival time from several geometric properties of breast tumors. @Lowsky2013 used CBR with non-parametric survival models on registry data to predict patient–graft survival times following kidney transplantation. @Lee2015 and @Lee2017 used localized logistic regression and random forest modeling to predict 30-day mortality following discharge for ICU patients. @Vilhena2016a developed a CBR cycle around a clustering-informed similarity matching procedure and an artificial neural network–based classifier to identify thrombophilia patients at high risk of thrombotic episodes. And @Ma2020a took a similarity cohort–based approach to predicting length of stay for ICU patients. Also of note, from a public health perspective, @Xu2008a used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.

Toward detection and diagnosis, @Wyns2004 proposed a hybrid neural net–CBR system to distinguish (with confidence bounds) arthritic versus control patients, based on several histological features. @Nicolas2014a used collaborative multilabel CBR to subtype melanoma patients based on confocal and dermoscopy images. @Wang2019 used localized models built from a multi-type additive similarity measure to distinguish type 2 diabetic versus control populations. Along the way, @Song2006c, @Kasabov2010, and @Verma2015a took an iterative model-building approach to several tasks: predicting glomerular filtration rate, a key indicator of renal function, from demographic and physiological variables; identifying patients with colon cancer using a large number of gene expression measurements; and identifying patients with type 2 diabetes based on demographic, physiological, and genetic variables.

While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking “decision points” as their units of analysis, @Tang2021d and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.

A partial exception to this focus was a 2014 study that also reported a decision support tool, in this case for early diagnosis of melanoma from clinical data and dermoscopy images [@Nicolas2014a]. While the stated objectives were analogous, specific emphasis was placed on the acquisition of new knowledge through the development of the tool, including the systematic generation of new data and creation of a clinical ontology. This study was included for its use of a collaborative classifier that drew from multiple modeling approaches. In keeping with this focus of the included literature, the stated objectives of the proposed methods were more often (or additionally) to predict outcomes or to recommend interventions than only to diagnose disease.

### Method development

Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), laboratory (2), and image (1) data. We suggest three reasons for this: First, this literature traces back to the 1990s, before -omic data could be generated cost-effectively at scale. Second, CBR in particular has a strong tradition in clinical decision support, where the focus of our sample remains throughout the review period. Third, because most -omic data are highly homogeneous—all measurements are made along or can be transformed to a common scale, e.g. greyscale pixellations for X-ray images and transcripts per million for RNA-seq data—more deeply theoretical analysis techniques have been developed and come into wide use. While variations on correlation-based approaches like EHR-based phenome-wide association studies and similarity-based methods like CBR itself have been developed, more mechanistic and probabilistic tools have not become a domain standard.

CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as “knowledge-light”, in that “the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases”. This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle. Several studies addressed these challenges: @Park2006, while excluded from the synthesis, were the earliest to propose that cohorts be bounded by a similarity threshold rather than by a number of cases, and this improved performance in their experiments. @CampilloGimenez2013 leveraged predictor weights obtained from logistic regression to inform the similarity calculation used in retrieval. @Lowsky2013 used non-parametric models to reduce the computational burden of revision (prediction). @Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time (“just-in-time learning”), and @Ng2021b and @Tang2021d partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.

Other studies addressed limitations of available tools. @Lopez2011 implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake. Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018a].

Because we excluded studies that used conventional NN prediction, many included studies reported new approaches to adaptation subsequent to retrieval. Very few proposed novel similarity measures, possibly because larger studies tended to be reported in separate articles detailing experiments with specific components of the process. However, this also suggests that few experimental studies have focused on the unified development of new retrieval and adaptation strategies.
We also note that the majority of studies evaluated and compared methods using leave-one-out cross-validation (LOOCV). This is an especially appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer. This likely follows from the older age of many included studies and from their consistent primary focus on clinical data, which is more costly to collect and comes with more restrictions on its use.

### Coherence

The motivational and methodological unity of these studies does not reflect a unified research program. Besides the lack of any primary journal of record, we observed collaborations only among the authors of smaller contiguous programs, including applications of the TWNFI methodology [@Song2006c; @Verma2015a], individualized mortality prediction for ICU patients [@Lee2015; @Lee2017] and the use of more explanatory models to prioritize predictors or treatments for chronic disease [@Ng2015; @Tang2021d; @Ng2021b].

These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as “targeted prognosis” [@Mariuzzi1997], “transductive inference” [@Song2006c], “personalized decision support” [@Lee2015], “personalized (predictive) models” [@Liang2015, in contrast to “local models”; @Ng2015; @Wang2019; @Ma2020a], and “precision cohort” analysis [@Wang2019; @Tang2021d; @Ng2021b]. We find uses of “targeted”, “personalized”, and “precision” generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML. Because local naturally contrasts with global, we propose “localized models” as a suitable term for this counterpart to globally-fitted models.

## Synthesis

The 25 studies included in our synthesis used composite techniques that we organized into two major types. One type, similarity learning, was used in 8 studies that tie in to the much larger literature on metric learning [@Yang2006]. These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure. Seven studies [@] used response values in training data sets to supervise similarity learning while one [@Wyns2004] used unsupervised learning. Based on the key properties of metric learning algorithms identified by @Bellet2014, most learned measures were non-linear while some were local, and some were optimized globally while others locally. However, not all qualified as metric learning, since the resulting measures would not necessarily satisfy the triangle inequality. Each then used its similarity measure to retrieve training cases relevant to each testing case. Once retrieved, most took a standard NN approach to generating predictions from the cohort; two exceptions [@Vilhena2016a; @Tang2021d] proceeded to fit predictive models to the retrieved cohorts and are discussed below.

### Similarity learning and threshold optimization

The similarity learning approaches took a variety of forms. Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007]. In particular, @Yearwood1997a defined similarity as a weighted sum of differences in predictor values and used linear regression to optimize the weights for the predictive accuracy of the cohort they retrieve. Others used supervised learning: @Song2006c proposed an iterative algorithm to optimize a set of fuzzy inference rules used to retrieve relevant cases and generate a prediction, using back-propagation on the rules’ parameters. Their method was used in several later studies returned by our search, which are not discussed here because they did not originate the technique. @Nicolas2014a explicitly appeal to a supervised distance metric learning technique [@Xing2002] to obtain a similarity measure on their set of dermoscopy and confocal images that most effectively separates malignant from benign melanoma tumors. We will discuss the remaining uses of supervised similarity learning shortly.

Distinct from but related to similarity learning was supervised cohort construction. This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called “statistical CBR”. The step used a heuristic procedure to locate a similarity threshold that (locally) maximizes predictive accuracy on the training set. This is analogous to optimizing the neighborhood size parameter in NN predictive modeling, so we will not discuss it further except to mention that the authors found statistical CBR to outperform conventional CBR on several data sets. @CampilloGimenez2012a combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the NN model using the previously optimized weights.

As noted above, these instances of similarity learning fit into a much larger literature on metric learning, which has seen widespread use in health informatics and modeling [@]. That these studies satisfied our review criteria is due to the novelty of the techniques at the time of publication and the detailed attention paid to their techniques by the authors. It also reflects a limitation of our study and of the literature we set out to retrieve: We are aware of no standard terms in use to identify the family of techniques that fit statistical models to similarity-based cohorts. The retrieved studies whose use of such techniques alone satisfied our criteria [@] variably termed them individualized models, personalized models, etc. (list all terms). We prefer the term localized models, which is both sensitive and specific to uses of a similarity or distance measure to retrieve a cohort to which a model is fit: It includes some properly included techniques we did not have in mind [@Vilhena2016a] yet excludes other techniques commonly referred to as individualized, personalized, or precision.

### Localized models

In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016a, @Tang2021d, and @Ng2021b, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015a, @Wang2019, and @Ma2020a. In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021d; @Ng2021b] later. The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016a] or via cross-validation [@Lowsky2013; @Verma2015a]. Further in the former direction, one study [@Ma2020a] set a fixed cohort size while two [@Liang2015; @Tang2021d; @Ng2021b] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.

Several themes emerged from this sample. Foremost was the optimization of retrieved cohort sizes for some measure of model performance. Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021d; @Ng2021b] in this corpus took this approach. Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.

An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015. More recently, @Tang2021d and @Ng2021b proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance. Notably, though their similarity measure was supervised, this optimization process was not.

With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population. This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021d; @Ng2021b]. These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].

### A general framework

This relatively small sample exhibited a wide range of approaches, both technically and conceptually. Discussion of the technical diversity of these approaches has been published in previous reviews [@], and we focus here on the conceptual. The design of a localized modeling approach can be decomposed into three largely independent choices: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort. The choices are the _similarity measure_, the _retrieval step_, and the _statistical model_, respectively. These are diagrammed in Figure X.

```{r fig:framework, fig.align='center', out.width="80%", fig.cap = "\\label{fig:framework}General framework for localized models.", echo=FALSE}
knitr::include_graphics("../../fig/fig-typology.pdf")
```

Each step can be unsupervised supervised: Unsupervised similarity measures include discrete measures like the Levenshtein distance and continuous measures like cosine similarity, while supervised measures include the Mahalanobis distance and random forest proximity as well as several composite measures with weights calculated from the data. Most retrieval steps obtained cohorts with a uniform size (cardinality) or similarity threshold, and the remaining were likewise only informed by predictors; thus, while in principle a retrieval step could be supervised, in this sample none were. Almost all models were supervised, being that most performed predictive tasks, though @Mariuzzi1997 modeled only survival rates within localized cohorts. Meanwhile, each step can be optimized in a ML fashion by having its parameters tuned to improve performance. We found no studies that tuned the calculation of similarity, though most uniform cohort sizes were tuned. Some earlier studies tuned generalized regression models, but more recent studies have not.

## Limitations

## Conclusions

We propose the term _localized modeling_ to encompass an approach derived from CBR in which parameterized models are fitted in a standardized way to nearest neighborhoods of past or training cases according to a measure of patient similarity. We conducted a systematic search for studies that apply localized models to tasks involving health data and synthesized these largely independently developed approaches into a general framework. While the search was limited by low inter-rater reliability and failure to recover several motivating examples, the included studies used many of the same underlying/fundamental tools to build, optimize, and evaluate their methods. We therefore believe that our framework can serve to taxonomize ongoing work of this type and inform the development of flexible/customizable implementations. Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth. Whereas precious few of the reviewed studies used these models, for any task other than prediction, despite widespread suspicion among clinicians of “black-box” models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.

# References {-}

### 2023-03-13 scr/sr-journals.r

@@ -1,0 +1,15 @@
+library(tidyverse)
+library(readxl)
+
+here::here("data/SR-journals.xlsx") %>%
+  read_xlsx() %>%
+  mutate(journal = str_remove(journal, "^[0-9]{4} *")) %>%
+  mutate(journal = str_remove(journal, "^[0-9]{1,2}[a-z]{2} *")) %>%
+  mutate(journal = str_remove(journal, " *\\([A-Z0-9]+\\)$")) %>%
+  mutate(journal = tolower(journal)) %>%
+  group_by(journal) %>%
+  count() %>%
+  drop_na() %>%
+  arrange(desc(n)) %>%
+  filter(n > 2L) %>%
+  print(n = Inf)

### 2023-03-13 scr/synthesis.r

@@ -1,0 +1,75 @@
+library(tidyverse)
+
+# Excel file downloaded from Google Sheet:
+# https://docs.google.com/spreadsheets/d/
+# 1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/
+here::here("data/IMSR composite techniques.xlsx") %>%
+  readxl::read_xlsx() %>%
+  select(seq(3L)) %>%
+  rename_with(~ str_replace(snakecase::to_snake_case(.), "_s$", "s")) %>%
+  mutate(across(c(approach_types, terms), ~ str_split(., "(;|,) *"))) %>%
+  print() -> composite_studies
+
+# years of publication
+composite_studies %>%
+  transmute(year = as.integer(str_remove(citation, "^.*, "))) %>%
+  ggplot(aes(x = year)) +
+  geom_bar() +
+  theme_bw() +
+  theme(
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.minor.y = element_blank()
+  ) +
+  labs(x = NULL, y = NULL) ->
+  year_plot
+ggsave(here::here("fig/fig-years.png"), year_plot, width = 5, height = 3)
+
+# approach types
+composite_studies %>%
+  select(citation, approach_types) %>%
+  unnest(approach_types) %>%
+  mutate(approach_types = str_remove_all(approach_types, "^\\[|\\]$")) %>%
+  print() -> composite_approaches
+# approaches to ignore (check whether each study has one unignored)
+drop_types <- c(
+  "NN?", "unclear",
+  "interactive implementation",
+  "similarity matching", "whole-population weight learning"
+)
+# frequency table
+composite_approaches %>%
+  count(approach_types) %>%
+  arrange(desc(n)) %>%
+  filter(! approach_types %in% drop_types) %>%
+  print(n = Inf)
+
+# similarity learning
+composite_approaches %>%
+  filter(str_detect(approach_types, "similarity learning")) %>%
+  select(citation) %>%
+  inner_join(composite_approaches, by = "citation") %>%
+  filter(! str_detect(approach_types, "similarity learning"))
+# modeling on similarity cohorts
+composite_approaches %>%
+  filter(str_detect(approach_types, "modeling on similarity cohorts")) %>%
+  select(citation) %>%
+  inner_join(composite_approaches, by = "citation") %>%
[…]

### 2023-03-17 docs/Manuscript/Manuscript.Rmd
add cohen details

@@
   - name: Adi Cohen
-    email: 
-    affiliation: Some Institute of Technology
-    footnote: 1
+    email: ac4569@my.nsu.edu
+    affiliation: 1
   - name: Patti McCall
     email: 
-    affiliation: Some Institute of Technology
+    affiliation: 2
   - name: Jason Cory Brunson
     email: jason.brunson@medicine.ufl.edu
-    affiliation: Laboratory for Systems Medicine, University of Florida
+    affiliation: 3
     correspondingauthor: true
 address:
-  - code: Some Institute of Technology
+  - code: 1
+    address: Nova Southeastern University College of Osteopathic Medicine, 3200 S University Drive, Davie, FL 33328
+  - code: 2
     address: Department, Street, City, State, Zip
-  - code: Laboratory for Systems Medicine, University of Florida
-    address: P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225
-footnote:
-  - code: 1
-    text: "This is an illustrative author footnote."
+  - code: 3
+    address: Laboratory for Systems Medicine, University of Florida, P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225
 abstract: |
@@
   - nearest neighbors
+  - clinical decision support
 journal: "Artificial Intelligence In Medicine"
@@
 classoption: preprint, 3p, authoryear
-bibliography: mybibfile.bib, SR-bibliography,
+bibliography: SR-bibliography,
   SR-0-Found, SR-2-Included, SR-3-Tracked
@@
 
-One memorable aspiration of clinical informatics was the “green button” that would, in response to clinical queries, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014.
+One memorable aspiration of clinical informatics was the "green button" that would, in response to clinical queries, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014.
 Localized modeling has recently been put forth as one answer to this appeal.
@@
 Clinical CBR emerged among rule-based approaches and other AI tools in the development of expert systems [@Aamodt1994].
-Early implementations realized a general workflow described as “the four REs”, later the “R4 cycle” [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system retrieves one or more past cases (solved problems) from a corpus, reuses these to generate an understanding of (solution to) the new case, revises this understanding to better fit the new case (sometimes called “adaptation”), and retains the new case and its eventual understanding in the corpus to be retrieved and reused in future.
+Early implementations realized a general workflow described as "the four REs", later the "R4 cycle" [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system retrieves one or more past cases (solved problems) from a corpus, reuses these to generate an understanding of (solution to) the new case, revises this understanding to better fit the new case (sometimes called "adaptation"), and retains the new case and its eventual understanding in the corpus to be retrieved and reused in future.
 In contrast to rule-based systems, which are variable-based and generally interpretable as a single rule applied to all new cases, case-based systems provide case-specific interpretations in the form of a number of more fully understood reference cases.
@@
 We derived the eligibility criteria itemized in the Introduction from a seed set of previously read studies.
-Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and Mathematical Reviews.
+Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
 We finalized the PubMed search first, then adapted it to the other platforms.
@@
 We initially included each term and their synonyms separately to evaluate what resulted.
-We pruned several terms that returned no results (“phrases not found”) or to reduce the number of results.
+We pruned several terms that returned no results ("phrases not found") or to reduce the number of results.
 At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.
@@
 Another dominant characteristic of included studies is their remit.
-We categorized studies as “science” or “quality improvement” based on whether their aim was to produce generalizable knowledge or to improve practice.
-We classified 7 of the 25 studies as “science”, making the majority of studies from a “quality improvement” standpoint.
+We categorized studies as "science" or "quality improvement" based on whether their aim was to produce generalizable knowledge or to improve practice.
+We classified 7 of the 25 studies as "science", making the majority of studies from a "quality improvement" standpoint.
 These studies aimed primarily to provide providers with tools to use in clinic or to improve outcomes for patients currently receiving care.
@@
 
-While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking “decision points” as their units of analysis, @Tang2021d and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
+While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking "decision points" as their units of analysis, @Tang2021d and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
 
@@
 
-CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as “knowledge-light”, in that “the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases”.
+CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
 This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle.
@@
 @Lowsky2013 used non-parametric models to reduce the computational burden of revision (prediction).
-@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time (“just-in-time learning”), and @Ng2021b and @Tang2021d partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
+@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning"), and @Ng2021b and @Tang2021d partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
 
@@
 
-These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as “targeted prognosis” [@Mariuzzi1997], “transductive inference” [@Song2006c], “personalized decision support” [@Lee2015], “personalized (predictive) models” [@Liang2015, in contrast to “local models”; @Ng2015; @Wang2019; @Ma2020a], and “precision cohort” analysis [@Wang2019; @Tang2021d; @Ng2021b].
-We find uses of “targeted”, “personalized”, and “precision” generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
-Because local naturally contrasts with global, we propose “localized models” as a suitable term for this counterpart to globally-fitted models.
+These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021d; @Ng2021b].
+We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
+Because local naturally contrasts with global, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.
 
@@
 Distinct from but related to similarity learning was supervised cohort construction.
-This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called “statistical CBR”.
+This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called "statistical CBR".
 The step used a heuristic procedure to locate a similarity threshold that (locally) maximizes predictive accuracy on the training set.
@@
 Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth.
-Whereas precious few of the reviewed studies used these models, for any task other than prediction, despite widespread suspicion among clinicians of “black-box” models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+Whereas precious few of the reviewed studies used these models, for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+
+## Appendix
+
+### Search strings
+
+Here we reproduce the search strings and platform specifications used in our literature search.
+We first finalized the **PubMed** search string below:
+
+```
+(
+  "case-based reasoning" [All Fields] OR
+  "case-based system" [All Fields]
+) OR (
+  "individualized modeling" [All Fields] OR
+  "personalized modeling" [All Fields] OR
+  "customized modeling" [All Fields]
+) OR (
+  "individualized cohort" [All Fields]
+) OR (
+  (
+    "patient similarity" [All Fields] OR
+    "patient distance" [All Fields] OR
+    "patient connection" [All Fields] OR
+    "patient affinity" [All Fields] OR
+    "patient clustering" [All Fields]
+  ) AND (
+    "cohort study" [All Fields]
+  )
+)
+```
+
+This search yielded 423 results.
+
+We then generated analogous search strings or search strategies for the Web of Science, Academic Search Premier, and MathSciNet platforms, based on the logics and syntaxes of their respective interfaces.
+
+For **Web of Science**, we searched for several separate disjunctions derived from the PubMed search string.
+The separate search strings and the number of results obtained using each are below.
+
+------------------------------------------------
+Search string                    Number of items
+------------------------------- ----------------
+`"case-based reasoning" OR`                     
+`"case-based system"`                      3,636
+
+`"individualized model" OR`                  444
+`"individualized modeling" OR`                  
+`"personalized model" OR`                    
+`"personalized modeling" OR`                    
+`"customized model"`                         
+`"customized modeling"`
+
+`"individualized cohort"`                      1
+
+`"patient similarity" OR`                     48
+`"patient distance" OR`                         
+`"patient connection" OR`                       
+`"patient affinity" OR`                         
+`"patient clustering"`                          
+Refined search: `"cohort"`                      
+------------------------------------------------
+
+We found the results of the first search string to be predominantly irrelevant. To reduce review time, these were dropped.
+The last search was refined with an additional term following the initial disjunctive search.
+Our searches on Web of Science thus yielded 493 results.
+
+When searching **Academic Search Premier**, we checked the option "Scholarly (Peer-Reviewed) Journals", unchecked the option "Apply equivalent subjects", and searched for several separate disjunctions and conjunctions of strings in the "TX All Text" field.
+We obtained the resulting citations via email in RIS format.
+
+------------------------------------------------
+Search string                    Number of items
+------------------------------- ----------------
+`"case-based reasoning" OR`                     
+`"case-based system"`                      3,283
+
+`"individualized modeling" OR`               100
+`"personalized modeling" OR`                    
+`"customized modeling"`
+
+`"individualized cohort"`                      2
+
+`"patient similarity" AND`                    14
+`"cohort study"`
+
+`"patient distance" AND`                      15
+`"cohort study"`
+
+`"patient connection" AND`                     6
+`"cohort study"`
+
+`"patient affinity" AND`                       0
+`"cohort study"`
+
+`"patient clustering" AND`                    50
+`"cohort study"`                                
+------------------------------------------------
+
+We found the results of the first search string to be predominantly irrelevant. To reduce review time, these were dropped.
+Our searches on Academic Search Premier thus yielded 187 results.
+
+For **MathSciNet**, the portal to the _Mathematical Reviews_ database, we used the same separate searches as for Web of Science, in some cases expanded to obtain more results.
+Those which yielded nonzero numbers of results are below:
+
+------------------------------------------------
+Search string                    Number of items
+------------------------------- ----------------
+`"case-based reasoning"`                     110
+
+`"individualized model`                        2
+
+`"patient similarity" AND`                     1
+`"cohort"`                                
+------------------------------------------------
+
+Our searches of _Mathematical Reviews_ therefore yielded 113 results.
+
+These totaled 1,422 sources from all platforms. We organized the full results in a Zotero collection alongside the seed set and created a single folder for the 25 results reviewed in detail.
 

### 2023-03-18 docs/Manuscript/Manuscript.Rmd
fill in citations + draft abstract

@@
 abstract: |
-  This is the abstract.
+  Background
 
-  It consists of two paragraphs.
+  Applications of classical case-based reasoning (CBR) to clinical and health tasks have given rise to a family of techniques in which an (often conventional) model is fitted to a subset of labeled cases, selected for relevance or similarity to an unlabeled case to which the model is applied.
+These localized models, as we propose to call them, may perform as well as "black-box" models while retaining direct interpretations.
+Their adoption has been hindered by fragmented development and implementation challenges.
+Our goal in this scoping review was to describe the clinical and health applications of localized models to date and to derive a general framework for their design and evaluation.
+
+  Methods
+
+  We searched five electronic bibliographic databases for publications that report applications of localized models to tasks involving analysis of clinical or health data.
+Two authors divided title/abstract screening and collaboratively reviewed screened entries for inclusion.
+We also reviewed a seed set that inspired the review and traced a subset of citations from  We discussed and summarized settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
+
+  Results
+
+  Of 1,309 search results, 328 were reviewed.
+Of these, combined with 43 publications that seeded the review and 1 obtained by citation tracking, 51 were included in the review and 25 of these in the methodological synthesis.
+The specificity of several search terms was poor, and inter-rater reliability was low.
+Entries were published from 1997 to 2021, with half of these since 2015.
+The most common tasks were prognosis and diagnosis, with only few recent studies focusing on care recommendation.
+The studies overwhelmingly focused on clinical data, occasionally also including laboratory and image-derived data.
+Only three studies used localized models for any task other than prediction, most recently to identify significant determinants behind recommendation outputs.
+Several studies commented on memory and runtime costs, and four proposed partial solutions to them.
+While most techniques were not replicated and did not cite predecessors, a general technique that specializes to most of those encountered was straightforward to describe.
+
+  Conclusions
+
+  Localized models hold promise to improve the performance of models used in clinical decision support tools while maintaining interpretability, provided their computational limitations can be overcome.
+Our search undoubtedly failed to capture all studies that used methods of this kind, but those reviewed and synthesized exhibit consistent enough design to be reproduced by a general-purpose tool.
+We hope that our review and synthesis will assist the development of more efficient implementations suitable for integration into practice.
 keywords: 
@@
 
-Finally, one author (JCB) applied the same review process to the seed set that inspired the review.
+Finally, one author (JCB) applied the same review process to the seed set of 43 studies that inspired the review.
 This obtained 6 additional studies that had not been retrieved by any of our database searches [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019].
@@
 
-The most common clinical motivations were individualized detection or diagnosis, prognosis or outcome prediction, and treatment or therapy recommendation.
+The most common clinical motivations were individualized detection or diagnosis, prognosis or outcome prediction, and treatment or care recommendation.
 The plurality focused on prognosis or outcome prediction, often using time-to-event analysis: @Mariuzzi1997 used CBR to predict survival time from several geometric properties of breast tumors.
@@
 These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
-Seven studies [@] used response values in training data sets to supervise similarity learning while one [@Wyns2004] used unsupervised learning.
+Seven studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
 Based on the key properties of metric learning algorithms identified by @Bellet2014, most learned measures were non-linear while some were local, and some were optimized globally while others locally.
@@
 
-As noted above, these instances of similarity learning fit into a much larger literature on metric learning, which has seen widespread use in health informatics and modeling [@].
+As noted above, these instances of similarity learning fit into a much larger literature on metric learning, which has seen widespread use in health informatics and modeling.
 That these studies satisfied our review criteria is due to the novelty of the techniques at the time of publication and the detailed attention paid to their techniques by the authors.
 It also reflects a limitation of our study and of the literature we set out to retrieve: We are aware of no standard terms in use to identify the family of techniques that fit statistical models to similarity-based cohorts.
-The retrieved studies whose use of such techniques alone satisfied our criteria [@] variably termed them individualized models, personalized models, etc.
-(list all terms).
+The retrieved studies whose use of such techniques alone satisfied our criteria variably termed them individualized, personalized, local, and patient-specific models, among other terms (Table \ref{tab:composite}).
 We prefer the term localized models, which is both sensitive and specific to uses of a similarity or distance measure to retrieve a cohort to which a model is fit: It includes some properly included techniques we did not have in mind [@Vilhena2016a] yet excludes other techniques commonly referred to as individualized, personalized, or precision.
@@
 This relatively small sample exhibited a wide range of approaches, both technically and conceptually.
-Discussion of the technical diversity of these approaches has been published in previous reviews [@], and we focus here on the conceptual.
+Discussion of the technical diversity of these approaches has been summarized above and published in greater detail in previous reviews [@Choudhury2016; @Sharafoddini2017; @Parimbelli2018], and we focus here on the conceptual.
 The design of a localized modeling approach can be decomposed into three largely independent choices: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
 The choices are the _similarity measure_, the _retrieval step_, and the _statistical model_, respectively.
-These are diagrammed in Figure X.
+These are diagrammed in Figure \ref{fig:framework}.
 
@@
 We conducted a systematic search for studies that apply localized models to tasks involving health data and synthesized these largely independently developed approaches into a general framework.
-While the search was limited by low inter-rater reliability and failure to recover several motivating examples, the included studies used many of the same underlying/fundamental tools to build, optimize, and evaluate their methods.
-We therefore believe that our framework can serve to taxonomize ongoing work of this type and inform the development of flexible/customizable implementations.
+While the search was limited by low inter-rater reliability and failure to recover several motivating examples, the included studies used many of the same underlying tools to build, optimize, and evaluate their methods.
+We therefore believe that our framework can serve to taxonomize ongoing work of this type and inform the development of customizable implementations.
 Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth.

### 2023-03-30 docs/Manuscript/Manuscript.Rmd
revise manuscript following discussions with collaborators

@@
     affiliation: 1
-  - name: Patti McCall
-    email: 
+  - name: Patti McCall-Junkin
+    email: pattimccall.junkin@gmail.com
     affiliation: 2
@@
   - code: 2
-    address: Department, Street, City, State, Zip
+    address: University at Florida, Smathers Libraries Academic Research & Consulting Services, P. O. Box 117000, Gainesville, FL 32611-7000
   - code: 3
@@
 abstract: |
-  Background
+  **Background**
 
@@
 
-  Methods
+  **Methods**
 
@@
 
-  Results
+  **Results**
 
@@
 
-  Conclusions
+  **Conclusions**
 
@@
 ---
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(dplyr)
+library(stringr)
+googlesheets4::gs4_auth(email = "cornelioid@gmail.com")
+Sys.sleep(10)
+```
 
@@
 
-We modeled our literature review on PRISMA guidelines (PRISMA Group, 2021). Because we focus on methodologies rather than approaches, we deviated in some ways from these guidelines.
+## PRISMA checklist
 
-## PRISMA-S checklist
-
-The PRISMA-S checklist comprises 16 items categorized into information sources and methods, search strategies, peer review, and managing records.
-We use the checklist to report the steps of our systematic review.
-Some items of the PRISMA-S checklist are intended for meta analysis and, therefore, did not apply to this study.
-We followed the checklist throughout the search process and explain the deviations from it in [Supplemental Table 1](https://docs.google.com/spreadsheets/d/1lKI5WWGIzo_hE9olkWWVCbLjCY9To5fiKPiefXUqgoU/edit).
+We include a PRISMA checklist as Supplemental Table 1.
+Because we focus on methodologies rather than approaches, we deviate in some ways from PRISMA guidelines.
+In particular, many items of the checklist, involving bias assessment and quantitative synthesis, are intended for meta-analyses and, therefore, did not apply to this study.
 
@@
 We finalized the PubMed search first, then adapted it to the other platforms.
-Through each platform, we searched those databases included by default.[^databases]
 
-[^databases]: This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts & Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.
+Through each platform, we searched those databases included by default.
+This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts & Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.
 
@@
 We conducted all searches over 2021 July 19–22.
-We organized the results into a Zotero collection with one subcollection for each database.
+We tentatively excluded results from Google Scholar because they were missing abstracts, and later agreed to discard these results due to the lack of reproducibility of the search.
+We organized the remaining results into a Zotero collection with one subcollection for each database.
 We then imported the pooled results to Covidence, which merged some duplicate entries.
@@
 
-Following the determination of 25 sources aligning with our systematic review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)).
+Following the determination of 25 sources aligning with our review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
@@
 ```{r tab:synthesis, echo=FALSE, message=FALSE, warning=FALSE}
-library(dplyr)
 googlesheets4::read_sheet(
@@
     Task = objective,
+    Aim = ifelse(generalizable_knowledge == "Science", "Knowledge", "Practice"),
     Source = type_of_data,
@@
   mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) |> 
-  mutate(across(
-    c(`Cases`, `Features`),
-    sapply, FUN = paste0, collapse = "; "
-  )) |> 
+  mutate(across(c(Cases, Features), sapply, FUN = format, big.mark = ",")) |>
+  mutate(across(c(Cases, Features), sapply, FUN = paste0, collapse = "; ")) |>
+  # mutate(across(c(Cases, Features), str_split, pattern = "; ")) |> 
+  # mutate(Dimensions = purrr::map2_chr(
+  #   Cases, Features,
+  #   ~ str_c(.x, .y, sep = " × ", collapse = "; ")
+  # )) |> 
   mutate(across(
@@
 tab_synthesis |> 
+  select(-Date) |> 
+  # prevent numeric list columns from dominating the horizontal space
+  mutate(across(c(-Cases, -Features), ~ str_c(., "\\hspace{6em}"))) |> 
   knitr::kable(
@@
       "\\label{tab:synthesis}",
-      "Studies included in the method synthesis.",
-      " 'Date' is the earliest the study is known to have been public."
+      "Studies included in the method synthesis,",
+      " arranged by the earliest the study is known to have been public."
     )
@@
 Another dominant characteristic of included studies is their remit.
-We categorized studies as "science" or "quality improvement" based on whether their aim was to produce generalizable knowledge or to improve practice.
-We classified 7 of the 25 studies as "science", making the majority of studies from a "quality improvement" standpoint.
+We categorized studies as "knowledge" or "practice" based on whether their aim was to produce generalizable knowledge or to improve practice.
+We classified 7 of the 25 studies as "knowledge", making the majority of studies from a "practice" standpoint.
 These studies aimed primarily to provide providers with tools to use in clinic or to improve outcomes for patients currently receiving care.
@@
 ```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
-library(dplyr)
-library(stringr)
 googlesheets4::read_sheet(
@@
   )) |>
+  mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
   knitr::kable(
@@
 
-## Limitations
-
 ## Conclusions
@@
 
-## Appendix
+# Appendix
 
-### Search strings
+## Search strings
 
 Here we reproduce the search strings and platform specifications used in our literature search.
-We first finalized the **PubMed** search string below:
+We first finalized the **PubMed** search string below:[^pubmed-url]
+
+[^pubmed-url]: <https://pubmed.ncbi.nlm.nih.gov/?term=(+"case-based+reasoning"+[All+Fields]+OR+"case-based+system"+[All+Fields]+)+OR+(+"individualized+modeling"+[All+Fields]+OR+"personalized+modeling"+[All+Fields]+OR+"customized+modeling"+[All+Fields]+)+OR+(+"individualized+cohort"+[All+Fields]+)+OR+(+(+"patient+similarity"+[All+Fields]+OR+"patient+distance"+[All+Fields]+OR+"patient+connection"+[All+Fields]+OR+"patient+affinity"+[All+Fields]+OR+"patient+clustering"+[All+Fields]+)+AND+(+"cohort+study"+[All+Fields]+)+)&sort=date>
 
@@
 `"patient affinity" OR`                         
-`"patient clustering"`                          
+`"patient clustering"`;                         
 Refined search: `"cohort"`                      

### 2023-04-03 docs/Manuscript/Manuscript.Rmd
revise manuscript per outstanding prisma items

@@
 
-  We searched five electronic bibliographic databases for publications that report applications of localized models to tasks involving analysis of clinical or health data.
+  We searched four electronic bibliographic platforms (PubMed, Web of Science, Academic Search Premier, and MathSciNet) during 2021 July 19–22.
+We screened and reviewed entries based on four criteria intended to specify applications of localized models to tasks involving analysis of clinical or health data.
 Two authors divided title/abstract screening and collaboratively reviewed screened entries for inclusion.
-We also reviewed a seed set that inspired the review and traced a subset of citations from  We discussed and summarized settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
+We also screened and reviewed a seed set that inspired the review and traced methodological citations from included entries.
+We discussed and summarized settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
+We received no specific funding for this work.
 
@@
 In particular, many items of the checklist, involving bias assessment and quantitative synthesis, are intended for meta-analyses and, therefore, did not apply to this study.
+Because we are not aware of any standard procedures for conducting reviews and syntheses of methodology, no protocol was prepared for this study.
 
@@
 
-## Title/Abstract screen
+## Title/abstract screen
 
@@
 
+# Acknowledgments
+
+# Support
+
+The authors received no specific funding or other support for this work.
+
+# Competing interests
+
+The authors declare that they have no competing interests.
+
+# Data and code availability
+
+Data collected or encoded for included studies are publicly available in two Google Sheets. Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
+
+* Properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
+* Composite techniques: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
+* Analysis and manuscript code: <https://github.com/corybrunson/imsr>
+
 # Appendix

### 2023-04-03 scr/synthesis.r
read data directly from google sheet

@@ -3,9 +3,12 @@ library(tidyverse)
 # Excel file downloaded from Google Sheet:
 # https://docs.google.com/spreadsheets/d/
 # 1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/
-here::here("data/IMSR composite techniques.xlsx") %>%
-  readxl::read_xlsx() %>%
-  select(seq(3L)) %>%
+# here::here("data/IMSR composite techniques.xlsx") %>%
+#   readxl::read_xlsx() %>%
+googlesheets4::read_sheet(
+  ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 1L
+) |> 
+  select(seq(2L, 4L)) %>%
   rename_with(~ str_replace(snakecase::to_snake_case(.), "_s$", "s")) %>%
   mutate(across(c(approach_types, terms), ~ str_split(., "(;|,) *"))) %>%
   print() -> composite_studies
@@ -23,7 +26,7 @@ composite_studies %>%
   ) +
   labs(x = NULL, y = NULL) ->
   year_plot
-ggsave(here::here("fig/fig-years.png"), year_plot, width = 5, height = 3)
+ggsave(here::here("fig/fig-years.png"), year_plot, width = 5, height = 2)
 
 # approach types
 composite_studies %>%

### 2023-04-04 docs/Manuscript/Manuscript.Rmd
additional changes to address prisma checklist items

@@
   Of 1,309 search results, 328 were reviewed.
-Of these, combined with 43 publications that seeded the review and 1 obtained by citation tracking, 51 were included in the review and 25 of these in the methodological synthesis.
+Of these, combined with 43 publications that seeded the review and 1 obtained by citation tracking, 25 were included in the review and methodological synthesis.
 The specificity of several search terms was poor, and inter-rater reliability was low.
@@
 The authors arrived at agreement on each manuscript through discussion.
+We calculated inter-rater reliability to evaluate our screening and review process.
+
+During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival [@Mariuzzi1997] and survival curves [@Lowsky2012].
+These approaches were novel to CBR and presage later developments, so were helpful to understanding the development of localized modeling.
+However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the responses of the cohort.
+A review of all studies that use nearest neighbors prediction, even in the clinical domain, would be impractical.
+Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to exclude those studies whose approach was equivalent to nearest neighbors prediction using a conventional similarity measure.
 
@@
 
-## Taxonomy of included studies
-
-We noticed during full text review that our inclusion criteria admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the responses of the cohort.
-Mean and plurality can be understood as statistical models, and other studies used equally simple though less conventional calculations.
-However, a review of all studies that use nearest neighbors prediction, even in the clinical domain, would be impractical.
-Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to analyze only those studies whose approach was not equivalent to nearest neighbors prediction using a conventional similarity measure.
-The unanalyzed studies comprised 25 of the 44 studies included from the search and 1 of those from the seed set, and our analysis focuses on the remaining 19 + 1 + 5 = 25 studies.
-
 ## Coding
 
-Following the determination of 25 sources aligning with our review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)).
+Following the selection sources aligning with our review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
@@
 Of these, 328 entries met the screening criteria and were assessed through full text review.
-Of these, 44 fit the criteria and were considered reports of included studies.
-Exclusion of NN models left 19 studies being included in the method synthesis.
+Of these, 44 fit the original criteria, and 19 were included as distinct from NN prediction.
 
@@
 
-We then reviewed 43 studies that comprised a seed set that inspired this review.
+We then reviewed 43 studies comprising a seed set that inspired this review.
 After removing duplicates and screening for eligibility, we were left with 6 additional studies, 1 of which was excluded from the synthesis for using nearest neighbors prediction.
 Finally, citations in reference lists from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
-This left us with 51 studies reviewed and 25 included in the synthesis.
-
+This left us with 19 + 1 + 5 = 25 studies included in the review and synthesis.
 Inter-rater reliability was low, at 82% relative to a 72% probability of random agreement.
-This was due, in parts, to different interpretations of some eligibility criteria by the authors, inconsistent terminology across the sample, and incomplete reporting of resources and methods in the sample.
-Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning.
-We will discuss the different domains, terminologies, and reporting issues of the sample in later sections.
 
@@
 # Discussion
+
+## Process
+
+The low inter-rater reliability was due, in parts, to different interpretations of some eligibility criteria by the authors, inconsistent terminology across the sample, and incomplete reporting of resources and methods in the sample.
+Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning.
+We will discuss the different domains, terminologies, and reporting issues of the sample in the remainder of this section.
 
@@
 
-* Properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
-* Composite techniques: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
-* Analysis and manuscript code: <https://github.com/corybrunson/imsr>
+* Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
+* Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
+* Code used to perform analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
 

### 2023-04-05 docs/Manuscript/Manuscript.Rmd
link to zotero group library

@@
 We tentatively excluded results from Google Scholar because they were missing abstracts, and later agreed to discard these results due to the lack of reproducibility of the search.
-We organized the remaining results into a Zotero collection with one subcollection for each database.
+We organized the remaining results into a public Zotero collection with one subcollection for each database.
 We then imported the pooled results to Covidence, which merged some duplicate entries.
@@
 
-Data collected or encoded for included studies are publicly available in two Google Sheets. Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
+Search results are publicly available in a Zotero group library.
+Data collected or encoded for included studies are publicly available in two Google Sheets.
+Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
 
+* Search results, included studies, and other bibliography: <https://www.zotero.org/groups/5017571/imsr/>
 * Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
@@
 
-These totaled 1,422 sources from all platforms. We organized the full results in a Zotero collection alongside the seed set and created a single folder for the 25 results reviewed in detail.
+These totaled 1,422 sources from all platforms. We organized the full results in a public Zotero collection alongside the seed set and created a single folder for the 25 results reviewed in detail.
 

### 2023-08-13 docs/Manuscript/Manuscript.Rmd
begin making edits in response to reviewers

@@
 The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching.
-The idea is to use patient similarity not to obtain an ideally homogeneous population matched to the index case, as in CBR and NN prediction, but to obtain a relevant yet heterogeneous population, each member of which deviates slightly from the index case.
-As a result, the variation of any single factor will be (1) centered near its value on the index case and (2) most different from the index case in cases that in other respects are more similar to it.
-This makes interpretative models like GLMs better able to identify and measure the effects of determining factors.
-Localized modeling shows promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.
+Whereas CBR and NN prediction use patient similarity to obtain an ideally homogeneous subpopulation matched to the index case, in localized modeling one seeks a relevant yet heterogeneous subpopulation, each member of which deviates slightly from the index case.
+In principle, this allows interpretable models like GLMs to identify and measure the effects of determining factors that are specific to the index case.
+In the reviewed studies, localized modeling shows some promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.
 
-One memorable aspiration of clinical informatics was the "green button" that would, in response to clinical queries, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014.
+One memorable aspiration of clinical informatics was the "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014.
 Localized modeling has recently been put forth as one answer to this appeal.
@@
 The motivation for this scoping review is to describe the settings in and problems with which localized models have been tasked to date, and to synthesize existing methods into a general framework.
-Along the way, we attempt to reconcile varying terminology, summarize evaluation techniques, propose needed follow-up work, and recommend best practices.
+Along the way, we attempt to reconcile terminology, summarize evaluation techniques, propose needed follow-up work, and recommend best practices.
 

### 2023-08-14 docs/Manuscript/Manuscript.Rmd
commit uncommitted edits from months ago

--- 2023-04-05 docs/Manuscript/Manuscript.Rmd
@@
   - code: 1
-    address: Nova Southeastern University College of Osteopathic Medicine, 3200 S University Drive, Davie, FL 33328
+    address: College of Osteopathic Medicine, Nova Southeastern University, 3200 S University Drive, Davie, FL 33328
   - code: 2
-    address: University at Florida, Smathers Libraries Academic Research & Consulting Services, P. O. Box 117000, Gainesville, FL 32611-7000
+    address: Smathers Libraries Academic Research & Consulting Services, University at Florida, P. O. Box 117000, Gainesville, FL 32611-7000
   - code: 3
@@
 abstract: |
-  **Background**
-
+  Background:
   Applications of classical case-based reasoning (CBR) to clinical and health tasks have given rise to a family of techniques in which an (often conventional) model is fitted to a subset of labeled cases, selected for relevance or similarity to an unlabeled case to which the model is applied.
@@
 
-  **Methods**
-
+  Methods:
   We searched four electronic bibliographic platforms (PubMed, Web of Science, Academic Search Premier, and MathSciNet) during 2021 July 19–22.
@@
 
-  **Results**
-
+  Results:
   Of 1,309 search results, 328 were reviewed.
@@
 
-  **Conclusions**
-
+  Conclusions:
   Localized models hold promise to improve the performance of models used in clinical decision support tools while maintaining interpretability, provided their computational limitations can be overcome.
@@
   - clinical decision support
+  - scoping review
 journal: "Artificial Intelligence In Medicine"
-date: "`r Sys.Date()`"
+date: "`r format(Sys.Date(), format = '%Y %b %d')`"
 classoption: preprint, 3p, authoryear
-bibliography: SR-bibliography,
-  SR-0-Found, SR-2-Included, SR-3-Tracked
+bibliography:
+  - SR-bibliography.bib
+  - SR-0-Found.bib
+  - SR-2-Included.bib
+  - SR-3-Tracked.bib
 linenumbers: true
@@
 # csl: https://www.zotero.org/styles/elsevier-harvard
+# csl: https://www.zotero.org/styles/artificial-intelligence-in-medicine
+csl: artificial-intelligence-in-medicine.csl
 output: 
@@
     latex_engine: xelatex
-    citation_package: natbib
+    # citation_package: natbib
+    citation_package: "default"
 ---
@@
 
-## PRISMA checklist
+## Protocol
 
-We include a PRISMA checklist as Supplemental Table 1.
+Our methods followed prepared plans, which we collect into a protocol in the Appendix.
+We also note deviations from these plans and the reasons for them.
+
+### PRISMA checklist
+
+We include a PRISMA checklist as Supplemental Table 1 and a PRISMA abstract checklist as Supplemental Table 2.
 Because we focus on methodologies rather than approaches, we deviate in some ways from PRISMA guidelines.
@@
 
-During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival [@Mariuzzi1997] and survival curves [@Lowsky2012].
+During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival [@Mariuzzi1997] and survival curves [@Lowsky2013].
 These approaches were novel to CBR and presage later developments, so were helpful to understanding the development of localized modeling.
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies ([Supplemental Table 2](https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/edit)).
+Following the selection sources aligning with our review, we collected characteristics of included studies (see the Appendix).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
@@
   rename_with(snakecase::to_snake_case) |>
+  filter(is.na(synthesis)) |> 
   arrange(date_of_publication) |> 
@@
   # prevent numeric list columns from dominating the horizontal space
-  mutate(across(c(-Cases, -Features), ~ str_c(., "\\hspace{6em}"))) |> 
+  mutate(across(
+    c(-Citation, -Cases, -Features),
+    ~ str_c(., "\\hspace{6em}")
+  )) |> 
   knitr::kable(
@@
   rename_with(snakecase::to_snake_case) |>
+  filter(is.na(synthesis)) |> 
+  # mutate(zotero_key = str_split(zotero_key, "; ")) |> 
+  # tidyr::unnest(zotero_key) |> 
   transmute(
-    Citation = paste0("@", zotero_key),
+    # Citation = paste0("@", zotero_key),
+    Citation = sapply(
+      str_split(zotero_key, "; "),
+      function(s) str_c("[", str_c("@", s, collapse = "; "), "]")
+    ),
     Elements = str_remove_all(approach_type_s, "\\[|\\]"),
@@
   )) |>
-  mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
+  # mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
   knitr::kable(
@@
 
+## Protocol
+
+### Research question
+
+How is patient similarity–based individualized modeling conducted using retrospective data?
+
+### Purpose of review
+
+1. Provide a summary of individualized models to date.
+2. Lay the groundwork for conducting a comparison study of individualized models.
+3. Provide a framework for future individualized modeling studies.
+
+### Search
+
+The procedure for formulating the search began with an evaluation of the research question.
+We highlighted specific elements within our topic of interest that we found critical to our search and listed them using an OR of ANDs, pairing terms we believed would provide our desired result.
+Following the solidification of the search, a thesaurus was created in which each term was expanded by synonyms that are similar enough to our core term to be applicable to our search.
+The expanded search string was then evaluated using the PubMed Advanced Search platform.
+We initially included each term and their synonyms separately to evaluate what resulted.
+Several terms were eliminated due to PubMed classifying them as “phrases not found” and other terms were removed to reduce the number of results, providing a more concise list of results.
+At the conclusions of this process for each individual term, we combined each search term using an OR of ANDs to ultimately form our search.
+
+The search will have been designed to recover studies of the kind reviewed by the review papers from which we obtained our “seed set”.
+To validate the final search design, we will determine how many of the papers in this seed set that are indexed by PubMed are actually recovered by our search.
+In most cases, the focus of a review paper is different from ours, so we will only perform this validation test on the seed set obtained from two review papers that are (a) closest in focus to ours and (b) use terminology associated with the two distinct sub-literatures relevant to our focus: Choudhury & Begum (2016), which focuses on case-based reasoning in medicine, and Sharafoddini, Dubin, & Lee (2017), which focuses on patient similarity–based prediction models on health data.
+The proportion of each PubMed-indexed seed set that is recovered from our PubMed search provides a rough and optimistic yet useful estimate of the proportion of the relevant literature that our full search strategy will recover.
+
+Once we have finalized the search as a logical pattern, we will take the following steps to generate the sample/corpus of literature that will be the starting point for our selection process.
+
+1.
+The logical pattern will be converted to a search string using the syntax appropriate to each database in our search strategy.
+These include PubMed (already done as part of the search design), Web of Science, Academic Search Premier/Elite, and Mathematical Reviews.
+2. The search will be conducted on each database and the results organized into a Zotero collection, with one subcollection for each database.
+3.
+Duplicate results will be identified and merged.
+(A result obtained from multiple databases should have only one Zotero entry but should be filed under the subcollection for each database in which it was found.)
+
+Following discussion among AC, PMJ, and JCB, we omitted results from Google Scholar due to missing abstracts, missing URLs, high overlap with other search results, and irreproducibility of the search process.
+
+### Screening process
+
+1. Deduplication (Covidence)
+    a. Specific to that source: title, creator, URL, DOI
+    b. May overlap: journal, publication date, ISSN
+2. Title
+    a. Exclude non clinical/non medical setting
+    b. Must be in English
+    c. Must be original study (not reviews, surveys, opinion, news)
+    d. Exclude if search term clearly has different meaning than intended
+3. Abstract
+4. Full Text
+    a. Study criteria (inclusion/exclusion)
+
+### Selection process
+
+Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
+
+* Uses labeled case-level (empirical) data set 
+* Defines a continuous-valued multivariate case similarity measure
+* Uses the similarity measure to select cohorts for index cases from the corpus
+* Fits statistical models to cohorts to make inferences about index cases
+
+We decided after concluding full-text review to perform one round of citation-tracking, of citations within the Methods (or analogous) sections of the included entries.
+
+### Analysis and synthesis
+
+Because we could not predict the scope of methodological approaches we would encounter, we did not prepare specific analyses or syntheses _a priori_.
+
 ## Search strings

### 2023-08-15 docs/Manuscript/Manuscript.Rmd
continue transcribing straightforward responses to reviewers

@@
 ---
-title: "Clinical prediction using similarity-based cohorts: A scoping review"
+title: "Clinical prediction with localized modeling using similarity-based cohorts: A scoping review"
 author:
@@
 
-## Protocol
-
-Our methods followed prepared plans, which we collect into a protocol in the Appendix.
+Our methods followed prepared plans, which we describe in detail in the Appendix.
 We also note deviations from these plans and the reasons for them.
 
-### PRISMA checklist
+## PRISMA checklist
 
@@
 Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
-We finalized the PubMed search first, then adapted it to the other platforms.
+We finalized the PubMed search first, then adapted it to the other platforms (Sections \ref{sec:appendix-search}).
 
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies (see the Appendix).
+Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{sec:availability}).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
@@
 
-```{r fig:prisma, fig.align='center', out.width="80%", fig.cap = "\\label{fig:prisma}PRISMA-S flow chart.", echo=FALSE}
+```{r fig:prisma, fig.align='center', out.width="80%", echo=FALSE}
+#| fig.cap =
+#| "\\label{fig:prisma}
+#| PRISMA-S flow chart.
+#|  Lowercase letters refer to items obtained from
+#|  the seed set (m),
+#|  the structured search (n), and
+#|  citation/reference tracking (s)."
 knitr::include_graphics("../../fig/fig-prisma.pdf")
@@
 
-## Overview
+## Bibliographic and methodological properties
 
-The 25 included studies were analyzed based on 20 characteristics (Table \ref{tab:synthesis}), and we describe some observations here.
-The studies were published in a variety of journals, with some of greater frequency, though no single journal published exceptionally many.
+The 25 included studies were analyzed based on 20 characteristics (see Section \ref{sec:availability}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
 The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, and _Hindawi - Journal of Healthcare Engineering_ published 3.
-Another distinguishing factor was the years of publication of the included studies (Figure \ref{fig:year}).
-While they stretch back to the late 1990s, they have become more common in recent years, which suggests that this is still an active, though not rapidly expanding methodological approach.
+Another interesting pattern lay in the years of publication of the included studies (Figure \ref{fig:year}).
+While they stretch back to the late 1990s, they have become more common, which suggests that this is still an active, though not rapidly expanding, methodological approach.
 We note that, while CBR in health and medicine originated as early as 1990, these studies did emerge early on once CBR had established itself.
@@
 ```{r tab:synthesis, echo=FALSE, message=FALSE, warning=FALSE}
-googlesheets4::read_sheet(
-  ss = "1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo", sheet = 1L
-) |>
-  rename_with(snakecase::to_snake_case) |>
-  filter(is.na(synthesis)) |> 
-  arrange(date_of_publication) |> 
-  transmute(
-    Citation = paste0("@", zotero_key),
-    Date = format(date_of_publication, "%Y %b %d"),
-    Task = objective,
-    Aim = ifelse(generalizable_knowledge == "Science", "Knowledge", "Practice"),
-    Source = type_of_data,
-    Type = range_of_data,
-    `Cases` = number_of_cases_incidents,
-    `Features` = number_of_predictors_features
-  ) |>
-  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) |> 
-  mutate(across(c(Cases, Features), sapply, FUN = format, big.mark = ",")) |>
-  mutate(across(c(Cases, Features), sapply, FUN = paste0, collapse = "; ")) |>
-  # mutate(across(c(Cases, Features), str_split, pattern = "; ")) |> 
-  # mutate(Dimensions = purrr::map2_chr(
-  #   Cases, Features,
-  #   ~ str_c(.x, .y, sep = " × ", collapse = "; ")
-  # )) |> 
-  mutate(across(
-    everything(),
-    # ~ stringr::str_replace_all(., "<U+001B>", "")
-    textclean::replace_non_ascii
-  )) ->
-  tab_synthesis
+tab_synthesis <- read_rds(here::here("tab/synthesis.rds"))
 tab_synthesis |> 
@@
 
-```{r fig:year, fig.align='center', out.width="80%", fig.cap = "\\label{fig:year}Number of publications in our sample each year.", echo=FALSE}
+```{r fig:year, fig.align='center', out.width="80%", echo=FALSE}
+#| fig.cap =
+#| "\\label{fig:year}
+#| Number of publications in our sample each year."
 knitr::include_graphics("../../fig/fig-years.png")
@@
 
-```{r fig:framework, fig.align='center', out.width="80%", fig.cap = "\\label{fig:framework}General framework for localized models.", echo=FALSE}
+```{r fig:framework, fig.align='center', out.width="80%", echo=FALSE}
+#| fig.cap =
+#| "\\label{fig:framework}
+#| General framework for localized models."
 knitr::include_graphics("../../fig/fig-typology.pdf")
@@
 # Data and code availability
+\label{sec:availability}
 
@@
 
-## Protocol
-
-### Research question
+## Research question
 
@@
 
-### Purpose of review
+## Purpose of review
+\label{sec:appendix-purpose}
 
@@
 
-### Search
+## Search
+\label{sec:appendix-search}
 
@@
 
-### Screening process
-
-1. Deduplication (Covidence)
-    a. Specific to that source: title, creator, URL, DOI
-    b. May overlap: journal, publication date, ISSN
-2. Title
-    a. Exclude non clinical/non medical setting
-    b. Must be in English
-    c. Must be original study (not reviews, surveys, opinion, news)
-    d. Exclude if search term clearly has different meaning than intended
-3. Abstract
-4. Full Text
-    a. Study criteria (inclusion/exclusion)
-
-### Selection process
-
-Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
-
-* Uses labeled case-level (empirical) data set 
-* Defines a continuous-valued multivariate case similarity measure
-* Uses the similarity measure to select cohorts for index cases from the corpus
-* Fits statistical models to cohorts to make inferences about index cases
-
-We decided after concluding full-text review to perform one round of citation-tracking, of citations within the Methods (or analogous) sections of the included entries.
-
-### Analysis and synthesis
-
-Because we could not predict the scope of methodological approaches we would encounter, we did not prepare specific analyses or syntheses _a priori_.
-
-## Search strings
+### Search strings
 
@@
 
+## Screening process
+\label{sec:appendix-screening}
+
+1. Deduplication (Covidence)
+    a. Specific to that source: title, creator, URL, DOI
+    b. May overlap: journal, publication date, ISSN
+2. Title
+    a. Exclude non clinical/non medical setting
+    b. Must be in English
+    c. Must be original study (not reviews, surveys, opinion, news)
+    d. Exclude if search term clearly has different meaning than intended
+3. Abstract
+4. Full Text
+    a. Study criteria (inclusion/exclusion)
+
+## Selection process
+\label{sec:appendix-selection}
+
+Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
+
+* Uses labeled case-level (empirical) data set 
+* Defines a continuous-valued multivariate case similarity measure
+* Uses the similarity measure to select cohorts for index cases from the corpus
+* Fits statistical models to cohorts to make inferences about index cases
+
+We decided after concluding full-text review to perform one round of citation-tracking, of citations within the Methods (or analogous) sections of the included entries.
+
+## Analysis and synthesis
+
+Because we could not predict the scope of methodological approaches we would encounter, we did not prepare specific analyses or syntheses _a priori_.
+
 # References {-}

### 2023-08-15 scr/synthesis.r

@@ -1,4 +1,69 @@
 library(tidyverse)
+library(ggmosaic)
+
+# scrape and format table summarizing studies included in synthesis
+# Excel file downloaded from Google Sheet:
+# https://docs.google.com/spreadsheets/d/
+# 1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/
+googlesheets4::read_sheet(
+  ss = "1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo", sheet = 1L
+) |>
+  rename_with(snakecase::to_snake_case) |>
+  filter(is.na(synthesis)) |> 
+  arrange(date_of_publication) |> 
+  transmute(
+    Citation = paste0("@", zotero_key),
+    Date = format(date_of_publication, "%Y %b %d"),
+    Task = objective,
+    Aim = ifelse(generalizable_knowledge == "Science", "Knowledge", "Practice"),
+    Source = type_of_data,
+    Type = range_of_data,
+    `Cases` = number_of_cases_incidents,
+    `Features` = number_of_predictors_features
+  ) |>
+  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) |> 
+  mutate(across(c(Cases, Features), sapply, FUN = format, big.mark = ",")) |>
+  mutate(across(c(Cases, Features), sapply, FUN = paste0, collapse = "; ")) |>
+  # mutate(across(c(Cases, Features), str_split, pattern = "; ")) |> 
+  # mutate(Dimensions = purrr::map2_chr(
+  #   Cases, Features,
+  #   ~ str_c(.x, .y, sep = " × ", collapse = "; ")
+  # )) |> 
+  mutate(across(
+    everything(),
+    # ~ stringr::str_replace_all(., "<U+001B>", "")
+    textclean::replace_non_ascii
+  )) ->
+  tab_synthesis
+# write to file, to be read into document
+write_rds(tab_synthesis, file = here::here("tab/synthesis.rds"))
+
+# mosaic plot of task, aim, and source
+tab_synthesis %>%
+  mutate(Source = factor(
+    Source,
+    c("Clinical", "Laboratory", "Imaging",
+      "Clinical, Laboratory", "Clinical, Imaging", "Patient-reported")
+  )) %>%
+  mutate(Task = factor(
+    Task,
[…]

### 2023-08-16 docs/Manuscript/Manuscript.Rmd
complete transcribing straightforward responses to reviewers

@@
 The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, and _Hindawi - Journal of Healthcare Engineering_ published 3.
+
 Another interesting pattern lay in the years of publication of the included studies (Figure \ref{fig:year}).
@@
 These studies aimed primarily to provide providers with tools to use in clinic or to improve outcomes for patients currently receiving care.
+
 Lastly, we observed a commonality among 6 of the 25 included studies in having evaluated methods using leave-one-out cross-validation (LOOCV).
@@
 
-In the Discussion, we discuss trends in the literature we reviewed, identify common limitations and gaps in understanding, and attempt to synthesize their methods into a general framework.
-
-# Discussion
-
-## Process
-
-The low inter-rater reliability was due, in parts, to different interpretations of some eligibility criteria by the authors, inconsistent terminology across the sample, and incomplete reporting of resources and methods in the sample.
-Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning.
-We will discuss the different domains, terminologies, and reporting issues of the sample in the remainder of this section.
-
-## Trends
-
-We draw several observations from our encoding of the included studies.
-We first focus on two driving questions: What challenges did investigators set out to solve?
-How did their approaches meet these challenges?
-We then comment on the connectedness of the included studies and the potential for recombination.
-See Table \ref{tab:composite} for details on approaches taken by the synthesized studies.
-
-(See [this sheet](https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/).)
+In addition to their remit, we coded several aspects of the design of each study, including the source and type of data and the clinical task the model performed (Table \ref{tab:synthesis}), as well as the specific methodology and terminology adopted (Table \ref{tab:composite}).
+In the following subsections, we take a closer look at these design elements.
 
@@
 
-### Application domain
+## Application domain
 
@@
 For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006].
-This was part of a much larger literature on CBR systems that was included here despite relying on NN prediction because it used a non-traditional voting scheme to generate recommendations.
+This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used a non-traditional voting scheme to generate recommendations.
 A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021a].
@@
 
-### Method development
+The mosaic plot in Figure \ref{fig:properties} summarizes the relation between data source, clinical task, and remit.
+
+```{r fig:properties, fig.align='center', out.width="100%", echo=FALSE}
+#| fig.cap =
+#| "\\label{fig:properties}
+#| Share of studies characterized by several design elements:
+#|  data source (row),
+#|  clinical task (fill color), and
+#|  remit (opacity)."
+knitr::include_graphics("../../fig/fig-properties.png")
+```
+
+## Method development
 
@@
 Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018a].
+
+The frequency plot in Figure \ref{fig:} compares the rates at which several methodological elements are invoked in the sample.
+
+```{r fig:methods, fig.align='center', out.width="100%", echo=FALSE}
+#| fig.cap =
+#| "\\label{fig:methods}
+#| Frequency of recurring methodological elements across included studies."
+# knitr::include_graphics("../../fig/fig-methods.png")
+```
+
+In the Discussion, we discuss trends in the literature we reviewed, identify common limitations and gaps in understanding, and attempt to synthesize their methods into a general framework.
+
+# Discussion
+
+## Process
+
+The low inter-rater reliability was due, in parts, to different interpretations of some eligibility criteria by the authors, inconsistent terminology across the sample, and incomplete reporting of resources and methods in the sample.
+Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning.
+We will discuss the different domains, terminologies, and reporting issues of the sample in the remainder of this section.
+
+## Trends
+
+We draw several observations from our encoding of the included studies.
+We first focus on two driving questions: What challenges did investigators set out to solve?
+How did their approaches meet these challenges?
+We then comment on the connectedness of the included studies and the potential for recombination.
+
+**Add 1--2 paragraphs.**
 

### 2023-08-16 scr/synthesis.r

@@ -54,16 +54,34 @@ tab_synthesis %>%
   count(Source, Task, Aim) %>%
   ggplot() +
   theme_bw() +
-  theme(panel.grid = element_blank()) +
-  geom_mosaic(
+  theme(panel.grid = element_blank(),
+        axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  # force vertical axis labels to respect binnings without reference to 'Aim'
+  stat_mosaic(
     aes(x = product(Task, Source), fill = Task, weight = n),
-    # aes(x = product(Task, Source), fill = Task, alpha = Aim, weight = n),
+    divider = mosaic("v"),
+    fill = "transparent"
+  ) +
+  stat_mosaic(
+    aes(x = product(Task, Source), fill = Task, alpha = Aim, weight = n),
     divider = mosaic("v")
   ) +
-  # coord_flip() +
+  # stat_mosaic_text(
+  #   geom = "label",
+  #   aes(x = product(Aim, Task, Source), weight = n, label = n),
+  #   divider = mosaic("v")
+  # ) +
   scale_y_productlist("Source", labels = ggmosaic:::product_labels()) +
   scale_fill_brewer(type = "qual", na.value = "#000000") +
-  scale_alpha_manual(values = c(Knowledge = .5, Practice = 1))
+  scale_alpha_manual(values = c(Knowledge = .5, Practice = 1)) ->
+  properties_mosaic
+print(properties_mosaic)
+ggsave(
+  here::here("fig/fig-properties.png"), properties_mosaic,
+  width = 8, height = 4
+)
 
 # Excel file downloaded from Google Sheet:
 # https://docs.google.com/spreadsheets/d/
@@ -113,6 +131,26 @@ composite_approaches %>%
   arrange(desc(n)) %>%
   filter(! approach_types %in% drop_types) %>%
   print(n = Inf)
+# frequency plot
+composite_approaches %>%
+  filter(! approach_types %in% c("NN?", "unclear")) %>%
+  mutate(year = as.integer(str_remove(citation, "^.*, "))) %>%
+  group_by(approach_types) %>%
+  summarize(count = n(), earliest_use = min(year)) %>%
+  mutate(approach_types = fct_reorder(approach_types, earliest_use)) %>%
+  ggplot(aes(x = approach_types, y = count)) +
+  theme_bw() +
+  geom_col() +
[…]

### 2023-08-16 docs/Manuscript/Manuscript.Rmd
include new plot in manuscript + debug use of read_rds

@@
 ```{r tab:synthesis, echo=FALSE, message=FALSE, warning=FALSE}
-tab_synthesis <- read_rds(here::here("tab/synthesis.rds"))
+tab_synthesis <- readr::read_rds(here::here("tab/synthesis.rds"))
 tab_synthesis |> 
@@
 
-The frequency plot in Figure \ref{fig:} compares the rates at which several methodological elements are invoked in the sample.
+The frequency plot in Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
 
@@
 #| Frequency of recurring methodological elements across included studies."
-# knitr::include_graphics("../../fig/fig-methods.png")
+knitr::include_graphics("../../fig/fig-methods.png")
 ```

### 2023-08-16 scr/synthesis.r
make pipes consistent

@@ -7,10 +7,10 @@ library(ggmosaic)
 # 1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/
 googlesheets4::read_sheet(
   ss = "1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo", sheet = 1L
-) |>
-  rename_with(snakecase::to_snake_case) |>
-  filter(is.na(synthesis)) |> 
-  arrange(date_of_publication) |> 
+) %>%
+  rename_with(snakecase::to_snake_case) %>%
+  filter(is.na(synthesis)) %>% 
+  arrange(date_of_publication) %>% 
   transmute(
     Citation = paste0("@", zotero_key),
     Date = format(date_of_publication, "%Y %b %d"),
@@ -20,15 +20,21 @@ googlesheets4::read_sheet(
     Type = range_of_data,
     `Cases` = number_of_cases_incidents,
     `Features` = number_of_predictors_features
-  ) |>
-  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) |> 
-  mutate(across(c(Cases, Features), sapply, FUN = format, big.mark = ",")) |>
-  mutate(across(c(Cases, Features), sapply, FUN = paste0, collapse = "; ")) |>
-  # mutate(across(c(Cases, Features), str_split, pattern = "; ")) |> 
+  ) %>%
+  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) %>% 
+  mutate(across(
+    c(Cases, Features),
+    \(x) sapply(x, FUN = format, big.mark = ",")
+  )) %>% 
+  mutate(across(
+    c(Cases, Features),
+    \(x) sapply(x, FUN = paste0, collapse = "; ")
+  )) %>%
+  # mutate(across(c(Cases, Features), str_split, pattern = "; ")) %>% 
   # mutate(Dimensions = purrr::map2_chr(
   #   Cases, Features,
   #   ~ str_c(.x, .y, sep = " × ", collapse = "; ")
-  # )) |> 
+  # )) %>% 
   mutate(across(
     everything(),
     # ~ stringr::str_replace_all(., "<U+001B>", "")
@@ -90,7 +96,7 @@ ggsave(
 #   readxl::read_xlsx() %>%
 googlesheets4::read_sheet(
[…]

### 2023-08-17 docs/Manuscript/Manuscript.Rmd
use compiled latex file to get citation numbers for mosaic plot

@@
 tab_synthesis |> 
-  select(-Date) |> 
+  select(-Number, -Date) |> 
   # prevent numeric list columns from dominating the horizontal space

### 2023-08-17 scr/synthesis.r

@@ -1,6 +1,17 @@
 library(tidyverse)
 library(ggmosaic)
 
+# add citation numbers from most recent LaTeX compilation
+here::here("docs/Manuscript/Manuscript.tex") %>%
+  read_lines() %>%
+  enframe(name = NULL, value = "line") %>%
+  filter(str_detect(line, "\\{ref-")) %>%
+  transmute(
+    number = row_number(),
+    ref = str_replace(line, "^.*\\{ref-([A-Za-z]+[0-9]{4}[a-z]*)\\}.*$", "\\1")
+  ) %>%
+  print() -> citation_numbers
+
 # scrape and format table summarizing studies included in synthesis
 # Excel file downloaded from Google Sheet:
 # https://docs.google.com/spreadsheets/d/
@@ -9,9 +20,11 @@ googlesheets4::read_sheet(
   ss = "1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo", sheet = 1L
 ) %>%
   rename_with(snakecase::to_snake_case) %>%
-  filter(is.na(synthesis)) %>% 
-  arrange(date_of_publication) %>% 
+  filter(is.na(synthesis)) %>%
+  arrange(date_of_publication) %>%
+  inner_join(citation_numbers, by = c("zotero_key" = "ref")) %>%
   transmute(
+    Number = number,
     Citation = paste0("@", zotero_key),
     Date = format(date_of_publication, "%Y %b %d"),
     Task = objective,
@@ -57,7 +70,18 @@ tab_synthesis %>%
   )) %>%
   mutate(Aim = factor(Aim, c("Knowledge", "Practice"))) %>%
   # mutate(across(c(Source, Task, Aim), fct_infreq)) %>%
-  count(Source, Task, Aim) %>%
+  # count(Source, Task, Aim) %>%
+  group_by(Source, Task, Aim) %>%
+  summarize(
+    n = n(),
+    Refs = str_c("[", str_c(sort(Number), collapse = ","), "]")
+  ) %>%
+  ungroup() %>%
+  print() ->
+  mosaic_data
+mosaic_labs <- with(mosaic_data, str_c(Aim, Task, Source, sep = "\n"))
+mosaic_refs <- mosaic_data$Refs
+mosaic_data %>%
   ggplot() +
   theme_bw() +
   theme(panel.grid = element_blank(),
@@ -66,7 +90,7 @@ tab_synthesis %>%
         axis.ticks.x = element_blank()) +
[…]

### 2023-08-17 docs/Manuscript/Manuscript.tex
track latex file

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Artificial Intelligence In Medicine} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Clinical prediction with localized modeling using similarity-based cohorts: A scoping review},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}

\begin{frontmatter}

  \title{Clinical prediction with localized modeling using
similarity-based cohorts: A scoping review}
    \author[1]{Adi Cohen%
  %
  }
   \ead{ac4569@my.nsu.edu} 
    \author[2]{Patti McCall-Junkin%
  %
  }
   \ead{pattimccall.junkin@gmail.com} 
    \author[3]{Jason Cory Brunson%
  \corref{cor1}%
  }
   \ead{jason.brunson@medicine.ufl.edu} 
      \affiliation[1]{College of Osteopathic Medicine, Nova Southeastern
University, 3200 S University Drive, Davie, FL 33328}
    \affiliation[2]{Smathers Libraries Academic Research \& Consulting
Services, University at Florida, P. O. Box 117000, Gainesville, FL
32611-7000}
    \affiliation[3]{Laboratory for Systems Medicine, University of
Florida, P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Background: Applications of classical case-based reasoning (CBR) to
  clinical and health tasks have given rise to a family of techniques in
  which an (often conventional) model is fitted to a subset of labeled
  cases, selected for relevance or similarity to an unlabeled case to
  which the model is applied. These localized models, as we propose to
  call them, may perform as well as ``black-box'' models while retaining
  direct interpretations. Their adoption has been hindered by fragmented
  development and implementation challenges. Our goal in this scoping
  review was to describe the clinical and health applications of
  localized models to date and to derive a general framework for their
  design and evaluation.

  Methods: We searched four electronic bibliographic platforms (PubMed,
  Web of Science, Academic Search Premier, and MathSciNet) during 2021
  July 19--22. We screened and reviewed entries based on four criteria
  intended to specify applications of localized models to tasks
  involving analysis of clinical or health data. Two authors divided
  title/abstract screening and collaboratively reviewed screened entries
  for inclusion. We also screened and reviewed a seed set that inspired
  the review and traced methodological citations from included entries.
  We discussed and summarized settings, tasks, and tools; identified and
  tabulated themes; and synthesized the methods into a general
  framework. We received no specific funding for this work.

  Results: Of 1,309 search results, 328 were reviewed. Of these,
  combined with 43 publications that seeded the review and 1 obtained by
  citation tracking, 25 were included in the review and methodological
  synthesis. The specificity of several search terms was poor, and
  inter-rater reliability was low. Entries were published from 1997 to
  2021, with half of these since 2015. The most common tasks were
  prognosis and diagnosis, with only few recent studies focusing on care
  recommendation. The studies overwhelmingly focused on clinical data,
  occasionally also including laboratory and image-derived data. Only
  three studies used localized models for any task other than
  prediction, most recently to identify significant determinants behind
  recommendation outputs. Several studies commented on memory and
  runtime costs, and four proposed partial solutions to them. While most
  techniques were not replicated and did not cite predecessors, a
  general technique that specializes to most of those encountered was
  straightforward to describe.

  Conclusions: Localized models hold promise to improve the performance
  of models used in clinical decision support tools while maintaining
  interpretability, provided their computational limitations can be
  overcome. Our search undoubtedly failed to capture all studies that
  used methods of this kind, but those reviewed and synthesized exhibit
  consistent enough design to be reproduced by a general-purpose tool.
  We hope that our review and synthesis will assist the development of
  more efficient implementations suitable for integration into practice.
  \end{abstract}
    \begin{keyword}
    case-based reasoning \sep localized modeling \sep nearest
neighbors \sep clinical decision support \sep 
    scoping review
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Clinical decision support (CDS) systems have been a fixture of
artificial intelligence (AI) research since inception. As machine
learning (ML) emerged, clinical modeling tasks became some of the most
common and celebrated applications. A technique core to clinical
applications of both AI and ML is \emph{similarity-matching}, which
relies on a \emph{patient similarity measure}\footnote{Similarity
  measures between more granular units of analysis, such as encounters
  and decision points, are often still referred to as patient similarity
  measures.} to extract a cohort of past or training cases used to
inform the diagnosis, prognosis, or care of a new or test case.
Similarity-matching powers the retrieval step of most case-based
reasoning (CBR) systems and provides the neighborhoods from which
nearest neighbor (NN) models derive predictions. Whereas NN predictions
are generated automatically via averaging (regression) or voting
(classification), most CBR tools return the retrieved cases to the user
to inform human decision-making.

We discuss in this review a hybrid computational approach we call
\emph{localized modeling} that borrows elements from both clinical
application areas and appears to have been introduced independently
several times. The approach fits ML models, including regularized
families of generalized regression models (GLMs), to cohorts retrieved
via similarity matching. Whereas CBR and NN prediction use patient
similarity to obtain an ideally homogeneous subpopulation matched to the
index case, in localized modeling one seeks a relevant yet heterogeneous
subpopulation, each member of which deviates slightly from the index
case. In principle, this allows interpretable models like GLMs to
identify and measure the effects of determining factors that are
specific to the index case. In the reviewed studies, localized modeling
shows some promise toward simultaneously achieving the performance
improvements of ML and the interpretability of classical models.

One memorable aspiration of clinical informatics was the ``green
button'' that would, in response to a query, automatically retrieve
patient data from an institution's records with which to conduct
on-demand retrospective studies for an individual patient {[}1{]}.
Localized modeling has recently been put forth as one answer to this
appeal. Studies using localized models have been limited by
computational cost---a naïve implementation involves fitting,
optimizing, and evaluating a model on a separate population for each
training case---and recent studies have described algorithms and
implementations designed to reduce this cost. As greater access to
health data, informatical knowledge, algorithmic efficiency, and
computational power continue to spur innovation in CDS, we may expect
localized modeling to become more accessible and efficient, its
advantages better understood, and its disadvantages ameliorated.

The motivation for this scoping review is to describe the settings in
and problems with which localized models have been tasked to date, and
to synthesize existing methods into a general framework. Along the way,
we attempt to reconcile terminology, summarize evaluation techniques,
propose needed follow-up work, and recommend best practices.

\hypertarget{related-work}{%
\subsection{Related work}\label{related-work}}

Clinical CBR emerged among rule-based approaches and other AI tools in
the development of expert systems {[}2{]}. Early implementations
realized a general workflow described as ``the four REs'', later the
``R4 cycle'' {[}2,3{]}: Given a new case (or problem), the system
retrieves one or more past cases (solved problems) from a corpus, reuses
these to generate an understanding of (solution to) the new case,
revises this understanding to better fit the new case (sometimes called
``adaptation''), and retains the new case and its eventual understanding
in the corpus to be retrieved and reused in future. In contrast to
rule-based systems, which are variable-based and generally interpretable
as a single rule applied to all new cases, case-based systems provide
case-specific interpretations in the form of a number of more fully
understood reference cases. While the similarity measure used in the
retrieval step need not be changed as the corpus grows, proposed
measures have been diverse, contested, and rarely systematically
compared.

{[}4{]} distinguished two styles of CBR: problem-solving, which is more
procedural and used when objectives are more clearly defined, and
interpretative, which provides categorizations and justifications for
possible solutions. A similar dichotomy is commonly used to distinguish
the performance criterion for the usefulness of predictive models from
the interpretability criterion for their usability. In these terms,
localized models are highly procedural: The step of fitting a predictive
model to a retrieved cohort can be understood as an automated adaptive
strategy {[}3{]}, and the parameters that govern cohort retrieval can be
tuned alongside model hyperparameters in a ML workflow. Accordingly,
localized models have been primarily used for and evaluated on their
ability to more accurately predict classes or outcomes. Nevertheless, as
some more recent applications have begun to show, they can be used to
draw individualized inferences about the importance of different risk
factors. While the patient similarity measure used to retrieve each
cohort may be complicated, the cohort itself can be directly inspected
by the user. Provided the model family fitted to the cohorts is
interpretable, the localized model as a whole inherits this property.
Thus, localized models may embody both styles of CBR.

We refer the interested reader to several previous reviews of CBR in
medicine {[}3,5,6{]}, of measures of patient similarity {[}7{]}, and of
uses of patient similarity in predictive models {[}8--10{]}. While the
reviews of CBR focus on applications using health data, the patient
similarity reviews encompass many additional types of data (various
molecular -omics, genetic tests, medical images, laboratory tests,
patient preferences, patient-reported outcomes, tracking devices, social
media) and survey a much broader scope of models (exploratory analysis /
dimension reduction / cluster analysis, risk evaluation / outcome
prediction, clinical decision support / software tools). For the present
review, we are interested in how similarity matching on patient-level
health data is used to construct local cohorts for predictive modeling.

\hypertarget{objectives}{%
\subsection{Objectives}\label{objectives}}

Our goals in this review are (1) to describe applications to date of
similarity-based localized models using patient-level health data and
(2) to provide a general framework for the design and evaluation of such
localized models. Our selection of relevant papers from the search
corpus will be based on the following inclusion/exclusion criteria:

\begin{itemize}
\tightlist
\item
  Pulls case-level data from a corpus of past cases with known response
  values\newline (response may be outcome, diagnosis, subtype, etc.)
\item
  Defines a numeric multivariate case similarity measure\newline
  (include integer-valued measures)
\item
  Uses the similarity measure to select cohorts for index cases from the
  corpus\newline (for example, based on a training--testing partition)
\item
  Fits statistical models to cohorts from which to draw inferences about
  index cases\newline (outcome predictions count, survival estimates,
  risk factor effect estimates, model evaluation statistics, etc.)
\end{itemize}

We expected the general framework to boil down to three choices: A
patient similarity measure, a cohort selection process, and a
statistical model family. Such a general method would be
straightforward, if laborious, to implement, and specializations of it
would account for the majority of real-world applications. A general
implementation based on this method would allow researchers to expedite
every step of the analysis process, including selection, optimization,
and evaluation, and enable sensitivity, robustness, and multiverse
analyses that help identify the most consequential choices along the
way.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Our methods followed prepared plans, which we describe in detail in the
Appendix. We also note deviations from these plans and the reasons for
them.

\hypertarget{prisma-checklist}{%
\subsection{PRISMA checklist}\label{prisma-checklist}}

We include a PRISMA checklist as Supplemental Table 1 and a PRISMA
abstract checklist as Supplemental Table 2. Because we focus on
methodologies rather than approaches, we deviate in some ways from
PRISMA guidelines. In particular, many items of the checklist, involving
bias assessment and quantitative synthesis, are intended for
meta-analyses and, therefore, did not apply to this study. Because we
are not aware of any standard procedures for conducting reviews and
syntheses of methodology, no protocol was prepared for this study.

\hypertarget{search}{%
\subsection{Search}\label{search}}

We derived the eligibility criteria itemized in the Introduction from a
seed set of previously read studies. Based on these criteria, we
formulated search strings for five platforms: PubMed, Web of Science,
Academic Search Premier, Google Scholar, and MathSciNet. We finalized
the PubMed search first, then adapted it to the other platforms
(Sections \ref{sec:appendix-search}).

Through each platform, we searched those databases included by default.
This means that we searched both MEDLINE and PubMed Central (PMC)
through Pubmed (we did not exclude other databases, but our earliest
included results post-date them) and that we searched the six indices of
the Core Collection (the Science Citation Index Expanded, the Social
Sciences Citation Index, the Arts \& Humanities Citation Index, the
Emerging Sources Citation Index, the Conference Proceedings Citation
Index, and the Book Citation Index) as well as several regional
databases through the Web of Science platform.

The structure and terms of our search strings were inspired in part by
previous reviews adjacent to our topic of interest {[}9,10{]}. We
organized the PubMed search in disjunctive normal form (an OR of ANDs).
Following the solidification of an outline, we expanded each term to
include synonyms that are similar enough to be applicable to our search.
We then evaluated the expanded search string using the PubMed Advanced
Search platform. We initially included each term and their synonyms
separately to evaluate what resulted. We pruned several terms that
returned no results (``phrases not found'') or to reduce the number of
results. At the conclusions of this process for each individual term, we
combined the search terms using disjunctive normal form.

We conducted all searches over 2021 July 19--22. We tentatively excluded
results from Google Scholar because they were missing abstracts, and
later agreed to discard these results due to the lack of reproducibility
of the search. We organized the remaining results into a public Zotero
collection with one subcollection for each database. We then imported
the pooled results to Covidence, which merged some duplicate entries.

\hypertarget{titleabstract-screen}{%
\subsection{Title/abstract screen}\label{titleabstract-screen}}

Within Covidence, we screened titles and abstracts for relevance. We
decided on four eligibility criteria to expedite the screening process.
These include articles written in English for readability, original
studies to strictly include primary sources with new
information/conclusions, medical or clinical related settings due to the
applicability of our data to this field, and alternative use of the key
terms within the search string, meaning that it was clear from the title
and abstract that the entry was included due to the use of one or more
search terms with a different meaning than ours. Each of two authors (AC
and JCB) screened roughly half of the entries. They regularly reviewed
each other's decisions to improve consistency. In cases of uncertainty,
an entry would be included.

\hypertarget{full-text-review}{%
\subsection{Full text review}\label{full-text-review}}

We set out 4 criteria for full-text review to restrict to studies that
used some form of localized modeling on health data:

\begin{itemize}
\tightlist
\item
  Pulls case-level data from a corpus of past cases with known classes
  or outcomes
\item
  Defines a numeric multivariate case similarity measure
\item
  Uses the similarity measure to select cohorts for index cases from the
  corpus
\item
  Fits statistical models to cohorts to make inferences about index
  cases
\end{itemize}

Note that classical CBR satisfies the first and third criteria by
definition and in most cases will satisfy the second.

In Covidence, two authors (AC and JCB) independently evaluated each
manuscript for the four eligibility criteria. The first criterion that a
manuscript failed was designated the reason for exclusion. An antecedent
criterion was used to exclude manuscripts that did not report the
results of original studies involving real-world experiments or
empirical data, for example surveys of prior work and proposals of
frameworks. In cases of disagreement between the authors on the reason
for exclusion, the first criterion was adopted. There were 60
disagreements over inclusion versus exclusion. The authors arrived at
agreement on each manuscript through discussion. We calculated
inter-rater reliability to evaluate our screening and review process.

During full text review, we decided to expand the conception of
statistical models (fourth criterion): Rather than restricting to models
that are fit and evaluated in separate steps, we chose to allow simple
statistical summaries such as mean survival {[}11{]} and survival curves
{[}12{]}. These approaches were novel to CBR and presage later
developments, so were helpful to understanding the development of
localized modeling. However, this then admitted studies that applied
conventional nearest neighbors prediction: Each retrieved cohort
consisted of the \(k\) most similar cases to the index case, and the
response for the index case was predicted to be either the mean
(continuous response) or the plurality (discrete response) of the
responses of the cohort. A review of all studies that use nearest
neighbors prediction, even in the clinical domain, would be impractical.
Because our focus is on novel approaches that combine similarity-based
retrieval and statistical modeling of retrieved cohorts, we chose to
exclude those studies whose approach was equivalent to nearest neighbors
prediction using a conventional similarity measure.

Finally, one author (JCB) applied the same review process to the seed
set of 43 studies that inspired the review. This obtained 6 additional
studies that had not been retrieved by any of our database searches
{[}12--17{]}.

\hypertarget{coding}{%
\subsection{Coding}\label{coding}}

Following the selection sources aligning with our review, we collected
characteristics of included studies (Section \ref{sec:availability}).
The features included bibliographic fields (date of publication,
journal, authors, title, keywords, DOI), study goals (objective,
generalizable knowledge, evaluation, clinical/medical domain), data sets
(data source, type of data, range of data, number of cases/incidences,
number of predictors/features), and methodological choices (types of
similarity measure, families of adaptation step/statistical model,
method(s) compared against, evaluation/performance/comparison measures
(\& their values), name given to modeling approach). We used these data
to detect and visualize trends amongst the included studies.

\hypertarget{synthesis}{%
\subsection{Synthesis}\label{synthesis}}

Rather than an evidence synthesis characteristic of most systematic
reviews, we here pursue a methodology synthesis to harmonize possibly
independent research efforts that have converged on a common modeling
technique. The goal will be to describe a unified framework for
localized models that can be used to guide future study designs and
implementations as well as more systematically evaluate variations on
the theme and measure the dependence of results on modeling choices.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{selection}{%
\subsection{Selection}\label{selection}}

Figure \ref{fig:prisma} depicts our identification of studies via
databases and registers. Following the completion and input of each
search string, there were a total of 1,422 sources within all of the
platforms used. De-duplication resulted in 1,309 entries, which were
added to the title/abstract screening for review. Of these, 328 entries
met the screening criteria and were assessed through full text review.
Of these, 44 fit the original criteria, and 19 were included as distinct
from NN prediction.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-prisma}

}

\caption{\label{fig:prisma}PRISMA-S flow chart. Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).}\label{fig:fig:prisma}
\end{figure}

We then reviewed 43 studies comprising a seed set that inspired this
review. After removing duplicates and screening for eligibility, we were
left with 6 additional studies, 1 of which was excluded from the
synthesis for using nearest neighbors prediction. Finally, citations in
reference lists from the 43 + 328 = 371 studies assessed for eligibility
led us to identify 1 additional study that met our criteria {[}18{]}.
This left us with 19 + 1 + 5 = 25 studies included in the review and
synthesis. Inter-rater reliability was low, at 82\% relative to a 72\%
probability of random agreement.

\hypertarget{bibliographic-and-methodological-properties}{%
\subsection{Bibliographic and methodological
properties}\label{bibliographic-and-methodological-properties}}

The 25 included studies were analyzed based on 20 characteristics (see
Section \ref{sec:availability}), and we report and describe some
observations here (Table \ref{tab:synthesis}). The studies were
published in a variety of journals, with some of greater frequency,
though no single journal published more than 3. The journal
\emph{Evolving Systems} published 2 of the included studies,
\emph{Artificial Intelligence in Medicine} published 3, and
\emph{Hindawi - Journal of Healthcare Engineering} published 3.

Another interesting pattern lay in the years of publication of the
included studies (Figure \ref{fig:year}). While they stretch back to the
late 1990s, they have become more common, which suggests that this is
still an active, though not rapidly expanding, methodological approach.
We note that, while CBR in health and medicine originated as early as
1990, these studies did emerge early on once CBR had established itself.
So the idea has been ``in the air'' for as long as CBR has been in use.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1215}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1602}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1215}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1823}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1547}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1050}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1547}}@{}}
\caption{\label{tab:synthesis}Studies included in the method synthesis,
arranged by the earliest the study is known to have been
public.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Task
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Aim
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Cases
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Features
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Task
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Aim
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Source
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Cases
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Features
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
{[}19{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,355 & 9 \\
{[}11{]} & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
Clinical\hspace{6em} & Longitudinal\hspace{6em} & 113 & 4 \\
{[}20{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 160 & 14 \\
{[}13{]} & Diagnosis\hspace{6em} & Practice\hspace{6em} & Clinical,
Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 366; 270; 560; 760 &
35; 14; 31; 8 \\
{[}21{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
6 \\
{[}22{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 2,620 & 6 \\
{[}23{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 67 & 14 \\
{[}24{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 871 & 4 \\
{[}18{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62 & 2,000 \\
{[}25{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62; 72; 77; 181 &
2,000; 7,129; 7,129; 12,533 \\
{[}12{]} & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 13,525 & 13 \\
{[}26{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 74 & 93 \\
{[}27{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Patient-reported\hspace{6em} & Cross-sectional\hspace{6em} & 1,647 &
18 \\
{[}28{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 150 & 83 \\
{[}15{]} & Clustering\hspace{6em} & Practice\hspace{6em} & Clinical,
Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 7,519 & 130 \\
{[}14{]} & Prediction\hspace{6em} & Knowledge\hspace{6em} & Clinical,
Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
{[}29{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,046 & 8 \\
{[}16{]} & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
{[}30{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 638 & 49,728 \\
{[}31{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
{[}17{]} & Diagnosis\hspace{6em} & Practice\hspace{6em} & Clinical,
Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 16,490 & 106 \\
{[}32{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 4,000 & 22 \\
{[}33{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\
{[}34{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 8 & 5 \\
{[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\
\end{longtable}

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-years}

}

\caption{\label{fig:year}Number of publications in our sample each year.}\label{fig:fig:year}
\end{figure}

Another dominant characteristic of included studies is their remit. We
categorized studies as ``knowledge'' or ``practice'' based on whether
their aim was to produce generalizable knowledge or to improve practice.
We classified 7 of the 25 studies as ``knowledge'', making the majority
of studies from a ``practice'' standpoint. These studies aimed primarily
to provide providers with tools to use in clinic or to improve outcomes
for patients currently receiving care.

Lastly, we observed a commonality among 6 of the 25 included studies in
having evaluated methods using leave-one-out cross-validation (LOOCV).
The purpose of this measure is to estimate the overall performance of
certain factors when used to make predictions, particularly utilized on
smaller data sets, where models benefit greatly from larger training
sets and additional model fitting is less costly.

In addition to their remit, we coded several aspects of the design of
each study, including the source and type of data and the clinical task
the model performed (Table \ref{tab:synthesis}), as well as the specific
methodology and terminology adopted (Table \ref{tab:composite}). In the
following subsections, we take a closer look at these design elements.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1218}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4365}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4416}}@{}}
\caption{\label{tab:composite}Methodological elements of studies
included in the synthesis.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
{[}11{]} & predictive modeling on similarity cohorts & CBR; targeted
subsampling \\
{[}19{]} & supervised similarity learning & CBR; case-structured
retrieval \\
{[}20{]} & unsupervised similarity learning & Combined Kohonen type
neural network-case-based reasoning \\
{[}21{]} & supervised similarity learning & Transductive inference \\
{[}22{]} & unsupervised similarity learning & CBR; entropic distance
measure \\
{[}23{]} & tiered constraint; similarity matching & CBR \\
{[}24{]} & interactive implementation & CBR; eXiT*CBR \\
{[}27{]} & supervised similarity learning & CBR \\
{[}28{]} & supervised similarity learning & CBR \\
{[}25{]} & predictive modeling on similarity cohorts & Local modeling;
personalized modeling; TWNFI \\
{[}26{]} & predictive modeling on similarity cohorts & Personalized
modeling; TWNFI \\
{[}29{]} & supervised similarity learning; predictive modeling on
similarity cohorts & CBR \\
{[}30{]} & unclear & CBR \\
{[}31{]} & whole-population weight learning & Gaussian processes \\
{[}32{]} & predictive modeling on similarity cohorts & Personalized
model \\
{[}34{]} & descriptive modeling on similarity cohorts & CBR \\
{[}33,35{]} & supervised similarity learning; predictive modeling on
similarity cohorts & Similarity model, personalized model, precision
cohort, personalized treatment options \\
{[}13{]} & supervised cohort construction & CBR, statistical case-based
reasoning \\
{[}12{]} & predictive modeling on similarity cohorts & K-nearest
neighbors survival \\
{[}14{]} & predictive modeling on similarity cohorts & Personalized
prediction \\
{[}15{]} & predictive modeling on similarity cohorts; explanatory
modeling on similarity cohorts & Personalized predictive model \\
{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
predictive model \\
{[}17{]} & predictive modeling on similarity cohorts & Personalized
predictive modeling \\
{[}18{]} & predictive modeling on similarity cohorts & Personalized
model \\
\end{longtable}

\hypertarget{application-domain}{%
\subsection{Application domain}\label{application-domain}}

While all included studies were reported in scientific and medical
journals, the vast majority were oriented toward clinical practice
rather than medical research. For example, a 2006 study specifically
evaluated the usefulness of CBR-based explanations for the purpose of
decision support {[}36{]}. This study was part of a much larger
literature on CBR systems and was included here despite relying on NN
prediction because it used a non-traditional voting scheme to generate
recommendations. A more recent study took essentially the same focus
with respect to a proposed clinical risk prediction model, which
amounted to CBR with a novel weighting scheme on predictors informed by
expert consensus {[}37{]}. In both cases a prototype implementation was
deployed in an experimental setting for evaluation.

The most common clinical motivations were individualized detection or
diagnosis, prognosis or outcome prediction, and treatment or care
recommendation. The plurality focused on prognosis or outcome
prediction, often using time-to-event analysis: {[}11{]} used CBR to
predict survival time from several geometric properties of breast
tumors. {[}12{]} used CBR with non-parametric survival models on
registry data to predict patient--graft survival times following kidney
transplantation. {[}14{]} and {[}16{]} used localized logistic
regression and random forest modeling to predict 30-day mortality
following discharge for ICU patients. {[}29{]} developed a CBR cycle
around a clustering-informed similarity matching procedure and an
artificial neural network--based classifier to identify thrombophilia
patients at high risk of thrombotic episodes. And {[}32{]} took a
similarity cohort--based approach to predicting length of stay for ICU
patients. Also of note, from a public health perspective, {[}23{]} used
CBR to predict rehabilitation time as well as disability risk for
unemployed workers experiencing chronic pain.

Toward detection and diagnosis, {[}20{]} proposed a hybrid neural
net--CBR system to distinguish (with confidence bounds) arthritic versus
control patients, based on several histological features. {[}28{]} used
collaborative multilabel CBR to subtype melanoma patients based on
confocal and dermoscopy images. {[}17{]} used localized models built
from a multi-type additive similarity measure to distinguish type 2
diabetic versus control populations. Along the way, {[}21{]}, {[}18{]},
and {[}26{]} took an iterative model-building approach to several tasks:
predicting glomerular filtration rate, a key indicator of renal
function, from demographic and physiological variables; identifying
patients with colon cancer using a large number of gene expression
measurements; and identifying patients with type 2 diabetes based on
demographic, physiological, and genetic variables.

While several studies emphasized the potential or actual value to
decision-making of their methods and tools, only one incorporated
treatment decisions into their approach: By taking ``decision points''
as their units of analysis, {[}33{]} and {[}35{]} built not classifiers
or predictors but comparative effectiveness models into a localized
framework, providing for the first time in our sample explicitly
prescriptive rather than descriptive clinical decision support.

A partial exception to this focus was a 2014 study that also reported a
decision support tool, in this case for early diagnosis of melanoma from
clinical data and dermoscopy images {[}28{]}. While the stated
objectives were analogous, specific emphasis was placed on the
acquisition of new knowledge through the development of the tool,
including the systematic generation of new data and creation of a
clinical ontology. This study was included for its use of a
collaborative classifier that drew from multiple modeling approaches. In
keeping with this focus of the included literature, the stated
objectives of the proposed methods were more often (or additionally) to
predict outcomes or to recommend interventions than only to diagnose
disease.

The mosaic plot in Figure \ref{fig:properties} summarizes the relation
between data source, clinical task, and remit.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../../fig/fig-properties}

}

\caption{\label{fig:properties}Share of studies characterized by several design elements: data source (row), clinical task (fill color), and remit (opacity).}\label{fig:fig:properties}
\end{figure}

\hypertarget{method-development}{%
\subsection{Method development}\label{method-development}}

Consistent with their orientation, almost all included studies were
conducted using clinical data, only occasionally together with
patient-reported (2), laboratory (2), and image (1) data. We suggest
three reasons for this: First, this literature traces back to the 1990s,
before -omic data could be generated cost-effectively at scale. Second,
CBR in particular has a strong tradition in clinical decision support,
where the focus of our sample remains throughout the review period.
Third, because most -omic data are highly homogeneous---all measurements
are made along or can be transformed to a common scale, e.g.~greyscale
pixellations for X-ray images and transcripts per million for RNA-seq
data---more deeply theoretical analysis techniques have been developed
and come into wide use. While variations on correlation-based approaches
like EHR-based phenome-wide association studies and similarity-based
methods like CBR itself have been developed, more mechanistic and
probabilistic tools have not become a domain standard.

CBR can be understood as the opposite side of a trade-off with
rule-based reasoning between model size and model complexity; {[}36{]}
describe their approach as ``knowledge-light'', in that ``the cases do
not contain explicit explanation structures; instead, explanation is
achieved by comparison of the query case with retrieved cases''. This
means that the greatest performance and efficiency challenges in CBR
have to do with the retrieval and revision phases in the R4 cycle.
Several studies addressed these challenges: {[}13{]}, while excluded
from the synthesis, were the earliest to propose that cohorts be bounded
by a similarity threshold rather than by a number of cases, and this
improved performance in their experiments. {[}38{]} leveraged predictor
weights obtained from logistic regression to inform the similarity
calculation used in retrieval. {[}12{]} used non-parametric models to
reduce the computational burden of revision (prediction). {[}32{]}
proposed to improve efficiency along the entire R4 cycle, but in
particular by only executing task-dependent steps in real time
(``just-in-time learning''), and {[}35{]} and {[}33{]} partitioned
multiple phases in the R4 cycle into offline and online components, only
the latter of which would be performed in real time as new data are
received.

Other studies addressed limitations of available tools. {[}24{]}
implemented a comprehensive CBR tool in response to the lack of
general-purpose software, to enable coupling with other tools as well as
to expedite development, experimentation, and uptake. Several other
teams also set out to develop more interoperable clinical support tools
{[}15,22,25,31{]}.

The frequency plot in Figure \ref{fig:methods} compares the rates at
which several methodological elements are invoked in the sample.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../../fig/fig-methods}

}

\caption{\label{fig:methods}Frequency of recurring methodological elements across included studies.}\label{fig:fig:methods}
\end{figure}

In the Discussion, we discuss trends in the literature we reviewed,
identify common limitations and gaps in understanding, and attempt to
synthesize their methods into a general framework.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{process}{%
\subsection{Process}\label{process}}

The low inter-rater reliability was due, in parts, to different
interpretations of some eligibility criteria by the authors,
inconsistent terminology across the sample, and incomplete reporting of
resources and methods in the sample. Regarding interpretation of
criteria, some wording of the criteria was adjusted following
discussions between the reviewing authors during full-text review to
better specify an agreed-upon meaning. We will discuss the different
domains, terminologies, and reporting issues of the sample in the
remainder of this section.

\hypertarget{trends}{%
\subsection{Trends}\label{trends}}

We draw several observations from our encoding of the included studies.
We first focus on two driving questions: What challenges did
investigators set out to solve? How did their approaches meet these
challenges? We then comment on the connectedness of the included studies
and the potential for recombination.

\textbf{Add 1--2 paragraphs.}

Because we excluded studies that used conventional NN prediction, many
included studies reported new approaches to adaptation subsequent to
retrieval. Very few proposed novel similarity measures, possibly because
larger studies tended to be reported in separate articles detailing
experiments with specific components of the process. However, this also
suggests that few experimental studies have focused on the unified
development of new retrieval and adaptation strategies. We also note
that the majority of studies evaluated and compared methods using
leave-one-out cross-validation (LOOCV). This is an especially
appropriate technique when available data are scarce, and indeed most
data sets used by included studies numbered in the hundreds of cases or
fewer. This likely follows from the older age of many included studies
and from their consistent primary focus on clinical data, which is more
costly to collect and comes with more restrictions on its use.

\hypertarget{coherence}{%
\subsubsection{Coherence}\label{coherence}}

The motivational and methodological unity of these studies does not
reflect a unified research program. Besides the lack of any primary
journal of record, we observed collaborations only among the authors of
smaller contiguous programs, including applications of the TWNFI
methodology {[}21,26{]}, individualized mortality prediction for ICU
patients {[}14,16{]} and the use of more explanatory models to
prioritize predictors or treatments for chronic disease {[}15,33,35{]}.

These programs used significantly varying terminology for common
concepts, and in particular no common term is in use for what we here
refer to as localized modeling; authors described this kind of method as
``targeted prognosis'' {[}11{]}, ``transductive inference'' {[}21{]},
``personalized decision support'' {[}14{]}, ``personalized (predictive)
models'' {[}25, in contrast to ``local models,''15,17,32{]}, and
``precision cohort'' analysis {[}17,33,35{]}. We find uses of
``targeted'', ``personalized'', and ``precision'' generic and imprecise
to this approach, while transductive inference is an established term
for a broader set of methods in ML. Because local naturally contrasts
with global, we propose ``localized models'' as a suitable term for this
counterpart to globally-fitted models.

\hypertarget{synthesis-1}{%
\subsection{Synthesis}\label{synthesis-1}}

The 25 studies included in our synthesis used composite techniques that
we organized into two major types. One type, similarity learning, was
used in 8 studies that tie in to the much larger literature on metric
learning {[}39{]}. These techniques are designed to detect structure in
data, especially associations between predictors and responses, and to
use this structure to inform the definition of a patient similarity
measure. Seven studies used response values in training data sets to
supervise similarity learning while two used unsupervised learning
(Table \ref{tab:composite}). Based on the key properties of metric
learning algorithms identified by {[}40{]}, most learned measures were
non-linear while some were local, and some were optimized globally while
others locally. However, not all qualified as metric learning, since the
resulting measures would not necessarily satisfy the triangle
inequality. Each then used its similarity measure to retrieve training
cases relevant to each testing case. Once retrieved, most took a
standard NN approach to generating predictions from the cohort; two
exceptions {[}29,33{]} proceeded to fit predictive models to the
retrieved cohorts and are discussed below.

\hypertarget{similarity-learning-and-threshold-optimization}{%
\subsubsection{Similarity learning and threshold
optimization}\label{similarity-learning-and-threshold-optimization}}

The similarity learning approaches took a variety of forms. Several
studies used unsupervised similarity learning, for example the
Mahalanobis distance {[}12{]} and the Kolmogorov entropy-based distance
{[}22{]}. In particular, {[}19{]} defined similarity as a weighted sum
of differences in predictor values and used linear regression to
optimize the weights for the predictive accuracy of the cohort they
retrieve. Others used supervised learning: {[}21{]} proposed an
iterative algorithm to optimize a set of fuzzy inference rules used to
retrieve relevant cases and generate a prediction, using
back-propagation on the rules' parameters. Their method was used in
several later studies returned by our search, which are not discussed
here because they did not originate the technique. {[}28{]} explicitly
appeal to a supervised distance metric learning technique {[}41{]} to
obtain a similarity measure on their set of dermoscopy and confocal
images that most effectively separates malignant from benign melanoma
tumors. We will discuss the remaining uses of supervised similarity
learning shortly.

Distinct from but related to similarity learning was supervised cohort
construction. This was an adaptive step taken by {[}13{]} to optimize
the predictive performance of models fitted to cohorts retrieved using a
fixed similarity rather than count threshold, a counterpart to
conventional CBR they called ``statistical CBR''. The step used a
heuristic procedure to locate a similarity threshold that (locally)
maximizes predictive accuracy on the training set. This is analogous to
optimizing the neighborhood size parameter in NN predictive modeling, so
we will not discuss it further except to mention that the authors found
statistical CBR to outperform conventional CBR on several data sets.
{[}27{]} combined these learning techniques: They used logistic
regression on the training set to obtain weights for the predictors (by
predicting the binary outcome of kidney transplant waitlist
registration) and for the cases (by predicting agreement of outcome
between an index case and other training cases), and they used
exhaustion to optimize the size of the retrieved cohort for the accuracy
of the NN model using the previously optimized weights.

As noted above, these instances of similarity learning fit into a much
larger literature on metric learning, which has seen widespread use in
health informatics and modeling. That these studies satisfied our review
criteria is due to the novelty of the techniques at the time of
publication and the detailed attention paid to their techniques by the
authors. It also reflects a limitation of our study and of the
literature we set out to retrieve: We are aware of no standard terms in
use to identify the family of techniques that fit statistical models to
similarity-based cohorts. The retrieved studies whose use of such
techniques alone satisfied our criteria variably termed them
individualized, personalized, local, and patient-specific models, among
other terms (Table \ref{tab:composite}). We prefer the term localized
models, which is both sensitive and specific to uses of a similarity or
distance measure to retrieve a cohort to which a model is fit: It
includes some properly included techniques we did not have in mind
{[}29{]} yet excludes other techniques commonly referred to as
individualized, personalized, or precision.

\hypertarget{localized-models}{%
\subsubsection{Localized models}\label{localized-models}}

In addition to {[}12{]}, {[}25{]}, {[}15{]}, {[}16{]}, {[}29{]},
{[}33{]}, and {[}35{]}, five other studies employed localized models:
{[}11{]}, {[}14{]}, {[}26{]}, {[}17{]}, and {[}32{]}. In most cases
these models were purely predictive in application; of the exceptions,
{[}11{]} produced purely descriptive models of survival outcomes, which
were evaluated for their precision rather than their accuracy, while
{[}15{]} used generalized regression models descriptively as well as
predictively, as did some of the same authors {[}33,35{]} later. The
most common approach to cohort selection was to optimize a size
threshold via manual exploration {[}11,14--17,29{]} or via
cross-validation {[}12,26{]}. Further in the former direction, one study
{[}32{]} set a fixed cohort size while two {[}25,33,35{]} devised more
sophisticated algorithms to balance multiple desiderata including
predictive accuracy.

Several themes emerged from this sample. Foremost was the optimization
of retrieved cohort sizes for some measure of model performance. Despite
the early recommendation by {[}13{]} to threshold cohorts by similarity
rather than by cardinality, only the one previously published study
{[}11{]} and the most recent study {[}33,35{]} in this corpus took this
approach. Viewed as a hyperparameter of the individualized modeling
approach, cohort size can be treated in the same way as neighborhood
size in NN prediction (viewed here as a special case of localized
modeling wherein the model is a simple summary statistic), so we will
not discuss it further.

An alternative approach was to optimize multiple hyperparameters
together: {[}18{]} proposed an iterative algorithm to settle on an
optimal number and set of predictors as well as number of training cases
(comprising the retrieved cohort) that was later used by {[}25{]}. More
recently, {[}33{]} and {[}35{]} proposed a three-step process for cohort
selection that filtered by exact match for one subset of predictors,
used domain-informed similarity measures to rank these, and optimized
the similarity threshold for a trade-off between cohort size and a
measure of bias called cohort balance. Notably, though their similarity
measure was supervised, this optimization process was not.

With the exception of the descriptive survival models of {[}11{]}, only
one thread in this corpus concerned itself with the interpretability of
localized models: {[}15{]} fit logistic regression models to
similarity-based cohorts and examined not only their predictive
performance but the localized sets of largest and most detectable
predictors, termed risk profiles, and how they differed across the
testing population. This approach informed their later use of localized
comparative effectiveness--style models to recommend courses of
treatment at decision points during a monitored patent's stay
{[}33,35{]}. These studies suggest much wider potential for localized
real-world evidence generation, which has long been promoted as a
promise of advances in clinical and health informatics {[}1{]}.

\hypertarget{a-general-framework}{%
\subsubsection{A general framework}\label{a-general-framework}}

This relatively small sample exhibited a wide range of approaches, both
technically and conceptually. Discussion of the technical diversity of
these approaches has been summarized above and published in greater
detail in previous reviews {[}6,9,10{]}, and we focus here on the
conceptual. The design of a localized modeling approach can be
decomposed into three largely independent choices: (1) how to measure
the similarity between cases, (2) how to retrieve a cohort of cases
similar to an index case, and (3) how to generate a prediction or other
statistical insight for the index case from the similarity cohort. The
choices are the \emph{similarity measure}, the \emph{retrieval step},
and the \emph{statistical model}, respectively. These are diagrammed in
Figure \ref{fig:framework}.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-typology}

}

\caption{\label{fig:framework}General framework for localized models.}\label{fig:fig:framework}
\end{figure}

Each step can be unsupervised supervised: Unsupervised similarity
measures include discrete measures like the Levenshtein distance and
continuous measures like cosine similarity, while supervised measures
include the Mahalanobis distance and random forest proximity as well as
several composite measures with weights calculated from the data. Most
retrieval steps obtained cohorts with a uniform size (cardinality) or
similarity threshold, and the remaining were likewise only informed by
predictors; thus, while in principle a retrieval step could be
supervised, in this sample none were. Almost all models were supervised,
being that most performed predictive tasks, though {[}11{]} modeled only
survival rates within localized cohorts. Meanwhile, each step can be
optimized in a ML fashion by having its parameters tuned to improve
performance. We found no studies that tuned the calculation of
similarity, though most uniform cohort sizes were tuned. Some earlier
studies tuned generalized regression models, but more recent studies
have not.

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

We propose the term \emph{localized modeling} to encompass an approach
derived from CBR in which parameterized models are fitted in a
standardized way to nearest neighborhoods of past or training cases
according to a measure of patient similarity. We conducted a systematic
search for studies that apply localized models to tasks involving health
data and synthesized these largely independently developed approaches
into a general framework. While the search was limited by low
inter-rater reliability and failure to recover several motivating
examples, the included studies used many of the same underlying tools to
build, optimize, and evaluate their methods. We therefore believe that
our framework can serve to taxonomize ongoing work of this type and
inform the development of customizable implementations. Indeed, the
availability of increasing computational power, the diversity of tasks
to which these models were applied and of technical specifications they
employed, and the apparent lack of any multi-group research program to
date suggest great potential for growth. Whereas precious few of the
reviewed studies used these models, for any task other than prediction,
despite widespread suspicion among clinicians of ``black-box'' models
and growing interest in interpretable alternatives, we recommend that
future work put greater emphasis on the development and validation of
interpretable localized models and on their reception by communities of
medical research and clinical practice.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

\hypertarget{support}{%
\section{Support}\label{support}}

The authors received no specific funding or other support for this work.

\hypertarget{competing-interests}{%
\section{Competing interests}\label{competing-interests}}

The authors declare that they have no competing interests.

\hypertarget{data-and-code-availability}{%
\section{Data and code availability}\label{data-and-code-availability}}

\label{sec:availability}

Search results are publicly available in a Zotero group library. Data
collected or encoded for included studies are publicly available in two
Google Sheets. Code used to analyze these data and generate this
manuscript is publicly available in a GitHub repository.

\begin{itemize}
\tightlist
\item
  Search results, included studies, and other bibliography:
  \url{https://www.zotero.org/groups/5017571/imsr/}
\item
  Bibliographic and methodological properties of included studies:
  \url{https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/}
\item
  Terminology and composite techniques of included studies:
  \url{https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/}
\item
  Code used to perform analyses and prepare the manuscript:
  \url{https://github.com/corybrunson/imsr}
\end{itemize}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{research-question}{%
\subsection{Research question}\label{research-question}}

How is patient similarity--based individualized modeling conducted using
retrospective data?

\hypertarget{purpose-of-review}{%
\subsection{Purpose of review}\label{purpose-of-review}}

\label{sec:appendix-purpose}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provide a summary of individualized models to date.
\item
  Lay the groundwork for conducting a comparison study of individualized
  models.
\item
  Provide a framework for future individualized modeling studies.
\end{enumerate}

\hypertarget{search-1}{%
\subsection{Search}\label{search-1}}

\label{sec:appendix-search}

The procedure for formulating the search began with an evaluation of the
research question. We highlighted specific elements within our topic of
interest that we found critical to our search and listed them using an
OR of ANDs, pairing terms we believed would provide our desired result.
Following the solidification of the search, a thesaurus was created in
which each term was expanded by synonyms that are similar enough to our
core term to be applicable to our search. The expanded search string was
then evaluated using the PubMed Advanced Search platform. We initially
included each term and their synonyms separately to evaluate what
resulted. Several terms were eliminated due to PubMed classifying them
as ``phrases not found'' and other terms were removed to reduce the
number of results, providing a more concise list of results. At the
conclusions of this process for each individual term, we combined each
search term using an OR of ANDs to ultimately form our search.

The search will have been designed to recover studies of the kind
reviewed by the review papers from which we obtained our ``seed set''.
To validate the final search design, we will determine how many of the
papers in this seed set that are indexed by PubMed are actually
recovered by our search. In most cases, the focus of a review paper is
different from ours, so we will only perform this validation test on the
seed set obtained from two review papers that are (a) closest in focus
to ours and (b) use terminology associated with the two distinct
sub-literatures relevant to our focus: Choudhury \& Begum (2016), which
focuses on case-based reasoning in medicine, and Sharafoddini, Dubin, \&
Lee (2017), which focuses on patient similarity--based prediction models
on health data. The proportion of each PubMed-indexed seed set that is
recovered from our PubMed search provides a rough and optimistic yet
useful estimate of the proportion of the relevant literature that our
full search strategy will recover.

Once we have finalized the search as a logical pattern, we will take the
following steps to generate the sample/corpus of literature that will be
the starting point for our selection process.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The logical pattern will be converted to a search string using the
  syntax appropriate to each database in our search strategy. These
  include PubMed (already done as part of the search design), Web of
  Science, Academic Search Premier/Elite, and Mathematical Reviews.
\item
  The search will be conducted on each database and the results
  organized into a Zotero collection, with one subcollection for each
  database.
\item
  Duplicate results will be identified and merged. (A result obtained
  from multiple databases should have only one Zotero entry but should
  be filed under the subcollection for each database in which it was
  found.)
\end{enumerate}

Following discussion among AC, PMJ, and JCB, we omitted results from
Google Scholar due to missing abstracts, missing URLs, high overlap with
other search results, and irreproducibility of the search process.

\hypertarget{search-strings}{%
\subsubsection{Search strings}\label{search-strings}}

Here we reproduce the search strings and platform specifications used in
our literature search. We first finalized the \textbf{PubMed} search
string below:\footnote{\url{https://pubmed.ncbi.nlm.nih.gov/?term=(+\%22case-based+reasoning\%22+\%5BAll+Fields\%5D+OR+\%22case-based+system\%22+\%5BAll+Fields\%5D+)+OR+(+\%22individualized+modeling\%22+\%5BAll+Fields\%5D+OR+\%22personalized+modeling\%22+\%5BAll+Fields\%5D+OR+\%22customized+modeling\%22+\%5BAll+Fields\%5D+)+OR+(+\%22individualized+cohort\%22+\%5BAll+Fields\%5D+)+OR+(+(+\%22patient+similarity\%22+\%5BAll+Fields\%5D+OR+\%22patient+distance\%22+\%5BAll+Fields\%5D+OR+\%22patient+connection\%22+\%5BAll+Fields\%5D+OR+\%22patient+affinity\%22+\%5BAll+Fields\%5D+OR+\%22patient+clustering\%22+\%5BAll+Fields\%5D+)+AND+(+\%22cohort+study\%22+\%5BAll+Fields\%5D+)+)\&sort=date}}

\begin{verbatim}
(
  "case-based reasoning" [All Fields] OR
  "case-based system" [All Fields]
) OR (
  "individualized modeling" [All Fields] OR
  "personalized modeling" [All Fields] OR
  "customized modeling" [All Fields]
) OR (
  "individualized cohort" [All Fields]
) OR (
  (
    "patient similarity" [All Fields] OR
    "patient distance" [All Fields] OR
    "patient connection" [All Fields] OR
    "patient affinity" [All Fields] OR
    "patient clustering" [All Fields]
  ) AND (
    "cohort study" [All Fields]
  )
)
\end{verbatim}

This search yielded 423 results.

We then generated analogous search strings or search strategies for the
Web of Science, Academic Search Premier, and MathSciNet platforms, based
on the logics and syntaxes of their respective interfaces.

For \textbf{Web of Science}, we searched for several separate
disjunctions derived from the PubMed search string. The separate search
strings and the number of results obtained using each are below.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4444}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Search string
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Number of items
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{"case-based\ reasoning"\ OR} \texttt{"case-based\ system"} &
3,636 \\
\texttt{"individualized\ model"\ OR}
\texttt{"individualized\ modeling"\ OR}
\texttt{"personalized\ model"\ OR} \texttt{"personalized\ modeling"\ OR}
\texttt{"customized\ model"} \texttt{"customized\ modeling"} & 444 \\
\texttt{"individualized\ cohort"} & 1 \\
\texttt{"patient\ similarity"\ OR} \texttt{"patient\ distance"\ OR}
\texttt{"patient\ connection"\ OR} \texttt{"patient\ affinity"\ OR}
\texttt{"patient\ clustering"}; Refined search: \texttt{"cohort"} &
48 \\
\end{longtable}

We found the results of the first search string to be predominantly
irrelevant. To reduce review time, these were dropped. The last search
was refined with an additional term following the initial disjunctive
search. Our searches on Web of Science thus yielded 493 results.

When searching \textbf{Academic Search Premier}, we checked the option
``Scholarly (Peer-Reviewed) Journals'', unchecked the option ``Apply
equivalent subjects'', and searched for several separate disjunctions
and conjunctions of strings in the ``TX All Text'' field. We obtained
the resulting citations via email in RIS format.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4444}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Search string
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Number of items
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{"case-based\ reasoning"\ OR} \texttt{"case-based\ system"} &
3,283 \\
\texttt{"individualized\ modeling"\ OR}
\texttt{"personalized\ modeling"\ OR} \texttt{"customized\ modeling"} &
100 \\
\texttt{"individualized\ cohort"} & 2 \\
\texttt{"patient\ similarity"\ AND} \texttt{"cohort\ study"} & 14 \\
\texttt{"patient\ distance"\ AND} \texttt{"cohort\ study"} & 15 \\
\texttt{"patient\ connection"\ AND} \texttt{"cohort\ study"} & 6 \\
\texttt{"patient\ affinity"\ AND} \texttt{"cohort\ study"} & 0 \\
\texttt{"patient\ clustering"\ AND} \texttt{"cohort\ study"} & 50 \\
\end{longtable}

We found the results of the first search string to be predominantly
irrelevant. To reduce review time, these were dropped. Our searches on
Academic Search Premier thus yielded 187 results.

For \textbf{MathSciNet}, the portal to the \emph{Mathematical Reviews}
database, we used the same separate searches as for Web of Science, in
some cases expanded to obtain more results. Those which yielded nonzero
numbers of results are below:

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4444}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Search string
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Number of items
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{"case-based\ reasoning"} & 110 \\
\texttt{"individualized\ model} & 2 \\
\texttt{"patient\ similarity"\ AND} \texttt{"cohort"} & 1 \\
\end{longtable}

Our searches of \emph{Mathematical Reviews} therefore yielded 113
results.

These totaled 1,422 sources from all platforms. We organized the full
results in a public Zotero collection alongside the seed set and created
a single folder for the 25 results reviewed in detail.

\hypertarget{screening-process}{%
\subsection{Screening process}\label{screening-process}}

\label{sec:appendix-screening}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Deduplication (Covidence)

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Specific to that source: title, creator, URL, DOI
  \item
    May overlap: journal, publication date, ISSN
  \end{enumerate}
\item
  Title

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Exclude non clinical/non medical setting
  \item
    Must be in English
  \item
    Must be original study (not reviews, surveys, opinion, news)
  \item
    Exclude if search term clearly has different meaning than intended
  \end{enumerate}
\item
  Abstract
\item
  Full Text

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Study criteria (inclusion/exclusion)
  \end{enumerate}
\end{enumerate}

\hypertarget{selection-process}{%
\subsection{Selection process}\label{selection-process}}

\label{sec:appendix-selection}

Our selection of relevant papers from the search corpus will be based on
the following inclusion/exclusion criteria:

\begin{itemize}
\tightlist
\item
  Uses labeled case-level (empirical) data set
\item
  Defines a continuous-valued multivariate case similarity measure
\item
  Uses the similarity measure to select cohorts for index cases from the
  corpus
\item
  Fits statistical models to cohorts to make inferences about index
  cases
\end{itemize}

We decided after concluding full-text review to perform one round of
citation-tracking, of citations within the Methods (or analogous)
sections of the included entries.

\hypertarget{analysis-and-synthesis}{%
\subsection{Analysis and synthesis}\label{analysis-and-synthesis}}

Because we could not predict the scope of methodological approaches we
would encounter, we did not prepare specific analyses or syntheses
\emph{a priori}.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Longhurst2014}{}}%
\CSLLeftMargin{{[}1{]} }%
\CSLRightInline{Longhurst CA, Harrington RA, Shah NH. A {``{Green
Button}''} {For Using Aggregate Patient Data At The Point Of Care}.
Health Affairs 2014;33:1229--35.
\url{https://doi.org/10.1377/hlthaff.2014.0099}.}

\leavevmode\vadjust pre{\hypertarget{ref-Aamodt1994}{}}%
\CSLLeftMargin{{[}2{]} }%
\CSLRightInline{Aamodt A, Plaza E. Case-{Based Reasoning}: {Foundational
Issues}, {Methodological Variations}, and {System Approaches}. AI
Communications 1994;7:39--59.
\url{https://doi.org/10.3233/AIC-1994-7104}.}

\leavevmode\vadjust pre{\hypertarget{ref-Begum2011}{}}%
\CSLLeftMargin{{[}3{]} }%
\CSLRightInline{Begum S, Ahmed MU, Funk P, Xiong N, Folke M. Case-{Based
Reasoning Systems} in the {Health Sciences}: {A Survey} of {Recent
Trends} and {Developments}. IEEE Transactions on Systems, Man, and
Cybernetics, Part C (Applications and Reviews) 2011;41:421--34.
\url{https://doi.org/10.1109/TSMCC.2010.2071862}.}

\leavevmode\vadjust pre{\hypertarget{ref-Kolodner1992}{}}%
\CSLLeftMargin{{[}4{]} }%
\CSLRightInline{Kolodner JL. An introduction to case-based reasoning.
Artificial Intelligence Review 1992;6:3--34.
\url{https://doi.org/10.1007/BF00155578}.}

\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
Heidelberg}; 1998, p. 273--97.
\url{https://doi.org/10.1007/3-540-69351-3_11}.}

\leavevmode\vadjust pre{\hypertarget{ref-Choudhury2016}{}}%
\CSLLeftMargin{{[}6{]} }%
\CSLRightInline{Choudhury N, Begum SA. A {Survey} on {Case-based
Reasoning} in {Medicine}. International Journal of Advanced Computer
Science and Applications 2016;7.
\url{https://doi.org/10.14569/IJACSA.2016.070820}.}

\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{Dai L, Zhu H, Liu D.
\href{https://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
applications}. arXiv:201201976 {[}Cs{]} 2020.}

\leavevmode\vadjust pre{\hypertarget{ref-Welch2013}{}}%
\CSLLeftMargin{{[}8{]} }%
\CSLRightInline{Welch BM, Kawamoto K. Clinical decision support for
genetically guided personalized medicine: A systematic review. Journal
of the American Medical Informatics Association 2013;20:388--400.
\url{https://doi.org/10.1136/amiajnl-2012-000892}.}

\leavevmode\vadjust pre{\hypertarget{ref-Sharafoddini2017}{}}%
\CSLLeftMargin{{[}9{]} }%
\CSLRightInline{Sharafoddini A, Dubin JA, Lee J. Patient {Similarity} in
{Prediction Models Based} on {Health Data}: {A Scoping Review}. JMIR
Medical Informatics 2017;5:e7.
\url{https://doi.org/10.2196/medinform.6730}.}

\leavevmode\vadjust pre{\hypertarget{ref-Parimbelli2018}{}}%
\CSLLeftMargin{{[}10{]} }%
\CSLRightInline{Parimbelli E, Marini S, Sacchi L, Bellazzi R. Patient
similarity for precision medicine: {A} systematic review. Journal of
Biomedical Informatics 2018;83:87--96.
\url{https://doi.org/10.1016/j.jbi.2018.06.001}.}

\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{Mariuzzi G, Mombello A, Mariuzzi L, Hamilton PW, Weber
JE, Thompson D, et al. Quantitative study of ductal breast
cancer--patient targeted prognosis: An exploration of case base
reasoning. Pathology, Research and Practice 1997;193:535--42.
\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-Lowsky2013}{}}%
\CSLLeftMargin{{[}12{]} }%
\CSLRightInline{Lowsky DJ, Ding Y, Lee DKK, McCulloch CE, Ross LF,
Thistlethwaite JR, et al. A {K-nearest} neighbors survival probability
prediction method. Statistics in Medicine 2013;32:2062--9.
\url{https://doi.org/10.1002/sim.5673}.}

\leavevmode\vadjust pre{\hypertarget{ref-Park2006}{}}%
\CSLLeftMargin{{[}13{]} }%
\CSLRightInline{Park Y-J, Kim B-C, Chun S-H. New knowledge extraction
technique using probability for case-based reasoning: Application to
medical diagnosis. Expert Systems 2006;23:2--20.
\url{https://doi.org/10.1111/j.1468-0394.2006.00321.x}.}

\leavevmode\vadjust pre{\hypertarget{ref-Lee2015}{}}%
\CSLLeftMargin{{[}14{]} }%
\CSLRightInline{Lee J, Maslove DM, Dubin JA. Personalized {Mortality
Prediction Driven} by {Electronic Medical Data} and a {Patient
Similarity Metric}. PLOS ONE 2015;10:e0127428.
\url{https://doi.org/10.1371/journal.pone.0127428}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ng2015}{}}%
\CSLLeftMargin{{[}15{]} }%
\CSLRightInline{Ng K, Sun J, Hu J, Wang F.
\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4525240}{Personalized
{Predictive Modeling} and {Risk Factor Identification} using {Patient
Similarity}}. AMIA Joint Summits on Translational Science Proceedings
AMIA Joint Summits on Translational Science 2015;2015:132--6.}

\leavevmode\vadjust pre{\hypertarget{ref-Lee2017}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{Lee J. Patient-{Specific Predictive Modeling Using
Random Forests}: {An Observational Study} for the {Critically Ill}. JMIR
Medical Informatics 2017;5:e3.
\url{https://doi.org/10.2196/medinform.6690}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2019}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{Wang N, Huang Y, Liu H, Fei X, Wei L, Zhao X, et al.
Measurement and application of patient similarity in personalized
predictive modeling based on electronic medical records. BioMedical
Engineering OnLine 2019;18.
\url{https://doi.org/10.1186/s12938-019-0718-2}.}

\leavevmode\vadjust pre{\hypertarget{ref-Kasabov2010}{}}%
\CSLLeftMargin{{[}18{]} }%
\CSLRightInline{Kasabov N, Hu Y. Integrated optimisation method for
personalised modelling and case studies for medical decision support.
International Journal of Functional Informatics and Personalised
Medicine 2010;3:236--56.
\url{https://doi.org/10.1504/IJFIPM.2010.039123}.}

\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997a}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{Yearwood J, Wilkinson R. Retrieving cases for treatment
advice in nursing using text representation and structured text
retrieval. Artif Intell Med 1997;9:79--99.
\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{Wyns B, Sette S, Boullart L, Baeten D, Hoffman IEA, De
Keyser F. Prediction of diagnosis in patients with early arthritis using
a combined {Kohonen} mapping and instance-based evaluation criterion.
Artif Intell Med 2004;31:45--55.
\url{https://doi.org/10.1016/j.artmed.2004.01.002}.}

\leavevmode\vadjust pre{\hypertarget{ref-Song2006c}{}}%
\CSLLeftMargin{{[}21{]} }%
\CSLRightInline{Song Q, Kasabov N. {TWNFI} \textemdash{} a transductive
neuro-fuzzy inference system with weighted data normalization for
personalized modeling. Neural Networks 2006;19:1591--6.}

\leavevmode\vadjust pre{\hypertarget{ref-Elter2007}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{Elter M, Schulz-Wendtland R, Wittenberg T. The
prediction of breast cancer biopsy outcomes using two {CAD} approaches
that both emphasize an intelligible decision process. Med Phys
2007;34:4164--72. \url{https://doi.org/10.1118/1.2786864}.}

\leavevmode\vadjust pre{\hypertarget{ref-Xu2008a}{}}%
\CSLLeftMargin{{[}23{]} }%
\CSLRightInline{Xu Y, Chan CCH, Lo KHY-L, Tang D. Prediction model for
the return to work of workers with injuries in {Hong Kong}. Work
2008;30:77--84.}

\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{López B, Pous C, Gay P, Pla A, Sanz J, Brunet J.
{eXiT}*{CBR}: {A} framework for case-based medical diagnosis development
and experimentation. Artif Intell Med 2011;51:81--91.
\url{https://doi.org/10.1016/j.artmed.2010.09.002}.}

\leavevmode\vadjust pre{\hypertarget{ref-Liang2015}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
system for integrated feature, neighborhood and parameter optimization
utilizing gravitational search algorithm. EVOLVING SYSTEMS 2015;6:1--14.
\url{https://doi.org/10.1007/s12530-013-9081-x}.}

\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{Verma A, Fiasche M, Cuzzola M, Irrera G. Comparative
study of existing personalized approaches for identifying important gene
markers and for risk estimation in {Type2 Diabetes} in {Italian}
population. EVOLVING SYSTEMS 2015;6:15--22.
\url{https://doi.org/10.1007/s12530-013-9083-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2012a}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{Campillo-Gimenez B, Bayat S, Cuggia M. Coupling
{K-nearest} neighbors with logistic regression in case-based reasoning.
Stud Health Technol Inform 2012;180:275--9.}

\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014a}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{Nicolas R, Fornells A, Golobardes E, Corral G, Puig S,
Malvehy J. {DERMA}: A melanoma diagnosis platform based on collaborative
multilabel analog reasoning. ScientificWorldJournal 2014;2014:351518.
\url{https://doi.org/10.1155/2014/351518}.}

\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016a}{}}%
\CSLLeftMargin{{[}29{]} }%
\CSLRightInline{Vilhena J, Vicente H, Martins MR, Grañeda JM, Caldeira
F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk. J
Biomed Inform 2016;62:265--75.
\url{https://doi.org/10.1016/j.jbi.2016.07.013}.}

\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
\CSLLeftMargin{{[}30{]} }%
\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
Decision-Making Based} on {Big Clinical Data}. J Healthc Eng
2018;2018:3917659. \url{https://doi.org/10.1155/2018/3917659}.}

\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
\CSLLeftMargin{{[}31{]} }%
\CSLRightInline{Zhang H, Zhu F, Dodge H, Higgins G, Omenn G, Guan Y, et
al. A similarity-based approach to leverage multi-cohort medical data on
the diagnosis and prognosis of {Alzheimer}'s disease. GIGASCIENCE
2018;7. \url{https://doi.org/10.1093/gigascience/giy085}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ma2020a}{}}%
\CSLLeftMargin{{[}32{]} }%
\CSLRightInline{Ma X, Si Y, Wang Z, Wang Y. Length of stay prediction
for {ICU} patients using individualized single classification algorithm.
COMPUTER METHODS AND PROGRAMS IN BIOMEDICINE 2020;186.
\url{https://doi.org/10.1016/j.cmpb.2019.105224}.}

\leavevmode\vadjust pre{\hypertarget{ref-Tang2021d}{}}%
\CSLLeftMargin{{[}33{]} }%
\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
Ng K. Precision population analytics: Population management at the
point-of-care. JOURNAL OF THE AMERICAN MEDICAL INFORMATICS ASSOCIATION
2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
\CSLLeftMargin{{[}34{]} }%
\CSLRightInline{Wang Y, Liang Y, Sun H, Yang Y. Emergency {Response} for
{COVID-19 Prevention} and {Control} in {Urban Rail Transit Based} on
{Case-Based Reasoning Method}. Discrete Dynamics in Nature \& Society
2020:1--9.}

\leavevmode\vadjust pre{\hypertarget{ref-Ng2021b}{}}%
\CSLLeftMargin{{[}35{]} }%
\CSLRightInline{Ng K, Kartoun U, Stavropoulos H, Zambrano J, Tang P.
Personalized treatment options for chronic diseases using precision
cohort analytics. SCIENTIFIC REPORTS 2021;11.
\url{https://doi.org/10.1038/s41598-021-80967-5}.}

\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006}{}}%
\CSLLeftMargin{{[}36{]} }%
\CSLRightInline{Doyle D, Cunningham P, Walsh P. An evaluation of the
usefulness of explanation in a case-based reasoning system for decision
support in bronchiolitis treatment. Comput Intelligence 2006;22:269--81.
\url{https://doi.org/10.1111/j.1467-8640.2006.00288.x}.}

\leavevmode\vadjust pre{\hypertarget{ref-Fang2021a}{}}%
\CSLLeftMargin{{[}37{]} }%
\CSLRightInline{Fang HSA, Tan NC, Tan WY, Oei RW, Lee ML, Hsu W. Patient
similarity analytics for explainable clinical risk prediction. BMC
Medical Informatics \& Decision Making 2021;21:1--12.}

\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2013}{}}%
\CSLLeftMargin{{[}38{]} }%
\CSLRightInline{Campillo-Gimenez B, Jouini W, Bayat S, Cuggia M.
Improving {Case-Based Reasoning Systems} by {Combining K-Nearest
Neighbour Algorithm} with {Logistic Regression} in the {Prediction} of
{Patients}' {Registration} on the {Renal Transplant Waiting List}. PLoS
ONE 2013;8:e71991. \url{https://doi.org/10.1371/journal.pone.0071991}.}

\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
\CSLLeftMargin{{[}39{]} }%
\CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
Survey}. PhD thesis. Michigan State University, 2006.}

\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
\CSLLeftMargin{{[}40{]} }%
\CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
Learning} for {Feature Vectors} and {Structured Data} 2014.
\url{https://doi.org/10.48550/arXiv.1306.6709}.}

\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
\CSLLeftMargin{{[}41{]} }%
\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
Learning} with {Application} to {Clustering} with {Side-Information}.
Advances in {Neural Information Processing Systems}, vol. 15, {MIT
Press}; 2002.}

\end{CSLReferences}

\end{document}

### 2023-08-17 docs/Manuscript/Manuscript.Rmd
add citation numbers to remaining figures + avoid googlesheets in manuscript

--- 2023-08-16 docs/Manuscript/Manuscript.Rmd
@@
 library(stringr)
-googlesheets4::gs4_auth(email = "cornelioid@gmail.com")
 Sys.sleep(10)
@@
 tab_synthesis |> 
-  select(-Date) |> 
+  select(Citation, Task, Aim, Source, Type, Cases, Features) |> 
   # prevent numeric list columns from dominating the horizontal space
@@
 ```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
-googlesheets4::read_sheet(
-  ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 1L
+composite_studies <- read_rds(here::here("tab/composite.rds"))
+composite_studies |> 
+  transmute(
+    # Citation = str_c("@", Citation),
+    Citation = sapply(
+      str_split(Citation, "; "),
+      function(s) str_c("@", s, collapse = "; ")
+    ),
+    Elements = str_replace_all(Elements, ",", ";"),
+    Terminology
   ) |> 
-  rename_with(snakecase::to_snake_case) |>
-  filter(is.na(synthesis)) |> 
-  # mutate(zotero_key = str_split(zotero_key, "; ")) |> 
-  # tidyr::unnest(zotero_key) |> 
-  transmute(
-    # Citation = paste0("@", zotero_key),
-    Citation = sapply(
-      str_split(zotero_key, "; "),
-      function(s) str_c("[", str_c("@", s, collapse = "; "), "]")
-    ),
-    Elements = str_remove_all(approach_type_s, "\\[|\\]"),
-    Terminology = term_s#,
-    # Outcomes = outcome
-  ) |> 
-  mutate(Elements = str_replace_all(Elements, ",", ";")) |> 
-  left_join(select(tab_synthesis, Citation, Date), by = "Citation") |> 
-  arrange(Date) |> 
-  select(-Date) |> 
   mutate(across(

### 2023-08-17 scr/synthesis.r

@@ -25,7 +25,7 @@ googlesheets4::read_sheet(
   inner_join(citation_numbers, by = c("zotero_key" = "ref")) %>%
   transmute(
     Number = number,
-    Citation = paste0("@", zotero_key),
+    Citation = zotero_key,
     Date = format(date_of_publication, "%Y %b %d"),
     Task = objective,
     Aim = ifelse(generalizable_knowledge == "Science", "Knowledge", "Practice"),
@@ -124,16 +124,42 @@ googlesheets4::read_sheet(
   ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 1L
 ) %>% 
   filter(is.na(Synthesis)) %>%
-  select(seq(3L, 5L)) %>%
-  rename_with(~ str_replace(snakecase::to_snake_case(.), "_s$", "s")) %>%
-  mutate(across(c(approach_types, terms), ~ str_split(., "(;|,) *"))) %>%
+  transmute(
+    Reference = Citation,
+    # Citation = paste0("@", `Zotero key`),
+    Citation = sapply(
+      str_split(`Zotero key`, "; "),
+      function(s) str_c(s, collapse = "; ")
+    ),
+    Elements = str_remove_all(`Approach Type(s)`, "\\[|\\]"),
+    Terminology = `Term(s)`,
+    Outcomes = Outcome
+  ) |> 
+  mutate(Elements = str_replace_all(Elements, ",", ";")) |> 
+  left_join(select(tab_synthesis, Citation, Date), by = "Citation") |> 
+  arrange(Date) |> 
   print() -> composite_studies
+# write to file, to be read into document
+write_rds(composite_studies, file = here::here("tab/composite.rds"))
 
 # years of publication
 composite_studies %>%
-  transmute(year = as.integer(str_remove(citation, "^.*, "))) %>%
-  ggplot(aes(x = year)) +
-  geom_bar() +
+  # only first of any multiple reports of same study
+  # mutate(across(c(Reference, Citation), ~ str_remove(., ";.*$"))) %>%
+  # un-nest multiple reports of same study
+  mutate(across(c(Reference, Citation), ~ str_split(., "; *"))) %>%
+  unnest(c(Reference, Citation)) %>%
+  mutate(Year = as.integer(str_remove(Reference, "^.*, "))) %>%
+  left_join(citation_numbers, by = c("Citation" = "ref")) %>%
+  # select(Citation, Year) %>% print(n = Inf)
[…]

### 2023-08-18 docs/Manuscript/Manuscript.Rmd
update zotero keys + make tables consistent

@@
 Sys.sleep(10)
+tab_synthesis <- readr::read_rds(here::here("data/synthesis.rds"))
+composite_studies <- readr::read_rds(here::here("data/composite.rds"))
+citation_id <- readr::read_rds(here::here("data/citations.rds"))
 ```
@@
 
-We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998a; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
+We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
 While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis / dimension reduction / cluster analysis, risk evaluation / outcome prediction, clinical decision support / software tools).
@@
 ```{r tab:synthesis, echo=FALSE, message=FALSE, warning=FALSE}
-tab_synthesis <- readr::read_rds(here::here("tab/synthesis.rds"))
 tab_synthesis |> 
+  # combine citations of the same study
+  left_join(citation_id, by = "Citation") |> 
+  group_by(across(c(-Number, -Citation, -Date))) |> 
+  summarize(Date = min(Date), Citation = str_c(Citation, collapse = "; ")) |> 
+  ungroup() |> 
+  arrange(Date) |> 
   select(Citation, Task, Aim, Source, Type, Cases, Features) |> 
+  # format citations for Markdown-LaTeX
+  # mutate(Citation = str_c("[@", Citation, "]")) |> 
+  mutate(Citation = sapply(
+    str_split(Citation, "; "),
+    function(s) str_c("@", s, collapse = "; ")
+  )) |> 
   # prevent numeric list columns from dominating the horizontal space
@@
 
-```{r fig:year, fig.align='center', out.width="80%", echo=FALSE}
+```{r fig:year, fig.align='center', out.width="75%", echo=FALSE}
 #| fig.cap =
@@
 ```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
-composite_studies <- read_rds(here::here("tab/composite.rds"))
 composite_studies |> 
@@
 While all included studies were reported in scientific and medical journals, the vast majority were oriented toward clinical practice rather than medical research.
-For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006].
+For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006a].
 This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used a non-traditional voting scheme to generate recommendations.
@@
 
-Toward detection and diagnosis, @Wyns2004 proposed a hybrid neural net–CBR system to distinguish (with confidence bounds) arthritic versus control patients, based on several histological features.
+Toward detection and diagnosis, @Wyns2004a proposed a hybrid neural net–CBR system to distinguish (with confidence bounds) arthritic versus control patients, based on several histological features.
 @Nicolas2014a used collaborative multilabel CBR to subtype melanoma patients based on confocal and dermoscopy images.
@@
 
-While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking "decision points" as their units of analysis, @Tang2021d and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
+While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking "decision points" as their units of analysis, @Tang2021c and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
 
@@
 
-CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
+CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006a describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
 This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle.
@@
 @Lowsky2013 used non-parametric models to reduce the computational burden of revision (prediction).
-@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning"), and @Ng2021b and @Tang2021d partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
+@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning"), and @Ng2021b and @Tang2021c partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
 
 Other studies addressed limitations of available tools.
-@Lopez2011 implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
-Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018a].
+@Lopez2011a implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
+Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007a; @Liang2015a; @Ng2015; @Zhang2018a].
 
@@
 The motivational and methodological unity of these studies does not reflect a unified research program.
-Besides the lack of any primary journal of record, we observed collaborations only among the authors of smaller contiguous programs, including applications of the TWNFI methodology [@Song2006c; @Verma2015a], individualized mortality prediction for ICU patients [@Lee2015; @Lee2017] and the use of more explanatory models to prioritize predictors or treatments for chronic disease [@Ng2015; @Tang2021d; @Ng2021b].
+Besides the lack of any primary journal of record, we observed collaborations only among the authors of smaller contiguous programs, including applications of the TWNFI methodology [@Song2006c; @Verma2015a], individualized mortality prediction for ICU patients [@Lee2015; @Lee2017] and the use of more explanatory models to prioritize predictors or treatments for chronic disease [@Ng2015; @Tang2021c; @Ng2021b].
 
-These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021d; @Ng2021b].
+These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015a, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021c; @Ng2021b].
 We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
@@
 Each then used its similarity measure to retrieve training cases relevant to each testing case.
-Once retrieved, most took a standard NN approach to generating predictions from the cohort; two exceptions [@Vilhena2016a; @Tang2021d] proceeded to fit predictive models to the retrieved cohorts and are discussed below.
+Once retrieved, most took a standard NN approach to generating predictions from the cohort; two exceptions [@Vilhena2016a; @Tang2021c] proceeded to fit predictive models to the retrieved cohorts and are discussed below.
 
@@
 The similarity learning approaches took a variety of forms.
-Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007].
+Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007a].
 In particular, @Yearwood1997a defined similarity as a weighted sum of differences in predictor values and used linear regression to optimize the weights for the predictive accuracy of the cohort they retrieve.
@@
 
-In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016a, @Tang2021d, and @Ng2021b, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015a, @Wang2019, and @Ma2020a.
-In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021d; @Ng2021b] later.
+In addition to @Lowsky2013, @Liang2015a, @Ng2015, @Lee2017, @Vilhena2016a, @Tang2021c, and @Ng2021b, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015a, @Wang2019, and @Ma2020a.
+In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021c; @Ng2021b] later.
 The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016a] or via cross-validation [@Lowsky2013; @Verma2015a].
-Further in the former direction, one study [@Ma2020a] set a fixed cohort size while two [@Liang2015; @Tang2021d; @Ng2021b] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
+Further in the former direction, one study [@Ma2020a] set a fixed cohort size while two [@Liang2015a; @Tang2021c; @Ng2021b] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
 
@@
 Foremost was the optimization of retrieved cohort sizes for some measure of model performance.
-Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021d; @Ng2021b] in this corpus took this approach.
+Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021c; @Ng2021b] in this corpus took this approach.
 Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.
 
-An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015.
-More recently, @Tang2021d and @Ng2021b proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance.
+An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015a.
+More recently, @Tang2021c and @Ng2021b proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance.
 Notably, though their similarity measure was supervised, this optimization process was not.
@@
 With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
-This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021d; @Ng2021b].
+This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021c; @Ng2021b].
 These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].

### 2023-08-18 docs/Manuscript/Manuscript.tex
update zotero keys + make tables consistent

@@
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62 & 2,000 \\
-{[}25{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
+{[}25{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 74 & 93 \\
+{[}26{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62; 72; 77; 181 &
@@
 Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 13,525 & 13 \\
-{[}26{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 74 & 93 \\
 {[}27{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
@@
 Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
-{[}30{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
+{[}30{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
+{[}31{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 638 & 49,728 \\
-{[}31{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
-Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
+{[}32{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 4,000 & 22 \\
 {[}17{]} & Diagnosis\hspace{6em} & Practice\hspace{6em} & Clinical,
 Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 16,490 & 106 \\
-{[}32{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 4,000 & 22 \\
 {[}33{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\
-{[}34{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 8 & 5 \\
-{[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\
+{[}34{]}; {[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em}
+& Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 &
+1,466,474 \\
 \end{longtable}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-years}
+{\centering \includegraphics[width=0.75\linewidth]{../../fig/fig-years}
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1218}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4365}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4416}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1128}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4410}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4462}}@{}}
 \caption{\label{tab:composite}Methodological elements of studies
@@
 \endlastfoot
+{[}19{]} & supervised similarity learning & CBR; case-structured
+retrieval \\
 {[}11{]} & predictive modeling on similarity cohorts & CBR; targeted
 subsampling \\
-{[}19{]} & supervised similarity learning & CBR; case-structured
-retrieval \\
 {[}20{]} & unsupervised similarity learning & Combined Kohonen type
 neural network-case-based reasoning \\
+{[}13{]} & supervised cohort construction & CBR, statistical case-based
+reasoning \\
 {[}21{]} & supervised similarity learning & Transductive inference \\
@@
 {[}24{]} & interactive implementation & CBR; eXiT*CBR \\
+{[}18{]} & predictive modeling on similarity cohorts & Personalized
+model \\
+{[}25{]} & predictive modeling on similarity cohorts & Personalized
+modeling; TWNFI \\
+{[}26{]} & predictive modeling on similarity cohorts & Local modeling;
+personalized modeling; TWNFI \\
+{[}12{]} & predictive modeling on similarity cohorts & K-nearest
+neighbors survival \\
 {[}27{]} & supervised similarity learning & CBR \\
 {[}28{]} & supervised similarity learning & CBR \\
-{[}25{]} & predictive modeling on similarity cohorts & Local modeling;
-personalized modeling; TWNFI \\
-{[}26{]} & predictive modeling on similarity cohorts & Personalized
-modeling; TWNFI \\
+{[}15{]} & predictive modeling on similarity cohorts; explanatory
+modeling on similarity cohorts & Personalized predictive model \\
+{[}14{]} & predictive modeling on similarity cohorts & Personalized
+prediction \\
 {[}29{]} & supervised similarity learning; predictive modeling on
 similarity cohorts & CBR \\
-{[}30{]} & unclear & CBR \\
-{[}31{]} & whole-population weight learning & Gaussian processes \\
+{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
+predictive model \\
+{[}30{]} & whole-population weight learning & Gaussian processes \\
+{[}31{]} & unclear & CBR \\
 {[}32{]} & predictive modeling on similarity cohorts & Personalized
 model \\
-{[}34{]} & descriptive modeling on similarity cohorts & CBR \\
-{[}33,35{]} & supervised similarity learning; predictive modeling on
-similarity cohorts & Similarity model, personalized model, precision
-cohort, personalized treatment options \\
-{[}13{]} & supervised cohort construction & CBR, statistical case-based
-reasoning \\
-{[}12{]} & predictive modeling on similarity cohorts & K-nearest
-neighbors survival \\
-{[}14{]} & predictive modeling on similarity cohorts & Personalized
-prediction \\
-{[}15{]} & predictive modeling on similarity cohorts; explanatory
-modeling on similarity cohorts & Personalized predictive model \\
-{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
-predictive model \\
 {[}17{]} & predictive modeling on similarity cohorts & Personalized
 predictive modeling \\
-{[}18{]} & predictive modeling on similarity cohorts & Personalized
-model \\
+{[}33{]} & descriptive modeling on similarity cohorts & CBR \\
+{[}34{]}; {[}35{]} & supervised similarity learning; predictive modeling
+on similarity cohorts & Similarity model, personalized model, precision
+cohort, personalized treatment options \\
 \end{longtable}
@@
 diabetic versus control populations. Along the way, {[}21{]}, {[}18{]},
-and {[}26{]} took an iterative model-building approach to several tasks:
+and {[}25{]} took an iterative model-building approach to several tasks:
 predicting glomerular filtration rate, a key indicator of renal
@@
 treatment decisions into their approach: By taking ``decision points''
-as their units of analysis, {[}33{]} and {[}35{]} built not classifiers
+as their units of analysis, {[}34{]} and {[}35{]} built not classifiers
 or predictors but comparative effectiveness models into a localized
@@
 particular by only executing task-dependent steps in real time
-(``just-in-time learning''), and {[}35{]} and {[}33{]} partitioned
+(``just-in-time learning''), and {[}35{]} and {[}34{]} partitioned
 multiple phases in the R4 cycle into offline and online components, only
@@
 teams also set out to develop more interoperable clinical support tools
-{[}15,22,25,31{]}.
+{[}15,22,26,30{]}.
 
@@
 smaller contiguous programs, including applications of the TWNFI
-methodology {[}21,26{]}, individualized mortality prediction for ICU
+methodology {[}21,25{]}, individualized mortality prediction for ICU
 patients {[}14,16{]} and the use of more explanatory models to
-prioritize predictors or treatments for chronic disease {[}15,33,35{]}.
+prioritize predictors or treatments for chronic disease {[}15,34,35{]}.
 
@@
 ``personalized decision support'' {[}14{]}, ``personalized (predictive)
-models'' {[}25, in contrast to ``local models,''15,17,32{]}, and
-``precision cohort'' analysis {[}17,33,35{]}. We find uses of
+models'' {[}26, in contrast to ``local models,''15,17,32{]}, and
+``precision cohort'' analysis {[}17,34,35{]}. We find uses of
 ``targeted'', ``personalized'', and ``precision'' generic and imprecise
@@
 standard NN approach to generating predictions from the cohort; two
-exceptions {[}29,33{]} proceeded to fit predictive models to the
+exceptions {[}29,34{]} proceeded to fit predictive models to the
 retrieved cohorts and are discussed below.
@@
 
-In addition to {[}12{]}, {[}25{]}, {[}15{]}, {[}16{]}, {[}29{]},
-{[}33{]}, and {[}35{]}, five other studies employed localized models:
-{[}11{]}, {[}14{]}, {[}26{]}, {[}17{]}, and {[}32{]}. In most cases
+In addition to {[}12{]}, {[}26{]}, {[}15{]}, {[}16{]}, {[}29{]},
+{[}34{]}, and {[}35{]}, five other studies employed localized models:
+{[}11{]}, {[}14{]}, {[}25{]}, {[}17{]}, and {[}32{]}. In most cases
 these models were purely predictive in application; of the exceptions,
@@
 {[}15{]} used generalized regression models descriptively as well as
-predictively, as did some of the same authors {[}33,35{]} later. The
+predictively, as did some of the same authors {[}34,35{]} later. The
 most common approach to cohort selection was to optimize a size
 threshold via manual exploration {[}11,14--17,29{]} or via
-cross-validation {[}12,26{]}. Further in the former direction, one study
-{[}32{]} set a fixed cohort size while two {[}25,33,35{]} devised more
+cross-validation {[}12,25{]}. Further in the former direction, one study
+{[}32{]} set a fixed cohort size while two {[}26,34,35{]} devised more
 sophisticated algorithms to balance multiple desiderata including
@@
 rather than by cardinality, only the one previously published study
-{[}11{]} and the most recent study {[}33,35{]} in this corpus took this
+{[}11{]} and the most recent study {[}34,35{]} in this corpus took this
 approach. Viewed as a hyperparameter of the individualized modeling
@@
 optimal number and set of predictors as well as number of training cases
-(comprising the retrieved cohort) that was later used by {[}25{]}. More
-recently, {[}33{]} and {[}35{]} proposed a three-step process for cohort
+(comprising the retrieved cohort) that was later used by {[}26{]}. More
+recently, {[}34{]} and {[}35{]} proposed a three-step process for cohort
 selection that filtered by exact match for one subset of predictors,
@@
 treatment at decision points during a monitored patent's stay
-{[}33,35{]}. These studies suggest much wider potential for localized
+{[}34,35{]}. These studies suggest much wider potential for localized
 real-world evidence generation, which has long been promoted as a
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
 \CSLLeftMargin{{[}5{]} }%
@@
 cancer--patient targeted prognosis: An exploration of case base
-reasoning. Pathology, Research and Practice 1997;193:535--42.
+reasoning. Pathology - Research and Practice 1997;193:535--42.
 \url{https://doi.org/10.1016/S0344-0338(97)80011-8}.}
@@
 advice in nursing using text representation and structured text
-retrieval. Artif Intell Med 1997;9:79--99.
+retrieval. Artificial Intelligence in Medicine 1997;9:79--99.
 \url{https://doi.org/10.1016/s0933-3657(96)00362-4}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004a}{}}%
 \CSLLeftMargin{{[}20{]} }%
@@
 a combined {Kohonen} mapping and instance-based evaluation criterion.
-Artif Intell Med 2004;31:45--55.
+Artificial Intelligence in Medicine 2004;31:45--55.
 \url{https://doi.org/10.1016/j.artmed.2004.01.002}.}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Elter2007}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Elter2007a}{}}%
 \CSLLeftMargin{{[}22{]} }%
@@
 prediction of breast cancer biopsy outcomes using two {CAD} approaches
-that both emphasize an intelligible decision process. Med Phys
+that both emphasize an intelligible decision process. Medical Physics
 2007;34:4164--72. \url{https://doi.org/10.1118/1.2786864}.}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011a}{}}%
 \CSLLeftMargin{{[}24{]} }%
@@
 {eXiT}*{CBR}: {A} framework for case-based medical diagnosis development
-and experimentation. Artif Intell Med 2011;51:81--91.
+and experimentation. Artificial Intelligence in Medicine 2011;51:81--91.
 \url{https://doi.org/10.1016/j.artmed.2010.09.002}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Liang2015}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
 \CSLLeftMargin{{[}25{]} }%
-\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
-system for integrated feature, neighborhood and parameter optimization
-utilizing gravitational search algorithm. EVOLVING SYSTEMS 2015;6:1--14.
-\url{https://doi.org/10.1007/s12530-013-9081-x}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
-\CSLLeftMargin{{[}26{]} }%
 \CSLRightInline{Verma A, Fiasche M, Cuzzola M, Irrera G. Comparative
@@
 markers and for risk estimation in {Type2 Diabetes} in {Italian}
-population. EVOLVING SYSTEMS 2015;6:15--22.
+population. Evolving Systems 2015;6:15--22.
 \url{https://doi.org/10.1007/s12530-013-9083-8}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Liang2015a}{}}%
+\CSLLeftMargin{{[}26{]} }%
+\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
+system for integrated feature, neighborhood and parameter optimization
+utilizing gravitational search algorithm. Evolving Systems 2015;6:1--14.
+\url{https://doi.org/10.1007/s12530-013-9081-x}.}
 
@@
 {K-nearest} neighbors with logistic regression in case-based reasoning.
-Stud Health Technol Inform 2012;180:275--9.}
+Studies in Health Technology and Informatics 2012;180:275--9.}
 
@@
 Malvehy J. {DERMA}: A melanoma diagnosis platform based on collaborative
-multilabel analog reasoning. ScientificWorldJournal 2014;2014:351518.
-\url{https://doi.org/10.1155/2014/351518}.}
+multilabel analog reasoning. The Scientific World Journal
+2014;2014:351518. \url{https://doi.org/10.1155/2014/351518}.}
 
@@
 \CSLRightInline{Vilhena J, Vicente H, Martins MR, Grañeda JM, Caldeira
-F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk. J
-Biomed Inform 2016;62:265--75.
+F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk.
+Journal of Biomedical Informatics 2016;62:265--75.
 \url{https://doi.org/10.1016/j.jbi.2016.07.013}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
 \CSLLeftMargin{{[}30{]} }%
-\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
-Decision-Making Based} on {Big Clinical Data}. J Healthc Eng
-2018;2018:3917659. \url{https://doi.org/10.1155/2018/3917659}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
-\CSLLeftMargin{{[}31{]} }%
 \CSLRightInline{Zhang H, Zhu F, Dodge H, Higgins G, Omenn G, Guan Y, et
 al. A similarity-based approach to leverage multi-cohort medical data on
-the diagnosis and prognosis of {Alzheimer}'s disease. GIGASCIENCE
+the diagnosis and prognosis of {Alzheimer}'s disease. GigaScience
 2018;7. \url{https://doi.org/10.1093/gigascience/giy085}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
+\CSLLeftMargin{{[}31{]} }%
+\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
+Decision-Making Based} on {Big Clinical Data}. Journal of Healthcare
+Engineering 2018;2018:3917659.
+\url{https://doi.org/10.1155/2018/3917659}.}
 
@@
 for {ICU} patients using individualized single classification algorithm.
-COMPUTER METHODS AND PROGRAMS IN BIOMEDICINE 2020;186.
+Computer Methods and Programs in Biomedicine 2020;186.
 \url{https://doi.org/10.1016/j.cmpb.2019.105224}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Tang2021d}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
 \CSLLeftMargin{{[}33{]} }%
-\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
-Ng K. Precision population analytics: Population management at the
-point-of-care. JOURNAL OF THE AMERICAN MEDICAL INFORMATICS ASSOCIATION
-2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
-\CSLLeftMargin{{[}34{]} }%
 \CSLRightInline{Wang Y, Liang Y, Sun H, Yang Y. Emergency {Response} for
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Tang2021c}{}}%
+\CSLLeftMargin{{[}34{]} }%
+\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
+Ng K. Precision population analytics: Population management at the
+point-of-care. Journal of the American Medical Informatics Association
+2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Ng2021b}{}}%
@@
 Personalized treatment options for chronic diseases using precision
-cohort analytics. SCIENTIFIC REPORTS 2021;11.
+cohort analytics. Scientific Reports 2021;11.
 \url{https://doi.org/10.1038/s41598-021-80967-5}.}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006a}{}}%
 \CSLLeftMargin{{[}36{]} }%
@@
 usefulness of explanation in a case-based reasoning system for decision
-support in bronchiolitis treatment. Comput Intelligence 2006;22:269--81.
+support in bronchiolitis treatment. Computational Intelligence
+2006;22:269--81.
 \url{https://doi.org/10.1111/j.1467-8640.2006.00288.x}.}
@@
 Neighbour Algorithm} with {Logistic Regression} in the {Prediction} of
-{Patients}' {Registration} on the {Renal Transplant Waiting List}. PLoS
+{Patients}' {Registration} on the {Renal Transplant Waiting List}. PLOS
 ONE 2013;8:e71991. \url{https://doi.org/10.1371/journal.pone.0071991}.}

### 2023-08-18 scr/synthesis.r

@@ -55,7 +55,7 @@ googlesheets4::read_sheet(
   )) ->
   tab_synthesis
 # write to file, to be read into document
-write_rds(tab_synthesis, file = here::here("tab/synthesis.rds"))
+write_rds(tab_synthesis, file = here::here("data/synthesis.rds"))
 
 # mosaic plot of task, aim, and source
 tab_synthesis %>%
@@ -140,7 +140,7 @@ googlesheets4::read_sheet(
   arrange(Date) |> 
   print() -> composite_studies
 # write to file, to be read into document
-write_rds(composite_studies, file = here::here("tab/composite.rds"))
+write_rds(composite_studies, file = here::here("data/composite.rds"))
 
 # years of publication
 composite_studies %>%
@@ -169,7 +169,7 @@ composite_studies %>%
   labs(x = NULL, y = NULL) ->
   year_plot
 print(year_plot)
-ggsave(here::here("fig/fig-years.png"), year_plot, width = 5, height = 2)
+ggsave(here::here("fig/fig-years.png"), year_plot, width = 6, height = 2)
 
 # approach types
 composite_studies %>%
@@ -236,4 +236,9 @@ composite_terms %>%
   arrange(desc(n), terms) %>%
   print(n = Inf)
 
-# 
+# save ID column for studies described by multiple citations
+composite_studies |> 
+  transmute(Citation = str_split(Citation, "; "), ID = row_number()) |> 
+  unnest(Citation) |> 
+  print(n = Inf) -> citation_id
+write_rds(citation_id, here::here("data/citations.rds"))

### 2023-08-19 docs/Manuscript/Manuscript.Rmd
draft rationales section

@@
 Similarity-matching powers the retrieval step of most case-based reasoning (CBR) systems and provides the neighborhoods from which nearest neighbor (NN) models derive predictions.
-Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), most CBR tools return the retrieved cases to the user to inform human decision-making.
+Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), traditional CBR tools return the retrieved cases to the user to inform human decision-making.
 
@@
 In principle, this allows interpretable models like GLMs to identify and measure the effects of determining factors that are specific to the index case.
+The use of similarity cohorts distinguishes localized modeling from classical GLMs that still dominate analysis in clinical trials and comparative effectiveness research, as well as from many cutting-edge ML models that are fitted to whole populations, while the cohorts' heterogeneity and its utility to the model distinguish the approach from CBR.
 In the reviewed studies, localized modeling shows some promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.
+
+<!--
+In multivariate settings, because similarity decreases with difference along each factor, included cases that differ more with respect to some factors will tend to differ less with respect to others.
+Consequently, (1) the distribution of a factor in the subpopulation will tend to be centered near its value at the index case, and (2) the joint distribution of a pair of factors will tend to be sparse away from both center axes, i.e.
+when both factors differ from their values at the index case.
+This should produce localized models that are less responsive to deviation along multiple uncorrelated factors, so that the influence of each factor in the model reflects more its independent effect than its interactions with other factors.
+-->
 
@@
 We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
-While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis / dimension reduction / cluster analysis, risk evaluation / outcome prediction, clinical decision support / software tools).
+While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools).
 For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive modeling.
@@
 In cases of disagreement between the authors on the reason for exclusion, the first criterion was adopted.
-There were 60 disagreements over inclusion versus exclusion.
-The authors arrived at agreement on each manuscript through discussion.
+The authors arrived at agreement on inclusion or exclusion through discussion.
 We calculated inter-rater reliability to evaluate our screening and review process.
@@
 Finally, one author (JCB) applied the same review process to the seed set of 43 studies that inspired the review.
-This obtained 6 additional studies that had not been retrieved by any of our database searches [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019].
 
@@
 
+There were 60 disagreements over inclusion versus exclusion.
+Inter-rater reliability was low, at 82% relative to a 72% probability of random agreement.
+
 We then reviewed 43 studies comprising a seed set that inspired this review.
-After removing duplicates and screening for eligibility, we were left with 6 additional studies, 1 of which was excluded from the synthesis for using nearest neighbors prediction.
+After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using nearest neighbors prediction.
 Finally, citations in reference lists from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
 This left us with 19 + 1 + 5 = 25 studies included in the review and synthesis.
-Inter-rater reliability was low, at 82% relative to a 72% probability of random agreement.
 
@@
 
-Another dominant characteristic of included studies is their remit.
+Another dominant characteristic of included studies is their broader remit.
 We categorized studies as "knowledge" or "practice" based on whether their aim was to produce generalizable knowledge or to improve practice.
@@
 
-In addition to their remit, we coded several aspects of the design of each study, including the source and type of data and the clinical task the model performed (Table \ref{tab:synthesis}), as well as the specific methodology and terminology adopted (Table \ref{tab:composite}).
+In addition to the aim of its analysis, we coded several aspects of the design of each study, including the source and type of data and the clinical task the model performed (Table \ref{tab:synthesis}), as well as the specific methodology and terminology adopted (Table \ref{tab:composite}).
 In the following subsections, we take a closer look at these design elements.
@@
 
-## Application domain
+## Application domains
 
@@
 For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006a].
-This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used a non-traditional voting scheme to generate recommendations.
+This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used an unconventional voting scheme to generate recommendations.
 A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021a].
@@
 
-The mosaic plot in Figure \ref{fig:properties} summarizes the relation between data source, clinical task, and remit.
+The mosaic plot in Figure \ref{fig:properties} summarizes the relation between data source, clinical task, and aim.
 
@@
 
+## Rationales
+
+The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{sec:availability}).
+The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015a; @Ng2015; @Lee2017].
+In particular, @Lee2015 hypothesized and confirmed that the value of each past case would be positively related to its similarity to the index case, an assumption built in to the weighting schemes of other approaches.
+@Lowsky2013 made a different case, that the fewer parametric assumptions and complications of a CBR-style model would allow for greater accuracy.
+Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to model performance [@Song2006c; @Xu2008a].
+
+In an interesting contrast, @Tang2021c and @Ng2021b argued that their localized approach using the larger and more heterogeneous populations covered by EHR-derived data could better capture the messy and diverse lessons of everyday practice, as a counterpart to guidelines based on randomized controlled trials.
+Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2012a; @Nicolas2014a; @Verma2015a; @Lee2017].
+@Verma2015a and @Zhang2018a additionally pointed out that localized models are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
+These properties, they said, tend to be more difficult for whole-population models to handle.
+
+The other frequent advantage attributed to localized modeling was interpretability.
+@Elter2007a and @Nicolas2014a emphasized the value of the detailed past cases, available to the user, on which predictions are based.
+This "self-explanation capability" made outputs more intelligible to physicians in the CDS setting.
+@Ng2015 additional pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
+
+The remaining coded rationales were for augmentations or hybridizations of then-conventional CBR, most of which defended the use of other tools to improve performance via cohort selection [@CampilloGimenez2012a; @Nicolas2014a; @Vilhena2016a] or to make models and outputs more interpretable [@Lopez2011a; @Wang2019].
+@Liang2015a argued for simultaneous optimization of cohort and feature selection with model parameterization; their TWNFI approach combines localization with regularization.
+
 ## Method development
+
+**Move part of this paragraph to Discussion.**
 
@@
 
-### Coherence
+## Coherence
 
@@
 Distinct from but related to similarity learning was supervised cohort construction.
-This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called "statistical CBR".
-The step used a heuristic procedure to locate a similarity threshold that (locally) maximizes predictive accuracy on the training set.
+This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called "statistical CBR".[^combinatorial] The step used a heuristic procedure to locate a similarity threshold that (locally) maximizes predictive accuracy on the training set.
 This is analogous to optimizing the neighborhood size parameter in NN predictive modeling, so we will not discuss it further except to mention that the authors found statistical CBR to outperform conventional CBR on several data sets.
 @CampilloGimenez2012a combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the NN model using the previously optimized weights.
+
+[^combinatorial]: Drawing from @Goyal2008, we suggest the term "combinatorial CBR" for the then-conventional approach using similarity cohorts of fixed cardinality.
 
@@
 Each step can be unsupervised supervised: Unsupervised similarity measures include discrete measures like the Levenshtein distance and continuous measures like cosine similarity, while supervised measures include the Mahalanobis distance and random forest proximity as well as several composite measures with weights calculated from the data.
-Most retrieval steps obtained cohorts with a uniform size (cardinality) or similarity threshold, and the remaining were likewise only informed by predictors; thus, while in principle a retrieval step could be supervised, in this sample none were.
+Most retrieval steps obtained cohorts with a uniform size (cardinality) or similarity threshold, and the remaining were likewise only informed by predictors; thus, while a retrieval step could in principle be supervised, in this sample none were.
 Almost all models were supervised, being that most performed predictive tasks, though @Mariuzzi1997 modeled only survival rates within localized cohorts.
@@
 # Acknowledgments
+
+We are grateful for the astute and supportive comments of three anonymous reviewers.
 
@@
 * Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
-* Code used to perform analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
+* Code used to conduct analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
 
@@
 
-Following discussion among AC, PMJ, and JCB, we omitted results from Google Scholar due to missing abstracts, missing URLs, high overlap with other search results, and irreproducibility of the search process.
+Following discussion among AC, PMJ, and JCB, we discarded results from Google Scholar due to missing abstracts, missing URLs, high overlap with other search results, and irreproducibility of the search process.
 
@@
 
+**Write this section!**
+
 1. Deduplication (Covidence)

### 2023-08-19 docs/Manuscript/Manuscript.tex
draft rationales section

@@
 are generated automatically via averaging (regression) or voting
-(classification), most CBR tools return the retrieved cases to the user
-to inform human decision-making.
+(classification), traditional CBR tools return the retrieved cases to
+the user to inform human decision-making.
 
@@
 identify and measure the effects of determining factors that are
-specific to the index case. In the reviewed studies, localized modeling
-shows some promise toward simultaneously achieving the performance
+specific to the index case. The use of similarity cohorts distinguishes
+localized modeling from classical GLMs that still dominate analysis in
+clinical trials and comparative effectiveness research, as well as from
+many cutting-edge ML models that are fitted to whole populations, while
+the cohorts' heterogeneity and its utility to the model distinguish the
+approach from CBR. In the reviewed studies, localized modeling shows
+some promise toward simultaneously achieving the performance
 improvements of ML and the interpretability of classical models.
@@
 patient preferences, patient-reported outcomes, tracking devices, social
-media) and survey a much broader scope of models (exploratory analysis /
-dimension reduction / cluster analysis, risk evaluation / outcome
-prediction, clinical decision support / software tools). For the present
-review, we are interested in how similarity matching on patient-level
-health data is used to construct local cohorts for predictive modeling.
+media) and survey a much broader scope of models (exploratory analysis
+via dimension reduction and cluster analysis; risk evaluation and
+outcome prediction; clinical decision support and software tools). For
+the present review, we are interested in how similarity matching on
+patient-level health data is used to construct local cohorts for
+predictive modeling.
 
@@
 frameworks. In cases of disagreement between the authors on the reason
-for exclusion, the first criterion was adopted. There were 60
-disagreements over inclusion versus exclusion. The authors arrived at
-agreement on each manuscript through discussion. We calculated
+for exclusion, the first criterion was adopted. The authors arrived at
+agreement on inclusion or exclusion through discussion. We calculated
 inter-rater reliability to evaluate our screening and review process.
@@
 Finally, one author (JCB) applied the same review process to the seed
-set of 43 studies that inspired the review. This obtained 6 additional
-studies that had not been retrieved by any of our database searches
-{[}12--17{]}.
+set of 43 studies that inspired the review.
 
@@
 
+There were 60 disagreements over inclusion versus exclusion. Inter-rater
+reliability was low, at 82\% relative to a 72\% probability of random
+agreement.
+
 We then reviewed 43 studies comprising a seed set that inspired this
 review. After removing duplicates and screening for eligibility, we were
-left with 6 additional studies, 1 of which was excluded from the
-synthesis for using nearest neighbors prediction. Finally, citations in
-reference lists from the 43 + 328 = 371 studies assessed for eligibility
-led us to identify 1 additional study that met our criteria {[}18{]}.
-This left us with 19 + 1 + 5 = 25 studies included in the review and
-synthesis. Inter-rater reliability was low, at 82\% relative to a 72\%
-probability of random agreement.
+left with 6 additional studies {[}12--17{]}, 1 of which was excluded
+from the synthesis for using nearest neighbors prediction. Finally,
+citations in reference lists from the 43 + 328 = 371 studies assessed
+for eligibility led us to identify 1 additional study that met our
+criteria {[}18{]}. This left us with 19 + 1 + 5 = 25 studies included in
+the review and synthesis.
 
@@
 
-Another dominant characteristic of included studies is their remit. We
-categorized studies as ``knowledge'' or ``practice'' based on whether
-their aim was to produce generalizable knowledge or to improve practice.
-We classified 7 of the 25 studies as ``knowledge'', making the majority
-of studies from a ``practice'' standpoint. These studies aimed primarily
-to provide providers with tools to use in clinic or to improve outcomes
-for patients currently receiving care.
+Another dominant characteristic of included studies is their broader
+remit. We categorized studies as ``knowledge'' or ``practice'' based on
+whether their aim was to produce generalizable knowledge or to improve
+practice. We classified 7 of the 25 studies as ``knowledge'', making the
+majority of studies from a ``practice'' standpoint. These studies aimed
+primarily to provide providers with tools to use in clinic or to improve
+outcomes for patients currently receiving care.
 
@@
 
-In addition to their remit, we coded several aspects of the design of
-each study, including the source and type of data and the clinical task
-the model performed (Table \ref{tab:synthesis}), as well as the specific
-methodology and terminology adopted (Table \ref{tab:composite}). In the
-following subsections, we take a closer look at these design elements.
+In addition to the aim of its analysis, we coded several aspects of the
+design of each study, including the source and type of data and the
+clinical task the model performed (Table \ref{tab:synthesis}), as well
+as the specific methodology and terminology adopted (Table
+\ref{tab:composite}). In the following subsections, we take a closer
+look at these design elements.
 
@@
 
-\hypertarget{application-domain}{%
-\subsection{Application domain}\label{application-domain}}
+\hypertarget{application-domains}{%
+\subsection{Application domains}\label{application-domains}}
 
@@
 literature on CBR systems and was included here despite relying on NN
-prediction because it used a non-traditional voting scheme to generate
+prediction because it used an unconventional voting scheme to generate
 recommendations. A more recent study took essentially the same focus
@@
 The mosaic plot in Figure \ref{fig:properties} summarizes the relation
-between data source, clinical task, and remit.
+between data source, clinical task, and aim.
 
@@
 
+\hypertarget{rationales}{%
+\subsection{Rationales}\label{rationales}}
+
+The included studies hypothesized, asserted, or assumed several benefits
+of localized modeling specific to clinical and medical settings and
+advantages over other modeling approaches (see Section
+\ref{sec:availability}). The most common was that the restriction to
+similar or relevant past cases would improve predictive performance for
+the index case {[}11,15,16,26{]}. In particular, {[}14{]} hypothesized
+and confirmed that the value of each past case would be positively
+related to its similarity to the index case, an assumption built in to
+the weighting schemes of other approaches. {[}12{]} made a different
+case, that the fewer parametric assumptions and complications of a
+CBR-style model would allow for greater accuracy. Additionally, several
+investigators asserted that the use of localized cohorts befit the
+clinical focus on the individual patient rather than the population,
+without reference to model performance {[}21,23{]}.
+
+In an interesting contrast, {[}34{]} and {[}35{]} argued that their
+localized approach using the larger and more heterogeneous populations
+covered by EHR-derived data could better capture the messy and diverse
+lessons of everyday practice, as a counterpart to guidelines based on
+randomized controlled trials. Their emphasis on the enabling role of
+EHRs to power methods well-suited to large, structured data repositories
+was shared by several others {[}16,25,27,28{]}. {[}25{]} and {[}30{]}
+additionally pointed out that localized models are adaptable to noise in
+the data, variation in patterns of missingness, and (harkening to the
+last step of the R4 cycle) addition of new cases with known outcomes to
+the corpus. These properties, they said, tend to be more difficult for
+whole-population models to handle.
+
+The other frequent advantage attributed to localized modeling was
+interpretability. {[}22{]} and {[}28{]} emphasized the value of the
+detailed past cases, available to the user, on which predictions are
+based. This ``self-explanation capability'' made outputs more
+intelligible to physicians in the CDS setting. {[}15{]} additional
+pointed out that localized GLMs yield localized effect estimates, which
+may help investigators identify individually relevant risk factors.
+
+The remaining coded rationales were for augmentations or hybridizations
+of then-conventional CBR, most of which defended the use of other tools
+to improve performance via cohort selection {[}27--29{]} or to make
+models and outputs more interpretable {[}17,24{]}. {[}26{]} argued for
+simultaneous optimization of cohort and feature selection with model
+parameterization; their TWNFI approach combines localization with
+regularization.
+
 \hypertarget{method-development}{%
 \subsection{Method development}\label{method-development}}
+
+\textbf{Move part of this paragraph to Discussion.}
 
@@
 \hypertarget{coherence}{%
-\subsubsection{Coherence}\label{coherence}}
+\subsection{Coherence}\label{coherence}}
 
@@
 fixed similarity rather than count threshold, a counterpart to
-conventional CBR they called ``statistical CBR''. The step used a
-heuristic procedure to locate a similarity threshold that (locally)
-maximizes predictive accuracy on the training set. This is analogous to
-optimizing the neighborhood size parameter in NN predictive modeling, so
-we will not discuss it further except to mention that the authors found
-statistical CBR to outperform conventional CBR on several data sets.
-{[}27{]} combined these learning techniques: They used logistic
-regression on the training set to obtain weights for the predictors (by
-predicting the binary outcome of kidney transplant waitlist
-registration) and for the cases (by predicting agreement of outcome
-between an index case and other training cases), and they used
-exhaustion to optimize the size of the retrieved cohort for the accuracy
-of the NN model using the previously optimized weights.
+conventional CBR they called ``statistical CBR''.\footnote{Drawing from
+  {[}42{]}, we suggest the term ``combinatorial CBR'' for the
+  then-conventional approach using similarity cohorts of fixed
+  cardinality.} The step used a heuristic procedure to locate a
+similarity threshold that (locally) maximizes predictive accuracy on the
+training set. This is analogous to optimizing the neighborhood size
+parameter in NN predictive modeling, so we will not discuss it further
+except to mention that the authors found statistical CBR to outperform
+conventional CBR on several data sets. {[}27{]} combined these learning
+techniques: They used logistic regression on the training set to obtain
+weights for the predictors (by predicting the binary outcome of kidney
+transplant waitlist registration) and for the cases (by predicting
+agreement of outcome between an index case and other training cases),
+and they used exhaustion to optimize the size of the retrieved cohort
+for the accuracy of the NN model using the previously optimized weights.
 
@@
 similarity threshold, and the remaining were likewise only informed by
-predictors; thus, while in principle a retrieval step could be
+predictors; thus, while a retrieval step could in principle be
 supervised, in this sample none were. Almost all models were supervised,
@@
 
+We are grateful for the astute and supportive comments of three
+anonymous reviewers.
+
 \hypertarget{support}{%
@@
 \item
-  Code used to perform analyses and prepare the manuscript:
+  Code used to conduct analyses and prepare the manuscript:
   \url{https://github.com/corybrunson/imsr}
@@
 
-Following discussion among AC, PMJ, and JCB, we omitted results from
+Following discussion among AC, PMJ, and JCB, we discarded results from
 Google Scholar due to missing abstracts, missing URLs, high overlap with
@@
 
+\textbf{Write this section!}
+
 \begin{enumerate}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
+\CSLLeftMargin{{[}42{]} }%
+\CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A
+combinatorial approach to nearest neighbor search. Proceedings of the
+2008 {International Conference} on {Web Search} and {Data Mining}, {New
+York, NY, USA}: {Association for Computing Machinery}; 2008, p. 25--32.
+\url{https://doi.org/10.1145/1341531.1341538}.}
+
 \end{CSLReferences}

### 2023-08-29 docs/Manuscript/Manuscript.Rmd
draft results subsection on identified needs for future work

@@
 
+## Identified needs
+
+The study authors focused their recommendations and their own plans for future work mostly on technical improvements and evaluations [@Mariuzzi1997; @Yearwood1997a].
+Urged improvements included full or partial automation of user-dependent steps such as predictor selection [@Mariuzzi1997; @Yearwood1997a], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006c; @Lee2017]; extensions to new data structures [@Lopez2011a], data types [@Liang2015a; @Verma2015a; @Malykh2018a], and reasoning systems [@Nicolas2014a]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015a; @Zhang2018a].
+Authors also urged validations and independent evaluations using larger or more comprehensive data sets [@Elter2007a; @Xu2008a; @Verma2015a; @Ng2015], using data aggregated from multiple health systems [@Lee2015; @Lee2017; @Tang2021c; @Ng2021b], and in other care settings or disease contexts [@Song2006c; @Zhang2018a; @Tang2021c; @Ng2021b].
+Less common were calls to strengthen the connection between the methods and the users.
+Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997a; @Wang2019].
+Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004a], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
+Most of this incremental work was indeed carried out by later researchers whose output appears in our sample.
+
 # Discussion

### 2023-08-29 docs/Manuscript/Manuscript.tex
draft results subsection on identified needs for future work

@@
 
+\hypertarget{identified-needs}{%
+\subsection{Identified needs}\label{identified-needs}}
+
+The study authors focused their recommendations and their own plans for
+future work mostly on technical improvements and evaluations
+{[}11,19{]}. Urged improvements included full or partial automation of
+user-dependent steps such as predictor selection {[}11,19{]}, similarity
+learning {[}11,17{]}, and parameter optimization {[}16,21{]}; extensions
+to new data structures {[}24{]}, data types {[}25,26,31{]}, and
+reasoning systems {[}28{]}; and the use of more advanced model
+components to improve accuracy or efficiency {[}12,14,26,30{]}. Authors
+also urged validations and independent evaluations using larger or more
+comprehensive data sets {[}15,22,23,25{]}, using data aggregated from
+multiple health systems {[}14,16,34,35{]}, and in other care settings or
+disease contexts {[}21,30,34,35{]}. Less common were calls to strengthen
+the connection between the methods and the users. Some authors urged the
+incorporation of (human-derived) domain knowledge into the data or
+models {[}17,19{]}. Others suggested new uses of their modeling
+approaches: to measure feature importance {[}20{]}, to make models more
+expressive {[}14{]}, and to combine information obtained both from
+global and from localized models {[}15{]}. Most of this incremental work
+was indeed carried out by later researchers whose output appears in our
+sample.
+
 \hypertarget{discussion}{%

### 2023-08-29 docs/Manuscript/Manuscript.Rmd
add footnote about federated models

--- 2023-08-19 docs/Manuscript/Manuscript.Rmd
@@
 
-These programs used significantly varying terminology for common concepts, and in particular no common term is in use for what we here refer to as localized modeling; authors described this kind of method as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015a, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021c; @Ng2021b].
+These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015a, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021c; @Ng2021b].
 We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
-Because local naturally contrasts with global, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.
+Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^federated]
+
+[^federated]: As one reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
+The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
+At present, however, we feel that these research areas are sufficiently disjoint to avoid confusion in practice.
 

### 2023-08-29 docs/Manuscript/Manuscript.tex
add references for federated models footnote

@@
 
-These programs used significantly varying terminology for common
-concepts, and in particular no common term is in use for what we here
-refer to as localized modeling; authors described this kind of method as
-``targeted prognosis'' {[}11{]}, ``transductive inference'' {[}21{]},
-``personalized decision support'' {[}14{]}, ``personalized (predictive)
-models'' {[}26, in contrast to ``local models,''15,17,32{]}, and
-``precision cohort'' analysis {[}17,34,35{]}. We find uses of
-``targeted'', ``personalized'', and ``precision'' generic and imprecise
-to this approach, while transductive inference is an established term
-for a broader set of methods in ML. Because local naturally contrasts
-with global, we propose ``localized models'' as a suitable term for this
-counterpart to globally-fitted models.
+These programs used varying terminology for common concepts, and no
+common term is in use for what we here refer to as localized modeling;
+authors described their approaches as ``targeted prognosis'' {[}11{]},
+``transductive inference'' {[}21{]}, ``personalized decision support''
+{[}14{]}, ``personalized (predictive) models'' {[}26, in contrast to
+``local models,''15,17,32{]}, and ``precision cohort'' analysis
+{[}17,34,35{]}. We find uses of ``targeted'', ``personalized'', and
+``precision'' generic and imprecise to this approach, while transductive
+inference is an established term for a broader set of methods in ML.
+Because \emph{local} naturally contrasts with \emph{global}, we propose
+``localized models'' as a suitable term for this counterpart to
+globally-fitted models.\footnote{As one reviewer pointed out, the
+  terminology of local and global is also used in the unrelated context
+  of federated ML to distinguish models trained using data stored in a
+  single storage device (local) from their aggregations (global)
+  {[}39,40{]}. The prospects for federated architecture to more
+  effectively implement similarity cohort--based models are intriguing,
+  and would require some reconciliation of terms. At present, however,
+  we feel that these research areas are sufficiently disjoint to avoid
+  confusion in practice.}
 
@@
 used in 8 studies that tie in to the much larger literature on metric
-learning {[}39{]}. These techniques are designed to detect structure in
+learning {[}41{]}. These techniques are designed to detect structure in
 data, especially associations between predictors and responses, and to
@@
 (Table \ref{tab:composite}). Based on the key properties of metric
-learning algorithms identified by {[}40{]}, most learned measures were
+learning algorithms identified by {[}42{]}, most learned measures were
 non-linear while some were local, and some were optimized globally while
@@
 here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning technique {[}41{]} to
+appeal to a supervised distance metric learning technique {[}43{]} to
 obtain a similarity measure on their set of dermoscopy and confocal
@@
 conventional CBR they called ``statistical CBR''.\footnote{Drawing from
-  {[}42{]}, we suggest the term ``combinatorial CBR'' for the
+  {[}44{]}, we suggest the term ``combinatorial CBR'' for the
   then-conventional approach using similarity cohorts of fixed
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
+\CSLLeftMargin{{[}39{]} }%
+\CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
+Reviewing {Federated Machine Learning} and {Its Use} in {Diseases
+Prediction}. Sensors 2023;23:2112.
+\url{https://doi.org/10.3390/s23042112}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
+\CSLLeftMargin{{[}40{]} }%
+\CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
+Völker U, Baumbach J, et al. Federated {Machine Learning},
+{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
+Research}: {Scoping Review}. Journal of Medical Internet Research
+2023;25:e41588. \url{https://doi.org/10.2196/41588}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-\CSLLeftMargin{{[}39{]} }%
+\CSLLeftMargin{{[}41{]} }%
 \CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-\CSLLeftMargin{{[}40{]} }%
+\CSLLeftMargin{{[}42{]} }%
 \CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-\CSLLeftMargin{{[}41{]} }%
+\CSLLeftMargin{{[}43{]} }%
 \CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-\CSLLeftMargin{{[}42{]} }%
+\CSLLeftMargin{{[}44{]} }%
 \CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A

### 2023-09-07 docs/Manuscript/Manuscript.Rmd
make further revisions following discussion with adi

@@
     citation_package: "default"
+header-includes:
+  \usepackage{lscape}
+  \newcommand{\landscapebegin}{\begin{landscape}}
+  \newcommand{\landscapeend}{\end{landscape}}
+
 ---
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(dplyr)
-library(stringr)
+library(tidyverse)
 Sys.sleep(10)
@@
 composite_studies <- readr::read_rds(here::here("data/composite.rds"))
+tab_eval_comp <- readr::read_rds(here::here("data/eval-comp.rds"))
 citation_id <- readr::read_rds(here::here("data/citations.rds"))
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{sec:availability}).
+Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{sec:data}).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
@@
 
-The 25 included studies were analyzed based on 20 characteristics (see Section \ref{sec:availability}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The 25 included studies were analyzed based on 20 characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
@@
 ## Rationales
+\label{sec:rationales}
 
-The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{sec:availability}).
+The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{sec:data}).
 The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015a; @Ng2015; @Lee2017].
@@
 
-## Method development
-
-**Move part of this paragraph to Discussion.**
-
-Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), laboratory (2), and image (1) data.
-We suggest three reasons for this: First, this literature traces back to the 1990s, before -omic data could be generated cost-effectively at scale.
-Second, CBR in particular has a strong tradition in clinical decision support, where the focus of our sample remains throughout the review period.
-Third, because most -omic data are highly homogeneous—all measurements are made along or can be transformed to a common scale, e.g.
-greyscale pixellations for X-ray images and transcripts per million for RNA-seq data—more deeply theoretical analysis techniques have been developed and come into wide use.
-While variations on correlation-based approaches like EHR-based phenome-wide association studies and similarity-based methods like CBR itself have been developed, more mechanistic and probabilistic tools have not become a domain standard.
+## Challenges
 
@@
 
+## Performance assessments
+\label{sec:performance}
+
+Most included studies quantitatively compared the predictive performance of their proposed method(s) to one or more comparators.
+What we took to be the signature results are collated in Table \ref{tab:performance} in the Appendix.
+Note that we exercised some judgment in classifying methods as proposals and comparators, as in some cases all methods were original but only some showcased main ideas.
+It would be impractical to meta-analyze these numbers due to the great variety of settings, problems, data types, and choices involved.
+
+However, we do observe one clear pattern: Of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], GSA [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]), all except the last are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
+Though @CampilloGimenez2012a and [@Ng2015] report exceptions, the pattern suggests the importance of the similarity measure to the retrieval step.
+Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model (statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]), they did not consistently outperform comparators.
+
 ## Identified needs
+\label{sec:needs}
 
@@
 Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004a], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
-Most of this incremental work was indeed carried out by later researchers whose output appears in our sample.
+Most of this incremental work was indeed carried out in later investigations in our sample.
+
+## Public availability
+\label{sec:availability}
+
+**Discuss public availability of code or implementations.**
 
 # Discussion
+
+We draw several observations from our encoding of the included studies.
+We first focus on two driving questions: What challenges did investigators set out to solve?
+How did their approaches meet these challenges?
+We then comment on the connectedness of the included studies and the potential for recombination.
 
@@
 
-## Trends
+## Study designs
 
-We draw several observations from our encoding of the included studies.
-We first focus on two driving questions: What challenges did investigators set out to solve?
-How did their approaches meet these challenges?
-We then comment on the connectedness of the included studies and the potential for recombination.
-
-**Add 1--2 paragraphs.**
+Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), laboratory (2), and image (1) data.
+We suggest three reasons for this: First, this literature traces back to the 1990s, before -omic data could be generated cost-effectively at scale.
+Second, CBR in particular has a strong tradition in clinical decision support, where the focus of our sample remains throughout the review period.
+Third, because most -omic data are highly homogeneous—all measurements are made along or can be transformed to a common scale, e.g.
+greyscale pixellations for X-ray images and transcripts per million for RNA-seq data—more deeply theoretical analysis techniques have been developed and come into wide use.
+While variations on correlation-based approaches like EHR-based phenome-wide association studies and similarity-based methods like CBR itself have been developed, more mechanistic and probabilistic tools have not become a domain standard.
 
@@
 Their method was used in several later studies returned by our search, which are not discussed here because they did not originate the technique.
-@Nicolas2014a explicitly appeal to a supervised distance metric learning technique [@Xing2002] to obtain a similarity measure on their set of dermoscopy and confocal images that most effectively separates malignant from benign melanoma tumors.
+@Nicolas2014a explicitly appeal to a supervised distance metric learning (DML) technique [@Xing2002] to obtain a similarity measure on their set of dermoscopy and confocal images that most effectively separates malignant from benign melanoma tumors.
 We will discuss the remaining uses of supervised similarity learning shortly.
@@
 
+## Directions and expectations
+
+As discussed in Section \ref{sec:rationales}, most studies were premised on the potential predictive value of localized models.
+As cautioned in Section \ref{sec:performance}, not all experiments affirmed this premise.
+Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
+Some studies showcased results using a variety of such specializations [@CampilloGimenez2012a; @Ng2015; @Zhang2018a].
+However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
+This would be a worthy goal of future work, though it might be limited by the lack of open-source code or public implementations (Section \ref{sec:availability}).
+
+While most studies identified outstanding technical needs (Section \ref{sec:needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
+Most studies also identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
+Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
+One valuable direction for future work would be to compare the ability of non-technical investigators or users to draw correct inferences from these and other interpretable methods and their levels of confidence in those inferences.
+
+We posit that an essential component of such "soft" performance goals is the ability of the user community to exert some control over the models.
+Given the impact on performance of the choice or optimization of the similarity measure, an important target for user input would be the importance of certain variables in the calculation of similarity, with an understanding of how it can impact not only the performance of the predictions but also the cohort retrieved for the model.
+For one example, a user may want to minimize the weight of rare diseases in medical history in order to retrieve a population with more such cases in order to better measure their associated risk to the outcome.
+In contrast, they may want to increase the weight of the indicating diagnosis in order to allow fewer patients from similar but distinct populations to influence the model.
+For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
+Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.
+
 ## Conclusions
@@
 
-# Data and code availability
-\label{sec:availability}
+# Data and code
+\label{sec:data}
 
@@
 
+## Performance evaluations and comparisons
+
+\landscapebegin
+
+```{r tab:performance, echo=FALSE, message=FALSE, warning=FALSE}
+# TODO: Place vertical separators between major columns.
+tab_eval_comp |> 
+  pivot_wider(
+    id_cols = c(zotero_key, date_of_publication, measure, task),
+    names_from = role, values_from = results
+  ) |> 
+  arrange(date_of_publication) |> 
+  mutate(len = map2_int(proposal, comparator, ~ max(length(.x), length(.y)))) |> 
+  mutate(across(c(proposal, comparator), ~ map2(., len, \(x, y) x[seq(y)]))) |> 
+  unnest(c(proposal, comparator)) |> 
+  separate(proposal, c("proposal", " "), sep = ": ") |> 
+  separate(comparator, c("comparator", "  "), sep = ": ") |> 
+  # use abbreviations throughout
+  mutate(across(
+    c(proposal, comparator),
+    ~ str_replace(., "[Ll]ogistic [Rr]egression", "LR")
+  )) |> 
+  mutate(across(
+    c(proposal, comparator),
+    ~ ifelse(. == "Logistic", "LR", .)
+  )) |> 
+  mutate(across(
+    c(proposal, comparator),
+    ~ str_replace(., "[Rr]andom [Ff]orest", "RF")
+  )) |> 
+  mutate(across(
+    c(proposal, comparator),
+    ~ str_replace(., "[Kk][\\- ][Nn]earest [Nn]eighbor(s){0,1}", "KNN")
+  )) |> 
+  mutate(across(
+    c(proposal, comparator),
+    ~ str_replace(., "([Aa]rtificial ){0,1}[Nn]eural [Nn]et(work){0,1}", "NN")
+  )) |> 
+  mutate(across(
+    c(proposal, comparator),
+    ~ str_replace(., "[Oo]ne-class", "One-class")
+  )) |> 
+  mutate(across(
+    c(proposal, ` `, comparator, `  `),
+    ~ ifelse(is.na(.), "", .)
+  )) |> 
+  transmute(
+    Citation = str_c("[@", zotero_key, "]"),
+    Measure = measure, Problem = task,
+    Proposals = proposal, ` `,
+    # insert blank column to separate right- and left-justified entries
+    # `   ` = "",
+    Comparators = comparator, `  `
+  ) |> 
+  mutate(across(
+    c(Citation, Measure, Problem),
+    ~ ifelse(is.na(lag(.)) | . != lag(.), ., "")
+  )) |>
+  # prevent numeric list columns from dominating the horizontal space
+  # mutate(across(c(Proposals, Comparators), ~ str_c("\\hspace{3em}", .))) |> 
+  mutate(across(c(` `, `  `), ~ str_c(., "\\hspace{6em}"))) |> 
+  knitr::kable(
+    align = "llllrlr",
+    caption = paste0(
+      "\\label{tab:performance}",
+      "Evaluations of proposed methods and comparisons to alternative methods.",
+      " (See original studies for full names and descriptions.)"
+    )
+  )
+```
+
+\landscapeend
+
 # References {-}

### 2023-09-07 docs/Manuscript/Manuscript.tex
make further revisions following discussion with adi

@@
 
+\usepackage{lscape} \newcommand{\landscapebegin}{\begin{landscape}} \newcommand{\landscapeend}{\end{landscape}}
+
 \begin{document}
@@
 Following the selection sources aligning with our review, we collected
-characteristics of included studies (Section \ref{sec:availability}).
-The features included bibliographic fields (date of publication,
-journal, authors, title, keywords, DOI), study goals (objective,
-generalizable knowledge, evaluation, clinical/medical domain), data sets
-(data source, type of data, range of data, number of cases/incidences,
-number of predictors/features), and methodological choices (types of
-similarity measure, families of adaptation step/statistical model,
-method(s) compared against, evaluation/performance/comparison measures
-(\& their values), name given to modeling approach). We used these data
-to detect and visualize trends amongst the included studies.
+characteristics of included studies (Section \ref{sec:data}). The
+features included bibliographic fields (date of publication, journal,
+authors, title, keywords, DOI), study goals (objective, generalizable
+knowledge, evaluation, clinical/medical domain), data sets (data source,
+type of data, range of data, number of cases/incidences, number of
+predictors/features), and methodological choices (types of similarity
+measure, families of adaptation step/statistical model, method(s)
+compared against, evaluation/performance/comparison measures (\& their
+values), name given to modeling approach). We used these data to detect
+and visualize trends amongst the included studies.
 
@@
 The 25 included studies were analyzed based on 20 characteristics (see
-Section \ref{sec:availability}), and we report and describe some
-observations here (Table \ref{tab:synthesis}). The studies were
-published in a variety of journals, with some of greater frequency,
-though no single journal published more than 3. The journal
-\emph{Evolving Systems} published 2 of the included studies,
-\emph{Artificial Intelligence in Medicine} published 3, and
-\emph{Hindawi - Journal of Healthcare Engineering} published 3.
+Section \ref{sec:data}), and we report and describe some observations
+here (Table \ref{tab:synthesis}). The studies were published in a
+variety of journals, with some of greater frequency, though no single
+journal published more than 3. The journal \emph{Evolving Systems}
+published 2 of the included studies, \emph{Artificial Intelligence in
+Medicine} published 3, and \emph{Hindawi - Journal of Healthcare
+Engineering} published 3.
 
@@
 
+\label{sec:rationales}
+
 The included studies hypothesized, asserted, or assumed several benefits
 of localized modeling specific to clinical and medical settings and
-advantages over other modeling approaches (see Section
-\ref{sec:availability}). The most common was that the restriction to
-similar or relevant past cases would improve predictive performance for
-the index case {[}11,15,16,26{]}. In particular, {[}14{]} hypothesized
-and confirmed that the value of each past case would be positively
-related to its similarity to the index case, an assumption built in to
-the weighting schemes of other approaches. {[}12{]} made a different
-case, that the fewer parametric assumptions and complications of a
-CBR-style model would allow for greater accuracy. Additionally, several
-investigators asserted that the use of localized cohorts befit the
-clinical focus on the individual patient rather than the population,
-without reference to model performance {[}21,23{]}.
+advantages over other modeling approaches (see Section \ref{sec:data}).
+The most common was that the restriction to similar or relevant past
+cases would improve predictive performance for the index case
+{[}11,15,16,26{]}. In particular, {[}14{]} hypothesized and confirmed
+that the value of each past case would be positively related to its
+similarity to the index case, an assumption built in to the weighting
+schemes of other approaches. {[}12{]} made a different case, that the
+fewer parametric assumptions and complications of a CBR-style model
+would allow for greater accuracy. Additionally, several investigators
+asserted that the use of localized cohorts befit the clinical focus on
+the individual patient rather than the population, without reference to
+model performance {[}21,23{]}.
 
@@
 
-\hypertarget{method-development}{%
-\subsection{Method development}\label{method-development}}
-
-\textbf{Move part of this paragraph to Discussion.}
-
-Consistent with their orientation, almost all included studies were
-conducted using clinical data, only occasionally together with
-patient-reported (2), laboratory (2), and image (1) data. We suggest
-three reasons for this: First, this literature traces back to the 1990s,
-before -omic data could be generated cost-effectively at scale. Second,
-CBR in particular has a strong tradition in clinical decision support,
-where the focus of our sample remains throughout the review period.
-Third, because most -omic data are highly homogeneous---all measurements
-are made along or can be transformed to a common scale, e.g.~greyscale
-pixellations for X-ray images and transcripts per million for RNA-seq
-data---more deeply theoretical analysis techniques have been developed
-and come into wide use. While variations on correlation-based approaches
-like EHR-based phenome-wide association studies and similarity-based
-methods like CBR itself have been developed, more mechanistic and
-probabilistic tools have not become a domain standard.
+\hypertarget{challenges}{%
+\subsection{Challenges}\label{challenges}}
 
@@
 
+\hypertarget{performance-assessments}{%
+\subsection{Performance assessments}\label{performance-assessments}}
+
+\label{sec:performance}
+
+Most included studies quantitatively compared the predictive performance
+of their proposed method(s) to one or more comparators. What we took to
+be the signature results are collated in Table \ref{tab:performance} in
+the Appendix. Note that we exercised some judgment in classifying
+methods as proposals and comparators, as in some cases all methods were
+original but only some showcased main ideas. It would be impractical to
+meta-analyze these numbers due to the great variety of settings,
+problems, data types, and choices involved.
+
+However, we do observe one clear pattern: Of the proposed methods that
+most evidently outperformed their comparators (Kohonen + CBR {[}20{]},
+TWNFI {[}18{]}, GSA {[}26{]}, CBR + rules (+ DML) {[}28{]}, Gaussian
+process regression {[}30{]}, JITL-ELM {[}32{]}), all except the last are
+hybrids of localized modeling (in some cases CBR) with other techniques,
+often metric learning. Though {[}27{]} and {[}15{]} report exceptions,
+the pattern suggests the importance of the similarity measure to the
+retrieval step. Meanwhile, when proposed approaches targeted cohort
+demarcation or choice of predictive model (statistical CBR {[}13{]};
+individualized logistic regression, decision tree, and random forest
+{[}14,16{]}), they did not consistently outperform comparators.
+
 \hypertarget{identified-needs}{%
 \subsection{Identified needs}\label{identified-needs}}
+
+\label{sec:needs}
 
@@
 global and from localized models {[}15{]}. Most of this incremental work
-was indeed carried out by later researchers whose output appears in our
-sample.
+was indeed carried out in later investigations in our sample.
+
+\hypertarget{public-availability}{%
+\subsection{Public availability}\label{public-availability}}
+
+\label{sec:availability}
+
+\textbf{Discuss public availability of code or implementations.}
 
@@
 \section{Discussion}\label{discussion}}
+
+We draw several observations from our encoding of the included studies.
+We first focus on two driving questions: What challenges did
+investigators set out to solve? How did their approaches meet these
+challenges? We then comment on the connectedness of the included studies
+and the potential for recombination.
 
@@
 
-\hypertarget{trends}{%
-\subsection{Trends}\label{trends}}
-
-We draw several observations from our encoding of the included studies.
-We first focus on two driving questions: What challenges did
-investigators set out to solve? How did their approaches meet these
-challenges? We then comment on the connectedness of the included studies
-and the potential for recombination.
-
-\textbf{Add 1--2 paragraphs.}
+\hypertarget{study-designs}{%
+\subsection{Study designs}\label{study-designs}}
+
+Consistent with their orientation, almost all included studies were
+conducted using clinical data, only occasionally together with
+patient-reported (2), laboratory (2), and image (1) data. We suggest
+three reasons for this: First, this literature traces back to the 1990s,
+before -omic data could be generated cost-effectively at scale. Second,
+CBR in particular has a strong tradition in clinical decision support,
+where the focus of our sample remains throughout the review period.
+Third, because most -omic data are highly homogeneous---all measurements
+are made along or can be transformed to a common scale, e.g.~greyscale
+pixellations for X-ray images and transcripts per million for RNA-seq
+data---more deeply theoretical analysis techniques have been developed
+and come into wide use. While variations on correlation-based approaches
+like EHR-based phenome-wide association studies and similarity-based
+methods like CBR itself have been developed, more mechanistic and
+probabilistic tools have not become a domain standard.
 
@@
 here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning technique {[}43{]} to
-obtain a similarity measure on their set of dermoscopy and confocal
+appeal to a supervised distance metric learning (DML) technique {[}43{]}
+to obtain a similarity measure on their set of dermoscopy and confocal
 images that most effectively separates malignant from benign melanoma
@@
 have not.
+
+\hypertarget{directions-and-expectations}{%
+\subsection{Directions and
+expectations}\label{directions-and-expectations}}
+
+As discussed in Section \ref{sec:rationales}, most studies were premised
+on the potential predictive value of localized models. As cautioned in
+Section \ref{sec:performance}, not all experiments affirmed this
+premise. Most of those that did involved using metric learning to
+improve retrieval and sometimes iterative optimization of the metric and
+of the localized model. Some studies showcased results using a variety
+of such specializations {[}15,27,30{]}. However, because these advanced
+implementations were mostly compared against their precursors or
+commonplace alternatives, we cannot speak to their performance or any
+trade-offs with respect to each other. This would be a worthy goal of
+future work, though it might be limited by the lack of open-source code
+or public implementations (Section \ref{sec:availability}).
+
+While most studies identified outstanding technical needs (Section
+\ref{sec:needs}), few emphasized the need to assess human-focused
+qualities like user interface and user experience, interpretability,
+meaningfulness, or trust. Most studies also identified interpretability
+as an advantage of localized models, though none evaluated model
+interpretability and few proposed new interpretative uses. Lack of trust
+in ML tools is a long-recognized problem that direct intepretability of
+model components could help alleviate, but this is more often assumed
+than demonstrated. One valuable direction for future work would be to
+compare the ability of non-technical investigators or users to draw
+correct inferences from these and other interpretable methods and their
+levels of confidence in those inferences.
+
+We posit that an essential component of such ``soft'' performance goals
+is the ability of the user community to exert some control over the
+models. Given the impact on performance of the choice or optimization of
+the similarity measure, an important target for user input would be the
+importance of certain variables in the calculation of similarity, with
+an understanding of how it can impact not only the performance of the
+predictions but also the cohort retrieved for the model. For one
+example, a user may want to minimize the weight of rare diseases in
+medical history in order to retrieve a population with more such cases
+in order to better measure their associated risk to the outcome. In
+contrast, they may want to increase the weight of the indicating
+diagnosis in order to allow fewer patients from similar but distinct
+populations to influence the model. For another example, a user may want
+to down-weight socioeconomic variables like race--ethnicity in order to
+ensure a more diverse modeling cohort. Future purely quantitative work
+could assess whether similarity-tuning can achieve these ends more
+flexibly than strict inclusion/exclusion criteria.
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\section{Data and code availability}\label{data-and-code-availability}}
-
-\label{sec:availability}
+\hypertarget{data-and-code}{%
+\section{Data and code}\label{data-and-code}}
+
+\label{sec:data}
 
@@
 
+\hypertarget{performance-evaluations-and-comparisons}{%
+\subsection{Performance evaluations and
+comparisons}\label{performance-evaluations-and-comparisons}}
+
+\begin{landscape}
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1224}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0714}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1633}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2398}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0969}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2092}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0969}}@{}}
+\caption{\label{tab:performance}Evaluations of proposed methods and
+comparisons to alternative methods. (See original studies for full names
+and descriptions.)}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Measure
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Proposals
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Comparators
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Measure
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Problem
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Proposals
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Comparators
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+{[}20{]} & Accuracy & Arthritis & Kohonen + CBR & 75.60\%\hspace{6em} &
+Kohonen mapping & 54.70\%\hspace{6em} \\
+& & & & \hspace{6em} & CBR & 60.40\%\hspace{6em} \\
+& & & & \hspace{6em} & Quest & 50.90\%\hspace{6em} \\
+& & & & \hspace{6em} & backpropagation & 54.70\%\hspace{6em} \\
+{[}13{]} & & Dermatology & Statistical CBR & 96.57\%\hspace{6em} & CBR &
+95.43\%\hspace{6em} \\
+& & & & \hspace{6em} & C5.0 & 93.71\%\hspace{6em} \\
+& & & & \hspace{6em} & CART & 92.00\%\hspace{6em} \\
+& & & & \hspace{6em} & LR & 91.14\%\hspace{6em} \\
+& & & & \hspace{6em} & NN & 88.29\%\hspace{6em} \\
+& & Heart & Statistical CBR & 83.70\%\hspace{6em} & LR &
+84.44\%\hspace{6em} \\
+& & & & \hspace{6em} & NN & 82.22\%\hspace{6em} \\
+& & & & \hspace{6em} & CBR & 81.85\%\hspace{6em} \\
+& & & & \hspace{6em} & CART & 76.67\%\hspace{6em} \\
+& & & & \hspace{6em} & C5.0 & 76.30\%\hspace{6em} \\
+& & Breast & Statistical CBR & 96.96\%\hspace{6em} & NN &
+97.50\%\hspace{6em} \\
+& & & & \hspace{6em} & CBR & 96.61\%\hspace{6em} \\
+& & & & \hspace{6em} & LR & 96.61\%\hspace{6em} \\
+& & & & \hspace{6em} & C5.0 & 94.29\%\hspace{6em} \\
+& & & & \hspace{6em} & CART & 92.14\%\hspace{6em} \\
+& & Diabetes & Statistical CBR & 76.32\%\hspace{6em} & LR &
+77.11\%\hspace{6em} \\
+& & & & \hspace{6em} & CBR & 73.16\%\hspace{6em} \\
+& & & & \hspace{6em} & C5.0 & 73.15\%\hspace{6em} \\
+& & & & \hspace{6em} & CART & 72.89\%\hspace{6em} \\
+& & & & \hspace{6em} & NN & 65.39\%\hspace{6em} \\
+& & Liver & Statistical CBR & 66.76\%\hspace{6em} & CART &
+67.94\%\hspace{6em} \\
+& & & & \hspace{6em} & LR & 67.35\%\hspace{6em} \\
+& & & & \hspace{6em} & C5.0 & 66.47\%\hspace{6em} \\
+& & & & \hspace{6em} & CBR & 60.88\%\hspace{6em} \\
+& & & & \hspace{6em} & NN & 55.59\%\hspace{6em} \\
+{[}21{]} & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
+7.74\hspace{6em} \\
+& & & & \hspace{6em} & MLP & 8.38\hspace{6em} \\
+& & & & \hspace{6em} & ANFIS & 7.40\hspace{6em} \\
+& & & & \hspace{6em} & DENFIS & 7.22\hspace{6em} \\
+& & & & \hspace{6em} & TNFI & 7.28\hspace{6em} \\
+{[}22{]} & AUROC & mammography mass region & CBR & 0.885\hspace{6em} &
+DT & 0.872\hspace{6em} \\
+& & & & \hspace{6em} & ANN & 0.882\hspace{6em} \\
+{[}23{]} & Accuracy & Return-to-work & CBR & 62.5\%\hspace{6em} & LR &
+71.2\%\hspace{6em} \\
+{[}18{]} & & Colon cancer diagnosis & TWNFI & 91.9\%\hspace{6em} & MLR
+(Personalized) & 82.3\%\hspace{6em} \\
+& & & & \hspace{6em} & SVM (Personalized) & 90.3\%\hspace{6em} \\
+& & & & \hspace{6em} & WKNN & 90.3\%\hspace{6em} \\
+& & Crohn's disease risk & TWNFI & 82.45\%\hspace{6em} & &
+\hspace{6em} \\
+{[}26{]} & & Colon cancer & knnGSA & 85.48\%\hspace{6em} & SVM &
+82.14\%\hspace{6em} \\
+& & & svmGSA & 87.10\%\hspace{6em} & ECF & 72.30\%\hspace{6em} \\
+& & & & \hspace{6em} & KNN & 82.14\%\hspace{6em} \\
+& & & & \hspace{6em} & WKNN & 82.14\%\hspace{6em} \\
+& & Leukemia & knnGSA & 97.22\%\hspace{6em} & SVM &
+95.83\%\hspace{6em} \\
+& & & svmGSA & 97.22\%\hspace{6em} & ECF & 94.44\%\hspace{6em} \\
+& & & & \hspace{6em} & KNN & 94.44\%\hspace{6em} \\
+& & & & \hspace{6em} & WKNN & 94.44\%\hspace{6em} \\
+& & Lymphoma & knnGSA & 94.81\%\hspace{6em} & SVM &
+93.51\%\hspace{6em} \\
+& & & svmGSA & 94.81\%\hspace{6em} & ECF & 92.21\%\hspace{6em} \\
+& & & & \hspace{6em} & KNN & 93.51\%\hspace{6em} \\
+& & & & \hspace{6em} & WKNN & 93.51\%\hspace{6em} \\
+& & Lung cancer & knnGSA & 98.34\%\hspace{6em} & SVM &
+95.31\%\hspace{6em} \\
+& & & svmGSA & 98.90\%\hspace{6em} & ECF & 92.87\%\hspace{6em} \\
+& & & & \hspace{6em} & KNN & 96.60\%\hspace{6em} \\
+& & & & \hspace{6em} & WKNN & 95.50\%\hspace{6em} \\
+{[}12{]} & Relative IPEC & Graft survival & MKNN & 0.973\hspace{6em} &
+RSF & 0.957\hspace{6em} \\
+{[}27{]} & AUROC & Waitlist registration, complete & CBR + Wj &
+89.8\%\hspace{6em} & LR & 92.0\%\hspace{6em} \\
+& & & CBR + Wk & 90.7\%\hspace{6em} & Standalone CBR &
+90.4\%\hspace{6em} \\
+& & & CBR + Wj + Wk & 87.4\%\hspace{6em} & & \hspace{6em} \\
+& & Waitlist registration, stepwise & CBR + Wj & 82.7\%\hspace{6em} & LR
+& 92.1\%\hspace{6em} \\
+& & & CBR + Wk & 86.2\%\hspace{6em} & Standalone CBR &
+91.4\%\hspace{6em} \\
+& & & CBR + Wj + Wk & 88.5\%\hspace{6em} & & \hspace{6em} \\
+{[}28{]} & Accuracy & Nevus & Collaborative + rules & 98\%\hspace{6em} &
+Dermoscopy CBR & 92\%\hspace{6em} \\
+& & & Collaborative + rules + DML & 100\%\hspace{6em} & Confocal CBR &
+96\%\hspace{6em} \\
+& & & & \hspace{6em} & Collaborative & 95\%\hspace{6em} \\
+{[}15{]} & AUROC & Diabetes onset & Personalized LR
+(LSML+FeatureFiltering) & 0.624\hspace{6em} & KNN & 0.617\hspace{6em} \\
+& & & Personalized LR (LSML) & 0.619\hspace{6em} & Global LR &
+0.611\hspace{6em} \\
+& & & Personalized LR (Euclidean) & 0.614\hspace{6em} & &
+\hspace{6em} \\
+& & & Personalized LR (Random) & 0.602\hspace{6em} & & \hspace{6em} \\
+{[}14{]} & & 30-day mortality & Individualized LR & 0.830\hspace{6em} &
+Individualized DC & 0.797\hspace{6em} \\
+& & & Individualized DT & 0.753\hspace{6em} & & \hspace{6em} \\
+{[}16{]} & & & Individualized LR & 0.824\hspace{6em} & Individualized DC
+& 0.801\hspace{6em} \\
+& & & Individualized DT & 0.779\hspace{6em} & & \hspace{6em} \\
+& & & Individualized RF & 0.839\hspace{6em} & & \hspace{6em} \\
+& & & Individualized CSRF & 0.832\hspace{6em} & & \hspace{6em} \\
+{[}31{]} & Accuracy & J13/pneumonia & Case-based NN & 81.6\%\hspace{6em}
+& & \hspace{6em} \\
+& & K80.1/gallbladder & Case-based NN & 76.7\%\hspace{6em} & &
+\hspace{6em} \\
+& & H25.1/age-related cataract & Case-based NN & 94.9\%\hspace{6em} & &
+\hspace{6em} \\
+& & H26.2/complicated cataract & Case-based NN & 91.4\%\hspace{6em} & &
+\hspace{6em} \\
+& & 167.4/encephalopathy & Case-based NN & 72.4\%\hspace{6em} & &
+\hspace{6em} \\
+& & 167.9/CBD & Case-based NN & 75.4\%\hspace{6em} & & \hspace{6em} \\
+& & N20.1/ureter & Case-based NN & 58.7\%\hspace{6em} & &
+\hspace{6em} \\
+{[}30{]} & AUROC & Cognitive impairment & Gaussian process regression
+with custom kernel & 0.92\hspace{6em} & linear regression without
+regularization & 0.81\hspace{6em} \\
+& & & & \hspace{6em} & LASSO regression & 0.85\hspace{6em} \\
+& & & & \hspace{6em} & ridge regression & 0.87\hspace{6em} \\
+& & & & \hspace{6em} & RF & 0.90\hspace{6em} \\
+& & & & \hspace{6em} & gradient boosting regression tree &
+0.91\hspace{6em} \\
+& & & & \hspace{6em} & XGBoost & 0.91\hspace{6em} \\
+& & & & \hspace{6em} & kernel support vector regressor &
+0.92\hspace{6em} \\
+& & Parkinson's disease & Gaussian process regression with custom kernel
+& 0.875\hspace{6em} & linear regression without regularization &
+0.820\hspace{6em} \\
+& & & & \hspace{6em} & ridge regression & 0.798\hspace{6em} \\
+& & & & \hspace{6em} & RF & 0.734\hspace{6em} \\
+& & & & \hspace{6em} & gradient boosting regression tree &
+0.825\hspace{6em} \\
+& & & & \hspace{6em} & kernel support vector regressor &
+0.778\hspace{6em} \\
+{[}17{]} & & Diabetes & Personalized RF & 0.90\hspace{6em} & &
+\hspace{6em} \\
+& & & Personalized KNN & 0.82\hspace{6em} & & \hspace{6em} \\
+& & & Personalized LR & 0.89\hspace{6em} & & \hspace{6em} \\
+{[}32{]} & & Discharge at 10 days & One-class JITL-ELM &
+0.8510\hspace{6em} & ELM & 0.4973\hspace{6em} \\
+& & & & \hspace{6em} & JITL-ELM & 0.2014\hspace{6em} \\
+& & & & \hspace{6em} & One-class ELM & 0.7588\hspace{6em} \\
+& & & & \hspace{6em} & One-class SVM & 0.4647\hspace{6em} \\
+\end{longtable}
+
+\end{landscape}
+
 \hypertarget{references}{%

### 2023-09-07 scr/synthesis.r

@@ -4,7 +4,7 @@ library(ggmosaic)
 #' Read data.
 
 # journals of references in previous reviews
-sr_journals <- read_xlsx(here::here("data/SR-journals.xlsx"))
+sr_journals <- readxl::read_xlsx(here::here("data/SR-journals.xlsx"))
 
 # scrape and format table summarizing studies included in synthesis
 # Excel file downloaded from Google Sheet:
@@ -33,49 +33,6 @@ here::here("docs/Manuscript/Manuscript.tex") %>%
     ref = str_replace(line, "^.*\\{ref-([A-Za-z]+[0-9]{4}[a-z]*)\\}.*$", "\\1")
   ) %>%
   print() -> citation_numbers
-
-#' Print summaries.
-
-# journal frequencies of previous reviews' references
-sr_journals %>%
-  mutate(journal = str_remove(journal, "^[0-9]{4} *")) %>%
-  mutate(journal = str_remove(journal, "^[0-9]{1,2}[a-z]{2} *")) %>%
-  mutate(journal = str_remove(journal, " *\\([A-Z0-9]+\\)$")) %>%
-  mutate(journal = tolower(journal)) %>%
-  group_by(journal) %>%
-  count() %>%
-  drop_na() %>%
-  arrange(desc(n)) %>%
-  filter(n > 2L) %>%
-  print(n = Inf)
-
-# journal frequencies of synthesized studies
-properties_inclusion |> 
-  select(journal = Journal) |> 
-  mutate(journal = str_remove(journal, "^[0-9]{4} *")) %>%
-  mutate(journal = str_remove(journal, "^[0-9]{1,2}[a-z]{2} *")) %>%
-  mutate(journal = str_remove(journal, " *\\([A-Z0-9]+\\)$")) %>%
-  mutate(journal = tolower(journal)) %>%
-  group_by(journal) %>%
-  count() %>%
-  drop_na() %>%
-  arrange(desc(n)) %>%
-  print(n = Inf)
-
-# terms
-composite_studies %>%
-  select(citation = Citation, terms = Terminology) %>%
-  unnest(terms) %>%
-  print() -> composite_terms
-# frequency table
-composite_terms %>%
-  mutate(terms = tolower(terms)) %>%
-  mutate(terms = str_replace(terms, "modeling", "model")) %>%
-  mutate(terms = str_replace(terms, "sampling", "sample")) %>%
[…]

### 2023-09-08 docs/Manuscript/Manuscript.Rmd
transcribe edits to paper copy

@@
 Entries were published from 1997 to 2021, with half of these since 2015.
-The most common tasks were prognosis and diagnosis, with only few recent studies focusing on care recommendation.
-The studies overwhelmingly focused on clinical data, occasionally also including laboratory and image-derived data.
-Only three studies used localized models for any task other than prediction, most recently to identify significant determinants behind recommendation outputs.
-Several studies commented on memory and runtime costs, and four proposed partial solutions to them.
-While most techniques were not replicated and did not cite predecessors, a general technique that specializes to most of those encountered was straightforward to describe.
+The most common tasks were prognosis and diagnosis.
+The studies overwhelmingly focused on clinical data, occasionally including laboratory and image-derived data.
+Only three studies applied localized models to tasks other than prediction.
+Several studies commented on memory and runtime costs, and four proposed partial solutions.
+While few methods were reused and no results were reproduced, a general technique that specializes to most of those encountered was straightforward to describe.
 
   Conclusions:
-  Localized models hold promise to improve the performance of models used in clinical decision support tools while maintaining interpretability, provided their computational limitations can be overcome.
-Our search undoubtedly failed to capture all studies that used methods of this kind, but those reviewed and synthesized exhibit consistent enough design to be reproduced by a general-purpose tool.
-We hope that our review and synthesis will assist the development of more efficient implementations suitable for integration into practice.
+  Localized models have potential to improve the performance of models used in clinical decision support tools while maintaining interpretability, provided computational hurdles can be overcome.
+Our search undoubtedly failed to capture all studies that used methods of this kind, but those we found follow a consistent enough design to be reproduced by a general-purpose tool.
+We hope that our review and synthesis will spur future work on efficiency, reproducibility, and user interactions and needs.
 keywords: 
@@
 Similarity-matching powers the retrieval step of most case-based reasoning (CBR) systems and provides the neighborhoods from which nearest neighbor (NN) models derive predictions.
-Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), traditional CBR tools return the retrieved cases to the user to inform human decision-making.
+Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), traditional CBR tools return the retrieved cases to inform human decision-making.
 
@@
 
-We discuss in this review a hybrid computational approach we call _localized modeling_ that borrows elements from both clinical application areas and appears to have been introduced independently several times.
+We discuss in this review a hybrid computational approach we call _localized modeling_ that appears to have been introduced independently several times.
 The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching.
-Whereas CBR and NN prediction use patient similarity to obtain an ideally homogeneous subpopulation matched to the index case, in localized modeling one seeks a relevant yet heterogeneous subpopulation, each member of which deviates slightly from the index case.
+Whereas CBR and NN prediction use patient similarity to obtain an ideally homogeneous subpopulation matched to the index case, in localized modeling one seeks a relevant yet heterogeneous subpopulation, each member of which may deviate from the index case.
+A prediction or inference is then obtained from a model fitted to this _similarity cohort_.
+The use of similarity cohorts distinguishes localized modeling from classical GLMs that still dominate analysis in clinical trials and comparative effectiveness research, as well as from many cutting-edge ML models that are fitted to whole populations, while the cohorts' heterogeneity and its utility to the model distinguish the approach from CBR.
 In principle, this allows interpretable models like GLMs to identify and measure the effects of determining factors that are specific to the index case.
-The use of similarity cohorts distinguishes localized modeling from classical GLMs that still dominate analysis in clinical trials and comparative effectiveness research, as well as from many cutting-edge ML models that are fitted to whole populations, while the cohorts' heterogeneity and its utility to the model distinguish the approach from CBR.
 In the reviewed studies, localized modeling shows some promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.
@@
 The motivation for this scoping review is to describe the settings in and problems with which localized models have been tasked to date, and to synthesize existing methods into a general framework.
-Along the way, we attempt to reconcile terminology, summarize evaluation techniques, propose needed follow-up work, and recommend best practices.
+Along the way, we attempt to reconcile terminology, summarize motivations and evaluations, and propose valuable follow-up work.
 
@@
 Clinical CBR emerged among rule-based approaches and other AI tools in the development of expert systems [@Aamodt1994].
-Early implementations realized a general workflow described as "the four REs", later the "R4 cycle" [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system retrieves one or more past cases (solved problems) from a corpus, reuses these to generate an understanding of (solution to) the new case, revises this understanding to better fit the new case (sometimes called "adaptation"), and retains the new case and its eventual understanding in the corpus to be retrieved and reused in future.
+Early implementations realized a general workflow described as "the four REs", later the "R4 cycle" [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system _retrieves_ one or more past cases (solved problems) from a corpus, _reuses_ these to generate an understanding of (solution to) the new case, _revises_ this understanding (solution) to better fit the new case (sometimes called "adaptation"), and _retains_ the new case and its eventual understanding (solution) in the corpus to be retrieved and reused in future.
 In contrast to rule-based systems, which are variable-based and generally interpretable as a single rule applied to all new cases, case-based systems provide case-specific interpretations in the form of a number of more fully understood reference cases.
@@
 In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort can be understood as an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a ML workflow.
-Accordingly, localized models have been primarily used for and evaluated on their ability to more accurately predict classes or outcomes.
-Nevertheless, as some more recent applications have begun to show, they can be used to draw individualized inferences about the importance of different risk factors.
+Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
+Nevertheless, as some recent applications have begun to show, they can also be used to draw individualized inferences about the importance of different risk factors.
+While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the effect estimates of a GLM.
 While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
@@
 While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools).
-For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive modeling.
+For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.
 
@@
 
-* Pulls case-level data from a corpus of past cases with known response values\newline
+* Uses case-level data from a corpus of past cases with known responses\newline
 (response may be outcome, diagnosis, subtype, etc.)
 * Defines a numeric multivariate case similarity measure\newline
-(include integer-valued measures)
-* Uses the similarity measure to select cohorts for index cases from the corpus\newline
+(allow integer-valued measures)
+* Uses the similarity measure to retrieve cohorts for index cases from the corpus\newline
 (for example, based on a training–testing partition)
-* Fits statistical models to cohorts from which to draw inferences about index cases\newline
-(outcome predictions count, survival estimates, risk factor effect estimates, model evaluation statistics, etc.)
+* Fits statistical models to cohorts from which to make predictions or draw inferences about index cases\newline
+(outcome predictions, survival estimates, risk factor contributions, model evaluation statistics, etc.)
 
@@
 We include a PRISMA checklist as Supplemental Table 1 and a PRISMA abstract checklist as Supplemental Table 2.
-Because we focus on methodologies rather than approaches, we deviate in some ways from PRISMA guidelines.
-In particular, many items of the checklist, involving bias assessment and quantitative synthesis, are intended for meta-analyses and, therefore, did not apply to this study.
+Because we focus on methodologies rather than conceptual approaches or evidence, we deviate in some ways from PRISMA guidelines.
+In particular, those items of the checklist involving bias assessment and quantitative synthesis are intended for meta-analyses so did not apply to this study.
 Because we are not aware of any standard procedures for conducting reviews and syntheses of methodology, no protocol was prepared for this study.
@@
 Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
-We finalized the PubMed search first, then adapted it to the other platforms (Sections \ref{sec:appendix-search}).
+We finalized the PubMed search first, then adapted it to the other platforms (Section \ref{sec:appendix-search}).
 
@@
 
-Within Covidence, we screened titles and abstracts for relevance.
-We decided on four eligibility criteria to expedite the screening process.
-These include articles written in English for readability, original studies to strictly include primary sources with new information/conclusions, medical or clinical related settings due to the applicability of our data to this field, and alternative use of the key terms within the search string, meaning that it was clear from the title and abstract that the entry was included due to the use of one or more search terms with a different meaning than ours.
-Each of two authors (AC and JCB) screened roughly half of the entries.
-They regularly reviewed each other’s decisions to improve consistency.
-In cases of uncertainty, an entry would be included.
+Within Covidence, we screened titles and abstracts for relevance. We decided on four eligibility criteria to expedite the screening process.
+Articles must be written in English, for readability; they must be original studies, to exclude secondary sources with duplicate information; their use settings must be medical, clinical, or related, to keep our review topical; and it must not be clear that their use of our search terms was different from our intended meaning.
+Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, entries were included.
 
@@
 * Uses the similarity measure to select cohorts for index cases from the corpus
-* Fits statistical models to cohorts to make inferences about index cases
+* Fits statistical models to cohorts to make predictions or inferences about index cases
 
@@
 
-In Covidence, two authors (AC and JCB) independently evaluated each manuscript for the four eligibility criteria.
+In Covidence, two authors (AC and JCB) independently evaluated each manuscript for these eligibility criteria.
 The first criterion that a manuscript failed was designated the reason for exclusion.
@@
 These approaches were novel to CBR and presage later developments, so were helpful to understanding the development of localized modeling.
-However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the responses of the cohort.
-A review of all studies that use nearest neighbors prediction, even in the clinical domain, would be impractical.
+However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the cohort's responses.
+A review of all studies that use nearest neighbors prediction would be impractical.
 Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to exclude those studies whose approach was equivalent to nearest neighbors prediction using a conventional similarity measure.
@@
 Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{sec:data}).
-The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, evaluation/performance/comparison measures (& their values), name given to modeling approach).
+The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, performance measures, results of evaluations and comparisons, name given to modeling approach).
 We used these data to detect and visualize trends amongst the included studies.
@@
 
-Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize possibly independent research efforts that have converged on a common modeling technique.
+Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common modeling technique.
 The goal will be to describe a unified framework for localized models that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.
@@
 We then reviewed 43 studies comprising a seed set that inspired this review.
-After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using nearest neighbors prediction.
-Finally, citations in reference lists from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
+After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using NN prediction.
+Finally, reference tracking from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
 This left us with 19 + 1 + 5 = 25 studies included in the review and synthesis.
@@
 ## Bibliographic and methodological properties
+
+**Update numbers below, perhaps dynamically.**
 
@@
 
-Another interesting pattern lay in the years of publication of the included studies (Figure \ref{fig:year}).
-While they stretch back to the late 1990s, they have become more common, which suggests that this is still an active, though not rapidly expanding, methodological approach.
+Another interesting pattern lay in the years of publication (Figure \ref{fig:year}).
+While they trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, methodological approach.
 We note that, while CBR in health and medicine originated as early as 1990, these studies did emerge early on once CBR had established itself.
@@
 We classified 7 of the 25 studies as "knowledge", making the majority of studies from a "practice" standpoint.
-These studies aimed primarily to provide providers with tools to use in clinic or to improve outcomes for patients currently receiving care.
+These studies aimed to provide tools for use in clinic or to improve outcomes.
 
@@
 @Lowsky2013 made a different case, that the fewer parametric assumptions and complications of a CBR-style model would allow for greater accuracy.
-Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to model performance [@Song2006c; @Xu2008a].
+Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to performance [@Song2006c; @Xu2008a].
 
@@
 This "self-explanation capability" made outputs more intelligible to physicians in the CDS setting.
-@Ng2015 additional pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
+@Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
 
@@
 @Lowsky2013 used non-parametric models to reduce the computational burden of revision (prediction).
-@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning"), and @Ng2021b and @Tang2021c partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
+@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021b and @Tang2021c partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
 
@@
 Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007a; @Liang2015a; @Ng2015; @Zhang2018a].
-
 The frequency plot in Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
@@
 
-In the Discussion, we discuss trends in the literature we reviewed, identify common limitations and gaps in understanding, and attempt to synthesize their methods into a general framework.
-
 ## Performance assessments
@@
 
-However, we do observe one clear pattern: Of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], GSA [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]), all except the last are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
-Though @CampilloGimenez2012a and [@Ng2015] report exceptions, the pattern suggests the importance of the similarity measure to the retrieval step.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], GSA [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
+Though @CampilloGimenez2012a and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
 Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model (statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]), they did not consistently outperform comparators.
@@
 The study authors focused their recommendations and their own plans for future work mostly on technical improvements and evaluations [@Mariuzzi1997; @Yearwood1997a].
-Urged improvements included full or partial automation of user-dependent steps such as predictor selection [@Mariuzzi1997; @Yearwood1997a], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006c; @Lee2017]; extensions to new data structures [@Lopez2011a], data types [@Liang2015a; @Verma2015a; @Malykh2018a], and reasoning systems [@Nicolas2014a]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015a; @Zhang2018a].
+Urged improvements included full or partial automation of predictor selection [@Mariuzzi1997; @Yearwood1997a], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006c; @Lee2017]; extensions to new data structures [@Lopez2011a], data types [@Liang2015a; @Verma2015a; @Malykh2018a], and reasoning systems [@Nicolas2014a]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015a; @Zhang2018a].
 Authors also urged validations and independent evaluations using larger or more comprehensive data sets [@Elter2007a; @Xu2008a; @Verma2015a; @Ng2015], using data aggregated from multiple health systems [@Lee2015; @Lee2017; @Tang2021c; @Ng2021b], and in other care settings or disease contexts [@Song2006c; @Zhang2018a; @Tang2021c; @Ng2021b].
+
 Less common were calls to strengthen the connection between the methods and the users.
@@
 
-## Public availability
-\label{sec:availability}
+## Reproducibility
+\label{sec:reproducibility}
 
@@
 However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
-This would be a worthy goal of future work, though it might be limited by the lack of open-source code or public implementations (Section \ref{sec:availability}).
+This would be a worthy goal of future work, though it might be limited by the lack of open-source code or public implementations (Section \ref{sec:reproducibility}).
 

### 2023-09-09 docs/Manuscript/Manuscript.Rmd
make or note changes following email with adi

@@
   Background:
-  Applications of classical case-based reasoning (CBR) to clinical and health tasks have given rise to a family of techniques in which an (often conventional) model is fitted to a subset of labeled cases, selected for relevance or similarity to an unlabeled case to which the model is applied.
+  Applications of classical case-based reasoning (CBR) to clinical and health tasks have given rise to a family of techniques in which a statistical model is fitted to a subset of labeled cases, selected for relevance or similarity to an unlabeled case to which the model is applied.
 These localized models, as we propose to call them, may perform as well as "black-box" models while retaining direct interpretations.
@@
 
-We expected the general framework to boil down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family.
+We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family.
 Such a general method would be straightforward, if laborious, to implement, and specializations of it would account for the majority of real-world applications.
@@
 Our methods followed prepared plans, which we describe in detail in the Appendix.
+**Additional details are provided in Section \label{sec:appendix-selection}.**
 We also note deviations from these plans and the reasons for them.
@@
 
+**Correct "n = PubMed" in PRISMA diagram.**
 We include a PRISMA checklist as Supplemental Table 1 and a PRISMA abstract checklist as Supplemental Table 2.
@@
 ## Full text review
+\label{sec:full-text}
 
@@
 
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], GSA [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
 Though @CampilloGimenez2012a and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
@@
 
-**Write this section!**
+Most deduplication was done automatically in Covidence. As full-text review was done in Zotero, some additional duplicates were noticed and merged.
 
-1. Deduplication (Covidence)
-    a. Specific to that source: title, creator, URL, DOI
-    b. May overlap: journal, publication date, ISSN
-2. Title
+When abstracts were not obtained by search or by Covidence, we attempted to find them online using DOIs; when an abstract could not be found, screening was based on the title alone.
+We decided to screen titles and abstracts conservatively, rejecting only studies that were clearly outside the scope of our review.
+The reasons for rejection were four, as discussed in the main text:
     a. Exclude non clinical/non medical setting
@@
     d. Exclude if search term clearly has different meaning than intended
-3. Abstract
-4. Full Text
-    a. Study criteria (inclusion/exclusion)
+(Two examples of (d) are the use of the term "personalized model" to refer in some cases to parameterized models tuned to individual patient measurements and in others to patient-centered models of care.) Studies that passed title/abstract screen were exported to Zotero.
 
@@
 
-Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
+Some PDFs were obtained using Zotero from a university workstation, and the remaining were obtained through university library services.
+The review process is detailed in Section \label{sec:full-text}.
+
+Our selection of relevant papers from the search corpus was based on the following inclusion/exclusion criteria:
 
@@
 
+**TODO: Place vertical separators between major columns.**
+
 ```{r tab:performance, echo=FALSE, message=FALSE, warning=FALSE}
-# TODO: Place vertical separators between major columns.
 tab_eval_comp |> 

### 2023-09-09 docs/Manuscript/Manuscript.Rmd
continue addressing reviewer comments

--- 2023-09-07 docs/Manuscript/Manuscript.Rmd
@@
 Accordingly, localized models have been primarily used for and evaluated on their ability to more accurately predict classes or outcomes.
-Nevertheless, as some more recent applications have begun to show, they can be used to draw individualized inferences about the importance of different risk factors.
-While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
+Nevertheless, as some more recent applications have begun to show, they can be used to draw inferences about the importance of different risk factors to specific individuals.
+**Distinguish between interpretable model components like effect estimates and informative but not directly interpretable importance measures as are commonly computed for 'black-box' models.** While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
 Provided the model family fitted to the cohorts is interpretable, the localized model as a whole inherits this property.
@@
 
-Our methods followed prepared plans, which we describe in detail in the Appendix.
-We also note deviations from these plans and the reasons for them.
+We describe the steps of our review in detail in the Appendix.
+We also note deviations from plans and the reasons for them.
 
@@
 **Discuss public availability of code or implementations.**
+**Methods, results, and inferential reproducibility [@Matarese2022].**
 
@@
 We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
-Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^federated]
+Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^local]
 
-[^federated]: As one reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
+[^local]: We note, in response to one reviewer's comment, that for many if not most models used in ML, excepting GLMs, a predictor may play a different and more or less important role for some cases than others, as quantified by importance measures or model-agnostic explanations [@Biecek2021; @Molnar2023].
+In this way, such models can also be viewed as "localized".
+We recognize that the use of global and local explanations is standard.
+As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
 The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
-At present, however, we feel that these research areas are sufficiently disjoint to avoid confusion in practice.
+At present, we feel that these methods are sufficiently disjoint to avoid confusion in practice, but we suggest the more specific term "similarity cohort--based models" when the term "local" is overloaded.
 

### 2023-09-09 docs/Manuscript/Manuscript.tex
continue addressing reviewer comments

@@
 some more recent applications have begun to show, they can be used to
-draw individualized inferences about the importance of different risk
-factors. While the patient similarity measure used to retrieve each
-cohort may be complicated, the cohort itself can be directly inspected
-by the user. Provided the model family fitted to the cohorts is
-interpretable, the localized model as a whole inherits this property.
-Thus, localized models may embody both styles of CBR.
+draw inferences about the importance of different risk factors to
+specific individuals. \textbf{Distinguish between interpretable model
+components like effect estimates and informative but not directly
+interpretable importance measures as are commonly computed for
+`black-box' models.} While the patient similarity measure used to
+retrieve each cohort may be complicated, the cohort itself can be
+directly inspected by the user. Provided the model family fitted to the
+cohorts is interpretable, the localized model as a whole inherits this
+property. Thus, localized models may embody both styles of CBR.
 
@@
 
-Our methods followed prepared plans, which we describe in detail in the
-Appendix. We also note deviations from these plans and the reasons for
-them.
+We describe the steps of our review in detail in the Appendix. We also
+note deviations from plans and the reasons for them.
 
@@
 \textbf{Discuss public availability of code or implementations.}
+\textbf{Methods, results, and inferential reproducibility
+{[}\textbf{Matarese2022?}{]}.}
 
@@
 ``localized models'' as a suitable term for this counterpart to
-globally-fitted models.\footnote{As one reviewer pointed out, the
-  terminology of local and global is also used in the unrelated context
-  of federated ML to distinguish models trained using data stored in a
-  single storage device (local) from their aggregations (global)
-  {[}39,40{]}. The prospects for federated architecture to more
+globally-fitted models.\footnote{We note, in response to one reviewer's
+  comment, that for many if not most models used in ML, excepting GLMs,
+  a predictor may play a different and more or less important role for
+  some cases than others, as quantified by importance measures or
+  model-agnostic explanations {[}39,40{]}. In this way, such models can
+  also be viewed as ``localized''. We recognize that the use of global
+  and local explanations is standard. As another reviewer pointed out,
+  the terminology of local and global is also used in the unrelated
+  context of federated ML to distinguish models trained using data
+  stored in a single storage device (local) from their aggregations
+  (global) {[}41,42{]}. The prospects for federated architecture to more
   effectively implement similarity cohort--based models are intriguing,
-  and would require some reconciliation of terms. At present, however,
-  we feel that these research areas are sufficiently disjoint to avoid
-  confusion in practice.}
+  and would require some reconciliation of terms. At present, we feel
+  that these methods are sufficiently disjoint to avoid confusion in
+  practice, but we suggest the more specific term ``similarity
+  cohort--based models'' when the term ``local'' is overloaded.}
 
@@
 used in 8 studies that tie in to the much larger literature on metric
-learning {[}41{]}. These techniques are designed to detect structure in
+learning {[}43{]}. These techniques are designed to detect structure in
 data, especially associations between predictors and responses, and to
@@
 (Table \ref{tab:composite}). Based on the key properties of metric
-learning algorithms identified by {[}42{]}, most learned measures were
+learning algorithms identified by {[}44{]}, most learned measures were
 non-linear while some were local, and some were optimized globally while
@@
 here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning (DML) technique {[}43{]}
+appeal to a supervised distance metric learning (DML) technique {[}45{]}
 to obtain a similarity measure on their set of dermoscopy and confocal
@@
 conventional CBR they called ``statistical CBR''.\footnote{Drawing from
-  {[}44{]}, we suggest the term ``combinatorial CBR'' for the
+  {[}46{]}, we suggest the term ``combinatorial CBR'' for the
   then-conventional approach using similarity cohorts of fixed
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
+\CSLLeftMargin{{[}39{]} }%
+\CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
+{Explore}, {Explain}, and {Examine Predictive Models}. {New York}:
+{Chapman and Hall/CRC}; 2021.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
+\CSLLeftMargin{{[}40{]} }%
+\CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
+for {Making Black Box Models Explainable}. 2023.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
-\CSLLeftMargin{{[}39{]} }%
+\CSLLeftMargin{{[}41{]} }%
 \CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
-\CSLLeftMargin{{[}40{]} }%
+\CSLLeftMargin{{[}42{]} }%
 \CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-\CSLLeftMargin{{[}41{]} }%
+\CSLLeftMargin{{[}43{]} }%
 \CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-\CSLLeftMargin{{[}42{]} }%
+\CSLLeftMargin{{[}44{]} }%
 \CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-\CSLLeftMargin{{[}43{]} }%
+\CSLLeftMargin{{[}45{]} }%
 \CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-\CSLLeftMargin{{[}44{]} }%
+\CSLLeftMargin{{[}46{]} }%
 \CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A

### 2023-09-09 docs/Manuscript/Manuscript.Rmd

--- 2023-09-09 docs/Manuscript/Manuscript.Rmd
@@
 
-We discuss in this review a hybrid computational approach we call _localized modeling_ that appears to have been introduced independently several times.
+We discuss in this review a hybrid computational approach we call _localized modeling_ (see Section \ref{sec:coherence}) that appears to have been introduced independently several times.
 The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching.
@@
 Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
-Nevertheless, as some recent applications have begun to show, they can also be used to draw individualized inferences about the importance of different risk factors.
+Nevertheless, as some recent applications have begun to show, they can be used to draw inferences about the importance of different risk factors to specific individuals.
 While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the effect estimates of a GLM.
@@
 
-Our methods followed prepared plans, which we describe in detail in the Appendix.
-**Additional details are provided in Section \label{sec:appendix-selection}.**
-We also note deviations from these plans and the reasons for them.
+Here we describe our review process, including deviations from plans and the reasons for them.
+More details are included in Section \label{sec:appendix-selection}.
 
@@
 
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often metric learning.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
 Though @CampilloGimenez2012a and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
@@
 
-We draw several observations from our encoding of the included studies.
-We first focus on two driving questions: What challenges did investigators set out to solve?
-How did their approaches meet these challenges?
-We then comment on the connectedness of the included studies and the potential for recombination.
+After commenting briefly on our review process, we draw several observations from our encoding of the included studies.
+We first focus on their methodological choices and innovations, then on their conventions and language.
+We spend most of the section discussing the differences and commonalities of the techniques used, toward a general method that encompasses most or all cases.
+We then suggest some avenues for future studies and conclude with our key takeaways.
 
@@
 ## Coherence
+\label{sec:coherence}
 
@@
 The 25 studies included in our synthesis used composite techniques that we organized into two major types.
-One type, similarity learning, was used in 8 studies that tie in to the much larger literature on metric learning [@Yang2006].
-These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
-Seven studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
-Based on the key properties of metric learning algorithms identified by @Bellet2014, most learned measures were non-linear while some were local, and some were optimized globally while others locally.
-However, not all qualified as metric learning, since the resulting measures would not necessarily satisfy the triangle inequality.
-Each then used its similarity measure to retrieve training cases relevant to each testing case.
-Once retrieved, most took a standard NN approach to generating predictions from the cohort; two exceptions [@Vilhena2016a; @Tang2021c] proceeded to fit predictive models to the retrieved cohorts and are discussed below.
+One type, _similarity learning_, was used in 8 studies that tie in to the much larger literature on DML [@Yang2006].
+The other, _cohort thresholding_, focused on choosing or optimizing the manner in which cohorts were demarcated using the similarity measure.
 
@@
 The similarity learning approaches took a variety of forms.
+These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
+Seven studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
+Based on the key properties of DML algorithms identified by @Bellet2014, most learned measures were non-linear while some were locally learned, and some were optimized globally while others locally.
+Not all qualified as DML, since the resulting measures would not necessarily satisfy the triangle inequality.
+Each used its similarity measure to retrieve training cases relevant to each testing case.
+Once retrieved, most took a standard NN approach to generating predictions; two exceptions [@Vilhena2016a; @Tang2021c] fit predictive models to the retrieved cohorts.
+
 Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007a].
@@
 
-As noted above, these instances of similarity learning fit into a much larger literature on metric learning, which has seen widespread use in health informatics and modeling.
+As noted above, these instances of similarity learning fit into a much larger literature on DML, which has seen widespread use in health informatics and modeling.
 That these studies satisfied our review criteria is due to the novelty of the techniques at the time of publication and the detailed attention paid to their techniques by the authors.
@@
 The design of a localized modeling approach can be decomposed into three largely independent choices: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
-The choices are the _similarity measure_, the _retrieval step_, and the _statistical model_, respectively.
+The choices are the _similarity measure_, the _cohort retrieval_, and the _statistical model_, respectively.
 These are diagrammed in Figure \ref{fig:framework}.
@@
 We found no studies that tuned the calculation of similarity, though most uniform cohort sizes were tuned.
-Some earlier studies tuned generalized regression models, but more recent studies have not.
+Some earlier studies tuned generalized regression models, but more recent studies did not.
 

### 2023-09-09 docs/Manuscript/Manuscript.tex

@@
   clinical and health tasks have given rise to a family of techniques in
-  which an (often conventional) model is fitted to a subset of labeled
-  cases, selected for relevance or similarity to an unlabeled case to
-  which the model is applied. These localized models, as we propose to
-  call them, may perform as well as ``black-box'' models while retaining
-  direct interpretations. Their adoption has been hindered by fragmented
+  which a statistical model is fitted to a subset of labeled cases,
+  selected for relevance or similarity to an unlabeled case to which the
+  model is applied. These localized models, as we propose to call them,
+  may perform as well as ``black-box'' models while retaining direct
+  interpretations. Their adoption has been hindered by fragmented
   development and implementation challenges. Our goal in this scoping
@@
   2021, with half of these since 2015. The most common tasks were
-  prognosis and diagnosis, with only few recent studies focusing on care
-  recommendation. The studies overwhelmingly focused on clinical data,
-  occasionally also including laboratory and image-derived data. Only
-  three studies used localized models for any task other than
-  prediction, most recently to identify significant determinants behind
-  recommendation outputs. Several studies commented on memory and
-  runtime costs, and four proposed partial solutions to them. While most
-  techniques were not replicated and did not cite predecessors, a
-  general technique that specializes to most of those encountered was
-  straightforward to describe.
-
-  Conclusions: Localized models hold promise to improve the performance
-  of models used in clinical decision support tools while maintaining
-  interpretability, provided their computational limitations can be
+  prognosis and diagnosis. The studies overwhelmingly focused on
+  clinical data, occasionally including laboratory and image-derived
+  data. Only three studies applied localized models to tasks other than
+  prediction. Several studies commented on memory and runtime costs, and
+  four proposed partial solutions. While few methods were reused and no
+  results were reproduced, a general technique that specializes to most
+  of those encountered was straightforward to describe.
+
+  Conclusions: Localized models have potential to improve the
+  performance of models used in clinical decision support tools while
+  maintaining interpretability, provided computational hurdles can be
   overcome. Our search undoubtedly failed to capture all studies that
-  used methods of this kind, but those reviewed and synthesized exhibit
-  consistent enough design to be reproduced by a general-purpose tool.
-  We hope that our review and synthesis will assist the development of
-  more efficient implementations suitable for integration into practice.
+  used methods of this kind, but those we found follow a consistent
+  enough design to be reproduced by a general-purpose tool. We hope that
+  our review and synthesis will spur future work on efficiency,
+  reproducibility, and user interactions and needs.
   \end{abstract}
@@
 (classification), traditional CBR tools return the retrieved cases to
-the user to inform human decision-making.
+inform human decision-making.
 
 We discuss in this review a hybrid computational approach we call
-\emph{localized modeling} that borrows elements from both clinical
-application areas and appears to have been introduced independently
-several times. The approach fits ML models, including regularized
-families of generalized regression models (GLMs), to cohorts retrieved
-via similarity matching. Whereas CBR and NN prediction use patient
-similarity to obtain an ideally homogeneous subpopulation matched to the
-index case, in localized modeling one seeks a relevant yet heterogeneous
-subpopulation, each member of which deviates slightly from the index
-case. In principle, this allows interpretable models like GLMs to
-identify and measure the effects of determining factors that are
-specific to the index case. The use of similarity cohorts distinguishes
-localized modeling from classical GLMs that still dominate analysis in
-clinical trials and comparative effectiveness research, as well as from
-many cutting-edge ML models that are fitted to whole populations, while
-the cohorts' heterogeneity and its utility to the model distinguish the
-approach from CBR. In the reviewed studies, localized modeling shows
-some promise toward simultaneously achieving the performance
-improvements of ML and the interpretability of classical models.
+\emph{localized modeling} that appears to have been introduced
+independently several times. The approach fits ML models, including
+regularized families of generalized regression models (GLMs), to cohorts
+retrieved via similarity matching. Whereas CBR and NN prediction use
+patient similarity to obtain an ideally homogeneous subpopulation
+matched to the index case, in localized modeling one seeks a relevant
+yet heterogeneous subpopulation, each member of which may deviate from
+the index case. A prediction or inference is then obtained from a model
+fitted to this \emph{similarity cohort}. The use of similarity cohorts
+distinguishes localized modeling from classical GLMs that still dominate
+analysis in clinical trials and comparative effectiveness research, as
+well as from many cutting-edge ML models that are fitted to whole
+populations, while the cohorts' heterogeneity and its utility to the
+model distinguish the approach from CBR. In principle, this allows
+interpretable models like GLMs to identify and measure the effects of
+determining factors that are specific to the index case. In the reviewed
+studies, localized modeling shows some promise toward simultaneously
+achieving the performance improvements of ML and the interpretability of
+classical models.
 
@@
 to synthesize existing methods into a general framework. Along the way,
-we attempt to reconcile terminology, summarize evaluation techniques,
-propose needed follow-up work, and recommend best practices.
+we attempt to reconcile terminology, summarize motivations and
+evaluations, and propose valuable follow-up work.
 
@@
 ``R4 cycle'' {[}2,3{]}: Given a new case (or problem), the system
-retrieves one or more past cases (solved problems) from a corpus, reuses
-these to generate an understanding of (solution to) the new case,
-revises this understanding to better fit the new case (sometimes called
-``adaptation''), and retains the new case and its eventual understanding
-in the corpus to be retrieved and reused in future. In contrast to
-rule-based systems, which are variable-based and generally interpretable
-as a single rule applied to all new cases, case-based systems provide
-case-specific interpretations in the form of a number of more fully
-understood reference cases. While the similarity measure used in the
-retrieval step need not be changed as the corpus grows, proposed
-measures have been diverse, contested, and rarely systematically
-compared.
+\emph{retrieves} one or more past cases (solved problems) from a corpus,
+\emph{reuses} these to generate an understanding of (solution to) the
+new case, \emph{revises} this understanding (solution) to better fit the
+new case (sometimes called ``adaptation''), and \emph{retains} the new
+case and its eventual understanding (solution) in the corpus to be
+retrieved and reused in future. In contrast to rule-based systems, which
+are variable-based and generally interpretable as a single rule applied
+to all new cases, case-based systems provide case-specific
+interpretations in the form of a number of more fully understood
+reference cases. While the similarity measure used in the retrieval step
+need not be changed as the corpus grows, proposed measures have been
+diverse, contested, and rarely systematically compared.
 
@@
 localized models have been primarily used for and evaluated on their
-ability to more accurately predict classes or outcomes. Nevertheless, as
-some more recent applications have begun to show, they can be used to
-draw inferences about the importance of different risk factors to
-specific individuals. \textbf{Distinguish between interpretable model
-components like effect estimates and informative but not directly
-interpretable importance measures as are commonly computed for
-`black-box' models.} While the patient similarity measure used to
+ability to predict classes or outcomes. Nevertheless, as some recent
+applications have begun to show, they can be used to draw inferences
+about the importance of different risk factors to specific individuals.
+While most ML models come equipped with measures of feature importance
+and model-agnostic tools can generate explanations for model
+predictions, these quantifications are not directly interpretable model
+components analogous to the split nodes of a decision tree (DT) or the
+effect estimates of a GLM. While the patient similarity measure used to
 retrieve each cohort may be complicated, the cohort itself can be
@@
 patient-level health data is used to construct local cohorts for
-predictive modeling.
+predictive or inferential modeling.
 
@@
 \item
-  Pulls case-level data from a corpus of past cases with known response
-  values\newline (response may be outcome, diagnosis, subtype, etc.)
+  Uses case-level data from a corpus of past cases with known
+  responses\newline (response may be outcome, diagnosis, subtype, etc.)
 \item
-  Defines a numeric multivariate case similarity measure\newline
-  (include integer-valued measures)
+  Defines a numeric multivariate case similarity measure\newline (allow
+  integer-valued measures)
 \item
-  Uses the similarity measure to select cohorts for index cases from the
-  corpus\newline (for example, based on a training--testing partition)
+  Uses the similarity measure to retrieve cohorts for index cases from
+  the corpus\newline (for example, based on a training--testing
+  partition)
 \item
-  Fits statistical models to cohorts from which to draw inferences about
-  index cases\newline (outcome predictions count, survival estimates,
-  risk factor effect estimates, model evaluation statistics, etc.)
+  Fits statistical models to cohorts from which to make predictions or
+  draw inferences about index cases\newline (outcome predictions,
+  survival estimates, risk factor contributions, model evaluation
+  statistics, etc.)
 \end{itemize}
 
-We expected the general framework to boil down to three choices: A
+We expected the general framework to come down to three choices: A
 patient similarity measure, a cohort selection process, and a
@@
 
-We describe the steps of our review in detail in the Appendix. We also
-note deviations from plans and the reasons for them.
+Here we describe our review process, including deviations from plans and
+the reasons for them. More details are included in Section
+\label{sec:appendix-selection}.
 
@@
 
-We include a PRISMA checklist as Supplemental Table 1 and a PRISMA
-abstract checklist as Supplemental Table 2. Because we focus on
-methodologies rather than approaches, we deviate in some ways from
-PRISMA guidelines. In particular, many items of the checklist, involving
-bias assessment and quantitative synthesis, are intended for
-meta-analyses and, therefore, did not apply to this study. Because we
-are not aware of any standard procedures for conducting reviews and
-syntheses of methodology, no protocol was prepared for this study.
+\textbf{Correct ``n = PubMed'' in PRISMA diagram.} We include a PRISMA
+checklist as Supplemental Table 1 and a PRISMA abstract checklist as
+Supplemental Table 2. Because we focus on methodologies rather than
+conceptual approaches or evidence, we deviate in some ways from PRISMA
+guidelines. In particular, those items of the checklist involving bias
+assessment and quantitative synthesis are intended for meta-analyses so
+did not apply to this study. Because we are not aware of any standard
+procedures for conducting reviews and syntheses of methodology, no
+protocol was prepared for this study.
 
@@
 Academic Search Premier, Google Scholar, and MathSciNet. We finalized
-the PubMed search first, then adapted it to the other platforms
-(Sections \ref{sec:appendix-search}).
+the PubMed search first, then adapted it to the other platforms (Section
+\ref{sec:appendix-search}).
 
@@
 decided on four eligibility criteria to expedite the screening process.
-These include articles written in English for readability, original
-studies to strictly include primary sources with new
-information/conclusions, medical or clinical related settings due to the
-applicability of our data to this field, and alternative use of the key
-terms within the search string, meaning that it was clear from the title
-and abstract that the entry was included due to the use of one or more
-search terms with a different meaning than ours. Each of two authors (AC
-and JCB) screened roughly half of the entries. They regularly reviewed
-each other's decisions to improve consistency. In cases of uncertainty,
-an entry would be included.
+Articles must be written in English, for readability; they must be
+original studies, to exclude secondary sources with duplicate
+information; their use settings must be medical, clinical, or related,
+to keep our review topical; and it must not be clear that their use of
+our search terms was different from our intended meaning. Each of two
+authors (AC and JCB) screened roughly half of the entries. They
+regularly reviewed each other's decisions to improve consistency. In
+cases of uncertainty, entries were included.
 
@@
 \subsection{Full text review}\label{full-text-review}}
+
+\label{sec:full-text}
 
@@
 \item
-  Fits statistical models to cohorts to make inferences about index
-  cases
+  Fits statistical models to cohorts to make predictions or inferences
+  about index cases
 \end{itemize}
@@
 In Covidence, two authors (AC and JCB) independently evaluated each
-manuscript for the four eligibility criteria. The first criterion that a
+manuscript for these eligibility criteria. The first criterion that a
 manuscript failed was designated the reason for exclusion. An antecedent
@@
 (continuous response) or the plurality (discrete response) of the
-responses of the cohort. A review of all studies that use nearest
-neighbors prediction, even in the clinical domain, would be impractical.
-Because our focus is on novel approaches that combine similarity-based
-retrieval and statistical modeling of retrieved cohorts, we chose to
-exclude those studies whose approach was equivalent to nearest neighbors
-prediction using a conventional similarity measure.
+cohort's responses. A review of all studies that use nearest neighbors
+prediction would be impractical. Because our focus is on novel
+approaches that combine similarity-based retrieval and statistical
+modeling of retrieved cohorts, we chose to exclude those studies whose
+approach was equivalent to nearest neighbors prediction using a
+conventional similarity measure.
 
@@
 measure, families of adaptation step/statistical model, method(s)
-compared against, evaluation/performance/comparison measures (\& their
-values), name given to modeling approach). We used these data to detect
-and visualize trends amongst the included studies.
+compared against, performance measures, results of evaluations and
+comparisons, name given to modeling approach). We used these data to
+detect and visualize trends amongst the included studies.
 
@@
 Rather than an evidence synthesis characteristic of most systematic
-reviews, we here pursue a methodology synthesis to harmonize possibly
+reviews, we here pursue a methodology synthesis to harmonize largely
 independent research efforts that have converged on a common modeling
@@
 left with 6 additional studies {[}12--17{]}, 1 of which was excluded
-from the synthesis for using nearest neighbors prediction. Finally,
-citations in reference lists from the 43 + 328 = 371 studies assessed
-for eligibility led us to identify 1 additional study that met our
-criteria {[}18{]}. This left us with 19 + 1 + 5 = 25 studies included in
-the review and synthesis.
+from the synthesis for using NN prediction. Finally, reference tracking
+from the 43 + 328 = 371 studies assessed for eligibility led us to
+identify 1 additional study that met our criteria {[}18{]}. This left us
+with 19 + 1 + 5 = 25 studies included in the review and synthesis.
 
@@
 properties}\label{bibliographic-and-methodological-properties}}
+
+\textbf{Update numbers below, perhaps dynamically.}
 
@@
 
-Another interesting pattern lay in the years of publication of the
-included studies (Figure \ref{fig:year}). While they stretch back to the
-late 1990s, they have become more common, which suggests that this is
-still an active, though not rapidly expanding, methodological approach.
-We note that, while CBR in health and medicine originated as early as
-1990, these studies did emerge early on once CBR had established itself.
-So the idea has been ``in the air'' for as long as CBR has been in use.
+Another interesting pattern lay in the years of publication (Figure
+\ref{fig:year}). While they trace back to the late 1990s, they have
+become more common, which suggests that this is an active, though not
+rapidly expanding, methodological approach. We note that, while CBR in
+health and medicine originated as early as 1990, these studies did
+emerge early on once CBR had established itself. So the idea has been
+``in the air'' for as long as CBR has been in use.
 
@@
 majority of studies from a ``practice'' standpoint. These studies aimed
-primarily to provide providers with tools to use in clinic or to improve
-outcomes for patients currently receiving care.
+to provide tools for use in clinic or to improve outcomes.
 
@@
 the individual patient rather than the population, without reference to
-model performance {[}21,23{]}.
+performance {[}21,23{]}.
 
@@
 based. This ``self-explanation capability'' made outputs more
-intelligible to physicians in the CDS setting. {[}15{]} additional
+intelligible to physicians in the CDS setting. {[}15{]} additionally
 pointed out that localized GLMs yield localized effect estimates, which
@@
 particular by only executing task-dependent steps in real time
-(``just-in-time learning''), and {[}35{]} and {[}34{]} partitioned
+(``just-in-time learning'', JITL), and {[}35{]} and {[}34{]} partitioned
 multiple phases in the R4 cycle into offline and online components, only
@@
 teams also set out to develop more interoperable clinical support tools
-{[}15,22,26,30{]}.
-
-The frequency plot in Figure \ref{fig:methods} compares the rates at
-which several methodological elements are invoked in the sample.
+{[}15,22,26,30{]}. The frequency plot in Figure \ref{fig:methods}
+compares the rates at which several methodological elements are invoked
+in the sample.
 
@@
 \end{figure}
-
-In the Discussion, we discuss trends in the literature we reviewed,
-identify common limitations and gaps in understanding, and attempt to
-synthesize their methods into a general framework.
 
@@
 
-However, we do observe one clear pattern: Of the proposed methods that
-most evidently outperformed their comparators (Kohonen + CBR {[}20{]},
-TWNFI {[}18{]}, GSA {[}26{]}, CBR + rules (+ DML) {[}28{]}, Gaussian
-process regression {[}30{]}, JITL-ELM {[}32{]}), all except the last are
-hybrids of localized modeling (in some cases CBR) with other techniques,
-often metric learning. Though {[}27{]} and {[}15{]} report exceptions,
-the pattern suggests the importance of the similarity measure to the
-retrieval step. Meanwhile, when proposed approaches targeted cohort
-demarcation or choice of predictive model (statistical CBR {[}13{]};
-individualized logistic regression, decision tree, and random forest
-{[}14,16{]}), they did not consistently outperform comparators.
+We do observe one clear pattern: All of the proposed methods that most
+evidently outperformed their comparators (Kohonen + CBR {[}20{]}, TWNFI
+{[}18{]}, gravitational search algorithms (GSA) {[}26{]}, CBR + rules (+
+DML) {[}28{]}, Gaussian process regression {[}30{]}, JITL-ELM {[}32{]})
+are hybrids of localized modeling (in some cases CBR) with other
+techniques, often metric learning. Though {[}27{]} and {[}15{]} report
+non-superior performance by such hybrids, the pattern suggests the
+importance of the similarity measure to the retrieval step. Meanwhile,
+when proposed approaches targeted cohort demarcation or choice of
+predictive model (statistical CBR {[}13{]}; individualized logistic
+regression, decision tree, and random forest {[}14,16{]}), they did not
+consistently outperform comparators.
 
@@
 {[}11,19{]}. Urged improvements included full or partial automation of
-user-dependent steps such as predictor selection {[}11,19{]}, similarity
-learning {[}11,17{]}, and parameter optimization {[}16,21{]}; extensions
-to new data structures {[}24{]}, data types {[}25,26,31{]}, and
-reasoning systems {[}28{]}; and the use of more advanced model
-components to improve accuracy or efficiency {[}12,14,26,30{]}. Authors
-also urged validations and independent evaluations using larger or more
-comprehensive data sets {[}15,22,23,25{]}, using data aggregated from
-multiple health systems {[}14,16,34,35{]}, and in other care settings or
-disease contexts {[}21,30,34,35{]}. Less common were calls to strengthen
-the connection between the methods and the users. Some authors urged the
-incorporation of (human-derived) domain knowledge into the data or
-models {[}17,19{]}. Others suggested new uses of their modeling
-approaches: to measure feature importance {[}20{]}, to make models more
-expressive {[}14{]}, and to combine information obtained both from
-global and from localized models {[}15{]}. Most of this incremental work
-was indeed carried out in later investigations in our sample.
-
-\hypertarget{public-availability}{%
-\subsection{Public availability}\label{public-availability}}
-
-\label{sec:availability}
+predictor selection {[}11,19{]}, similarity learning {[}11,17{]}, and
+parameter optimization {[}16,21{]}; extensions to new data structures
+{[}24{]}, data types {[}25,26,31{]}, and reasoning systems {[}28{]}; and
+the use of more advanced model components to improve accuracy or
+efficiency {[}12,14,26,30{]}. Authors also urged validations and
+independent evaluations using larger or more comprehensive data sets
+{[}15,22,23,25{]}, using data aggregated from multiple health systems
+{[}14,16,34,35{]}, and in other care settings or disease contexts
+{[}21,30,34,35{]}.
+
+Less common were calls to strengthen the connection between the methods
+and the users. Some authors urged the incorporation of (human-derived)
+domain knowledge into the data or models {[}17,19{]}. Others suggested
+new uses of their modeling approaches: to measure feature importance
+{[}20{]}, to make models more expressive {[}14{]}, and to combine
+information obtained both from global and from localized models
+{[}15{]}. Most of this incremental work was indeed carried out in later
+investigations in our sample.
+
+\hypertarget{reproducibility}{%
+\subsection{Reproducibility}\label{reproducibility}}
+
+\label{sec:reproducibility}
 
@@
 future work, though it might be limited by the lack of open-source code
-or public implementations (Section \ref{sec:availability}).
+or public implementations (Section \ref{sec:reproducibility}).
 
@@
 
-\textbf{Write this section!}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Deduplication (Covidence)
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Specific to that source: title, creator, URL, DOI
-  \item
-    May overlap: journal, publication date, ISSN
-  \end{enumerate}
-\item
-  Title
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Exclude non clinical/non medical setting
-  \item
-    Must be in English
-  \item
-    Must be original study (not reviews, surveys, opinion, news)
-  \item
-    Exclude if search term clearly has different meaning than intended
-  \end{enumerate}
-\item
-  Abstract
-\item
-  Full Text
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii}.}
-  \tightlist
-  \item
-    Study criteria (inclusion/exclusion)
-  \end{enumerate}
-\end{enumerate}
+Most deduplication was done automatically in Covidence. As full-text
+review was done in Zotero, some additional duplicates were noticed and
+merged.
+
+When abstracts were not obtained by search or by Covidence, we attempted
+to find them online using DOIs; when an abstract could not be found,
+screening was based on the title alone. We decided to screen titles and
+abstracts conservatively, rejecting only studies that were clearly
+outside the scope of our review. The reasons for rejection were four, as
+discussed in the main text: a. Exclude non clinical/non medical setting
+b. Must be in English c.~Must be original study (not reviews, surveys,
+opinion, news) d.~Exclude if search term clearly has different meaning
+than intended (Two examples of (d) are the use of the term
+``personalized model'' to refer in some cases to parameterized models
+tuned to individual patient measurements and in others to
+patient-centered models of care.) Studies that passed title/abstract
+screen were exported to Zotero.
 
@@
 
-Our selection of relevant papers from the search corpus will be based on
-the following inclusion/exclusion criteria:
+Some PDFs were obtained using Zotero from a university workstation, and
+the remaining were obtained through university library services. The
+review process is detailed in Section \label{sec:full-text}.
+
+Our selection of relevant papers from the search corpus was based on the
+following inclusion/exclusion criteria:
 
@@
 \begin{landscape}
+
+\textbf{TODO: Place vertical separators between major columns.}
 

### 2023-09-09 docs/Manuscript/Manuscript.Rmd
transcribe revisions to discussion + format performance table + update figures

@@
 Here we describe our review process, including deviations from plans and the reasons for them.
-More details are included in Section \label{sec:appendix-selection}.
+More details are included in Section \ref{sec:appendix-selection}.
 
@@
 
-**Correct "n = PubMed" in PRISMA diagram.**
 We include a PRISMA checklist as Supplemental Table 1 and a PRISMA abstract checklist as Supplemental Table 2.
@@
 
-**Update numbers below, perhaps dynamically.**
-
-The 25 included studies were analyzed based on 20 characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The `r nrow(tab_synthesis)` included studies were analyzed based on 20 characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
-The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, and _Hindawi - Journal of Healthcare Engineering_ published 3.
+The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, and _PLOS One_ published 2.
 
@@
 
-Some PDFs were obtained using Zotero from a university workstation, and the remaining were obtained through university library services.
-The review process is detailed in Section \label{sec:full-text}.
+Some PDFs were obtained using Zotero from a university workstation, and the remaining were obtained through university library services. The review process is detailed in Section \ref{sec:full-text}.
 
@@
   ) |> 
+  mutate(Measure = ifelse(
+    is.na(lag(Measure)) | Measure != lag(Measure) | Citation != lag(Citation),
+    Measure, ""
+  )) |> 
   mutate(across(
-    c(Citation, Measure, Problem),
+    c(Citation, Problem),
     ~ ifelse(is.na(lag(.)) | . != lag(.), ., "")

### 2023-09-09 docs/Manuscript/Manuscript.tex
transcribe revisions to discussion + format performance table + update figures

@@
 We discuss in this review a hybrid computational approach we call
-\emph{localized modeling} that appears to have been introduced
-independently several times. The approach fits ML models, including
-regularized families of generalized regression models (GLMs), to cohorts
-retrieved via similarity matching. Whereas CBR and NN prediction use
-patient similarity to obtain an ideally homogeneous subpopulation
-matched to the index case, in localized modeling one seeks a relevant
-yet heterogeneous subpopulation, each member of which may deviate from
-the index case. A prediction or inference is then obtained from a model
-fitted to this \emph{similarity cohort}. The use of similarity cohorts
-distinguishes localized modeling from classical GLMs that still dominate
-analysis in clinical trials and comparative effectiveness research, as
-well as from many cutting-edge ML models that are fitted to whole
-populations, while the cohorts' heterogeneity and its utility to the
-model distinguish the approach from CBR. In principle, this allows
-interpretable models like GLMs to identify and measure the effects of
-determining factors that are specific to the index case. In the reviewed
-studies, localized modeling shows some promise toward simultaneously
-achieving the performance improvements of ML and the interpretability of
-classical models.
+\emph{localized modeling} (see Section \ref{sec:coherence}) that appears
+to have been introduced independently several times. The approach fits
+ML models, including regularized families of generalized regression
+models (GLMs), to cohorts retrieved via similarity matching. Whereas CBR
+and NN prediction use patient similarity to obtain an ideally
+homogeneous subpopulation matched to the index case, in localized
+modeling one seeks a relevant yet heterogeneous subpopulation, each
+member of which may deviate from the index case. A prediction or
+inference is then obtained from a model fitted to this \emph{similarity
+cohort}. The use of similarity cohorts distinguishes localized modeling
+from classical GLMs that still dominate analysis in clinical trials and
+comparative effectiveness research, as well as from many cutting-edge ML
+models that are fitted to whole populations, while the cohorts'
+heterogeneity and its utility to the model distinguish the approach from
+CBR. In principle, this allows interpretable models like GLMs to
+identify and measure the effects of determining factors that are
+specific to the index case. In the reviewed studies, localized modeling
+shows some promise toward simultaneously achieving the performance
+improvements of ML and the interpretability of classical models.
 
@@
 the reasons for them. More details are included in Section
-\label{sec:appendix-selection}.
+\ref{sec:appendix-selection}.
 
@@
 
-\textbf{Correct ``n = PubMed'' in PRISMA diagram.} We include a PRISMA
-checklist as Supplemental Table 1 and a PRISMA abstract checklist as
-Supplemental Table 2. Because we focus on methodologies rather than
-conceptual approaches or evidence, we deviate in some ways from PRISMA
-guidelines. In particular, those items of the checklist involving bias
-assessment and quantitative synthesis are intended for meta-analyses so
-did not apply to this study. Because we are not aware of any standard
-procedures for conducting reviews and syntheses of methodology, no
-protocol was prepared for this study.
+We include a PRISMA checklist as Supplemental Table 1 and a PRISMA
+abstract checklist as Supplemental Table 2. Because we focus on
+methodologies rather than conceptual approaches or evidence, we deviate
+in some ways from PRISMA guidelines. In particular, those items of the
+checklist involving bias assessment and quantitative synthesis are
+intended for meta-analyses so did not apply to this study. Because we
+are not aware of any standard procedures for conducting reviews and
+syntheses of methodology, no protocol was prepared for this study.
 
@@
 
-\textbf{Update numbers below, perhaps dynamically.}
-
 The 25 included studies were analyzed based on 20 characteristics (see
@@
 published 2 of the included studies, \emph{Artificial Intelligence in
-Medicine} published 3, and \emph{Hindawi - Journal of Healthcare
-Engineering} published 3.
+Medicine} published 3, \emph{Hindawi - Journal of Healthcare
+Engineering} published 3, and \emph{PLOS One} published 2.
 
@@
 are hybrids of localized modeling (in some cases CBR) with other
-techniques, often metric learning. Though {[}27{]} and {[}15{]} report
-non-superior performance by such hybrids, the pattern suggests the
-importance of the similarity measure to the retrieval step. Meanwhile,
-when proposed approaches targeted cohort demarcation or choice of
-predictive model (statistical CBR {[}13{]}; individualized logistic
-regression, decision tree, and random forest {[}14,16{]}), they did not
-consistently outperform comparators.
+techniques, often DML. Though {[}27{]} and {[}15{]} report non-superior
+performance by such hybrids, the pattern suggests the importance of the
+similarity measure to the retrieval step. Meanwhile, when proposed
+approaches targeted cohort demarcation or choice of predictive model
+(statistical CBR {[}13{]}; individualized logistic regression, decision
+tree, and random forest {[}14,16{]}), they did not consistently
+outperform comparators.
 
@@
 \textbf{Discuss public availability of code or implementations.}
-\textbf{Methods, results, and inferential reproducibility
-{[}\textbf{Matarese2022?}{]}.}
+\textbf{Methods, results, and inferential reproducibility {[}39{]}.}
 
@@
 
-We draw several observations from our encoding of the included studies.
-We first focus on two driving questions: What challenges did
-investigators set out to solve? How did their approaches meet these
-challenges? We then comment on the connectedness of the included studies
-and the potential for recombination.
+After commenting briefly on our review process, we draw several
+observations from our encoding of the included studies. We first focus
+on their methodological choices and innovations, then on their
+conventions and language. We spend most of the section discussing the
+differences and commonalities of the techniques used, toward a general
+method that encompasses most or all cases. We then suggest some avenues
+for future studies and conclude with our key takeaways.
 
@@
 
+\label{sec:coherence}
+
 The motivational and methodological unity of these studies does not
@@
   some cases than others, as quantified by importance measures or
-  model-agnostic explanations {[}39,40{]}. In this way, such models can
+  model-agnostic explanations {[}40,41{]}. In this way, such models can
   also be viewed as ``localized''. We recognize that the use of global
@@
   stored in a single storage device (local) from their aggregations
-  (global) {[}41,42{]}. The prospects for federated architecture to more
+  (global) {[}42,43{]}. The prospects for federated architecture to more
   effectively implement similarity cohort--based models are intriguing,
@@
 The 25 studies included in our synthesis used composite techniques that
-we organized into two major types. One type, similarity learning, was
-used in 8 studies that tie in to the much larger literature on metric
-learning {[}43{]}. These techniques are designed to detect structure in
-data, especially associations between predictors and responses, and to
-use this structure to inform the definition of a patient similarity
-measure. Seven studies used response values in training data sets to
-supervise similarity learning while two used unsupervised learning
-(Table \ref{tab:composite}). Based on the key properties of metric
-learning algorithms identified by {[}44{]}, most learned measures were
-non-linear while some were local, and some were optimized globally while
-others locally. However, not all qualified as metric learning, since the
-resulting measures would not necessarily satisfy the triangle
-inequality. Each then used its similarity measure to retrieve training
-cases relevant to each testing case. Once retrieved, most took a
-standard NN approach to generating predictions from the cohort; two
-exceptions {[}29,34{]} proceeded to fit predictive models to the
-retrieved cohorts and are discussed below.
+we organized into two major types. One type, \emph{similarity learning},
+was used in 8 studies that tie in to the much larger literature on DML
+{[}44{]}. The other, \emph{cohort thresholding}, focused on choosing or
+optimizing the manner in which cohorts were demarcated using the
+similarity measure.
 
@@
 
-The similarity learning approaches took a variety of forms. Several
-studies used unsupervised similarity learning, for example the
+The similarity learning approaches took a variety of forms. These
+techniques are designed to detect structure in data, especially
+associations between predictors and responses, and to use this structure
+to inform the definition of a patient similarity measure. Seven studies
+used response values in training data sets to supervise similarity
+learning while two used unsupervised learning (Table
+\ref{tab:composite}). Based on the key properties of DML algorithms
+identified by {[}45{]}, most learned measures were non-linear while some
+were locally learned, and some were optimized globally while others
+locally. Not all qualified as DML, since the resulting measures would
+not necessarily satisfy the triangle inequality. Each used its
+similarity measure to retrieve training cases relevant to each testing
+case. Once retrieved, most took a standard NN approach to generating
+predictions; two exceptions {[}29,34{]} fit predictive models to the
+retrieved cohorts.
+
+Several studies used unsupervised similarity learning, for example the
 Mahalanobis distance {[}12{]} and the Kolmogorov entropy-based distance
@@
 here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning (DML) technique {[}45{]}
+appeal to a supervised distance metric learning (DML) technique {[}46{]}
 to obtain a similarity measure on their set of dermoscopy and confocal
@@
 conventional CBR they called ``statistical CBR''.\footnote{Drawing from
-  {[}46{]}, we suggest the term ``combinatorial CBR'' for the
+  {[}47{]}, we suggest the term ``combinatorial CBR'' for the
   then-conventional approach using similarity cohorts of fixed
@@
 As noted above, these instances of similarity learning fit into a much
-larger literature on metric learning, which has seen widespread use in
-health informatics and modeling. That these studies satisfied our review
+larger literature on DML, which has seen widespread use in health
+informatics and modeling. That these studies satisfied our review
 criteria is due to the novelty of the techniques at the time of
@@
 statistical insight for the index case from the similarity cohort. The
-choices are the \emph{similarity measure}, the \emph{retrieval step},
+choices are the \emph{similarity measure}, the \emph{cohort retrieval},
 and the \emph{statistical model}, respectively. These are diagrammed in
@@
 similarity, though most uniform cohort sizes were tuned. Some earlier
-studies tuned generalized regression models, but more recent studies
-have not.
+studies tuned generalized regression models, but more recent studies did
+not.
 
@@
 the remaining were obtained through university library services. The
-review process is detailed in Section \label{sec:full-text}.
+review process is detailed in Section \ref{sec:full-text}.
 
@@
 & & & & \hspace{6em} & backpropagation & 54.70\%\hspace{6em} \\
-{[}13{]} & & Dermatology & Statistical CBR & 96.57\%\hspace{6em} & CBR &
-95.43\%\hspace{6em} \\
+{[}13{]} & Accuracy & Dermatology & Statistical CBR &
+96.57\%\hspace{6em} & CBR & 95.43\%\hspace{6em} \\
 & & & & \hspace{6em} & C5.0 & 93.71\%\hspace{6em} \\
@@
 71.2\%\hspace{6em} \\
-{[}18{]} & & Colon cancer diagnosis & TWNFI & 91.9\%\hspace{6em} & MLR
-(Personalized) & 82.3\%\hspace{6em} \\
+{[}18{]} & Accuracy & Colon cancer diagnosis & TWNFI &
+91.9\%\hspace{6em} & MLR (Personalized) & 82.3\%\hspace{6em} \\
 & & & & \hspace{6em} & SVM (Personalized) & 90.3\%\hspace{6em} \\
@@
 \hspace{6em} \\
-{[}26{]} & & Colon cancer & knnGSA & 85.48\%\hspace{6em} & SVM &
-82.14\%\hspace{6em} \\
+{[}26{]} & Accuracy & Colon cancer & knnGSA & 85.48\%\hspace{6em} & SVM
+& 82.14\%\hspace{6em} \\
 & & & svmGSA & 87.10\%\hspace{6em} & ECF & 72.30\%\hspace{6em} \\
@@
 & & & Personalized LR (Random) & 0.602\hspace{6em} & & \hspace{6em} \\
-{[}14{]} & & 30-day mortality & Individualized LR & 0.830\hspace{6em} &
-Individualized DC & 0.797\hspace{6em} \\
+{[}14{]} & AUROC & 30-day mortality & Individualized LR &
+0.830\hspace{6em} & Individualized DC & 0.797\hspace{6em} \\
 & & & Individualized DT & 0.753\hspace{6em} & & \hspace{6em} \\
-{[}16{]} & & & Individualized LR & 0.824\hspace{6em} & Individualized DC
-& 0.801\hspace{6em} \\
+{[}16{]} & AUROC & & Individualized LR & 0.824\hspace{6em} &
+Individualized DC & 0.801\hspace{6em} \\
 & & & Individualized DT & 0.779\hspace{6em} & & \hspace{6em} \\
@@
 0.778\hspace{6em} \\
-{[}17{]} & & Diabetes & Personalized RF & 0.90\hspace{6em} & &
+{[}17{]} & AUROC & Diabetes & Personalized RF & 0.90\hspace{6em} & &
 \hspace{6em} \\
@@
 & & & Personalized LR & 0.89\hspace{6em} & & \hspace{6em} \\
-{[}32{]} & & Discharge at 10 days & One-class JITL-ELM &
+{[}32{]} & AUROC & Discharge at 10 days & One-class JITL-ELM &
 0.8510\hspace{6em} & ELM & 0.4973\hspace{6em} \\
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
+\CSLLeftMargin{{[}39{]} }%
+\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
+and {Different Functions}. Axiomathes 2022;32:647--70.
+\url{https://doi.org/10.1007/s10516-021-09610-2}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
-\CSLLeftMargin{{[}39{]} }%
+\CSLLeftMargin{{[}40{]} }%
 \CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
-\CSLLeftMargin{{[}40{]} }%
+\CSLLeftMargin{{[}41{]} }%
 \CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
-\CSLLeftMargin{{[}41{]} }%
+\CSLLeftMargin{{[}42{]} }%
 \CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
-\CSLLeftMargin{{[}42{]} }%
+\CSLLeftMargin{{[}43{]} }%
 \CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-\CSLLeftMargin{{[}43{]} }%
+\CSLLeftMargin{{[}44{]} }%
 \CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-\CSLLeftMargin{{[}44{]} }%
+\CSLLeftMargin{{[}45{]} }%
 \CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-\CSLLeftMargin{{[}45{]} }%
+\CSLLeftMargin{{[}46{]} }%
 \CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-\CSLLeftMargin{{[}46{]} }%
+\CSLLeftMargin{{[}47{]} }%
 \CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A

### 2023-09-10 docs/Manuscript/Manuscript.Rmd
make final rmarkdown revisions

@@
 
-The `r nrow(tab_synthesis)` included studies were analyzed based on 20 characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
@@
 Most of this incremental work was indeed carried out in later investigations in our sample.
-
-## Reproducibility
-\label{sec:reproducibility}
-
-**Discuss public availability of code or implementations.**
-**Methods, results, and inferential reproducibility [@Matarese2022].**
+That said, we did not find reports of successful implementations of these tools in clinical practice. (Several excluded studies did showcase uses of CBR in practice.)
 
@@
 We also note that the majority of studies evaluated and compared methods using leave-one-out cross-validation (LOOCV).
-This is an especially appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer.
+This is an appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer.
 This likely follows from the older age of many included studies and from their consistent primary focus on clinical data, which is more costly to collect and comes with more restrictions on its use.
@@
 However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
-This would be a worthy goal of future work, though it might be limited by the lack of open-source code or public implementations (Section \ref{sec:reproducibility}).
+This would be a worthy goal of future work.
 
 While most studies identified outstanding technical needs (Section \ref{sec:needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
-Most studies also identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
+Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
 Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
-One valuable direction for future work would be to compare the ability of non-technical investigators or users to draw correct inferences from these and other interpretable methods and their levels of confidence in those inferences.
+One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
 
-We posit that an essential component of such "soft" performance goals is the ability of the user community to exert some control over the models.
+The primary goal of these studies was to establish that localized models, and certain strategies within this paradigm, perform at or above the level of other predictive models or strategies.
+From a functionalist approach to reproducibility [@Matarese2022], then, later studies were broadly successful at reproducing earlier studies, and later studies provide sufficient detail to support ongoing reproduction efforts---though these might be limited by the lack of open-source code or public implementations.
+However, the immediate goal of most studes was to improve care, and an implied need, often enjoined but neither obtained nor reproduced, was a demonstration of practical use.
+
+We posit that an essential component of practical usefulness is the ability of the user community to exert some control over the models.
 Given the impact on performance of the choice or optimization of the similarity measure, an important target for user input would be the importance of certain variables in the calculation of similarity, with an understanding of how it can impact not only the performance of the predictions but also the cohort retrieved for the model.

### 2023-09-10 docs/Manuscript/Manuscript.tex
make final rmarkdown revisions

@@
 
-The 25 included studies were analyzed based on 20 characteristics (see
-Section \ref{sec:data}), and we report and describe some observations
-here (Table \ref{tab:synthesis}). The studies were published in a
-variety of journals, with some of greater frequency, though no single
-journal published more than 3. The journal \emph{Evolving Systems}
-published 2 of the included studies, \emph{Artificial Intelligence in
-Medicine} published 3, \emph{Hindawi - Journal of Healthcare
-Engineering} published 3, and \emph{PLOS One} published 2.
+The 25 included studies were analyzed based on several characteristics
+(see Section \ref{sec:data}), and we report and describe some
+observations here (Table \ref{tab:synthesis}). The studies were
+published in a variety of journals, with some of greater frequency,
+though no single journal published more than 3. The journal
+\emph{Evolving Systems} published 2 of the included studies,
+\emph{Artificial Intelligence in Medicine} published 3, \emph{Hindawi -
+Journal of Healthcare Engineering} published 3, and \emph{PLOS One}
+published 2.
 
@@
 {[}15{]}. Most of this incremental work was indeed carried out in later
-investigations in our sample.
-
-\hypertarget{reproducibility}{%
-\subsection{Reproducibility}\label{reproducibility}}
-
-\label{sec:reproducibility}
-
-\textbf{Discuss public availability of code or implementations.}
-\textbf{Methods, results, and inferential reproducibility {[}39{]}.}
+investigations in our sample. That said, we did not find reports of
+successful implementations of these tools in clinical practice. (Several
+excluded studies did showcase uses of CBR in practice.)
 
@@
 that the majority of studies evaluated and compared methods using
-leave-one-out cross-validation (LOOCV). This is an especially
-appropriate technique when available data are scarce, and indeed most
-data sets used by included studies numbered in the hundreds of cases or
-fewer. This likely follows from the older age of many included studies
-and from their consistent primary focus on clinical data, which is more
-costly to collect and comes with more restrictions on its use.
+leave-one-out cross-validation (LOOCV). This is an appropriate technique
+when available data are scarce, and indeed most data sets used by
+included studies numbered in the hundreds of cases or fewer. This likely
+follows from the older age of many included studies and from their
+consistent primary focus on clinical data, which is more costly to
+collect and comes with more restrictions on its use.
 
@@
   some cases than others, as quantified by importance measures or
-  model-agnostic explanations {[}40,41{]}. In this way, such models can
+  model-agnostic explanations {[}39,40{]}. In this way, such models can
   also be viewed as ``localized''. We recognize that the use of global
@@
   stored in a single storage device (local) from their aggregations
-  (global) {[}42,43{]}. The prospects for federated architecture to more
+  (global) {[}41,42{]}. The prospects for federated architecture to more
   effectively implement similarity cohort--based models are intriguing,
@@
 was used in 8 studies that tie in to the much larger literature on DML
-{[}44{]}. The other, \emph{cohort thresholding}, focused on choosing or
+{[}43{]}. The other, \emph{cohort thresholding}, focused on choosing or
 optimizing the manner in which cohorts were demarcated using the
@@
 \ref{tab:composite}). Based on the key properties of DML algorithms
-identified by {[}45{]}, most learned measures were non-linear while some
+identified by {[}44{]}, most learned measures were non-linear while some
 were locally learned, and some were optimized globally while others
@@
 here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning (DML) technique {[}46{]}
+appeal to a supervised distance metric learning (DML) technique {[}45{]}
 to obtain a similarity measure on their set of dermoscopy and confocal
@@
 conventional CBR they called ``statistical CBR''.\footnote{Drawing from
-  {[}47{]}, we suggest the term ``combinatorial CBR'' for the
+  {[}46{]}, we suggest the term ``combinatorial CBR'' for the
   then-conventional approach using similarity cohorts of fixed
@@
 trade-offs with respect to each other. This would be a worthy goal of
-future work, though it might be limited by the lack of open-source code
-or public implementations (Section \ref{sec:reproducibility}).
+future work.
 
@@
 qualities like user interface and user experience, interpretability,
-meaningfulness, or trust. Most studies also identified interpretability
-as an advantage of localized models, though none evaluated model
+meaningfulness, or trust. Most studies identified interpretability as an
+advantage of localized models, though none evaluated model
 interpretability and few proposed new interpretative uses. Lack of trust
@@
 than demonstrated. One valuable direction for future work would be to
-compare the ability of non-technical investigators or users to draw
-correct inferences from these and other interpretable methods and their
-levels of confidence in those inferences.
-
-We posit that an essential component of such ``soft'' performance goals
-is the ability of the user community to exert some control over the
-models. Given the impact on performance of the choice or optimization of
-the similarity measure, an important target for user input would be the
+measure the utility of interpretable components in research and in
+clinical practice and the correctness and confidence of users in their
+interpretations.
+
+The primary goal of these studies was to establish that localized
+models, and certain strategies within this paradigm, perform at or above
+the level of other predictive models or strategies. From a functionalist
+approach to reproducibility {[}47{]}, then, later studies were broadly
+successful at reproducing earlier studies, and later studies provide
+sufficient detail to support ongoing reproduction efforts---though these
+might be limited by the lack of open-source code or public
+implementations. However, the immediate goal of most studes was to
+improve care, and an implied need, often enjoined but neither obtained
+nor reproduced, was a demonstration of practical use.
+
+We posit that an essential component of practical usefulness is the
+ability of the user community to exert some control over the models.
+Given the impact on performance of the choice or optimization of the
+similarity measure, an important target for user input would be the
 importance of certain variables in the calculation of similarity, with
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
 \CSLLeftMargin{{[}39{]} }%
-\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
-and {Different Functions}. Axiomathes 2022;32:647--70.
-\url{https://doi.org/10.1007/s10516-021-09610-2}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
-\CSLLeftMargin{{[}40{]} }%
 \CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
-\CSLLeftMargin{{[}41{]} }%
+\CSLLeftMargin{{[}40{]} }%
 \CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
-\CSLLeftMargin{{[}42{]} }%
+\CSLLeftMargin{{[}41{]} }%
 \CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
-\CSLLeftMargin{{[}43{]} }%
+\CSLLeftMargin{{[}42{]} }%
 \CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-\CSLLeftMargin{{[}44{]} }%
+\CSLLeftMargin{{[}43{]} }%
 \CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-\CSLLeftMargin{{[}45{]} }%
+\CSLLeftMargin{{[}44{]} }%
 \CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-\CSLLeftMargin{{[}46{]} }%
+\CSLLeftMargin{{[}45{]} }%
 \CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-\CSLLeftMargin{{[}47{]} }%
+\CSLLeftMargin{{[}46{]} }%
 \CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
+\CSLLeftMargin{{[}47{]} }%
+\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
+and {Different Functions}. Axiomathes 2022;32:647--70.
+\url{https://doi.org/10.1007/s10516-021-09610-2}.}
+
 \end{CSLReferences}

### 2023-09-10 docs/Manuscript/Manuscript.Rmd
make additional edits during resubmission process

@@
 
-We discuss in this review a hybrid computational approach we call _localized modeling_ (see Section \ref{sec:coherence}) that appears to have been introduced independently several times.
+We discuss in this review a hybrid computational approach we call _localized modeling_ (see Section \ref{coherence}) that appears to have been introduced independently several times.
 The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching.
@@
 Here we describe our review process, including deviations from plans and the reasons for them.
-More details are included in Section \ref{sec:appendix-selection}.
+More details are included in Section \ref{selection-process}.
 
@@
 Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
-We finalized the PubMed search first, then adapted it to the other platforms (Section \ref{sec:appendix-search}).
+We finalized the PubMed search first, then adapted it to the other platforms (Section \ref{search-design}).
 
@@
 ## Full text review
-\label{sec:full-text}
 
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{sec:data}).
+Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{data-and-code}).
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, performance measures, results of evaluations and comparisons, name given to modeling approach).
@@
 
-```{r fig:prisma, fig.align='center', out.width="80%", echo=FALSE}
+```{r prisma, fig.align='center', out.width="80%", echo=FALSE}
 #| fig.cap =
-#| "\\label{fig:prisma}
-#| PRISMA-S flow chart.
+#| "PRISMA-S flow chart.
 #|  Lowercase letters refer to items obtained from
@@
 
-The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics (see Section \ref{sec:data}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics (see Section \ref{data-and-code}), and we report and describe some observations here (Table \ref{tab:synthesis}).
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
@@
 
-```{r fig:year, fig.align='center', out.width="75%", echo=FALSE}
+```{r year, fig.align='center', out.width="75%", echo=FALSE}
 #| fig.cap =
-#| "\\label{fig:year}
-#| Number of publications in our sample each year."
+#| "Number of publications in our sample each year."
 knitr::include_graphics("../../fig/fig-years.png")
@@
 
-```{r fig:properties, fig.align='center', out.width="100%", echo=FALSE}
+```{r properties, fig.align='center', out.width="100%", echo=FALSE}
 #| fig.cap =
-#| "\\label{fig:properties}
-#| Share of studies characterized by several design elements:
+#| "Share of studies characterized by several design elements:
 #|  data source (row),
@@
 ## Rationales
-\label{sec:rationales}
+\label{rationales}
 
-The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{sec:data}).
+The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{data-and-code}).
 The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015a; @Ng2015; @Lee2017].
@@
 
-```{r fig:methods, fig.align='center', out.width="100%", echo=FALSE}
+```{r methods, fig.align='center', out.width="100%", echo=FALSE}
 #| fig.cap =
-#| "\\label{fig:methods}
-#| Frequency of recurring methodological elements across included studies."
+#| "Frequency of recurring methodological elements across included studies."
 knitr::include_graphics("../../fig/fig-methods.png")
@@
 ## Performance assessments
-\label{sec:performance}
+\label{performance-assessments}
 
@@
 ## Identified needs
-\label{sec:needs}
 
@@
 ## Coherence
-\label{sec:coherence}
 
@@
 
-```{r fig:framework, fig.align='center', out.width="80%", echo=FALSE}
+```{r framework, fig.align='center', out.width="80%", echo=FALSE}
 #| fig.cap =
-#| "\\label{fig:framework}
-#| General framework for localized models."
+#| "General framework for localized models."
 knitr::include_graphics("../../fig/fig-typology.pdf")
@@
 
-As discussed in Section \ref{sec:rationales}, most studies were premised on the potential predictive value of localized models.
-As cautioned in Section \ref{sec:performance}, not all experiments affirmed this premise.
+As discussed in Section \ref{rationales}, most studies were premised on the potential predictive value of localized models.
+As cautioned in Section \ref{performance-assessments}, not all experiments affirmed this premise.
 Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
@@
 
-While most studies identified outstanding technical needs (Section \ref{sec:needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
+While most studies identified outstanding technical needs (Section \ref{identified-needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
 Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
@@
 # Data and code
-\label{sec:data}
 
@@
 ## Purpose of review
-\label{sec:appendix-purpose}
 
@@
 
-## Search
-\label{sec:appendix-search}
+## Search design
 
@@
 ## Screening process
-\label{sec:appendix-screening}
 
@@
 ## Selection process
-\label{sec:appendix-selection}
 
-Some PDFs were obtained using Zotero from a university workstation, and the remaining were obtained through university library services. The review process is detailed in Section \ref{sec:full-text}.
+Some PDFs were obtained using Zotero from a university workstation, and the remaining were obtained through university library services.
+The review process is detailed in Section \ref{full-text-review}.
 

### 2023-09-10 docs/Manuscript/Manuscript.tex
make additional edits during resubmission process

@@
 We discuss in this review a hybrid computational approach we call
-\emph{localized modeling} (see Section \ref{sec:coherence}) that appears
-to have been introduced independently several times. The approach fits
-ML models, including regularized families of generalized regression
-models (GLMs), to cohorts retrieved via similarity matching. Whereas CBR
-and NN prediction use patient similarity to obtain an ideally
-homogeneous subpopulation matched to the index case, in localized
-modeling one seeks a relevant yet heterogeneous subpopulation, each
-member of which may deviate from the index case. A prediction or
-inference is then obtained from a model fitted to this \emph{similarity
-cohort}. The use of similarity cohorts distinguishes localized modeling
-from classical GLMs that still dominate analysis in clinical trials and
-comparative effectiveness research, as well as from many cutting-edge ML
-models that are fitted to whole populations, while the cohorts'
-heterogeneity and its utility to the model distinguish the approach from
-CBR. In principle, this allows interpretable models like GLMs to
-identify and measure the effects of determining factors that are
-specific to the index case. In the reviewed studies, localized modeling
-shows some promise toward simultaneously achieving the performance
-improvements of ML and the interpretability of classical models.
+\emph{localized modeling} (see Section \ref{coherence}) that appears to
+have been introduced independently several times. The approach fits ML
+models, including regularized families of generalized regression models
+(GLMs), to cohorts retrieved via similarity matching. Whereas CBR and NN
+prediction use patient similarity to obtain an ideally homogeneous
+subpopulation matched to the index case, in localized modeling one seeks
+a relevant yet heterogeneous subpopulation, each member of which may
+deviate from the index case. A prediction or inference is then obtained
+from a model fitted to this \emph{similarity cohort}. The use of
+similarity cohorts distinguishes localized modeling from classical GLMs
+that still dominate analysis in clinical trials and comparative
+effectiveness research, as well as from many cutting-edge ML models that
+are fitted to whole populations, while the cohorts' heterogeneity and
+its utility to the model distinguish the approach from CBR. In
+principle, this allows interpretable models like GLMs to identify and
+measure the effects of determining factors that are specific to the
+index case. In the reviewed studies, localized modeling shows some
+promise toward simultaneously achieving the performance improvements of
+ML and the interpretability of classical models.
 
@@
 the reasons for them. More details are included in Section
-\ref{sec:appendix-selection}.
+\ref{selection-process}.
 
@@
 the PubMed search first, then adapted it to the other platforms (Section
-\ref{sec:appendix-search}).
+\ref{search-design}).
 
@@
 \subsection{Full text review}\label{full-text-review}}
-
-\label{sec:full-text}
 
@@
 Following the selection sources aligning with our review, we collected
-characteristics of included studies (Section \ref{sec:data}). The
+characteristics of included studies (Section \ref{data-and-code}). The
 features included bibliographic fields (date of publication, journal,
@@
 
-\caption{\label{fig:prisma}PRISMA-S flow chart. Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).}\label{fig:fig:prisma}
+\caption{PRISMA-S flow chart. Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).}\label{fig:prisma}
 \end{figure}
@@
 The 25 included studies were analyzed based on several characteristics
-(see Section \ref{sec:data}), and we report and describe some
+(see Section \ref{data-and-code}), and we report and describe some
 observations here (Table \ref{tab:synthesis}). The studies were
@@
 
-\caption{\label{fig:year}Number of publications in our sample each year.}\label{fig:fig:year}
+\caption{Number of publications in our sample each year.}\label{fig:year}
 \end{figure}
@@
 
-\caption{\label{fig:properties}Share of studies characterized by several design elements: data source (row), clinical task (fill color), and remit (opacity).}\label{fig:fig:properties}
+\caption{Share of studies characterized by several design elements: data source (row), clinical task (fill color), and remit (opacity).}\label{fig:properties}
 \end{figure}
@@
 
-\label{sec:rationales}
+\label{rationales}
 
@@
 of localized modeling specific to clinical and medical settings and
-advantages over other modeling approaches (see Section \ref{sec:data}).
-The most common was that the restriction to similar or relevant past
-cases would improve predictive performance for the index case
-{[}11,15,16,26{]}. In particular, {[}14{]} hypothesized and confirmed
-that the value of each past case would be positively related to its
-similarity to the index case, an assumption built in to the weighting
-schemes of other approaches. {[}12{]} made a different case, that the
-fewer parametric assumptions and complications of a CBR-style model
-would allow for greater accuracy. Additionally, several investigators
-asserted that the use of localized cohorts befit the clinical focus on
-the individual patient rather than the population, without reference to
-performance {[}21,23{]}.
+advantages over other modeling approaches (see Section
+\ref{data-and-code}). The most common was that the restriction to
+similar or relevant past cases would improve predictive performance for
+the index case {[}11,15,16,26{]}. In particular, {[}14{]} hypothesized
+and confirmed that the value of each past case would be positively
+related to its similarity to the index case, an assumption built in to
+the weighting schemes of other approaches. {[}12{]} made a different
+case, that the fewer parametric assumptions and complications of a
+CBR-style model would allow for greater accuracy. Additionally, several
+investigators asserted that the use of localized cohorts befit the
+clinical focus on the individual patient rather than the population,
+without reference to performance {[}21,23{]}.
 
@@
 
-\caption{\label{fig:methods}Frequency of recurring methodological elements across included studies.}\label{fig:fig:methods}
+\caption{Frequency of recurring methodological elements across included studies.}\label{fig:methods}
 \end{figure}
@@
 
-\label{sec:performance}
+\label{performance-assessments}
 
@@
 \subsection{Identified needs}\label{identified-needs}}
-
-\label{sec:needs}
 
@@
 \subsection{Coherence}\label{coherence}}
-
-\label{sec:coherence}
 
@@
 
-\caption{\label{fig:framework}General framework for localized models.}\label{fig:fig:framework}
+\caption{General framework for localized models.}\label{fig:framework}
 \end{figure}
@@
 
-As discussed in Section \ref{sec:rationales}, most studies were premised
-on the potential predictive value of localized models. As cautioned in
-Section \ref{sec:performance}, not all experiments affirmed this
+As discussed in Section \ref{rationales}, most studies were premised on
+the potential predictive value of localized models. As cautioned in
+Section \ref{performance-assessments}, not all experiments affirmed this
 premise. Most of those that did involved using metric learning to
@@
 While most studies identified outstanding technical needs (Section
-\ref{sec:needs}), few emphasized the need to assess human-focused
+\ref{identified-needs}), few emphasized the need to assess human-focused
 qualities like user interface and user experience, interpretability,
@@
 
-\label{sec:data}
-
 Search results are publicly available in a Zotero group library. Data
@@
 
-\label{sec:appendix-purpose}
-
 \begin{enumerate}
@@
 
-\hypertarget{search-1}{%
-\subsection{Search}\label{search-1}}
-
-\label{sec:appendix-search}
+\hypertarget{search-design}{%
+\subsection{Search design}\label{search-design}}
 
@@
 
-\label{sec:appendix-screening}
-
 Most deduplication was done automatically in Covidence. As full-text
@@
 
-\label{sec:appendix-selection}
-
 Some PDFs were obtained using Zotero from a university workstation, and
 the remaining were obtained through university library services. The
-review process is detailed in Section \ref{sec:full-text}.
+review process is detailed in Section \ref{full-text-review}.
 

### 2023-09-10 scr/synthesis.r

@@ -257,6 +257,10 @@ ggsave(
   here::here("fig/fig-properties.png"), properties_mosaic,
   width = 8, height = 4
 )
+ggsave(
+  here::here("fig/fig-properties.jpg"), properties_mosaic,
+  width = 8, height = 4
+)
 
 # years of publication
 composite_studies %>%
@@ -286,6 +290,7 @@ composite_studies %>%
   year_plot
 print(year_plot)
 ggsave(here::here("fig/fig-years.png"), year_plot, width = 6, height = 2)
+ggsave(here::here("fig/fig-years.jpg"), year_plot, width = 6, height = 2)
 
 # approach types
 composite_studies %>%
@@ -350,3 +355,4 @@ composite_approaches %>%
   method_freq
 print(method_freq)
 ggsave(here::here("fig/fig-methods.png"), method_freq, width = 8, height = 3)
+ggsave(here::here("fig/fig-methods.jpg"), method_freq, width = 8, height = 3)

### 2023-10-28 docs/Manuscript/Manuscript.Rmd
add table of specializations from general framework to studies

@@
 composite_studies <- readr::read_rds(here::here("data/composite.rds"))
+framework_specs <- readr::read_rds(here::here("data/framework.rds"))
 tab_eval_comp <- readr::read_rds(here::here("data/eval-comp.rds"))
@@
 
-We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
+We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998a; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
 While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools).
@@
       "\\label{tab:composite}",
-      "Methodological elements of studies included in the synthesis. "
+      "Methodological elements of studies included in the synthesis."
     )
@@
 Some earlier studies tuned generalized regression models, but more recent studies did not.
+
+```{r tab:framework, echo=FALSE, message=FALSE, warning=FALSE}
+framework_specs |> 
+  mutate(
+    Citation = sapply(
+      str_split(Citation, "; "),
+      function(s) str_c("@", s, collapse = "; ")
+    )
+  ) |> 
+  mutate(Citation = str_c("[", Citation, "]")) |> 
+  # prevent numeric list columns from dominating the horizontal space
+  mutate(across(c(-Citation), ~ str_c(., "\\hspace{18em}"))) |> 
+  knitr::kable(
+    align = "llll",
+    caption = paste0(
+      "\\label{tab:framework}",
+      "Specializations of general framework to included studies.",
+      " Flags: R = recurse, S = supervised, T = tuned (optimized)."
+    )
+  )
+```
 

### 2023-10-28 docs/Manuscript/Manuscript.tex
add table of specializations from general framework to studies

@@
 
-\usepackage{lineno} % add
-  \linenumbers % turns line numbering on
-
 \usepackage{graphicx}
@@
 \usepackage{amssymb,amsmath}
+% TODO: Currently lineno needs to be loaded after amsmath because of conflict
+% https://github.com/latex-lineno/lineno/issues/5
+\usepackage{lineno} % add
+  \linenumbers % turns line numbering on
 \usepackage{ifxetex,ifluatex}
@@
 not.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2087}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2795}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1535}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3583}}@{}}
+\caption{\label{tab:framework}Specializations of general framework to
+included studies. Flags: R = recurse, S = supervised, T = tuned
+(optimized).}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Relevance
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Retrieval
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Adaptation
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Relevance
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Retrieval
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Adaptation
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+{[}11{]} & Levenshtein distance\hspace{18em} & Statistical\hspace{18em}
+& Survival (S)\hspace{18em} \\
+{[}19{]} & Weighted sum (S)\hspace{18em} & Combinatorial\hspace{18em} &
+Proportion (S)\hspace{18em} \\
+{[}20{]} & Kohonen mapping\hspace{18em} & Adaptive\hspace{18em} &
+Representation (S)\hspace{18em} \\
+{[}13{]} & Euclidean distance\hspace{18em} & Statistical
+(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
+{[}22{]} & Entropic distance\hspace{18em} & Combinatorial\hspace{18em} &
+Proportion (S)\hspace{18em} \\
+{[}23{]} & Euclidean distance\hspace{18em} & Combinatorial\hspace{18em}
+& Proportion (S)\hspace{18em} \\
+{[}18,21,25,26{]} & Feature selection (R,S)\hspace{18em} & Combinatorial
+(R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
+{[}24{]} & User-defined\hspace{18em} & User-defined\hspace{18em} &
+User-defined\hspace{18em} \\
+{[}12{]} & Mahalanobis distance (S)\hspace{18em} & Combinatorial
+(T)\hspace{18em} & Proportional hazards (S)\hspace{18em} \\
+{[}27{]} & Weighted overlap\hspace{18em} & Combinatorial
+(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
+{[}28{]} & Distance metric learning (S)\hspace{18em} &
+Combinatorial\hspace{18em} & Association rules (S,T)\hspace{18em} \\
+{[}15{]} & Locally supervised metric learning (S); Feature
+selection\hspace{18em} & Combinatorial\hspace{18em} & Logistic model
+(S)\hspace{18em} \\
+{[}14{]} & Cosine similarity\hspace{18em} & Combinatorial
+(T)\hspace{18em} & Logistic model (S); Decision tree (S)\hspace{18em} \\
+{[}29{]} & Range overlap\hspace{18em} & Clustering (S)\hspace{18em} &
+Artificial neural network (S)\hspace{18em} \\
+{[}16{]} & Cosine similarity; Random forest proximity (S)\hspace{18em} &
+Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree (S);
+Random forest (S)\hspace{18em} \\
+{[}31{]} & Unclear\hspace{18em} & Unclear\hspace{18em} &
+Unclear\hspace{18em} \\
+{[}30{]} & Gaussian process model\hspace{18em} & Complete\hspace{18em} &
+Weighted support vector machine (S); Weighted Gaussian process
+regression (S)\hspace{18em} \\
+{[}17{]} & Weighted sum\hspace{18em} & Combinatorial (T)\hspace{18em} &
+Logistic model (S); Random forest (S); Nearest neighbors
+(S)\hspace{18em} \\
+{[}32{]} & Weighted sum\hspace{18em} & Combinatorial\hspace{18em} &
+Extreme learning machine (S)\hspace{18em} \\
+{[}33{]} & Weighted sum\hspace{18em} & Statistical\hspace{18em} & Linear
+model\hspace{18em} \\
+{[}34,35{]} & Locally supervised metric learning (S)\hspace{18em} &
+Causal inference matching\hspace{18em} & Statistical test
+(S)\hspace{18em} \\
+\end{longtable}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
 \CSLLeftMargin{{[}5{]} }%
@@
 \CSLRightInline{Dai L, Zhu H, Liu D.
-\href{https://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
+\href{http://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
 applications}. arXiv:201201976 {[}Cs{]} 2020.}
@@
 \CSLLeftMargin{{[}39{]} }%
-\CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
-{Explore}, {Explain}, and {Examine Predictive Models}. {New York}:
-{Chapman and Hall/CRC}; 2021.}
+\CSLRightInline{Biecek P, Burzykowski T.
+\href{https://ema.drwhy.ai/}{Explanatory {Model Analysis}: {Explore},
+{Explain}, and {Examine Predictive Models}}. {New York}: {Chapman and
+Hall/CRC}; 2021.}
 
@@
 \CSLLeftMargin{{[}40{]} }%
-\CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
-for {Making Black Box Models Explainable}. 2023.}
+\CSLRightInline{Molnar C.
+\href{https://christophm.github.io/interpretable-ml-book}{Interpretable
+{Machine Learning}: {A Guide} for {Making Black Box Models
+Explainable}}. 2023.}
 
@@
 \CSLLeftMargin{{[}43{]} }%
-\CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
-Survey}. PhD thesis. Michigan State University, 2006.}
+\CSLRightInline{Yang L.
+\href{http://www.cs.cmu.edu/~./liuy/frame_survey_v2.pdf}{Distance
+{Metric Learning}: {A Comprehensive Survey}}. PhD thesis. Michigan State
+University, 2006.}
 
@@
 \CSLLeftMargin{{[}45{]} }%
-\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
-Learning} with {Application} to {Clustering} with {Side-Information}.
-Advances in {Neural Information Processing Systems}, vol. 15, {MIT
-Press}; 2002.}
+\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A.
+\href{https://proceedings.neurips.cc/paper/2002/hash/c3e4035af2a1cde9f21e1ae1951ac80b-Abstract.html}{Distance
+{Metric Learning} with {Application} to {Clustering} with
+{Side-Information}}. Advances in {Neural Information Processing
+Systems}, vol. 15, {MIT Press}; 2002.}
 

### 2023-10-28 scr/synthesis.r

@@ -1,6 +1,8 @@
 library(tidyverse)
 library(ggmosaic)
 
+googlesheets4::gs4_auth()
+
 #' Read data.
 
 # journals of references in previous reviews
@@ -21,6 +23,11 @@ properties_inclusion <- googlesheets4::read_sheet(
 #   readxl::read_xlsx() %>%
 composite_techniques <- googlesheets4::read_sheet(
   ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 1L
+)
+
+# methodological specifications within general framework
+framework_specializations <- googlesheets4::read_sheet(
+  ss = "1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs", sheet = 2L
 )
 
 # add citation numbers from most recent LaTeX compilation
@@ -154,6 +161,21 @@ properties_inclusion %>%
   print() -> tab_eval_comp
 # write to file, to be read into document
 write_rds(tab_eval_comp, file = here::here("data/eval-comp.rds"))
+
+# table of specifications from general framework to retrieved studies
+framework_specializations |> 
+  rename(Citation = `Zotero key`) |> 
+  mutate(across(
+    c(Relevance, Retrieval, Adaptation),
+    \(x) str_replace_all(x, "\\) \\(", ",")
+  )) |> 
+  mutate(across(
+    c(Relevance, Retrieval, Adaptation),
+    \(x) str_replace_all(x, "\\n", "; ")
+  )) |> 
+  print() -> framework_specs
+# write to file, to be read into document
+write_rds(framework_specs, file = here::here("data/framework.rds"))
 
 #' Print summaries.
 

### 2023-11-25 docs/Localized-Modeling/Localized-Modeling.Rmd
relocate bibliography files + begin submission to air

--- 2023-10-28 docs/Manuscript/Manuscript.Rmd
@@
 ---
+# Supported options: 
+#   sn-nature:       Style for submissions to Nature Portfolio journals
+#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
+#   sn-mathphys:     Math and Physical Sciences Reference Style
+#   sn-aps:          American Physical Society (APS) Reference Style
+#   sn-vancouver:    Vancouver Reference Style
+#   sn-apa:          APA Reference Style 
+#   sn-chicago:      Chicago-based Humanities Reference Style
+#   default:         Default
+
+classoptions: 
+  - sn-nature      
+  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
+  # - referee       # Optional: Use double line spacing 
+  # - lineno        # Optional: Add line numbers
+  # - iicol         # Optional: Double column layour
+
 title: "Clinical prediction with localized modeling using similarity-based cohorts: A scoping review"
-author:
-  - name: Adi Cohen
+titlerunning: Clinical Prediction with Localized Modeling
+authors: 
+  - firstname: Adi
+    lastname: Cohen
+    degrees: BS?
     email: ac4569@my.nsu.edu
     affiliation: 1
-  - name: Patti McCall-Junkin
+  - firstname: Patti
+    lastname: McCall-Junkin
+    degrees: MA, MLS
     email: pattimccall.junkin@gmail.com
     affiliation: 2
-  - name: Jason Cory Brunson
+  - firstname: Jason Cory
+    lastname: Brunson
+    degrees: PhD
     email: jason.brunson@medicine.ufl.edu
+    orcid: 0000-0003-3126-9494
     affiliation: 3
-    correspondingauthor: true
+    corresponding: TRUE
+affiliations:
+  - number: 1
+    info:
+      orgdiv: College of Osteopathic Medicine
+      orgname: Nova Southeastern University
     address:
-  - code: 1
-    address: College of Osteopathic Medicine, Nova Southeastern University, 3200 S University Drive, Davie, FL 33328
-  - code: 2
-    address: Smathers Libraries Academic Research & Consulting Services, University at Florida, P. O. Box 117000, Gainesville, FL 32611-7000
-  - code: 3
-    address: Laboratory for Systems Medicine, University of Florida, P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225
-abstract: |
-  Background:
-  Applications of classical case-based reasoning (CBR) to clinical and health tasks have given rise to a family of techniques in which a statistical model is fitted to a subset of labeled cases, selected for relevance or similarity to an unlabeled case to which the model is applied.
-These localized models, as we propose to call them, may perform as well as "black-box" models while retaining direct interpretations.
-Their adoption has been hindered by fragmented development and implementation challenges.
-Our goal in this scoping review was to describe the clinical and health applications of localized models to date and to derive a general framework for their design and evaluation.
-
-  Methods:
-  We searched four electronic bibliographic platforms (PubMed, Web of Science, Academic Search Premier, and MathSciNet) during 2021 July 19–22.
-We screened and reviewed entries based on four criteria intended to specify applications of localized models to tasks involving analysis of clinical or health data.
-Two authors divided title/abstract screening and collaboratively reviewed screened entries for inclusion.
-We also screened and reviewed a seed set that inspired the review and traced methodological citations from included entries.
-We discussed and summarized settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
-We received no specific funding for this work.
-
-  Results:
-  Of 1,309 search results, 328 were reviewed.
-Of these, combined with 43 publications that seeded the review and 1 obtained by citation tracking, 25 were included in the review and methodological synthesis.
-The specificity of several search terms was poor, and inter-rater reliability was low.
-Entries were published from 1997 to 2021, with half of these since 2015.
-The most common tasks were prognosis and diagnosis.
-The studies overwhelmingly focused on clinical data, occasionally including laboratory and image-derived data.
-Only three studies applied localized models to tasks other than prediction.
-Several studies commented on memory and runtime costs, and four proposed partial solutions.
-While few methods were reused and no results were reproduced, a general technique that specializes to most of those encountered was straightforward to describe.
-
-  Conclusions:
-  Localized models have potential to improve the performance of models used in clinical decision support tools while maintaining interpretability, provided computational hurdles can be overcome.
-Our search undoubtedly failed to capture all studies that used methods of this kind, but those we found follow a consistent enough design to be reproduced by a general-purpose tool.
-We hope that our review and synthesis will spur future work on efficiency, reproducibility, and user interactions and needs.
+        street: 3200 S University Drive
+        city: Davie
+        postcode: 33328
+        state: FL
+        country: USA
+  - number: 2
+    info:
+      orgdiv: Smathers Libraries Academic Research & Consulting Services
+      orgname: University of Florida
+    address:
+        street: P.O. Box 117000
+        city: Gainesville
+        postcode: 32611-7000
+        state: FL
+        country: USA
+  - number: 3
+    corresponding: TRUE
+    info:
+      orgdiv: Laboratory for Systems Medicine
+      orgname: University of Florida
+    address:
+        street: P.O. Box 100225 JHMHC
+        city: Gainesville
+        postcode: 32610-0225
+        state: FL
+        country: USA
 keywords:
@@
   - scoping review
-journal: "Artificial Intelligence In Medicine"
-date: "`r format(Sys.Date(), format = '%Y %b %d')`"
-classoption: preprint, 3p, authoryear
+# pacs: 
+#   jel:
+#     - "D8"
+#     - "H51"
+#   msc:
+#     - "35A01"
+#     - "65L10"
+
+# Sample for structured abstract
+abstract: |
+  Background:
+  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of similar labeled cases.
+We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
+
+  Methods:
+  We searched four bibliographic platforms during 2021 July 19–22.
+We set four eligibility criteria to identify applications of localized models to clinical and health tasks.
+Two authors divided title/abstract screening and reviewed screened entries for inclusion.
+We discussed settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
+
+  Results:
+  Of 1,309 search results, 328 were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
+25 were included, published 1997–2021.
+The specificity of search terms was poor, and inter-rater reliability was low.
+Almost all models were predictive, the most common tasks being prognosis and diagnosis.
+Most studies used clinical, occasionally laboratory and image, data.
+Several addressed memory and runtime costs.
+A general technique that specializes to most of those reviewed was straightforward to describe.
+
+  Conclusions:
+  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, provided computational hurdles are overcome, and they follow a consistent pattern that could be implemented in a general-purpose way.
+We hope that our review will spur future work on efficiency, reproducibility, and user needs.
+
 bibliography:
-  - SR-bibliography.bib
-  - SR-0-Found.bib
-  - SR-2-Included.bib
-  - SR-3-Tracked.bib
-linenumbers: true
-numbersections: true
-# Use a CSL with `citation_package = "default"`
-# csl: https://www.zotero.org/styles/elsevier-harvard
-# csl: https://www.zotero.org/styles/artificial-intelligence-in-medicine
-csl: artificial-intelligence-in-medicine.csl
+  - ../bibliography/SR-bibliography.bib
+  - ../bibliography/SR-0-Found.bib
+  - ../bibliography/SR-2-Included.bib
+  - ../bibliography/SR-3-Tracked.bib
+
+header-includes: |
+  \usepackage{lscape}
+  \newcommand{\landscapebegin}{\begin{landscape}}
+  \newcommand{\landscapeend}{\end{landscape}}
+
+  %% Per the spinger doc, new theorem styles can be included using built in style, 
+  %% but it seems the don't work so commented below
+  %\theoremstyle{thmstyleone}%
+  \newtheorem{theorem}{Theorem}%  meant for continuous numbers
+  %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+  %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+  \newtheorem{proposition}[theorem]{Proposition}%
+  %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+  %% \theoremstyle{thmstyletwo}%
+  \theoremstyle{remark}
+  \newtheorem{example}{Example}%
+  \newtheorem{remark}{Remark}%
+
+  %% \theoremstyle{thmstylethree}%
+  \theoremstyle{definition}
+  \newtheorem{definition}{Definition}%
+
 output:
-  rticles::elsevier_article:
+  rticles::springer_article:
     keep_tex: true
@@
     citation_package: "default"
-header-includes:
-  \usepackage{lscape}
-  \newcommand{\landscapebegin}{\begin{landscape}}
-  \newcommand{\landscapeend}{\end{landscape}}
-
 ---
@@
 
+### Acknowledgments {-}
+
+We are grateful for the astute and supportive comments of three anonymous reviewers.
+
 # Introduction
 
-Clinical decision support (CDS) systems have been a fixture of artificial intelligence (AI) research since inception.
-As machine learning (ML) emerged, clinical modeling tasks became some of the most common and celebrated applications.
-A technique core to clinical applications of both AI and ML is _similarity-matching_, which relies on a _patient similarity measure_[^psm] to extract a cohort of past or training cases used to inform the diagnosis, prognosis, or care of a new or test case.
-Similarity-matching powers the retrieval step of most case-based reasoning (CBR) systems and provides the neighborhoods from which nearest neighbor (NN) models derive predictions.
-Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), traditional CBR tools return the retrieved cases to inform human decision-making.
+A main driver of the development of clinical information systems (CIS) and clinical decision support (CDS) tools has been to leverage the advantages of routinely collected health data toward improving care.
+These advantages include their immediate accessibility through the institutional EHR, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
+Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, with the monicker "practice-based evidence", as an important complement to the paradigm of evidence-based practice.
+
+Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools, including artificial intelligence (AI).
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and their interplay with respect to outcomes.
+The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models.
+
+One of few methodologies to address all three of these established needs is one we call _localized modeling_, which appears to have been introduced several times independently under different names, and in slightly different forms.
+The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
+While traditional CBR returns these retrieved cases to inform human decision-making, the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+In contrast, localized modeling fits families of predictive models, most commonly generalized linear models (GLMs), to these retrieved similarity cohorts in order to generate predictions.
+Localized models thus provide a way to fit fully interpretable models to large population data in a way that is personalized to each case.
 
@@
 
-We discuss in this review a hybrid computational approach we call _localized modeling_ (see Section \ref{coherence}) that appears to have been introduced independently several times.
-The approach fits ML models, including regularized families of generalized regression models (GLMs), to cohorts retrieved via similarity matching.
-Whereas CBR and NN prediction use patient similarity to obtain an ideally homogeneous subpopulation matched to the index case, in localized modeling one seeks a relevant yet heterogeneous subpopulation, each member of which may deviate from the index case.
-A prediction or inference is then obtained from a model fitted to this _similarity cohort_.
-The use of similarity cohorts distinguishes localized modeling from classical GLMs that still dominate analysis in clinical trials and comparative effectiveness research, as well as from many cutting-edge ML models that are fitted to whole populations, while the cohorts' heterogeneity and its utility to the model distinguish the approach from CBR.
-In principle, this allows interpretable models like GLMs to identify and measure the effects of determining factors that are specific to the index case.
-In the reviewed studies, localized modeling shows some promise toward simultaneously achieving the performance improvements of ML and the interpretability of classical models.
-
-<!--
-In multivariate settings, because similarity decreases with difference along each factor, included cases that differ more with respect to some factors will tend to differ less with respect to others.
-Consequently, (1) the distribution of a factor in the subpopulation will tend to be centered near its value at the index case, and (2) the joint distribution of a pair of factors will tend to be sparse away from both center axes, i.e.
-when both factors differ from their values at the index case.
-This should produce localized models that are less responsive to deviation along multiple uncorrelated factors, so that the influence of each factor in the model reflects more its independent effect than its interactions with other factors.
--->
-
-One memorable aspiration of clinical informatics was the "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient @Longhurst2014.
-Localized modeling has recently been put forth as one answer to this appeal.
-Studies using localized models have been limited by computational cost—a naïve implementation involves fitting, optimizing, and evaluating a model on a separate population for each training case—and recent studies have described algorithms and implementations designed to reduce this cost.
-As greater access to health data, informatical knowledge, algorithmic efficiency, and computational power continue to spur innovation in CDS, we may expect localized modeling to become more accessible and efficient, its advantages better understood, and its disadvantages ameliorated.
-
-The motivation for this scoping review is to describe the settings in and problems with which localized models have been tasked to date, and to synthesize existing methods into a general framework.
+This approach harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
+Our motivations in this scoping review are to describe the settings in and problems with which this approach has been tasked, to attempt to organize them within a common framework, and to evaluate the promise they show toward achieving this or other needs in clinical informatics.
 Along the way, we attempt to reconcile terminology, summarize motivations and evaluations, and propose valuable follow-up work.
@@
 #| fig.cap =
-#| "General framework for localized models."
+#| "General framework for localized models.
+#|  Dashed arrows indicate optional steps."
 knitr::include_graphics("../../fig/fig-typology.pdf")
@@
 Whereas precious few of the reviewed studies used these models, for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
-
-# Acknowledgments
-
-We are grateful for the astute and supportive comments of three anonymous reviewers.
 

### 2023-11-25 docs/Manuscript/Manuscript.tex
relocate bibliography files + begin submission to air

@@
     \affiliation[2]{Smathers Libraries Academic Research \& Consulting
-Services, University at Florida, P. O. Box 117000, Gainesville, FL
+Services, University of Florida, P.O. Box 117000, Gainesville, FL
 32611-7000}
@@
 
-Clinical decision support (CDS) systems have been a fixture of
-artificial intelligence (AI) research since inception. As machine
-learning (ML) emerged, clinical modeling tasks became some of the most
-common and celebrated applications. A technique core to clinical
-applications of both AI and ML is \emph{similarity-matching}, which
-relies on a \emph{patient similarity measure}\footnote{Similarity
-  measures between more granular units of analysis, such as encounters
-  and decision points, are often still referred to as patient similarity
-  measures.} to extract a cohort of past or training cases used to
-inform the diagnosis, prognosis, or care of a new or test case.
-Similarity-matching powers the retrieval step of most case-based
-reasoning (CBR) systems and provides the neighborhoods from which
-nearest neighbor (NN) models derive predictions. Whereas NN predictions
-are generated automatically via averaging (regression) or voting
-(classification), traditional CBR tools return the retrieved cases to
-inform human decision-making.
-
-We discuss in this review a hybrid computational approach we call
-\emph{localized modeling} (see Section \ref{coherence}) that appears to
-have been introduced independently several times. The approach fits ML
-models, including regularized families of generalized regression models
-(GLMs), to cohorts retrieved via similarity matching. Whereas CBR and NN
-prediction use patient similarity to obtain an ideally homogeneous
-subpopulation matched to the index case, in localized modeling one seeks
-a relevant yet heterogeneous subpopulation, each member of which may
-deviate from the index case. A prediction or inference is then obtained
-from a model fitted to this \emph{similarity cohort}. The use of
-similarity cohorts distinguishes localized modeling from classical GLMs
-that still dominate analysis in clinical trials and comparative
-effectiveness research, as well as from many cutting-edge ML models that
-are fitted to whole populations, while the cohorts' heterogeneity and
-its utility to the model distinguish the approach from CBR. In
-principle, this allows interpretable models like GLMs to identify and
-measure the effects of determining factors that are specific to the
-index case. In the reviewed studies, localized modeling shows some
-promise toward simultaneously achieving the performance improvements of
-ML and the interpretability of classical models.
-
-One memorable aspiration of clinical informatics was the ``green
-button'' that would, in response to a query, automatically retrieve
-patient data from an institution's records with which to conduct
-on-demand retrospective studies for an individual patient {[}1{]}.
-Localized modeling has recently been put forth as one answer to this
-appeal. Studies using localized models have been limited by
-computational cost---a naïve implementation involves fitting,
-optimizing, and evaluating a model on a separate population for each
-training case---and recent studies have described algorithms and
-implementations designed to reduce this cost. As greater access to
-health data, informatical knowledge, algorithmic efficiency, and
-computational power continue to spur innovation in CDS, we may expect
-localized modeling to become more accessible and efficient, its
-advantages better understood, and its disadvantages ameliorated.
-
-The motivation for this scoping review is to describe the settings in
-and problems with which localized models have been tasked to date, and
-to synthesize existing methods into a general framework. Along the way,
-we attempt to reconcile terminology, summarize motivations and
-evaluations, and propose valuable follow-up work.
+A main driver of the development of clinical information systems (CIS)
+and clinical decision support (CDS) tools has been to leverage the
+advantages of routinely collected health data toward improving care.
+These advantages include their immediate accessibility through the
+institutional EHR, billing database, or other source; their specificity
+to the institution and the population it serves; and the closeness in
+time of the available data. Though by definition not collected for
+research use, the secondary research use of routinely collected health
+data has been put forth, with the monicker ``practice-based evidence'',
+as an important complement to the paradigm of evidence-based practice.
+
+Over the same period of time, interest in individualizing care from
+population-derived evidence-based recommendations to specific patient
+needs has driven the application of advanced computational tools,
+including artificial intelligence (AI). Whereas classical models
+produced formulae involving a limited set of data elements that would be
+uniformly applied to all patients, AI models often process hundreds or
+more variables in opaque ways to yield predictions that depend on not
+only their values but their associations and their interplay with
+respect to outcomes. The loss of straightforward interpretations of
+these models has limited their practical uptake and motivated the
+construction of explanatory statistics for opaque models as well as the
+development of more interpretable complex models.
+
+One of few methodologies to address all three of these established needs
+is one we call \emph{localized modeling}, which appears to have been
+introduced several times independently under different names, and in
+slightly different forms. The approach is a specialized form of
+case-based reasoning (CBR) that relies on a patient similarity
+measure\footnote{Similarity measures between more granular units of
+  analysis, such as encounters and decision points, are often still
+  referred to as patient similarity measures.} to extract a cohort of
+past or training cases that then inform the diagnosis, prognosis, or
+care of a new or test case. While traditional CBR returns these
+retrieved cases to inform human decision-making, the related nearest
+neighbors (NN) technique generates predictions from cohorts
+automatically via averaging (regression) or voting (classification) of
+their outcomes. In contrast, localized modeling fits families of
+predictive models, most commonly generalized linear models (GLMs), to
+these retrieved similarity cohorts in order to generate predictions.
+Localized models thus provide a way to fit fully interpretable models to
+large population data in a way that is personalized to each case.
+
+This approach harkens to the aspirational ``green button'' that would,
+in response to a query, automatically retrieve patient data from an
+institution's records with which to conduct on-demand retrospective
+studies for an individual patient {[}1{]}. Our motivations in this
+scoping review are to describe the settings in and problems with which
+this approach has been tasked, to attempt to organize them within a
+common framework, and to evaluate the promise they show toward achieving
+this or other needs in clinical informatics. Along the way, we attempt
+to reconcile terminology, summarize motivations and evaluations, and
+propose valuable follow-up work.
 

### 2023-11-28 docs/Localized-Modeling/Localized-Modeling.Rmd
track second submission + redistribute bibliography files + script to copy figures

@@
     lastname: Cohen
-    degrees: BS?
+    degrees: BS
     email: ac4569@my.nsu.edu
@@
 bibliography:
-  - ../bibliography/SR-bibliography.bib
-  - ../bibliography/SR-0-Found.bib
-  - ../bibliography/SR-2-Included.bib
-  - ../bibliography/SR-3-Tracked.bib
+  - SR-bibliography.bib
+  - SR-0-Found.bib
+  - SR-2-Included.bib
+  - SR-3-Tracked.bib
 
@@
 
-### Acknowledgments {-}
-
-We are grateful for the astute and supportive comments of three anonymous reviewers.
+\pagebreak
 
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies (Section \ref{data-and-code}).
+Following the selection sources aligning with our review, we collected characteristics of included studies.
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, performance measures, results of evaluations and comparisons, name given to modeling approach).
@@
 #|  citation/reference tracking (s)."
-knitr::include_graphics("../../fig/fig-prisma.pdf")
+knitr::include_graphics("Fig1.pdf")
 ```
@@
 
-The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics (see Section \ref{data-and-code}), and we report and describe some observations here (Table \ref{tab:synthesis}).
+The `r nrow(tab_synthesis)` included studies were analyzed based on several characteristics, and we report and describe some observations here (Table \ref{tab:synthesis}).
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
@@
 So the idea has been "in the air" for as long as CBR has been in use.
+
+\small
 
@@
 
-```{r year, fig.align='center', out.width="75%", echo=FALSE}
+\normalsize
+
+```{r year, fig.align='center', out.width="100%", echo=FALSE}
 #| fig.cap =
 #| "Number of publications in our sample each year."
-knitr::include_graphics("../../fig/fig-years.png")
+knitr::include_graphics("Fig2.png")
 ```
@@
 In the following subsections, we take a closer look at these design elements.
+
+\small
 
@@
 ```
+
+\normalsize
 
@@
 #|  data source (row),
-#|  clinical task (fill color), and
+#|  clinical task (fill), and
 #|  remit (opacity)."
-knitr::include_graphics("../../fig/fig-properties.png")
+knitr::include_graphics("Fig3.png")
 ```
@@
 ## Rationales
-\label{rationales}
 
-The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches (see Section \ref{data-and-code}).
+The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches.
 The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015a; @Ng2015; @Lee2017].
@@
 #| "Frequency of recurring methodological elements across included studies."
-knitr::include_graphics("../../fig/fig-methods.png")
+knitr::include_graphics("Fig4.png")
 ```
@@
 ## Performance assessments
-\label{performance-assessments}
 
@@
 
-```{r framework, fig.align='center', out.width="80%", echo=FALSE}
+```{r framework, fig.align='center', out.width="100%", echo=FALSE}
 #| fig.cap =
@@
 #|  Dashed arrows indicate optional steps."
-knitr::include_graphics("../../fig/fig-typology.pdf")
+knitr::include_graphics("Fig5.pdf")
 ```
@@
 Some earlier studies tuned generalized regression models, but more recent studies did not.
+
+\small
 
@@
 
-# Support
+\pagebreak
 
-The authors received no specific funding or other support for this work.
+# Appendix {-}
 
-# Competing interests
-
-The authors declare that they have no competing interests.
-
-# Data and code
-
-Search results are publicly available in a Zotero group library.
-Data collected or encoded for included studies are publicly available in two Google Sheets.
-Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
-
-* Search results, included studies, and other bibliography: <https://www.zotero.org/groups/5017571/imsr/>
-* Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
-* Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
-* Code used to conduct analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
-
-# Appendix
-
-## Research question
+## Research question {-}
 
@@
 
-## Purpose of review
+## Purpose of review {-}
 
@@
 
-## Search design
+## Search design {-}
 
@@
 
-### Search strings
+### Search strings {-}
 
@@
 
-## Screening process
+## Screening process {-}
 
@@
 
-## Selection process
+## Selection process {-}
 
@@
 
-## Analysis and synthesis
+## Analysis and synthesis {-}
 
@@
 
-## Performance evaluations and comparisons
+## Performance evaluations and comparisons {-}
 
@@
 **TODO: Place vertical separators between major columns.**
+
+\small
 
@@
 
+\normalsize
+
 \landscapeend

### 2023-11-28 docs/Manuscript/Manuscript.Rmd
track second submission + redistribute bibliography files + script to copy figures

@@
   - code: 2
-    address: Smathers Libraries Academic Research & Consulting Services, University at Florida, P. O. Box 117000, Gainesville, FL 32611-7000
+    address: Smathers Libraries Academic Research & Consulting Services, University of Florida, P.O. Box 117000, Gainesville, FL 32611-7000
   - code: 3
@@
 
-Clinical decision support (CDS) systems have been a fixture of artificial intelligence (AI) research since inception.
-As machine learning (ML) emerged, clinical modeling tasks became some of the most common and celebrated applications.
-A technique core to clinical applications of both AI and ML is _similarity-matching_, which relies on a _patient similarity measure_[^psm] to extract a cohort of past or training cases used to inform the diagnosis, prognosis, or care of a new or test case.
-Similarity-matching powers the retrieval step of most case-based reasoning (CBR) systems and provides the neighborhoods from which nearest neighbor (NN) models derive predictions.
-Whereas NN predictions are generated automatically via averaging (regression) or voting (classification), traditional CBR tools return the retrieved cases to inform human decision-making.
+A main driver of the development of clinical information systems (CIS) and clinical decision support (CDS) tools has been to leverage the advantages of routinely collected health data toward improving care.
+These advantages include their immediate accessibility through the institutional EHR, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
+Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, with the monicker "practice-based evidence", as an important complement to the paradigm of evidence-based practice.
+
+Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools, including artificial intelligence (AI).
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and their interplay with respect to outcomes.
+The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models.
+
+One of few methodologies to address all three of these established needs is one we call _localized modeling_, which appears to have been introduced several times independently under different names, and in slightly different forms.
+The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
+While traditional CBR returns these retrieved cases to inform human decision-making, the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+In contrast, localized modeling fits families of predictive models, most commonly generalized linear models (GLMs), to these retrieved similarity cohorts in order to generate predictions.
+Localized models thus provide a way to fit fully interpretable models to large population data in a way that is personalized to each case.
 

### 2023-11-28 scr/submission.r

@@ -1,0 +1,28 @@
+
+# copy figures to _AI Review_ folder
+file.copy(
+  here::here("fig/fig-prisma.pdf"),
+  here::here("docs/Localized-Modeling/Fig1.pdf"),
+  overwrite = TRUE
+)
+file.copy(
+  here::here("fig/fig-years.png"),
+  here::here("docs/Localized-Modeling/Fig2.png"),
+  overwrite = TRUE
+)
+file.copy(
+  here::here("fig/fig-properties.png"),
+  here::here("docs/Localized-Modeling/Fig3.png"),
+  overwrite = TRUE
+)
+file.copy(
+  here::here("fig/fig-methods.png"),
+  here::here("docs/Localized-Modeling/Fig4.png"),
+  overwrite = TRUE
+)
+file.copy(
+  here::here("fig/fig-typology.pdf"),
+  here::here("docs/Localized-Modeling/Fig5.pdf"),
+  overwrite = TRUE
+)
+

### 2023-11-28 scr/synthesis.r

@@ -263,12 +263,13 @@ mosaic_data %>%
     aes(x = product(Task, Source), fill = Task, alpha = Aim, weight = n),
     divider = mosaic("v")
   ) +
-  stat_mosaic_text(
-    # geom = "label",
-    aes(x = product(Aim, Task, Source), weight = n,
-        label = mosaic_refs[match(after_stat(label), mosaic_labs)]),
-    divider = mosaic("v"), size = 3
-  ) +
+  # FIXME: remove reference annotations if journal does not enumerate them
+  # stat_mosaic_text(
+  #   # geom = "label",
+  #   aes(x = product(Aim, Task, Source), weight = n,
+  #       label = mosaic_refs[match(after_stat(label), mosaic_labs)]),
+  #   divider = mosaic("v"), size = 3
+  # ) +
   scale_y_productlist("Source") +
   # scale_y_productlist("Source", labels = ggmosaic:::product_labels()) +
   scale_fill_brewer(type = "qual", na.value = "#000000") +
@@ -311,8 +312,8 @@ composite_studies %>%
   labs(x = NULL, y = NULL) ->
   year_plot
 print(year_plot)
-ggsave(here::here("fig/fig-years.png"), year_plot, width = 6, height = 2)
-ggsave(here::here("fig/fig-years.jpg"), year_plot, width = 6, height = 2)
+ggsave(here::here("fig/fig-years.png"), year_plot, width = 8, height = 2)
+ggsave(here::here("fig/fig-years.jpg"), year_plot, width = 8, height = 2)
 
 # approach types
 composite_studies %>%
@@ -364,8 +365,8 @@ composite_approaches %>%
   ggplot(aes(x = Elements, y = count, label = refs)) +
   theme_bw() + theme(panel.grid.major.y = element_blank()) +
   geom_col() +
-  geom_text(hjust = 0, size = 3) +
-  expand_limits(y = c(NA_real_, 17)) +
+  # geom_text(hjust = 0, size = 3) +
+  # expand_limits(y = c(NA_real_, 17)) +
   # geom_text(hjust = 1, size = 3, color = "white") +
   scale_x_discrete(limits = rev) +
   scale_y_continuous(

### 2024-03-11 docs/Localized-Modeling/Localized-Modeling.Rmd
update bibliography

@@
   - SR-bibliography.bib
+  - SR-reviewers.bib
   - SR-0-Found.bib
@@
   - SR-3-Tracked.bib
+  - SR-5-Screened.bib
 
@@
 library(tidyverse)
-Sys.sleep(10)
+Sys.sleep(2)
 tab_synthesis <- readr::read_rds(here::here("data/synthesis.rds"))
@@
 
-We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998a; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
+We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
 While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools).
@@
 While all included studies were reported in scientific and medical journals, the vast majority were oriented toward clinical practice rather than medical research.
-For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006a].
+For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006].
 This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used an unconventional voting scheme to generate recommendations.
-A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021a].
+A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021].
 In both cases a prototype implementation was deployed in an experimental setting for evaluation.
@@
 @Lee2015 and @Lee2017 used localized logistic regression and random forest modeling to predict 30-day mortality following discharge for ICU patients.
-@Vilhena2016a developed a CBR cycle around a clustering-informed similarity matching procedure and an artificial neural network–based classifier to identify thrombophilia patients at high risk of thrombotic episodes.
-And @Ma2020a took a similarity cohort–based approach to predicting length of stay for ICU patients.
-Also of note, from a public health perspective, @Xu2008a used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
+@Vilhena2016 developed a CBR cycle around a clustering-informed similarity matching procedure and an artificial neural network–based classifier to identify thrombophilia patients at high risk of thrombotic episodes.
+And @Ma2020 took a similarity cohort–based approach to predicting length of stay for ICU patients.
+Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
 
-Toward detection and diagnosis, @Wyns2004a proposed a hybrid neural net–CBR system to distinguish (with confidence bounds) arthritic versus control patients, based on several histological features.
-@Nicolas2014a used collaborative multilabel CBR to subtype melanoma patients based on confocal and dermoscopy images.
+Toward detection and diagnosis, @Wyns2004 proposed a hybrid neural net–CBR system to distinguish (with confidence bounds) arthritic versus control patients, based on several histological features.
+@Nicolas2014 used collaborative multilabel CBR to subtype melanoma patients based on confocal and dermoscopy images.
 @Wang2019 used localized models built from a multi-type additive similarity measure to distinguish type 2 diabetic versus control populations.
-Along the way, @Song2006c, @Kasabov2010, and @Verma2015a took an iterative model-building approach to several tasks: predicting glomerular filtration rate, a key indicator of renal function, from demographic and physiological variables; identifying patients with colon cancer using a large number of gene expression measurements; and identifying patients with type 2 diabetes based on demographic, physiological, and genetic variables.
+Along the way, @Song2006, @Kasabov2010, and @Verma2015 took an iterative model-building approach to several tasks: predicting glomerular filtration rate, a key indicator of renal function, from demographic and physiological variables; identifying patients with colon cancer using a large number of gene expression measurements; and identifying patients with type 2 diabetes based on demographic, physiological, and genetic variables.
 
-While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking "decision points" as their units of analysis, @Tang2021c and @Ng2021b built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
+While several studies emphasized the potential or actual value to decision-making of their methods and tools, only one incorporated treatment decisions into their approach: By taking "decision points" as their units of analysis, @Tang2021 and @Ng2021 built not classifiers or predictors but comparative effectiveness models into a localized framework, providing for the first time in our sample explicitly prescriptive rather than descriptive clinical decision support.
 
-A partial exception to this focus was a 2014 study that also reported a decision support tool, in this case for early diagnosis of melanoma from clinical data and dermoscopy images [@Nicolas2014a].
+A partial exception to this focus was a 2014 study that also reported a decision support tool, in this case for early diagnosis of melanoma from clinical data and dermoscopy images [@Nicolas2014].
 While the stated objectives were analogous, specific emphasis was placed on the acquisition of new knowledge through the development of the tool, including the systematic generation of new data and creation of a clinical ontology.
@@
 The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches.
-The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015a; @Ng2015; @Lee2017].
+The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015; @Ng2015; @Lee2017].
 In particular, @Lee2015 hypothesized and confirmed that the value of each past case would be positively related to its similarity to the index case, an assumption built in to the weighting schemes of other approaches.
 @Lowsky2013 made a different case, that the fewer parametric assumptions and complications of a CBR-style model would allow for greater accuracy.
-Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to performance [@Song2006c; @Xu2008a].
+Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to performance [@Song2006; @Xu2008].
 
-In an interesting contrast, @Tang2021c and @Ng2021b argued that their localized approach using the larger and more heterogeneous populations covered by EHR-derived data could better capture the messy and diverse lessons of everyday practice, as a counterpart to guidelines based on randomized controlled trials.
-Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2012a; @Nicolas2014a; @Verma2015a; @Lee2017].
-@Verma2015a and @Zhang2018a additionally pointed out that localized models are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
+In an interesting contrast, @Tang2021 and @Ng2021 argued that their localized approach using the larger and more heterogeneous populations covered by EHR-derived data could better capture the messy and diverse lessons of everyday practice, as a counterpart to guidelines based on randomized controlled trials.
+Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2013; @Nicolas2014; @Verma2015; @Lee2017].
+@Verma2015 and @Zhang2018 additionally pointed out that localized models are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
 These properties, they said, tend to be more difficult for whole-population models to handle.
@@
 The other frequent advantage attributed to localized modeling was interpretability.
-@Elter2007a and @Nicolas2014a emphasized the value of the detailed past cases, available to the user, on which predictions are based.
+@Elter2007 and @Nicolas2014 emphasized the value of the detailed past cases, available to the user, on which predictions are based.
 This "self-explanation capability" made outputs more intelligible to physicians in the CDS setting.
@@
 
-The remaining coded rationales were for augmentations or hybridizations of then-conventional CBR, most of which defended the use of other tools to improve performance via cohort selection [@CampilloGimenez2012a; @Nicolas2014a; @Vilhena2016a] or to make models and outputs more interpretable [@Lopez2011a; @Wang2019].
-@Liang2015a argued for simultaneous optimization of cohort and feature selection with model parameterization; their TWNFI approach combines localization with regularization.
+The remaining coded rationales were for augmentations or hybridizations of then-conventional CBR, most of which defended the use of other tools to improve performance via cohort selection [@CampilloGimenez2013; @Nicolas2014; @Vilhena2016] or to make models and outputs more interpretable [@Lopez2011; @Wang2019].
+@Liang2015 argued for simultaneous optimization of cohort and feature selection with model parameterization; their TWNFI approach combines localization with regularization.
 
@@
 
-CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006a describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
+CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
 This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle.
@@
 @Lowsky2013 used non-parametric models to reduce the computational burden of revision (prediction).
-@Ma2020a proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021b and @Tang2021c partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
+@Ma2020 proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021 and @Tang2021 partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
 
 Other studies addressed limitations of available tools.
-@Lopez2011a implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
-Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007a; @Liang2015a; @Ng2015; @Zhang2018a].
+@Lopez2011 implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
+Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
 The frequency plot in Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
@@
 
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004a], TWNFI [@Song2006c, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015a], CBR + rules (+ DML) [@Nicolas2014a], Gaussian process regression [@Zhang2018a], JITL-ELM [@Ma2020a]) are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
-Though @CampilloGimenez2012a and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]) are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
+Though @CampilloGimenez2013 and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
 Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model (statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]), they did not consistently outperform comparators.
@@
 
-The study authors focused their recommendations and their own plans for future work mostly on technical improvements and evaluations [@Mariuzzi1997; @Yearwood1997a].
-Urged improvements included full or partial automation of predictor selection [@Mariuzzi1997; @Yearwood1997a], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006c; @Lee2017]; extensions to new data structures [@Lopez2011a], data types [@Liang2015a; @Verma2015a; @Malykh2018a], and reasoning systems [@Nicolas2014a]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015a; @Zhang2018a].
-Authors also urged validations and independent evaluations using larger or more comprehensive data sets [@Elter2007a; @Xu2008a; @Verma2015a; @Ng2015], using data aggregated from multiple health systems [@Lee2015; @Lee2017; @Tang2021c; @Ng2021b], and in other care settings or disease contexts [@Song2006c; @Zhang2018a; @Tang2021c; @Ng2021b].
+The study authors focused their recommendations and their own plans for future work mostly on technical improvements and evaluations [@Mariuzzi1997; @Yearwood1997].
+Urged improvements included full or partial automation of predictor selection [@Mariuzzi1997; @Yearwood1997], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006; @Lee2017]; extensions to new data structures [@Lopez2011], data types [@Liang2015; @Verma2015; @Malykh2018], and reasoning systems [@Nicolas2014]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015; @Zhang2018].
+Authors also urged validations and independent evaluations using larger or more comprehensive data sets [@Elter2007; @Xu2008; @Verma2015; @Ng2015], using data aggregated from multiple health systems [@Lee2015; @Lee2017; @Tang2021; @Ng2021], and in other care settings or disease contexts [@Song2006; @Zhang2018; @Tang2021; @Ng2021].
 
 Less common were calls to strengthen the connection between the methods and the users.
-Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997a; @Wang2019].
-Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004a], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
+Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997; @Wang2019].
+Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
 Most of this incremental work was indeed carried out in later investigations in our sample.
@@
 The motivational and methodological unity of these studies does not reflect a unified research program.
-Besides the lack of any primary journal of record, we observed collaborations only among the authors of smaller contiguous programs, including applications of the TWNFI methodology [@Song2006c; @Verma2015a], individualized mortality prediction for ICU patients [@Lee2015; @Lee2017] and the use of more explanatory models to prioritize predictors or treatments for chronic disease [@Ng2015; @Tang2021c; @Ng2021b].
+Besides the lack of any primary journal of record, we observed collaborations only among the authors of smaller contiguous programs, including applications of the TWNFI methodology [@Song2006; @Verma2015], individualized mortality prediction for ICU patients [@Lee2015; @Lee2017] and the use of more explanatory models to prioritize predictors or treatments for chronic disease [@Ng2015; @Tang2021; @Ng2021].
 
-These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006c], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015a, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020a], and "precision cohort" analysis [@Wang2019; @Tang2021c; @Ng2021b].
+These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
 We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
@@
 Each used its similarity measure to retrieve training cases relevant to each testing case.
-Once retrieved, most took a standard NN approach to generating predictions; two exceptions [@Vilhena2016a; @Tang2021c] fit predictive models to the retrieved cohorts.
+Once retrieved, most took a standard NN approach to generating predictions; two exceptions [@Vilhena2016; @Tang2021] fit predictive models to the retrieved cohorts.
 
-Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007a].
-In particular, @Yearwood1997a defined similarity as a weighted sum of differences in predictor values and used linear regression to optimize the weights for the predictive accuracy of the cohort they retrieve.
-Others used supervised learning: @Song2006c proposed an iterative algorithm to optimize a set of fuzzy inference rules used to retrieve relevant cases and generate a prediction, using back-propagation on the rules’ parameters.
+Several studies used unsupervised similarity learning, for example the Mahalanobis distance [@Lowsky2013] and the Kolmogorov entropy-based distance [@Elter2007].
+In particular, @Yearwood1997 defined similarity as a weighted sum of differences in predictor values and used linear regression to optimize the weights for the predictive accuracy of the cohort they retrieve.
+Others used supervised learning: @Song2006 proposed an iterative algorithm to optimize a set of fuzzy inference rules used to retrieve relevant cases and generate a prediction, using back-propagation on the rules’ parameters.
 Their method was used in several later studies returned by our search, which are not discussed here because they did not originate the technique.
-@Nicolas2014a explicitly appeal to a supervised distance metric learning (DML) technique [@Xing2002] to obtain a similarity measure on their set of dermoscopy and confocal images that most effectively separates malignant from benign melanoma tumors.
+@Nicolas2014 explicitly appeal to a supervised distance metric learning (DML) technique [@Xing2002] to obtain a similarity measure on their set of dermoscopy and confocal images that most effectively separates malignant from benign melanoma tumors.
 We will discuss the remaining uses of supervised similarity learning shortly.
@@
 This is analogous to optimizing the neighborhood size parameter in NN predictive modeling, so we will not discuss it further except to mention that the authors found statistical CBR to outperform conventional CBR on several data sets.
-@CampilloGimenez2012a combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the NN model using the previously optimized weights.
+@CampilloGimenez2013 combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the NN model using the previously optimized weights.
 
@@
 The retrieved studies whose use of such techniques alone satisfied our criteria variably termed them individualized, personalized, local, and patient-specific models, among other terms (Table \ref{tab:composite}).
-We prefer the term localized models, which is both sensitive and specific to uses of a similarity or distance measure to retrieve a cohort to which a model is fit: It includes some properly included techniques we did not have in mind [@Vilhena2016a] yet excludes other techniques commonly referred to as individualized, personalized, or precision.
+We prefer the term localized models, which is both sensitive and specific to uses of a similarity or distance measure to retrieve a cohort to which a model is fit: It includes some properly included techniques we did not have in mind [@Vilhena2016] yet excludes other techniques commonly referred to as individualized, personalized, or precision.
 
@@
 
-In addition to @Lowsky2013, @Liang2015a, @Ng2015, @Lee2017, @Vilhena2016a, @Tang2021c, and @Ng2021b, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015a, @Wang2019, and @Ma2020a.
-In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021c; @Ng2021b] later.
-The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016a] or via cross-validation [@Lowsky2013; @Verma2015a].
-Further in the former direction, one study [@Ma2020a] set a fixed cohort size while two [@Liang2015a; @Tang2021c; @Ng2021b] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
+In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, and @Ma2020.
+In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021; @Ng2021] later.
+The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016] or via cross-validation [@Lowsky2013; @Verma2015].
+Further in the former direction, one study [@Ma2020] set a fixed cohort size while two [@Liang2015; @Tang2021; @Ng2021] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
 
@@
 Foremost was the optimization of retrieved cohort sizes for some measure of model performance.
-Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021c; @Ng2021b] in this corpus took this approach.
+Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach.
 Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.
 
-An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015a.
-More recently, @Tang2021c and @Ng2021b proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance.
+An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015.
+More recently, @Tang2021 and @Ng2021 proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance.
 Notably, though their similarity measure was supervised, this optimization process was not.
@@
 With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
-This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021c; @Ng2021b].
+This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021].
 These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].
@@
 Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
-Some studies showcased results using a variety of such specializations [@CampilloGimenez2012a; @Ng2015; @Zhang2018a].
+Some studies showcased results using a variety of such specializations [@CampilloGimenez2013; @Ng2015; @Zhang2018].
 However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.

### 2024-03-11 scr/synthesis.r

@@ -31,7 +31,8 @@ framework_specializations <- googlesheets4::read_sheet(
 )
 
 # add citation numbers from most recent LaTeX compilation
-here::here("docs/Manuscript/Manuscript.tex") %>%
+# here::here("docs/Manuscript/Manuscript.tex") %>%
+here::here("docs/Localized-Modeling/Localized-Modeling.tex") %>%
   read_lines() %>%
   enframe(name = NULL, value = "line") %>%
   filter(str_detect(line, "\\{ref-")) %>%
@@ -113,10 +114,11 @@ write_rds(citation_id, here::here("data/citations.rds"))
 
 # table of evaluations and comparisons
 eval_comp_incl <- c(
-  "Wyns2004a", "Song2006c", "Elter2007a", "Xu2008a", "Kasabov2010",
-  "Liang2015a", "CampilloGimenez2012a", "Malykh2018a", "Zhang2018a",
-  "Nicolas2014a", "Ma2020a",# "Tang2021c",
-  "Park2006", "Lowsky2013", "Ng2015", "Lee2015", "Lee2017", "Wang2019"
+  "Wyns2004", "Song2006", "Elter2007", "Xu2008", "Kasabov2010",
+  "Liang2015", "CampilloGimenez2013", "Malykh2018", "Zhang2018",
+  "Nicolas2014", "Ma2020",# "Tang2021",
+  "Park2006", "Lowsky2013", "Ng2015", "Lee2015", "Lee2017", "Wang2019",
+  "Doborjeh2022", "Liu2022"
 )
 properties_inclusion %>%
   rename_with(snakecase::to_snake_case) %>%

### 2024-03-11 docs/Localized-Modeling/Localized-Modeling.Rmd
update prisma flowchart + begin adding text revisions

@@
   Background:
-  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of similar labeled cases.
+  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of \hl{labeled cases matched by similarity to a target case}.
 We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
@@
   Methods:
-  We searched four bibliographic platforms during 2021 July 19–22.
+  We searched four bibliographic platforms during 2021 July 19–22\hl{, updated 2024 January 24}.
 We set four eligibility criteria to identify applications of localized models to clinical and health tasks.
@@
   Results:
-  Of 1,309 search results, 328 were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
-25 were included, published 1997–2021.
+  Of \hl{1,657} search results, \hl{360} were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
+\hl{27} were included, published 1997–\hl{2022}.
 The specificity of search terms was poor, and inter-rater reliability was low.
@@
 Several addressed memory and runtime costs.
-A general technique that specializes to most of those reviewed was straightforward to describe.
+A general technique that specializes to most of those reviewed \hl{involved matching, retrieval, fitting, and evaluation steps that could optionally be supervised, optimized, or recursively performed}.
 
   Conclusions:
-  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, provided computational hurdles are overcome, and they follow a consistent pattern that could be implemented in a general-purpose way.
-We hope that our review will spur future work on efficiency, reproducibility, and user needs.
+  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, \hl{but rigorous comparisons to competing methods must be conducted and} computational hurdles \hl{must be} \hl{overcome.} We hope that our review will spur future work on efficiency, reproducibility, and user needs.
 
@@
   \newcommand{\landscapeend}{\end{landscape}}
+
+  \usepackage{soul}
+  \definecolor{lightblue}{rgb}{.80,.90,1}
+  \sethlcolor{lightblue}
 
@@
 These advantages include their immediate accessibility through the institutional EHR, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
-Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, with the monicker "practice-based evidence", as an important complement to the paradigm of evidence-based practice.
+Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, \hl{as} "practice-based evidence", \hl{a} complement to the paradigm of evidence-based practice.
 
 Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools, including artificial intelligence (AI).
-Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and their interplay with respect to outcomes.
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and \hl{interplay with each other}.
 The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models.
@@
 The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
-While traditional CBR returns these retrieved cases to inform human decision-making, the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
-In contrast, localized modeling fits families of predictive models, most commonly generalized linear models (GLMs), to these retrieved similarity cohorts in order to generate predictions.
-Localized models thus provide a way to fit fully interpretable models to large population data in a way that is personalized to each case.
+\hl{Traditional} CBR returns these retrieved cases to inform human decision-making\hl{;} the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+In contrast, localized modeling fits families of predictive models, \hl{for example} generalized linear models (GLMs), \hl{to retrieved} similarity cohorts in order to generate predictions.
+Localized models thus provide \hl{an individualized way to fit fully interpretable models to large population data}.
 
@@
 This approach harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
+\hl{Several hurdles face the adoption of such an approach in practice, but its benefits must first be demonstrated.}
 Our motivations in this scoping review are to describe the settings in and problems with which this approach has been tasked, to attempt to organize them within a common framework, and to evaluate the promise they show toward achieving this or other needs in clinical informatics.
@@
 A similar dichotomy is commonly used to distinguish the performance criterion for the usefulness of predictive models from the interpretability criterion for their usability.
-In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort can be understood as an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a ML workflow.
+In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort \hl{is} an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a \hl{machine learning (ML)} workflow.
 Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
-Nevertheless, as some recent applications have begun to show, they can be used to draw inferences about the importance of different risk factors to specific individuals.
-While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the effect estimates of a GLM.
+Nevertheless, as some recent applications have \hl{shown}, they can be used to draw inferences about the importance of different risk factors to specific individuals.
+While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the \hl{coefficients} of a GLM.
 While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
-Provided the model family fitted to the cohorts is interpretable, the localized model as a whole inherits this property.
+Provided the model family fitted to the cohorts is interpretable, the localized model \hl{inherits} this property.
 Thus, localized models may embody both styles of CBR.
+
+\hl{TODO: Discuss implementation; add review citations to next paragraph.}
 
@@
 #|  the structured search (n), and
-#|  citation/reference tracking (s)."
+#|  citation/reference tracking (s).
+#|  \\hl{Within the structured search results,
+#|  summands correspond to original and update.}"
 knitr::include_graphics("Fig1.pdf")

### 2024-03-11 scr/submission.r

@@ -1,7 +1,7 @@
 
 # copy figures to _AI Review_ folder
 file.copy(
-  here::here("fig/fig-prisma.pdf"),
+  here::here("fig/fig-prisma-update.pdf"),
   here::here("docs/Localized-Modeling/Fig1.pdf"),
   overwrite = TRUE
 )

### 2024-03-12 docs/Localized-Modeling/Localized-Modeling.Rmd
transcribe more edits

@@
 
-Our goals in this review are (1) to describe applications to date of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
+Our goals in this review are (1) to describe \hl{applications} of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
 Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
@@
 
-We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family.
-Such a general method would be straightforward, if laborious, to implement, and specializations of it would account for the majority of real-world applications.
-A general implementation based on this method would allow researchers to expedite every step of the analysis process, including selection, optimization, and evaluation, and enable sensitivity, robustness, and multiverse analyses that help identify the most consequential choices along the way.
+We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family\hl{.
+}A general implementation based on this method would allow researchers to expedite every step of the analysis process, including \hl{retrieval}, optimization, and evaluation, and enable sensitivity, robustness, and multiverse analyses that help identify the most consequential choices along the way.
 
@@
 
+\hl{We repeated the search on 2024 January 24 on the PubMed and Web of Science databases to bring the results up to date.}
+
 ## Title/abstract screen
@@
 Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, entries were included.
+\hl{For the update, on author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.}
 
@@
 Finally, one author (JCB) applied the same review process to the seed set of 43 studies that inspired the review.
+\hl{The same author later reviewed the updated results.}
 
@@
 
-Following the selection sources aligning with our review, we collected characteristics of included studies.
+\hl{We next} collected characteristics of included studies.
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, performance measures, results of evaluations and comparisons, name given to modeling approach).
@@
 
-Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common modeling technique.
+Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common \hl{family of techniques}.
 The goal will be to describe a unified framework for localized models that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.
@@
 
+## \hl{Update} {-}
+
+\hl{
+Of the 25 studies included, 4 showed up in no database searches (but included from the seed set), 1 was obtained via reference tracking, and, of the remaining 20 found through the database searches, only 1 was not found in PubMed or Web of Science.
+That one was instead found in Mathematical Reviews.
+Since MR returns about the same volume of results as WoS and PM, we omitted it from the update.
+
+We applied the search terms exactly as before but restricted the dates to 2021 July 19 (the beginning of the date range for our original searches) or later.
+We omitted the WoS search that returned impractically many results the first time around and was omitted then.
+
+We ran the update search on 2024 January 24.
+
+The PubMed update returned 97 results. The Web of Science refresher returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so that 348 studies were slated for screening.
+
+Screening by title and abstract was done in two waves.
+First, PM excluded only on account of a study not being (a) clinical/medical, (b) English, and (c) original.
+Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.
+}
+
 ## Analysis and synthesis {-}

### 2024-03-12 docs/Localized-Modeling/Localized-Modeling.Rmd
continue updates

@@
 Figure \ref{fig:prisma} depicts our identification of studies via databases and registers.
-Following the completion and input of each search string, there were a total of 1,422 sources within all of the platforms used.
-De-duplication resulted in 1,309 entries, which were added to the title/abstract screening for review.
-Of these, 328 entries met the screening criteria and were assessed through full text review.
-Of these, 44 fit the original criteria, and 19 were included as distinct from NN prediction.
+Following the completion and input of each search string, there were a total of \hl{1,817} sources within all of the platforms used.
+De-duplication resulted in \hl{1,657} entries, which were added to the title/abstract screening for review.
+Of these, \hl{360} entries met the screening criteria and were assessed through full text review.
+Of these, \hl{51} fit the original criteria, and \hl{21} were included as distinct from NN prediction.
 
@@
 
-There were 60 disagreements over inclusion versus exclusion.
+From the first search results, there were 60 disagreements over inclusion versus exclusion.
 Inter-rater reliability was low, at 82% relative to a 72% probability of random agreement.
+Only one author was available to review full texts from the update search.
 
@@
 After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using NN prediction.
-Finally, reference tracking from the 43 + 328 = 371 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
-This left us with 19 + 1 + 5 = 25 studies included in the review and synthesis.
+\hl{Reference} tracking from the \hl{43 + 360 = 403} studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
+This left us with \hl{21 + 1 + 5 = 27} studies included in the review and synthesis.
+
+```{r, echo=FALSE}
+stopifnot(nrow(tab_synthesis) == 27L)
+```
 
@@
 Another interesting pattern lay in the years of publication (Figure \ref{fig:year}).
-While they trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, methodological approach.
-We note that, while CBR in health and medicine originated as early as 1990, these studies did emerge early on once CBR had established itself.
-So the idea has been "in the air" for as long as CBR has been in use.
+While \hl{these studies} trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, \hl{approach}.
+We note \hl{that} CBR in health and medicine originated as early as 1990, \hl{and localized modeling emerged soon after} CBR had established itself\hl{;} the idea has been "in the air" for as long as CBR has been in use.
 
@@
 We categorized studies as "knowledge" or "practice" based on whether their aim was to produce generalizable knowledge or to improve practice.
-We classified 7 of the 25 studies as "knowledge", making the majority of studies from a "practice" standpoint.
+We classified 8 of the studies as "knowledge", making the majority of studies from a "practice" standpoint.
 These studies aimed to provide tools for use in clinic or to improve outcomes.
 
-Lastly, we observed a commonality among 6 of the 25 included studies in having evaluated methods using leave-one-out cross-validation (LOOCV).
+Lastly, we observed a commonality among 6 of the included studies in having evaluated methods using leave-one-out cross-validation (LOOCV).
 The purpose of this measure is to estimate the overall performance of certain factors when used to make predictions, particularly utilized on smaller data sets, where models benefit greatly from larger training sets and additional model fitting is less costly.
@@
 In addition to the aim of its analysis, we coded several aspects of the design of each study, including the source and type of data and the clinical task the model performed (Table \ref{tab:synthesis}), as well as the specific methodology and terminology adopted (Table \ref{tab:composite}).
-In the following subsections, we take a closer look at these design elements.
-
-\small
-
-```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
-composite_studies |> 
-  transmute(
-    # Citation = str_c("@", Citation),
-    Citation = sapply(
-      str_split(Citation, "; "),
-      function(s) str_c("@", s, collapse = "; ")
-    ),
-    Elements = str_replace_all(Elements, ",", ";"),
-    Terminology
-  ) |> 
-  mutate(across(
-    everything(),
-    # ~ str_replace_all(., "<U+001B>", "")
-    textclean::replace_non_ascii
-  )) |>
-  # mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
-  knitr::kable(
-    align = "lllllrr",
-    caption = paste0(
-      "\\label{tab:composite}",
-      "Methodological elements of studies included in the synthesis."
-    )
-  )
-```
-
-\normalsize
+In the following subsections, we take a closer look at these design elements \hl{and their reported justifications and limitations}.
 
@@
 @Vilhena2016 developed a CBR cycle around a clustering-informed similarity matching procedure and an artificial neural network–based classifier to identify thrombophilia patients at high risk of thrombotic episodes.
-And @Ma2020 took a similarity cohort–based approach to predicting length of stay for ICU patients.
-Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
+@Ma2020 took a similarity cohort–based approach to predicting length of stay for ICU patients.
+@Doborjeh2022 \hl{localized spiking neural networks in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.} @Liu2022 \hl{incorporated global-to-local transfer learning into localized models of acute kidney injury risk in hospitalized patients.} Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
 
@@
 @Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
+@Liu2022 \hl{took this idea further, systematically comparing regression coefficients for specific risk factors across dozens of diagnostic subpopulations.}
+@Doborjeh2022 \hl{used localized time series models to identify environmental changes associated with increased stroke risk.}
 
@@
 
-CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity; @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
+CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity\hl{:} @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
 This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle.
-Several studies addressed these challenges: @Park2006, while excluded from the synthesis, were the earliest to propose that cohorts be bounded by a similarity threshold rather than by a number of cases, and this improved performance in their experiments.
+Several studies addressed these challenges: @Park2006, while excluded from the synthesis, were the earliest to propose that cohorts be bounded by a similarity threshold rather than by \hl{the} number of cases, \hl{which} improved performance in their experiments.
 @CampilloGimenez2013 leveraged predictor weights obtained from logistic regression to inform the similarity calculation used in retrieval.
@@
 @Ma2020 proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021 and @Tang2021 partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
+@Liu2022 \hl{used transfer learning from globally-fitted models to improve the efficiency of localized models.}
 
@@
 Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
-The frequency plot in Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
-
-```{r methods, fig.align='center', out.width="100%", echo=FALSE}
-#| fig.cap =
-#| "Frequency of recurring methodological elements across included studies."
-knitr::include_graphics("Fig4.png")
-```
-
-## Performance assessments
+Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
+
+## Evaluations
 
@@
 Note that we exercised some judgment in classifying methods as proposals and comparators, as in some cases all methods were original but only some showcased main ideas.
-It would be impractical to meta-analyze these numbers due to the great variety of settings, problems, data types, and choices involved.
-
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators (Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]) are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
-Though @CampilloGimenez2013 and [@Ng2015] report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
-Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model (statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]), they did not consistently outperform comparators.
+It would be impractical to meta-analyze these numbers due to the great variety of settings, problems, data types\hl{, techniques}, and choices involved.
+
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators\hl{---}Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]\hl{---}are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
+Though @CampilloGimenez2013 and @Ng2015 report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
+Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model\hl{---}statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]\hl{---}they did not consistently outperform comparators.
 
@@
 
-Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), laboratory (2), and image (1) data.
+Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), \hl{laboratory (3), image (1), and environmental (1)} data.
 We suggest three reasons for this: First, this literature traces back to the 1990s, before -omic data could be generated cost-effectively at scale.
@@
 
-These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
-We find uses of "targeted", "personalized", and "precision" generic and imprecise to this approach, while transductive inference is an established term for a broader set of methods in ML.
+These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
+We find uses of "targeted", "personalized", and "precision" generic and \hl{imprecise,} while transductive inference is an established term for a broader set of methods in ML.
 Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^local]
@@
 
-The 25 studies included in our synthesis used composite techniques that we organized into two major types.
+The 27 studies included in our synthesis used composite techniques that we organized into two major types.
 One type, _similarity learning_, was used in 8 studies that tie in to the much larger literature on DML [@Yang2006].
@@
 These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
-Seven studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
+\hl{Eight} studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
 Based on the key properties of DML algorithms identified by @Bellet2014, most learned measures were non-linear while some were locally learned, and some were optimized globally while others locally.
@@
 
-In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, five other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, and @Ma2020.
-In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did some of the same authors [@Tang2021; @Ng2021] later.
+In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, \hl{seven} other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, @Ma2020, @Liu2022, and @Doborjeh2022.
+In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, \hl{as did} @Tang2021, @Ng2021, and @Liu2022 later.
 The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016] or via cross-validation [@Lowsky2013; @Verma2015].
@@
 Foremost was the optimization of retrieved cohort sizes for some measure of model performance.
-Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study @Mariuzzi1997 and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach.
+Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study [@Mariuzzi1997] and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach\hl{, though} @Liu2022 \hl{took a third option by retrieving fixed proportions of training data}.
 Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.
@@
 With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
-This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021].
+This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021]\hl{, and it was later used by} @Liu2022 \hl{to show a dependency of predictor importance on the patient disease subgroup}.
 These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].
@@
 Meanwhile, each step can be optimized in a ML fashion by having its parameters tuned to improve performance.
-We found no studies that tuned the calculation of similarity, though most uniform cohort sizes were tuned.
+We found \hl{one study} that tuned the calculation of similarity [@Liu2022], though most uniform cohort sizes were also tuned.
 Some earlier studies tuned generalized regression models, but more recent studies did not.
@@
 
-## Directions and expectations
+## Directions and expectations for future work
+
+###
 
@@
 
+###
+
 While most studies identified outstanding technical needs (Section \ref{identified-needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
@@
 One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
+
+###
 
@@
 
-## \hl{Update} {-}
-
-\hl{
-Of the 25 studies included, 4 showed up in no database searches (but included from the seed set), 1 was obtained via reference tracking, and, of the remaining 20 found through the database searches, only 1 was not found in PubMed or Web of Science.
+## Update {-}
+
+\hl{Of the 25 studies included, 4 showed up in no database searches (but included from the seed set), 1 was obtained via reference tracking, and, of the remaining 20 found through the database searches, only 1 was not found in PubMed or Web of Science.
 That one was instead found in Mathematical Reviews.
-Since MR returns about the same volume of results as WoS and PM, we omitted it from the update.
-
-We applied the search terms exactly as before but restricted the dates to 2021 July 19 (the beginning of the date range for our original searches) or later.
-We omitted the WoS search that returned impractically many results the first time around and was omitted then.
-
-We ran the update search on 2024 January 24.
-
-The PubMed update returned 97 results. The Web of Science refresher returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so that 348 studies were slated for screening.
-
-Screening by title and abstract was done in two waves.
+Since MR returns about the same volume of results as WoS and PM, we omitted it from the update.}
+
+\hl{We applied the search terms exactly as before but restricted the dates to 2021 July 19 (the beginning of the date range for our original searches) or later.
+We omitted the WoS search that returned impractically many results the first time around and was omitted then.}
+
+\hl{We ran the update search on 2024 January 24.}
+
+\hl{The PubMed update returned 97 results. The Web of Science refresher returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so that 348 studies were slated for screening.}
+
+\hl{Screening by title and abstract was done in two waves.
 First, PM excluded only on account of a study not being (a) clinical/medical, (b) English, and (c) original.
-Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.
-}
+Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.}
 
@@
 Because we could not predict the scope of methodological approaches we would encounter, we did not prepare specific analyses or syntheses _a priori_.
+
+## Methodological elements {-}
+
+Table \ref{tab:composite} and Figure \ref{fig:methods} summarize the techniques and terminology used by the included studies, as referred to in the main text.
+
+\small
+
+```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
+composite_studies |> 
+  transmute(
+    # Citation = str_c("@", Citation),
+    Citation = sapply(
+      str_split(Citation, "; "),
+      function(s) str_c("@", s, collapse = "; ")
+    ),
+    Elements = str_replace_all(Elements, ",", ";"),
+    Terminology
+  ) |> 
+  mutate(across(
+    everything(),
+    # ~ str_replace_all(., "<U+001B>", "")
+    textclean::replace_non_ascii
+  )) |>
+  # mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
+  knitr::kable(
+    align = "lllllrr",
+    caption = paste0(
+      "\\label{tab:composite}",
+      "Methodological elements of studies included in the synthesis."
+    )
+  )
+```
+
+\normalsize
+
+```{r methods, fig.align='center', out.width="100%", echo=FALSE}
+#| fig.cap =
+#| "Frequency of recurring methodological elements across included studies."
+knitr::include_graphics("Fig4.png")
+```
 

### 2024-03-13 docs/Localized-Modeling/Localized-Modeling.Rmd
finish transcribing revisions

@@
 This approach harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
-\hl{Several hurdles face the adoption of such an approach in practice, but its benefits must first be demonstrated.}
+\hl{Several hurdles face the deployment of such an approach in practice, but its benefits must first be demonstrated.}
 Our motivations in this scoping review are to describe the settings in and problems with which this approach has been tasked, to attempt to organize them within a common framework, and to evaluate the promise they show toward achieving this or other needs in clinical informatics.
@@
 
-\hl{TODO: Discuss implementation; add review citations to next paragraph.}
-
 We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
@@
 Our goals in this review are (1) to describe \hl{applications} of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
+\hl{We focus narrowly on localized models, rather than broadly on patient similarity--based clinical decision support tools, so that we may thoroughly assess their value in terms of reported evaluations and comparisons to other methods.
+This also allows us to devise a high-level framework that specializes to the majority of these methods, which we then use to frame our discussion and recommendations.}
+
 Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
@@
 @Lopez2011 implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
-Several other teams also set out to develop  more interoperable clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
+Several other teams also set out to develop more \hl{generalizable and data-agnostic} clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
+\hl{However, more informatical and legal challenges to implementation, including interoperability of systems and data and regulations concerning security and privacy, were rarely addressed.
+We discuss this further in Section} \ref{feasibility}\hl{.}
+
 Figure \ref{fig:methods} compares the rates at which several methodological elements are invoked in the sample.
@@
 
-###
+### Performance
 
 As discussed in Section \ref{rationales}, most studies were premised on the potential predictive value of localized models.
-As cautioned in Section \ref{performance-assessments}, not all experiments affirmed this premise.
+As cautioned in Section \ref{evaluations}, not all experiments affirmed this premise.
 Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
-Some studies showcased results using a variety of such specializations [@CampilloGimenez2013; @Ng2015; @Zhang2018].
+Some studies showcased results using a variety of such specializations [@CampilloGimenez2013; @Ng2015; @Zhang2018; @Liu2022].
 However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
@@
 
-###
+\hl{
+The value of this work would be enhanced by standardization and modularization.
+First, studies using localized models should locate the proposed approach and any variations in a shared specification space.
+For this purpose, we propose the framework outlined above, at least as a starting point.
+When a specification invokes complex techniques, for example supervised DML or neural network classifiers, it should be compared at least to alternative specifications using simpler alternatives.
+}
+@CampilloGimenez2013 and @Liu2022 \hl{provide excellent examples in which all possible choices are toggled and yield a neatly nested comparison.
+Additionally, some state-of-the-art models, including others that obtain individualized predictions, should be included in comparisons.
+Finally, implementations should be not only published on public repositories but designed in such a way that users familiar with the underlying language can substitute a technique of their choosing for any part of the specification (similarity, retrieval, fitting, evaluation).
+We believe these principles will make reported results clearer and reproduction and extension easier.
+}
+
+### Interpretability
 
@@
 Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
+
 One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
-
-###
+\hl{
+This is a necessary precondition for practical use but would also be a valuable contribution to the experimental literature.
+Another would be to compare the localized predictor importance measures obtained from localized models to the local importance measures used to explain predictions made by opaque models.
+This would test both measures for concurrent validity, while differences between them would inform what settings or cases are poorly served by one or the other.
+}
+
+### Feasibility
+
+\hl{Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
+This need demands not only methodologies but also architectures that are robust, versatile, and compliant in the face of diverse data models and use restrictions, yet no studies in our sample addressed the challenges of system interoperability and regulatory requirements head-on.}
+
+\hl{CBR and, by extension, localized modeling are especially susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures as well.
+Individual systems exhibit many dimensions of incompleteness} [@Weiskopf2013] \hl{and vary along other dimensions of quality} [@Kohane2021]\hl{, and their aggregation for modeling purposes depends on several kinds of interoperability} [@Weber2015]\hl{.
+Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties} [@Haendel2021]\hl{.
+While most studies in our sample focused on improving care, increasingly many over time focused on generalizable knowledge.
+Therefore, as studies using localized modeling shift from proof of concept to feasibility, they will also need to demonstrate practical and regulatory feasibility.
+}
+
+### Customizability
 
@@
 For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
+@Lopez2011 \hl{took an important step in this direction with an adjustable and adaptable implementation.}
 Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.
@@
 Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth.
-Whereas precious few of the reviewed studies used these models, for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+Whereas precious few of the reviewed studies used these \hl{models for} any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
 
@@
 
-Table \ref{tab:composite} and Figure \ref{fig:methods} summarize the techniques and terminology used by the included studies, as referred to in the main text.
+\hl{Table} \ref{tab:composite} \hl{and Figure} \ref{fig:methods} \hl{summarize the techniques and terminology used by the included studies, as referred to in the main text.}
 
@@
 ## Performance evaluations and comparisons {-}
+
+\hl{Table} \ref{tab:performance} \hl{summarizes the evaluations and comparisons made of methods proposed in the included studies.
+In most cases, not all performance results are included; for the sake of concision, only what we judged to be the headline results, using the primary evaluation metrics, are reproduced here.
+The table contains no original results.}
 

### 2024-03-17 docs/Localized-Modeling/Localized-Modeling.Rmd
make corrections noticed by adi

@@
     degrees: BS
-    email: ac4569@my.nsu.edu
+    email: ac4569@mynsu.nova.edu
     affiliation: 1
@@
 Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, entries were included.
-\hl{For the update, on author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.}
+\hl{For the update, one author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.}
 

### 2024-03-19 docs/Localized-Modeling/Localized-Modeling.Rmd
update analysis & finalize manuscript for resubmission

@@
   \usepackage{soul}
-  \definecolor{lightblue}{rgb}{.80,.90,1}
-  \sethlcolor{lightblue}
+  %%\definecolor{hlcolor}{rgb}{.80,.90,1}
+  \definecolor{hlcolor}{rgb}{1,1,1}
+  \sethlcolor{hlcolor}
 
@@
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
-The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, and _PLOS One_ published 2.
+The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3\hl{, \emph{Evolving Systems} published 2}, and _PLOS One_ published 2.
 
@@
 #|  Dashed arrows indicate optional steps."
-knitr::include_graphics("Fig5.pdf")
+knitr::include_graphics("Fig4.pdf")
 ```
@@
 #| "Frequency of recurring methodological elements across included studies."
-knitr::include_graphics("Fig4.png")
+knitr::include_graphics("Fig5.png")
 ```

### 2024-03-19 scr/submission.r

@@ -16,13 +16,12 @@ file.copy(
   overwrite = TRUE
 )
 file.copy(
-  here::here("fig/fig-methods.png"),
-  here::here("docs/Localized-Modeling/Fig4.png"),
+  here::here("fig/fig-typology.pdf"),
+  here::here("docs/Localized-Modeling/Fig4.pdf"),
   overwrite = TRUE
 )
 file.copy(
-  here::here("fig/fig-typology.pdf"),
-  here::here("docs/Localized-Modeling/Fig5.pdf"),
+  here::here("fig/fig-methods.png"),
+  here::here("docs/Localized-Modeling/Fig5.png"),
   overwrite = TRUE
 )
-

### 2024-03-19 scr/synthesis.r

@@ -229,7 +229,8 @@ tab_synthesis %>%
   mutate(Source = factor(
     Source,
     c("Clinical", "Laboratory", "Imaging",
-      "Clinical, Laboratory", "Clinical, Imaging", "Patient-reported")
+      "Clinical, Laboratory", "Clinical, Imaging", "Clinical, Environmental",
+      "Patient-reported")
   )) %>%
   mutate(Task = factor(
     Task,

### 2024-06-17 docs/localized-modeling-medRxiv/localized-modeling-medRxiv.Rmd
prepare arxiv manuscript

--- 2024-03-13 docs/Localized-Modeling/Localized-Modeling.Rmd
@@
 ---
-# Supported options: 
-#   sn-nature:       Style for submissions to Nature Portfolio journals
-#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
-#   sn-mathphys:     Math and Physical Sciences Reference Style
-#   sn-aps:          American Physical Society (APS) Reference Style
-#   sn-vancouver:    Vancouver Reference Style
-#   sn-apa:          APA Reference Style 
-#   sn-chicago:      Chicago-based Humanities Reference Style
-#   default:         Default
-
-classoptions: 
-  - sn-nature      
-  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
-  # - referee       # Optional: Use double line spacing 
-  # - lineno        # Optional: Add line numbers
-  # - iicol         # Optional: Double column layour
-
 title: "Clinical prediction with localized modeling using similarity-based cohorts: A scoping review"
-titlerunning: Clinical Prediction with Localized Modeling
 authors: 
-  - firstname: Adi
-    lastname: Cohen
-    degrees: BS
-    email: ac4569@my.nsu.edu
-    affiliation: 1
-  - firstname: Patti
-    lastname: McCall-Junkin
-    degrees: MA, MLS
+  - name: Adi Cohen
+    email: ac4569@mynsu.nova.edu
+    affiliation: College of Osteopathic Medicine, Nova Southeastern University
+    location: Davie, FL
+  - name: Patti McCall-Junkin
     email: pattimccall.junkin@gmail.com
-    affiliation: 2
-  - firstname: Jason Cory
-    lastname: Brunson
-    degrees: PhD
+    affiliation: Smathers Libraries Academic Research & Consulting Services, University of Florida
+    location: Gainesville, FL
+  - name: Jason Cory Brunson
     email: jason.brunson@medicine.ufl.edu
-    orcid: 0000-0003-3126-9494
-    affiliation: 3
-    corresponding: TRUE
-affiliations:
-  - number: 1
-    info:
-      orgdiv: College of Osteopathic Medicine
-      orgname: Nova Southeastern University
-    address:
-        street: 3200 S University Drive
-        city: Davie
-        postcode: 33328
-        state: FL
-        country: USA
-  - number: 2
-    info:
-      orgdiv: Smathers Libraries Academic Research & Consulting Services
-      orgname: University of Florida
-    address:
-        street: P.O. Box 117000
-        city: Gainesville
-        postcode: 32611-7000
-        state: FL
-        country: USA
-  - number: 3
-    corresponding: TRUE
-    info:
-      orgdiv: Laboratory for Systems Medicine
-      orgname: University of Florida
-    address:
-        street: P.O. Box 100225 JHMHC
-        city: Gainesville
-        postcode: 32610-0225
-        state: FL
-        country: USA
+    affiliation: Laboratory for Systems Medicine, University of Florida
+    location: Gainesville, FL
+abstract: |
+  Background:
+  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of labeled cases matched by similarity to a target case.
+We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
+
+  Methods:
+  We searched four bibliographic platforms during 2021 July 19–22, updated 2024 January 24.
+We set four eligibility criteria to identify applications of localized models to clinical and health tasks.
+Two authors divided title/abstract screening and reviewed screened entries for inclusion.
+We discussed settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
+
+  Results:
+  Of 1,657 search results, 360 were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
+27 were included, published 1997–2022.
+The specificity of search terms was poor, and inter-rater reliability was low.
+Almost all models were predictive, the most common tasks being prognosis and diagnosis.
+Most studies used clinical, occasionally laboratory and image, data.
+Several addressed memory and runtime costs.
+A general technique that specializes to most of those reviewed involved matching, retrieval, fitting, and evaluation steps that could optionally be supervised, optimized, or recursively performed.
+
+  Conclusions:
+  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, but rigorous comparisons to competing methods must be conducted and computational hurdles must be overcome.
+We hope that our review will spur future work on efficiency, reproducibility, and user needs.
+
 keywords:
@@
   - scoping review
-# pacs: 
-#   jel:
-#     - "D8"
-#     - "H51"
-#   msc:
-#     - "35A01"
-#     - "65L10"
-
-# Sample for structured abstract
-abstract: |
-  Background:
-  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of \hl{labeled cases matched by similarity to a target case}.
-We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
-
-  Methods:
-  We searched four bibliographic platforms during 2021 July 19–22\hl{, updated 2024 January 24}.
-We set four eligibility criteria to identify applications of localized models to clinical and health tasks.
-Two authors divided title/abstract screening and reviewed screened entries for inclusion.
-We discussed settings, tasks, and tools; identified and tabulated themes; and synthesized the methods into a general framework.
-
-  Results:
-  Of \hl{1,657} search results, \hl{360} were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
-\hl{27} were included, published 1997–\hl{2022}.
-The specificity of search terms was poor, and inter-rater reliability was low.
-Almost all models were predictive, the most common tasks being prognosis and diagnosis.
-Most studies used clinical, occasionally laboratory and image, data.
-Several addressed memory and runtime costs.
-A general technique that specializes to most of those reviewed \hl{involved matching, retrieval, fitting, and evaluation steps that could optionally be supervised, optimized, or recursively performed}.
-
-  Conclusions:
-  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, \hl{but rigorous comparisons to competing methods must be conducted and} computational hurdles \hl{must be} \hl{overcome.} We hope that our review will spur future work on efficiency, reproducibility, and user needs.
-
 bibliography:
@@
   - SR-5-Screened.bib
-
+biblio-style: unsrt
+output:
+  rticles::arxiv_article:
+    keep_tex: true
+    latex_engine: xelatex
+    # citation_package: natbib
+    citation_package: "default"
 header-includes: |
@@
   \newcommand{\landscapeend}{\end{landscape}}
-
-  \usepackage{soul}
-  \definecolor{lightblue}{rgb}{.80,.90,1}
-  \sethlcolor{lightblue}
-
-  %% Per the spinger doc, new theorem styles can be included using built in style, 
-  %% but it seems the don't work so commented below
-  %\theoremstyle{thmstyleone}%
-  \newtheorem{theorem}{Theorem}%  meant for continuous numbers
-  %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
-  %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
-  \newtheorem{proposition}[theorem]{Proposition}%
-  %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
-
-  %% \theoremstyle{thmstyletwo}%
-  \theoremstyle{remark}
-  \newtheorem{example}{Example}%
-  \newtheorem{remark}{Remark}%
-
-  %% \theoremstyle{thmstylethree}%
-  \theoremstyle{definition}
-  \newtheorem{definition}{Definition}%
-
-output:
-  rticles::springer_article:
-    keep_tex: true
-    latex_engine: xelatex
-    # citation_package: natbib
-    citation_package: "default"
 ---
@@
 These advantages include their immediate accessibility through the institutional EHR, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
-Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, \hl{as} "practice-based evidence", \hl{a} complement to the paradigm of evidence-based practice.
+Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of evidence-based practice.
 
 Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools, including artificial intelligence (AI).
-Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and \hl{interplay with each other}.
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other.
 The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models.
@@
 The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
-\hl{Traditional} CBR returns these retrieved cases to inform human decision-making\hl{;} the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
-In contrast, localized modeling fits families of predictive models, \hl{for example} generalized linear models (GLMs), \hl{to retrieved} similarity cohorts in order to generate predictions.
-Localized models thus provide \hl{an individualized way to fit fully interpretable models to large population data}.
+Traditional CBR returns these retrieved cases to inform human decision-making; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+In contrast, localized modeling fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
+Localized models thus provide an individualized way to fit fully interpretable models to large population data.
 
@@
 This approach harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
-\hl{Several hurdles face the deployment of such an approach in practice, but its benefits must first be demonstrated.}
+Several hurdles face the deployment of such an approach in practice, but its benefits must first be demonstrated.
 Our motivations in this scoping review are to describe the settings in and problems with which this approach has been tasked, to attempt to organize them within a common framework, and to evaluate the promise they show toward achieving this or other needs in clinical informatics.
@@
 A similar dichotomy is commonly used to distinguish the performance criterion for the usefulness of predictive models from the interpretability criterion for their usability.
-In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort \hl{is} an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a \hl{machine learning (ML)} workflow.
+In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
 Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
-Nevertheless, as some recent applications have \hl{shown}, they can be used to draw inferences about the importance of different risk factors to specific individuals.
-While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the \hl{coefficients} of a GLM.
+Nevertheless, as some recent applications have shown, they can be used to draw inferences about the importance of different risk factors to specific individuals.
+While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the coefficients of a GLM.
 While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
-Provided the model family fitted to the cohorts is interpretable, the localized model \hl{inherits} this property.
+Provided the model family fitted to the cohorts is interpretable, the localized model inherits this property.
 Thus, localized models may embody both styles of CBR.
@@
 
-Our goals in this review are (1) to describe \hl{applications} of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
-\hl{We focus narrowly on localized models, rather than broadly on patient similarity--based clinical decision support tools, so that we may thoroughly assess their value in terms of reported evaluations and comparisons to other methods.
-This also allows us to devise a high-level framework that specializes to the majority of these methods, which we then use to frame our discussion and recommendations.}
+Our goals in this review are (1) to describe applications of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
+We focus narrowly on localized models, rather than broadly on patient similarity--based clinical decision support tools, so that we may thoroughly assess their value in terms of reported evaluations and comparisons to other methods.
+This also allows us to devise a high-level framework that specializes to the majority of these methods, which we then use to frame our discussion and recommendations.
 
@@
 
-We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family\hl{.
-}A general implementation based on this method would allow researchers to expedite every step of the analysis process, including \hl{retrieval}, optimization, and evaluation, and enable sensitivity, robustness, and multiverse analyses that help identify the most consequential choices along the way.
+We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family.
+A general implementation based on this method would allow researchers to expedite every step of the analysis process, including retrieval, optimization, and evaluation, and enable sensitivity, robustness, and multiverse analyses that help identify the most consequential choices along the way.
 
@@
 
-\hl{We repeated the search on 2024 January 24 on the PubMed and Web of Science databases to bring the results up to date.}
+We repeated the search on 2024 January 24 on the PubMed and Web of Science databases to bring the results up to date.
 
@@
 Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, entries were included.
-\hl{For the update, on author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.}
+For the update, one author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.
 
@@
 Finally, one author (JCB) applied the same review process to the seed set of 43 studies that inspired the review.
-\hl{The same author later reviewed the updated results.}
+The same author later reviewed the updated results.
 
@@
 
-\hl{We next} collected characteristics of included studies.
+We next collected characteristics of included studies.
 The features included bibliographic fields (date of publication, journal, authors, title, keywords, DOI), study goals (objective, generalizable knowledge, evaluation, clinical/medical domain), data sets (data source, type of data, range of data, number of cases/incidences, number of predictors/features), and methodological choices (types of similarity measure, families of adaptation step/statistical model, method(s) compared against, performance measures, results of evaluations and comparisons, name given to modeling approach).
@@
 
-Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common \hl{family of techniques}.
+Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common family of techniques.
 The goal will be to describe a unified framework for localized models that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.
@@
 Figure \ref{fig:prisma} depicts our identification of studies via databases and registers.
-Following the completion and input of each search string, there were a total of \hl{1,817} sources within all of the platforms used.
-De-duplication resulted in \hl{1,657} entries, which were added to the title/abstract screening for review.
-Of these, \hl{360} entries met the screening criteria and were assessed through full text review.
-Of these, \hl{51} fit the original criteria, and \hl{21} were included as distinct from NN prediction.
+Following the completion and input of each search string, there were a total of 1,817 sources within all of the platforms used.
+De-duplication resulted in 1,657 entries, which were added to the title/abstract screening for review.
+Of these, 360 entries met the screening criteria and were assessed through full text review.
+Of these, 51 fit the original criteria, and 21 were included as distinct from NN prediction.
 
@@
 #|  citation/reference tracking (s).
-#|  \\hl{Within the structured search results,
-#|  summands correspond to original and update.}"
+#|  Within the structured search results,
+#|  summands correspond to original and update."
 knitr::include_graphics("Fig1.pdf")
@@
 After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using NN prediction.
-\hl{Reference} tracking from the \hl{43 + 360 = 403} studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
-This left us with \hl{21 + 1 + 5 = 27} studies included in the review and synthesis.
+Reference tracking from the 43 + 360 = 403 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
+This left us with 21 + 1 + 5 = 27 studies included in the review and synthesis.
 
@@
 The studies were published in a variety of journals, with some of greater frequency, though no single journal published more than 3.
-The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, and _PLOS One_ published 2.
+The journal _Evolving Systems_ published 2 of the included studies, _Artificial Intelligence in Medicine_ published 3, _Hindawi - Journal of Healthcare Engineering_ published 3, \emph{Evolving Systems} published 2, and _PLOS One_ published 2.
 
 Another interesting pattern lay in the years of publication (Figure \ref{fig:year}).
-While \hl{these studies} trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, \hl{approach}.
-We note \hl{that} CBR in health and medicine originated as early as 1990, \hl{and localized modeling emerged soon after} CBR had established itself\hl{;} the idea has been "in the air" for as long as CBR has been in use.
+While these studies trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, approach.
+We note that CBR in health and medicine originated as early as 1990, and localized modeling emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
 
@@
 In addition to the aim of its analysis, we coded several aspects of the design of each study, including the source and type of data and the clinical task the model performed (Table \ref{tab:synthesis}), as well as the specific methodology and terminology adopted (Table \ref{tab:composite}).
-In the following subsections, we take a closer look at these design elements \hl{and their reported justifications and limitations}.
+In the following subsections, we take a closer look at these design elements and their reported justifications and limitations.
 
@@
 @Ma2020 took a similarity cohort–based approach to predicting length of stay for ICU patients.
-@Doborjeh2022 \hl{localized spiking neural networks in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.} @Liu2022 \hl{incorporated global-to-local transfer learning into localized models of acute kidney injury risk in hospitalized patients.} Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
+@Doborjeh2022 localized spiking neural networks in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.
+@Liu2022 incorporated global-to-local transfer learning into localized models of acute kidney injury risk in hospitalized patients.
+Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
 
@@
 @Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
-@Liu2022 \hl{took this idea further, systematically comparing regression coefficients for specific risk factors across dozens of diagnostic subpopulations.}
-@Doborjeh2022 \hl{used localized time series models to identify environmental changes associated with increased stroke risk.}
+@Liu2022 took this idea further, systematically comparing regression coefficients for specific risk factors across dozens of diagnostic subpopulations.
+@Doborjeh2022 used localized time series models to identify environmental changes associated with increased stroke risk.
 
@@
 
-CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity\hl{:} @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
+CBR can be understood as the opposite side of a trade-off with rule-based reasoning between model size and model complexity: @Doyle2006 describe their approach as "knowledge-light", in that "the cases do not contain explicit explanation structures; instead, explanation is achieved by comparison of the query case with retrieved cases".
 This means that the greatest performance and efficiency challenges in CBR have to do with the retrieval and revision phases in the R4 cycle.
-Several studies addressed these challenges: @Park2006, while excluded from the synthesis, were the earliest to propose that cohorts be bounded by a similarity threshold rather than by \hl{the} number of cases, \hl{which} improved performance in their experiments.
+Several studies addressed these challenges: @Park2006, while excluded from the synthesis, were the earliest to propose that cohorts be bounded by a similarity threshold rather than by the number of cases, which improved performance in their experiments.
 @CampilloGimenez2013 leveraged predictor weights obtained from logistic regression to inform the similarity calculation used in retrieval.
@@
 @Ma2020 proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021 and @Tang2021 partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
-@Liu2022 \hl{used transfer learning from globally-fitted models to improve the efficiency of localized models.}
+@Liu2022 used transfer learning from globally-fitted models to improve the efficiency of localized models.
 
@@
 @Lopez2011 implemented a comprehensive CBR tool in response to the lack of general-purpose software, to enable coupling with other tools as well as to expedite development, experimentation, and uptake.
-Several other teams also set out to develop more \hl{generalizable and data-agnostic} clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
-\hl{However, more informatical and legal challenges to implementation, including interoperability of systems and data and regulations concerning security and privacy, were rarely addressed.
-We discuss this further in Section} \ref{feasibility}\hl{.}
+Several other teams also set out to develop more generalizable and data-agnostic clinical support tools [@Elter2007; @Liang2015; @Ng2015; @Zhang2018].
+However, more informatical and legal challenges to implementation, including interoperability of systems and data and regulations concerning security and privacy, were rarely addressed.
+We discuss this further in Section \ref{feasibility}.
 
@@
 Note that we exercised some judgment in classifying methods as proposals and comparators, as in some cases all methods were original but only some showcased main ideas.
-It would be impractical to meta-analyze these numbers due to the great variety of settings, problems, data types\hl{, techniques}, and choices involved.
-
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators\hl{---}Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]\hl{---}are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
+It would be impractical to meta-analyze these numbers due to the great variety of settings, problems, data types, techniques, and choices involved.
+
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators---Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]---are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
 Though @CampilloGimenez2013 and @Ng2015 report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
-Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model\hl{---}statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]\hl{---}they did not consistently outperform comparators.
+Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model---statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]---they did not consistently outperform comparators.
 
@@
 
-Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), \hl{laboratory (3), image (1), and environmental (1)} data.
+Consistent with their orientation, almost all included studies were conducted using clinical data, only occasionally together with patient-reported (2), laboratory (3), image (1), and environmental (1) data.
 We suggest three reasons for this: First, this literature traces back to the 1990s, before -omic data could be generated cost-effectively at scale.
@@
 These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
-We find uses of "targeted", "personalized", and "precision" generic and \hl{imprecise,} while transductive inference is an established term for a broader set of methods in ML.
+We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
 Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^local]
@@
 These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
-\hl{Eight} studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
+Eight studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
 Based on the key properties of DML algorithms identified by @Bellet2014, most learned measures were non-linear while some were locally learned, and some were optimized globally while others locally.
@@
 
-In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, \hl{seven} other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, @Ma2020, @Liu2022, and @Doborjeh2022.
-In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, \hl{as did} @Tang2021, @Ng2021, and @Liu2022 later.
+In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, seven other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, @Ma2020, @Liu2022, and @Doborjeh2022.
+In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did @Tang2021, @Ng2021, and @Liu2022 later.
 The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016] or via cross-validation [@Lowsky2013; @Verma2015].
@@
 Foremost was the optimization of retrieved cohort sizes for some measure of model performance.
-Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study [@Mariuzzi1997] and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach\hl{, though} @Liu2022 \hl{took a third option by retrieving fixed proportions of training data}.
+Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study [@Mariuzzi1997] and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach, though @Liu2022 took a third option by retrieving fixed proportions of training data.
 Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.
@@
 With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
-This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021]\hl{, and it was later used by} @Liu2022 \hl{to show a dependency of predictor importance on the patient disease subgroup}.
+This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021], and it was later used by @Liu2022 to show a dependency of predictor importance on the patient disease subgroup.
 These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].
@@
 #|  Dashed arrows indicate optional steps."
-knitr::include_graphics("Fig5.pdf")
+knitr::include_graphics("Fig5.png")
 ```
@@
 Meanwhile, each step can be optimized in a ML fashion by having its parameters tuned to improve performance.
-We found \hl{one study} that tuned the calculation of similarity [@Liu2022], though most uniform cohort sizes were also tuned.
+We found one study that tuned the calculation of similarity [@Liu2022], though most uniform cohort sizes were also tuned.
 Some earlier studies tuned generalized regression models, but more recent studies did not.
@@
   ) |> 
-  mutate(Citation = str_c("[", Citation, "]")) |> 
+  # mutate(Citation = str_c("[", Citation, "]")) |> 
   # prevent numeric list columns from dominating the horizontal space
@@
 
-\hl{
 The value of this work would be enhanced by standardization and modularization.
@@
 When a specification invokes complex techniques, for example supervised DML or neural network classifiers, it should be compared at least to alternative specifications using simpler alternatives.
-}
-@CampilloGimenez2013 and @Liu2022 \hl{provide excellent examples in which all possible choices are toggled and yield a neatly nested comparison.
+@CampilloGimenez2013 and @Liu2022 provide excellent examples in which all possible choices are toggled and yield a neatly nested comparison.
 Additionally, some state-of-the-art models, including others that obtain individualized predictions, should be included in comparisons.
@@
 We believe these principles will make reported results clearer and reproduction and extension easier.
-}
 
@@
 One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
-\hl{
 This is a necessary precondition for practical use but would also be a valuable contribution to the experimental literature.
@@
 This would test both measures for concurrent validity, while differences between them would inform what settings or cases are poorly served by one or the other.
-}
 
@@
 
-\hl{Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
-This need demands not only methodologies but also architectures that are robust, versatile, and compliant in the face of diverse data models and use restrictions, yet no studies in our sample addressed the challenges of system interoperability and regulatory requirements head-on.}
-
-\hl{CBR and, by extension, localized modeling are especially susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures as well.
-Individual systems exhibit many dimensions of incompleteness} [@Weiskopf2013] \hl{and vary along other dimensions of quality} [@Kohane2021]\hl{, and their aggregation for modeling purposes depends on several kinds of interoperability} [@Weber2015]\hl{.
-Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties} [@Haendel2021]\hl{.
+Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
+This need demands not only methodologies but also architectures that are robust, versatile, and compliant in the face of diverse data models and use restrictions, yet no studies in our sample addressed the challenges of system interoperability and regulatory requirements head-on.
+
+CBR and, by extension, localized modeling are especially susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures as well.
+Individual systems exhibit many dimensions of incompleteness [@Weiskopf2013] and vary along other dimensions of quality [@Kohane2021], and their aggregation for modeling purposes depends on several kinds of interoperability [@Weber2015].
+Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties [@Haendel2021].
 While most studies in our sample focused on improving care, increasingly many over time focused on generalizable knowledge.
 Therefore, as studies using localized modeling shift from proof of concept to feasibility, they will also need to demonstrate practical and regulatory feasibility.
-}
 
@@
 For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
-@Lopez2011 \hl{took an important step in this direction with an adjustable and adaptable implementation.}
+@Lopez2011 took an important step in this direction with an adjustable and adaptable implementation.
 Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.
@@
 Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth.
-Whereas precious few of the reviewed studies used these \hl{models for} any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+Whereas precious few of the reviewed studies used these models for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
 
@@
 
-[^pubmed-url]: <https://pubmed.ncbi.nlm.nih.gov/?term=(+"case-based+reasoning"+[All+Fields]+OR+"case-based+system"+[All+Fields]+)+OR+(+"individualized+modeling"+[All+Fields]+OR+"personalized+modeling"+[All+Fields]+OR+"customized+modeling"+[All+Fields]+)+OR+(+"individualized+cohort"+[All+Fields]+)+OR+(+(+"patient+similarity"+[All+Fields]+OR+"patient+distance"+[All+Fields]+OR+"patient+connection"+[All+Fields]+OR+"patient+affinity"+[All+Fields]+OR+"patient+clustering"+[All+Fields]+)+AND+(+"cohort+study"+[All+Fields]+)+)&sort=date>
+[^pubmed-url]: \url{https://pubmed.ncbi.nlm.nih.gov/?term=(+"case-based+reasoning"+[All+Fields]+OR+"case-based+system"+[All+Fields]+)+OR+(+"individualized+modeling"+[All+Fields]+OR+"personalized+modeling"+[All+Fields]+OR+"customized+modeling"+[All+Fields]+)+OR+(+"individualized+cohort"+[All+Fields]+)+OR+(+(+"patient+similarity"+[All+Fields]+OR+"patient+distance"+[All+Fields]+OR+"patient+connection"+[All+Fields]+OR+"patient+affinity"+[All+Fields]+OR+"patient+clustering"+[All+Fields]+)+AND+(+"cohort+study"+[All+Fields]+)+)&sort=date}
 
@@
 
-\hl{Of the 25 studies included, 4 showed up in no database searches (but included from the seed set), 1 was obtained via reference tracking, and, of the remaining 20 found through the database searches, only 1 was not found in PubMed or Web of Science.
+Of the 25 studies included, 4 showed up in no database searches (but included from the seed set), 1 was obtained via reference tracking, and, of the remaining 20 found through the database searches, only 1 was not found in PubMed or Web of Science.
 That one was instead found in Mathematical Reviews.
-Since MR returns about the same volume of results as WoS and PM, we omitted it from the update.}
-
-\hl{We applied the search terms exactly as before but restricted the dates to 2021 July 19 (the beginning of the date range for our original searches) or later.
-We omitted the WoS search that returned impractically many results the first time around and was omitted then.}
-
-\hl{We ran the update search on 2024 January 24.}
-
-\hl{The PubMed update returned 97 results. The Web of Science refresher returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so that 348 studies were slated for screening.}
-
-\hl{Screening by title and abstract was done in two waves.
+Since MR returns about the same volume of results as WoS and PM, we omitted it from the update.
+
+We applied the search terms exactly as before but restricted the dates to 2021 July 19 (the beginning of the date range for our original searches) or later.
+We omitted the WoS search that returned impractically many results the first time around and was omitted then.
+
+We ran the update search on 2024 January 24.
+
+The PubMed update returned 97 results. The Web of Science refresher returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so that 348 studies were slated for screening.
+
+Screening by title and abstract was done in two waves.
 First, PM excluded only on account of a study not being (a) clinical/medical, (b) English, and (c) original.
-Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.}
+Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.
 
@@
 
-\hl{Table} \ref{tab:composite} \hl{and Figure} \ref{fig:methods} \hl{summarize the techniques and terminology used by the included studies, as referred to in the main text.}
+Table \ref{tab:composite} and Figure \ref{fig:methods} summarize the techniques and terminology used by the included studies, as referred to in the main text.
 
@@
 
-\hl{Table} \ref{tab:performance} \hl{summarizes the evaluations and comparisons made of methods proposed in the included studies.
+Table \ref{tab:performance} summarizes the evaluations and comparisons made of methods proposed in the included studies.
 In most cases, not all performance results are included; for the sake of concision, only what we judged to be the headline results, using the primary evaluation metrics, are reproduced here.
-The table contains no original results.}
+The table contains no original results.
 
@@
   transmute(
-    Citation = str_c("[@", zotero_key, "]"),
+    Citation = str_c("@", zotero_key, ""),
     Measure = measure, Problem = task,
@@
 
+## Data and code
+
+Search results are publicly available in a Zotero group library.
+Data collected or encoded for included studies are publicly available in two Google Sheets.
+Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
+
+* Search results, included studies, and other bibliography: <https://www.zotero.org/groups/5017571/imsr/>
+* Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
+* Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
+* Code used to conduct analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
+
 # References {-}

### 2024-06-17 docs/localized-modeling-medRxiv/localized-modeling-medRxiv.tex
prepare arxiv manuscript

--- 2023-11-25 docs/Manuscript/Manuscript.tex
@@
-\documentclass[preprint, 3p,
-authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-
-\usepackage[hyphens]{url}
-
-  \journal{Artificial Intelligence In Medicine} % Sets Journal name
-
+\documentclass{article}
+
+\usepackage{arxiv}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
 \usepackage{graphicx}
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-% TODO: Currently lineno needs to be loaded after amsmath because of conflict
-% https://github.com/latex-lineno/lineno/issues/5
-\usepackage{lineno} % add
-  \linenumbers % turns line numbering on
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Clinical prediction with localized modeling using similarity-based cohorts: A scoping review},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
+
+\title{Clinical prediction with localized modeling using
+similarity-based cohorts: A scoping review}
+
+\author{
+    Adi Cohen
+   \\
+    College of Osteopathic Medicine, Nova Southeastern University \\
+  Davie, FL \\
+  \texttt{\href{mailto:ac4569@mynsu.nova.edu}{\nolinkurl{ac4569@mynsu.nova.edu}}} \\
+   \And
+    Patti McCall-Junkin
+   \\
+    Smathers Libraries Academic Research \& Consulting Services,
+University of Florida \\
+  Gainesville, FL \\
+  \texttt{\href{mailto:pattimccall.junkin@gmail.com}{\nolinkurl{pattimccall.junkin@gmail.com}}} \\
+   \And
+    Jason Cory Brunson
+   \\
+    Laboratory for Systems Medicine, University of Florida \\
+  Gainesville, FL \\
+  \texttt{\href{mailto:jason.brunson@medicine.ufl.edu}{\nolinkurl{jason.brunson@medicine.ufl.edu}}} \\
+  }
 
@@
 
-\usepackage{lscape} \newcommand{\landscapebegin}{\begin{landscape}} \newcommand{\landscapeend}{\end{landscape}}
-
+\usepackage{lscape}
+\newcommand{\landscapebegin}{\begin{landscape}}
+\newcommand{\landscapeend}{\end{landscape}}
 \begin{document}
-
-\begin{frontmatter}
-
-  \title{Clinical prediction with localized modeling using
-similarity-based cohorts: A scoping review}
-    \author[1]{Adi Cohen%
-  %
+\maketitle
+
+\begin{abstract}
+Background: Applications of classical case-based reasoning (CBR) have
+given rise to a family of techniques we call ``localized models'', in
+which a statistical model is fitted to a neighborhood of labeled cases
+matched by similarity to a target case. We aim to describe clinical and
+health applications of localized models to date and propose a general
+framework for their design and evaluation.
+
+Methods: We searched four bibliographic platforms during 2021 July
+19--22, updated 2024 January 24. We set four eligibility criteria to
+identify applications of localized models to clinical and health tasks.
+Two authors divided title/abstract screening and reviewed screened
+entries for inclusion. We discussed settings, tasks, and tools;
+identified and tabulated themes; and synthesized the methods into a
+general framework.
+
+Results: Of 1,657 search results, 360 were reviewed, then combined with
+43 publications that seeded the review and 1 obtained by citation
+tracking. 27 were included, published 1997--2022. The specificity of
+search terms was poor, and inter-rater reliability was low. Almost all
+models were predictive, the most common tasks being prognosis and
+diagnosis. Most studies used clinical, occasionally laboratory and
+image, data. Several addressed memory and runtime costs. A general
+technique that specializes to most of those reviewed involved matching,
+retrieval, fitting, and evaluation steps that could optionally be
+supervised, optimized, or recursively performed.
+
+Conclusions: Localized models have potential to improve the performance
+of clinical decision support tools while maintaining interpretability,
+but rigorous comparisons to competing methods must be conducted and
+computational hurdles must be overcome. We hope that our review will
+spur future work on efficiency, reproducibility, and user needs.
+\end{abstract}
+
+\keywords{
+    case-based reasoning
+   \and
+    localized modeling
+   \and
+    nearest neighbors
+   \and
+    clinical decision support
+   \and
+    scoping review
   }
-   \ead{ac4569@my.nsu.edu} 
-    \author[2]{Patti McCall-Junkin%
-  %
-  }
-   \ead{pattimccall.junkin@gmail.com} 
-    \author[3]{Jason Cory Brunson%
-  \corref{cor1}%
-  }
-   \ead{jason.brunson@medicine.ufl.edu} 
-      \affiliation[1]{College of Osteopathic Medicine, Nova Southeastern
-University, 3200 S University Drive, Davie, FL 33328}
-    \affiliation[2]{Smathers Libraries Academic Research \& Consulting
-Services, University of Florida, P.O. Box 117000, Gainesville, FL
-32611-7000}
-    \affiliation[3]{Laboratory for Systems Medicine, University of
-Florida, P.O. Box 100225 JHMHC, Gainesville, FL 32610-0225}
-    \cortext[cor1]{Corresponding author}
-
-  \begin{abstract}
-  Background: Applications of classical case-based reasoning (CBR) to
-  clinical and health tasks have given rise to a family of techniques in
-  which a statistical model is fitted to a subset of labeled cases,
-  selected for relevance or similarity to an unlabeled case to which the
-  model is applied. These localized models, as we propose to call them,
-  may perform as well as ``black-box'' models while retaining direct
-  interpretations. Their adoption has been hindered by fragmented
-  development and implementation challenges. Our goal in this scoping
-  review was to describe the clinical and health applications of
-  localized models to date and to derive a general framework for their
-  design and evaluation.
-
-  Methods: We searched four electronic bibliographic platforms (PubMed,
-  Web of Science, Academic Search Premier, and MathSciNet) during 2021
-  July 19--22. We screened and reviewed entries based on four criteria
-  intended to specify applications of localized models to tasks
-  involving analysis of clinical or health data. Two authors divided
-  title/abstract screening and collaboratively reviewed screened entries
-  for inclusion. We also screened and reviewed a seed set that inspired
-  the review and traced methodological citations from included entries.
-  We discussed and summarized settings, tasks, and tools; identified and
-  tabulated themes; and synthesized the methods into a general
-  framework. We received no specific funding for this work.
-
-  Results: Of 1,309 search results, 328 were reviewed. Of these,
-  combined with 43 publications that seeded the review and 1 obtained by
-  citation tracking, 25 were included in the review and methodological
-  synthesis. The specificity of several search terms was poor, and
-  inter-rater reliability was low. Entries were published from 1997 to
-  2021, with half of these since 2015. The most common tasks were
-  prognosis and diagnosis. The studies overwhelmingly focused on
-  clinical data, occasionally including laboratory and image-derived
-  data. Only three studies applied localized models to tasks other than
-  prediction. Several studies commented on memory and runtime costs, and
-  four proposed partial solutions. While few methods were reused and no
-  results were reproduced, a general technique that specializes to most
-  of those encountered was straightforward to describe.
-
-  Conclusions: Localized models have potential to improve the
-  performance of models used in clinical decision support tools while
-  maintaining interpretability, provided computational hurdles can be
-  overcome. Our search undoubtedly failed to capture all studies that
-  used methods of this kind, but those we found follow a consistent
-  enough design to be reproduced by a general-purpose tool. We hope that
-  our review and synthesis will spur future work on efficiency,
-  reproducibility, and user interactions and needs.
-  \end{abstract}
-    \begin{keyword}
-    case-based reasoning \sep localized modeling \sep nearest
-neighbors \sep clinical decision support \sep 
-    scoping review
-  \end{keyword}
-
- \end{frontmatter}
+
+\pagebreak
 
@@
 research use, the secondary research use of routinely collected health
-data has been put forth, with the monicker ``practice-based evidence'',
-as an important complement to the paradigm of evidence-based practice.
+data has been put forth, as ``practice-based evidence'', a complement to
+the paradigm of evidence-based practice.
 
@@
 more variables in opaque ways to yield predictions that depend on not
-only their values but their associations and their interplay with
-respect to outcomes. The loss of straightforward interpretations of
-these models has limited their practical uptake and motivated the
-construction of explanatory statistics for opaque models as well as the
-development of more interpretable complex models.
+only their values but their associations and interplay with each other.
+The loss of straightforward interpretations of these models has limited
+their practical uptake and motivated the construction of explanatory
+statistics for opaque models as well as the development of more
+interpretable complex models.
 
@@
 past or training cases that then inform the diagnosis, prognosis, or
-care of a new or test case. While traditional CBR returns these
-retrieved cases to inform human decision-making, the related nearest
-neighbors (NN) technique generates predictions from cohorts
-automatically via averaging (regression) or voting (classification) of
-their outcomes. In contrast, localized modeling fits families of
-predictive models, most commonly generalized linear models (GLMs), to
-these retrieved similarity cohorts in order to generate predictions.
-Localized models thus provide a way to fit fully interpretable models to
-large population data in a way that is personalized to each case.
+care of a new or test case. Traditional CBR returns these retrieved
+cases to inform human decision-making; the related nearest neighbors
+(NN) technique generates predictions from cohorts automatically via
+averaging (regression) or voting (classification) of their outcomes. In
+contrast, localized modeling fits families of predictive models, for
+example generalized linear models (GLMs), to retrieved similarity
+cohorts in order to generate predictions. Localized models thus provide
+an individualized way to fit fully interpretable models to large
+population data.
 
@@
 institution's records with which to conduct on-demand retrospective
-studies for an individual patient {[}1{]}. Our motivations in this
-scoping review are to describe the settings in and problems with which
-this approach has been tasked, to attempt to organize them within a
-common framework, and to evaluate the promise they show toward achieving
-this or other needs in clinical informatics. Along the way, we attempt
-to reconcile terminology, summarize motivations and evaluations, and
-propose valuable follow-up work.
+studies for an individual patient (Longhurst, Harrington, and Shah
+2014). Several hurdles face the deployment of such an approach in
+practice, but its benefits must first be demonstrated. Our motivations
+in this scoping review are to describe the settings in and problems with
+which this approach has been tasked, to attempt to organize them within
+a common framework, and to evaluate the promise they show toward
+achieving this or other needs in clinical informatics. Along the way, we
+attempt to reconcile terminology, summarize motivations and evaluations,
+and propose valuable follow-up work.
 
@@
 Clinical CBR emerged among rule-based approaches and other AI tools in
-the development of expert systems {[}2{]}. Early implementations
-realized a general workflow described as ``the four REs'', later the
-``R4 cycle'' {[}2,3{]}: Given a new case (or problem), the system
-\emph{retrieves} one or more past cases (solved problems) from a corpus,
-\emph{reuses} these to generate an understanding of (solution to) the
-new case, \emph{revises} this understanding (solution) to better fit the
-new case (sometimes called ``adaptation''), and \emph{retains} the new
-case and its eventual understanding (solution) in the corpus to be
-retrieved and reused in future. In contrast to rule-based systems, which
-are variable-based and generally interpretable as a single rule applied
-to all new cases, case-based systems provide case-specific
-interpretations in the form of a number of more fully understood
-reference cases. While the similarity measure used in the retrieval step
-need not be changed as the corpus grows, proposed measures have been
-diverse, contested, and rarely systematically compared.
-
-{[}4{]} distinguished two styles of CBR: problem-solving, which is more
-procedural and used when objectives are more clearly defined, and
-interpretative, which provides categorizations and justifications for
-possible solutions. A similar dichotomy is commonly used to distinguish
-the performance criterion for the usefulness of predictive models from
-the interpretability criterion for their usability. In these terms,
-localized models are highly procedural: The step of fitting a predictive
-model to a retrieved cohort can be understood as an automated adaptive
-strategy {[}3{]}, and the parameters that govern cohort retrieval can be
-tuned alongside model hyperparameters in a ML workflow. Accordingly,
-localized models have been primarily used for and evaluated on their
-ability to predict classes or outcomes. Nevertheless, as some recent
-applications have begun to show, they can be used to draw inferences
-about the importance of different risk factors to specific individuals.
-While most ML models come equipped with measures of feature importance
-and model-agnostic tools can generate explanations for model
+the development of expert systems (Aamodt and Plaza 1994). Early
+implementations realized a general workflow described as ``the four
+REs'', later the ``R4 cycle'' (Aamodt and Plaza 1994; Begum et al.
+2011): Given a new case (or problem), the system \emph{retrieves} one or
+more past cases (solved problems) from a corpus, \emph{reuses} these to
+generate an understanding of (solution to) the new case, \emph{revises}
+this understanding (solution) to better fit the new case (sometimes
+called ``adaptation''), and \emph{retains} the new case and its eventual
+understanding (solution) in the corpus to be retrieved and reused in
+future. In contrast to rule-based systems, which are variable-based and
+generally interpretable as a single rule applied to all new cases,
+case-based systems provide case-specific interpretations in the form of
+a number of more fully understood reference cases. While the similarity
+measure used in the retrieval step need not be changed as the corpus
+grows, proposed measures have been diverse, contested, and rarely
+systematically compared.
+
+Kolodner (1992) distinguished two styles of CBR: problem-solving, which
+is more procedural and used when objectives are more clearly defined,
+and interpretative, which provides categorizations and justifications
+for possible solutions. A similar dichotomy is commonly used to
+distinguish the performance criterion for the usefulness of predictive
+models from the interpretability criterion for their usability. In these
+terms, localized models are highly procedural: The step of fitting a
+predictive model to a retrieved cohort is an automated adaptive strategy
+(Begum et al. 2011), and the parameters that govern cohort retrieval can
+be tuned alongside model hyperparameters in a machine learning (ML)
+workflow. Accordingly, localized models have been primarily used for and
+evaluated on their ability to predict classes or outcomes. Nevertheless,
+as some recent applications have shown, they can be used to draw
+inferences about the importance of different risk factors to specific
+individuals. While most ML models come equipped with measures of feature
+importance and model-agnostic tools can generate explanations for model
 predictions, these quantifications are not directly interpretable model
 components analogous to the split nodes of a decision tree (DT) or the
-effect estimates of a GLM. While the patient similarity measure used to
+coefficients of a GLM. While the patient similarity measure used to
 retrieve each cohort may be complicated, the cohort itself can be
 directly inspected by the user. Provided the model family fitted to the
-cohorts is interpretable, the localized model as a whole inherits this
-property. Thus, localized models may embody both styles of CBR.
+cohorts is interpretable, the localized model inherits this property.
+Thus, localized models may embody both styles of CBR.
 
 We refer the interested reader to several previous reviews of CBR in
-medicine {[}3,5,6{]}, of measures of patient similarity {[}7{]}, and of
-uses of patient similarity in predictive models {[}8--10{]}. While the
-reviews of CBR focus on applications using health data, the patient
-similarity reviews encompass many additional types of data (various
-molecular -omics, genetic tests, medical images, laboratory tests,
-patient preferences, patient-reported outcomes, tracking devices, social
-media) and survey a much broader scope of models (exploratory analysis
-via dimension reduction and cluster analysis; risk evaluation and
-outcome prediction; clinical decision support and software tools). For
-the present review, we are interested in how similarity matching on
+medicine (Gierl, Bull, and Schmidt 1998; Begum et al. 2011; Choudhury
+and Begum 2016), of measures of patient similarity (Dai, Zhu, and Liu
+2020), and of uses of patient similarity in predictive models (Welch and
+Kawamoto 2013; Sharafoddini, Dubin, and Lee 2017; Parimbelli et al.
+2018). While the reviews of CBR focus on applications using health data,
+the patient similarity reviews encompass many additional types of data
+(various molecular -omics, genetic tests, medical images, laboratory
+tests, patient preferences, patient-reported outcomes, tracking devices,
+social media) and survey a much broader scope of models (exploratory
+analysis via dimension reduction and cluster analysis; risk evaluation
+and outcome prediction; clinical decision support and software tools).
+For the present review, we are interested in how similarity matching on
 patient-level health data is used to construct local cohorts for
@@
 
-Our goals in this review are (1) to describe applications to date of
+Our goals in this review are (1) to describe applications of
 similarity-based localized models using patient-level health data and
 (2) to provide a general framework for the design and evaluation of such
-localized models. Our selection of relevant papers from the search
-corpus will be based on the following inclusion/exclusion criteria:
+localized models. We focus narrowly on localized models, rather than
+broadly on patient similarity--based clinical decision support tools, so
+that we may thoroughly assess their value in terms of reported
+evaluations and comparisons to other methods. This also allows us to
+devise a high-level framework that specializes to the majority of these
+methods, which we then use to frame our discussion and recommendations.
+
+Our selection of relevant papers from the search corpus will be based on
+the following inclusion/exclusion criteria:
 
@@
 patient similarity measure, a cohort selection process, and a
-statistical model family. Such a general method would be
-straightforward, if laborious, to implement, and specializations of it
-would account for the majority of real-world applications. A general
-implementation based on this method would allow researchers to expedite
-every step of the analysis process, including selection, optimization,
-and evaluation, and enable sensitivity, robustness, and multiverse
-analyses that help identify the most consequential choices along the
-way.
+statistical model family. A general implementation based on this method
+would allow researchers to expedite every step of the analysis process,
+including retrieval, optimization, and evaluation, and enable
+sensitivity, robustness, and multiverse analyses that help identify the
+most consequential choices along the way.
 
@@
 The structure and terms of our search strings were inspired in part by
-previous reviews adjacent to our topic of interest {[}9,10{]}. We
-organized the PubMed search in disjunctive normal form (an OR of ANDs).
-Following the solidification of an outline, we expanded each term to
-include synonyms that are similar enough to be applicable to our search.
-We then evaluated the expanded search string using the PubMed Advanced
-Search platform. We initially included each term and their synonyms
-separately to evaluate what resulted. We pruned several terms that
-returned no results (``phrases not found'') or to reduce the number of
-results. At the conclusions of this process for each individual term, we
-combined the search terms using disjunctive normal form.
+previous reviews adjacent to our topic of interest (Sharafoddini, Dubin,
+and Lee 2017; Parimbelli et al. 2018). We organized the PubMed search in
+disjunctive normal form (an OR of ANDs). Following the solidification of
+an outline, we expanded each term to include synonyms that are similar
+enough to be applicable to our search. We then evaluated the expanded
+search string using the PubMed Advanced Search platform. We initially
+included each term and their synonyms separately to evaluate what
+resulted. We pruned several terms that returned no results (``phrases
+not found'') or to reduce the number of results. At the conclusions of
+this process for each individual term, we combined the search terms
+using disjunctive normal form.
 
@@
 the pooled results to Covidence, which merged some duplicate entries.
+
+We repeated the search on 2024 January 24 on the PubMed and Web of
+Science databases to bring the results up to date.
 
@@
 regularly reviewed each other's decisions to improve consistency. In
-cases of uncertainty, entries were included.
+cases of uncertainty, entries were included. For the update, one author
+(PM) screened all results for three criteria and another (JCB) screened
+the survivors for the fourth criterion.
 
@@
 that are fit and evaluated in separate steps, we chose to allow simple
-statistical summaries such as mean survival {[}11{]} and survival curves
-{[}12{]}. These approaches were novel to CBR and presage later
-developments, so were helpful to understanding the development of
-localized modeling. However, this then admitted studies that applied
-conventional nearest neighbors prediction: Each retrieved cohort
-consisted of the \(k\) most similar cases to the index case, and the
-response for the index case was predicted to be either the mean
+statistical summaries such as mean survival (Mariuzzi et al. 1997) and
+survival curves (Lowsky et al. 2013). These approaches were novel to CBR
+and presage later developments, so were helpful to understanding the
+development of localized modeling. However, this then admitted studies
+that applied conventional nearest neighbors prediction: Each retrieved
+cohort consisted of the \(k\) most similar cases to the index case, and
+the response for the index case was predicted to be either the mean
 (continuous response) or the plurality (discrete response) of the
@@
 Finally, one author (JCB) applied the same review process to the seed
-set of 43 studies that inspired the review.
+set of 43 studies that inspired the review. The same author later
+reviewed the updated results.
 
@@
 
-Following the selection sources aligning with our review, we collected
-characteristics of included studies (Section \ref{data-and-code}). The
-features included bibliographic fields (date of publication, journal,
-authors, title, keywords, DOI), study goals (objective, generalizable
-knowledge, evaluation, clinical/medical domain), data sets (data source,
-type of data, range of data, number of cases/incidences, number of
+We next collected characteristics of included studies. The features
+included bibliographic fields (date of publication, journal, authors,
+title, keywords, DOI), study goals (objective, generalizable knowledge,
+evaluation, clinical/medical domain), data sets (data source, type of
+data, range of data, number of cases/incidences, number of
 predictors/features), and methodological choices (types of similarity
@@
 reviews, we here pursue a methodology synthesis to harmonize largely
-independent research efforts that have converged on a common modeling
-technique. The goal will be to describe a unified framework for
+independent research efforts that have converged on a common family of
+techniques. The goal will be to describe a unified framework for
 localized models that can be used to guide future study designs and
@@
 databases and registers. Following the completion and input of each
-search string, there were a total of 1,422 sources within all of the
-platforms used. De-duplication resulted in 1,309 entries, which were
-added to the title/abstract screening for review. Of these, 328 entries
+search string, there were a total of 1,817 sources within all of the
+platforms used. De-duplication resulted in 1,657 entries, which were
+added to the title/abstract screening for review. Of these, 360 entries
 met the screening criteria and were assessed through full text review.
-Of these, 44 fit the original criteria, and 19 were included as distinct
+Of these, 51 fit the original criteria, and 21 were included as distinct
 from NN prediction.
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-prisma}
+{\centering \includegraphics[width=0.8\linewidth]{Fig1}
 
@@
 
-\caption{PRISMA-S flow chart. Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).}\label{fig:prisma}
+\caption{PRISMA-S flow chart.
+Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).
+Within the structured search results, summands correspond to original and update.}\label{fig:prisma}
 \end{figure}
 
-There were 60 disagreements over inclusion versus exclusion. Inter-rater
-reliability was low, at 82\% relative to a 72\% probability of random
-agreement.
+From the first search results, there were 60 disagreements over
+inclusion versus exclusion. Inter-rater reliability was low, at 82\%
+relative to a 72\% probability of random agreement. Only one author was
+available to review full texts from the update search.
 
@@
 review. After removing duplicates and screening for eligibility, we were
-left with 6 additional studies {[}12--17{]}, 1 of which was excluded
-from the synthesis for using NN prediction. Finally, reference tracking
-from the 43 + 328 = 371 studies assessed for eligibility led us to
-identify 1 additional study that met our criteria {[}18{]}. This left us
-with 19 + 1 + 5 = 25 studies included in the review and synthesis.
+left with 6 additional studies (Park, Kim, and Chun 2006; Lowsky et al.
+2013; Lee, Maslove, and Dubin 2015; Ng et al. 2015; Lee 2017; N. Wang et
+al. 2019), 1 of which was excluded from the synthesis for using NN
+prediction. Reference tracking from the 43 + 360 = 403 studies assessed
+for eligibility led us to identify 1 additional study that met our
+criteria (Kasabov and Hu 2010). This left us with 21 + 1 + 5 = 27
+studies included in the review and synthesis.
 
@@
 
-The 25 included studies were analyzed based on several characteristics
-(see Section \ref{data-and-code}), and we report and describe some
-observations here (Table \ref{tab:synthesis}). The studies were
-published in a variety of journals, with some of greater frequency,
-though no single journal published more than 3. The journal
-\emph{Evolving Systems} published 2 of the included studies,
-\emph{Artificial Intelligence in Medicine} published 3, \emph{Hindawi -
-Journal of Healthcare Engineering} published 3, and \emph{PLOS One}
+The 27 included studies were analyzed based on several characteristics,
+and we report and describe some observations here (Table
+\ref{tab:synthesis}). The studies were published in a variety of
+journals, with some of greater frequency, though no single journal
+published more than 3. The journal \emph{Evolving Systems} published 2
+of the included studies, \emph{Artificial Intelligence in Medicine}
+published 3, \emph{Hindawi - Journal of Healthcare Engineering}
+published 3, \emph{Evolving Systems} published 2, and \emph{PLOS One}
 published 2.
@@
 Another interesting pattern lay in the years of publication (Figure
-\ref{fig:year}). While they trace back to the late 1990s, they have
-become more common, which suggests that this is an active, though not
-rapidly expanding, methodological approach. We note that, while CBR in
-health and medicine originated as early as 1990, these studies did
-emerge early on once CBR had established itself. So the idea has been
-``in the air'' for as long as CBR has been in use.
+\ref{fig:year}). While these studies trace back to the late 1990s, they
+have become more common, which suggests that this is an active, though
+not rapidly expanding, approach. We note that CBR in health and medicine
+originated as early as 1990, and localized modeling emerged soon after
+CBR had established itself; the idea has been ``in the air'' for as long
+as CBR has been in use.
+
+\small
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1215}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1602}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1215}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1823}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1547}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1050}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1547}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1148}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1585}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1202}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1967}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1530}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1038}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1530}}@{}}
 \caption{\label{tab:synthesis}Studies included in the method synthesis,
@@
 \endlastfoot
-{[}19{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,355 & 9 \\
-{[}11{]} & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
+Yearwood and Wilkinson (1997) & Prediction\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 1,355 & 9 \\
+Mariuzzi et al. (1997) & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Longitudinal\hspace{6em} & 113 & 4 \\
-{[}20{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
+Wyns et al. (2004) & Prediction\hspace{6em} & Practice\hspace{6em} &
 Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 160 & 14 \\
-{[}13{]} & Diagnosis\hspace{6em} & Practice\hspace{6em} & Clinical,
-Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 366; 270; 560; 760 &
-35; 14; 31; 8 \\
-{[}21{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
+Park, Kim, and Chun (2006) & Diagnosis\hspace{6em} &
+Practice\hspace{6em} & Clinical, Imaging\hspace{6em} &
+Cross-sectional\hspace{6em} & 366; 270; 560; 760 & 35; 14; 31; 8 \\
+Song and Kasabov (2006) & Prediction\hspace{6em} & Practice\hspace{6em}
+& Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
 6 \\
-{[}22{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 2,620 & 6 \\
-{[}23{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
+Elter, Schulz-Wendtland, and Wittenberg (2007) & Prediction\hspace{6em}
+& Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 2,620 & 6 \\
+Xu et al. (2008) & Prediction\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 67 & 14 \\
-{[}24{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+López et al. (2011) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 871 & 4 \\
-{[}18{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+Kasabov and Hu (2010) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62 & 2,000 \\
-{[}25{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 74 & 93 \\
-{[}26{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62; 72; 77; 181 &
-2,000; 7,129; 7,129; 12,533 \\
-{[}12{]} & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
-Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 13,525 & 13 \\
-{[}27{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Patient-reported\hspace{6em} & Cross-sectional\hspace{6em} & 1,647 &
-18 \\
-{[}28{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+Verma et al. (2015) & Decision Support\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 74 & 93 \\
+Liang, Hu, and Kasabov (2015) & Prediction\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 62; 72; 77; 181 & 2,000; 7,129; 7,129;
+12,533 \\
+Lowsky et al. (2013) & Prediction\hspace{6em} & Practice\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 13,525 &
+13 \\
+Campillo-Gimenez et al. (2013) & Decision Support\hspace{6em} &
+Practice\hspace{6em} & Patient-reported\hspace{6em} &
+Cross-sectional\hspace{6em} & 1,647 & 18 \\
+Nicolas et al. (2014) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 150 & 83 \\
-{[}15{]} & Clustering\hspace{6em} & Practice\hspace{6em} & Clinical,
-Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 7,519 & 130 \\
-{[}14{]} & Prediction\hspace{6em} & Knowledge\hspace{6em} & Clinical,
+Ng et al. (2015) & Clustering\hspace{6em} & Practice\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 7,519 &
+130 \\
+Lee, Maslove, and Dubin (2015) & Prediction\hspace{6em} &
+Knowledge\hspace{6em} & Clinical, Laboratory\hspace{6em} &
+Cross-sectional\hspace{6em} & 17,152 & 75 \\
+Vilhena et al. (2016) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,046 & 8 \\
+Lee (2017) & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
 Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
-{[}29{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,046 & 8 \\
-{[}16{]} & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
-Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
-{[}30{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+Zhang et al. (2018) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
-{[}31{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 638 & 49,728 \\
-{[}32{]} & Prediction\hspace{6em} & Practice\hspace{6em} &
+Malykh and Rudetskiy (2018) & Decision Support\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 638 & 49,728 \\
+Ma et al. (2020) & Prediction\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 4,000 & 22 \\
-{[}17{]} & Diagnosis\hspace{6em} & Practice\hspace{6em} & Clinical,
-Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 16,490 & 106 \\
-{[}33{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 8 & 5 \\
-{[}34{]}; {[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em}
-& Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 &
-1,466,474 \\
+N. Wang et al. (2019) & Diagnosis\hspace{6em} & Practice\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 16,490
+& 106 \\
+Y. Wang et al. (2020) & Decision Support\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} &
+Cross-sectional\hspace{6em} & 8 & 5 \\
+Tang et al. (2021); Ng et al. (2021) & Decision Support\hspace{6em} &
+Practice\hspace{6em} & Clinical\hspace{6em} & Longitudinal\hspace{6em} &
+245,825 & 1,466,474 \\
+Liu et al. (2022) & Prediction\hspace{6em} & Practice\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 76,597
+& 1,892 \\
+Doborjeh et al. (2022) & Prediction\hspace{6em} & Practice\hspace{6em} &
+Clinical, Environmental\hspace{6em} & Longitudinal\hspace{6em} & 804 &
+46 \\
 \end{longtable}
 
+\normalsize
+
 \begin{figure}
 
-{\centering \includegraphics[width=0.75\linewidth]{../../fig/fig-years}
+{\centering \includegraphics[width=1\linewidth]{Fig2}
 
@@
 whether their aim was to produce generalizable knowledge or to improve
-practice. We classified 7 of the 25 studies as ``knowledge'', making the
+practice. We classified 8 of the studies as ``knowledge'', making the
 majority of studies from a ``practice'' standpoint. These studies aimed
@@
 
-Lastly, we observed a commonality among 6 of the 25 included studies in
+Lastly, we observed a commonality among 6 of the included studies in
 having evaluated methods using leave-one-out cross-validation (LOOCV).
@@
 \ref{tab:composite}). In the following subsections, we take a closer
-look at these design elements.
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1128}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4410}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4462}}@{}}
-\caption{\label{tab:composite}Methodological elements of studies
-included in the synthesis.}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Citation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Elements
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Terminology
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Citation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Elements
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Terminology
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-{[}19{]} & supervised similarity learning & CBR; case-structured
-retrieval \\
-{[}11{]} & predictive modeling on similarity cohorts & CBR; targeted
-subsampling \\
-{[}20{]} & unsupervised similarity learning & Combined Kohonen type
-neural network-case-based reasoning \\
-{[}13{]} & supervised cohort construction & CBR, statistical case-based
-reasoning \\
-{[}21{]} & supervised similarity learning & Transductive inference \\
-{[}22{]} & unsupervised similarity learning & CBR; entropic distance
-measure \\
-{[}23{]} & tiered constraint; similarity matching & CBR \\
-{[}24{]} & interactive implementation & CBR; eXiT*CBR \\
-{[}18{]} & predictive modeling on similarity cohorts & Personalized
-model \\
-{[}25{]} & predictive modeling on similarity cohorts & Personalized
-modeling; TWNFI \\
-{[}26{]} & predictive modeling on similarity cohorts & Local modeling;
-personalized modeling; TWNFI \\
-{[}12{]} & predictive modeling on similarity cohorts & K-nearest
-neighbors survival \\
-{[}27{]} & supervised similarity learning & CBR \\
-{[}28{]} & supervised similarity learning & CBR \\
-{[}15{]} & predictive modeling on similarity cohorts; explanatory
-modeling on similarity cohorts & Personalized predictive model \\
-{[}14{]} & predictive modeling on similarity cohorts & Personalized
-prediction \\
-{[}29{]} & supervised similarity learning; predictive modeling on
-similarity cohorts & CBR \\
-{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
-predictive model \\
-{[}30{]} & whole-population weight learning & Gaussian processes \\
-{[}31{]} & unclear & CBR \\
-{[}32{]} & predictive modeling on similarity cohorts & Personalized
-model \\
-{[}17{]} & predictive modeling on similarity cohorts & Personalized
-predictive modeling \\
-{[}33{]} & descriptive modeling on similarity cohorts & CBR \\
-{[}34{]}; {[}35{]} & supervised similarity learning; predictive modeling
-on similarity cohorts & Similarity model, personalized model, precision
-cohort, personalized treatment options \\
-\end{longtable}
+look at these design elements and their reported justifications and
+limitations.
 
@@
 evaluated the usefulness of CBR-based explanations for the purpose of
-decision support {[}36{]}. This study was part of a much larger
-literature on CBR systems and was included here despite relying on NN
-prediction because it used an unconventional voting scheme to generate
-recommendations. A more recent study took essentially the same focus
-with respect to a proposed clinical risk prediction model, which
-amounted to CBR with a novel weighting scheme on predictors informed by
-expert consensus {[}37{]}. In both cases a prototype implementation was
-deployed in an experimental setting for evaluation.
+decision support (Doyle, Cunningham, and Walsh 2006). This study was
+part of a much larger literature on CBR systems and was included here
+despite relying on NN prediction because it used an unconventional
+voting scheme to generate recommendations. A more recent study took
+essentially the same focus with respect to a proposed clinical risk
+prediction model, which amounted to CBR with a novel weighting scheme on
+predictors informed by expert consensus (Fang et al. 2021). In both
+cases a prototype implementation was deployed in an experimental setting
+for evaluation.
 
@@
 recommendation. The plurality focused on prognosis or outcome
-prediction, often using time-to-event analysis: {[}11{]} used CBR to
-predict survival time from several geometric properties of breast
-tumors. {[}12{]} used CBR with non-parametric survival models on
-registry data to predict patient--graft survival times following kidney
-transplantation. {[}14{]} and {[}16{]} used localized logistic
-regression and random forest modeling to predict 30-day mortality
-following discharge for ICU patients. {[}29{]} developed a CBR cycle
-around a clustering-informed similarity matching procedure and an
-artificial neural network--based classifier to identify thrombophilia
-patients at high risk of thrombotic episodes. And {[}32{]} took a
-similarity cohort--based approach to predicting length of stay for ICU
-patients. Also of note, from a public health perspective, {[}23{]} used
-CBR to predict rehabilitation time as well as disability risk for
-unemployed workers experiencing chronic pain.
-
-Toward detection and diagnosis, {[}20{]} proposed a hybrid neural
-net--CBR system to distinguish (with confidence bounds) arthritic versus
-control patients, based on several histological features. {[}28{]} used
-collaborative multilabel CBR to subtype melanoma patients based on
-confocal and dermoscopy images. {[}17{]} used localized models built
-from a multi-type additive similarity measure to distinguish type 2
-diabetic versus control populations. Along the way, {[}21{]}, {[}18{]},
-and {[}25{]} took an iterative model-building approach to several tasks:
+prediction, often using time-to-event analysis: Mariuzzi et al. (1997)
+used CBR to predict survival time from several geometric properties of
+breast tumors. Lowsky et al. (2013) used CBR with non-parametric
+survival models on registry data to predict patient--graft survival
+times following kidney transplantation. Lee, Maslove, and Dubin (2015)
+and Lee (2017) used localized logistic regression and random forest
+modeling to predict 30-day mortality following discharge for ICU
+patients. Vilhena et al. (2016) developed a CBR cycle around a
+clustering-informed similarity matching procedure and an artificial
+neural network--based classifier to identify thrombophilia patients at
+high risk of thrombotic episodes. Ma et al. (2020) took a similarity
+cohort--based approach to predicting length of stay for ICU patients.
+Doborjeh et al. (2022) localized spiking neural networks in order to
+assess stroke risk from up to 7-day time series of clinical and
+environmental factors. Liu et al. (2022) incorporated global-to-local
+transfer learning into localized models of acute kidney injury risk in
+hospitalized patients. Also of note, from a public health perspective,
+Xu et al. (2008) used CBR to predict rehabilitation time as well as
+disability risk for unemployed workers experiencing chronic pain.
+
+Toward detection and diagnosis, Wyns et al. (2004) proposed a hybrid
+neural net--CBR system to distinguish (with confidence bounds) arthritic
+versus control patients, based on several histological features. Nicolas
+et al. (2014) used collaborative multilabel CBR to subtype melanoma
+patients based on confocal and dermoscopy images. N. Wang et al. (2019)
+used localized models built from a multi-type additive similarity
+measure to distinguish type 2 diabetic versus control populations. Along
+the way, Song and Kasabov (2006), Kasabov and Hu (2010), and Verma et
+al. (2015) took an iterative model-building approach to several tasks:
 predicting glomerular filtration rate, a key indicator of renal
@@
 treatment decisions into their approach: By taking ``decision points''
-as their units of analysis, {[}34{]} and {[}35{]} built not classifiers
-or predictors but comparative effectiveness models into a localized
-framework, providing for the first time in our sample explicitly
-prescriptive rather than descriptive clinical decision support.
+as their units of analysis, Tang et al. (2021) and Ng et al. (2021)
+built not classifiers or predictors but comparative effectiveness models
+into a localized framework, providing for the first time in our sample
+explicitly prescriptive rather than descriptive clinical decision
+support.
 
@@
 decision support tool, in this case for early diagnosis of melanoma from
-clinical data and dermoscopy images {[}28{]}. While the stated
-objectives were analogous, specific emphasis was placed on the
+clinical data and dermoscopy images (Nicolas et al. 2014). While the
+stated objectives were analogous, specific emphasis was placed on the
 acquisition of new knowledge through the development of the tool,
@@
 
-{\centering \includegraphics[width=1\linewidth]{../../fig/fig-properties}
+{\centering \includegraphics[width=1\linewidth]{Fig3}
 
@@
 
-\caption{Share of studies characterized by several design elements: data source (row), clinical task (fill color), and remit (opacity).}\label{fig:properties}
+\caption{Share of studies characterized by several design elements: data source (row), clinical task (fill), and remit (opacity).}\label{fig:properties}
 \end{figure}
@@
 
-\label{rationales}
-
 The included studies hypothesized, asserted, or assumed several benefits
 of localized modeling specific to clinical and medical settings and
-advantages over other modeling approaches (see Section
-\ref{data-and-code}). The most common was that the restriction to
-similar or relevant past cases would improve predictive performance for
-the index case {[}11,15,16,26{]}. In particular, {[}14{]} hypothesized
-and confirmed that the value of each past case would be positively
-related to its similarity to the index case, an assumption built in to
-the weighting schemes of other approaches. {[}12{]} made a different
-case, that the fewer parametric assumptions and complications of a
-CBR-style model would allow for greater accuracy. Additionally, several
-investigators asserted that the use of localized cohorts befit the
-clinical focus on the individual patient rather than the population,
-without reference to performance {[}21,23{]}.
-
-In an interesting contrast, {[}34{]} and {[}35{]} argued that their
-localized approach using the larger and more heterogeneous populations
-covered by EHR-derived data could better capture the messy and diverse
-lessons of everyday practice, as a counterpart to guidelines based on
-randomized controlled trials. Their emphasis on the enabling role of
-EHRs to power methods well-suited to large, structured data repositories
-was shared by several others {[}16,25,27,28{]}. {[}25{]} and {[}30{]}
-additionally pointed out that localized models are adaptable to noise in
-the data, variation in patterns of missingness, and (harkening to the
-last step of the R4 cycle) addition of new cases with known outcomes to
-the corpus. These properties, they said, tend to be more difficult for
+advantages over other modeling approaches. The most common was that the
+restriction to similar or relevant past cases would improve predictive
+performance for the index case (Mariuzzi et al. 1997; Liang, Hu, and
+Kasabov 2015; Ng et al. 2015; Lee 2017). In particular, Lee, Maslove,
+and Dubin (2015) hypothesized and confirmed that the value of each past
+case would be positively related to its similarity to the index case, an
+assumption built in to the weighting schemes of other approaches. Lowsky
+et al. (2013) made a different case, that the fewer parametric
+assumptions and complications of a CBR-style model would allow for
+greater accuracy. Additionally, several investigators asserted that the
+use of localized cohorts befit the clinical focus on the individual
+patient rather than the population, without reference to performance
+(Song and Kasabov 2006; Xu et al. 2008).
+
+In an interesting contrast, Tang et al. (2021) and Ng et al. (2021)
+argued that their localized approach using the larger and more
+heterogeneous populations covered by EHR-derived data could better
+capture the messy and diverse lessons of everyday practice, as a
+counterpart to guidelines based on randomized controlled trials. Their
+emphasis on the enabling role of EHRs to power methods well-suited to
+large, structured data repositories was shared by several others
+(Campillo-Gimenez et al. 2013; Nicolas et al. 2014; Verma et al. 2015;
+Lee 2017). Verma et al. (2015) and Zhang et al. (2018) additionally
+pointed out that localized models are adaptable to noise in the data,
+variation in patterns of missingness, and (harkening to the last step of
+the R4 cycle) addition of new cases with known outcomes to the corpus.
+These properties, they said, tend to be more difficult for
 whole-population models to handle.
@@
 The other frequent advantage attributed to localized modeling was
-interpretability. {[}22{]} and {[}28{]} emphasized the value of the
-detailed past cases, available to the user, on which predictions are
-based. This ``self-explanation capability'' made outputs more
-intelligible to physicians in the CDS setting. {[}15{]} additionally
-pointed out that localized GLMs yield localized effect estimates, which
-may help investigators identify individually relevant risk factors.
+interpretability. Elter, Schulz-Wendtland, and Wittenberg (2007) and
+Nicolas et al. (2014) emphasized the value of the detailed past cases,
+available to the user, on which predictions are based. This
+``self-explanation capability'' made outputs more intelligible to
+physicians in the CDS setting. Ng et al. (2015) additionally pointed out
+that localized GLMs yield localized effect estimates, which may help
+investigators identify individually relevant risk factors. Liu et al.
+(2022) took this idea further, systematically comparing regression
+coefficients for specific risk factors across dozens of diagnostic
+subpopulations. Doborjeh et al. (2022) used localized time series models
+to identify environmental changes associated with increased stroke risk.
 
@@
 of then-conventional CBR, most of which defended the use of other tools
-to improve performance via cohort selection {[}27--29{]} or to make
-models and outputs more interpretable {[}17,24{]}. {[}26{]} argued for
-simultaneous optimization of cohort and feature selection with model
-parameterization; their TWNFI approach combines localization with
-regularization.
+to improve performance via cohort selection (Campillo-Gimenez et al.
+2013; Nicolas et al. 2014; Vilhena et al. 2016) or to make models and
+outputs more interpretable (López et al. 2011; N. Wang et al. 2019).
+Liang, Hu, and Kasabov (2015) argued for simultaneous optimization of
+cohort and feature selection with model parameterization; their TWNFI
+approach combines localization with regularization.
 
@@
 CBR can be understood as the opposite side of a trade-off with
-rule-based reasoning between model size and model complexity; {[}36{]}
-describe their approach as ``knowledge-light'', in that ``the cases do
-not contain explicit explanation structures; instead, explanation is
-achieved by comparison of the query case with retrieved cases''. This
-means that the greatest performance and efficiency challenges in CBR
-have to do with the retrieval and revision phases in the R4 cycle.
-Several studies addressed these challenges: {[}13{]}, while excluded
-from the synthesis, were the earliest to propose that cohorts be bounded
-by a similarity threshold rather than by a number of cases, and this
-improved performance in their experiments. {[}38{]} leveraged predictor
-weights obtained from logistic regression to inform the similarity
-calculation used in retrieval. {[}12{]} used non-parametric models to
-reduce the computational burden of revision (prediction). {[}32{]}
-proposed to improve efficiency along the entire R4 cycle, but in
-particular by only executing task-dependent steps in real time
-(``just-in-time learning'', JITL), and {[}35{]} and {[}34{]} partitioned
-multiple phases in the R4 cycle into offline and online components, only
-the latter of which would be performed in real time as new data are
-received.
-
-Other studies addressed limitations of available tools. {[}24{]}
-implemented a comprehensive CBR tool in response to the lack of
+rule-based reasoning between model size and model complexity: Doyle,
+Cunningham, and Walsh (2006) describe their approach as
+``knowledge-light'', in that ``the cases do not contain explicit
+explanation structures; instead, explanation is achieved by comparison
+of the query case with retrieved cases''. This means that the greatest
+performance and efficiency challenges in CBR have to do with the
+retrieval and revision phases in the R4 cycle. Several studies addressed
+these challenges: Park, Kim, and Chun (2006), while excluded from the
+synthesis, were the earliest to propose that cohorts be bounded by a
+similarity threshold rather than by the number of cases, which improved
+performance in their experiments. Campillo-Gimenez et al. (2013)
+leveraged predictor weights obtained from logistic regression to inform
+the similarity calculation used in retrieval. Lowsky et al. (2013) used
+non-parametric models to reduce the computational burden of revision
+(prediction). Ma et al. (2020) proposed to improve efficiency along the
+entire R4 cycle, but in particular by only executing task-dependent
+steps in real time (``just-in-time learning'', JITL), and Ng et al.
+(2021) and Tang et al. (2021) partitioned multiple phases in the R4
+cycle into offline and online components, only the latter of which would
+be performed in real time as new data are received. Liu et al. (2022)
+used transfer learning from globally-fitted models to improve the
+efficiency of localized models.
+
+Other studies addressed limitations of available tools. López et al.
+(2011) implemented a comprehensive CBR tool in response to the lack of
 general-purpose software, to enable coupling with other tools as well as
 to expedite development, experimentation, and uptake. Several other
-teams also set out to develop more interoperable clinical support tools
-{[}15,22,26,30{]}. The frequency plot in Figure \ref{fig:methods}
-compares the rates at which several methodological elements are invoked
-in the sample.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{../../fig/fig-methods}
-
-}
-
-\caption{Frequency of recurring methodological elements across included studies.}\label{fig:methods}
-\end{figure}
-
-\hypertarget{performance-assessments}{%
-\subsection{Performance assessments}\label{performance-assessments}}
-
-\label{performance-assessments}
+teams also set out to develop more generalizable and data-agnostic
+clinical support tools (Elter, Schulz-Wendtland, and Wittenberg 2007;
+Liang, Hu, and Kasabov 2015; Ng et al. 2015; Zhang et al. 2018).
+However, more informatical and legal challenges to implementation,
+including interoperability of systems and data and regulations
+concerning security and privacy, were rarely addressed. We discuss this
+further in Section \ref{feasibility}.
+
+Figure \ref{fig:methods} compares the rates at which several
+methodological elements are invoked in the sample.
+
+\hypertarget{evaluations}{%
+\subsection{Evaluations}\label{evaluations}}
 
@@
 meta-analyze these numbers due to the great variety of settings,
-problems, data types, and choices involved.
+problems, data types, techniques, and choices involved.
 
 We do observe one clear pattern: All of the proposed methods that most
-evidently outperformed their comparators (Kohonen + CBR {[}20{]}, TWNFI
-{[}18{]}, gravitational search algorithms (GSA) {[}26{]}, CBR + rules (+
-DML) {[}28{]}, Gaussian process regression {[}30{]}, JITL-ELM {[}32{]})
-are hybrids of localized modeling (in some cases CBR) with other
-techniques, often DML. Though {[}27{]} and {[}15{]} report non-superior
-performance by such hybrids, the pattern suggests the importance of the
-similarity measure to the retrieval step. Meanwhile, when proposed
-approaches targeted cohort demarcation or choice of predictive model
-(statistical CBR {[}13{]}; individualized logistic regression, decision
-tree, and random forest {[}14,16{]}), they did not consistently
-outperform comparators.
+evidently outperformed their comparators---Kohonen + CBR (Wyns et al.
+2004), TWNFI Kasabov and Hu (2010), gravitational search algorithms
+(GSA) (Liang, Hu, and Kasabov 2015), CBR + rules (+ DML) (Nicolas et al.
+2014), Gaussian process regression (Zhang et al. 2018), JITL-ELM (Ma et
+al. 2020)---are hybrids of localized modeling (in some cases CBR) with
+other techniques, often DML. Though Campillo-Gimenez et al. (2013) and
+Ng et al. (2015) report non-superior performance by such hybrids, the
+pattern suggests the importance of the similarity measure to the
+retrieval step. Meanwhile, when proposed approaches targeted cohort
+demarcation or choice of predictive model---statistical CBR (Park, Kim,
+and Chun 2006); individualized logistic regression, decision tree, and
+random forest (Lee, Maslove, and Dubin 2015; Lee 2017)---they did not
+consistently outperform comparators.
 
@@
 The study authors focused their recommendations and their own plans for
-future work mostly on technical improvements and evaluations
-{[}11,19{]}. Urged improvements included full or partial automation of
-predictor selection {[}11,19{]}, similarity learning {[}11,17{]}, and
-parameter optimization {[}16,21{]}; extensions to new data structures
-{[}24{]}, data types {[}25,26,31{]}, and reasoning systems {[}28{]}; and
-the use of more advanced model components to improve accuracy or
-efficiency {[}12,14,26,30{]}. Authors also urged validations and
+future work mostly on technical improvements and evaluations (Mariuzzi
+et al. 1997; Yearwood and Wilkinson 1997). Urged improvements included
+full or partial automation of predictor selection (Mariuzzi et al. 1997;
+Yearwood and Wilkinson 1997), similarity learning (Mariuzzi et al. 1997;
+N. Wang et al. 2019), and parameter optimization (Song and Kasabov 2006;
+Lee 2017); extensions to new data structures (López et al. 2011), data
+types (Liang, Hu, and Kasabov 2015; Verma et al. 2015; Malykh and
+Rudetskiy 2018), and reasoning systems (Nicolas et al. 2014); and the
+use of more advanced model components to improve accuracy or efficiency
+(Lowsky et al. 2013; Lee, Maslove, and Dubin 2015; Liang, Hu, and
+Kasabov 2015; Zhang et al. 2018). Authors also urged validations and
 independent evaluations using larger or more comprehensive data sets
-{[}15,22,23,25{]}, using data aggregated from multiple health systems
-{[}14,16,34,35{]}, and in other care settings or disease contexts
-{[}21,30,34,35{]}.
+(Elter, Schulz-Wendtland, and Wittenberg 2007; Xu et al. 2008; Verma et
+al. 2015; Ng et al. 2015), using data aggregated from multiple health
+systems (Lee, Maslove, and Dubin 2015; Lee 2017; Tang et al. 2021; Ng et
+al. 2021), and in other care settings or disease contexts (Song and
+Kasabov 2006; Zhang et al. 2018; Tang et al. 2021; Ng et al. 2021).
 
@@
 and the users. Some authors urged the incorporation of (human-derived)
-domain knowledge into the data or models {[}17,19{]}. Others suggested
-new uses of their modeling approaches: to measure feature importance
-{[}20{]}, to make models more expressive {[}14{]}, and to combine
-information obtained both from global and from localized models
-{[}15{]}. Most of this incremental work was indeed carried out in later
+domain knowledge into the data or models (Yearwood and Wilkinson 1997;
+N. Wang et al. 2019). Others suggested new uses of their modeling
+approaches: to measure feature importance (Wyns et al. 2004), to make
+models more expressive (Lee, Maslove, and Dubin 2015), and to combine
+information obtained both from global and from localized models (Ng et
+al. 2015). Most of this incremental work was indeed carried out in later
 investigations in our sample. That said, we did not find reports of
@@
 conducted using clinical data, only occasionally together with
-patient-reported (2), laboratory (2), and image (1) data. We suggest
-three reasons for this: First, this literature traces back to the 1990s,
-before -omic data could be generated cost-effectively at scale. Second,
-CBR in particular has a strong tradition in clinical decision support,
-where the focus of our sample remains throughout the review period.
-Third, because most -omic data are highly homogeneous---all measurements
-are made along or can be transformed to a common scale, e.g.~greyscale
-pixellations for X-ray images and transcripts per million for RNA-seq
-data---more deeply theoretical analysis techniques have been developed
-and come into wide use. While variations on correlation-based approaches
-like EHR-based phenome-wide association studies and similarity-based
-methods like CBR itself have been developed, more mechanistic and
-probabilistic tools have not become a domain standard.
+patient-reported (2), laboratory (3), image (1), and environmental (1)
+data. We suggest three reasons for this: First, this literature traces
+back to the 1990s, before -omic data could be generated cost-effectively
+at scale. Second, CBR in particular has a strong tradition in clinical
+decision support, where the focus of our sample remains throughout the
+review period. Third, because most -omic data are highly
+homogeneous---all measurements are made along or can be transformed to a
+common scale, e.g.~greyscale pixellations for X-ray images and
+transcripts per million for RNA-seq data---more deeply theoretical
+analysis techniques have been developed and come into wide use. While
+variations on correlation-based approaches like EHR-based phenome-wide
+association studies and similarity-based methods like CBR itself have
+been developed, more mechanistic and probabilistic tools have not become
+a domain standard.
 
@@
 smaller contiguous programs, including applications of the TWNFI
-methodology {[}21,25{]}, individualized mortality prediction for ICU
-patients {[}14,16{]} and the use of more explanatory models to
-prioritize predictors or treatments for chronic disease {[}15,34,35{]}.
+methodology (Song and Kasabov 2006; Verma et al. 2015), individualized
+mortality prediction for ICU patients (Lee, Maslove, and Dubin 2015; Lee
+2017) and the use of more explanatory models to prioritize predictors or
+treatments for chronic disease (Ng et al. 2015; Tang et al. 2021; Ng et
+al. 2021).
 
@@
 common term is in use for what we here refer to as localized modeling;
-authors described their approaches as ``targeted prognosis'' {[}11{]},
-``transductive inference'' {[}21{]}, ``personalized decision support''
-{[}14{]}, ``personalized (predictive) models'' {[}26, in contrast to
-``local models,''15,17,32{]}, and ``precision cohort'' analysis
-{[}17,34,35{]}. We find uses of ``targeted'', ``personalized'', and
-``precision'' generic and imprecise to this approach, while transductive
-inference is an established term for a broader set of methods in ML.
-Because \emph{local} naturally contrasts with \emph{global}, we propose
-``localized models'' as a suitable term for this counterpart to
-globally-fitted models.\footnote{We note, in response to one reviewer's
-  comment, that for many if not most models used in ML, excepting GLMs,
-  a predictor may play a different and more or less important role for
-  some cases than others, as quantified by importance measures or
-  model-agnostic explanations {[}39,40{]}. In this way, such models can
-  also be viewed as ``localized''. We recognize that the use of global
-  and local explanations is standard. As another reviewer pointed out,
-  the terminology of local and global is also used in the unrelated
+authors described their approaches as ``targeted prognosis'' (Mariuzzi
+et al. 1997), ``transductive inference'' (Song and Kasabov 2006),
+``personalized decision support'' (Lee, Maslove, and Dubin 2015),
+``personalized (predictive) models'' (Liang, Hu, and Kasabov 2015, in
+contrast to ``local models''; Ng et al. 2015; N. Wang et al. 2019; Ma et
+al. 2020; Liu et al. 2022; Doborjeh et al. 2022), and ``precision
+cohort'' analysis (N. Wang et al. 2019; Tang et al. 2021; Ng et al.
+2021). We find uses of ``targeted'', ``personalized'', and ``precision''
+generic and imprecise, while transductive inference is an established
+term for a broader set of methods in ML. Because \emph{local} naturally
+contrasts with \emph{global}, we propose ``localized models'' as a
+suitable term for this counterpart to globally-fitted models.\footnote{We
+  note, in response to one reviewer's comment, that for many if not most
+  models used in ML, excepting GLMs, a predictor may play a different
+  and more or less important role for some cases than others, as
+  quantified by importance measures or model-agnostic explanations
+  (Biecek and Burzykowski 2021; Molnar 2023). In this way, such models
+  can also be viewed as ``localized''. We recognize that the use of
+  global and local explanations is standard. As another reviewer pointed
+  out, the terminology of local and global is also used in the unrelated
   context of federated ML to distinguish models trained using data
   stored in a single storage device (local) from their aggregations
-  (global) {[}41,42{]}. The prospects for federated architecture to more
-  effectively implement similarity cohort--based models are intriguing,
-  and would require some reconciliation of terms. At present, we feel
-  that these methods are sufficiently disjoint to avoid confusion in
-  practice, but we suggest the more specific term ``similarity
-  cohort--based models'' when the term ``local'' is overloaded.}
+  (global) (Moshawrab et al. 2023; Brauneck et al. 2023). The prospects
+  for federated architecture to more effectively implement similarity
+  cohort--based models are intriguing, and would require some
+  reconciliation of terms. At present, we feel that these methods are
+  sufficiently disjoint to avoid confusion in practice, but we suggest
+  the more specific term ``similarity cohort--based models'' when the
+  term ``local'' is overloaded.}
 
@@
 
-The 25 studies included in our synthesis used composite techniques that
+The 27 studies included in our synthesis used composite techniques that
 we organized into two major types. One type, \emph{similarity learning},
 was used in 8 studies that tie in to the much larger literature on DML
-{[}43{]}. The other, \emph{cohort thresholding}, focused on choosing or
-optimizing the manner in which cohorts were demarcated using the
+(Yang 2006). The other, \emph{cohort thresholding}, focused on choosing
+or optimizing the manner in which cohorts were demarcated using the
 similarity measure.
@@
 associations between predictors and responses, and to use this structure
-to inform the definition of a patient similarity measure. Seven studies
+to inform the definition of a patient similarity measure. Eight studies
 used response values in training data sets to supervise similarity
@@
 \ref{tab:composite}). Based on the key properties of DML algorithms
-identified by {[}44{]}, most learned measures were non-linear while some
-were locally learned, and some were optimized globally while others
-locally. Not all qualified as DML, since the resulting measures would
-not necessarily satisfy the triangle inequality. Each used its
-similarity measure to retrieve training cases relevant to each testing
-case. Once retrieved, most took a standard NN approach to generating
-predictions; two exceptions {[}29,34{]} fit predictive models to the
-retrieved cohorts.
+identified by Bellet, Habrard, and Sebban (2014), most learned measures
+were non-linear while some were locally learned, and some were optimized
+globally while others locally. Not all qualified as DML, since the
+resulting measures would not necessarily satisfy the triangle
+inequality. Each used its similarity measure to retrieve training cases
+relevant to each testing case. Once retrieved, most took a standard NN
+approach to generating predictions; two exceptions (Vilhena et al. 2016;
+Tang et al. 2021) fit predictive models to the retrieved cohorts.
 
 Several studies used unsupervised similarity learning, for example the
-Mahalanobis distance {[}12{]} and the Kolmogorov entropy-based distance
-{[}22{]}. In particular, {[}19{]} defined similarity as a weighted sum
-of differences in predictor values and used linear regression to
-optimize the weights for the predictive accuracy of the cohort they
-retrieve. Others used supervised learning: {[}21{]} proposed an
-iterative algorithm to optimize a set of fuzzy inference rules used to
-retrieve relevant cases and generate a prediction, using
-back-propagation on the rules' parameters. Their method was used in
-several later studies returned by our search, which are not discussed
-here because they did not originate the technique. {[}28{]} explicitly
-appeal to a supervised distance metric learning (DML) technique {[}45{]}
-to obtain a similarity measure on their set of dermoscopy and confocal
-images that most effectively separates malignant from benign melanoma
-tumors. We will discuss the remaining uses of supervised similarity
-learning shortly.
+Mahalanobis distance (Lowsky et al. 2013) and the Kolmogorov
+entropy-based distance (Elter, Schulz-Wendtland, and Wittenberg 2007).
+In particular, Yearwood and Wilkinson (1997) defined similarity as a
+weighted sum of differences in predictor values and used linear
+regression to optimize the weights for the predictive accuracy of the
+cohort they retrieve. Others used supervised learning: Song and Kasabov
+(2006) proposed an iterative algorithm to optimize a set of fuzzy
+inference rules used to retrieve relevant cases and generate a
+prediction, using back-propagation on the rules' parameters. Their
+method was used in several later studies returned by our search, which
+are not discussed here because they did not originate the technique.
+Nicolas et al. (2014) explicitly appeal to a supervised distance metric
+learning (DML) technique (Xing et al. 2002) to obtain a similarity
+measure on their set of dermoscopy and confocal images that most
+effectively separates malignant from benign melanoma tumors. We will
+discuss the remaining uses of supervised similarity learning shortly.
 
 Distinct from but related to similarity learning was supervised cohort
-construction. This was an adaptive step taken by {[}13{]} to optimize
-the predictive performance of models fitted to cohorts retrieved using a
-fixed similarity rather than count threshold, a counterpart to
-conventional CBR they called ``statistical CBR''.\footnote{Drawing from
-  {[}46{]}, we suggest the term ``combinatorial CBR'' for the
-  then-conventional approach using similarity cohorts of fixed
-  cardinality.} The step used a heuristic procedure to locate a
-similarity threshold that (locally) maximizes predictive accuracy on the
-training set. This is analogous to optimizing the neighborhood size
-parameter in NN predictive modeling, so we will not discuss it further
-except to mention that the authors found statistical CBR to outperform
-conventional CBR on several data sets. {[}27{]} combined these learning
-techniques: They used logistic regression on the training set to obtain
-weights for the predictors (by predicting the binary outcome of kidney
-transplant waitlist registration) and for the cases (by predicting
-agreement of outcome between an index case and other training cases),
-and they used exhaustion to optimize the size of the retrieved cohort
-for the accuracy of the NN model using the previously optimized weights.
+construction. This was an adaptive step taken by Park, Kim, and Chun
+(2006) to optimize the predictive performance of models fitted to
+cohorts retrieved using a fixed similarity rather than count threshold,
+a counterpart to conventional CBR they called ``statistical
+CBR''.\footnote{Drawing from Goyal, Lifshits, and Schütze (2008), we
+  suggest the term ``combinatorial CBR'' for the then-conventional
+  approach using similarity cohorts of fixed cardinality.} The step used
+a heuristic procedure to locate a similarity threshold that (locally)
+maximizes predictive accuracy on the training set. This is analogous to
+optimizing the neighborhood size parameter in NN predictive modeling, so
+we will not discuss it further except to mention that the authors found
+statistical CBR to outperform conventional CBR on several data sets.
+Campillo-Gimenez et al. (2013) combined these learning techniques: They
+used logistic regression on the training set to obtain weights for the
+predictors (by predicting the binary outcome of kidney transplant
+waitlist registration) and for the cases (by predicting agreement of
+outcome between an index case and other training cases), and they used
+exhaustion to optimize the size of the retrieved cohort for the accuracy
+of the NN model using the previously optimized weights.
 
@@
 includes some properly included techniques we did not have in mind
-{[}29{]} yet excludes other techniques commonly referred to as
-individualized, personalized, or precision.
+(Vilhena et al. 2016) yet excludes other techniques commonly referred to
+as individualized, personalized, or precision.
 
@@
 
-In addition to {[}12{]}, {[}26{]}, {[}15{]}, {[}16{]}, {[}29{]},
-{[}34{]}, and {[}35{]}, five other studies employed localized models:
-{[}11{]}, {[}14{]}, {[}25{]}, {[}17{]}, and {[}32{]}. In most cases
-these models were purely predictive in application; of the exceptions,
-{[}11{]} produced purely descriptive models of survival outcomes, which
-were evaluated for their precision rather than their accuracy, while
-{[}15{]} used generalized regression models descriptively as well as
-predictively, as did some of the same authors {[}34,35{]} later. The
-most common approach to cohort selection was to optimize a size
-threshold via manual exploration {[}11,14--17,29{]} or via
-cross-validation {[}12,25{]}. Further in the former direction, one study
-{[}32{]} set a fixed cohort size while two {[}26,34,35{]} devised more
-sophisticated algorithms to balance multiple desiderata including
-predictive accuracy.
+In addition to Lowsky et al. (2013), Liang, Hu, and Kasabov (2015), Ng
+et al. (2015), Lee (2017), Vilhena et al. (2016), Tang et al. (2021),
+and Ng et al. (2021), seven other studies employed localized models:
+Mariuzzi et al. (1997), Lee, Maslove, and Dubin (2015), Verma et al.
+(2015), N. Wang et al. (2019), Ma et al. (2020), Liu et al. (2022), and
+Doborjeh et al. (2022). In most cases these models were purely
+predictive in application; of the exceptions, Mariuzzi et al. (1997)
+produced purely descriptive models of survival outcomes, which were
+evaluated for their precision rather than their accuracy, while Ng et
+al. (2015) used generalized regression models descriptively as well as
+predictively, as did Tang et al. (2021), Ng et al. (2021), and Liu et
+al. (2022) later. The most common approach to cohort selection was to
+optimize a size threshold via manual exploration (Mariuzzi et al. 1997;
+Lee, Maslove, and Dubin 2015; Ng et al. 2015; Lee 2017; N. Wang et al.
+2019; Vilhena et al. 2016) or via cross-validation (Lowsky et al. 2013;
+Verma et al. 2015). Further in the former direction, one study (Ma et
+al. 2020) set a fixed cohort size while two (Liang, Hu, and Kasabov
+2015; Tang et al. 2021; Ng et al. 2021) devised more sophisticated
+algorithms to balance multiple desiderata including predictive accuracy.
 
@@
 of retrieved cohort sizes for some measure of model performance. Despite
-the early recommendation by {[}13{]} to threshold cohorts by similarity
-rather than by cardinality, only the one previously published study
-{[}11{]} and the most recent study {[}34,35{]} in this corpus took this
-approach. Viewed as a hyperparameter of the individualized modeling
-approach, cohort size can be treated in the same way as neighborhood
-size in NN prediction (viewed here as a special case of localized
-modeling wherein the model is a simple summary statistic), so we will
-not discuss it further.
+the early recommendation by Park, Kim, and Chun (2006) to threshold
+cohorts by similarity rather than by cardinality, only the one
+previously published study (Mariuzzi et al. 1997) and the most recent
+study (Tang et al. 2021; Ng et al. 2021) in this corpus took this
+approach, though Liu et al. (2022) took a third option by retrieving
+fixed proportions of training data. Viewed as a hyperparameter of the
+individualized modeling approach, cohort size can be treated in the same
+way as neighborhood size in NN prediction (viewed here as a special case
+of localized modeling wherein the model is a simple summary statistic),
+so we will not discuss it further.
 
 An alternative approach was to optimize multiple hyperparameters
-together: {[}18{]} proposed an iterative algorithm to settle on an
-optimal number and set of predictors as well as number of training cases
-(comprising the retrieved cohort) that was later used by {[}26{]}. More
-recently, {[}34{]} and {[}35{]} proposed a three-step process for cohort
-selection that filtered by exact match for one subset of predictors,
-used domain-informed similarity measures to rank these, and optimized
-the similarity threshold for a trade-off between cohort size and a
-measure of bias called cohort balance. Notably, though their similarity
-measure was supervised, this optimization process was not.
-
-With the exception of the descriptive survival models of {[}11{]}, only
-one thread in this corpus concerned itself with the interpretability of
-localized models: {[}15{]} fit logistic regression models to
-similarity-based cohorts and examined not only their predictive
-performance but the localized sets of largest and most detectable
-predictors, termed risk profiles, and how they differed across the
-testing population. This approach informed their later use of localized
-comparative effectiveness--style models to recommend courses of
-treatment at decision points during a monitored patent's stay
-{[}34,35{]}. These studies suggest much wider potential for localized
+together: Kasabov and Hu (2010) proposed an iterative algorithm to
+settle on an optimal number and set of predictors as well as number of
+training cases (comprising the retrieved cohort) that was later used by
+Liang, Hu, and Kasabov (2015). More recently, Tang et al. (2021) and Ng
+et al. (2021) proposed a three-step process for cohort selection that
+filtered by exact match for one subset of predictors, used
+domain-informed similarity measures to rank these, and optimized the
+similarity threshold for a trade-off between cohort size and a measure
+of bias called cohort balance. Notably, though their similarity measure
+was supervised, this optimization process was not.
+
+With the exception of the descriptive survival models of Mariuzzi et al.
+(1997), only one thread in this corpus concerned itself with the
+interpretability of localized models: Ng et al. (2015) fit logistic
+regression models to similarity-based cohorts and examined not only
+their predictive performance but the localized sets of largest and most
+detectable predictors, termed risk profiles, and how they differed
+across the testing population. This approach informed their later use of
+localized comparative effectiveness--style models to recommend courses
+of treatment at decision points during a monitored patent's stay (Tang
+et al. 2021; Ng et al. 2021), and it was later used by Liu et al. (2022)
+to show a dependency of predictor importance on the patient disease
+subgroup. These studies suggest much wider potential for localized
 real-world evidence generation, which has long been promoted as a
-promise of advances in clinical and health informatics {[}1{]}.
+promise of advances in clinical and health informatics (Longhurst,
+Harrington, and Shah 2014).
 
@@
 these approaches has been summarized above and published in greater
-detail in previous reviews {[}6,9,10{]}, and we focus here on the
+detail in previous reviews (Choudhury and Begum 2016; Sharafoddini,
+Dubin, and Lee 2017; Parimbelli et al. 2018), and we focus here on the
 conceptual. The design of a localized modeling approach can be
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-typology}
+{\centering \includegraphics[width=1\linewidth]{Fig5}
 
@@
 
-\caption{General framework for localized models.}\label{fig:framework}
+\caption{General framework for localized models. Dashed arrows indicate optional steps.}\label{fig:framework}
 \end{figure}
@@
 supervised, in this sample none were. Almost all models were supervised,
-being that most performed predictive tasks, though {[}11{]} modeled only
-survival rates within localized cohorts. Meanwhile, each step can be
-optimized in a ML fashion by having its parameters tuned to improve
-performance. We found no studies that tuned the calculation of
-similarity, though most uniform cohort sizes were tuned. Some earlier
-studies tuned generalized regression models, but more recent studies did
-not.
+being that most performed predictive tasks, though Mariuzzi et al.
+(1997) modeled only survival rates within localized cohorts. Meanwhile,
+each step can be optimized in a ML fashion by having its parameters
+tuned to improve performance. We found one study that tuned the
+calculation of similarity (Liu et al. 2022), though most uniform cohort
+sizes were also tuned. Some earlier studies tuned generalized regression
+models, but more recent studies did not.
+
+\small
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2087}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2795}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1535}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3583}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2851}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1566}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3655}}@{}}
 \caption{\label{tab:framework}Specializations of general framework to
@@
 \endlastfoot
-{[}11{]} & Levenshtein distance\hspace{18em} & Statistical\hspace{18em}
-& Survival (S)\hspace{18em} \\
-{[}19{]} & Weighted sum (S)\hspace{18em} & Combinatorial\hspace{18em} &
-Proportion (S)\hspace{18em} \\
-{[}20{]} & Kohonen mapping\hspace{18em} & Adaptive\hspace{18em} &
-Representation (S)\hspace{18em} \\
-{[}13{]} & Euclidean distance\hspace{18em} & Statistical
-(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
-{[}22{]} & Entropic distance\hspace{18em} & Combinatorial\hspace{18em} &
-Proportion (S)\hspace{18em} \\
-{[}23{]} & Euclidean distance\hspace{18em} & Combinatorial\hspace{18em}
-& Proportion (S)\hspace{18em} \\
-{[}18,21,25,26{]} & Feature selection (R,S)\hspace{18em} & Combinatorial
-(R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
-{[}24{]} & User-defined\hspace{18em} & User-defined\hspace{18em} &
-User-defined\hspace{18em} \\
-{[}12{]} & Mahalanobis distance (S)\hspace{18em} & Combinatorial
-(T)\hspace{18em} & Proportional hazards (S)\hspace{18em} \\
-{[}27{]} & Weighted overlap\hspace{18em} & Combinatorial
-(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
-{[}28{]} & Distance metric learning (S)\hspace{18em} &
+Mariuzzi et al. (1997) & Levenshtein distance\hspace{18em} &
+Statistical\hspace{18em} & Survival (S)\hspace{18em} \\
+Yearwood and Wilkinson (1997) & Weighted sum (S)\hspace{18em} &
+Combinatorial\hspace{18em} & Proportion (S)\hspace{18em} \\
+Wyns et al. (2004) & Kohonen mapping\hspace{18em} &
+Adaptive\hspace{18em} & Representation (S)\hspace{18em} \\
+Park, Kim, and Chun (2006) & Euclidean distance\hspace{18em} &
+Statistical (T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
+Elter, Schulz-Wendtland, and Wittenberg (2007) & Entropic
+distance\hspace{18em} & Combinatorial\hspace{18em} & Proportion
+(S)\hspace{18em} \\
+Xu et al. (2008) & Euclidean distance\hspace{18em} &
+Combinatorial\hspace{18em} & Proportion (S)\hspace{18em} \\
+Song and Kasabov (2006); Kasabov and Hu (2010); Liang, Hu, and Kasabov
+(2015); Verma et al. (2015) & Feature selection (R,S)\hspace{18em} &
+Combinatorial (R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
+López et al. (2011) & User-defined\hspace{18em} &
+User-defined\hspace{18em} & User-defined\hspace{18em} \\
+Lowsky et al. (2013) & Mahalanobis distance (S)\hspace{18em} &
+Combinatorial (T)\hspace{18em} & Proportional hazards
+(S)\hspace{18em} \\
+Campillo-Gimenez et al. (2013) & Weighted overlap\hspace{18em} &
+Combinatorial (T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
+Nicolas et al. (2014) & Distance metric learning (S)\hspace{18em} &
 Combinatorial\hspace{18em} & Association rules (S,T)\hspace{18em} \\
-{[}15{]} & Locally supervised metric learning (S); Feature
+Ng et al. (2015) & Locally supervised metric learning (S); Feature
 selection\hspace{18em} & Combinatorial\hspace{18em} & Logistic model
 (S)\hspace{18em} \\
-{[}14{]} & Cosine similarity\hspace{18em} & Combinatorial
-(T)\hspace{18em} & Logistic model (S); Decision tree (S)\hspace{18em} \\
-{[}29{]} & Range overlap\hspace{18em} & Clustering (S)\hspace{18em} &
-Artificial neural network (S)\hspace{18em} \\
-{[}16{]} & Cosine similarity; Random forest proximity (S)\hspace{18em} &
-Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree (S);
-Random forest (S)\hspace{18em} \\
-{[}31{]} & Unclear\hspace{18em} & Unclear\hspace{18em} &
-Unclear\hspace{18em} \\
-{[}30{]} & Gaussian process model\hspace{18em} & Complete\hspace{18em} &
-Weighted support vector machine (S); Weighted Gaussian process
-regression (S)\hspace{18em} \\
-{[}17{]} & Weighted sum\hspace{18em} & Combinatorial (T)\hspace{18em} &
-Logistic model (S); Random forest (S); Nearest neighbors
+Lee, Maslove, and Dubin (2015) & Cosine similarity\hspace{18em} &
+Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree
 (S)\hspace{18em} \\
-{[}32{]} & Weighted sum\hspace{18em} & Combinatorial\hspace{18em} &
-Extreme learning machine (S)\hspace{18em} \\
-{[}33{]} & Weighted sum\hspace{18em} & Statistical\hspace{18em} & Linear
-model\hspace{18em} \\
-{[}34,35{]} & Locally supervised metric learning (S)\hspace{18em} &
-Causal inference matching\hspace{18em} & Statistical test
+Vilhena et al. (2016) & Range overlap\hspace{18em} & Clustering
+(S)\hspace{18em} & Artificial neural network (S)\hspace{18em} \\
+Lee (2017) & Cosine similarity; Random forest proximity (S)\hspace{18em}
+& Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree
+(S); Random forest (S)\hspace{18em} \\
+Malykh and Rudetskiy (2018) & Unclear\hspace{18em} &
+Unclear\hspace{18em} & Unclear\hspace{18em} \\
+Zhang et al. (2018) & Gaussian process model\hspace{18em} &
+Complete\hspace{18em} & Weighted support vector machine (S); Weighted
+Gaussian process regression (S)\hspace{18em} \\
+N. Wang et al. (2019) & Weighted sum\hspace{18em} & Combinatorial
+(T)\hspace{18em} & Logistic model (S); Random forest (S); Nearest
+neighbors (S)\hspace{18em} \\
+Ma et al. (2020) & Weighted sum\hspace{18em} &
+Combinatorial\hspace{18em} & Extreme learning machine
 (S)\hspace{18em} \\
+Y. Wang et al. (2020) & Weighted sum\hspace{18em} &
+Statistical\hspace{18em} & Linear model\hspace{18em} \\
+Tang et al. (2021); Ng et al. (2021) & Locally supervised metric
+learning (S)\hspace{18em} & Causal inference matching\hspace{18em} &
+Statistical test (S)\hspace{18em} \\
+Doborjeh et al. (2022) & Euclidean distance; Signal-to-noise ratio
+(S)\hspace{18em} & Statistical (T)\hspace{18em} & Spiking neural network
+(S)\hspace{18em} \\
+Liu et al. (2022) & Weighted Manhattan distance (R,T)\hspace{18em} &
+Proportional (T)\hspace{18em} & Logistic model (S)\hspace{18em} \\
 \end{longtable}
 
-\hypertarget{directions-and-expectations}{%
-\subsection{Directions and
-expectations}\label{directions-and-expectations}}
+\normalsize
+
+\hypertarget{directions-and-expectations-for-future-work}{%
+\subsection{Directions and expectations for future
+work}\label{directions-and-expectations-for-future-work}}
+
+\hypertarget{performance}{%
+\subsubsection{Performance}\label{performance}}
 
@@
 the potential predictive value of localized models. As cautioned in
-Section \ref{performance-assessments}, not all experiments affirmed this
-premise. Most of those that did involved using metric learning to
-improve retrieval and sometimes iterative optimization of the metric and
-of the localized model. Some studies showcased results using a variety
-of such specializations {[}15,27,30{]}. However, because these advanced
+Section \ref{evaluations}, not all experiments affirmed this premise.
+Most of those that did involved using metric learning to improve
+retrieval and sometimes iterative optimization of the metric and of the
+localized model. Some studies showcased results using a variety of such
+specializations (Campillo-Gimenez et al. 2013; Ng et al. 2015; Zhang et
+al. 2018; Liu et al. 2022). However, because these advanced
 implementations were mostly compared against their precursors or
@@
 future work.
+
+The value of this work would be enhanced by standardization and
+modularization. First, studies using localized models should locate the
+proposed approach and any variations in a shared specification space.
+For this purpose, we propose the framework outlined above, at least as a
+starting point. When a specification invokes complex techniques, for
+example supervised DML or neural network classifiers, it should be
+compared at least to alternative specifications using simpler
+alternatives. Campillo-Gimenez et al. (2013) and Liu et al. (2022)
+provide excellent examples in which all possible choices are toggled and
+yield a neatly nested comparison. Additionally, some state-of-the-art
+models, including others that obtain individualized predictions, should
+be included in comparisons. Finally, implementations should be not only
+published on public repositories but designed in such a way that users
+familiar with the underlying language can substitute a technique of
+their choosing for any part of the specification (similarity, retrieval,
+fitting, evaluation). We believe these principles will make reported
+results clearer and reproduction and extension easier.
+
+\hypertarget{interpretability}{%
+\subsubsection{Interpretability}\label{interpretability}}
 
@@
 model components could help alleviate, but this is more often assumed
-than demonstrated. One valuable direction for future work would be to
-measure the utility of interpretable components in research and in
-clinical practice and the correctness and confidence of users in their
-interpretations.
+than demonstrated.
+
+One valuable direction for future work would be to measure the utility
+of interpretable components in research and in clinical practice and the
+correctness and confidence of users in their interpretations. This is a
+necessary precondition for practical use but would also be a valuable
+contribution to the experimental literature. Another would be to compare
+the localized predictor importance measures obtained from localized
+models to the local importance measures used to explain predictions made
+by opaque models. This would test both measures for concurrent validity,
+while differences between them would inform what settings or cases are
+poorly served by one or the other.
+
+\hypertarget{feasibility}{%
+\subsubsection{Feasibility}\label{feasibility}}
+
+Much of the reviewed work was motivated by the need for modeling
+paradigms that perform well on a variety of tasks and in a variety of
+settings. This need demands not only methodologies but also
+architectures that are robust, versatile, and compliant in the face of
+diverse data models and use restrictions, yet no studies in our sample
+addressed the challenges of system interoperability and regulatory
+requirements head-on.
+
+CBR and, by extension, localized modeling are especially susceptible to
+these challenges, as corpora of past cases must be aggregated from
+multiple institutions and from multiple systems within institutions, and
+models built on them must then be applicable to out-of-box data
+structures as well. Individual systems exhibit many dimensions of
+incompleteness (Weiskopf et al. 2013) and vary along other dimensions of
+quality (Kohane et al. 2021), and their aggregation for modeling
+purposes depends on several kinds of interoperability (Weber 2015).
+Successful deployment also depends on satisfying regulatory regimes
+designed to protect the privacy of patients, the security of
+communications, and trust between parties (Haendel et al. 2021). While
+most studies in our sample focused on improving care, increasingly many
+over time focused on generalizable knowledge. Therefore, as studies
+using localized modeling shift from proof of concept to feasibility,
+they will also need to demonstrate practical and regulatory feasibility.
+
+\hypertarget{customizability}{%
+\subsubsection{Customizability}\label{customizability}}
 
@@
 the level of other predictive models or strategies. From a functionalist
-approach to reproducibility {[}47{]}, then, later studies were broadly
-successful at reproducing earlier studies, and later studies provide
-sufficient detail to support ongoing reproduction efforts---though these
-might be limited by the lack of open-source code or public
-implementations. However, the immediate goal of most studes was to
-improve care, and an implied need, often enjoined but neither obtained
-nor reproduced, was a demonstration of practical use.
+approach to reproducibility (Matarese 2022), then, later studies were
+broadly successful at reproducing earlier studies, and later studies
+provide sufficient detail to support ongoing reproduction
+efforts---though these might be limited by the lack of open-source code
+or public implementations. However, the immediate goal of most studes
+was to improve care, and an implied need, often enjoined but neither
+obtained nor reproduced, was a demonstration of practical use.
 
@@
 to down-weight socioeconomic variables like race--ethnicity in order to
-ensure a more diverse modeling cohort. Future purely quantitative work
-could assess whether similarity-tuning can achieve these ends more
-flexibly than strict inclusion/exclusion criteria.
+ensure a more diverse modeling cohort. López et al. (2011) took an
+important step in this direction with an adjustable and adaptable
+implementation. Future purely quantitative work could assess whether
+similarity-tuning can achieve these ends more flexibly than strict
+inclusion/exclusion criteria.
 
@@
 date suggest great potential for growth. Whereas precious few of the
-reviewed studies used these models, for any task other than prediction,
+reviewed studies used these models for any task other than prediction,
 despite widespread suspicion among clinicians of ``black-box'' models
@@
 
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-We are grateful for the astute and supportive comments of three
-anonymous reviewers.
-
-\hypertarget{support}{%
-\section{Support}\label{support}}
-
-The authors received no specific funding or other support for this work.
-
-\hypertarget{competing-interests}{%
-\section{Competing interests}\label{competing-interests}}
-
-The authors declare that they have no competing interests.
-
-\hypertarget{data-and-code}{%
-\section{Data and code}\label{data-and-code}}
-
-Search results are publicly available in a Zotero group library. Data
-collected or encoded for included studies are publicly available in two
-Google Sheets. Code used to analyze these data and generate this
-manuscript is publicly available in a GitHub repository.
-
-\begin{itemize}
-\tightlist
-\item
-  Search results, included studies, and other bibliography:
-  \url{https://www.zotero.org/groups/5017571/imsr/}
-\item
-  Bibliographic and methodological properties of included studies:
-  \url{https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/}
-\item
-  Terminology and composite techniques of included studies:
-  \url{https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/}
-\item
-  Code used to conduct analyses and prepare the manuscript:
-  \url{https://github.com/corybrunson/imsr}
-\end{itemize}
+\pagebreak
 
 \hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
 \hypertarget{research-question}{%
-\subsection{Research question}\label{research-question}}
+\subsection*{Research question}\label{research-question}}
+\addcontentsline{toc}{subsection}{Research question}
 
@@
 \hypertarget{purpose-of-review}{%
-\subsection{Purpose of review}\label{purpose-of-review}}
+\subsection*{Purpose of review}\label{purpose-of-review}}
+\addcontentsline{toc}{subsection}{Purpose of review}
 
@@
 \hypertarget{search-design}{%
-\subsection{Search design}\label{search-design}}
+\subsection*{Search design}\label{search-design}}
+\addcontentsline{toc}{subsection}{Search design}
 
@@
 \hypertarget{search-strings}{%
-\subsubsection{Search strings}\label{search-strings}}
+\subsubsection*{Search strings}\label{search-strings}}
+\addcontentsline{toc}{subsubsection}{Search strings}
 
@@
 our literature search. We first finalized the \textbf{PubMed} search
-string below:\footnote{\url{https://pubmed.ncbi.nlm.nih.gov/?term=(+\%22case-based+reasoning\%22+\%5BAll+Fields\%5D+OR+\%22case-based+system\%22+\%5BAll+Fields\%5D+)+OR+(+\%22individualized+modeling\%22+\%5BAll+Fields\%5D+OR+\%22personalized+modeling\%22+\%5BAll+Fields\%5D+OR+\%22customized+modeling\%22+\%5BAll+Fields\%5D+)+OR+(+\%22individualized+cohort\%22+\%5BAll+Fields\%5D+)+OR+(+(+\%22patient+similarity\%22+\%5BAll+Fields\%5D+OR+\%22patient+distance\%22+\%5BAll+Fields\%5D+OR+\%22patient+connection\%22+\%5BAll+Fields\%5D+OR+\%22patient+affinity\%22+\%5BAll+Fields\%5D+OR+\%22patient+clustering\%22+\%5BAll+Fields\%5D+)+AND+(+\%22cohort+study\%22+\%5BAll+Fields\%5D+)+)\&sort=date}}
+string below:\footnote{\url{https://pubmed.ncbi.nlm.nih.gov/?term=(+"case-based+reasoning"+[All+Fields]+OR+"case-based+system"+[All+Fields]+)+OR+(+"individualized+modeling"+[All+Fields]+OR+"personalized+modeling"+[All+Fields]+OR+"customized+modeling"+[All+Fields]+)+OR+(+"individualized+cohort"+[All+Fields]+)+OR+(+(+"patient+similarity"+[All+Fields]+OR+"patient+distance"+[All+Fields]+OR+"patient+connection"+[All+Fields]+OR+"patient+affinity"+[All+Fields]+OR+"patient+clustering"+[All+Fields]+)+AND+(+"cohort+study"+[All+Fields]+)+)&sort=date}}
 
@@
 \hypertarget{screening-process}{%
-\subsection{Screening process}\label{screening-process}}
+\subsection*{Screening process}\label{screening-process}}
+\addcontentsline{toc}{subsection}{Screening process}
 
@@
 \hypertarget{selection-process}{%
-\subsection{Selection process}\label{selection-process}}
+\subsection*{Selection process}\label{selection-process}}
+\addcontentsline{toc}{subsection}{Selection process}
 
@@
 
+\hypertarget{update}{%
+\subsection*{Update}\label{update}}
+\addcontentsline{toc}{subsection}{Update}
+
+Of the 25 studies included, 4 showed up in no database searches (but
+included from the seed set), 1 was obtained via reference tracking, and,
+of the remaining 20 found through the database searches, only 1 was not
+found in PubMed or Web of Science. That one was instead found in
+Mathematical Reviews. Since MR returns about the same volume of results
+as WoS and PM, we omitted it from the update.
+
+We applied the search terms exactly as before but restricted the dates
+to 2021 July 19 (the beginning of the date range for our original
+searches) or later. We omitted the WoS search that returned
+impractically many results the first time around and was omitted then.
+
+We ran the update search on 2024 January 24.
+
+The PubMed update returned 97 results. The Web of Science refresher
+returned 272 + 0 + 26 = 298 results. Covidence removed duplicates, so
+that 348 studies were slated for screening.
+
+Screening by title and abstract was done in two waves. First, PM
+excluded only on account of a study not being (a) clinical/medical, (b)
+English, and (c) original. Then, JCB then excluded on account of a study
+being (d) indicated by the intended meanings of our search terms.
+
 \hypertarget{analysis-and-synthesis}{%
-\subsection{Analysis and synthesis}\label{analysis-and-synthesis}}
+\subsection*{Analysis and synthesis}\label{analysis-and-synthesis}}
+\addcontentsline{toc}{subsection}{Analysis and synthesis}
 
@@
 
+\hypertarget{methodological-elements}{%
+\subsection*{Methodological elements}\label{methodological-elements}}
+\addcontentsline{toc}{subsection}{Methodological elements}
+
+Table \ref{tab:composite} and Figure \ref{fig:methods} summarize the
+techniques and terminology used by the included studies, as referred to
+in the main text.
+
+\small
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1082}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4433}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4485}}@{}}
+\caption{\label{tab:composite}Methodological elements of studies
+included in the synthesis.}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Elements
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Terminology
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+Citation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Elements
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Terminology
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Yearwood and Wilkinson (1997) & supervised similarity learning & CBR;
+case-structured retrieval \\
+Mariuzzi et al. (1997) & predictive modeling on similarity cohorts &
+CBR; targeted subsampling \\
+Wyns et al. (2004) & unsupervised similarity learning & Combined Kohonen
+type neural network-case-based reasoning \\
+Park, Kim, and Chun (2006) & supervised cohort construction & CBR,
+statistical case-based reasoning \\
+Song and Kasabov (2006) & supervised similarity learning & Transductive
+inference \\
+Elter, Schulz-Wendtland, and Wittenberg (2007) & unsupervised similarity
+learning & CBR; entropic distance measure \\
+Xu et al. (2008) & tiered constraint; similarity matching & CBR \\
+López et al. (2011) & interactive implementation & CBR; eXiT*CBR \\
+Kasabov and Hu (2010) & predictive modeling on similarity cohorts &
+Personalized model \\
+Verma et al. (2015) & predictive modeling on similarity cohorts &
+Personalized modeling; TWNFI \\
+Liang, Hu, and Kasabov (2015) & predictive modeling on similarity
+cohorts & Local modeling; personalized modeling; TWNFI \\
+Lowsky et al. (2013) & predictive modeling on similarity cohorts &
+K-nearest neighbors survival \\
+Campillo-Gimenez et al. (2013) & supervised similarity learning & CBR \\
+Nicolas et al. (2014) & supervised similarity learning & CBR \\
+Ng et al. (2015) & predictive modeling on similarity cohorts;
+explanatory modeling on similarity cohorts & Personalized predictive
+model \\
+Lee, Maslove, and Dubin (2015) & predictive modeling on similarity
+cohorts & Personalized prediction \\
+Vilhena et al. (2016) & supervised similarity learning; predictive
+modeling on similarity cohorts & CBR \\
+Lee (2017) & predictive modeling on similarity cohorts &
+Patient-specific predictive model \\
+Zhang et al. (2018) & whole-population weight learning & Gaussian
+processes \\
+Malykh and Rudetskiy (2018) & unclear & CBR \\
+Ma et al. (2020) & predictive modeling on similarity cohorts &
+Personalized model \\
+N. Wang et al. (2019) & predictive modeling on similarity cohorts &
+Personalized predictive modeling \\
+Y. Wang et al. (2020) & descriptive modeling on similarity cohorts &
+CBR \\
+Liu et al. (2022) & NA & Personalized model with transfer learning \\
+Doborjeh et al. (2022) & NA & Personalized spiking neural network \\
+Tang et al. (2021); Ng et al. (2021) & supervised similarity learning;
+predictive modeling on similarity cohorts & Similarity model,
+personalized model, precision cohort, personalized treatment options \\
+\end{longtable}
+
+\normalsize
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Fig4}
+
+}
+
+\caption{Frequency of recurring methodological elements across included studies.}\label{fig:methods}
+\end{figure}
+
 \hypertarget{performance-evaluations-and-comparisons}{%
-\subsection{Performance evaluations and
+\subsection*{Performance evaluations and
 comparisons}\label{performance-evaluations-and-comparisons}}
+\addcontentsline{toc}{subsection}{Performance evaluations and
+comparisons}
+
+Table \ref{tab:performance} summarizes the evaluations and comparisons
+made of methods proposed in the included studies. In most cases, not all
+performance results are included; for the sake of concision, only what
+we judged to be the headline results, using the primary evaluation
+metrics, are reproduced here. The table contains no original results.
 
@@
 
-\textbf{TODO: Place vertical separators between major columns.}
+%\textbf{TODO: Place vertical separators between major columns.}
+
+\small
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1224}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0714}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1633}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2398}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0969}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2092}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0969}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1088}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0725}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1658}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2435}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2124}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}}@{}}
 \caption{\label{tab:performance}Evaluations of proposed methods and
@@
 \endlastfoot
-{[}20{]} & Accuracy & Arthritis & Kohonen + CBR & 75.60\%\hspace{6em} &
-Kohonen mapping & 54.70\%\hspace{6em} \\
+Wyns et al. (2004) & Accuracy & Arthritis & Kohonen + CBR &
+75.60\%\hspace{6em} & Kohonen mapping & 54.70\%\hspace{6em} \\
 & & & & \hspace{6em} & CBR & 60.40\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & backpropagation & 54.70\%\hspace{6em} \\
-{[}13{]} & Accuracy & Dermatology & Statistical CBR &
+\midrule\noalign{}
+Park, Kim, and Chun (2006) & Accuracy & Dermatology & Statistical CBR &
 96.57\%\hspace{6em} & CBR & 95.43\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & NN & 55.59\%\hspace{6em} \\
-{[}21{]} & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
+\midrule\noalign{}
+Song and Kasabov (2006) & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
 7.74\hspace{6em} \\
@@
 & & & & \hspace{6em} & TNFI & 7.28\hspace{6em} \\
-{[}22{]} & AUROC & mammography mass region & CBR & 0.885\hspace{6em} &
-DT & 0.872\hspace{6em} \\
+\midrule\noalign{}
+Elter, Schulz-Wendtland, and Wittenberg (2007) & AUROC & mammography
+mass region & CBR & 0.885\hspace{6em} & DT & 0.872\hspace{6em} \\
 & & & & \hspace{6em} & ANN & 0.882\hspace{6em} \\
-{[}23{]} & Accuracy & Return-to-work & CBR & 62.5\%\hspace{6em} & LR &
-71.2\%\hspace{6em} \\
-{[}18{]} & Accuracy & Colon cancer diagnosis & TWNFI &
+\midrule\noalign{}
+Xu et al. (2008) & Accuracy & Return-to-work & CBR & 62.5\%\hspace{6em}
+& LR & 71.2\%\hspace{6em} \\
+\midrule\noalign{}
+Kasabov and Hu (2010) & Accuracy & Colon cancer diagnosis & TWNFI &
 91.9\%\hspace{6em} & MLR (Personalized) & 82.3\%\hspace{6em} \\
@@
 \hspace{6em} \\
-{[}26{]} & Accuracy & Colon cancer & knnGSA & 85.48\%\hspace{6em} & SVM
-& 82.14\%\hspace{6em} \\
+\midrule\noalign{}
+Liang, Hu, and Kasabov (2015) & Accuracy & Colon cancer & knnGSA &
+85.48\%\hspace{6em} & SVM & 82.14\%\hspace{6em} \\
 & & & svmGSA & 87.10\%\hspace{6em} & ECF & 72.30\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & WKNN & 95.50\%\hspace{6em} \\
-{[}12{]} & Relative IPEC & Graft survival & MKNN & 0.973\hspace{6em} &
-RSF & 0.957\hspace{6em} \\
-{[}27{]} & AUROC & Waitlist registration, complete & CBR + Wj &
-89.8\%\hspace{6em} & LR & 92.0\%\hspace{6em} \\
+\midrule\noalign{}
+Lowsky et al. (2013) & Relative IPEC & Graft survival & MKNN &
+0.973\hspace{6em} & RSF & 0.957\hspace{6em} \\
+\midrule\noalign{}
+Campillo-Gimenez et al. (2013) & AUROC & Waitlist registration, complete
+& CBR + Wj & 89.8\%\hspace{6em} & LR & 92.0\%\hspace{6em} \\
 & & & CBR + Wk & 90.7\%\hspace{6em} & Standalone CBR &
@@
 & & & CBR + Wj + Wk & 88.5\%\hspace{6em} & & \hspace{6em} \\
-{[}28{]} & Accuracy & Nevus & Collaborative + rules & 98\%\hspace{6em} &
-Dermoscopy CBR & 92\%\hspace{6em} \\
+\midrule\noalign{}
+Nicolas et al. (2014) & Accuracy & Nevus & Collaborative + rules &
+98\%\hspace{6em} & Dermoscopy CBR & 92\%\hspace{6em} \\
 & & & Collaborative + rules + DML & 100\%\hspace{6em} & Confocal CBR &
@@
 & & & & \hspace{6em} & Collaborative & 95\%\hspace{6em} \\
-{[}15{]} & AUROC & Diabetes onset & Personalized LR
+\midrule\noalign{}
+Ng et al. (2015) & AUROC & Diabetes onset & Personalized LR
 (LSML+FeatureFiltering) & 0.624\hspace{6em} & KNN & 0.617\hspace{6em} \\
@@
 & & & Personalized LR (Random) & 0.602\hspace{6em} & & \hspace{6em} \\
-{[}14{]} & AUROC & 30-day mortality & Individualized LR &
-0.830\hspace{6em} & Individualized DC & 0.797\hspace{6em} \\
+\midrule\noalign{}
+Lee, Maslove, and Dubin (2015) & AUROC & 30-day mortality &
+Individualized LR & 0.830\hspace{6em} & Individualized DC &
+0.797\hspace{6em} \\
 & & & Individualized DT & 0.753\hspace{6em} & & \hspace{6em} \\
-{[}16{]} & AUROC & & Individualized LR & 0.824\hspace{6em} &
+\midrule\noalign{}
+Lee (2017) & AUROC & & Individualized LR & 0.824\hspace{6em} &
 Individualized DC & 0.801\hspace{6em} \\
@@
 & & & Individualized CSRF & 0.832\hspace{6em} & & \hspace{6em} \\
-{[}31{]} & Accuracy & J13/pneumonia & Case-based NN & 81.6\%\hspace{6em}
-& & \hspace{6em} \\
+\midrule\noalign{}
+Malykh and Rudetskiy (2018) & Accuracy & J13/pneumonia & Case-based NN &
+81.6\%\hspace{6em} & & \hspace{6em} \\
 & & K80.1/gallbladder & Case-based NN & 76.7\%\hspace{6em} & &
@@
 \hspace{6em} \\
-{[}30{]} & AUROC & Cognitive impairment & Gaussian process regression
-with custom kernel & 0.92\hspace{6em} & linear regression without
-regularization & 0.81\hspace{6em} \\
+\midrule\noalign{}
+Zhang et al. (2018) & AUROC & Cognitive impairment & Gaussian process
+regression with custom kernel & 0.92\hspace{6em} & linear regression
+without regularization & 0.81\hspace{6em} \\
 & & & & \hspace{6em} & LASSO regression & 0.85\hspace{6em} \\
@@
 0.778\hspace{6em} \\
-{[}17{]} & AUROC & Diabetes & Personalized RF & 0.90\hspace{6em} & &
-\hspace{6em} \\
+\midrule\noalign{}
+N. Wang et al. (2019) & AUROC & Diabetes & Personalized RF &
+0.90\hspace{6em} & & \hspace{6em} \\
 & & & Personalized KNN & 0.82\hspace{6em} & & \hspace{6em} \\
 & & & Personalized LR & 0.89\hspace{6em} & & \hspace{6em} \\
-{[}32{]} & AUROC & Discharge at 10 days & One-class JITL-ELM &
+\midrule\noalign{}
+Ma et al. (2020) & AUROC & Discharge at 10 days & One-class JITL-ELM &
 0.8510\hspace{6em} & ELM & 0.4973\hspace{6em} \\
@@
 & & & & \hspace{6em} & One-class SVM & 0.4647\hspace{6em} \\
+\midrule\noalign{}
+Liu et al. (2022) & AUROC & Acute kidney injury & & \hspace{6em} &
+Global & 0.684\hspace{6em} \\
+& & & & \hspace{6em} & Risk-ranking & 0.708\hspace{6em} \\
+& & & & \hspace{6em} & PM-Kmeans & 0.721\hspace{6em} \\
+& & & & \hspace{6em} & PM-Kmeans \& TL & 0.759\hspace{6em} \\
+& & & & \hspace{6em} & PM-kNN & 0.732\hspace{6em} \\
+& & & & \hspace{6em} & PM-kNN \& WS & 0.738\hspace{6em} \\
+& & & & \hspace{6em} & PM-kNN \& TL & 0.769\hspace{6em} \\
+& & & & \hspace{6em} & PM-kNN \& TL \& WS & 0.771\hspace{6em} \\
+& & & & \hspace{6em} & PM-kNN \& WS \& SL & 0.744\hspace{6em} \\
+& & & & \hspace{6em} & PMTL & 0.779\hspace{6em} \\
+\midrule\noalign{}
+Doborjeh et al. (2022) & Accuracy & Stroke & Personalized SNN &
+60.70\%\hspace{6em} & & \hspace{6em} \\
 \end{longtable}
 
+\normalsize
+
 \end{landscape}
+
+\hypertarget{data-and-code}{%
+\subsection{Data and code}\label{data-and-code}}
+
+Search results are publicly available in a Zotero group library. Data
+collected or encoded for included studies are publicly available in two
+Google Sheets. Code used to analyze these data and generate this
+manuscript is publicly available in a GitHub repository.
+
+\begin{itemize}
+\tightlist
+\item
+  Search results, included studies, and other bibliography:
+  \url{https://www.zotero.org/groups/5017571/imsr/}
+\item
+  Bibliographic and methodological properties of included studies:
+  \url{https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/}
+\item
+  Terminology and composite techniques of included studies:
+  \url{https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/}
+\item
+  Code used to conduct analyses and prepare the manuscript:
+  \url{https://github.com/corybrunson/imsr}
+\end{itemize}
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Aamodt1994}{}}%
+Aamodt, Agnar, and Enric Plaza. 1994. {``Case-{Based Reasoning}:
+{Foundational Issues}, {Methodological Variations}, and {System
+Approaches}.''} \emph{AI Communications} 7 (1): 39--59.
+\url{https://doi.org/10.3233/AIC-1994-7104}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Begum2011}{}}%
+Begum, Shahina, Mobyen Uddin Ahmed, Peter Funk, Ning Xiong, and Mia
+Folke. 2011. {``Case-{Based Reasoning Systems} in the {Health Sciences}:
+{A Survey} of {Recent Trends} and {Developments}.''} \emph{IEEE
+Transactions on Systems, Man, and Cybernetics, Part C (Applications and
+Reviews)} 41 (4): 421--34.
+\url{https://doi.org/10.1109/TSMCC.2010.2071862}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
+Bellet, Aurélien, Amaury Habrard, and Marc Sebban. 2014. {``A {Survey}
+on {Metric Learning} for {Feature Vectors} and {Structured Data}.''}
+arXiv. \url{https://doi.org/10.48550/arXiv.1306.6709}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
+Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
+{Model Analysis}: {Explore}, {Explain}, and {Examine Predictive
+Models}}. New York: {Chapman and Hall/CRC}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
+Brauneck, Alissa, Louisa Schmalhorst, Mohammad Mahdi Kazemi Majdabadi,
+Mohammad Bakhtiari, Uwe Völker, Jan Baumbach, Linda Baumbach, and
+Gabriele Buchholtz. 2023. {``Federated {Machine Learning},
+{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
+Research}: {Scoping Review}.''} \emph{Journal of Medical Internet
+Research} 25 (1): e41588. \url{https://doi.org/10.2196/41588}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2013}{}}%
+Campillo-Gimenez, Boris, Wassim Jouini, Sahar Bayat, and Marc Cuggia.
+2013. {``Improving {Case-Based Reasoning Systems} by {Combining
+K-Nearest Neighbour Algorithm} with {Logistic Regression} in the
+{Prediction} of {Patients}' {Registration} on the {Renal Transplant
+Waiting List}.''} Edited by Randen Lee Patterson. \emph{PLOS ONE} 8 (9):
+e71991. \url{https://doi.org/10.1371/journal.pone.0071991}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Choudhury2016}{}}%
+Choudhury, Nabanita, and Shahin Ara Begum. 2016. {``A {Survey} on
+{Case-based Reasoning} in {Medicine}.''} \emph{International Journal of
+Advanced Computer Science and Applications} 7 (8).
+\url{https://doi.org/10.14569/IJACSA.2016.070820}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
+Dai, Leyu, He Zhu, and Dianbo Liu. 2020. {``Patient Similarity: Methods
+and Applications.''} \emph{arXiv:2012.01976 {[}Cs{]}}, December.
+\url{https://arxiv.org/abs/2012.01976}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Doborjeh2022}{}}%
+Doborjeh, Maryam, Zohreh Doborjeh, Alexander Merkin, Rita Krishnamurthi,
+Reza Enayatollahi, Valery Feigin, and Nikola Kasabov. 2022.
+{``Personalized {Spiking Neural Network Models} of {Clinical} and
+{Environmental Factors} to {Predict Stroke}.''} \emph{Cognitive
+Computation} 14 (6): 2187--2202.
+\url{https://doi.org/10.1007/s12559-021-09975-x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006}{}}%
+Doyle, Dónal, Pádraig Cunningham, and Paul Walsh. 2006. {``An Evaluation
+of the Usefulness of Explanation in a Case-Based Reasoning System for
+Decision Support in Bronchiolitis Treatment.''} \emph{Computational
+Intelligence} 22 (3-4): 269--81.
+\url{https://doi.org/10.1111/j.1467-8640.2006.00288.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Elter2007}{}}%
+Elter, M., R. Schulz-Wendtland, and T. Wittenberg. 2007. {``The
+Prediction of Breast Cancer Biopsy Outcomes Using Two {CAD} Approaches
+That Both Emphasize an Intelligible Decision Process.''} \emph{Medical
+Physics} 34 (11): 4164--72. \url{https://doi.org/10.1118/1.2786864}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fang2021}{}}%
+Fang, Hao Sen Andrew, Ngiap Chuan Tan, Wei Ying Tan, Ronald Wihal Oei,
+Mong Li Lee, and Wynne Hsu. 2021. {``Patient Similarity Analytics for
+Explainable Clinical Risk Prediction.''} \emph{BMC Medical Informatics
+\& Decision Making} 21 (1): 1--12.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
+Gierl, Lothar, Mathias Bull, and Rainer Schmidt. 1998. {``{CBR} in
+{Medicine}.''} In \emph{Case-{Based Reasoning Technology}}, edited by
+Mario Lenz, Hans-Dieter Burkhard, Brigitte Bartsch-Spörl, and Stefan
+Wess, 1400:273--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
+\url{https://doi.org/10.1007/3-540-69351-3_11}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
+Goyal, Navin, Yury Lifshits, and Hinrich Schütze. 2008. {``Disorder
+Inequality: A Combinatorial Approach to Nearest Neighbor Search.''} In
+\emph{Proceedings of the 2008 {International Conference} on {Web Search}
+and {Data Mining}}, 25--32. {WSDM} '08. New York, NY, USA: Association
+for Computing Machinery. \url{https://doi.org/10.1145/1341531.1341538}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Haendel2021}{}}%
+Haendel, Melissa A, Christopher G Chute, Tellen D Bennett, David A
+Eichmann, Justin Guinney, Warren A Kibbe, Philip R O Payne, et al. 2021.
+{``The {National COVID Cohort Collaborative} ({N3C}): {Rationale},
+Design, Infrastructure, and Deployment.''} \emph{Journal of the American
+Medical Informatics Association} 28 (3): 427--43.
+\url{https://doi.org/10.1093/jamia/ocaa196}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kasabov2010}{}}%
+Kasabov, Nikola, and Yingjie Hu. 2010. {``Integrated Optimisation Method
+for Personalised Modelling and Case Studies for Medical Decision
+Support.''} \emph{International Journal of Functional Informatics and
+Personalised Medicine} 3 (3): 236--56.
+\url{https://doi.org/10.1504/IJFIPM.2010.039123}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kohane2021}{}}%
+Kohane, Isaac S., Bruce J. Aronow, Paul Avillach, Brett K.
+Beaulieu-Jones, Riccardo Bellazzi, Robert L. Bradford, Gabriel A. Brat,
+et al. 2021. {``What {Every Reader Should Know About Studies Using
+Electronic Health Record Data} but {May Be Afraid} to {Ask}.''}
+\emph{Journal of Medical Internet Research} 23 (3): e22219.
+\url{https://doi.org/10.2196/22219}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kolodner1992}{}}%
+Kolodner, Janet L. 1992. {``An Introduction to Case-Based Reasoning.''}
+\emph{Artificial Intelligence Review} 6 (1): 3--34.
+\url{https://doi.org/10.1007/BF00155578}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lee2017}{}}%
+Lee, Joon. 2017. {``Patient-{Specific Predictive Modeling Using Random
+Forests}: {An Observational Study} for the {Critically Ill}.''}
+\emph{JMIR Medical Informatics} 5 (1): e3.
+\url{https://doi.org/10.2196/medinform.6690}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lee2015}{}}%
+Lee, Joon, David M. Maslove, and Joel A. Dubin. 2015. {``Personalized
+{Mortality Prediction Driven} by {Electronic Medical Data} and a
+{Patient Similarity Metric}.''} Edited by Frank Emmert-Streib.
+\emph{PLOS ONE} 10 (5): e0127428.
+\url{https://doi.org/10.1371/journal.pone.0127428}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Liang2015}{}}%
+Liang, W, YJ Hu, and N Kasabov. 2015. {``Evolving Personalized Modeling
+System for Integrated Feature, Neighborhood and Parameter Optimization
+Utilizing Gravitational Search Algorithm.''} \emph{Evolving Systems} 6
+(1): 1--14. \url{https://doi.org/10.1007/s12530-013-9081-x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Liu2022}{}}%
+Liu, Kang, Xiangzhou Zhang, Weiqi Chen, Alan S. L. Yu, John A. Kellum,
+Michael E. Matheny, Steven Q. Simpson, Yong Hu, and Mei Liu. 2022.
+{``Development and {Validation} of a {Personalized Model With Transfer
+Learning} for {Acute Kidney Injury Risk Estimation Using Electronic
+Health Records}.''} \emph{JAMA Network Open} 5 (7): e2219776.
+\url{https://doi.org/10.1001/jamanetworkopen.2022.19776}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Longhurst2014}{}}%
-\CSLLeftMargin{{[}1{]} }%
-\CSLRightInline{Longhurst CA, Harrington RA, Shah NH. A {``{Green
-Button}''} {For Using Aggregate Patient Data At The Point Of Care}.
-Health Affairs 2014;33:1229--35.
-\url{https://doi.org/10.1377/hlthaff.2014.0099}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aamodt1994}{}}%
-\CSLLeftMargin{{[}2{]} }%
-\CSLRightInline{Aamodt A, Plaza E. Case-{Based Reasoning}: {Foundational
-Issues}, {Methodological Variations}, and {System Approaches}. AI
-Communications 1994;7:39--59.
-\url{https://doi.org/10.3233/AIC-1994-7104}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Begum2011}{}}%
-\CSLLeftMargin{{[}3{]} }%
-\CSLRightInline{Begum S, Ahmed MU, Funk P, Xiong N, Folke M. Case-{Based
-Reasoning Systems} in the {Health Sciences}: {A Survey} of {Recent
-Trends} and {Developments}. IEEE Transactions on Systems, Man, and
-Cybernetics, Part C (Applications and Reviews) 2011;41:421--34.
-\url{https://doi.org/10.1109/TSMCC.2010.2071862}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kolodner1992}{}}%
-\CSLLeftMargin{{[}4{]} }%
-\CSLRightInline{Kolodner JL. An introduction to case-based reasoning.
-Artificial Intelligence Review 1992;6:3--34.
-\url{https://doi.org/10.1007/BF00155578}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
-\CSLLeftMargin{{[}5{]} }%
-\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
-Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
-Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
-Heidelberg}; 1998, p. 273--97.
-\url{https://doi.org/10.1007/3-540-69351-3_11}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Choudhury2016}{}}%
-\CSLLeftMargin{{[}6{]} }%
-\CSLRightInline{Choudhury N, Begum SA. A {Survey} on {Case-based
-Reasoning} in {Medicine}. International Journal of Advanced Computer
-Science and Applications 2016;7.
-\url{https://doi.org/10.14569/IJACSA.2016.070820}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
-\CSLLeftMargin{{[}7{]} }%
-\CSLRightInline{Dai L, Zhu H, Liu D.
-\href{http://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
-applications}. arXiv:201201976 {[}Cs{]} 2020.}
+Longhurst, Christopher A., Robert A. Harrington, and Nigam H. Shah.
+2014. {``A {`{Green Button}'} {For Using Aggregate Patient Data At The
+Point Of Care}.''} \emph{Health Affairs} 33 (7): 1229--35.
+\url{https://doi.org/10.1377/hlthaff.2014.0099}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011}{}}%
+López, Beatriz, Carles Pous, Pablo Gay, Albert Pla, Judith Sanz, and
+Joan Brunet. 2011. {``{eXiT}*{CBR}: {A} Framework for Case-Based Medical
+Diagnosis Development and Experimentation.''} \emph{Artificial
+Intelligence in Medicine} 51 (2): 81--91.
+\url{https://doi.org/10.1016/j.artmed.2010.09.002}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lowsky2013}{}}%
+Lowsky, D. J., Y. Ding, D. K. K. Lee, C. E. McCulloch, L. F. Ross, J. R.
+Thistlethwaite, and S. A. Zenios. 2013. {``A {K-nearest} Neighbors
+Survival Probability Prediction Method.''} \emph{Statistics in Medicine}
+32 (12): 2062--69. \url{https://doi.org/10.1002/sim.5673}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ma2020}{}}%
+Ma, X, YB Si, ZF Wang, and YQ Wang. 2020. {``Length of Stay Prediction
+for {ICU} Patients Using Individualized Single Classification
+Algorithm.''} \emph{Computer Methods and Programs in Biomedicine} 186
+(April). \url{https://doi.org/10.1016/j.cmpb.2019.105224}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018}{}}%
+Malykh, V. L., and S. V. Rudetskiy. 2018. {``Approaches to {Medical
+Decision-Making Based} on {Big Clinical Data}.''} \emph{Journal of
+Healthcare Engineering} 2018: 3917659.
+\url{https://doi.org/10.1155/2018/3917659}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
+Mariuzzi, G., A. Mombello, L. Mariuzzi, P. W. Hamilton, J. E. Weber, D.
+Thompson, and P. H. Bartels. 1997. {``Quantitative Study of Ductal
+Breast Cancer--Patient Targeted Prognosis: An Exploration of Case Base
+Reasoning.''} \emph{Pathology - Research and Practice} 193 (8): 535--42.
+\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
+Matarese, Vera. 2022. {``Kinds of {Replicability}: {Different Terms} and
+{Different Functions}.''} \emph{Axiomathes} 32 (2): 647--70.
+\url{https://doi.org/10.1007/s10516-021-09610-2}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
+Molnar, Christoph. 2023. \emph{Interpretable {Machine Learning}: {A
+Guide} for {Making Black Box Models Explainable}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
+Moshawrab, Mohammad, Mehdi Adda, Abdenour Bouzouane, Hussein Ibrahim,
+and Ali Raad. 2023. {``Reviewing {Federated Machine Learning} and {Its
+Use} in {Diseases Prediction}.''} \emph{Sensors} 23 (4): 2112.
+\url{https://doi.org/10.3390/s23042112}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ng2021}{}}%
+Ng, Kenney, Uri Kartoun, Harry Stavropoulos, John A Zambrano, and Paul C
+Tang. 2021. {``Personalized Treatment Options for Chronic Diseases Using
+Precision Cohort Analytics.''} \emph{Scientific Reports} 11 (1): 1--13.
+\url{https://doi.org/10.1038/s41598-021-80967-5}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ng2015}{}}%
+Ng, Kenney, Jimeng Sun, Jianying Hu, and Fei Wang. 2015.
+{``\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4525240}{Personalized
+{Predictive Modeling} and {Risk Factor Identification} Using {Patient
+Similarity}}.''} \emph{AMIA Joint Summits on Translational Science
+Proceedings. AMIA Joint Summits on Translational Science} 2015: 132--36.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014}{}}%
+Nicolas, Ruben, Albert Fornells, Elisabet Golobardes, Guiomar Corral,
+Susana Puig, and Josep Malvehy. 2014. {``{DERMA}: A Melanoma Diagnosis
+Platform Based on Collaborative Multilabel Analog Reasoning.''}
+\emph{The Scientific World Journal} 2014: 351518.
+\url{https://doi.org/10.1155/2014/351518}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Parimbelli2018}{}}%
+Parimbelli, E., S. Marini, L. Sacchi, and R. Bellazzi. 2018. {``Patient
+Similarity for Precision Medicine: {A} Systematic Review.''}
+\emph{Journal of Biomedical Informatics} 83 (July): 87--96.
+\url{https://doi.org/10.1016/j.jbi.2018.06.001}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Park2006}{}}%
+Park, Yoon-Joo, Byung-Chun Kim, and Se-Hak Chun. 2006. {``New Knowledge
+Extraction Technique Using Probability for Case-Based Reasoning:
+Application to Medical Diagnosis.''} \emph{Expert Systems} 23 (1):
+2--20. \url{https://doi.org/10.1111/j.1468-0394.2006.00321.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Sharafoddini2017}{}}%
+Sharafoddini, Anis, Joel A Dubin, and Joon Lee. 2017. {``Patient
+{Similarity} in {Prediction Models Based} on {Health Data}: {A Scoping
+Review}.''} \emph{JMIR Medical Informatics} 5 (1): e7.
+\url{https://doi.org/10.2196/medinform.6730}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Song2006}{}}%
+Song, Qun, and Nikola Kasabov. 2006. {``{TWNFI} --- a Transductive
+Neuro-Fuzzy Inference System with Weighted Data Normalization for
+Personalized Modeling.''} \emph{Neural Networks} 19 (10): 1591--96.
+\url{https://doi.org/10.1016/j.neunet.2006.05.028}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tang2021}{}}%
+Tang, Paul C, Sarah Miller, Harry Stavropoulos, Uri Kartoun, John
+Zambrano, and Kenney Ng. 2021. {``Precision Population Analytics:
+Population Management at the Point-of-Care.''} \emph{Journal of the
+American Medical Informatics Association} 28 (3): 588--95.
+\url{https://doi.org/10.1093/jamia/ocaa247}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Verma2015}{}}%
+Verma, A, M Fiasche, M Cuzzola, and G Irrera. 2015. {``Comparative Study
+of Existing Personalized Approaches for Identifying Important Gene
+Markers and for Risk Estimation in {Type2 Diabetes} in {Italian}
+Population.''} \emph{Evolving Systems} 6 (1): 15--22.
+\url{https://doi.org/10.1007/s12530-013-9083-8}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016}{}}%
+Vilhena, João, Henrique Vicente, M. Rosário Martins, José M. Grañeda,
+Filomena Caldeira, Rodrigo Gusmão, João Neves, and José Neves. 2016.
+{``A Case-Based Reasoning View of Thrombophilia Risk.''} \emph{Journal
+of Biomedical Informatics} 62 (August): 265--75.
+\url{https://doi.org/10.1016/j.jbi.2016.07.013}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wang2019}{}}%
+Wang, Ni, Yanqun Huang, Honglei Liu, Xiaolu Fei, Lan Wei, Xiangkun Zhao,
+and Hui Chen. 2019. {``Measurement and Application of Patient Similarity
+in Personalized Predictive Modeling Based on Electronic Medical
+Records.''} \emph{BioMedical Engineering OnLine} 18 (1).
+\url{https://doi.org/10.1186/s12938-019-0718-2}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
+Wang, Yuning, Yingzi Liang, Hui Sun, and Yufei Yang. 2020. {``Emergency
+{Response} for {COVID-19 Prevention} and {Control} in {Urban Rail
+Transit Based} on {Case-Based Reasoning Method}.''} \emph{Discrete
+Dynamics in Nature \& Society}, December, 1--9.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Weber2015}{}}%
+Weber, Griffin M. 2015. {``Federated Queries of Clinical Data
+Repositories: {Scaling} to a National Network.''} \emph{Journal of
+Biomedical Informatics} 55 (June): 231--36.
+\url{https://doi.org/10.1016/j.jbi.2015.04.012}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Weiskopf2013}{}}%
+Weiskopf, Nicole G., George Hripcsak, Sushmita Swaminathan, and Chunhua
+Weng. 2013. {``Defining and Measuring Completeness of Electronic Health
+Records for Secondary Use.''} \emph{Journal of Biomedical Informatics}
+46 (5): 830--36. \url{https://doi.org/10.1016/j.jbi.2013.06.010}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Welch2013}{}}%
-\CSLLeftMargin{{[}8{]} }%
-\CSLRightInline{Welch BM, Kawamoto K. Clinical decision support for
-genetically guided personalized medicine: A systematic review. Journal
-of the American Medical Informatics Association 2013;20:388--400.
-\url{https://doi.org/10.1136/amiajnl-2012-000892}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Sharafoddini2017}{}}%
-\CSLLeftMargin{{[}9{]} }%
-\CSLRightInline{Sharafoddini A, Dubin JA, Lee J. Patient {Similarity} in
-{Prediction Models Based} on {Health Data}: {A Scoping Review}. JMIR
-Medical Informatics 2017;5:e7.
-\url{https://doi.org/10.2196/medinform.6730}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Parimbelli2018}{}}%
-\CSLLeftMargin{{[}10{]} }%
-\CSLRightInline{Parimbelli E, Marini S, Sacchi L, Bellazzi R. Patient
-similarity for precision medicine: {A} systematic review. Journal of
-Biomedical Informatics 2018;83:87--96.
-\url{https://doi.org/10.1016/j.jbi.2018.06.001}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
-\CSLLeftMargin{{[}11{]} }%
-\CSLRightInline{Mariuzzi G, Mombello A, Mariuzzi L, Hamilton PW, Weber
-JE, Thompson D, et al. Quantitative study of ductal breast
-cancer--patient targeted prognosis: An exploration of case base
-reasoning. Pathology - Research and Practice 1997;193:535--42.
-\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lowsky2013}{}}%
-\CSLLeftMargin{{[}12{]} }%
-\CSLRightInline{Lowsky DJ, Ding Y, Lee DKK, McCulloch CE, Ross LF,
-Thistlethwaite JR, et al. A {K-nearest} neighbors survival probability
-prediction method. Statistics in Medicine 2013;32:2062--9.
-\url{https://doi.org/10.1002/sim.5673}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Park2006}{}}%
-\CSLLeftMargin{{[}13{]} }%
-\CSLRightInline{Park Y-J, Kim B-C, Chun S-H. New knowledge extraction
-technique using probability for case-based reasoning: Application to
-medical diagnosis. Expert Systems 2006;23:2--20.
-\url{https://doi.org/10.1111/j.1468-0394.2006.00321.x}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lee2015}{}}%
-\CSLLeftMargin{{[}14{]} }%
-\CSLRightInline{Lee J, Maslove DM, Dubin JA. Personalized {Mortality
-Prediction Driven} by {Electronic Medical Data} and a {Patient
-Similarity Metric}. PLOS ONE 2015;10:e0127428.
-\url{https://doi.org/10.1371/journal.pone.0127428}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ng2015}{}}%
-\CSLLeftMargin{{[}15{]} }%
-\CSLRightInline{Ng K, Sun J, Hu J, Wang F.
-\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4525240}{Personalized
-{Predictive Modeling} and {Risk Factor Identification} using {Patient
-Similarity}}. AMIA Joint Summits on Translational Science Proceedings
-AMIA Joint Summits on Translational Science 2015;2015:132--6.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lee2017}{}}%
-\CSLLeftMargin{{[}16{]} }%
-\CSLRightInline{Lee J. Patient-{Specific Predictive Modeling Using
-Random Forests}: {An Observational Study} for the {Critically Ill}. JMIR
-Medical Informatics 2017;5:e3.
-\url{https://doi.org/10.2196/medinform.6690}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wang2019}{}}%
-\CSLLeftMargin{{[}17{]} }%
-\CSLRightInline{Wang N, Huang Y, Liu H, Fei X, Wei L, Zhao X, et al.
-Measurement and application of patient similarity in personalized
-predictive modeling based on electronic medical records. BioMedical
-Engineering OnLine 2019;18.
-\url{https://doi.org/10.1186/s12938-019-0718-2}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kasabov2010}{}}%
-\CSLLeftMargin{{[}18{]} }%
-\CSLRightInline{Kasabov N, Hu Y. Integrated optimisation method for
-personalised modelling and case studies for medical decision support.
-International Journal of Functional Informatics and Personalised
-Medicine 2010;3:236--56.
-\url{https://doi.org/10.1504/IJFIPM.2010.039123}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997a}{}}%
-\CSLLeftMargin{{[}19{]} }%
-\CSLRightInline{Yearwood J, Wilkinson R. Retrieving cases for treatment
-advice in nursing using text representation and structured text
-retrieval. Artificial Intelligence in Medicine 1997;9:79--99.
-\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004a}{}}%
-\CSLLeftMargin{{[}20{]} }%
-\CSLRightInline{Wyns B, Sette S, Boullart L, Baeten D, Hoffman IEA, De
-Keyser F. Prediction of diagnosis in patients with early arthritis using
-a combined {Kohonen} mapping and instance-based evaluation criterion.
-Artificial Intelligence in Medicine 2004;31:45--55.
-\url{https://doi.org/10.1016/j.artmed.2004.01.002}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Song2006c}{}}%
-\CSLLeftMargin{{[}21{]} }%
-\CSLRightInline{Song Q, Kasabov N. {TWNFI} \textemdash{} a transductive
-neuro-fuzzy inference system with weighted data normalization for
-personalized modeling. Neural Networks 2006;19:1591--6.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Elter2007a}{}}%
-\CSLLeftMargin{{[}22{]} }%
-\CSLRightInline{Elter M, Schulz-Wendtland R, Wittenberg T. The
-prediction of breast cancer biopsy outcomes using two {CAD} approaches
-that both emphasize an intelligible decision process. Medical Physics
-2007;34:4164--72. \url{https://doi.org/10.1118/1.2786864}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Xu2008a}{}}%
-\CSLLeftMargin{{[}23{]} }%
-\CSLRightInline{Xu Y, Chan CCH, Lo KHY-L, Tang D. Prediction model for
-the return to work of workers with injuries in {Hong Kong}. Work
-2008;30:77--84.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011a}{}}%
-\CSLLeftMargin{{[}24{]} }%
-\CSLRightInline{López B, Pous C, Gay P, Pla A, Sanz J, Brunet J.
-{eXiT}*{CBR}: {A} framework for case-based medical diagnosis development
-and experimentation. Artificial Intelligence in Medicine 2011;51:81--91.
-\url{https://doi.org/10.1016/j.artmed.2010.09.002}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
-\CSLLeftMargin{{[}25{]} }%
-\CSLRightInline{Verma A, Fiasche M, Cuzzola M, Irrera G. Comparative
-study of existing personalized approaches for identifying important gene
-markers and for risk estimation in {Type2 Diabetes} in {Italian}
-population. Evolving Systems 2015;6:15--22.
-\url{https://doi.org/10.1007/s12530-013-9083-8}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Liang2015a}{}}%
-\CSLLeftMargin{{[}26{]} }%
-\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
-system for integrated feature, neighborhood and parameter optimization
-utilizing gravitational search algorithm. Evolving Systems 2015;6:1--14.
-\url{https://doi.org/10.1007/s12530-013-9081-x}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2012a}{}}%
-\CSLLeftMargin{{[}27{]} }%
-\CSLRightInline{Campillo-Gimenez B, Bayat S, Cuggia M. Coupling
-{K-nearest} neighbors with logistic regression in case-based reasoning.
-Studies in Health Technology and Informatics 2012;180:275--9.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014a}{}}%
-\CSLLeftMargin{{[}28{]} }%
-\CSLRightInline{Nicolas R, Fornells A, Golobardes E, Corral G, Puig S,
-Malvehy J. {DERMA}: A melanoma diagnosis platform based on collaborative
-multilabel analog reasoning. The Scientific World Journal
-2014;2014:351518. \url{https://doi.org/10.1155/2014/351518}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016a}{}}%
-\CSLLeftMargin{{[}29{]} }%
-\CSLRightInline{Vilhena J, Vicente H, Martins MR, Grañeda JM, Caldeira
-F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk.
-Journal of Biomedical Informatics 2016;62:265--75.
-\url{https://doi.org/10.1016/j.jbi.2016.07.013}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
-\CSLLeftMargin{{[}30{]} }%
-\CSLRightInline{Zhang H, Zhu F, Dodge H, Higgins G, Omenn G, Guan Y, et
-al. A similarity-based approach to leverage multi-cohort medical data on
-the diagnosis and prognosis of {Alzheimer}'s disease. GigaScience
-2018;7. \url{https://doi.org/10.1093/gigascience/giy085}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
-\CSLLeftMargin{{[}31{]} }%
-\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
-Decision-Making Based} on {Big Clinical Data}. Journal of Healthcare
-Engineering 2018;2018:3917659.
-\url{https://doi.org/10.1155/2018/3917659}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ma2020a}{}}%
-\CSLLeftMargin{{[}32{]} }%
-\CSLRightInline{Ma X, Si Y, Wang Z, Wang Y. Length of stay prediction
-for {ICU} patients using individualized single classification algorithm.
-Computer Methods and Programs in Biomedicine 2020;186.
-\url{https://doi.org/10.1016/j.cmpb.2019.105224}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
-\CSLLeftMargin{{[}33{]} }%
-\CSLRightInline{Wang Y, Liang Y, Sun H, Yang Y. Emergency {Response} for
-{COVID-19 Prevention} and {Control} in {Urban Rail Transit Based} on
-{Case-Based Reasoning Method}. Discrete Dynamics in Nature \& Society
-2020:1--9.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tang2021c}{}}%
-\CSLLeftMargin{{[}34{]} }%
-\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
-Ng K. Precision population analytics: Population management at the
-point-of-care. Journal of the American Medical Informatics Association
-2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ng2021b}{}}%
-\CSLLeftMargin{{[}35{]} }%
-\CSLRightInline{Ng K, Kartoun U, Stavropoulos H, Zambrano J, Tang P.
-Personalized treatment options for chronic diseases using precision
-cohort analytics. Scientific Reports 2021;11.
-\url{https://doi.org/10.1038/s41598-021-80967-5}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006a}{}}%
-\CSLLeftMargin{{[}36{]} }%
-\CSLRightInline{Doyle D, Cunningham P, Walsh P. An evaluation of the
-usefulness of explanation in a case-based reasoning system for decision
-support in bronchiolitis treatment. Computational Intelligence
-2006;22:269--81.
-\url{https://doi.org/10.1111/j.1467-8640.2006.00288.x}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fang2021a}{}}%
-\CSLLeftMargin{{[}37{]} }%
-\CSLRightInline{Fang HSA, Tan NC, Tan WY, Oei RW, Lee ML, Hsu W. Patient
-similarity analytics for explainable clinical risk prediction. BMC
-Medical Informatics \& Decision Making 2021;21:1--12.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2013}{}}%
-\CSLLeftMargin{{[}38{]} }%
-\CSLRightInline{Campillo-Gimenez B, Jouini W, Bayat S, Cuggia M.
-Improving {Case-Based Reasoning Systems} by {Combining K-Nearest
-Neighbour Algorithm} with {Logistic Regression} in the {Prediction} of
-{Patients}' {Registration} on the {Renal Transplant Waiting List}. PLOS
-ONE 2013;8:e71991. \url{https://doi.org/10.1371/journal.pone.0071991}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
-\CSLLeftMargin{{[}39{]} }%
-\CSLRightInline{Biecek P, Burzykowski T.
-\href{https://ema.drwhy.ai/}{Explanatory {Model Analysis}: {Explore},
-{Explain}, and {Examine Predictive Models}}. {New York}: {Chapman and
-Hall/CRC}; 2021.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
-\CSLLeftMargin{{[}40{]} }%
-\CSLRightInline{Molnar C.
-\href{https://christophm.github.io/interpretable-ml-book}{Interpretable
-{Machine Learning}: {A Guide} for {Making Black Box Models
-Explainable}}. 2023.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
-\CSLLeftMargin{{[}41{]} }%
-\CSLRightInline{Moshawrab M, Adda M, Bouzouane A, Ibrahim H, Raad A.
-Reviewing {Federated Machine Learning} and {Its Use} in {Diseases
-Prediction}. Sensors 2023;23:2112.
-\url{https://doi.org/10.3390/s23042112}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
-\CSLLeftMargin{{[}42{]} }%
-\CSLRightInline{Brauneck A, Schmalhorst L, Majdabadi MMK, Bakhtiari M,
-Völker U, Baumbach J, et al. Federated {Machine Learning},
-{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
-Research}: {Scoping Review}. Journal of Medical Internet Research
-2023;25:e41588. \url{https://doi.org/10.2196/41588}.}
+Welch, Brandon M, and Kensaku Kawamoto. 2013. {``Clinical Decision
+Support for Genetically Guided Personalized Medicine: A Systematic
+Review.''} \emph{Journal of the American Medical Informatics
+Association} 20 (2): 388--400.
+\url{https://doi.org/10.1136/amiajnl-2012-000892}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004}{}}%
+Wyns, B., S. Sette, L. Boullart, D. Baeten, I. E. A. Hoffman, and F. De
+Keyser. 2004. {``Prediction of Diagnosis in Patients with Early
+Arthritis Using a Combined {Kohonen} Mapping and Instance-Based
+Evaluation Criterion.''} \emph{Artificial Intelligence in Medicine} 31
+(1): 45--55. \url{https://doi.org/10.1016/j.artmed.2004.01.002}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
+Xing, Eric, Michael Jordan, Stuart J Russell, and Andrew Ng. 2002.
+{``Distance {Metric Learning} with {Application} to {Clustering} with
+{Side-Information}.''} In \emph{Advances in {Neural Information
+Processing Systems}}. Vol. 15. MIT Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Xu2008}{}}%
+Xu, Yanwen, Chetwyn C. H. Chan, Karen Hui Yu-Ling Lo, and Dan Tang.
+2008. {``\href{https://www.ncbi.nlm.nih.gov/pubmed/18198444}{Prediction
+Model for the Return to Work of Workers with Injuries in {Hong
+Kong}.}''} \emph{Work} 30 (1): 77--84.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-\CSLLeftMargin{{[}43{]} }%
-\CSLRightInline{Yang L.
-\href{http://www.cs.cmu.edu/~./liuy/frame_survey_v2.pdf}{Distance
-{Metric Learning}: {A Comprehensive Survey}}. PhD thesis. Michigan State
-University, 2006.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-\CSLLeftMargin{{[}44{]} }%
-\CSLRightInline{Bellet A, Habrard A, Sebban M. A {Survey} on {Metric
-Learning} for {Feature Vectors} and {Structured Data} 2014.
-\url{https://doi.org/10.48550/arXiv.1306.6709}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-\CSLLeftMargin{{[}45{]} }%
-\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A.
-\href{https://proceedings.neurips.cc/paper/2002/hash/c3e4035af2a1cde9f21e1ae1951ac80b-Abstract.html}{Distance
-{Metric Learning} with {Application} to {Clustering} with
-{Side-Information}}. Advances in {Neural Information Processing
-Systems}, vol. 15, {MIT Press}; 2002.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-\CSLLeftMargin{{[}46{]} }%
-\CSLRightInline{Goyal N, Lifshits Y, Schütze H. Disorder inequality: A
-combinatorial approach to nearest neighbor search. Proceedings of the
-2008 {International Conference} on {Web Search} and {Data Mining}, {New
-York, NY, USA}: {Association for Computing Machinery}; 2008, p. 25--32.
-\url{https://doi.org/10.1145/1341531.1341538}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
-\CSLLeftMargin{{[}47{]} }%
-\CSLRightInline{Matarese V. Kinds of {Replicability}: {Different Terms}
-and {Different Functions}. Axiomathes 2022;32:647--70.
-\url{https://doi.org/10.1007/s10516-021-09610-2}.}
+Yang, Liu. 2006. {``Distance {Metric Learning}: {A Comprehensive
+Survey}.''} PhD thesis, Department of Computer Science; Engineering:
+Michigan State University.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997}{}}%
+Yearwood, J., and R. Wilkinson. 1997. {``Retrieving Cases for Treatment
+Advice in Nursing Using Text Representation and Structured Text
+Retrieval.''} \emph{Artificial Intelligence in Medicine} 9 (1): 79--99.
+\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018}{}}%
+Zhang, HJ, F Zhu, HH Dodge, GA Higgins, GS Omenn, YF Guan, and
+Alzheimer's Dis Neuroimaging Initi. 2018. {``A Similarity-Based Approach
+to Leverage Multi-Cohort Medical Data on the Diagnosis and Prognosis of
+{Alzheimer}'s Disease.''} \emph{GigaScience} 7 (7).
+\url{https://doi.org/10.1093/gigascience/giy085}.
 
@@
 
+\bibliographystyle{unsrt}
+\bibliography{SR-bibliography.bibSR-reviewers.bibSR-0-Found.bibSR-2-Included.bibSR-3-Tracked.bibSR-5-Screened.bib}
+
 \end{document}

### 2025-03-03 scr/synthesis.r
revise synthesis per update to tasks

@@ -61,7 +61,6 @@ properties_inclusion %>%
     `Cases` = number_of_cases_incidents,
     `Features` = number_of_predictors_features
   ) %>%
-  mutate(Task = ifelse(Task == "Recommendation", "Decision Support", Task)) %>% 
   mutate(across(
     c(Cases, Features),
     \(x) sapply(x, FUN = format, big.mark = ",")
@@ -234,7 +233,7 @@ tab_synthesis %>%
   )) %>%
   mutate(Task = factor(
     Task,
-    c("Prediction", "Clustering", "Diagnosis", "Prognosis", "Decision Support")
+    c("Diagnosis", "Prognosis", "Forecast", "Recommendation")
   )) %>%
   mutate(Aim = factor(Aim, c("Knowledge", "Practice"))) %>%
   # mutate(across(c(Source, Task, Aim), fct_infreq)) %>%

### 2025-03-03 docs/Localized-Modeling-BMCAI/localized-modeling-bmcai.Rmd
track bmcai submission & dai revision

--- 2024-06-17 docs/localized-modeling-medRxiv/localized-modeling-medRxiv.Rmd
@@
 ---
-title: "Clinical prediction with localized modeling using similarity-based cohorts: A scoping review"
+# Supported options: 
+#   sn-nature:       Style for submissions to Nature Portfolio journals
+#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
+#   sn-mathphys:     Math and Physical Sciences Reference Style
+#   sn-aps:          American Physical Society (APS) Reference Style
+#   sn-vancouver:    Vancouver Reference Style
+#   sn-apa:          APA Reference Style 
+#   sn-chicago:      Chicago-based Humanities Reference Style
+#   default:         Default
+
+classoptions: 
+  - sn-mathphys      
+  - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
+  # - referee       # Optional: Use double line spacing
+  # - lineno        # Optional: Add line numbers
+  # - iicol         # Optional: Double column layour
+
+title: "Clinical prediction with localized modeling using similarity-based cohorts: A systematic scoping review"
+titlerunning: "Clinical prediction with localized modeling: A scoping review"
+
 authors: 
-  - name: Adi Cohen
-    email: ac4569@mynsu.nova.edu
-    affiliation: College of Osteopathic Medicine, Nova Southeastern University
-    location: Davie, FL
-  - name: Patti McCall-Junkin
-    email: pattimccall.junkin@gmail.com
-    affiliation: Smathers Libraries Academic Research & Consulting Services, University of Florida
-    location: Gainesville, FL
-  - name: Jason Cory Brunson
+  - firstname: Adi
+    lastname: Cohen
+    degrees: BS
+    affiliation: 1
+
+  - firstname: Patti
+    lastname: McCall-Junkin
+    degrees: MA, MLS
+    affiliation: 2
+
+  - firstname: Jason Cory
+    lastname: Brunson
+    degrees: PhD
     email: jason.brunson@medicine.ufl.edu
-    affiliation: Laboratory for Systems Medicine, University of Florida
-    location: Gainesville, FL
+    affiliation: 3
+    corresponding: TRUE
+
+affiliations:
+  - number: 1
+    info:
+      orgdiv: College of Osteopathic Medicine
+      orgname: Nova Southeastern University
+  - number: 2
+    info:
+      orgdiv: Smathers Libraries Academic Research & Consulting Services
+      orgname: University of Florida
+  - number: 3
+    corresponding: TRUE
+    info:
+      orgdiv: Laboratory for Systems Medicine
+      orgname: University of Florida
+    address:
+        street: P.O. Box 100225 JHMHC
+        city: Gainesville
+        state: Florida
+        country: USA
+
+keywords:
+  - case-based reasoning
+  - localized modeling
+  - nearest neighbors
+  - clinical decision support
+  - scoping review
+
+pacs: 
+  jel:
+    - "C38"
+    - "C65"
+  msc:
+    - "62P10"
+
+# Sample for unstructured abstract 
+# abstract: |
+#   The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.
+
+# Sample for structured abstract
 abstract: |
-  Background:
+  **Background:**
   Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of labeled cases matched by similarity to a target case.
@@
 
-  Methods:
+  **Methods:**
   We searched four bibliographic platforms during 2021 July 19–22, updated 2024 January 24.
@@
 
-  Results:
+  **Results:**
   Of 1,657 search results, 360 were reviewed, then combined with 43 publications that seeded the review and 1 obtained by citation tracking.
@@
 
-  Conclusions:
+  **Conclusions:**
   Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, but rigorous comparisons to competing methods must be conducted and computational hurdles must be overcome.
@@
 
-keywords:
-  - case-based reasoning
-  - localized modeling
-  - nearest neighbors
-  - clinical decision support
-  - scoping review
 bibliography:
@@
   - SR-5-Screened.bib
-biblio-style: unsrt
-output:
-  rticles::arxiv_article:
-    keep_tex: true
-    latex_engine: xelatex
-    # citation_package: natbib
-    citation_package: "default"
+
 header-includes: |
+  %% Per the spinger doc, new theorem styles can be included using built in style, 
+  %% but it seems the don't work so commented below
+  %\theoremstyle{thmstyleone}%
+  \newtheorem{theorem}{Theorem}%  meant for continuous numbers
+  %%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+  %% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+  \newtheorem{proposition}[theorem]{Proposition}%
+  %%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+  %% \theoremstyle{thmstyletwo}%
+  \theoremstyle{remark}
+  \newtheorem{example}{Example}%
+  \newtheorem{remark}{Remark}%
+
+  %% \theoremstyle{thmstylethree}%
+  \theoremstyle{definition}
+  \newtheorem{definition}{Definition}%
+
   \usepackage{lscape}
@@
   \newcommand{\landscapeend}{\end{landscape}}
+
+output: rticles::springer_article
 ---
@@
 
-# Introduction
+# Background
 
@@
 
-# Discussion
+# Conclusions
 
@@
 However, this also suggests that few experimental studies have focused on the unified development of new retrieval and adaptation strategies.
-We also note that the majority of studies evaluated and compared methods using leave-one-out cross-validation (LOOCV).
+We also note that the majority of studies evaluated and compared methods using LOOCV.
 This is an appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer.
@@
 
-## Conclusions
+## Closing
 
@@
 Whereas precious few of the reviewed studies used these models for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+
+# List of abbreviations {-}
+
+* CBR: case-based reasoning
+* CIS: clinical information system
+* CDS: clinical decision support
+* AI: artificial intelligence
+* NN: nearest neighbors
+* GLM: generalized linear model
+* ML: machine learning
+* DT: decision tree
+* PMC: PubMed Central
+* LOOCV: leave-one-out cross-validation
+* GSA: gravitational search algorithm
+* DML: distance metric learning
+
+# Declarations {-}
+
+## Ethics approval and consent to participate {-}
+
+Not applicable.
+
+## Consent for publication {-}
+
+Not applicable.
+
+## Availability of data and materials {-}
+
+Search results, included studies, and other cited sources are publicly available in a Zotero group library:
+<https://www.zotero.org/groups/5017571/imsr/>
+
+Data collected or encoded for included studies are publicly available in two Google Sheets:
+
+* Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
+* Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
+
+Code used to analyze these data and prepare this manuscript is publicly available in a GitHub repository:
+<https://github.com/corybrunson/imsr>
+
+## Competing interests {-}
+
+The authors declare that they have no competing interests.
+
+## Funding {-}
+
+Not applicable.
+
+## Authors' contributions {-}
+
+JCB conceived of the study.
+All authors contributed to the study design---search, screen, review, and synthesis.
+AC and JCB conducted the search, screen, review, and synthesis (coded and analyzed bibliographic and methodological data).
+AC and JCB drafted, and all authors contributed to, read, and approved, the manuscript.
+
+## Acknowledgements {-}
+
+Two anonymous reviewers provided valuable feedback on a previous version of this manuscript.
 
@@
 
-## Data and code
-
-Search results are publicly available in a Zotero group library.
-Data collected or encoded for included studies are publicly available in two Google Sheets.
-Code used to analyze these data and generate this manuscript is publicly available in a GitHub repository.
-
-* Search results, included studies, and other bibliography: <https://www.zotero.org/groups/5017571/imsr/>
-* Bibliographic and methodological properties of included studies: <https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/>
-* Terminology and composite techniques of included studies: <https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/>
-* Code used to conduct analyses and prepare the manuscript: <https://github.com/corybrunson/imsr>
-
 # References {-}

### 2025-03-03 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
track bmcai submission & dai revision

--- 2025-03-03 docs/Localized-Modeling-BMCAI/localized-modeling-bmcai.Rmd
@@
   Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of labeled cases matched by similarity to a target case.
+
+  \hl{**Objective:**}
   We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
@@
   **Methods:**
-  We searched four bibliographic platforms during 2021 July 19–22, updated 2024 January 24.
-We set four eligibility criteria to identify applications of localized models to clinical and health tasks.
+  We \hl{conducted searches of} four bibliographic platforms during 2021 July 19–22\hl{ and updated them on} 2024 January 24.
+We set four eligibility criteria to identify applications of localized models to clinical and health tasks\hl{, which amounted to using a multivariate similarity measure to retrieve cases from a retrospective corpus relevant to new cases and fitting statistical models to these cohorts}.
 Two authors divided title/abstract screening and reviewed screened entries for inclusion.
@@
 27 were included, published 1997–2022.
-The specificity of search terms was poor, and inter-rater reliability was low.
-Almost all models were predictive, the most common tasks being prognosis and diagnosis.
-Most studies used clinical, occasionally laboratory and image, data.
-Several addressed memory and runtime costs.
-A general technique that specializes to most of those reviewed involved matching, retrieval, fitting, and evaluation steps that could optionally be supervised, optimized, or recursively performed.
+The specificity of search terms was poor\hl{ (51/1,657)}, and inter-rater reliability was low\hl{ (82%)}.
+Almost all models were predictive, the most common tasks being prognosis\hl{ (11)} and diagnosis\hl{ (10)}.
+Most studies used clinical\hl{ (22)}, occasionally laboratory\hl{ (7)} or imaging\hl{ (2)}, data.
+\hl{The most competitive performances were reported for models that further hybridized the basic approach.} A general technique that specializes to most of those reviewed involved matching, retrieval, fitting, and evaluation steps that could optionally be supervised, optimized, or recursively performed.
 
@@
 
+  \usepackage{soul}
+
 output: rticles::springer_article
@@
 
-A main driver of the development of clinical information systems (CIS) and clinical decision support (CDS) tools has been to leverage the advantages of routinely collected health data toward improving care.
-These advantages include their immediate accessibility through the institutional EHR, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
-Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of evidence-based practice.
-
-Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools, including artificial intelligence (AI).
-Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other.
-The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models.
-
-One of few methodologies to address all three of these established needs is one we call _localized modeling_, which appears to have been introduced several times independently under different names, and in slightly different forms.
+A main driver of the development of clinical information systems (CIS) and \hl{computerized clinical decision support (CCDS)} tools has been to leverage the advantages of routinely collected health data toward improving care\hl{ [@]}.
+These advantages include their immediate accessibility through the institutional \hl{electronic health record (EHR)}, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
+Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of evidence-based practice\hl{ [@]}.
+
+Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools\hl{ [@]}, including artificial intelligence (AI).
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other\hl{ [@]}.
+The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models\hl{ [@]}.
+
+One of few methodologies to address all three of these established needs is one we call _localized modeling_, which \hl{we find in this review} to have been introduced several times independently under different names, and in slightly different forms.
 The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
-Traditional CBR returns these retrieved cases to inform human decision-making; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+Traditional CBR returns these retrieved cases to inform human decision-making\hl{ [@]}; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
 In contrast, localized modeling fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
+\hl{While many ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions [@Molnar], these quantifications are not directly interpretable model components analogous to the split nodes of a DT or the coefficients of a GLM.
+While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected.}
+
+\hl{This approach offers an alternative paradigm for interpretable machine learning that combines the simplicity and transparency of classical models like decision trees (DTs) and GLMs with the sensitivity to nonlinearity and interactions of advanced opaque models like random forests (RFs) and artificial neural networks (ANNs).
+The sensitivity is enabled by the fitting of the model to the local cohort, whose size and scope are governed by hyperparameters that may be tuned like any others via grid or Bayesian optimization.
+Some similarity measures or other retrieval methods may be more interpretable than others, in the same way that **TODO-matching** [@Rudin].
+The interpretability of the localized modeling approach would then depend on that of its retrieval (similarity matching) and revision (predictive modeling) steps.
+Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures exist for similarity matching [@kdtree] and the local models would benefit from increased numbers of relevant cases.}
 Localized models thus provide an individualized way to fit fully interpretable models to large population data.
 
-[^psm]: Similarity measures between more granular units of analysis, such as encounters and decision points, are often still referred to as patient similarity measures.
+[^psm]: Similarity measures between more granular units of analysis, such as encounters and decision points, are often still referred to as patient similarity measures\hl{ [@]}.
 
@@
 
-## Related work
+## Related \hl{survey }work
 
@@
 In contrast to rule-based systems, which are variable-based and generally interpretable as a single rule applied to all new cases, case-based systems provide case-specific interpretations in the form of a number of more fully understood reference cases.
-While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.
-
-@Kolodner1992 distinguished two styles of CBR: problem-solving, which is more procedural and used when objectives are more clearly defined, and interpretative, which provides categorizations and justifications for possible solutions.
-A similar dichotomy is commonly used to distinguish the performance criterion for the usefulness of predictive models from the interpretability criterion for their usability.
-In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
+
+@Kolodner1992 distinguished two styles of CBR: _problem-solving_, which is more procedural and used when objectives are more clearly defined, and _interpretative_, which provides categorizations and justifications for possible solutions.
+A similar dichotomy is commonly used to distinguish the _performance_ \hl{criterion }of predictive models from the _interpretability_ criterion\hl{.} In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
 Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
-Nevertheless, as some recent applications have shown, they can be used to draw inferences about the importance of different risk factors to specific individuals.
-While most ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions, these quantifications are not directly interpretable model components analogous to the split nodes of a decision tree (DT) or the coefficients of a GLM.
-While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected by the user.
-Provided the model family fitted to the cohorts is interpretable, the localized model inherits this property.
+Nevertheless, as \hl{we find in this review}, they can be used to draw inferences about the importance of different risk factors to specific individuals.
+
+\hl{We also note that a large variety of AI tools, besides NN, use similarity measures or other relevance-based retrieval methods.
+Implicitly, back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
+that small distances between cases yield small changes in prediction.
+More explicitly, kernel-based methods **TODO**.
+The key difference in localized modeling is that the retrieved cohort is treated as a component of the model and made available to the user.
+Indeed, retrieval procedures were the original algorithmic accomplishments of CBR [@] and the retrieved cohorts may be reused for numerous modeling tasks.}
 Thus, localized models may embody both styles of CBR.
@@
 We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
-While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools).
-For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.
+While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; \hl{CCDS} and software tools).
+For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.}
 
@@
 
-Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:
-
-* Uses case-level data from a corpus of past cases with known responses\newline
-(response may be outcome, diagnosis, subtype, etc.)
-* Defines a numeric multivariate case similarity measure\newline
-(allow integer-valued measures)
-* Uses the similarity measure to retrieve cohorts for index cases from the corpus\newline
-(for example, based on a training–testing partition)
-* Fits statistical models to cohorts from which to make predictions or draw inferences about index cases\newline
-(outcome predictions, survival estimates, risk factor contributions, model evaluation statistics, etc.)
-
 We expected the general framework to come down to three choices: A patient similarity measure, a cohort selection process, and a statistical model family.
@@
 Here we describe our review process, including deviations from plans and the reasons for them.
-More details are included in Section \ref{selection-process}.
+More details are included in \hl{the Appendix}.
 
@@
 We include a PRISMA checklist as Supplemental Table 1 and a PRISMA abstract checklist as Supplemental Table 2.
-Because we focus on methodologies rather than conceptual approaches or evidence, we deviate in some ways from PRISMA guidelines.
+Because we focus on methodologies rather than\hl{ evidence}, we deviate in some ways from PRISMA guidelines.
 In particular, those items of the checklist involving bias assessment and quantitative synthesis are intended for meta-analyses so did not apply to this study.
@@
 
+## \hl{Eligibility criteria}
+
+\hl{Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:}
+
+* \hl{Uses case-level data from a corpus of past cases with known responses}\newline
+\hl{(response may be outcome, diagnosis, subtype, etc.)}
+* \hl{Defines a numeric multivariate case similarity measure}\newline
+\hl{(allow integer-valued measures)}
+* \hl{Uses the similarity measure to retrieve cohorts for index cases from the corpus}\newline
+\hl{(for example, based on a training--testing partition)}
+* \hl{Fits statistical models to cohorts from which to make predictions or draw inferences about index cases}\newline
+\hl{(outcome predictions, survival estimates, risk factor contributions, model evaluation statistics, etc.)}
+
+\hl{Note that classical CBR satisfies the first and third criteria by definition and in most cases will satisfy the second.}
+
 ## Search
 
-We derived the eligibility criteria itemized in the Introduction from a seed set of previously read studies.
+We derived the eligibility \hl{criteria }from a seed set of previously read studies.
 Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
-We finalized the PubMed search first, then adapted it to the other platforms (Section \ref{search-design}).
-
-Through each platform, we searched those databases included by default.
-This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts & Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.
-
-The structure and terms of our search strings were inspired in part by previous reviews adjacent to our topic of interest [@Sharafoddini2017; @Parimbelli2018].
-We organized the PubMed search in disjunctive normal form (an OR of ANDs).
-Following the solidification of an outline, we expanded each term to include synonyms that are similar enough to be applicable to our search.
-We then evaluated the expanded search string using the PubMed Advanced Search platform.
-We initially included each term and their synonyms separately to evaluate what resulted.
-We pruned several terms that returned no results ("phrases not found") or to reduce the number of results.
-At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.
+We finalized the PubMed search first, then adapted it to the other platforms\hl{.}
 
@@
 
-Within Covidence, we screened titles and abstracts for relevance. We decided on four eligibility criteria to expedite the screening process.
+\hl{We} screened titles and abstracts for relevance. We decided on four eligibility criteria to expedite the screening process.
 Articles must be written in English, for readability; they must be original studies, to exclude secondary sources with duplicate information; their use settings must be medical, clinical, or related, to keep our review topical; and it must not be clear that their use of our search terms was different from our intended meaning.
 Each of two authors (AC and JCB) screened roughly half of the entries. They regularly reviewed each other’s decisions to improve consistency. In cases of uncertainty, entries were included.
-For the update, one author (PM) screened all results for three criteria and another (JCB) screened the survivors for the fourth criterion.
+For the update, one author (PM) screened all results for \hl{the first }three criteria and another (JCB) screened the survivors for the fourth criterion.
 
@@
 
-We set out 4 criteria for full-text review to restrict to studies that used some form of localized modeling on health data:
-
-* Pulls case-level data from a corpus of past cases with known classes or outcomes
-* Defines a numeric multivariate case similarity measure
-* Uses the similarity measure to select cohorts for index cases from the corpus
-* Fits statistical models to cohorts to make predictions or inferences about index cases
-
-Note that classical CBR satisfies the first and third criteria by definition and in most cases will satisfy the second.
-
-In Covidence, two authors (AC and JCB) independently evaluated each manuscript for these eligibility criteria.
+We \hl{used the} 4 criteria\hl{ from Section }\ref{eligibility-criteria} for full-text review\hl{.}
+
+\hl{Two} authors (AC and JCB) independently evaluated each manuscript for these eligibility criteria.
 The first criterion that a manuscript failed was designated the reason for exclusion.
@@
 
-During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival [@Mariuzzi1997] and survival curves [@Lowsky2013].
-These approaches were novel to CBR and presage later developments, so were helpful to understanding the development of localized modeling.
-However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the cohort's responses.
-A review of all studies that use nearest neighbors prediction would be impractical.
-Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to exclude those studies whose approach was equivalent to nearest neighbors prediction using a conventional similarity measure.
-
 Finally, one author (JCB) applied the same review process to the seed set of 43 studies that inspired the review.
@@
 #| fig.cap =
-#| "PRISMA-S flow chart.
+#| "\\hl{PRISMA-S flow chart.}
 #|  Lowercase letters refer to items obtained from
@@
 
-We then reviewed 43 studies comprising a seed set that inspired this review.
+We then reviewed \hl{a seed set comprising 43 studies} that inspired this review.
 After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using NN prediction.
@@
 Another interesting pattern lay in the years of publication (Figure \ref{fig:year}).
-While these studies trace back to the late 1990s, they have become more common, which suggests that this is an active, though not rapidly expanding, approach.
+While these studies trace back to the late 1990s, they have become more common\hl{ [@]}, which suggests that this is an active, though not rapidly expanding, approach.
 We note that CBR in health and medicine originated as early as 1990, and localized modeling emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
@@
   )) |> 
+  # allow text wrapping
+  mutate(Task = ifelse(Task == "Recommendation", "Recommend-ation", Task)) |> 
   # prevent numeric list columns from dominating the horizontal space
@@
       "\\label{tab:synthesis}",
-      "Studies included in the method synthesis,",
-      " arranged by the earliest the study is known to have been public."
+      "\\hl{Studies included in the method synthesis,",
+      " arranged by the earliest the study is known to have been public.}"
     )
@@
 We categorized studies as "knowledge" or "practice" based on whether their aim was to produce generalizable knowledge or to improve practice.
-We classified 8 of the studies as "knowledge", making the majority of studies from a "practice" standpoint.
+We classified `r tab_synthesis |> filter(Aim == "Knowledge") |> nrow()` of the studies as "knowledge", making the majority of studies from a "practice" standpoint.
 These studies aimed to provide tools for use in clinic or to improve outcomes.
+
+\hl{Most studies used clinical data elements alone
+(`r tab_synthesis |> filter(Source == "Clinical") |> nrow()`),
+and most of the remainder used these in combination with laboratory
+(`r tab_synthesis |> filter(str_detect(Source, "Clinical.*Laboratory")) |> nrow()`),
+imaging
+(`r tab_synthesis |> filter(str_detect(Source, "Clinical.*Imaging")) |> nrow()`),
+or environmental
+(`r tab_synthesis |> filter(str_detect(Source, "Clinical.*Environmental")) |> nrow()`)
+data. The data source was related to the modeling task, for instance in that all decision support models used either clinical or patient-reported data.
+The mosaic plot in Figure }\ref{fig:properties}\hl{ summarizes the relation between data source, clinical task, and aim.}
 
@@
 A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021].
-In both cases a prototype implementation was deployed in an experimental setting for evaluation.
+In both cases a prototype implementation was \hl{evaluated} in an experimental \hl{setting.}
 
@@
 @Lowsky2013 used CBR with non-parametric survival models on registry data to predict patient–graft survival times following kidney transplantation.
-@Lee2015 and @Lee2017 used localized logistic regression and random forest modeling to predict 30-day mortality following discharge for ICU patients.
-@Vilhena2016 developed a CBR cycle around a clustering-informed similarity matching procedure and an artificial neural network–based classifier to identify thrombophilia patients at high risk of thrombotic episodes.
+@Lee2015 and @Lee2017 used localized \hl{GLM} and RF modeling to predict 30-day mortality following discharge for ICU patients.
+@Vilhena2016 developed a CBR cycle around a clustering-informed similarity matching procedure and an \hl{ANN}–based classifier to identify thrombophilia patients at high risk of thrombotic episodes.
 @Ma2020 took a similarity cohort–based approach to predicting length of stay for ICU patients.
-@Doborjeh2022 localized spiking neural networks in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.
+@Doborjeh2022 localized spiking \hl{ANN}s in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.
 @Liu2022 incorporated global-to-local transfer learning into localized models of acute kidney injury risk in hospitalized patients.
@@
 
-The mosaic plot in Figure \ref{fig:properties} summarizes the relation between data source, clinical task, and aim.
-
 ```{r properties, fig.align='center', out.width="100%", echo=FALSE}
 #| fig.cap =
-#| "Share of studies characterized by several design elements:
+#| "\\hl{Share of studies characterized by several design elements:}
 #|  data source (row),
@@
 
-## Rationales
+## Rationales\hl{ for methods}
 
@@
 @Elter2007 and @Nicolas2014 emphasized the value of the detailed past cases, available to the user, on which predictions are based.
-This "self-explanation capability" made outputs more intelligible to physicians in the CDS setting.
+This "self-explanation capability" made outputs more intelligible to physicians in the \hl{CCDS} setting.
 @Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
@@
 
-## Challenges
+## Challenges\hl{ with methods}
 
@@
 Though @CampilloGimenez2013 and @Ng2015 report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
-Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model---statistical CBR [@Park2006]; individualized logistic regression, decision tree, and random forest [@Lee2015; @Lee2017]---they did not consistently outperform comparators.
+Meanwhile, when proposed approaches targeted cohort demarcation or choice of predictive model---statistical CBR [@Park2006]; individualized \hl{GLM, DT, and RF} [@Lee2015; @Lee2017]---they did not consistently outperform comparators.
 
@@
 
-# Conclusions
+# \hl{Discussion}
 
@@
 #|  Dashed arrows indicate optional steps."
-knitr::include_graphics("Fig5.png")
-```
-
-Each step can be unsupervised supervised: Unsupervised similarity measures include discrete measures like the Levenshtein distance and continuous measures like cosine similarity, while supervised measures include the Mahalanobis distance and random forest proximity as well as several composite measures with weights calculated from the data.
+knitr::include_graphics("Fig5.pdf")
+```
+
+Each step can be unsupervised supervised: Unsupervised similarity measures include discrete measures like the Levenshtein distance and continuous measures like cosine similarity, while supervised measures include the Mahalanobis distance and \hl{RF} proximity as well as several composite measures with weights calculated from the data.
 Most retrieval steps obtained cohorts with a uniform size (cardinality) or similarity threshold, and the remaining were likewise only informed by predictors; thus, while a retrieval step could in principle be supervised, in this sample none were.
@@
 For this purpose, we propose the framework outlined above, at least as a starting point.
-When a specification invokes complex techniques, for example supervised DML or neural network classifiers, it should be compared at least to alternative specifications using simpler alternatives.
+When a specification invokes complex techniques, for example supervised DML or \hl{ANN} classifiers, it should be compared at least to alternative specifications using simpler alternatives.
 @CampilloGimenez2013 and @Liu2022 provide excellent examples in which all possible choices are toggled and yield a neatly nested comparison.
@@
 
-## Closing
+## \hl{Conclusion}
 
@@
 * CIS: clinical information system
-* CDS: clinical decision support
+* \hl{CCDS: computerized} clinical decision support
 * AI: artificial intelligence
@@
 * GLM: generalized linear model
+* \hl{DT: decision tree}
+* \hl{RF: random forest}
+* \hl{ANN: artificial neural network}
 * ML: machine learning
-* DT: decision tree
 * PMC: PubMed Central
@@
 (A result obtained from multiple databases should have only one Zotero entry but should be filed under the subcollection for each database in which it was found.)
+
+### \hl{Databases}
+
+\hl{Through each platform, we searched those databases included by default.
+This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts \& Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.}
+
+\hl{The structure and terms of our search strings were inspired in part by previous reviews adjacent to our topic of interest }[@Sharafoddini2017; @Parimbelli2018]\hl{.
+We organized the PubMed search in disjunctive normal form (an OR of ANDs).
+Following the solidification of an outline, we expanded each term to include synonyms that are similar enough to be applicable to our search.
+We then evaluated the expanded search string using the PubMed Advanced Search platform.
+We initially included each term and their synonyms separately to evaluate what resulted.
+We pruned several terms that returned no results ("phrases not found") or to reduce the number of results.
+At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.}
 
@@
 * Fits statistical models to cohorts to make inferences about index cases
+
+\hl{During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival }[@Mariuzzi1997]\hl{ and survival curves }[@Lowsky2013]\hl{.
+These approaches were novel to CBR and presaged later developments, so they were helpful to understanding the development of localized modeling.
+However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the cohort's responses.
+A review of all studies that use nearest neighbors prediction would be impractical.
+Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to exclude those studies whose approach was equivalent to nearest neighbors prediction using a conventional similarity measure.}
 

### 2025-03-05 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
revise discussion section

@@
 
-  \hl{**Objective:**}
+  **\hl{Objective:}**
   We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
@@
 
-## \hl{Eligibility criteria}
+## Eligibility criteria
 
@@
 
-### Performance
-
-As discussed in Section \ref{rationales}, most studies were premised on the potential predictive value of localized models.
-As cautioned in Section \ref{evaluations}, not all experiments affirmed this premise.
-Most of those that did involved using metric learning to improve retrieval and sometimes iterative optimization of the metric and of the localized model.
-Some studies showcased results using a variety of such specializations [@CampilloGimenez2013; @Ng2015; @Zhang2018; @Liu2022].
-However, because these advanced implementations were mostly compared against their precursors or commonplace alternatives, we cannot speak to their performance or any trade-offs with respect to each other.
-This would be a worthy goal of future work.
-
-The value of this work would be enhanced by standardization and modularization.
-First, studies using localized models should locate the proposed approach and any variations in a shared specification space.
-For this purpose, we propose the framework outlined above, at least as a starting point.
-When a specification invokes complex techniques, for example supervised DML or \hl{ANN} classifiers, it should be compared at least to alternative specifications using simpler alternatives.
-@CampilloGimenez2013 and @Liu2022 provide excellent examples in which all possible choices are toggled and yield a neatly nested comparison.
-Additionally, some state-of-the-art models, including others that obtain individualized predictions, should be included in comparisons.
-Finally, implementations should be not only published on public repositories but designed in such a way that users familiar with the underlying language can substitute a technique of their choosing for any part of the specification (similarity, retrieval, fitting, evaluation).
-We believe these principles will make reported results clearer and reproduction and extension easier.
-
-### Interpretability
-
-While most studies identified outstanding technical needs (Section \ref{identified-needs}), few emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
+\hl{We suggest four avenues for future work: evaluation of predictive performance, usability and interpretability, practical and regulatory feasibility, and customizability.}
+
+\hl{Most studies were premised on the potential predictive value of localized models (Section }\ref{rationales}\hl{), but not all experiments affirmed this premise (Section }\ref{evaluations}\hl{).
+Those that did mostly used DML to improve retrieval and sometimes iteratively optimized the metric and the model.
+However, even those that showcased several such specializations }[@CampilloGimenez2013; @Ng2015; @Zhang2018; @Liu2022]\hl{ mostly used their own precursors or commonplace alternatives as comparators.
+Future work should compare competitive alternatives to assess their performance and other trade-offs.
+This work would be facilitated by standardization and modularization using a framework like that outlined above.
+For example, }@CampilloGimenez2013\hl{ and }@Liu2022\hl{ use a modular design to toggle multiple choices and obtain neatly nested comparisons.}
+
+\hl{While most studies identified outstanding technical needs (Section }\ref{identified-needs}\hl{), few (though some) emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
 Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
-Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate, but this is more often assumed than demonstrated.
-
+Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate.
 One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
-This is a necessary precondition for practical use but would also be a valuable contribution to the experimental literature.
-Another would be to compare the localized predictor importance measures obtained from localized models to the local importance measures used to explain predictions made by opaque models.
-This would test both measures for concurrent validity, while differences between them would inform what settings or cases are poorly served by one or the other.
-
-### Feasibility
-
-Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
-This need demands not only methodologies but also architectures that are robust, versatile, and compliant in the face of diverse data models and use restrictions, yet no studies in our sample addressed the challenges of system interoperability and regulatory requirements head-on.
-
-CBR and, by extension, localized modeling are especially susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures as well.
-Individual systems exhibit many dimensions of incompleteness [@Weiskopf2013] and vary along other dimensions of quality [@Kohane2021], and their aggregation for modeling purposes depends on several kinds of interoperability [@Weber2015].
-Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties [@Haendel2021].
-While most studies in our sample focused on improving care, increasingly many over time focused on generalizable knowledge.
-Therefore, as studies using localized modeling shift from proof of concept to feasibility, they will also need to demonstrate practical and regulatory feasibility.
-
-### Customizability
-
-The primary goal of these studies was to establish that localized models, and certain strategies within this paradigm, perform at or above the level of other predictive models or strategies.
-From a functionalist approach to reproducibility [@Matarese2022], then, later studies were broadly successful at reproducing earlier studies, and later studies provide sufficient detail to support ongoing reproduction efforts---though these might be limited by the lack of open-source code or public implementations.
-However, the immediate goal of most studes was to improve care, and an implied need, often enjoined but neither obtained nor reproduced, was a demonstration of practical use.
-
-We posit that an essential component of practical usefulness is the ability of the user community to exert some control over the models.
-Given the impact on performance of the choice or optimization of the similarity measure, an important target for user input would be the importance of certain variables in the calculation of similarity, with an understanding of how it can impact not only the performance of the predictions but also the cohort retrieved for the model.
+Another would be to compare the localized predictor importance measures obtained from localized models to the local importance measures used to explain predictions made by opaque models.}
+
+\hl{Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
+Such software architectures must be robust, versatile, and compliant in the face of diverse data models and use restrictions.
+Localized modeling is susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures.
+Individual systems exhibit many dimensions of incompleteness }[@Weiskopf2013]\hl{ and vary along other dimensions of quality }[@Kohane2021]\hl{, and their aggregation for modeling purposes depends on several kinds of interoperability }[@Weber2015]\hl{.
+Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties }[@Haendel2021]\hl{.
+Future work may therefore eventually need to directly address the challenges of system interoperability and regulatory requirements.}
+
+\hl{From a functionalist approach to reproducibility }[@Matarese2022]\hl{, later studies were broadly successful at reproducing earlier studies, and later studies provide sufficient detail to support ongoing reproduction efforts---though these might be limited by the lack of open-source code or public implementations.
+However, the immediate goal of most studes was to improve care.
+We expect that this practical utility will depend on the ability of the user community to exert control over certain inputs, components, and behaviors of the models.
 For one example, a user may want to minimize the weight of rare diseases in medical history in order to retrieve a population with more such cases in order to better measure their associated risk to the outcome.
 In contrast, they may want to increase the weight of the indicating diagnosis in order to allow fewer patients from similar but distinct populations to influence the model.
-For another example, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
-@Lopez2011 took an important step in this direction with an adjustable and adaptable implementation.
-Future purely quantitative work could assess whether similarity-tuning can achieve these ends more flexibly than strict inclusion/exclusion criteria.
+Alternatively, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
+}@Lopez2011\hl{ took an important step in this direction with an adjustable and adaptable implementation.}
 
@@
 
-We propose the term _localized modeling_ to encompass an approach derived from CBR in which parameterized models are fitted in a standardized way to nearest neighborhoods of past or training cases according to a measure of patient similarity.
-We conducted a systematic search for studies that apply localized models to tasks involving health data and synthesized these largely independently developed approaches into a general framework.
-While the search was limited by low inter-rater reliability and failure to recover several motivating examples, the included studies used many of the same underlying tools to build, optimize, and evaluate their methods.
-We therefore believe that our framework can serve to taxonomize ongoing work of this type and inform the development of customizable implementations.
-Indeed, the availability of increasing computational power, the diversity of tasks to which these models were applied and of technical specifications they employed, and the apparent lack of any multi-group research program to date suggest great potential for growth.
-Whereas precious few of the reviewed studies used these models for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work put greater emphasis on the development and validation of interpretable localized models and on their reception by communities of medical research and clinical practice.
+We conducted a systematic search for studies that \hl{systematically fit parameterized models to similarity-based cohorts pursuant} to tasks involving \hl{medical and }health data\hl{.
+}While the search \hl{strategy showed serious limitations}, the included studies used many of the same underlying tools to build, optimize, and evaluate their methods.
+\hl{We synthesized these largely independently developed approaches into a general framework that we believe} can serve to taxonomize ongoing \hl{work }and inform \hl{future development}.
+Indeed, the availability of increasing computational power\hl and} the diversity of tasks to which these models were applied and of technical specifications they \hl{employed suggest }potential for growth.
+\hl{Whereas} few of the reviewed studies used these models for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work \hl{emphasize the} validation of interpretable localized models \hl{and their accessibility to} communities of medical research and clinical practice.
 

### 2025-03-06 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
add some references and change some language

@@
 
-title: "Clinical prediction with localized modeling using similarity-based cohorts: A systematic scoping review"
-titlerunning: "Clinical prediction with localized modeling: A scoping review"
+title: "Clinical prediction using similarity cohort--based models: A systematic scoping review"
+titlerunning: "Clinical prediction using similarity cohort--based models"
 
@@
 
-A main driver of the development of clinical information systems (CIS) and \hl{computerized clinical decision support (CCDS)} tools has been to leverage the advantages of routinely collected health data toward improving care\hl{ [@]}.
+A main driver of the development of clinical information systems (CIS) and \hl{computerized clinical decision support (CCDS)} tools has been to leverage the advantages of routinely collected health data toward improving care\hl{ }[@].
 These advantages include their immediate accessibility through the institutional \hl{electronic health record (EHR)}, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
-Though by definition not collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of evidence-based practice\hl{ [@]}.
-
-Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools\hl{ [@]}, including artificial intelligence (AI).
-Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other\hl{ [@]}.
-The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models\hl{ [@]}.
+\hl{Though not} collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of \hl{evidence-based practice }[@Holmqvist2015].
+
+Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools\hl{ }[@], including artificial intelligence (AI).
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other\hl{ }[@].
+The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models\hl{ }[@].
 
 One of few methodologies to address all three of these established needs is one we call _localized modeling_, which \hl{we find in this review} to have been introduced several times independently under different names, and in slightly different forms.
-The approach is a specialized form of case-based reasoning (CBR) that relies on a patient similarity measure[^psm] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.
-Traditional CBR returns these retrieved cases to inform human decision-making\hl{ [@]}; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+The approach is a specialized form of case-based reasoning (CBR) that relies on a \hl{patient similarity measure }[@Brown2016] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.[^psm]
+\hl{Some of the earliest and simplest CBR tools returned} these retrieved cases \hl{ to inform human decision-making}[@Aamodt1994]; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
 In contrast, localized modeling fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
-\hl{While many ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions [@Molnar], these quantifications are not directly interpretable model components analogous to the split nodes of a DT or the coefficients of a GLM.
+\hl{While many ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions }[@Molnar2023]\hl{, these quantifications are not directly interpretable model components analogous to the split nodes of a DT or the coefficients of a GLM.
 While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected.}
@@
 
-[^psm]: Similarity measures between more granular units of analysis, such as encounters and decision points, are often still referred to as patient similarity measures\hl{ [@]}.
+[^psm]: Similarity measures between more granular units of analysis, such as encounters and decision points\hl{, or between patients' time series}, are often still referred to as patient similarity measures\hl{ in our sample}.
 
@@
 The key difference in localized modeling is that the retrieved cohort is treated as a component of the model and made available to the user.
-Indeed, retrieval procedures were the original algorithmic accomplishments of CBR [@] and the retrieved cohorts may be reused for numerous modeling tasks.}
+Indeed, retrieval procedures were the most significant algorithmic requirements and accomplishments of CBR }[@Kolodner1992; @Gierl1998]\hl{ and the retrieved cohorts may be reused for numerous modeling tasks.}
 Thus, localized models may embody both styles of CBR.
@@
 Another interesting pattern lay in the years of publication (Figure \ref{fig:year}).
-While these studies trace back to the late 1990s, they have become more common\hl{ [@]}, which suggests that this is an active, though not rapidly expanding, approach.
+While these studies trace back to the late 1990s, they have become \hl{slightly }more common, which suggests that this is an active, though not rapidly expanding, approach.
 We note that CBR in health and medicine originated as early as 1990, and localized modeling emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
@@
 However, more informatical and legal challenges to implementation, including interoperability of systems and data and regulations concerning security and privacy, were rarely addressed.
-We discuss this further in Section \ref{feasibility}.
+We discuss this further in Section \ref{future-work}.
 
@@
 
-## Directions and expectations for future work
+## Future work
 

### 2025-03-07 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
finalize revisions

@@
   - case-based reasoning
-  - localized modeling
+  - similarity cohort
   - nearest neighbors
@@
   **Background:**
-  Applications of classical case-based reasoning (CBR) have given rise to a family of techniques we call “localized models”, in which a statistical model is fitted to a neighborhood of labeled cases matched by similarity to a target case.
+  Applications of classical case-based reasoning (CBR) have given rise to a family of modeling techniques \hl{reliant on similarity-based cohorts}, in which a statistical model is fitted to a neighborhood of labeled cases matched by similarity to a target case.
 
   **\hl{Objective:}**
-  We aim to describe clinical and health applications of localized models to date and propose a general framework for their design and evaluation.
+  We aim to describe clinical and health applications of \hl{similarity cohort--based models (SCBMs)} to date and propose a general framework for their design and evaluation.
 
@@
   We \hl{conducted searches of} four bibliographic platforms during 2021 July 19–22\hl{ and updated them on} 2024 January 24.
-We set four eligibility criteria to identify applications of localized models to clinical and health tasks\hl{, which amounted to using a multivariate similarity measure to retrieve cases from a retrospective corpus relevant to new cases and fitting statistical models to these cohorts}.
+We set four eligibility criteria to identify applications of \hl{SCBMs} to clinical and health tasks\hl{, which amounted to using a multivariate similarity measure to retrieve cases from a retrospective corpus relevant to new cases and fitting statistical models to these cohorts}.
 Two authors divided title/abstract screening and reviewed screened entries for inclusion.
@@
 27 were included, published 1997–2022.
-The specificity of search terms was poor\hl{ (51/1,657)}, and inter-rater reliability was low\hl{ (82%)}.
+The specificity of search terms was poor\hl{ (51/1,657)}, and inter-rater reliability was low\hl{ (82\%)}.
 Almost all models were predictive, the most common tasks being prognosis\hl{ (11)} and diagnosis\hl{ (10)}.
@@
   **Conclusions:**
-  Localized models have potential to improve the performance of clinical decision support tools while maintaining interpretability, but rigorous comparisons to competing methods must be conducted and computational hurdles must be overcome.
+  \hl{SCBMs} have potential to improve the performance of clinical decision support tools while maintaining interpretability, but rigorous comparisons to competing methods must be conducted and computational hurdles must be overcome.
 We hope that our review will spur future work on efficiency, reproducibility, and user needs.
@@
 
-A main driver of the development of clinical information systems (CIS) and \hl{computerized clinical decision support (CCDS)} tools has been to leverage the advantages of routinely collected health data toward improving care\hl{ }[@].
+A main driver of the development of clinical information systems (CIS) and \hl{clinical decision support systems (CDSS)} has been to leverage the advantages of routinely collected health data toward \hl{improving care }[@Benchimol2015; @Wasylewicz2019].
 These advantages include their immediate accessibility through the institutional \hl{electronic health record (EHR)}, billing database, or other source; their specificity to the institution and the population it serves; and the closeness in time of the available data.
-\hl{Though not} collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of \hl{evidence-based practice }[@Holmqvist2015].
-
-Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of advanced computational tools\hl{ }[@], including artificial intelligence (AI).
-Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but their associations and interplay with each other\hl{ }[@].
-The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of more interpretable complex models\hl{ }[@].
-
-One of few methodologies to address all three of these established needs is one we call _localized modeling_, which \hl{we find in this review} to have been introduced several times independently under different names, and in slightly different forms.
+\hl{Though not} collected for research use, the secondary research use of routinely collected health data has been put forth, as "practice-based evidence", a complement to the paradigm of \hl{evidence-based practice }[@Sim2001; @Holmqvist2015].
+
+Over the same period of time, interest in individualizing care from population-derived evidence-based recommendations to specific patient needs has driven the application of \hl{advanced computational tools }[@Rosella2022; @Elhaddad2024], including artificial intelligence (AI).
+Whereas classical models produced formulae involving a limited set of data elements that would be uniformly applied to all patients, AI models often process hundreds or more variables in opaque ways to yield predictions that depend on not only their values but \hl{their associations and interplay with each other }[@Molnar2023].
+The loss of straightforward interpretations of these models has limited their practical uptake and motivated the construction of explanatory statistics for opaque models as well as the development of \hl{more interpretable complex models }[@Rudin2022; @Molnar2023].
+
+One of few methodologies to address all three of these established needs is \hl{similarity cohort--based modeling (SCBM)}, which \hl{we find in this review} to have been introduced several times independently under different names, and in slightly different forms.
 The approach is a specialized form of case-based reasoning (CBR) that relies on a \hl{patient similarity measure }[@Brown2016] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.[^psm]
 \hl{Some of the earliest and simplest CBR tools returned} these retrieved cases \hl{ to inform human decision-making}[@Aamodt1994]; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
-In contrast, localized modeling fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
+In contrast, \hl{SCBM} fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
 \hl{While many ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions }[@Molnar2023]\hl{, these quantifications are not directly interpretable model components analogous to the split nodes of a DT or the coefficients of a GLM.
@@
 The sensitivity is enabled by the fitting of the model to the local cohort, whose size and scope are governed by hyperparameters that may be tuned like any others via grid or Bayesian optimization.
-Some similarity measures or other retrieval methods may be more interpretable than others, in the same way that **TODO-matching** [@Rudin].
-The interpretability of the localized modeling approach would then depend on that of its retrieval (similarity matching) and revision (predictive modeling) steps.
-Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures exist for similarity matching [@kdtree] and the local models would benefit from increased numbers of relevant cases.}
-Localized models thus provide an individualized way to fit fully interpretable models to large population data.
+Some similarity measures or other retrieval methods may be more interpretable than others, so that the interpretability of the SCBM approach would then depend on that of its retrieval (similarity matching) and revision (predictive modeling) steps.
+Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures and algorithms exist for similarity matching }[@Halder2024]\hl{ and the local models would benefit from increased numbers of relevant cases.}
+\hl{SCBM}s thus provide an individualized way to fit fully interpretable models to large population data.
 
@@
 @Kolodner1992 distinguished two styles of CBR: _problem-solving_, which is more procedural and used when objectives are more clearly defined, and _interpretative_, which provides categorizations and justifications for possible solutions.
-A similar dichotomy is commonly used to distinguish the _performance_ \hl{criterion }of predictive models from the _interpretability_ criterion\hl{.} In these terms, localized models are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
-Accordingly, localized models have been primarily used for and evaluated on their ability to predict classes or outcomes.
+A similar dichotomy is commonly used to distinguish the _performance_ \hl{criterion }of predictive models from the _interpretability_ criterion\hl{.} In these terms, \hl{SCBM}s are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
+Accordingly, \hl{SCBM}s have been primarily used for and evaluated on their ability to predict classes or outcomes.
 Nevertheless, as \hl{we find in this review}, they can be used to draw inferences about the importance of different risk factors to specific individuals.
 
-\hl{We also note that a large variety of AI tools, besides NN, use similarity measures or other relevance-based retrieval methods.
-Implicitly, back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
-that small distances between cases yield small changes in prediction.
-More explicitly, kernel-based methods **TODO**.
-The key difference in localized modeling is that the retrieved cohort is treated as a component of the model and made available to the user.
+\hl{We also note that a large variety of AI tools, besides NN, use similarity measures or other relevance-based retrieval methods: Back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
+that small distances between cases yield small changes in prediction; and kernel-based methods that rely directly on pairwise distances computed via inner product, rather than high-dimensional coordinates, extend far beyond their use in CBR.
+The key difference in SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
 Indeed, retrieval procedures were the most significant algorithmic requirements and accomplishments of CBR }[@Kolodner1992; @Gierl1998]\hl{ and the retrieved cohorts may be reused for numerous modeling tasks.}
-Thus, localized models may embody both styles of CBR.
+Thus, \hl{SCBM}s may embody both styles of CBR.
 
 We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
-While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; \hl{CCDS} and software tools).
+While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; \hl{CDSS} and software tools).
 For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.}
@@
 
-Our goals in this review are (1) to describe applications of similarity-based localized models using patient-level health data and (2) to provide a general framework for the design and evaluation of such localized models.
-We focus narrowly on localized models, rather than broadly on patient similarity--based clinical decision support tools, so that we may thoroughly assess their value in terms of reported evaluations and comparisons to other methods.
+Our goals in this review are (1) to describe applications of similarity cohort--based models using patient-level health data and (2) to provide a general framework for the design and evaluation of such \hl{SCBM}s.
+We focus narrowly on \hl{SCBM}s, rather than broadly on patient similarity--based clinical decision support tools, so that we may thoroughly assess their value in terms of reported evaluations and comparisons to other methods.
 This also allows us to devise a high-level framework that specializes to the majority of these methods, which we then use to frame our discussion and recommendations.
@@
 Rather than an evidence synthesis characteristic of most systematic reviews, we here pursue a methodology synthesis to harmonize largely independent research efforts that have converged on a common family of techniques.
-The goal will be to describe a unified framework for localized models that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.
+The goal will be to describe a unified framework for \hl{SCBM}s that can be used to guide future study designs and implementations as well as more systematically evaluate variations on the theme and measure the dependence of results on modeling choices.
 
@@
 While these studies trace back to the late 1990s, they have become \hl{slightly }more common, which suggests that this is an active, though not rapidly expanding, approach.
-We note that CBR in health and medicine originated as early as 1990, and localized modeling emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
+We note that CBR in health and medicine originated as early as 1990, and \hl{SCBM} emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
 
@@
 @Doborjeh2022 localized spiking \hl{ANN}s in order to assess stroke risk from up to 7-day time series of clinical and environmental factors.
-@Liu2022 incorporated global-to-local transfer learning into localized models of acute kidney injury risk in hospitalized patients.
+@Liu2022 incorporated global-to-local transfer learning into \hl{SCBM}s of acute kidney injury risk in hospitalized patients.
 Also of note, from a public health perspective, @Xu2008 used CBR to predict rehabilitation time as well as disability risk for unemployed workers experiencing chronic pain.
@@
 @Nicolas2014 used collaborative multilabel CBR to subtype melanoma patients based on confocal and dermoscopy images.
-@Wang2019 used localized models built from a multi-type additive similarity measure to distinguish type 2 diabetic versus control populations.
+@Wang2019 used \hl{SCBM}s built from a multi-type additive similarity measure to distinguish type 2 diabetic versus control populations.
 Along the way, @Song2006, @Kasabov2010, and @Verma2015 took an iterative model-building approach to several tasks: predicting glomerular filtration rate, a key indicator of renal function, from demographic and physiological variables; identifying patients with colon cancer using a large number of gene expression measurements; and identifying patients with type 2 diabetes based on demographic, physiological, and genetic variables.
@@
 
-The included studies hypothesized, asserted, or assumed several benefits of localized modeling specific to clinical and medical settings and advantages over other modeling approaches.
+The included studies hypothesized, asserted, or assumed several benefits of \hl{SCBM} specific to clinical and medical settings and advantages over other modeling approaches.
 The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015; @Ng2015; @Lee2017].
@@
 Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2013; @Nicolas2014; @Verma2015; @Lee2017].
-@Verma2015 and @Zhang2018 additionally pointed out that localized models are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
+@Verma2015 and @Zhang2018 additionally pointed out that \hl{SCBM}s are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
 These properties, they said, tend to be more difficult for whole-population models to handle.
 
-The other frequent advantage attributed to localized modeling was interpretability.
+The other frequent advantage attributed to \hl{SCBM} was interpretability.
 @Elter2007 and @Nicolas2014 emphasized the value of the detailed past cases, available to the user, on which predictions are based.
-This "self-explanation capability" made outputs more intelligible to physicians in the \hl{CCDS} setting.
+This "self-explanation capability" made outputs more intelligible to physicians in the \hl{CDSS} setting.
 @Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
@@
 @Ma2020 proposed to improve efficiency along the entire R4 cycle, but in particular by only executing task-dependent steps in real time ("just-in-time learning", JITL), and @Ng2021 and @Tang2021 partitioned multiple phases in the R4 cycle into offline and online components, only the latter of which would be performed in real time as new data are received.
-@Liu2022 used transfer learning from globally-fitted models to improve the efficiency of localized models.
+@Liu2022 used transfer learning from globally-fitted models to improve the efficiency of \hl{SCBM}s.
 
@@
 
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators---Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]---are hybrids of localized modeling (in some cases CBR) with other techniques, often DML.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators---Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]---are hybrids of \hl{SCBM} (in some cases CBR) with other techniques, often DML.
 Though @CampilloGimenez2013 and @Ng2015 report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
@@
 Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997; @Wang2019].
-Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004], to make models more expressive [@Lee2015], and to combine information obtained both from global and from localized models [@Ng2015].
+Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004], to make models more expressive [@Lee2015], and to combine information obtained both from global and from \hl{SCBM}s [@Ng2015].
 Most of this incremental work was indeed carried out in later investigations in our sample.
@@
 
-Because we excluded studies that used conventional NN prediction, many included studies reported new approaches to adaptation subsequent to retrieval.
-Very few proposed novel similarity measures, possibly because larger studies tended to be reported in separate articles detailing experiments with specific components of the process.
-However, this also suggests that few experimental studies have focused on the unified development of new retrieval and adaptation strategies.
-We also note that the majority of studies evaluated and compared methods using LOOCV.
-This is an appropriate technique when available data are scarce, and indeed most data sets used by included studies numbered in the hundreds of cases or fewer.
-This likely follows from the older age of many included studies and from their consistent primary focus on clinical data, which is more costly to collect and comes with more restrictions on its use.
-
 ## Coherence
@@
 
-These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as localized modeling; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
+These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as \hl{similarity cohort--based modeling}; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
 We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
-Because _local_ naturally contrasts with _global_, we propose "localized models" as a suitable term for this counterpart to globally-fitted models.[^local]
-
-[^local]: We note, in response to one reviewer's comment, that for many if not most models used in ML, excepting GLMs, a predictor may play a different and more or less important role for some cases than others, as quantified by importance measures or model-agnostic explanations [@Biecek2021; @Molnar2023].
-In this way, such models can also be viewed as "localized".
-We recognize that the use of global and local explanations is standard.
-As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
-The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
-At present, we feel that these methods are sufficiently disjoint to avoid confusion in practice, but we suggest the more specific term "similarity cohort--based models" when the term "local" is overloaded.
+\hl{While "local" naturally contrasts with "global", the overloaded term "localized models" stands to cause more confusion than it resolves, so we use "similarity cohort--based models" to refer to this specific approach.}
 
@@
 
-The similarity learning approaches took a variety of forms.
-These techniques are designed to detect structure in data, especially associations between predictors and responses, and to use this structure to inform the definition of a patient similarity measure.
-Eight studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
+\hl{Eight} studies used response values in training data sets to supervise similarity learning while two used unsupervised learning (Table \ref{tab:composite}).
 Based on the key properties of DML algorithms identified by @Bellet2014, most learned measures were non-linear while some were locally learned, and some were optimized globally while others locally.
@@
 Each used its similarity measure to retrieve training cases relevant to each testing case.
-Once retrieved, most took a standard NN approach to generating predictions; two exceptions [@Vilhena2016; @Tang2021] fit predictive models to the retrieved cohorts.
+\hl{Most} took a standard NN approach to generating predictions; \hl{two} [@Vilhena2016; @Tang2021] fit predictive models to the retrieved cohorts.
 
@@
 
-As noted above, these instances of similarity learning fit into a much larger literature on DML, which has seen widespread use in health informatics and modeling.
-That these studies satisfied our review criteria is due to the novelty of the techniques at the time of publication and the detailed attention paid to their techniques by the authors.
-It also reflects a limitation of our study and of the literature we set out to retrieve: We are aware of no standard terms in use to identify the family of techniques that fit statistical models to similarity-based cohorts.
-The retrieved studies whose use of such techniques alone satisfied our criteria variably termed them individualized, personalized, local, and patient-specific models, among other terms (Table \ref{tab:composite}).
-We prefer the term localized models, which is both sensitive and specific to uses of a similarity or distance measure to retrieve a cohort to which a model is fit: It includes some properly included techniques we did not have in mind [@Vilhena2016] yet excludes other techniques commonly referred to as individualized, personalized, or precision.
-
 ### Localized models
 
-In addition to @Lowsky2013, @Liang2015, @Ng2015, @Lee2017, @Vilhena2016, @Tang2021, and @Ng2021, seven other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, @Ma2020, @Liu2022, and @Doborjeh2022.
-In most cases these models were purely predictive in application; of the exceptions, @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did @Tang2021, @Ng2021, and @Liu2022 later.
-The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016] or via cross-validation [@Lowsky2013; @Verma2015].
-Further in the former direction, one study [@Ma2020] set a fixed cohort size while two [@Liang2015; @Tang2021; @Ng2021] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
-
-Several themes emerged from this sample.
-Foremost was the optimization of retrieved cohort sizes for some measure of model performance.
+\hl{Seven} other studies employed localized models: @Mariuzzi1997, @Lee2015, @Verma2015, @Wang2019, @Ma2020, @Liu2022, and @Doborjeh2022.
+\hl{Most} models were purely predictive\hl{, though} @Mariuzzi1997 produced purely descriptive models of survival outcomes, which were evaluated for their precision rather than their accuracy, while @Ng2015 used generalized regression models descriptively as well as predictively, as did @Tang2021, @Ng2021, and @Liu2022\hl{.} The most common approach to cohort selection was to optimize a size threshold via manual exploration [@Mariuzzi1997; @Lee2015; @Ng2015; @Lee2017; @Wang2019; @Vilhena2016] or via cross-validation [@Lowsky2013; @Verma2015].
+\hl{One} study [@Ma2020] set a fixed cohort size while two [@Liang2015; @Tang2021; @Ng2021] devised more sophisticated algorithms to balance multiple desiderata including predictive accuracy.
+
+\hl{One theme} was the optimization of retrieved cohort sizes for some measure of model performance.
 Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study [@Mariuzzi1997] and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach, though @Liu2022 took a third option by retrieving fixed proportions of training data.
-Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN prediction (viewed here as a special case of localized modeling wherein the model is a simple summary statistic), so we will not discuss it further.
-
-An alternative approach was to optimize multiple hyperparameters together: @Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015.
+Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN \hl{prediction}, so we will not discuss it further.
+
+@Kasabov2010 proposed an iterative algorithm to settle on an optimal number and set of predictors as well as number of training cases (comprising the retrieved cohort) that was later used by @Liang2015.
 More recently, @Tang2021 and @Ng2021 proposed a three-step process for cohort selection that filtered by exact match for one subset of predictors, used domain-informed similarity measures to rank these, and optimized the similarity threshold for a trade-off between cohort size and a measure of bias called cohort balance.
-Notably, though their similarity measure was supervised, this optimization process was not.
-
-With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus concerned itself with the interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
+\hl{T}hough their similarity measure was supervised, this optimization process was not.
+
+With the exception of the descriptive survival models of @Mariuzzi1997, only one thread in this corpus \hl{concerned the} interpretability of localized models: @Ng2015 fit logistic regression models to similarity-based cohorts and examined not only their predictive performance but the localized sets of largest and most detectable predictors, termed risk profiles, and how they differed across the testing population.
 This approach informed their later use of localized comparative effectiveness–style models to recommend courses of treatment at decision points during a monitored patent’s stay [@Tang2021; @Ng2021], and it was later used by @Liu2022 to show a dependency of predictor importance on the patient disease subgroup.
-These studies suggest much wider potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].
+These studies suggest \hl{wider} potential for localized real-world evidence generation, which has long been promoted as a promise of advances in clinical and health informatics [@Longhurst2014].
 
@@
 
-This relatively small sample exhibited a wide range of approaches, both technically and conceptually.
-Discussion of the technical diversity of these approaches has been summarized above and published in greater detail in previous reviews [@Choudhury2016; @Sharafoddini2017; @Parimbelli2018], and we focus here on the conceptual.
-The design of a localized modeling approach can be decomposed into three largely independent choices: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
-The choices are the _similarity measure_, the _cohort retrieval_, and the _statistical model_, respectively.
-These are diagrammed in Figure \ref{fig:framework}.
+This relatively small sample exhibited a wide range of \hl{designs in both the assembly of pre-processing, retrieval, reuse, and interpretation steps and the specific tools deployed at each}.
+Discussion of the \hl{diversity} of \hl{tools} has been summarized above and published in greater detail in previous reviews [@Choudhury2016; @Sharafoddini2017; @Parimbelli2018]\hl{.
+In this section, we focus on the diversity of assemblies.
+These depend on} three largely independent choices \hl{(Figure }\ref{fig:framework}\hl{})}: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
 
@@
 
-Each step can be unsupervised supervised: Unsupervised similarity measures include discrete measures like the Levenshtein distance and continuous measures like cosine similarity, while supervised measures include the Mahalanobis distance and \hl{RF} proximity as well as several composite measures with weights calculated from the data.
+Each step can be unsupervised supervised: Unsupervised similarity measures include \hl{the} Levenshtein distance and \hl{cosine} similarity, while supervised measures include the Mahalanobis distance and \hl{RF} proximity as well as several composite measures with weights calculated from the data.
 Most retrieval steps obtained cohorts with a uniform size (cardinality) or similarity threshold, and the remaining were likewise only informed by predictors; thus, while a retrieval step could in principle be supervised, in this sample none were.
@@
       "Specializations of general framework to included studies.",
-      " Flags: R = recurse, S = supervised, T = tuned (optimized)."
+      " Flags: R = recurse\\hl{d}, S = supervised, T = tuned (optimized)."
     )
@@
 Such software architectures must be robust, versatile, and compliant in the face of diverse data models and use restrictions.
-Localized modeling is susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures.
+SCBM is susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures.
 Individual systems exhibit many dimensions of incompleteness }[@Weiskopf2013]\hl{ and vary along other dimensions of quality }[@Kohane2021]\hl{, and their aggregation for modeling purposes depends on several kinds of interoperability }[@Weber2015]\hl{.
@@
 \hl{We synthesized these largely independently developed approaches into a general framework that we believe} can serve to taxonomize ongoing \hl{work }and inform \hl{future development}.
-Indeed, the availability of increasing computational power\hl and} the diversity of tasks to which these models were applied and of technical specifications they \hl{employed suggest }potential for growth.
+Indeed, the availability of increasing computational power\hl{ and} the diversity of tasks to which these models were applied and of technical specifications they \hl{employed suggest }potential for growth.
 \hl{Whereas} few of the reviewed studies used these models for any task other than prediction, despite widespread suspicion among clinicians of "black-box" models and growing interest in interpretable alternatives, we recommend that future work \hl{emphasize the} validation of interpretable localized models \hl{and their accessibility to} communities of medical research and clinical practice.
@@
 * CIS: clinical information system
-* \hl{CCDS: computerized} clinical decision support
+* \hl{CDSS:} clinical decision support\hl{ system}
 * AI: artificial intelligence
@@
 \hl{During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival }[@Mariuzzi1997]\hl{ and survival curves }[@Lowsky2013]\hl{.
-These approaches were novel to CBR and presaged later developments, so they were helpful to understanding the development of localized modeling.
+These approaches were novel to CBR and presaged later developments, so they were helpful to understanding the development of SCBM.
 However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the cohort's responses.
@@
 Then, JCB then excluded on account of a study being (d) indicated by the intended meanings of our search terms.
+
+## Terminology {-}
+
+\hl{As one reviewer pointed out for many if not most models used in ML a predictor may play a different and more or less important role for some cases than others, as quantified by importance measures or model-agnostic explanations }[@Biecek2021; @Molnar2023]\hl{.
+In this way, such models can also be viewed as "localized".
+We recognize, for example, that the use of global and local explanations is standard.
+As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
+The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
+A third reviewer pointed out that "local" may describe models fitted to data collected at a specific site or using features extracted from a specific patch of an image.
+Taken together, these issues convinced us to drop the shorthand "localized model" and adopt the more specific term "similarity cohort--based model".}
 

### 2025-03-07 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
fix typos & update latex source

@@
 
-We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators---Kohonen + CBR [@Wyns2004], TWNFI [@Song2006, @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]---are hybrids of \hl{SCBM} (in some cases CBR) with other techniques, often DML.
+We do observe one clear pattern: All of the proposed methods that most evidently outperformed their comparators---Kohonen + CBR [@Wyns2004], TWNFI [@Song2006; @Kasabov2010], gravitational search algorithms (GSA) [@Liang2015], CBR + rules (+ DML) [@Nicolas2014], Gaussian process regression [@Zhang2018], JITL-ELM [@Ma2020]---are hybrids of \hl{SCBM} (in some cases CBR) with other techniques, often DML.
 Though @CampilloGimenez2013 and @Ng2015 report non-superior performance by such hybrids, the pattern suggests the importance of the similarity measure to the retrieval step.
@@
 We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
-\hl{While "local" naturally contrasts with "global", the overloaded term "localized models" stands to cause more confusion than it resolves, so we use "similarity cohort--based models" to refer to this specific approach.}
+\hl{While ``local'' naturally contrasts with ``global'', the overloaded term ``localized models'' stands to cause more confusion than it resolves, so we use ``similarity cohort--based models'' to refer to this specific approach.}
 
@@
 We initially included each term and their synonyms separately to evaluate what resulted.
-We pruned several terms that returned no results ("phrases not found") or to reduce the number of results.
+We pruned several terms that returned no results (``phrases not found'') or to reduce the number of results.
 At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.}
@@
 \hl{As one reviewer pointed out for many if not most models used in ML a predictor may play a different and more or less important role for some cases than others, as quantified by importance measures or model-agnostic explanations }[@Biecek2021; @Molnar2023]\hl{.
-In this way, such models can also be viewed as "localized".
+In this way, such models can also be viewed as ``localized''.
 We recognize, for example, that the use of global and local explanations is standard.
-As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) [@Moshawrab2023; @Brauneck2023].
+As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) }[@Moshawrab2023; @Brauneck2023]\hl{.
 The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
-A third reviewer pointed out that "local" may describe models fitted to data collected at a specific site or using features extracted from a specific patch of an image.
-Taken together, these issues convinced us to drop the shorthand "localized model" and adopt the more specific term "similarity cohort--based model".}
+A third reviewer pointed out that ``local'' may describe models fitted to data collected at a specific site or using features extracted from a specific patch of an image.
+Taken together, these issues convinced us to drop the shorthand ``localized model'' and adopt the more specific term ``similarity cohort--based model''.}
 

### 2025-03-07 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
fix typos & update latex source

--- 2024-06-17 docs/localized-modeling-medRxiv/localized-modeling-medRxiv.tex
@@
-\documentclass{article}
-
-\usepackage{arxiv}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{graphicx}
-
-\title{Clinical prediction with localized modeling using
-similarity-based cohorts: A scoping review}
-
-\author{
-    Adi Cohen
-   \\
-    College of Osteopathic Medicine, Nova Southeastern University \\
-  Davie, FL \\
-  \texttt{\href{mailto:ac4569@mynsu.nova.edu}{\nolinkurl{ac4569@mynsu.nova.edu}}} \\
-   \And
-    Patti McCall-Junkin
-   \\
-    Smathers Libraries Academic Research \& Consulting Services,
-University of Florida \\
-  Gainesville, FL \\
-  \texttt{\href{mailto:pattimccall.junkin@gmail.com}{\nolinkurl{pattimccall.junkin@gmail.com}}} \\
-   \And
-    Jason Cory Brunson
-   \\
-    Laboratory for Systems Medicine, University of Florida \\
-  Gainesville, FL \\
-  \texttt{\href{mailto:jason.brunson@medicine.ufl.edu}{\nolinkurl{jason.brunson@medicine.ufl.edu}}} \\
-  }
+%Version 2.1 April 2023
+% See section 11 of the User Manual for version history
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[sn-mathphys,Numbered,pdflatex]{sn-jnl}
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published
+%%%%  by Springer Nature. The guidance has been prepared in partnership with
+%%%%  production teams to conform to Springer Nature technical requirements.
+%%%%  Editorial and presentation requirements differ among journal portfolios and
+%%%%  research disciplines. You may find sections in this template are irrelevant
+%%%%  to your work and are empowered to omit any such section if allowed by the
+%%%%  journal you intend to submit to. The submission guidelines and policies
+%%%%  of the journal take precedence. A detailed User Manual is available in the
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+%% Per the spinger doc, new theorem styles can be included using built in style, 
+%% but it seems the don't work so commented below
+%\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}%
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+%% \theoremstyle{thmstyletwo}%
+\theoremstyle{remark}
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+%% \theoremstyle{thmstylethree}%
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}%
+
+\usepackage{lscape}
+\newcommand{\landscapebegin}{\begin{landscape}}
+\newcommand{\landscapeend}{\end{landscape}}
+
+\usepackage{soul}
+
+\raggedbottom
 
@@
 
-% Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-\usepackage{lscape}
-\newcommand{\landscapebegin}{\begin{landscape}}
-\newcommand{\landscapeend}{\end{landscape}}
 \begin{document}
-\maketitle
-
-\begin{abstract}
-Background: Applications of classical case-based reasoning (CBR) have
-given rise to a family of techniques we call ``localized models'', in
-which a statistical model is fitted to a neighborhood of labeled cases
-matched by similarity to a target case. We aim to describe clinical and
-health applications of localized models to date and propose a general
-framework for their design and evaluation.
-
-Methods: We searched four bibliographic platforms during 2021 July
-19--22, updated 2024 January 24. We set four eligibility criteria to
-identify applications of localized models to clinical and health tasks.
+
+\title[Clinical prediction using similarity cohort--based
+models]{Clinical prediction using similarity cohort--based models: A
+systematic scoping review}
+
+%%=============================================================%%
+%% Prefix	-> \pfx{Dr}
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% NatureName	-> \tanm{Poet Laureate} -> Title after name
+%% Degrees	-> \dgr{MSc, PhD}
+%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
+%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\author[1]{\fnm{Adi} \sur{Cohen} \dgr{BS}}
+
+\author[2]{\fnm{Patti} \sur{McCall-Junkin} \dgr{MA, MLS}}
+
+\author*[3]{\fnm{Jason
+Cory} \sur{Brunson} \dgr{PhD}}\email{\href{mailto:jason.brunson@medicine.ufl.edu}{\nolinkurl{jason.brunson@medicine.ufl.edu}}}
+
+  \affil[1]{\orgdiv{College of Osteopathic Medicine}, \orgname{Nova
+Southeastern University}}
+  \affil[2]{\orgdiv{Smathers Libraries Academic Research \& Consulting
+Services}, \orgname{University of Florida}}
+  \affil*[3]{\orgdiv{Laboratory for Systems
+Medicine}, \orgname{University of
+Florida}, \orgaddress{\city{Gainesville}, \country{USA}, \state{Florida}, \street{P.O.
+Box 100225 JHMHC}}}
+
+\abstract{\textbf{Background:} Applications of classical case-based
+reasoning (CBR) have given rise to a family of modeling techniques
+\hl{reliant on similarity-based cohorts}, in which a statistical model
+is fitted to a neighborhood of labeled cases matched by similarity to a
+target case.
+
+\textbf{\hl{Objective:}} We aim to describe clinical and health
+applications of \hl{similarity cohort--based models (SCBMs)} to date and
+propose a general framework for their design and evaluation.
+
+\textbf{Methods:} We \hl{conducted searches of} four bibliographic
+platforms during 2021 July 19--22\hl{ and updated them on} 2024 January
+24. We set four eligibility criteria to identify applications of
+\hl{SCBMs} to clinical and health
+tasks\hl{, which amounted to using a multivariate similarity measure to retrieve cases from a retrospective corpus relevant to new cases and fitting statistical models to these cohorts}.
 Two authors divided title/abstract screening and reviewed screened
@@
 
-Results: Of 1,657 search results, 360 were reviewed, then combined with
-43 publications that seeded the review and 1 obtained by citation
-tracking. 27 were included, published 1997--2022. The specificity of
-search terms was poor, and inter-rater reliability was low. Almost all
-models were predictive, the most common tasks being prognosis and
-diagnosis. Most studies used clinical, occasionally laboratory and
-image, data. Several addressed memory and runtime costs. A general
-technique that specializes to most of those reviewed involved matching,
-retrieval, fitting, and evaluation steps that could optionally be
-supervised, optimized, or recursively performed.
-
-Conclusions: Localized models have potential to improve the performance
-of clinical decision support tools while maintaining interpretability,
-but rigorous comparisons to competing methods must be conducted and
-computational hurdles must be overcome. We hope that our review will
-spur future work on efficiency, reproducibility, and user needs.
-\end{abstract}
-
-\keywords{
-    case-based reasoning
-   \and
-    localized modeling
-   \and
-    nearest neighbors
-   \and
-    clinical decision support
-   \and
-    scoping review
-  }
+\textbf{Results:} Of 1,657 search results, 360 were reviewed, then
+combined with 43 publications that seeded the review and 1 obtained by
+citation tracking. 27 were included, published 1997--2022. The
+specificity of search terms was poor\hl{ (51/1,657)}, and inter-rater
+reliability was low\hl{ (82\%)}. Almost all models were predictive, the
+most common tasks being prognosis\hl{ (11)} and diagnosis\hl{ (10)}.
+Most studies used clinical\hl{ (22)}, occasionally laboratory\hl{ (7)}
+or imaging\hl{ (2)}, data.
+\hl{The most competitive performances were reported for models that further hybridized the basic approach.}
+A general technique that specializes to most of those reviewed involved
+matching, retrieval, fitting, and evaluation steps that could optionally
+be supervised, optimized, or recursively performed.
+
+\textbf{Conclusions:} \hl{SCBMs} have potential to improve the
+performance of clinical decision support tools while maintaining
+interpretability, but rigorous comparisons to competing methods must be
+conducted and computational hurdles must be overcome. We hope that our
+review will spur future work on efficiency, reproducibility, and user
+needs.}
+
+\keywords{case-based reasoning, similarity cohort, nearest
+neighbors, clinical decision support, scoping review}
+
+\pacs[JEL Classification]{C38, C65}
+\pacs[MSC Classification]{62P10}
+
+\maketitle
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Background}\label{background}
 
 A main driver of the development of clinical information systems (CIS)
-and clinical decision support (CDS) tools has been to leverage the
-advantages of routinely collected health data toward improving care.
-These advantages include their immediate accessibility through the
-institutional EHR, billing database, or other source; their specificity
-to the institution and the population it serves; and the closeness in
-time of the available data. Though by definition not collected for
-research use, the secondary research use of routinely collected health
-data has been put forth, as ``practice-based evidence'', a complement to
-the paradigm of evidence-based practice.
+and \hl{clinical decision support systems (CDSS)} has been to leverage
+the advantages of routinely collected health data toward
+\hl{improving care }\citep{Benchimol2015, Wasylewicz2019}. These
+advantages include their immediate accessibility through the
+institutional \hl{electronic health record (EHR)}, billing database, or
+other source; their specificity to the institution and the population it
+serves; and the closeness in time of the available data. \hl{Though not}
+collected for research use, the secondary research use of routinely
+collected health data has been put forth, as ``practice-based
+evidence'', a complement to the paradigm of
+\hl{evidence-based practice }\citep{Sim2001, Holmqvist2015}.
 
@@
 population-derived evidence-based recommendations to specific patient
-needs has driven the application of advanced computational tools,
+needs has driven the application of
+\hl{advanced computational tools }\citep{Rosella2022, Elhaddad2024},
 including artificial intelligence (AI). Whereas classical models
@@
 more variables in opaque ways to yield predictions that depend on not
-only their values but their associations and interplay with each other.
+only their values but
+\hl{their associations and interplay with each other }\citep{Molnar2023}.
 The loss of straightforward interpretations of these models has limited
 their practical uptake and motivated the construction of explanatory
-statistics for opaque models as well as the development of more
-interpretable complex models.
+statistics for opaque models as well as the development of
+\hl{more interpretable complex models }\citep{Rudin2022, Molnar2023}.
 
 One of few methodologies to address all three of these established needs
-is one we call \emph{localized modeling}, which appears to have been
-introduced several times independently under different names, and in
-slightly different forms. The approach is a specialized form of
-case-based reasoning (CBR) that relies on a patient similarity
-measure\footnote{Similarity measures between more granular units of
-  analysis, such as encounters and decision points, are often still
-  referred to as patient similarity measures.} to extract a cohort of
-past or training cases that then inform the diagnosis, prognosis, or
-care of a new or test case. Traditional CBR returns these retrieved
-cases to inform human decision-making; the related nearest neighbors
-(NN) technique generates predictions from cohorts automatically via
-averaging (regression) or voting (classification) of their outcomes. In
-contrast, localized modeling fits families of predictive models, for
-example generalized linear models (GLMs), to retrieved similarity
-cohorts in order to generate predictions. Localized models thus provide
-an individualized way to fit fully interpretable models to large
-population data.
+is \hl{similarity cohort--based modeling (SCBM)}, which
+\hl{we find in this review} to have been introduced several times
+independently under different names, and in slightly different forms.
+The approach is a specialized form of case-based reasoning (CBR) that
+relies on a \hl{patient similarity measure }\citep{Brown2016} to extract
+a cohort of past or training cases that then inform the diagnosis,
+prognosis, or care of a new or test case.\footnote{Similarity measures
+  between more granular units of analysis, such as encounters and
+  decision points\hl{, or between patients' time series}, are often
+  still referred to as patient similarity measures\hl{ in our sample}.}
+\hl{Some of the earliest and simplest CBR tools returned} these
+retrieved cases \hl{ to inform human decision-making}\citep{Aamodt1994};
+the related nearest neighbors (NN) technique generates predictions from
+cohorts automatically via averaging (regression) or voting
+(classification) of their outcomes. In contrast, \hl{SCBM} fits families
+of predictive models, for example generalized linear models (GLMs), to
+retrieved similarity cohorts in order to generate predictions.
+\hl{While many ML models come equipped with measures of feature importance and model-agnostic tools can generate explanations for model predictions }\citep{Molnar2023}\hl{, these quantifications are not directly interpretable model components analogous to the split nodes of a DT or the coefficients of a GLM.
+While the patient similarity measure used to retrieve each cohort may be complicated, the cohort itself can be directly inspected.}
+
+\hl{This approach offers an alternative paradigm for interpretable machine learning that combines the simplicity and transparency of classical models like decision trees (DTs) and GLMs with the sensitivity to nonlinearity and interactions of advanced opaque models like random forests (RFs) and artificial neural networks (ANNs).
+The sensitivity is enabled by the fitting of the model to the local cohort, whose size and scope are governed by hyperparameters that may be tuned like any others via grid or Bayesian optimization.
+Some similarity measures or other retrieval methods may be more interpretable than others, so that the interpretability of the SCBM approach would then depend on that of its retrieval (similarity matching) and revision (predictive modeling) steps.
+Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures and algorithms exist for similarity matching }\citep{Halder2024}\hl{ and the local models would benefit from increased numbers of relevant cases.}
+\hl{SCBM}s thus provide an individualized way to fit fully interpretable
+models to large population data.
 
@@
 institution's records with which to conduct on-demand retrospective
-studies for an individual patient (Longhurst, Harrington, and Shah
-2014). Several hurdles face the deployment of such an approach in
-practice, but its benefits must first be demonstrated. Our motivations
-in this scoping review are to describe the settings in and problems with
-which this approach has been tasked, to attempt to organize them within
-a common framework, and to evaluate the promise they show toward
-achieving this or other needs in clinical informatics. Along the way, we
-attempt to reconcile terminology, summarize motivations and evaluations,
-and propose valuable follow-up work.
-
-\hypertarget{related-work}{%
-\subsection{Related work}\label{related-work}}
+studies for an individual patient \citep{Longhurst2014}. Several hurdles
+face the deployment of such an approach in practice, but its benefits
+must first be demonstrated. Our motivations in this scoping review are
+to describe the settings in and problems with which this approach has
+been tasked, to attempt to organize them within a common framework, and
+to evaluate the promise they show toward achieving this or other needs
+in clinical informatics. Along the way, we attempt to reconcile
+terminology, summarize motivations and evaluations, and propose valuable
+follow-up work.
+
+\subsection{\texorpdfstring{Related
+\hl{survey }work}{Related work}}\label{related-work}
 
 Clinical CBR emerged among rule-based approaches and other AI tools in
-the development of expert systems (Aamodt and Plaza 1994). Early
+the development of expert systems \citep{Aamodt1994}. Early
 implementations realized a general workflow described as ``the four
-REs'', later the ``R4 cycle'' (Aamodt and Plaza 1994; Begum et al.
-2011): Given a new case (or problem), the system \emph{retrieves} one or
-more past cases (solved problems) from a corpus, \emph{reuses} these to
-generate an understanding of (solution to) the new case, \emph{revises}
-this understanding (solution) to better fit the new case (sometimes
-called ``adaptation''), and \emph{retains} the new case and its eventual
+REs'', later the ``R4 cycle'' \citep{Aamodt1994, Begum2011}: Given a new
+case (or problem), the system \emph{retrieves} one or more past cases
+(solved problems) from a corpus, \emph{reuses} these to generate an
+understanding of (solution to) the new case, \emph{revises} this
+understanding (solution) to better fit the new case (sometimes called
+``adaptation''), and \emph{retains} the new case and its eventual
 understanding (solution) in the corpus to be retrieved and reused in
@@
 case-based systems provide case-specific interpretations in the form of
-a number of more fully understood reference cases. While the similarity
-measure used in the retrieval step need not be changed as the corpus
-grows, proposed measures have been diverse, contested, and rarely
-systematically compared.
-
-Kolodner (1992) distinguished two styles of CBR: problem-solving, which
-is more procedural and used when objectives are more clearly defined,
-and interpretative, which provides categorizations and justifications
-for possible solutions. A similar dichotomy is commonly used to
-distinguish the performance criterion for the usefulness of predictive
-models from the interpretability criterion for their usability. In these
-terms, localized models are highly procedural: The step of fitting a
-predictive model to a retrieved cohort is an automated adaptive strategy
-(Begum et al. 2011), and the parameters that govern cohort retrieval can
-be tuned alongside model hyperparameters in a machine learning (ML)
-workflow. Accordingly, localized models have been primarily used for and
-evaluated on their ability to predict classes or outcomes. Nevertheless,
-as some recent applications have shown, they can be used to draw
-inferences about the importance of different risk factors to specific
-individuals. While most ML models come equipped with measures of feature
-importance and model-agnostic tools can generate explanations for model
-predictions, these quantifications are not directly interpretable model
-components analogous to the split nodes of a decision tree (DT) or the
-coefficients of a GLM. While the patient similarity measure used to
-retrieve each cohort may be complicated, the cohort itself can be
-directly inspected by the user. Provided the model family fitted to the
-cohorts is interpretable, the localized model inherits this property.
-Thus, localized models may embody both styles of CBR.
+a number of more fully understood reference cases.
+
+\citet{Kolodner1992} distinguished two styles of CBR:
+\emph{problem-solving}, which is more procedural and used when
+objectives are more clearly defined, and \emph{interpretative}, which
+provides categorizations and justifications for possible solutions. A
+similar dichotomy is commonly used to distinguish the \emph{performance}
+\hl{criterion }of predictive models from the \emph{interpretability}
+criterion\hl{.} In these terms, \hl{SCBM}s are highly procedural: The
+step of fitting a predictive model to a retrieved cohort is an automated
+adaptive strategy \citep{Begum2011}, and the parameters that govern
+cohort retrieval can be tuned alongside model hyperparameters in a
+machine learning (ML) workflow. Accordingly, \hl{SCBM}s have been
+primarily used for and evaluated on their ability to predict classes or
+outcomes. Nevertheless, as \hl{we find in this review}, they can be used
+to draw inferences about the importance of different risk factors to
+specific individuals.
+
+\hl{We also note that a large variety of AI tools, besides NN, use similarity measures or other relevance-based retrieval methods: Back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
+that small distances between cases yield small changes in prediction; and kernel-based methods that rely directly on pairwise distances computed via inner product, rather than high-dimensional coordinates, extend far beyond their use in CBR.
+The key difference in SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
+Indeed, retrieval procedures were the most significant algorithmic requirements and accomplishments of CBR }\citep{Kolodner1992, Gierl1998}\hl{ and the retrieved cohorts may be reused for numerous modeling tasks.}
+Thus, \hl{SCBM}s may embody both styles of CBR.
 
 We refer the interested reader to several previous reviews of CBR in
-medicine (Gierl, Bull, and Schmidt 1998; Begum et al. 2011; Choudhury
-and Begum 2016), of measures of patient similarity (Dai, Zhu, and Liu
-2020), and of uses of patient similarity in predictive models (Welch and
-Kawamoto 2013; Sharafoddini, Dubin, and Lee 2017; Parimbelli et al.
-2018). While the reviews of CBR focus on applications using health data,
-the patient similarity reviews encompass many additional types of data
+medicine \citep{Gierl1998, Begum2011, Choudhury2016}, of measures of
+patient similarity \citep{Dai2020}, and of uses of patient similarity in
+predictive models \citep{Welch2013, Sharafoddini2017, Parimbelli2018}.
+While the reviews of CBR focus on applications using health data, the
+patient similarity reviews encompass many additional types of data
 (various molecular -omics, genetic tests, medical images, laboratory
@@
 analysis via dimension reduction and cluster analysis; risk evaluation
-and outcome prediction; clinical decision support and software tools).
-For the present review, we are interested in how similarity matching on
-patient-level health data is used to construct local cohorts for
-predictive or inferential modeling.
-
-\hypertarget{objectives}{%
-\subsection{Objectives}\label{objectives}}
-
-Our goals in this review are (1) to describe applications of
-similarity-based localized models using patient-level health data and
-(2) to provide a general framework for the design and evaluation of such
-localized models. We focus narrowly on localized models, rather than
-broadly on patient similarity--based clinical decision support tools, so
-that we may thoroughly assess their value in terms of reported
-evaluations and comparisons to other methods. This also allows us to
-devise a high-level framework that specializes to the majority of these
-methods, which we then use to frame our discussion and recommendations.
-
-Our selection of relevant papers from the search corpus will be based on
-the following inclusion/exclusion criteria:
-
-\begin{itemize}
-\tightlist
-\item
-  Uses case-level data from a corpus of past cases with known
-  responses\newline (response may be outcome, diagnosis, subtype, etc.)
-\item
-  Defines a numeric multivariate case similarity measure\newline (allow
-  integer-valued measures)
-\item
-  Uses the similarity measure to retrieve cohorts for index cases from
-  the corpus\newline (for example, based on a training--testing
-  partition)
-\item
-  Fits statistical models to cohorts from which to make predictions or
-  draw inferences about index cases\newline (outcome predictions,
-  survival estimates, risk factor contributions, model evaluation
-  statistics, etc.)
-\end{itemize}
+and outcome prediction; \hl{CDSS} and software tools). For the present
+review, we are interested in how similarity matching on patient-level
+health data is used to construct local cohorts for predictive or
+inferential
+modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.}
+
+\subsection{Objectives}\label{objectives}
+
+Our goals in this review are (1) to describe applications of similarity
+cohort--based models using patient-level health data and (2) to provide
+a general framework for the design and evaluation of such \hl{SCBM}s. We
+focus narrowly on \hl{SCBM}s, rather than broadly on patient
+similarity--based clinical decision support tools, so that we may
+thoroughly assess their value in terms of reported evaluations and
+comparisons to other methods. This also allows us to devise a high-level
+framework that specializes to the majority of these methods, which we
+then use to frame our discussion and recommendations.
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section{Methods}\label{methods}
 
 Here we describe our review process, including deviations from plans and
-the reasons for them. More details are included in Section
-\ref{selection-process}.
-
-\hypertarget{prisma-checklist}{%
-\subsection{PRISMA checklist}\label{prisma-checklist}}
+the reasons for them. More details are included in \hl{the Appendix}.
+
+\subsection{PRISMA checklist}\label{prisma-checklist}
 
@@
 abstract checklist as Supplemental Table 2. Because we focus on
-methodologies rather than conceptual approaches or evidence, we deviate
-in some ways from PRISMA guidelines. In particular, those items of the
-checklist involving bias assessment and quantitative synthesis are
-intended for meta-analyses so did not apply to this study. Because we
-are not aware of any standard procedures for conducting reviews and
-syntheses of methodology, no protocol was prepared for this study.
-
-\hypertarget{search}{%
-\subsection{Search}\label{search}}
-
-We derived the eligibility criteria itemized in the Introduction from a
-seed set of previously read studies. Based on these criteria, we
-formulated search strings for five platforms: PubMed, Web of Science,
-Academic Search Premier, Google Scholar, and MathSciNet. We finalized
-the PubMed search first, then adapted it to the other platforms (Section
-\ref{search-design}).
-
-Through each platform, we searched those databases included by default.
-This means that we searched both MEDLINE and PubMed Central (PMC)
-through Pubmed (we did not exclude other databases, but our earliest
-included results post-date them) and that we searched the six indices of
-the Core Collection (the Science Citation Index Expanded, the Social
-Sciences Citation Index, the Arts \& Humanities Citation Index, the
-Emerging Sources Citation Index, the Conference Proceedings Citation
-Index, and the Book Citation Index) as well as several regional
-databases through the Web of Science platform.
-
-The structure and terms of our search strings were inspired in part by
-previous reviews adjacent to our topic of interest (Sharafoddini, Dubin,
-and Lee 2017; Parimbelli et al. 2018). We organized the PubMed search in
-disjunctive normal form (an OR of ANDs). Following the solidification of
-an outline, we expanded each term to include synonyms that are similar
-enough to be applicable to our search. We then evaluated the expanded
-search string using the PubMed Advanced Search platform. We initially
-included each term and their synonyms separately to evaluate what
-resulted. We pruned several terms that returned no results (``phrases
-not found'') or to reduce the number of results. At the conclusions of
-this process for each individual term, we combined the search terms
-using disjunctive normal form.
+methodologies rather than\hl{ evidence}, we deviate in some ways from
+PRISMA guidelines. In particular, those items of the checklist involving
+bias assessment and quantitative synthesis are intended for
+meta-analyses so did not apply to this study. Because we are not aware
+of any standard procedures for conducting reviews and syntheses of
+methodology, no protocol was prepared for this study.
+
+\subsection{Eligibility criteria}\label{eligibility-criteria}
+
+\hl{Our selection of relevant papers from the search corpus will be based on the following inclusion/exclusion criteria:}
+
+\begin{itemize}
+\tightlist
+\item
+  \hl{Uses case-level data from a corpus of past cases with known responses}\newline
+  \hl{(response may be outcome, diagnosis, subtype, etc.)}
+\item
+  \hl{Defines a numeric multivariate case similarity measure}\newline
+  \hl{(allow integer-valued measures)}
+\item
+  \hl{Uses the similarity measure to retrieve cohorts for index cases from the corpus}\newline
+  \hl{(for example, based on a training--testing partition)}
+\item
+  \hl{Fits statistical models to cohorts from which to make predictions or draw inferences about index cases}\newline
+  \hl{(outcome predictions, survival estimates, risk factor contributions, model evaluation statistics, etc.)}
+\end{itemize}
+
+\hl{Note that classical CBR satisfies the first and third criteria by definition and in most cases will satisfy the second.}
+
+\subsection{Search}\label{search}
+
+We derived the eligibility \hl{criteria }from a seed set of previously
+read studies. Based on these criteria, we formulated search strings for
+five platforms: PubMed, Web of Science, Academic Search Premier, Google
+Scholar, and MathSciNet. We finalized the PubMed search first, then
+adapted it to the other platforms\hl{.}
 
@@
 
-\hypertarget{titleabstract-screen}{%
-\subsection{Title/abstract screen}\label{titleabstract-screen}}
-
-Within Covidence, we screened titles and abstracts for relevance. We
-decided on four eligibility criteria to expedite the screening process.
-Articles must be written in English, for readability; they must be
-original studies, to exclude secondary sources with duplicate
-information; their use settings must be medical, clinical, or related,
-to keep our review topical; and it must not be clear that their use of
-our search terms was different from our intended meaning. Each of two
-authors (AC and JCB) screened roughly half of the entries. They
-regularly reviewed each other's decisions to improve consistency. In
-cases of uncertainty, entries were included. For the update, one author
-(PM) screened all results for three criteria and another (JCB) screened
-the survivors for the fourth criterion.
-
-\hypertarget{full-text-review}{%
-\subsection{Full text review}\label{full-text-review}}
-
-We set out 4 criteria for full-text review to restrict to studies that
-used some form of localized modeling on health data:
-
-\begin{itemize}
-\tightlist
-\item
-  Pulls case-level data from a corpus of past cases with known classes
-  or outcomes
-\item
-  Defines a numeric multivariate case similarity measure
-\item
-  Uses the similarity measure to select cohorts for index cases from the
-  corpus
-\item
-  Fits statistical models to cohorts to make predictions or inferences
-  about index cases
-\end{itemize}
-
-Note that classical CBR satisfies the first and third criteria by
-definition and in most cases will satisfy the second.
-
-In Covidence, two authors (AC and JCB) independently evaluated each
-manuscript for these eligibility criteria. The first criterion that a
-manuscript failed was designated the reason for exclusion. An antecedent
-criterion was used to exclude manuscripts that did not report the
-results of original studies involving real-world experiments or
-empirical data, for example surveys of prior work and proposals of
-frameworks. In cases of disagreement between the authors on the reason
-for exclusion, the first criterion was adopted. The authors arrived at
-agreement on inclusion or exclusion through discussion. We calculated
-inter-rater reliability to evaluate our screening and review process.
-
-During full text review, we decided to expand the conception of
-statistical models (fourth criterion): Rather than restricting to models
-that are fit and evaluated in separate steps, we chose to allow simple
-statistical summaries such as mean survival (Mariuzzi et al. 1997) and
-survival curves (Lowsky et al. 2013). These approaches were novel to CBR
-and presage later developments, so were helpful to understanding the
-development of localized modeling. However, this then admitted studies
-that applied conventional nearest neighbors prediction: Each retrieved
-cohort consisted of the \(k\) most similar cases to the index case, and
-the response for the index case was predicted to be either the mean
-(continuous response) or the plurality (discrete response) of the
-cohort's responses. A review of all studies that use nearest neighbors
-prediction would be impractical. Because our focus is on novel
-approaches that combine similarity-based retrieval and statistical
-modeling of retrieved cohorts, we chose to exclude those studies whose
-approach was equivalent to nearest neighbors prediction using a
-conventional similarity measure.
+\subsection{Title/abstract screen}\label{titleabstract-screen}
+
+\hl{We} screened titles and abstracts for relevance. We decided on four
+eligibility criteria to expedite the screening process. Articles must be
+written in English, for readability; they must be original studies, to
+exclude secondary sources with duplicate information; their use settings
+must be medical, clinical, or related, to keep our review topical; and
+it must not be clear that their use of our search terms was different
+from our intended meaning. Each of two authors (AC and JCB) screened
+roughly half of the entries. They regularly reviewed each other's
+decisions to improve consistency. In cases of uncertainty, entries were
+included. For the update, one author (PM) screened all results for
+\hl{the first }three criteria and another (JCB) screened the survivors
+for the fourth criterion.
+
+\subsection{Full text review}\label{full-text-review}
+
+We \hl{used the} 4 criteria\hl{ from Section }\ref{eligibility-criteria}
+for full-text review\hl{.}
+
+\hl{Two} authors (AC and JCB) independently evaluated each manuscript
+for these eligibility criteria. The first criterion that a manuscript
+failed was designated the reason for exclusion. An antecedent criterion
+was used to exclude manuscripts that did not report the results of
+original studies involving real-world experiments or empirical data, for
+example surveys of prior work and proposals of frameworks. In cases of
+disagreement between the authors on the reason for exclusion, the first
+criterion was adopted. The authors arrived at agreement on inclusion or
+exclusion through discussion. We calculated inter-rater reliability to
+evaluate our screening and review process.
 
@@
 
-\hypertarget{coding}{%
-\subsection{Coding}\label{coding}}
+\subsection{Coding}\label{coding}
 
@@
 
-\hypertarget{synthesis}{%
-\subsection{Synthesis}\label{synthesis}}
+\subsection{Synthesis}\label{synthesis}
 
@@
 techniques. The goal will be to describe a unified framework for
-localized models that can be used to guide future study designs and
+\hl{SCBM}s that can be used to guide future study designs and
 implementations as well as more systematically evaluate variations on
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{selection}{%
-\subsection{Selection}\label{selection}}
+\section{Results}\label{results}
+
+\subsection{Selection}\label{selection}
 
@@
 
-\caption{PRISMA-S flow chart.
-Lowercase letters refer to items obtained from the seed set (m), the structured search (n), and citation/reference tracking (s).
-Within the structured search results, summands correspond to original and update.}\label{fig:prisma}
+\caption{\hl{PRISMA-S flow chart.}
+ Lowercase letters refer to items obtained from
+ the seed set (m),
+ the structured search (n), and
+ citation/reference tracking (s).
+ Within the structured search results,
+ summands correspond to original and update.}\label{fig:prisma}
 \end{figure}
@@
 
-We then reviewed 43 studies comprising a seed set that inspired this
-review. After removing duplicates and screening for eligibility, we were
-left with 6 additional studies (Park, Kim, and Chun 2006; Lowsky et al.
-2013; Lee, Maslove, and Dubin 2015; Ng et al. 2015; Lee 2017; N. Wang et
-al. 2019), 1 of which was excluded from the synthesis for using NN
-prediction. Reference tracking from the 43 + 360 = 403 studies assessed
-for eligibility led us to identify 1 additional study that met our
-criteria (Kasabov and Hu 2010). This left us with 21 + 1 + 5 = 27
-studies included in the review and synthesis.
-
-\hypertarget{bibliographic-and-methodological-properties}{%
+We then reviewed \hl{a seed set comprising 43 studies} that inspired
+this review. After removing duplicates and screening for eligibility, we
+were left with 6 additional studies
+\citep{Park2006, Lowsky2013, Lee2015, Ng2015, Lee2017, Wang2019}, 1 of
+which was excluded from the synthesis for using NN prediction. Reference
+tracking from the 43 + 360 = 403 studies assessed for eligibility led us
+to identify 1 additional study that met our criteria
+\citep{Kasabov2010}. This left us with 21 + 1 + 5 = 27 studies included
+in the review and synthesis.
+
 \subsection{Bibliographic and methodological
-properties}\label{bibliographic-and-methodological-properties}}
+properties}\label{bibliographic-and-methodological-properties}
 
@@
 \ref{fig:year}). While these studies trace back to the late 1990s, they
-have become more common, which suggests that this is an active, though
-not rapidly expanding, approach. We note that CBR in health and medicine
-originated as early as 1990, and localized modeling emerged soon after
-CBR had established itself; the idea has been ``in the air'' for as long
-as CBR has been in use.
+have become \hl{slightly }more common, which suggests that this is an
+active, though not rapidly expanding, approach. We note that CBR in
+health and medicine originated as early as 1990, and \hl{SCBM} emerged
+soon after CBR had established itself; the idea has been ``in the air''
+for as long as CBR has been in use.
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1148}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1585}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1202}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1967}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1530}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1038}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1530}}@{}}
-\caption{\label{tab:synthesis}Studies included in the method synthesis,
-arranged by the earliest the study is known to have been
-public.}\tabularnewline
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1154}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1209}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1978}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1044}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}@{}}
+\caption{\label{tab:synthesis}\hl{Studies included in the method synthesis, arranged by the earliest the study is known to have been public.}}\tabularnewline
 \toprule\noalign{}
@@
 \endlastfoot
-Yearwood and Wilkinson (1997) & Prediction\hspace{6em} &
+\citet{Yearwood1997} & Recommend-ation\hspace{6em} &
 Practice\hspace{6em} & Clinical\hspace{6em} &
 Cross-sectional\hspace{6em} & 1,355 & 9 \\
-Mariuzzi et al. (1997) & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Mariuzzi1997} & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Longitudinal\hspace{6em} & 113 & 4 \\
-Wyns et al. (2004) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Wyns2004} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
 Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 160 & 14 \\
-Park, Kim, and Chun (2006) & Diagnosis\hspace{6em} &
-Practice\hspace{6em} & Clinical, Imaging\hspace{6em} &
-Cross-sectional\hspace{6em} & 366; 270; 560; 760 & 35; 14; 31; 8 \\
-Song and Kasabov (2006) & Prediction\hspace{6em} & Practice\hspace{6em}
-& Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
+\citet{Park2006} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
+Clinical, Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 366; 270;
+560; 760 & 35; 14; 31; 8 \\
+\citet{Song2006} & Prognosis\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
 6 \\
-Elter, Schulz-Wendtland, and Wittenberg (2007) & Prediction\hspace{6em}
-& Practice\hspace{6em} & Clinical\hspace{6em} &
-Cross-sectional\hspace{6em} & 2,620 & 6 \\
-Xu et al. (2008) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Elter2007} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 2,620 & 6 \\
+\citet{Xu2008} & Prognosis\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 67 & 14 \\
-López et al. (2011) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Lopez2011} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 871 & 4 \\
-Kasabov and Hu (2010) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Kasabov2010} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62 & 2,000 \\
-Verma et al. (2015) & Decision Support\hspace{6em} &
-Practice\hspace{6em} & Clinical\hspace{6em} &
-Cross-sectional\hspace{6em} & 74 & 93 \\
-Liang, Hu, and Kasabov (2015) & Prediction\hspace{6em} &
-Practice\hspace{6em} & Clinical\hspace{6em} &
-Cross-sectional\hspace{6em} & 62; 72; 77; 181 & 2,000; 7,129; 7,129;
-12,533 \\
-Lowsky et al. (2013) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Verma2015} & Prognosis\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 74 & 93 \\
+\citet{Liang2015} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 62; 72; 77; 181 &
+2,000; 7,129; 7,129; 12,533 \\
+\citet{Lowsky2013} & Prognosis\hspace{6em} & Practice\hspace{6em} &
 Clinical, Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 13,525 &
 13 \\
-Campillo-Gimenez et al. (2013) & Decision Support\hspace{6em} &
+\citet{CampilloGimenez2013} & Recommend-ation\hspace{6em} &
 Practice\hspace{6em} & Patient-reported\hspace{6em} &
 Cross-sectional\hspace{6em} & 1,647 & 18 \\
-Nicolas et al. (2014) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Nicolas2014} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 150 & 83 \\
-Ng et al. (2015) & Clustering\hspace{6em} & Practice\hspace{6em} &
+\citet{Ng2015} & Prognosis\hspace{6em} & Practice\hspace{6em} &
 Clinical, Laboratory\hspace{6em} & Longitudinal\hspace{6em} & 7,519 &
 130 \\
-Lee, Maslove, and Dubin (2015) & Prediction\hspace{6em} &
-Knowledge\hspace{6em} & Clinical, Laboratory\hspace{6em} &
-Cross-sectional\hspace{6em} & 17,152 & 75 \\
-Vilhena et al. (2016) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Lee2015} & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152
+& 75 \\
+\citet{Vilhena2016} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,046 & 8 \\
-Lee (2017) & Prediction\hspace{6em} & Practice\hspace{6em} & Clinical,
-Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152 & 75 \\
-Zhang et al. (2018) & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
+\citet{Lee2017} & Prognosis\hspace{6em} & Practice\hspace{6em} &
+Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 17,152
+& 75 \\
+\citet{Zhang2018} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
 Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
-Malykh and Rudetskiy (2018) & Decision Support\hspace{6em} &
-Practice\hspace{6em} & Clinical\hspace{6em} &
-Cross-sectional\hspace{6em} & 638 & 49,728 \\
-Ma et al. (2020) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Malykh2018} & Recommend-ation\hspace{6em} & Practice\hspace{6em}
+& Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 638 & 49,728 \\
+\citet{Ma2020} & Prognosis\hspace{6em} & Practice\hspace{6em} &
 Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 4,000 & 22 \\
-N. Wang et al. (2019) & Diagnosis\hspace{6em} & Practice\hspace{6em} &
+\citet{Wang2019} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
 Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 16,490
 & 106 \\
-Y. Wang et al. (2020) & Decision Support\hspace{6em} &
-Practice\hspace{6em} & Clinical\hspace{6em} &
-Cross-sectional\hspace{6em} & 8 & 5 \\
-Tang et al. (2021); Ng et al. (2021) & Decision Support\hspace{6em} &
+\citet{Wang2020} & Recommend-ation\hspace{6em} & Practice\hspace{6em} &
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 8 & 5 \\
+\citet{Tang2021}; \citet{Ng2021} & Recommend-ation\hspace{6em} &
 Practice\hspace{6em} & Clinical\hspace{6em} & Longitudinal\hspace{6em} &
 245,825 & 1,466,474 \\
-Liu et al. (2022) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Liu2022} & Prognosis\hspace{6em} & Knowledge\hspace{6em} &
 Clinical, Laboratory\hspace{6em} & Cross-sectional\hspace{6em} & 76,597
 & 1,892 \\
-Doborjeh et al. (2022) & Prediction\hspace{6em} & Practice\hspace{6em} &
+\citet{Doborjeh2022} & Prognosis\hspace{6em} & Practice\hspace{6em} &
 Clinical, Environmental\hspace{6em} & Longitudinal\hspace{6em} & 804 &
@@
 to provide tools for use in clinic or to improve outcomes.
+
+\hl{Most studies used clinical data elements alone
+(16),
+and most of the remainder used these in combination with laboratory
+(6),
+imaging
+(1),
+or environmental
+(1)
+data. The data source was related to the modeling task, for instance in that all decision support models used either clinical or patient-reported data.
+The mosaic plot in Figure }\ref{fig:properties}\hl{ summarizes the relation between data source, clinical task, and aim.}
 
@@
 
-\hypertarget{application-domains}{%
-\subsection{Application domains}\label{application-domains}}
+\subsection{Application domains}\label{application-domains}
 
@@
 evaluated the usefulness of CBR-based explanations for the purpose of
-decision support (Doyle, Cunningham, and Walsh 2006). This study was
-part of a much larger literature on CBR systems and was included here
-despite relying on NN prediction because it used an unconventional
-voting scheme to generate recommendations. A more recent study took
-essentially the same focus with respect to a proposed clinical risk
-prediction model, which amounted to CBR with a novel weighting scheme on
-predictors informed by expert consensus (Fang et al. 2021). In both
-cases a prototype implementation was deployed in an experimental setting
-for evaluation.
+decision support \citep{Doyle2006}. This study was part of a much larger
+literature on CBR systems and was included here despite relying on NN
+prediction because it used an unconventional voting scheme to generate
+recommendations. A more recent study took essentially the same focus
+with respect to a proposed clinical risk prediction model, which
+amounted to CBR with a novel weighting scheme on predictors informed by
+expert consensus \citep{Fang2021}. In both cases a prototype
+implementation was \hl{evaluated} in an experimental \hl{setting.}
 
@@
 recommendation. The plurality focused on prognosis or outcome
-prediction, often using time-to-event analysis: Mariuzzi et al. (1997)
+prediction, often using time-to-event analysis: \citet{Mariuzzi1997}
 used CBR to predict survival time from several geometric properties of
-breast tumors. Lowsky et al. (2013) used CBR with non-parametric
-survival models on registry data to predict patient--graft survival
-times following kidney transplantation. Lee, Maslove, and Dubin (2015)
-and Lee (2017) used localized logistic regression and random forest
-modeling to predict 30-day mortality following discharge for ICU
-patients. Vilhena et al. (2016) developed a CBR cycle around a
-clustering-informed similarity matching procedure and an artificial
-neural network--based classifier to identify thrombophilia patients at
-high risk of thrombotic episodes. Ma et al. (2020) took a similarity
+breast tumors. \citet{Lowsky2013} used CBR with non-parametric survival
+models on registry data to predict patient--graft survival times
+following kidney transplantation. \citet{Lee2015} and \citet{Lee2017}
+used localized \hl{GLM} and RF modeling to predict 30-day mortality
+following discharge for ICU patients. \citet{Vilhena2016} developed a
+CBR cycle around a clustering-informed similarity matching procedure and
+an \hl{ANN}--based classifier to identify thrombophilia patients at high
+risk of thrombotic episodes. \citet{Ma2020} took a similarity
 cohort--based approach to predicting length of stay for ICU patients.
-Doborjeh et al. (2022) localized spiking neural networks in order to
-assess stroke risk from up to 7-day time series of clinical and
-environmental factors. Liu et al. (2022) incorporated global-to-local
-transfer learning into localized models of acute kidney injury risk in
-hospitalized patients. Also of note, from a public health perspective,
-Xu et al. (2008) used CBR to predict rehabilitation time as well as
-disability risk for unemployed workers experiencing chronic pain.
-
-Toward detection and diagnosis, Wyns et al. (2004) proposed a hybrid
+\citet{Doborjeh2022} localized spiking \hl{ANN}s in order to assess
+stroke risk from up to 7-day time series of clinical and environmental
+factors. \citet{Liu2022} incorporated global-to-local transfer learning
+into \hl{SCBM}s of acute kidney injury risk in hospitalized patients.
+Also of note, from a public health perspective, \citet{Xu2008} used CBR
+to predict rehabilitation time as well as disability risk for unemployed
+workers experiencing chronic pain.
+
+Toward detection and diagnosis, \citet{Wyns2004} proposed a hybrid
 neural net--CBR system to distinguish (with confidence bounds) arthritic
-versus control patients, based on several histological features. Nicolas
-et al. (2014) used collaborative multilabel CBR to subtype melanoma
-patients based on confocal and dermoscopy images. N. Wang et al. (2019)
-used localized models built from a multi-type additive similarity
-measure to distinguish type 2 diabetic versus control populations. Along
-the way, Song and Kasabov (2006), Kasabov and Hu (2010), and Verma et
-al. (2015) took an iterative model-building approach to several tasks:
-predicting glomerular filtration rate, a key indicator of renal
+versus control patients, based on several histological features.
+\citet{Nicolas2014} used collaborative multilabel CBR to subtype
+melanoma patients based on confocal and dermoscopy images.
+\citet{Wang2019} used \hl{SCBM}s built from a multi-type additive
+similarity measure to distinguish type 2 diabetic versus control
+populations. Along the way, \citet{Song2006}, \citet{Kasabov2010}, and
+\citet{Verma2015} took an iterative model-building approach to several
+tasks: predicting glomerular filtration rate, a key indicator of renal
 function, from demographic and physiological variables; identifying
@@
 treatment decisions into their approach: By taking ``decision points''
-as their units of analysis, Tang et al. (2021) and Ng et al. (2021)
-built not classifiers or predictors but comparative effectiveness models
-into a localized framework, providing for the first time in our sample
+as their units of analysis, \citet{Tang2021} and \citet{Ng2021} built
+not classifiers or predictors but comparative effectiveness models into
+a localized framework, providing for the first time in our sample
 explicitly prescriptive rather than descriptive clinical decision
@@
 decision support tool, in this case for early diagnosis of melanoma from
-clinical data and dermoscopy images (Nicolas et al. 2014). While the
+clinical data and dermoscopy images \citep{Nicolas2014}. While the
 stated objectives were analogous, specific emphasis was placed on the
@@
 
-The mosaic plot in Figure \ref{fig:properties} summarizes the relation
-between data source, clinical task, and aim.
-
 \begin{figure}
@@
 
-\caption{Share of studies characterized by several design elements: data source (row), clinical task (fill), and remit (opacity).}\label{fig:properties}
+\caption{\hl{Share of studies characterized by several design elements:}
+ data source (row),
+ clinical task (fill), and
+ remit (opacity).}\label{fig:properties}
 \end{figure}
 
-\hypertarget{rationales}{%
-\subsection{Rationales}\label{rationales}}
+\subsection{\texorpdfstring{Rationales\hl{ for methods}}{Rationales}}\label{rationales}
 
 The included studies hypothesized, asserted, or assumed several benefits
-of localized modeling specific to clinical and medical settings and
-advantages over other modeling approaches. The most common was that the
-restriction to similar or relevant past cases would improve predictive
-performance for the index case (Mariuzzi et al. 1997; Liang, Hu, and
-Kasabov 2015; Ng et al. 2015; Lee 2017). In particular, Lee, Maslove,
-and Dubin (2015) hypothesized and confirmed that the value of each past
-case would be positively related to its similarity to the index case, an
-assumption built in to the weighting schemes of other approaches. Lowsky
-et al. (2013) made a different case, that the fewer parametric
-assumptions and complications of a CBR-style model would allow for
-greater accuracy. Additionally, several investigators asserted that the
-use of localized cohorts befit the clinical focus on the individual
-patient rather than the population, without reference to performance
-(Song and Kasabov 2006; Xu et al. 2008).
-
-In an interesting contrast, Tang et al. (2021) and Ng et al. (2021)
-argued that their localized approach using the larger and more
-heterogeneous populations covered by EHR-derived data could better
-capture the messy and diverse lessons of everyday practice, as a
-counterpart to guidelines based on randomized controlled trials. Their
-emphasis on the enabling role of EHRs to power methods well-suited to
-large, structured data repositories was shared by several others
-(Campillo-Gimenez et al. 2013; Nicolas et al. 2014; Verma et al. 2015;
-Lee 2017). Verma et al. (2015) and Zhang et al. (2018) additionally
-pointed out that localized models are adaptable to noise in the data,
-variation in patterns of missingness, and (harkening to the last step of
-the R4 cycle) addition of new cases with known outcomes to the corpus.
-These properties, they said, tend to be more difficult for
-whole-population models to handle.
-
-The other frequent advantage attributed to localized modeling was
-interpretability. Elter, Schulz-Wendtland, and Wittenberg (2007) and
-Nicolas et al. (2014) emphasized the value of the detailed past cases,
-available to the user, on which predictions are based. This
-``self-explanation capability'' made outputs more intelligible to
-physicians in the CDS setting. Ng et al. (2015) additionally pointed out
-that localized GLMs yield localized effect estimates, which may help
-investigators identify individually relevant risk factors. Liu et al.
-(2022) took this idea further, systematically comparing regression
-coefficients for specific risk factors across dozens of diagnostic
-subpopulations. Doborjeh et al. (2022) used localized time series models
-to identify environmental changes associated with increased stroke risk.
+of \hl{SCBM} specific to clinical and medical settings and advantages
+over other modeling approaches. The most common was that the restriction
+to similar or relevant past cases would improve predictive performance
+for the index case \citep{Mariuzzi1997, Liang2015, Ng2015, Lee2017}. In
+particular, \citet{Lee2015} hypothesized and confirmed that the value of
+each past case would be positively related to its similarity to the
+index case, an assumption built in to the weighting schemes of other
+approaches. \citet{Lowsky2013} made a different case, that the fewer
+parametric assumptions and complications of a CBR-style model would
+allow for greater accuracy. Additionally, several investigators asserted
+that the use of localized cohorts befit the clinical focus on the
+individual patient rather than the population, without reference to
+performance \citep{Song2006, Xu2008}.
+
+In an interesting contrast, \citet{Tang2021} and \citet{Ng2021} argued
+that their localized approach using the larger and more heterogeneous
+populations covered by EHR-derived data could better capture the messy
+and diverse lessons of everyday practice, as a counterpart to guidelines
+based on randomized controlled trials. Their emphasis on the enabling
+role of EHRs to power methods well-suited to large, structured data
+repositories was shared by several others
+\citep{CampilloGimenez2013, Nicolas2014, Verma2015, Lee2017}.
+\citet{Verma2015} and \citet{Zhang2018} additionally pointed out that
+\hl{SCBM}s are adaptable to noise in the data, variation in patterns of
+missingness, and (harkening to the last step of the R4 cycle) addition
+of new cases with known outcomes to the corpus. These properties, they
+said, tend to be more difficult for whole-population models to handle.
+
+The other frequent advantage attributed to \hl{SCBM} was
+interpretability. \citet{Elter2007} and \citet{Nicolas2014} emphasized
+the value of the detailed past cases, available to the user, on which
+predictions are based. This ``self-explanation capability'' made outputs
+more intelligible to physicians in the \hl{CDSS} setting. \citet{Ng2015}
+additionally pointed out that localized GLMs yield localized effect
+estimates, which may help investigators identify individually relevant
+risk factors. \citet{Liu2022} took this idea further, systematically
+comparing regression coefficients for specific risk factors across
+dozens of diagnostic subpopulations. \citet{Doborjeh2022} used localized
+time series models to identify environmental changes associated with
+increased stroke risk.
 
@@
 of then-conventional CBR, most of which defended the use of other tools
-to improve performance via cohort selection (Campillo-Gimenez et al.
-2013; Nicolas et al. 2014; Vilhena et al. 2016) or to make models and
-outputs more interpretable (López et al. 2011; N. Wang et al. 2019).
-Liang, Hu, and Kasabov (2015) argued for simultaneous optimization of
-cohort and feature selection with model parameterization; their TWNFI
-approach combines localization with regularization.
-
-\hypertarget{challenges}{%
-\subsection{Challenges}\label{challenges}}
+to improve performance via cohort selection
+\citep{CampilloGimenez2013, Nicolas2014, Vilhena2016} or to make models
+and outputs more interpretable \citep{Lopez2011, Wang2019}.
+\citet{Liang2015} argued for simultaneous optimization of cohort and
+feature selection with model parameterization; their TWNFI approach
+combines localization with regularization.
+
+\subsection{\texorpdfstring{Challenges\hl{ with methods}}{Challenges}}\label{challenges}
 
 CBR can be understood as the opposite side of a trade-off with
-rule-based reasoning between model size and model complexity: Doyle,
-Cunningham, and Walsh (2006) describe their approach as
-``knowledge-light'', in that ``the cases do not contain explicit
-explanation structures; instead, explanation is achieved by comparison
-of the query case with retrieved cases''. This means that the greatest
-performance and efficiency challenges in CBR have to do with the
-retrieval and revision phases in the R4 cycle. Several studies addressed
-these challenges: Park, Kim, and Chun (2006), while excluded from the
-synthesis, were the earliest to propose that cohorts be bounded by a
-similarity threshold rather than by the number of cases, which improved
-performance in their experiments. Campillo-Gimenez et al. (2013)
-leveraged predictor weights obtained from logistic regression to inform
-the similarity calculation used in retrieval. Lowsky et al. (2013) used
-non-parametric models to reduce the computational burden of revision
-(prediction). Ma et al. (2020) proposed to improve efficiency along the
-entire R4 cycle, but in particular by only executing task-dependent
-steps in real time (``just-in-time learning'', JITL), and Ng et al.
-(2021) and Tang et al. (2021) partitioned multiple phases in the R4
-cycle into offline and online components, only the latter of which would
-be performed in real time as new data are received. Liu et al. (2022)
-used transfer learning from globally-fitted models to improve the
-efficiency of localized models.
-
-Other studies addressed limitations of available tools. López et al.
-(2011) implemented a comprehensive CBR tool in response to the lack of
-general-purpose software, to enable coupling with other tools as well as
-to expedite development, experimentation, and uptake. Several other
-teams also set out to develop more generalizable and data-agnostic
-clinical support tools (Elter, Schulz-Wendtland, and Wittenberg 2007;
-Liang, Hu, and Kasabov 2015; Ng et al. 2015; Zhang et al. 2018).
-However, more informatical and legal challenges to implementation,
-including interoperability of systems and data and regulations
-concerning security and privacy, were rarely addressed. We discuss this
-further in Section \ref{feasibility}.
+rule-based reasoning between model size and model complexity:
+\citet{Doyle2006} describe their approach as ``knowledge-light'', in
+that ``the cases do not contain explicit explanation structures;
+instead, explanation is achieved by comparison of the query case with
+retrieved cases''. This means that the greatest performance and
+efficiency challenges in CBR have to do with the retrieval and revision
+phases in the R4 cycle. Several studies addressed these challenges:
+\citet{Park2006}, while excluded from the synthesis, were the earliest
+to propose that cohorts be bounded by a similarity threshold rather than
+by the number of cases, which improved performance in their experiments.
+\citet{CampilloGimenez2013} leveraged predictor weights obtained from
+logistic regression to inform the similarity calculation used in
+retrieval. \citet{Lowsky2013} used non-parametric models to reduce the
+computational burden of revision (prediction). \citet{Ma2020} proposed
+to improve efficiency along the entire R4 cycle, but in particular by
+only executing task-dependent steps in real time (``just-in-time
+learning'', JITL), and \citet{Ng2021} and \citet{Tang2021} partitioned
+multiple phases in the R4 cycle into offline and online components, only
+the latter of which would be performed in real time as new data are
+received. \citet{Liu2022} used transfer learning from globally-fitted
+models to improve the efficiency of \hl{SCBM}s.
+
+Other studies addressed limitations of available tools.
+\citet{Lopez2011} implemented a comprehensive CBR tool in response to
+the lack of general-purpose software, to enable coupling with other
+tools as well as to expedite development, experimentation, and uptake.
+Several other teams also set out to develop more generalizable and
+data-agnostic clinical support tools
+\citep{Elter2007, Liang2015, Ng2015, Zhang2018}. However, more
+informatical and legal challenges to implementation, including
+interoperability of systems and data and regulations concerning security
+and privacy, were rarely addressed. We discuss this further in Section
+\ref{future-work}.
 
@@
 
-\hypertarget{evaluations}{%
-\subsection{Evaluations}\label{evaluations}}
+\subsection{Evaluations}\label{evaluations}
 
@@
 We do observe one clear pattern: All of the proposed methods that most
-evidently outperformed their comparators---Kohonen + CBR (Wyns et al.
-2004), TWNFI Kasabov and Hu (2010), gravitational search algorithms
-(GSA) (Liang, Hu, and Kasabov 2015), CBR + rules (+ DML) (Nicolas et al.
-2014), Gaussian process regression (Zhang et al. 2018), JITL-ELM (Ma et
-al. 2020)---are hybrids of localized modeling (in some cases CBR) with
-other techniques, often DML. Though Campillo-Gimenez et al. (2013) and
-Ng et al. (2015) report non-superior performance by such hybrids, the
+evidently outperformed their comparators---Kohonen + CBR
+\citep{Wyns2004}, TWNFI \citep{Song2006, Kasabov2010}, gravitational
+search algorithms (GSA) \citep{Liang2015}, CBR + rules (+ DML)
+\citep{Nicolas2014}, Gaussian process regression \citep{Zhang2018},
+JITL-ELM \citep{Ma2020}---are hybrids of \hl{SCBM} (in some cases CBR)
+with other techniques, often DML. Though \citet{CampilloGimenez2013} and
+\citet{Ng2015} report non-superior performance by such hybrids, the
 pattern suggests the importance of the similarity measure to the
 retrieval step. Meanwhile, when proposed approaches targeted cohort
-demarcation or choice of predictive model---statistical CBR (Park, Kim,
-and Chun 2006); individualized logistic regression, decision tree, and
-random forest (Lee, Maslove, and Dubin 2015; Lee 2017)---they did not
-consistently outperform comparators.
-
-\hypertarget{identified-needs}{%
-\subsection{Identified needs}\label{identified-needs}}
+demarcation or choice of predictive model---statistical CBR
+\citep{Park2006}; individualized \hl{GLM, DT, and RF}
+\citep{Lee2015, Lee2017}---they did not consistently outperform
+comparators.
+
+\subsection{Identified needs}\label{identified-needs}
 
 The study authors focused their recommendations and their own plans for
-future work mostly on technical improvements and evaluations (Mariuzzi
-et al. 1997; Yearwood and Wilkinson 1997). Urged improvements included
-full or partial automation of predictor selection (Mariuzzi et al. 1997;
-Yearwood and Wilkinson 1997), similarity learning (Mariuzzi et al. 1997;
-N. Wang et al. 2019), and parameter optimization (Song and Kasabov 2006;
-Lee 2017); extensions to new data structures (López et al. 2011), data
-types (Liang, Hu, and Kasabov 2015; Verma et al. 2015; Malykh and
-Rudetskiy 2018), and reasoning systems (Nicolas et al. 2014); and the
-use of more advanced model components to improve accuracy or efficiency
-(Lowsky et al. 2013; Lee, Maslove, and Dubin 2015; Liang, Hu, and
-Kasabov 2015; Zhang et al. 2018). Authors also urged validations and
-independent evaluations using larger or more comprehensive data sets
-(Elter, Schulz-Wendtland, and Wittenberg 2007; Xu et al. 2008; Verma et
-al. 2015; Ng et al. 2015), using data aggregated from multiple health
-systems (Lee, Maslove, and Dubin 2015; Lee 2017; Tang et al. 2021; Ng et
-al. 2021), and in other care settings or disease contexts (Song and
-Kasabov 2006; Zhang et al. 2018; Tang et al. 2021; Ng et al. 2021).
+future work mostly on technical improvements and evaluations
+\citep{Mariuzzi1997, Yearwood1997}. Urged improvements included full or
+partial automation of predictor selection
+\citep{Mariuzzi1997, Yearwood1997}, similarity learning
+\citep{Mariuzzi1997, Wang2019}, and parameter optimization
+\citep{Song2006, Lee2017}; extensions to new data structures
+\citep{Lopez2011}, data types \citep{Liang2015, Verma2015, Malykh2018},
+and reasoning systems \citep{Nicolas2014}; and the use of more advanced
+model components to improve accuracy or efficiency
+\citep{Lowsky2013, Lee2015, Liang2015, Zhang2018}. Authors also urged
+validations and independent evaluations using larger or more
+comprehensive data sets \citep{Elter2007, Xu2008, Verma2015, Ng2015},
+using data aggregated from multiple health systems
+\citep{Lee2015, Lee2017, Tang2021, Ng2021}, and in other care settings
+or disease contexts \citep{Song2006, Zhang2018, Tang2021, Ng2021}.
 
@@
 and the users. Some authors urged the incorporation of (human-derived)
-domain knowledge into the data or models (Yearwood and Wilkinson 1997;
-N. Wang et al. 2019). Others suggested new uses of their modeling
-approaches: to measure feature importance (Wyns et al. 2004), to make
-models more expressive (Lee, Maslove, and Dubin 2015), and to combine
-information obtained both from global and from localized models (Ng et
-al. 2015). Most of this incremental work was indeed carried out in later
-investigations in our sample. That said, we did not find reports of
-successful implementations of these tools in clinical practice. (Several
-excluded studies did showcase uses of CBR in practice.)
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+domain knowledge into the data or models \citep{Yearwood1997, Wang2019}.
+Others suggested new uses of their modeling approaches: to measure
+feature importance \citep{Wyns2004}, to make models more expressive
+\citep{Lee2015}, and to combine information obtained both from global
+and from \hl{SCBM}s \citep{Ng2015}. Most of this incremental work was
+indeed carried out in later investigations in our sample. That said, we
+did not find reports of successful implementations of these tools in
+clinical practice. (Several excluded studies did showcase uses of CBR in
+practice.)
+
+\section{\texorpdfstring{\hl{Discussion}}{}}\label{section}
 
@@
 
-\hypertarget{process}{%
-\subsection{Process}\label{process}}
+\subsection{Process}\label{process}
 
@@
 
-\hypertarget{study-designs}{%
-\subsection{Study designs}\label{study-designs}}
+\subsection{Study designs}\label{study-designs}
 
@@
 
-Because we excluded studies that used conventional NN prediction, many
-included studies reported new approaches to adaptation subsequent to
-retrieval. Very few proposed novel similarity measures, possibly because
-larger studies tended to be reported in separate articles detailing
-experiments with specific components of the process. However, this also
-suggests that few experimental studies have focused on the unified
-development of new retrieval and adaptation strategies. We also note
-that the majority of studies evaluated and compared methods using
-leave-one-out cross-validation (LOOCV). This is an appropriate technique
-when available data are scarce, and indeed most data sets used by
-included studies numbered in the hundreds of cases or fewer. This likely
-follows from the older age of many included studies and from their
-consistent primary focus on clinical data, which is more costly to
-collect and comes with more restrictions on its use.
-
-\hypertarget{coherence}{%
-\subsection{Coherence}\label{coherence}}
+\subsection{Coherence}\label{coherence}
 
@@
 smaller contiguous programs, including applications of the TWNFI
-methodology (Song and Kasabov 2006; Verma et al. 2015), individualized
-mortality prediction for ICU patients (Lee, Maslove, and Dubin 2015; Lee
-2017) and the use of more explanatory models to prioritize predictors or
-treatments for chronic disease (Ng et al. 2015; Tang et al. 2021; Ng et
-al. 2021).
+methodology \citep{Song2006, Verma2015}, individualized mortality
+prediction for ICU patients \citep{Lee2015, Lee2017} and the use of more
+explanatory models to prioritize predictors or treatments for chronic
+disease \citep{Ng2015, Tang2021, Ng2021}.
 
 These programs used varying terminology for common concepts, and no
-common term is in use for what we here refer to as localized modeling;
-authors described their approaches as ``targeted prognosis'' (Mariuzzi
-et al. 1997), ``transductive inference'' (Song and Kasabov 2006),
-``personalized decision support'' (Lee, Maslove, and Dubin 2015),
-``personalized (predictive) models'' (Liang, Hu, and Kasabov 2015, in
-contrast to ``local models''; Ng et al. 2015; N. Wang et al. 2019; Ma et
-al. 2020; Liu et al. 2022; Doborjeh et al. 2022), and ``precision
-cohort'' analysis (N. Wang et al. 2019; Tang et al. 2021; Ng et al.
-2021). We find uses of ``targeted'', ``personalized'', and ``precision''
-generic and imprecise, while transductive inference is an established
-term for a broader set of methods in ML. Because \emph{local} naturally
-contrasts with \emph{global}, we propose ``localized models'' as a
-suitable term for this counterpart to globally-fitted models.\footnote{We
-  note, in response to one reviewer's comment, that for many if not most
-  models used in ML, excepting GLMs, a predictor may play a different
-  and more or less important role for some cases than others, as
-  quantified by importance measures or model-agnostic explanations
-  (Biecek and Burzykowski 2021; Molnar 2023). In this way, such models
-  can also be viewed as ``localized''. We recognize that the use of
-  global and local explanations is standard. As another reviewer pointed
-  out, the terminology of local and global is also used in the unrelated
-  context of federated ML to distinguish models trained using data
-  stored in a single storage device (local) from their aggregations
-  (global) (Moshawrab et al. 2023; Brauneck et al. 2023). The prospects
-  for federated architecture to more effectively implement similarity
-  cohort--based models are intriguing, and would require some
-  reconciliation of terms. At present, we feel that these methods are
-  sufficiently disjoint to avoid confusion in practice, but we suggest
-  the more specific term ``similarity cohort--based models'' when the
-  term ``local'' is overloaded.}
-
-\hypertarget{synthesis-1}{%
-\subsection{Synthesis}\label{synthesis-1}}
+common term is in use for what we here refer to as
+\hl{similarity cohort--based modeling}; authors described their
+approaches as ``targeted prognosis'' \citep{Mariuzzi1997},
+``transductive inference'' \citep{Song2006}, ``personalized decision
+support'' \citep{Lee2015}, ``personalized (predictive) models''
+\citetext{\citealp[in contrast to ``local
+models'']{Liang2015}; \citealp{Ng2015}; \citealp{Wang2019}; \citealp{Ma2020}; \citealp{Liu2022}; \citealp{Doborjeh2022}},
+and ``precision cohort'' analysis \citep{Wang2019, Tang2021, Ng2021}. We
+find uses of ``targeted'', ``personalized'', and ``precision'' generic
+and imprecise, while transductive inference is an established term for a
+broader set of methods in ML.
+\hl{While ``local'' naturally contrasts with ``global'', the overloaded term ``localized models'' stands to cause more confusion than it resolves, so we use ``similarity cohort--based models'' to refer to this specific approach.}
+
+\subsection{Synthesis}\label{synthesis-1}
 
@@
 was used in 8 studies that tie in to the much larger literature on DML
-(Yang 2006). The other, \emph{cohort thresholding}, focused on choosing
-or optimizing the manner in which cohorts were demarcated using the
-similarity measure.
-
-\hypertarget{similarity-learning-and-threshold-optimization}{%
+\citep{Yang2006}. The other, \emph{cohort thresholding}, focused on
+choosing or optimizing the manner in which cohorts were demarcated using
+the similarity measure.
+
 \subsubsection{Similarity learning and threshold
-optimization}\label{similarity-learning-and-threshold-optimization}}
-
-The similarity learning approaches took a variety of forms. These
-techniques are designed to detect structure in data, especially
-associations between predictors and responses, and to use this structure
-to inform the definition of a patient similarity measure. Eight studies
-used response values in training data sets to supervise similarity
-learning while two used unsupervised learning (Table
-\ref{tab:composite}). Based on the key properties of DML algorithms
-identified by Bellet, Habrard, and Sebban (2014), most learned measures
-were non-linear while some were locally learned, and some were optimized
+optimization}\label{similarity-learning-and-threshold-optimization}
+
+\hl{Eight} studies used response values in training data sets to
+supervise similarity learning while two used unsupervised learning
+(Table \ref{tab:composite}). Based on the key properties of DML
+algorithms identified by \citet{Bellet2014}, most learned measures were
+non-linear while some were locally learned, and some were optimized
 globally while others locally. Not all qualified as DML, since the
@@
 inequality. Each used its similarity measure to retrieve training cases
-relevant to each testing case. Once retrieved, most took a standard NN
-approach to generating predictions; two exceptions (Vilhena et al. 2016;
-Tang et al. 2021) fit predictive models to the retrieved cohorts.
+relevant to each testing case. \hl{Most} took a standard NN approach to
+generating predictions; \hl{two} \citep{Vilhena2016, Tang2021} fit
+predictive models to the retrieved cohorts.
 
 Several studies used unsupervised similarity learning, for example the
-Mahalanobis distance (Lowsky et al. 2013) and the Kolmogorov
-entropy-based distance (Elter, Schulz-Wendtland, and Wittenberg 2007).
-In particular, Yearwood and Wilkinson (1997) defined similarity as a
-weighted sum of differences in predictor values and used linear
-regression to optimize the weights for the predictive accuracy of the
-cohort they retrieve. Others used supervised learning: Song and Kasabov
-(2006) proposed an iterative algorithm to optimize a set of fuzzy
-inference rules used to retrieve relevant cases and generate a
+Mahalanobis distance \citep{Lowsky2013} and the Kolmogorov entropy-based
+distance \citep{Elter2007}. In particular, \citet{Yearwood1997} defined
+similarity as a weighted sum of differences in predictor values and used
+linear regression to optimize the weights for the predictive accuracy of
+the cohort they retrieve. Others used supervised learning:
+\citet{Song2006} proposed an iterative algorithm to optimize a set of
+fuzzy inference rules used to retrieve relevant cases and generate a
 prediction, using back-propagation on the rules' parameters. Their
@@
 are not discussed here because they did not originate the technique.
-Nicolas et al. (2014) explicitly appeal to a supervised distance metric
-learning (DML) technique (Xing et al. 2002) to obtain a similarity
-measure on their set of dermoscopy and confocal images that most
-effectively separates malignant from benign melanoma tumors. We will
-discuss the remaining uses of supervised similarity learning shortly.
+\citet{Nicolas2014} explicitly appeal to a supervised distance metric
+learning (DML) technique \citep{Xing2002} to obtain a similarity measure
+on their set of dermoscopy and confocal images that most effectively
+separates malignant from benign melanoma tumors. We will discuss the
+remaining uses of supervised similarity learning shortly.
 
 Distinct from but related to similarity learning was supervised cohort
-construction. This was an adaptive step taken by Park, Kim, and Chun
-(2006) to optimize the predictive performance of models fitted to
-cohorts retrieved using a fixed similarity rather than count threshold,
-a counterpart to conventional CBR they called ``statistical
-CBR''.\footnote{Drawing from Goyal, Lifshits, and Schütze (2008), we
-  suggest the term ``combinatorial CBR'' for the then-conventional
-  approach using similarity cohorts of fixed cardinality.} The step used
-a heuristic procedure to locate a similarity threshold that (locally)
-maximizes predictive accuracy on the training set. This is analogous to
-optimizing the neighborhood size parameter in NN predictive modeling, so
-we will not discuss it further except to mention that the authors found
+construction. This was an adaptive step taken by \citet{Park2006} to
+optimize the predictive performance of models fitted to cohorts
+retrieved using a fixed similarity rather than count threshold, a
+counterpart to conventional CBR they called ``statistical
+CBR''.\footnote{Drawing from \citet{Goyal2008}, we suggest the term
+  ``combinatorial CBR'' for the then-conventional approach using
+  similarity cohorts of fixed cardinality.} The step used a heuristic
+procedure to locate a similarity threshold that (locally) maximizes
+predictive accuracy on the training set. This is analogous to optimizing
+the neighborhood size parameter in NN predictive modeling, so we will
+not discuss it further except to mention that the authors found
 statistical CBR to outperform conventional CBR on several data sets.
-Campillo-Gimenez et al. (2013) combined these learning techniques: They
+\citet{CampilloGimenez2013} combined these learning techniques: They
 used logistic regression on the training set to obtain weights for the
@@
 
-As noted above, these instances of similarity learning fit into a much
-larger literature on DML, which has seen widespread use in health
-informatics and modeling. That these studies satisfied our review
-criteria is due to the novelty of the techniques at the time of
-publication and the detailed attention paid to their techniques by the
-authors. It also reflects a limitation of our study and of the
-literature we set out to retrieve: We are aware of no standard terms in
-use to identify the family of techniques that fit statistical models to
-similarity-based cohorts. The retrieved studies whose use of such
-techniques alone satisfied our criteria variably termed them
-individualized, personalized, local, and patient-specific models, among
-other terms (Table \ref{tab:composite}). We prefer the term localized
-models, which is both sensitive and specific to uses of a similarity or
-distance measure to retrieve a cohort to which a model is fit: It
-includes some properly included techniques we did not have in mind
-(Vilhena et al. 2016) yet excludes other techniques commonly referred to
-as individualized, personalized, or precision.
-
-\hypertarget{localized-models}{%
-\subsubsection{Localized models}\label{localized-models}}
-
-In addition to Lowsky et al. (2013), Liang, Hu, and Kasabov (2015), Ng
-et al. (2015), Lee (2017), Vilhena et al. (2016), Tang et al. (2021),
-and Ng et al. (2021), seven other studies employed localized models:
-Mariuzzi et al. (1997), Lee, Maslove, and Dubin (2015), Verma et al.
-(2015), N. Wang et al. (2019), Ma et al. (2020), Liu et al. (2022), and
-Doborjeh et al. (2022). In most cases these models were purely
-predictive in application; of the exceptions, Mariuzzi et al. (1997)
-produced purely descriptive models of survival outcomes, which were
-evaluated for their precision rather than their accuracy, while Ng et
-al. (2015) used generalized regression models descriptively as well as
-predictively, as did Tang et al. (2021), Ng et al. (2021), and Liu et
-al. (2022) later. The most common approach to cohort selection was to
-optimize a size threshold via manual exploration (Mariuzzi et al. 1997;
-Lee, Maslove, and Dubin 2015; Ng et al. 2015; Lee 2017; N. Wang et al.
-2019; Vilhena et al. 2016) or via cross-validation (Lowsky et al. 2013;
-Verma et al. 2015). Further in the former direction, one study (Ma et
-al. 2020) set a fixed cohort size while two (Liang, Hu, and Kasabov
-2015; Tang et al. 2021; Ng et al. 2021) devised more sophisticated
+\subsubsection{Localized models}\label{localized-models}
+
+\hl{Seven} other studies employed localized models:
+\citet{Mariuzzi1997}, \citet{Lee2015}, \citet{Verma2015},
+\citet{Wang2019}, \citet{Ma2020}, \citet{Liu2022}, and
+\citet{Doborjeh2022}. \hl{Most} models were purely
+predictive\hl{, though} \citet{Mariuzzi1997} produced purely descriptive
+models of survival outcomes, which were evaluated for their precision
+rather than their accuracy, while \citet{Ng2015} used generalized
+regression models descriptively as well as predictively, as did
+\citet{Tang2021}, \citet{Ng2021}, and \citet{Liu2022}\hl{.} The most
+common approach to cohort selection was to optimize a size threshold via
+manual exploration
+\citep{Mariuzzi1997, Lee2015, Ng2015, Lee2017, Wang2019, Vilhena2016} or
+via cross-validation \citep{Lowsky2013, Verma2015}. \hl{One} study
+\citep{Ma2020} set a fixed cohort size while two
+\citep{Liang2015, Tang2021, Ng2021} devised more sophisticated
 algorithms to balance multiple desiderata including predictive accuracy.
 
-Several themes emerged from this sample. Foremost was the optimization
-of retrieved cohort sizes for some measure of model performance. Despite
-the early recommendation by Park, Kim, and Chun (2006) to threshold
-cohorts by similarity rather than by cardinality, only the one
-previously published study (Mariuzzi et al. 1997) and the most recent
-study (Tang et al. 2021; Ng et al. 2021) in this corpus took this
-approach, though Liu et al. (2022) took a third option by retrieving
-fixed proportions of training data. Viewed as a hyperparameter of the
-individualized modeling approach, cohort size can be treated in the same
-way as neighborhood size in NN prediction (viewed here as a special case
-of localized modeling wherein the model is a simple summary statistic),
+\hl{One theme} was the optimization of retrieved cohort sizes for some
+measure of model performance. Despite the early recommendation by
+\citet{Park2006} to threshold cohorts by similarity rather than by
+cardinality, only the one previously published study
+\citep{Mariuzzi1997} and the most recent study \citep{Tang2021, Ng2021}
+in this corpus took this approach, though \citet{Liu2022} took a third
+option by retrieving fixed proportions of training data. Viewed as a
+hyperparameter of the individualized modeling approach, cohort size can
+be treated in the same way as neighborhood size in NN \hl{prediction},
 so we will not discuss it further.
 
-An alternative approach was to optimize multiple hyperparameters
-together: Kasabov and Hu (2010) proposed an iterative algorithm to
-settle on an optimal number and set of predictors as well as number of
-training cases (comprising the retrieved cohort) that was later used by
-Liang, Hu, and Kasabov (2015). More recently, Tang et al. (2021) and Ng
-et al. (2021) proposed a three-step process for cohort selection that
-filtered by exact match for one subset of predictors, used
-domain-informed similarity measures to rank these, and optimized the
-similarity threshold for a trade-off between cohort size and a measure
-of bias called cohort balance. Notably, though their similarity measure
-was supervised, this optimization process was not.
-
-With the exception of the descriptive survival models of Mariuzzi et al.
-(1997), only one thread in this corpus concerned itself with the
-interpretability of localized models: Ng et al. (2015) fit logistic
+\citet{Kasabov2010} proposed an iterative algorithm to settle on an
+optimal number and set of predictors as well as number of training cases
+(comprising the retrieved cohort) that was later used by
+\citet{Liang2015}. More recently, \citet{Tang2021} and \citet{Ng2021}
+proposed a three-step process for cohort selection that filtered by
+exact match for one subset of predictors, used domain-informed
+similarity measures to rank these, and optimized the similarity
+threshold for a trade-off between cohort size and a measure of bias
+called cohort balance. \hl{T}hough their similarity measure was
+supervised, this optimization process was not.
+
+With the exception of the descriptive survival models of
+\citet{Mariuzzi1997}, only one thread in this corpus \hl{concerned the}
+interpretability of localized models: \citet{Ng2015} fit logistic
 regression models to similarity-based cohorts and examined not only
@@
 localized comparative effectiveness--style models to recommend courses
-of treatment at decision points during a monitored patent's stay (Tang
-et al. 2021; Ng et al. 2021), and it was later used by Liu et al. (2022)
-to show a dependency of predictor importance on the patient disease
-subgroup. These studies suggest much wider potential for localized
+of treatment at decision points during a monitored patent's stay
+\citep{Tang2021, Ng2021}, and it was later used by \citet{Liu2022} to
+show a dependency of predictor importance on the patient disease
+subgroup. These studies suggest \hl{wider} potential for localized
 real-world evidence generation, which has long been promoted as a
-promise of advances in clinical and health informatics (Longhurst,
-Harrington, and Shah 2014).
-
-\hypertarget{a-general-framework}{%
-\subsubsection{A general framework}\label{a-general-framework}}
-
-This relatively small sample exhibited a wide range of approaches, both
-technically and conceptually. Discussion of the technical diversity of
-these approaches has been summarized above and published in greater
-detail in previous reviews (Choudhury and Begum 2016; Sharafoddini,
-Dubin, and Lee 2017; Parimbelli et al. 2018), and we focus here on the
-conceptual. The design of a localized modeling approach can be
-decomposed into three largely independent choices: (1) how to measure
-the similarity between cases, (2) how to retrieve a cohort of cases
-similar to an index case, and (3) how to generate a prediction or other
-statistical insight for the index case from the similarity cohort. The
-choices are the \emph{similarity measure}, the \emph{cohort retrieval},
-and the \emph{statistical model}, respectively. These are diagrammed in
-Figure \ref{fig:framework}.
+promise of advances in clinical and health informatics
+\citep{Longhurst2014}.
+
+\subsubsection{A general framework}\label{a-general-framework}
+
+This relatively small sample exhibited a wide range of
+\hl{designs in both the assembly of pre-processing, retrieval, reuse, and interpretation steps and the specific tools deployed at each}.
+Discussion of the \hl{diversity} of \hl{tools} has been summarized above
+and published in greater detail in previous reviews
+\citep{Choudhury2016, Sharafoddini2017, Parimbelli2018}\hl{. In this section, we focus on the diversity of assemblies. These depend on}
+three largely independent choices
+\hl{(Figure }\ref{fig:framework}\hl{})\}: (1) how to measure the
+similarity between cases, (2) how to retrieve a cohort of cases similar
+to an index case, and (3) how to generate a prediction or other
+statistical insight for the index case from the similarity cohort.
 
@@
 
-\caption{General framework for localized models. Dashed arrows indicate optional steps.}\label{fig:framework}
+\caption{General framework for localized models.
+ Dashed arrows indicate optional steps.}\label{fig:framework}
 \end{figure}
@@
 Each step can be unsupervised supervised: Unsupervised similarity
-measures include discrete measures like the Levenshtein distance and
-continuous measures like cosine similarity, while supervised measures
-include the Mahalanobis distance and random forest proximity as well as
-several composite measures with weights calculated from the data. Most
-retrieval steps obtained cohorts with a uniform size (cardinality) or
-similarity threshold, and the remaining were likewise only informed by
-predictors; thus, while a retrieval step could in principle be
-supervised, in this sample none were. Almost all models were supervised,
-being that most performed predictive tasks, though Mariuzzi et al.
-(1997) modeled only survival rates within localized cohorts. Meanwhile,
-each step can be optimized in a ML fashion by having its parameters
-tuned to improve performance. We found one study that tuned the
-calculation of similarity (Liu et al. 2022), though most uniform cohort
-sizes were also tuned. Some earlier studies tuned generalized regression
-models, but more recent studies did not.
+measures include \hl{the} Levenshtein distance and \hl{cosine}
+similarity, while supervised measures include the Mahalanobis distance
+and \hl{RF} proximity as well as several composite measures with weights
+calculated from the data. Most retrieval steps obtained cohorts with a
+uniform size (cardinality) or similarity threshold, and the remaining
+were likewise only informed by predictors; thus, while a retrieval step
+could in principle be supervised, in this sample none were. Almost all
+models were supervised, being that most performed predictive tasks,
+though \citet{Mariuzzi1997} modeled only survival rates within localized
+cohorts. Meanwhile, each step can be optimized in a ML fashion by having
+its parameters tuned to improve performance. We found one study that
+tuned the calculation of similarity \citep{Liu2022}, though most uniform
+cohort sizes were also tuned. Some earlier studies tuned generalized
+regression models, but more recent studies did not.
 
@@
 \caption{\label{tab:framework}Specializations of general framework to
-included studies. Flags: R = recurse, S = supervised, T = tuned
+included studies. Flags: R = recurse\hl{d}, S = supervised, T = tuned
 (optimized).}\tabularnewline
@@
 \endlastfoot
-Mariuzzi et al. (1997) & Levenshtein distance\hspace{18em} &
+\citet{Mariuzzi1997} & Levenshtein distance\hspace{18em} &
 Statistical\hspace{18em} & Survival (S)\hspace{18em} \\
-Yearwood and Wilkinson (1997) & Weighted sum (S)\hspace{18em} &
+\citet{Yearwood1997} & Weighted sum (S)\hspace{18em} &
 Combinatorial\hspace{18em} & Proportion (S)\hspace{18em} \\
-Wyns et al. (2004) & Kohonen mapping\hspace{18em} &
-Adaptive\hspace{18em} & Representation (S)\hspace{18em} \\
-Park, Kim, and Chun (2006) & Euclidean distance\hspace{18em} &
-Statistical (T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
-Elter, Schulz-Wendtland, and Wittenberg (2007) & Entropic
-distance\hspace{18em} & Combinatorial\hspace{18em} & Proportion
-(S)\hspace{18em} \\
-Xu et al. (2008) & Euclidean distance\hspace{18em} &
+\citet{Wyns2004} & Kohonen mapping\hspace{18em} & Adaptive\hspace{18em}
+& Representation (S)\hspace{18em} \\
+\citet{Park2006} & Euclidean distance\hspace{18em} & Statistical
+(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
+\citet{Elter2007} & Entropic distance\hspace{18em} &
 Combinatorial\hspace{18em} & Proportion (S)\hspace{18em} \\
-Song and Kasabov (2006); Kasabov and Hu (2010); Liang, Hu, and Kasabov
-(2015); Verma et al. (2015) & Feature selection (R,S)\hspace{18em} &
-Combinatorial (R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
-López et al. (2011) & User-defined\hspace{18em} &
+\citet{Xu2008} & Euclidean distance\hspace{18em} &
+Combinatorial\hspace{18em} & Proportion (S)\hspace{18em} \\
+\citet{Song2006}; \citet{Kasabov2010}; \citet{Liang2015};
+\citet{Verma2015} & Feature selection (R,S)\hspace{18em} & Combinatorial
+(R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
+\citet{Lopez2011} & User-defined\hspace{18em} &
 User-defined\hspace{18em} & User-defined\hspace{18em} \\
-Lowsky et al. (2013) & Mahalanobis distance (S)\hspace{18em} &
+\citet{Lowsky2013} & Mahalanobis distance (S)\hspace{18em} &
 Combinatorial (T)\hspace{18em} & Proportional hazards
 (S)\hspace{18em} \\
-Campillo-Gimenez et al. (2013) & Weighted overlap\hspace{18em} &
+\citet{CampilloGimenez2013} & Weighted overlap\hspace{18em} &
 Combinatorial (T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
-Nicolas et al. (2014) & Distance metric learning (S)\hspace{18em} &
+\citet{Nicolas2014} & Distance metric learning (S)\hspace{18em} &
 Combinatorial\hspace{18em} & Association rules (S,T)\hspace{18em} \\
-Ng et al. (2015) & Locally supervised metric learning (S); Feature
+\citet{Ng2015} & Locally supervised metric learning (S); Feature
 selection\hspace{18em} & Combinatorial\hspace{18em} & Logistic model
 (S)\hspace{18em} \\
-Lee, Maslove, and Dubin (2015) & Cosine similarity\hspace{18em} &
-Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree
-(S)\hspace{18em} \\
-Vilhena et al. (2016) & Range overlap\hspace{18em} & Clustering
+\citet{Lee2015} & Cosine similarity\hspace{18em} & Combinatorial
+(T)\hspace{18em} & Logistic model (S); Decision tree (S)\hspace{18em} \\
+\citet{Vilhena2016} & Range overlap\hspace{18em} & Clustering
 (S)\hspace{18em} & Artificial neural network (S)\hspace{18em} \\
-Lee (2017) & Cosine similarity; Random forest proximity (S)\hspace{18em}
-& Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree
-(S); Random forest (S)\hspace{18em} \\
-Malykh and Rudetskiy (2018) & Unclear\hspace{18em} &
-Unclear\hspace{18em} & Unclear\hspace{18em} \\
-Zhang et al. (2018) & Gaussian process model\hspace{18em} &
+\citet{Lee2017} & Cosine similarity; Random forest proximity
+(S)\hspace{18em} & Combinatorial (T)\hspace{18em} & Logistic model (S);
+Decision tree (S); Random forest (S)\hspace{18em} \\
+\citet{Malykh2018} & Unclear\hspace{18em} & Unclear\hspace{18em} &
+Unclear\hspace{18em} \\
+\citet{Zhang2018} & Gaussian process model\hspace{18em} &
 Complete\hspace{18em} & Weighted support vector machine (S); Weighted
 Gaussian process regression (S)\hspace{18em} \\
-N. Wang et al. (2019) & Weighted sum\hspace{18em} & Combinatorial
+\citet{Wang2019} & Weighted sum\hspace{18em} & Combinatorial
 (T)\hspace{18em} & Logistic model (S); Random forest (S); Nearest
 neighbors (S)\hspace{18em} \\
-Ma et al. (2020) & Weighted sum\hspace{18em} &
-Combinatorial\hspace{18em} & Extreme learning machine
-(S)\hspace{18em} \\
-Y. Wang et al. (2020) & Weighted sum\hspace{18em} &
-Statistical\hspace{18em} & Linear model\hspace{18em} \\
-Tang et al. (2021); Ng et al. (2021) & Locally supervised metric
-learning (S)\hspace{18em} & Causal inference matching\hspace{18em} &
-Statistical test (S)\hspace{18em} \\
-Doborjeh et al. (2022) & Euclidean distance; Signal-to-noise ratio
+\citet{Ma2020} & Weighted sum\hspace{18em} & Combinatorial\hspace{18em}
+& Extreme learning machine (S)\hspace{18em} \\
+\citet{Wang2020} & Weighted sum\hspace{18em} & Statistical\hspace{18em}
+& Linear model\hspace{18em} \\
+\citet{Tang2021}; \citet{Ng2021} & Locally supervised metric learning
+(S)\hspace{18em} & Causal inference matching\hspace{18em} & Statistical
+test (S)\hspace{18em} \\
+\citet{Doborjeh2022} & Euclidean distance; Signal-to-noise ratio
 (S)\hspace{18em} & Statistical (T)\hspace{18em} & Spiking neural network
 (S)\hspace{18em} \\
-Liu et al. (2022) & Weighted Manhattan distance (R,T)\hspace{18em} &
+\citet{Liu2022} & Weighted Manhattan distance (R,T)\hspace{18em} &
 Proportional (T)\hspace{18em} & Logistic model (S)\hspace{18em} \\
@@
 
-\hypertarget{directions-and-expectations-for-future-work}{%
-\subsection{Directions and expectations for future
-work}\label{directions-and-expectations-for-future-work}}
-
-\hypertarget{performance}{%
-\subsubsection{Performance}\label{performance}}
-
-As discussed in Section \ref{rationales}, most studies were premised on
-the potential predictive value of localized models. As cautioned in
-Section \ref{evaluations}, not all experiments affirmed this premise.
-Most of those that did involved using metric learning to improve
-retrieval and sometimes iterative optimization of the metric and of the
-localized model. Some studies showcased results using a variety of such
-specializations (Campillo-Gimenez et al. 2013; Ng et al. 2015; Zhang et
-al. 2018; Liu et al. 2022). However, because these advanced
-implementations were mostly compared against their precursors or
-commonplace alternatives, we cannot speak to their performance or any
-trade-offs with respect to each other. This would be a worthy goal of
-future work.
-
-The value of this work would be enhanced by standardization and
-modularization. First, studies using localized models should locate the
-proposed approach and any variations in a shared specification space.
-For this purpose, we propose the framework outlined above, at least as a
-starting point. When a specification invokes complex techniques, for
-example supervised DML or neural network classifiers, it should be
-compared at least to alternative specifications using simpler
-alternatives. Campillo-Gimenez et al. (2013) and Liu et al. (2022)
-provide excellent examples in which all possible choices are toggled and
-yield a neatly nested comparison. Additionally, some state-of-the-art
-models, including others that obtain individualized predictions, should
-be included in comparisons. Finally, implementations should be not only
-published on public repositories but designed in such a way that users
-familiar with the underlying language can substitute a technique of
-their choosing for any part of the specification (similarity, retrieval,
-fitting, evaluation). We believe these principles will make reported
-results clearer and reproduction and extension easier.
-
-\hypertarget{interpretability}{%
-\subsubsection{Interpretability}\label{interpretability}}
-
-While most studies identified outstanding technical needs (Section
-\ref{identified-needs}), few emphasized the need to assess human-focused
-qualities like user interface and user experience, interpretability,
-meaningfulness, or trust. Most studies identified interpretability as an
-advantage of localized models, though none evaluated model
-interpretability and few proposed new interpretative uses. Lack of trust
-in ML tools is a long-recognized problem that direct intepretability of
-model components could help alleviate, but this is more often assumed
-than demonstrated.
-
-One valuable direction for future work would be to measure the utility
-of interpretable components in research and in clinical practice and the
-correctness and confidence of users in their interpretations. This is a
-necessary precondition for practical use but would also be a valuable
-contribution to the experimental literature. Another would be to compare
-the localized predictor importance measures obtained from localized
-models to the local importance measures used to explain predictions made
-by opaque models. This would test both measures for concurrent validity,
-while differences between them would inform what settings or cases are
-poorly served by one or the other.
-
-\hypertarget{feasibility}{%
-\subsubsection{Feasibility}\label{feasibility}}
-
-Much of the reviewed work was motivated by the need for modeling
-paradigms that perform well on a variety of tasks and in a variety of
-settings. This need demands not only methodologies but also
-architectures that are robust, versatile, and compliant in the face of
-diverse data models and use restrictions, yet no studies in our sample
-addressed the challenges of system interoperability and regulatory
-requirements head-on.
-
-CBR and, by extension, localized modeling are especially susceptible to
-these challenges, as corpora of past cases must be aggregated from
-multiple institutions and from multiple systems within institutions, and
-models built on them must then be applicable to out-of-box data
-structures as well. Individual systems exhibit many dimensions of
-incompleteness (Weiskopf et al. 2013) and vary along other dimensions of
-quality (Kohane et al. 2021), and their aggregation for modeling
-purposes depends on several kinds of interoperability (Weber 2015).
-Successful deployment also depends on satisfying regulatory regimes
-designed to protect the privacy of patients, the security of
-communications, and trust between parties (Haendel et al. 2021). While
-most studies in our sample focused on improving care, increasingly many
-over time focused on generalizable knowledge. Therefore, as studies
-using localized modeling shift from proof of concept to feasibility,
-they will also need to demonstrate practical and regulatory feasibility.
-
-\hypertarget{customizability}{%
-\subsubsection{Customizability}\label{customizability}}
-
-The primary goal of these studies was to establish that localized
-models, and certain strategies within this paradigm, perform at or above
-the level of other predictive models or strategies. From a functionalist
-approach to reproducibility (Matarese 2022), then, later studies were
-broadly successful at reproducing earlier studies, and later studies
-provide sufficient detail to support ongoing reproduction
-efforts---though these might be limited by the lack of open-source code
-or public implementations. However, the immediate goal of most studes
-was to improve care, and an implied need, often enjoined but neither
-obtained nor reproduced, was a demonstration of practical use.
-
-We posit that an essential component of practical usefulness is the
-ability of the user community to exert some control over the models.
-Given the impact on performance of the choice or optimization of the
-similarity measure, an important target for user input would be the
-importance of certain variables in the calculation of similarity, with
-an understanding of how it can impact not only the performance of the
-predictions but also the cohort retrieved for the model. For one
-example, a user may want to minimize the weight of rare diseases in
-medical history in order to retrieve a population with more such cases
-in order to better measure their associated risk to the outcome. In
-contrast, they may want to increase the weight of the indicating
-diagnosis in order to allow fewer patients from similar but distinct
-populations to influence the model. For another example, a user may want
-to down-weight socioeconomic variables like race--ethnicity in order to
-ensure a more diverse modeling cohort. López et al. (2011) took an
-important step in this direction with an adjustable and adaptable
-implementation. Future purely quantitative work could assess whether
-similarity-tuning can achieve these ends more flexibly than strict
-inclusion/exclusion criteria.
-
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
-
-We propose the term \emph{localized modeling} to encompass an approach
-derived from CBR in which parameterized models are fitted in a
-standardized way to nearest neighborhoods of past or training cases
-according to a measure of patient similarity. We conducted a systematic
-search for studies that apply localized models to tasks involving health
-data and synthesized these largely independently developed approaches
-into a general framework. While the search was limited by low
-inter-rater reliability and failure to recover several motivating
-examples, the included studies used many of the same underlying tools to
-build, optimize, and evaluate their methods. We therefore believe that
-our framework can serve to taxonomize ongoing work of this type and
-inform the development of customizable implementations. Indeed, the
-availability of increasing computational power, the diversity of tasks
-to which these models were applied and of technical specifications they
-employed, and the apparent lack of any multi-group research program to
-date suggest great potential for growth. Whereas precious few of the
+\subsection{Future work}\label{future-work}
+
+\hl{We suggest four avenues for future work: evaluation of predictive performance, usability and interpretability, practical and regulatory feasibility, and customizability.}
+
+\hl{Most studies were premised on the potential predictive value of localized models (Section }\ref{rationales}\hl{), but not all experiments affirmed this premise (Section }\ref{evaluations}\hl{).
+Those that did mostly used DML to improve retrieval and sometimes iteratively optimized the metric and the model.
+However, even those that showcased several such specializations }\citep{CampilloGimenez2013, Ng2015, Zhang2018, Liu2022}\hl{ mostly used their own precursors or commonplace alternatives as comparators.
+Future work should compare competitive alternatives to assess their performance and other trade-offs.
+This work would be facilitated by standardization and modularization using a framework like that outlined above.
+For example, }\citet{CampilloGimenez2013}\hl{ and }\citet{Liu2022}\hl{ use a modular design to toggle multiple choices and obtain neatly nested comparisons.}
+
+\hl{While most studies identified outstanding technical needs (Section }\ref{identified-needs}\hl{), few (though some) emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
+Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
+Lack of trust in ML tools is a long-recognized problem that direct intepretability of model components could help alleviate.
+One valuable direction for future work would be to measure the utility of interpretable components in research and in clinical practice and the correctness and confidence of users in their interpretations.
+Another would be to compare the localized predictor importance measures obtained from localized models to the local importance measures used to explain predictions made by opaque models.}
+
+\hl{Much of the reviewed work was motivated by the need for modeling paradigms that perform well on a variety of tasks and in a variety of settings.
+Such software architectures must be robust, versatile, and compliant in the face of diverse data models and use restrictions.
+SCBM is susceptible to these challenges, as corpora of past cases must be aggregated from multiple institutions and from multiple systems within institutions, and models built on them must then be applicable to out-of-box data structures.
+Individual systems exhibit many dimensions of incompleteness }\citep{Weiskopf2013}\hl{ and vary along other dimensions of quality }\citep{Kohane2021}\hl{, and their aggregation for modeling purposes depends on several kinds of interoperability }\citep{Weber2015}\hl{.
+Successful deployment also depends on satisfying regulatory regimes designed to protect the privacy of patients, the security of communications, and trust between parties }\citep{Haendel2021}\hl{.
+Future work may therefore eventually need to directly address the challenges of system interoperability and regulatory requirements.}
+
+\hl{From a functionalist approach to reproducibility }\citep{Matarese2022}\hl{, later studies were broadly successful at reproducing earlier studies, and later studies provide sufficient detail to support ongoing reproduction efforts---though these might be limited by the lack of open-source code or public implementations.
+However, the immediate goal of most studes was to improve care.
+We expect that this practical utility will depend on the ability of the user community to exert control over certain inputs, components, and behaviors of the models.
+For one example, a user may want to minimize the weight of rare diseases in medical history in order to retrieve a population with more such cases in order to better measure their associated risk to the outcome.
+In contrast, they may want to increase the weight of the indicating diagnosis in order to allow fewer patients from similar but distinct populations to influence the model.
+Alternatively, a user may want to down-weight socioeconomic variables like race--ethnicity in order to ensure a more diverse modeling cohort.
+}\citet{Lopez2011}\hl{ took an important step in this direction with an adjustable and adaptable implementation.}
+
+\subsection{\texorpdfstring{\hl{Conclusion}}{}}\label{section-1}
+
+We conducted a systematic search for studies that
+\hl{systematically fit parameterized models to similarity-based cohorts pursuant}
+to tasks involving \hl{medical and }health data\hl{. }While the search
+\hl{strategy showed serious limitations}, the included studies used many
+of the same underlying tools to build, optimize, and evaluate their
+methods.
+\hl{We synthesized these largely independently developed approaches into a general framework that we believe}
+can serve to taxonomize ongoing \hl{work }and inform
+\hl{future development}. Indeed, the availability of increasing
+computational power\hl{ and} the diversity of tasks to which these
+models were applied and of technical specifications they
+\hl{employed suggest }potential for growth. \hl{Whereas} few of the
 reviewed studies used these models for any task other than prediction,
@@
 and growing interest in interpretable alternatives, we recommend that
-future work put greater emphasis on the development and validation of
-interpretable localized models and on their reception by communities of
-medical research and clinical practice.
+future work \hl{emphasize the} validation of interpretable localized
+models \hl{and their accessibility to} communities of medical research
+and clinical practice.
+
+\section*{List of abbreviations}\label{list-of-abbreviations}
+\addcontentsline{toc}{section}{List of abbreviations}
+
+\begin{itemize}
+\tightlist
+\item
+  CBR: case-based reasoning
+\item
+  CIS: clinical information system
+\item
+  \hl{CDSS:} clinical decision support\hl{ system}
+\item
+  AI: artificial intelligence
+\item
+  NN: nearest neighbors
+\item
+  GLM: generalized linear model
+\item
+  \hl{DT: decision tree}
+\item
+  \hl{RF: random forest}
+\item
+  \hl{ANN: artificial neural network}
+\item
+  ML: machine learning
+\item
+  PMC: PubMed Central
+\item
+  LOOCV: leave-one-out cross-validation
+\item
+  GSA: gravitational search algorithm
+\item
+  DML: distance metric learning
+\end{itemize}
+
+\section*{Declarations}\label{declarations}
+\addcontentsline{toc}{section}{Declarations}
+
+\subsection*{Ethics approval and consent to
+participate}\label{ethics-approval-and-consent-to-participate}
+\addcontentsline{toc}{subsection}{Ethics approval and consent to
+participate}
+
+Not applicable.
+
+\subsection*{Consent for publication}\label{consent-for-publication}
+\addcontentsline{toc}{subsection}{Consent for publication}
+
+Not applicable.
+
+\subsection*{Availability of data and
+materials}\label{availability-of-data-and-materials}
+\addcontentsline{toc}{subsection}{Availability of data and materials}
+
+Search results, included studies, and other cited sources are publicly
+available in a Zotero group library:
+\url{https://www.zotero.org/groups/5017571/imsr/}
+
+Data collected or encoded for included studies are publicly available in
+two Google Sheets:
+
+\begin{itemize}
+\tightlist
+\item
+  Bibliographic and methodological properties of included studies:
+  \url{https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/}
+\item
+  Terminology and composite techniques of included studies:
+  \url{https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/}
+\end{itemize}
+
+Code used to analyze these data and prepare this manuscript is publicly
+available in a GitHub repository:
+\url{https://github.com/corybrunson/imsr}
+
+\subsection*{Competing interests}\label{competing-interests}
+\addcontentsline{toc}{subsection}{Competing interests}
+
+The authors declare that they have no competing interests.
+
+\subsection*{Funding}\label{funding}
+\addcontentsline{toc}{subsection}{Funding}
+
+Not applicable.
+
+\subsection*{Authors' contributions}\label{authors-contributions}
+\addcontentsline{toc}{subsection}{Authors' contributions}
+
+JCB conceived of the study. All authors contributed to the study
+design---search, screen, review, and synthesis. AC and JCB conducted the
+search, screen, review, and synthesis (coded and analyzed bibliographic
+and methodological data). AC and JCB drafted, and all authors
+contributed to, read, and approved, the manuscript.
+
+\subsection*{Acknowledgements}\label{acknowledgements}
+\addcontentsline{toc}{subsection}{Acknowledgements}
+
+Two anonymous reviewers provided valuable feedback on a previous version
+of this manuscript.
 
@@
 
-\hypertarget{appendix}{%
-\section*{Appendix}\label{appendix}}
+\section*{Appendix}\label{appendix}
 \addcontentsline{toc}{section}{Appendix}
 
-\hypertarget{research-question}{%
-\subsection*{Research question}\label{research-question}}
+\subsection*{Research question}\label{research-question}
 \addcontentsline{toc}{subsection}{Research question}
@@
 
-\hypertarget{purpose-of-review}{%
-\subsection*{Purpose of review}\label{purpose-of-review}}
+\subsection*{Purpose of review}\label{purpose-of-review}
 \addcontentsline{toc}{subsection}{Purpose of review}
@@
 
-\hypertarget{search-design}{%
-\subsection*{Search design}\label{search-design}}
+\subsection*{Search design}\label{search-design}
 \addcontentsline{toc}{subsection}{Search design}
@@
 
+\subsubsection{\texorpdfstring{\hl{Databases}}{}}\label{section-2}
+
+\hl{Through each platform, we searched those databases included by default.
+This means that we searched both MEDLINE and PubMed Central (PMC) through Pubmed (we did not exclude other databases, but our earliest included results post-date them) and that we searched the six indices of the Core Collection (the Science Citation Index Expanded, the Social Sciences Citation Index, the Arts \& Humanities Citation Index, the Emerging Sources Citation Index, the Conference Proceedings Citation Index, and the Book Citation Index) as well as several regional databases through the Web of Science platform.}
+
+\hl{The structure and terms of our search strings were inspired in part by previous reviews adjacent to our topic of interest }\citep{Sharafoddini2017, Parimbelli2018}\hl{.
+We organized the PubMed search in disjunctive normal form (an OR of ANDs).
+Following the solidification of an outline, we expanded each term to include synonyms that are similar enough to be applicable to our search.
+We then evaluated the expanded search string using the PubMed Advanced Search platform.
+We initially included each term and their synonyms separately to evaluate what resulted.
+We pruned several terms that returned no results (``phrases not found'') or to reduce the number of results.
+At the conclusions of this process for each individual term, we combined the search terms using disjunctive normal form.}
+
 Following discussion among AC, PMJ, and JCB, we discarded results from
@@
 
-\hypertarget{search-strings}{%
-\subsubsection*{Search strings}\label{search-strings}}
+\subsubsection*{Search strings}\label{search-strings}
 \addcontentsline{toc}{subsubsection}{Search strings}
@@
 
-\hypertarget{screening-process}{%
-\subsection*{Screening process}\label{screening-process}}
+\subsection*{Screening process}\label{screening-process}
 \addcontentsline{toc}{subsection}{Screening process}
@@
 
-\hypertarget{selection-process}{%
-\subsection*{Selection process}\label{selection-process}}
+\subsection*{Selection process}\label{selection-process}
 \addcontentsline{toc}{subsection}{Selection process}
@@
 
+\hl{During full text review, we decided to expand the conception of statistical models (fourth criterion): Rather than restricting to models that are fit and evaluated in separate steps, we chose to allow simple statistical summaries such as mean survival }\citep{Mariuzzi1997}\hl{ and survival curves }\citep{Lowsky2013}\hl{.
+These approaches were novel to CBR and presaged later developments, so they were helpful to understanding the development of SCBM.
+However, this then admitted studies that applied conventional nearest neighbors prediction: Each retrieved cohort consisted of the $k$ most similar cases to the index case, and the response for the index case was predicted to be either the mean (continuous response) or the plurality (discrete response) of the cohort's responses.
+A review of all studies that use nearest neighbors prediction would be impractical.
+Because our focus is on novel approaches that combine similarity-based retrieval and statistical modeling of retrieved cohorts, we chose to exclude those studies whose approach was equivalent to nearest neighbors prediction using a conventional similarity measure.}
+
 We decided after concluding full-text review to perform one round of
@@
 
-\hypertarget{update}{%
-\subsection*{Update}\label{update}}
+\subsection*{Update}\label{update}
 \addcontentsline{toc}{subsection}{Update}
@@
 
-\hypertarget{analysis-and-synthesis}{%
-\subsection*{Analysis and synthesis}\label{analysis-and-synthesis}}
+\subsection*{Terminology}\label{terminology}
+\addcontentsline{toc}{subsection}{Terminology}
+
+\hl{As one reviewer pointed out for many if not most models used in ML a predictor may play a different and more or less important role for some cases than others, as quantified by importance measures or model-agnostic explanations }\citep{Biecek2021, Molnar2023}\hl{.
+In this way, such models can also be viewed as ``localized''.
+We recognize, for example, that the use of global and local explanations is standard.
+As another reviewer pointed out, the terminology of local and global is also used in the unrelated context of federated ML to distinguish models trained using data stored in a single storage device (local) from their aggregations (global) }\citep{Moshawrab2023, Brauneck2023}\hl{.
+The prospects for federated architecture to more effectively implement similarity cohort--based models are intriguing, and would require some reconciliation of terms.
+A third reviewer pointed out that ``local'' may describe models fitted to data collected at a specific site or using features extracted from a specific patch of an image.
+Taken together, these issues convinced us to drop the shorthand ``localized model'' and adopt the more specific term ``similarity cohort--based model''.}
+
+\subsection*{Analysis and synthesis}\label{analysis-and-synthesis}
 \addcontentsline{toc}{subsection}{Analysis and synthesis}
@@
 
-\hypertarget{methodological-elements}{%
-\subsection*{Methodological elements}\label{methodological-elements}}
+\subsection*{Methodological elements}\label{methodological-elements}
 \addcontentsline{toc}{subsection}{Methodological elements}
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1082}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4433}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4485}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1045}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4627}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4328}}@{}}
 \caption{\label{tab:composite}Methodological elements of studies
@@
 \endlastfoot
-Yearwood and Wilkinson (1997) & supervised similarity learning & CBR;
+\citet{Yearwood1997} & supervised similarity learning & CBR;
 case-structured retrieval \\
-Mariuzzi et al. (1997) & predictive modeling on similarity cohorts &
-CBR; targeted subsampling \\
-Wyns et al. (2004) & unsupervised similarity learning & Combined Kohonen
+\citet{Mariuzzi1997} & predictive modeling on similarity cohorts & CBR;
+targeted subsampling \\
+\citet{Wyns2004} & unsupervised similarity learning & Combined Kohonen
 type neural network-case-based reasoning \\
-Park, Kim, and Chun (2006) & supervised cohort construction & CBR,
-statistical case-based reasoning \\
-Song and Kasabov (2006) & supervised similarity learning & Transductive
+\citet{Park2006} & supervised cohort construction & CBR, statistical
+case-based reasoning \\
+\citet{Song2006} & supervised similarity learning & Transductive
 inference \\
-Elter, Schulz-Wendtland, and Wittenberg (2007) & unsupervised similarity
-learning & CBR; entropic distance measure \\
-Xu et al. (2008) & tiered constraint; similarity matching & CBR \\
-López et al. (2011) & interactive implementation & CBR; eXiT*CBR \\
-Kasabov and Hu (2010) & predictive modeling on similarity cohorts &
+\citet{Elter2007} & unsupervised similarity learning & CBR; entropic
+distance measure \\
+\citet{Xu2008} & tiered constraint; similarity matching & CBR \\
+\citet{Lopez2011} & interactive implementation & CBR; eXiT*CBR \\
+\citet{Kasabov2010} & predictive modeling on similarity cohorts &
 Personalized model \\
-Verma et al. (2015) & predictive modeling on similarity cohorts &
+\citet{Verma2015} & predictive modeling on similarity cohorts &
 Personalized modeling; TWNFI \\
-Liang, Hu, and Kasabov (2015) & predictive modeling on similarity
-cohorts & Local modeling; personalized modeling; TWNFI \\
-Lowsky et al. (2013) & predictive modeling on similarity cohorts &
+\citet{Liang2015} & predictive modeling on similarity cohorts & Local
+modeling; personalized modeling; TWNFI \\
+\citet{Lowsky2013} & predictive modeling on similarity cohorts &
 K-nearest neighbors survival \\
-Campillo-Gimenez et al. (2013) & supervised similarity learning & CBR \\
-Nicolas et al. (2014) & supervised similarity learning & CBR \\
-Ng et al. (2015) & predictive modeling on similarity cohorts;
-explanatory modeling on similarity cohorts & Personalized predictive
-model \\
-Lee, Maslove, and Dubin (2015) & predictive modeling on similarity
-cohorts & Personalized prediction \\
-Vilhena et al. (2016) & supervised similarity learning; predictive
+\citet{CampilloGimenez2013} & supervised similarity learning & CBR \\
+\citet{Nicolas2014} & supervised similarity learning & CBR \\
+\citet{Ng2015} & predictive modeling on similarity cohorts; explanatory
+modeling on similarity cohorts & Personalized predictive model \\
+\citet{Lee2015} & predictive modeling on similarity cohorts &
+Personalized prediction \\
+\citet{Vilhena2016} & supervised similarity learning; predictive
 modeling on similarity cohorts & CBR \\
-Lee (2017) & predictive modeling on similarity cohorts &
+\citet{Lee2017} & predictive modeling on similarity cohorts &
 Patient-specific predictive model \\
-Zhang et al. (2018) & whole-population weight learning & Gaussian
+\citet{Zhang2018} & whole-population weight learning & Gaussian
 processes \\
-Malykh and Rudetskiy (2018) & unclear & CBR \\
-Ma et al. (2020) & predictive modeling on similarity cohorts &
+\citet{Malykh2018} & unclear & CBR \\
+\citet{Ma2020} & predictive modeling on similarity cohorts &
 Personalized model \\
-N. Wang et al. (2019) & predictive modeling on similarity cohorts &
+\citet{Wang2019} & predictive modeling on similarity cohorts &
 Personalized predictive modeling \\
-Y. Wang et al. (2020) & descriptive modeling on similarity cohorts &
-CBR \\
-Liu et al. (2022) & NA & Personalized model with transfer learning \\
-Doborjeh et al. (2022) & NA & Personalized spiking neural network \\
-Tang et al. (2021); Ng et al. (2021) & supervised similarity learning;
+\citet{Wang2020} & descriptive modeling on similarity cohorts & CBR \\
+\citet{Liu2022} & predictive modeling on similarity cohorts; transfer
+learning; supervised similarity learning & Personalized model with
+transfer learning \\
+\citet{Doborjeh2022} & predictive modeling on similarity cohorts &
+Personalized spiking neural network \\
+\citet{Tang2021}; \citet{Ng2021} & supervised similarity learning;
 predictive modeling on similarity cohorts & Similarity model,
@@
 
-\hypertarget{performance-evaluations-and-comparisons}{%
 \subsection*{Performance evaluations and
-comparisons}\label{performance-evaluations-and-comparisons}}
+comparisons}\label{performance-evaluations-and-comparisons}
 \addcontentsline{toc}{subsection}{Performance evaluations and
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1088}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1088}} |
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0725}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1658}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1658}} |
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2435}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}} |
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2124}}
@@
 \endlastfoot
-Wyns et al. (2004) & Accuracy & Arthritis & Kohonen + CBR &
+\citet{Wyns2004} & Accuracy & Arthritis & Kohonen + CBR &
 75.60\%\hspace{6em} & Kohonen mapping & 54.70\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & backpropagation & 54.70\%\hspace{6em} \\
-\midrule\noalign{}
-Park, Kim, and Chun (2006) & Accuracy & Dermatology & Statistical CBR &
+\citet{Park2006} & Accuracy & Dermatology & Statistical CBR &
 96.57\%\hspace{6em} & CBR & 95.43\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & NN & 55.59\%\hspace{6em} \\
-\midrule\noalign{}
-Song and Kasabov (2006) & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
+\citet{Song2006} & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
 7.74\hspace{6em} \\
@@
 & & & & \hspace{6em} & TNFI & 7.28\hspace{6em} \\
-\midrule\noalign{}
-Elter, Schulz-Wendtland, and Wittenberg (2007) & AUROC & mammography
-mass region & CBR & 0.885\hspace{6em} & DT & 0.872\hspace{6em} \\
+\citet{Elter2007} & AUROC & mammography mass region & CBR &
+0.885\hspace{6em} & DT & 0.872\hspace{6em} \\
 & & & & \hspace{6em} & ANN & 0.882\hspace{6em} \\
-\midrule\noalign{}
-Xu et al. (2008) & Accuracy & Return-to-work & CBR & 62.5\%\hspace{6em}
-& LR & 71.2\%\hspace{6em} \\
-\midrule\noalign{}
-Kasabov and Hu (2010) & Accuracy & Colon cancer diagnosis & TWNFI &
+\citet{Xu2008} & Accuracy & Return-to-work & CBR & 62.5\%\hspace{6em} &
+LR & 71.2\%\hspace{6em} \\
+\citet{Kasabov2010} & Accuracy & Colon cancer diagnosis & TWNFI &
 91.9\%\hspace{6em} & MLR (Personalized) & 82.3\%\hspace{6em} \\
@@
 \hspace{6em} \\
-\midrule\noalign{}
-Liang, Hu, and Kasabov (2015) & Accuracy & Colon cancer & knnGSA &
+\citet{Liang2015} & Accuracy & Colon cancer & knnGSA &
 85.48\%\hspace{6em} & SVM & 82.14\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & WKNN & 95.50\%\hspace{6em} \\
-\midrule\noalign{}
-Lowsky et al. (2013) & Relative IPEC & Graft survival & MKNN &
+\citet{Lowsky2013} & Relative IPEC & Graft survival & MKNN &
 0.973\hspace{6em} & RSF & 0.957\hspace{6em} \\
-\midrule\noalign{}
-Campillo-Gimenez et al. (2013) & AUROC & Waitlist registration, complete
-& CBR + Wj & 89.8\%\hspace{6em} & LR & 92.0\%\hspace{6em} \\
+\citet{CampilloGimenez2013} & AUROC & Waitlist registration, complete &
+CBR + Wj & 89.8\%\hspace{6em} & LR & 92.0\%\hspace{6em} \\
 & & & CBR + Wk & 90.7\%\hspace{6em} & Standalone CBR &
@@
 & & & CBR + Wj + Wk & 88.5\%\hspace{6em} & & \hspace{6em} \\
-\midrule\noalign{}
-Nicolas et al. (2014) & Accuracy & Nevus & Collaborative + rules &
+\citet{Nicolas2014} & Accuracy & Nevus & Collaborative + rules &
 98\%\hspace{6em} & Dermoscopy CBR & 92\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & Collaborative & 95\%\hspace{6em} \\
-\midrule\noalign{}
-Ng et al. (2015) & AUROC & Diabetes onset & Personalized LR
+\citet{Ng2015} & AUROC & Diabetes onset & Personalized LR
 (LSML+FeatureFiltering) & 0.624\hspace{6em} & KNN & 0.617\hspace{6em} \\
@@
 & & & Personalized LR (Random) & 0.602\hspace{6em} & & \hspace{6em} \\
-\midrule\noalign{}
-Lee, Maslove, and Dubin (2015) & AUROC & 30-day mortality &
-Individualized LR & 0.830\hspace{6em} & Individualized DC &
-0.797\hspace{6em} \\
+\citet{Lee2015} & AUROC & 30-day mortality & Individualized LR &
+0.830\hspace{6em} & Individualized DC & 0.797\hspace{6em} \\
 & & & Individualized DT & 0.753\hspace{6em} & & \hspace{6em} \\
-\midrule\noalign{}
-Lee (2017) & AUROC & & Individualized LR & 0.824\hspace{6em} &
+\citet{Lee2017} & AUROC & & Individualized LR & 0.824\hspace{6em} &
 Individualized DC & 0.801\hspace{6em} \\
@@
 & & & Individualized CSRF & 0.832\hspace{6em} & & \hspace{6em} \\
-\midrule\noalign{}
-Malykh and Rudetskiy (2018) & Accuracy & J13/pneumonia & Case-based NN &
+\citet{Malykh2018} & Accuracy & J13/pneumonia & Case-based NN &
 81.6\%\hspace{6em} & & \hspace{6em} \\
@@
 \hspace{6em} \\
-\midrule\noalign{}
-Zhang et al. (2018) & AUROC & Cognitive impairment & Gaussian process
+\citet{Zhang2018} & AUROC & Cognitive impairment & Gaussian process
 regression with custom kernel & 0.92\hspace{6em} & linear regression
@@
 0.778\hspace{6em} \\
-\midrule\noalign{}
-N. Wang et al. (2019) & AUROC & Diabetes & Personalized RF &
-0.90\hspace{6em} & & \hspace{6em} \\
+\citet{Wang2019} & AUROC & Diabetes & Personalized RF & 0.90\hspace{6em}
+& & \hspace{6em} \\
 & & & Personalized KNN & 0.82\hspace{6em} & & \hspace{6em} \\
 & & & Personalized LR & 0.89\hspace{6em} & & \hspace{6em} \\
-\midrule\noalign{}
-Ma et al. (2020) & AUROC & Discharge at 10 days & One-class JITL-ELM &
+\citet{Ma2020} & AUROC & Discharge at 10 days & One-class JITL-ELM &
 0.8510\hspace{6em} & ELM & 0.4973\hspace{6em} \\
@@
 & & & & \hspace{6em} & One-class SVM & 0.4647\hspace{6em} \\
-\midrule\noalign{}
-Liu et al. (2022) & AUROC & Acute kidney injury & & \hspace{6em} &
-Global & 0.684\hspace{6em} \\
+\citet{Liu2022} & AUROC & Acute kidney injury & & \hspace{6em} & Global
+& 0.684\hspace{6em} \\
 & & & & \hspace{6em} & Risk-ranking & 0.708\hspace{6em} \\
@@
 & & & & \hspace{6em} & PMTL & 0.779\hspace{6em} \\
-\midrule\noalign{}
-Doborjeh et al. (2022) & Accuracy & Stroke & Personalized SNN &
+\citet{Doborjeh2022} & Accuracy & Stroke & Personalized SNN &
 60.70\%\hspace{6em} & & \hspace{6em} \\
@@
 
-\hypertarget{data-and-code}{%
-\subsection{Data and code}\label{data-and-code}}
-
-Search results are publicly available in a Zotero group library. Data
-collected or encoded for included studies are publicly available in two
-Google Sheets. Code used to analyze these data and generate this
-manuscript is publicly available in a GitHub repository.
-
-\begin{itemize}
-\tightlist
-\item
-  Search results, included studies, and other bibliography:
-  \url{https://www.zotero.org/groups/5017571/imsr/}
-\item
-  Bibliographic and methodological properties of included studies:
-  \url{https://docs.google.com/spreadsheets/d/1tpWMhYH2pyRT55K7n2J2XFs-kEV_JTuCDmXzJ4BBgNo/}
-\item
-  Terminology and composite techniques of included studies:
-  \url{https://docs.google.com/spreadsheets/d/1xvDJwiLBoI2oz8fxHJ5MjNmiju_RAlK7RJv-wXe1DAs/}
-\item
-  Code used to conduct analyses and prepare the manuscript:
-  \url{https://github.com/corybrunson/imsr}
-\end{itemize}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Aamodt1994}{}}%
-Aamodt, Agnar, and Enric Plaza. 1994. {``Case-{Based Reasoning}:
-{Foundational Issues}, {Methodological Variations}, and {System
-Approaches}.''} \emph{AI Communications} 7 (1): 39--59.
-\url{https://doi.org/10.3233/AIC-1994-7104}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Begum2011}{}}%
-Begum, Shahina, Mobyen Uddin Ahmed, Peter Funk, Ning Xiong, and Mia
-Folke. 2011. {``Case-{Based Reasoning Systems} in the {Health Sciences}:
-{A Survey} of {Recent Trends} and {Developments}.''} \emph{IEEE
-Transactions on Systems, Man, and Cybernetics, Part C (Applications and
-Reviews)} 41 (4): 421--34.
-\url{https://doi.org/10.1109/TSMCC.2010.2071862}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bellet2014}{}}%
-Bellet, Aurélien, Amaury Habrard, and Marc Sebban. 2014. {``A {Survey}
-on {Metric Learning} for {Feature Vectors} and {Structured Data}.''}
-arXiv. \url{https://doi.org/10.48550/arXiv.1306.6709}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
-Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
-{Model Analysis}: {Explore}, {Explain}, and {Examine Predictive
-Models}}. New York: {Chapman and Hall/CRC}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Brauneck2023}{}}%
-Brauneck, Alissa, Louisa Schmalhorst, Mohammad Mahdi Kazemi Majdabadi,
-Mohammad Bakhtiari, Uwe Völker, Jan Baumbach, Linda Baumbach, and
-Gabriele Buchholtz. 2023. {``Federated {Machine Learning},
-{Privacy-Enhancing Technologies}, and {Data Protection Laws} in {Medical
-Research}: {Scoping Review}.''} \emph{Journal of Medical Internet
-Research} 25 (1): e41588. \url{https://doi.org/10.2196/41588}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2013}{}}%
-Campillo-Gimenez, Boris, Wassim Jouini, Sahar Bayat, and Marc Cuggia.
-2013. {``Improving {Case-Based Reasoning Systems} by {Combining
-K-Nearest Neighbour Algorithm} with {Logistic Regression} in the
-{Prediction} of {Patients}' {Registration} on the {Renal Transplant
-Waiting List}.''} Edited by Randen Lee Patterson. \emph{PLOS ONE} 8 (9):
-e71991. \url{https://doi.org/10.1371/journal.pone.0071991}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Choudhury2016}{}}%
-Choudhury, Nabanita, and Shahin Ara Begum. 2016. {``A {Survey} on
-{Case-based Reasoning} in {Medicine}.''} \emph{International Journal of
-Advanced Computer Science and Applications} 7 (8).
-\url{https://doi.org/10.14569/IJACSA.2016.070820}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
-Dai, Leyu, He Zhu, and Dianbo Liu. 2020. {``Patient Similarity: Methods
-and Applications.''} \emph{arXiv:2012.01976 {[}Cs{]}}, December.
-\url{https://arxiv.org/abs/2012.01976}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Doborjeh2022}{}}%
-Doborjeh, Maryam, Zohreh Doborjeh, Alexander Merkin, Rita Krishnamurthi,
-Reza Enayatollahi, Valery Feigin, and Nikola Kasabov. 2022.
-{``Personalized {Spiking Neural Network Models} of {Clinical} and
-{Environmental Factors} to {Predict Stroke}.''} \emph{Cognitive
-Computation} 14 (6): 2187--2202.
-\url{https://doi.org/10.1007/s12559-021-09975-x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Doyle2006}{}}%
-Doyle, Dónal, Pádraig Cunningham, and Paul Walsh. 2006. {``An Evaluation
-of the Usefulness of Explanation in a Case-Based Reasoning System for
-Decision Support in Bronchiolitis Treatment.''} \emph{Computational
-Intelligence} 22 (3-4): 269--81.
-\url{https://doi.org/10.1111/j.1467-8640.2006.00288.x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Elter2007}{}}%
-Elter, M., R. Schulz-Wendtland, and T. Wittenberg. 2007. {``The
-Prediction of Breast Cancer Biopsy Outcomes Using Two {CAD} Approaches
-That Both Emphasize an Intelligible Decision Process.''} \emph{Medical
-Physics} 34 (11): 4164--72. \url{https://doi.org/10.1118/1.2786864}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fang2021}{}}%
-Fang, Hao Sen Andrew, Ngiap Chuan Tan, Wei Ying Tan, Ronald Wihal Oei,
-Mong Li Lee, and Wynne Hsu. 2021. {``Patient Similarity Analytics for
-Explainable Clinical Risk Prediction.''} \emph{BMC Medical Informatics
-\& Decision Making} 21 (1): 1--12.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
-Gierl, Lothar, Mathias Bull, and Rainer Schmidt. 1998. {``{CBR} in
-{Medicine}.''} In \emph{Case-{Based Reasoning Technology}}, edited by
-Mario Lenz, Hans-Dieter Burkhard, Brigitte Bartsch-Spörl, and Stefan
-Wess, 1400:273--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
-\url{https://doi.org/10.1007/3-540-69351-3_11}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Goyal2008}{}}%
-Goyal, Navin, Yury Lifshits, and Hinrich Schütze. 2008. {``Disorder
-Inequality: A Combinatorial Approach to Nearest Neighbor Search.''} In
-\emph{Proceedings of the 2008 {International Conference} on {Web Search}
-and {Data Mining}}, 25--32. {WSDM} '08. New York, NY, USA: Association
-for Computing Machinery. \url{https://doi.org/10.1145/1341531.1341538}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Haendel2021}{}}%
-Haendel, Melissa A, Christopher G Chute, Tellen D Bennett, David A
-Eichmann, Justin Guinney, Warren A Kibbe, Philip R O Payne, et al. 2021.
-{``The {National COVID Cohort Collaborative} ({N3C}): {Rationale},
-Design, Infrastructure, and Deployment.''} \emph{Journal of the American
-Medical Informatics Association} 28 (3): 427--43.
-\url{https://doi.org/10.1093/jamia/ocaa196}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kasabov2010}{}}%
-Kasabov, Nikola, and Yingjie Hu. 2010. {``Integrated Optimisation Method
-for Personalised Modelling and Case Studies for Medical Decision
-Support.''} \emph{International Journal of Functional Informatics and
-Personalised Medicine} 3 (3): 236--56.
-\url{https://doi.org/10.1504/IJFIPM.2010.039123}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kohane2021}{}}%
-Kohane, Isaac S., Bruce J. Aronow, Paul Avillach, Brett K.
-Beaulieu-Jones, Riccardo Bellazzi, Robert L. Bradford, Gabriel A. Brat,
-et al. 2021. {``What {Every Reader Should Know About Studies Using
-Electronic Health Record Data} but {May Be Afraid} to {Ask}.''}
-\emph{Journal of Medical Internet Research} 23 (3): e22219.
-\url{https://doi.org/10.2196/22219}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kolodner1992}{}}%
-Kolodner, Janet L. 1992. {``An Introduction to Case-Based Reasoning.''}
-\emph{Artificial Intelligence Review} 6 (1): 3--34.
-\url{https://doi.org/10.1007/BF00155578}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lee2017}{}}%
-Lee, Joon. 2017. {``Patient-{Specific Predictive Modeling Using Random
-Forests}: {An Observational Study} for the {Critically Ill}.''}
-\emph{JMIR Medical Informatics} 5 (1): e3.
-\url{https://doi.org/10.2196/medinform.6690}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lee2015}{}}%
-Lee, Joon, David M. Maslove, and Joel A. Dubin. 2015. {``Personalized
-{Mortality Prediction Driven} by {Electronic Medical Data} and a
-{Patient Similarity Metric}.''} Edited by Frank Emmert-Streib.
-\emph{PLOS ONE} 10 (5): e0127428.
-\url{https://doi.org/10.1371/journal.pone.0127428}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Liang2015}{}}%
-Liang, W, YJ Hu, and N Kasabov. 2015. {``Evolving Personalized Modeling
-System for Integrated Feature, Neighborhood and Parameter Optimization
-Utilizing Gravitational Search Algorithm.''} \emph{Evolving Systems} 6
-(1): 1--14. \url{https://doi.org/10.1007/s12530-013-9081-x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Liu2022}{}}%
-Liu, Kang, Xiangzhou Zhang, Weiqi Chen, Alan S. L. Yu, John A. Kellum,
-Michael E. Matheny, Steven Q. Simpson, Yong Hu, and Mei Liu. 2022.
-{``Development and {Validation} of a {Personalized Model With Transfer
-Learning} for {Acute Kidney Injury Risk Estimation Using Electronic
-Health Records}.''} \emph{JAMA Network Open} 5 (7): e2219776.
-\url{https://doi.org/10.1001/jamanetworkopen.2022.19776}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Longhurst2014}{}}%
-Longhurst, Christopher A., Robert A. Harrington, and Nigam H. Shah.
-2014. {``A {`{Green Button}'} {For Using Aggregate Patient Data At The
-Point Of Care}.''} \emph{Health Affairs} 33 (7): 1229--35.
-\url{https://doi.org/10.1377/hlthaff.2014.0099}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011}{}}%
-López, Beatriz, Carles Pous, Pablo Gay, Albert Pla, Judith Sanz, and
-Joan Brunet. 2011. {``{eXiT}*{CBR}: {A} Framework for Case-Based Medical
-Diagnosis Development and Experimentation.''} \emph{Artificial
-Intelligence in Medicine} 51 (2): 81--91.
-\url{https://doi.org/10.1016/j.artmed.2010.09.002}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Lowsky2013}{}}%
-Lowsky, D. J., Y. Ding, D. K. K. Lee, C. E. McCulloch, L. F. Ross, J. R.
-Thistlethwaite, and S. A. Zenios. 2013. {``A {K-nearest} Neighbors
-Survival Probability Prediction Method.''} \emph{Statistics in Medicine}
-32 (12): 2062--69. \url{https://doi.org/10.1002/sim.5673}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ma2020}{}}%
-Ma, X, YB Si, ZF Wang, and YQ Wang. 2020. {``Length of Stay Prediction
-for {ICU} Patients Using Individualized Single Classification
-Algorithm.''} \emph{Computer Methods and Programs in Biomedicine} 186
-(April). \url{https://doi.org/10.1016/j.cmpb.2019.105224}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018}{}}%
-Malykh, V. L., and S. V. Rudetskiy. 2018. {``Approaches to {Medical
-Decision-Making Based} on {Big Clinical Data}.''} \emph{Journal of
-Healthcare Engineering} 2018: 3917659.
-\url{https://doi.org/10.1155/2018/3917659}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
-Mariuzzi, G., A. Mombello, L. Mariuzzi, P. W. Hamilton, J. E. Weber, D.
-Thompson, and P. H. Bartels. 1997. {``Quantitative Study of Ductal
-Breast Cancer--Patient Targeted Prognosis: An Exploration of Case Base
-Reasoning.''} \emph{Pathology - Research and Practice} 193 (8): 535--42.
-\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Matarese2022}{}}%
-Matarese, Vera. 2022. {``Kinds of {Replicability}: {Different Terms} and
-{Different Functions}.''} \emph{Axiomathes} 32 (2): 647--70.
-\url{https://doi.org/10.1007/s10516-021-09610-2}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
-Molnar, Christoph. 2023. \emph{Interpretable {Machine Learning}: {A
-Guide} for {Making Black Box Models Explainable}}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Moshawrab2023}{}}%
-Moshawrab, Mohammad, Mehdi Adda, Abdenour Bouzouane, Hussein Ibrahim,
-and Ali Raad. 2023. {``Reviewing {Federated Machine Learning} and {Its
-Use} in {Diseases Prediction}.''} \emph{Sensors} 23 (4): 2112.
-\url{https://doi.org/10.3390/s23042112}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ng2021}{}}%
-Ng, Kenney, Uri Kartoun, Harry Stavropoulos, John A Zambrano, and Paul C
-Tang. 2021. {``Personalized Treatment Options for Chronic Diseases Using
-Precision Cohort Analytics.''} \emph{Scientific Reports} 11 (1): 1--13.
-\url{https://doi.org/10.1038/s41598-021-80967-5}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ng2015}{}}%
-Ng, Kenney, Jimeng Sun, Jianying Hu, and Fei Wang. 2015.
-{``\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4525240}{Personalized
-{Predictive Modeling} and {Risk Factor Identification} Using {Patient
-Similarity}}.''} \emph{AMIA Joint Summits on Translational Science
-Proceedings. AMIA Joint Summits on Translational Science} 2015: 132--36.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014}{}}%
-Nicolas, Ruben, Albert Fornells, Elisabet Golobardes, Guiomar Corral,
-Susana Puig, and Josep Malvehy. 2014. {``{DERMA}: A Melanoma Diagnosis
-Platform Based on Collaborative Multilabel Analog Reasoning.''}
-\emph{The Scientific World Journal} 2014: 351518.
-\url{https://doi.org/10.1155/2014/351518}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Parimbelli2018}{}}%
-Parimbelli, E., S. Marini, L. Sacchi, and R. Bellazzi. 2018. {``Patient
-Similarity for Precision Medicine: {A} Systematic Review.''}
-\emph{Journal of Biomedical Informatics} 83 (July): 87--96.
-\url{https://doi.org/10.1016/j.jbi.2018.06.001}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Park2006}{}}%
-Park, Yoon-Joo, Byung-Chun Kim, and Se-Hak Chun. 2006. {``New Knowledge
-Extraction Technique Using Probability for Case-Based Reasoning:
-Application to Medical Diagnosis.''} \emph{Expert Systems} 23 (1):
-2--20. \url{https://doi.org/10.1111/j.1468-0394.2006.00321.x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Sharafoddini2017}{}}%
-Sharafoddini, Anis, Joel A Dubin, and Joon Lee. 2017. {``Patient
-{Similarity} in {Prediction Models Based} on {Health Data}: {A Scoping
-Review}.''} \emph{JMIR Medical Informatics} 5 (1): e7.
-\url{https://doi.org/10.2196/medinform.6730}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Song2006}{}}%
-Song, Qun, and Nikola Kasabov. 2006. {``{TWNFI} --- a Transductive
-Neuro-Fuzzy Inference System with Weighted Data Normalization for
-Personalized Modeling.''} \emph{Neural Networks} 19 (10): 1591--96.
-\url{https://doi.org/10.1016/j.neunet.2006.05.028}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tang2021}{}}%
-Tang, Paul C, Sarah Miller, Harry Stavropoulos, Uri Kartoun, John
-Zambrano, and Kenney Ng. 2021. {``Precision Population Analytics:
-Population Management at the Point-of-Care.''} \emph{Journal of the
-American Medical Informatics Association} 28 (3): 588--95.
-\url{https://doi.org/10.1093/jamia/ocaa247}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Verma2015}{}}%
-Verma, A, M Fiasche, M Cuzzola, and G Irrera. 2015. {``Comparative Study
-of Existing Personalized Approaches for Identifying Important Gene
-Markers and for Risk Estimation in {Type2 Diabetes} in {Italian}
-Population.''} \emph{Evolving Systems} 6 (1): 15--22.
-\url{https://doi.org/10.1007/s12530-013-9083-8}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016}{}}%
-Vilhena, João, Henrique Vicente, M. Rosário Martins, José M. Grañeda,
-Filomena Caldeira, Rodrigo Gusmão, João Neves, and José Neves. 2016.
-{``A Case-Based Reasoning View of Thrombophilia Risk.''} \emph{Journal
-of Biomedical Informatics} 62 (August): 265--75.
-\url{https://doi.org/10.1016/j.jbi.2016.07.013}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wang2019}{}}%
-Wang, Ni, Yanqun Huang, Honglei Liu, Xiaolu Fei, Lan Wei, Xiangkun Zhao,
-and Hui Chen. 2019. {``Measurement and Application of Patient Similarity
-in Personalized Predictive Modeling Based on Electronic Medical
-Records.''} \emph{BioMedical Engineering OnLine} 18 (1).
-\url{https://doi.org/10.1186/s12938-019-0718-2}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
-Wang, Yuning, Yingzi Liang, Hui Sun, and Yufei Yang. 2020. {``Emergency
-{Response} for {COVID-19 Prevention} and {Control} in {Urban Rail
-Transit Based} on {Case-Based Reasoning Method}.''} \emph{Discrete
-Dynamics in Nature \& Society}, December, 1--9.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Weber2015}{}}%
-Weber, Griffin M. 2015. {``Federated Queries of Clinical Data
-Repositories: {Scaling} to a National Network.''} \emph{Journal of
-Biomedical Informatics} 55 (June): 231--36.
-\url{https://doi.org/10.1016/j.jbi.2015.04.012}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Weiskopf2013}{}}%
-Weiskopf, Nicole G., George Hripcsak, Sushmita Swaminathan, and Chunhua
-Weng. 2013. {``Defining and Measuring Completeness of Electronic Health
-Records for Secondary Use.''} \emph{Journal of Biomedical Informatics}
-46 (5): 830--36. \url{https://doi.org/10.1016/j.jbi.2013.06.010}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Welch2013}{}}%
-Welch, Brandon M, and Kensaku Kawamoto. 2013. {``Clinical Decision
-Support for Genetically Guided Personalized Medicine: A Systematic
-Review.''} \emph{Journal of the American Medical Informatics
-Association} 20 (2): 388--400.
-\url{https://doi.org/10.1136/amiajnl-2012-000892}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004}{}}%
-Wyns, B., S. Sette, L. Boullart, D. Baeten, I. E. A. Hoffman, and F. De
-Keyser. 2004. {``Prediction of Diagnosis in Patients with Early
-Arthritis Using a Combined {Kohonen} Mapping and Instance-Based
-Evaluation Criterion.''} \emph{Artificial Intelligence in Medicine} 31
-(1): 45--55. \url{https://doi.org/10.1016/j.artmed.2004.01.002}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
-Xing, Eric, Michael Jordan, Stuart J Russell, and Andrew Ng. 2002.
-{``Distance {Metric Learning} with {Application} to {Clustering} with
-{Side-Information}.''} In \emph{Advances in {Neural Information
-Processing Systems}}. Vol. 15. MIT Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Xu2008}{}}%
-Xu, Yanwen, Chetwyn C. H. Chan, Karen Hui Yu-Ling Lo, and Dan Tang.
-2008. {``\href{https://www.ncbi.nlm.nih.gov/pubmed/18198444}{Prediction
-Model for the Return to Work of Workers with Injuries in {Hong
-Kong}.}''} \emph{Work} 30 (1): 77--84.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
-Yang, Liu. 2006. {``Distance {Metric Learning}: {A Comprehensive
-Survey}.''} PhD thesis, Department of Computer Science; Engineering:
-Michigan State University.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997}{}}%
-Yearwood, J., and R. Wilkinson. 1997. {``Retrieving Cases for Treatment
-Advice in Nursing Using Text Representation and Structured Text
-Retrieval.''} \emph{Artificial Intelligence in Medicine} 9 (1): 79--99.
-\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018}{}}%
-Zhang, HJ, F Zhu, HH Dodge, GA Higgins, GS Omenn, YF Guan, and
-Alzheimer's Dis Neuroimaging Initi. 2018. {``A Similarity-Based Approach
-to Leverage Multi-Cohort Medical Data on the Diagnosis and Prognosis of
-{Alzheimer}'s Disease.''} \emph{GigaScience} 7 (7).
-\url{https://doi.org/10.1093/gigascience/giy085}.
-
-\end{CSLReferences}
-
-\bibliographystyle{unsrt}
-\bibliography{SR-bibliography.bibSR-reviewers.bibSR-0-Found.bibSR-2-Included.bibSR-3-Tracked.bibSR-5-Screened.bib}
+\renewcommand\refname{References}
+\bibliography{SR-bibliography.bib,SR-reviewers.bib,SR-0-Found.bib,SR-2-Included.bib,SR-3-Tracked.bib,SR-5-Screened.bib}
 

### 2025-03-09 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
edit latex document for reverse-inclusion into rmarkdown document

@@
 retrieved cases \hl{ to inform human decision-making}\citep{Aamodt1994};
-the related nearest neighbors (NN) technique generates predictions from
+the related nearest neighbors (\hl{kNN}) technique generates predictions from
 cohorts automatically via averaging (regression) or voting
@@
 Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures and algorithms exist for similarity matching }\citep{Halder2024}\hl{ and the local models would benefit from increased numbers of relevant cases.}
-\hl{SCBM}s thus provide an individualized way to fit fully interpretable
-models to large population data.
 
@@
 
-\hl{We also note that a large variety of AI tools, besides NN, use similarity measures or other relevance-based retrieval methods: Back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
+\hl{We also note that a large variety of AI tools, besides kNN, use similarity measures or other relevance-based retrieval methods: Back-propagation in ANNs relies on differentiation, which requires that all network nodes' activation functions and, therefore, the final model are continuous and differentiable, i.e.
 that small distances between cases yield small changes in prediction; and kernel-based methods that rely directly on pairwise distances computed via inner product, rather than high-dimensional coordinates, extend far beyond their use in CBR.
@@
 Of these, 51 fit the original criteria, and 21 were included as distinct
-from NN prediction.
+from \hl{kNN} prediction.
 
@@
 \citep{Park2006, Lowsky2013, Lee2015, Ng2015, Lee2017, Wang2019}, 1 of
-which was excluded from the synthesis for using NN prediction. Reference
+which was excluded from the synthesis for using \hl{kNN} prediction. Reference
 tracking from the 43 + 360 = 403 studies assessed for eligibility led us
@@
 soon after CBR had established itself; the idea has been ``in the air''
-for as long as CBR has been in use.
+for as long as CBR has been in use.\hl{ Based on the motivations and methodologies observed in our sample (discussed in later sections), we speculate that the higher frequency of studies since 2013 is attributable in large part to the availability of vaster clinical databases with improved pre-processing tools, following the rapid uptake of EHRs; and of more efficient and powerful tools for retrieval (similarity-matching), the traditional bottleneck of CBR.
+Another possibility is an increased interest in interpretable machine learning }\citep{Rudin2022}\hl{.
+Each of these factors should apply with no less force to other CDSS and clinical applications of ML, and the rise we see may simply be commensurate with an overall increase in the volume of articles in these fields.}
 
@@
 decision support \citep{Doyle2006}. This study was part of a much larger
-literature on CBR systems and was included here despite relying on NN
+literature on CBR systems and was included here despite relying on \hl{kNN}
 prediction because it used an unconventional voting scheme to generate
@@
 inequality. Each used its similarity measure to retrieve training cases
-relevant to each testing case. \hl{Most} took a standard NN approach to
+relevant to each testing case. \hl{Most} took a standard \hl{kNN} approach to
 generating predictions; \hl{two} \citep{Vilhena2016, Tang2021} fit
@@
 predictive accuracy on the training set. This is analogous to optimizing
-the neighborhood size parameter in NN predictive modeling, so we will
+the neighborhood size parameter in \hl{kNN} predictive modeling, so we will
 not discuss it further except to mention that the authors found
@@
 exhaustion to optimize the size of the retrieved cohort for the accuracy
-of the NN model using the previously optimized weights.
+of the \hl{kNN} model using the previously optimized weights.
 
@@
 hyperparameter of the individualized modeling approach, cohort size can
-be treated in the same way as neighborhood size in NN \hl{prediction},
+be treated in the same way as neighborhood size in \hl{kNN} \hl{prediction},
 so we will not discuss it further.
@@
 
-\hl{While most studies identified outstanding technical needs (Section }\ref{identified-needs}\hl{), few (though some) emphasized the need to assess human-focused qualities like user interface and user experience, interpretability, meaningfulness, or trust.
+\hl{While most studies identified outstanding technical needs (Section }\ref{identified-needs}\hl{), few (though some) emphasized the need to assess interpretability, meaningfulness, or trust.
 Most studies identified interpretability as an advantage of localized models, though none evaluated model interpretability and few proposed new interpretative uses.
@@
 \item
-  NN: nearest neighbors
+  \hl{kNN}: nearest neighbors
 \item
@@
 & & & & \hspace{6em} & LR & 91.14\%\hspace{6em} \\
-& & & & \hspace{6em} & NN & 88.29\%\hspace{6em} \\
+& & & & \hspace{6em} & \hl{kNN} & 88.29\%\hspace{6em} \\
 & & Heart & Statistical CBR & 83.70\%\hspace{6em} & LR &
 84.44\%\hspace{6em} \\
-& & & & \hspace{6em} & NN & 82.22\%\hspace{6em} \\
+& & & & \hspace{6em} & \hl{kNN} & 82.22\%\hspace{6em} \\
 & & & & \hspace{6em} & CBR & 81.85\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & C5.0 & 76.30\%\hspace{6em} \\
-& & Breast & Statistical CBR & 96.96\%\hspace{6em} & NN &
+& & Breast & Statistical CBR & 96.96\%\hspace{6em} & \hl{kNN} &
 97.50\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & CART & 72.89\%\hspace{6em} \\
-& & & & \hspace{6em} & NN & 65.39\%\hspace{6em} \\
+& & & & \hspace{6em} & \hl{kNN} & 65.39\%\hspace{6em} \\
 & & Liver & Statistical CBR & 66.76\%\hspace{6em} & CART &
@@
 & & & & \hspace{6em} & CBR & 60.88\%\hspace{6em} \\
-& & & & \hspace{6em} & NN & 55.59\%\hspace{6em} \\
+& & & & \hspace{6em} & \hl{kNN} & 55.59\%\hspace{6em} \\
 \citet{Song2006} & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
@@
 & & & Individualized CSRF & 0.832\hspace{6em} & & \hspace{6em} \\
-\citet{Malykh2018} & Accuracy & J13/pneumonia & Case-based NN &
+\citet{Malykh2018} & Accuracy & J13/pneumonia & Case-based \hl{ANN} &
 81.6\%\hspace{6em} & & \hspace{6em} \\
-& & K80.1/gallbladder & Case-based NN & 76.7\%\hspace{6em} & &
+& & K80.1/gallbladder & Case-based \hl{ANN} & 76.7\%\hspace{6em} & &
 \hspace{6em} \\
-& & H25.1/age-related cataract & Case-based NN & 94.9\%\hspace{6em} & &
+& & H25.1/age-related cataract & Case-based \hl{ANN} & 94.9\%\hspace{6em} & &
 \hspace{6em} \\
-& & H26.2/complicated cataract & Case-based NN & 91.4\%\hspace{6em} & &
+& & H26.2/complicated cataract & Case-based \hl{ANN} & 91.4\%\hspace{6em} & &
 \hspace{6em} \\
-& & 167.4/encephalopathy & Case-based NN & 72.4\%\hspace{6em} & &
+& & 167.4/encephalopathy & Case-based \hl{ANN} & 72.4\%\hspace{6em} & &
 \hspace{6em} \\
-& & 167.9/CBD & Case-based NN & 75.4\%\hspace{6em} & & \hspace{6em} \\
-& & N20.1/ureter & Case-based NN & 58.7\%\hspace{6em} & &
+& & 167.9/CBD & Case-based \hl{ANN} & 75.4\%\hspace{6em} & & \hspace{6em} \\
+& & N20.1/ureter & Case-based \hl{ANN} & 58.7\%\hspace{6em} & &
 \hspace{6em} \\

### 2025-03-09 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
write additional paragraphs + visualize performance dependencies

@@
 The approach is a specialized form of case-based reasoning (CBR) that relies on a \hl{patient similarity measure }[@Brown2016] to extract a cohort of past or training cases that then inform the diagnosis, prognosis, or care of a new or test case.[^psm]
-\hl{Some of the earliest and simplest CBR tools returned} these retrieved cases \hl{ to inform human decision-making}[@Aamodt1994]; the related nearest neighbors (NN) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
+\hl{Some of the earliest and simplest CBR tools returned} these retrieved cases \hl{ to inform human decision-making}[@Aamodt1994]; the related nearest neighbors (\hl{kNN}) technique generates predictions from cohorts automatically via averaging (regression) or voting (classification) of their outcomes.
 In contrast, \hl{SCBM} fits families of predictive models, for example generalized linear models (GLMs), to retrieved similarity cohorts in order to generate predictions.
@@
 Moreover, this approach would benefit computationally as well as practically from the availability of a large corpus of cases, as efficient data structures and algorithms exist for similarity matching }[@Halder2024]\hl{ and the local models would benefit from increased numbers of relevant cases.}
-\hl{SCBM}s thus provide an individualized way to fit fully interpretable models to large population data.
 
@@
 
-This approach harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
+\hl{Interpretability is often coupled and contrasted with \emph{explainability}.
+We follow }@Lisboa2023\hl{ in distinguishing these properties in terms of \emph{ante-hoc} versus \emph{post-hoc} model design and analysis choices, though we note that this usage is not universal }[@Molnar2023]\hl{.
+Work on explainable ML (or AI) is driven by many of the same needs as interpretable ML, including SCBM, in particular extracting knowledge from routinely collected data and balancing improvements in efficiency and accuracy with user or client comprehension and trust }[@Lisboa2023]\hl{.
+However, we find compelling the case that interpretable models are preferable to explainable models when their performance is commensurate, especially for high-stakes decisions, due to the often shallow and fragile insights offered by explainers in comparison to models with transparent and comprehensible structure }[@Rudin2022]\hl{.
+This review is driven by our interest specifically in interpretable ML, and by our suspicion that cohort retrieval has more potential for exploratory and predictive modeling of EHR and other routinely collected health data than is yet being exploited.}
+
+\hl{SCBM} harkens to the aspirational "green button" that would, in response to a query, automatically retrieve patient data from an institution’s records with which to conduct on-demand retrospective studies for an individual patient [@Longhurst2014].
 Several hurdles face the deployment of such an approach in practice, but its benefits must first be demonstrated.
@@
 Clinical CBR emerged among rule-based approaches and other AI tools in the development of expert systems [@Aamodt1994].
-Early implementations realized a general workflow described as "the four REs", later the "R4 cycle" [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system _retrieves_ one or more past cases (solved problems) from a corpus, _reuses_ these to generate an understanding of (solution to) the new case, _revises_ this understanding (solution) to better fit the new case (sometimes called "adaptation"), and _retains_ the new case and its eventual understanding (solution) in the corpus to be retrieved and reused in future.
+Early implementations realized a general workflow described as "the four REs", later the "R4 cycle" [@Aamodt1994; @Begum2011]: Given a new case (or problem), the system _retrieves_ one or more past cases (solved problems) from a corpus, _reuses_ these to generate an understanding of (solution to) the new case, _revises_ \hl{or \emph{adapts}}this understanding (solution) to better fit the new \hl{case,} and _retains_ the new case and its eventual understanding (solution) in the corpus to be retrieved and reused in future.
 In contrast to rule-based systems, which are variable-based and generally interpretable as a single rule applied to all new cases, case-based systems provide case-specific interpretations in the form of a number of more fully understood reference cases.
@@
 @Kolodner1992 distinguished two styles of CBR: _problem-solving_, which is more procedural and used when objectives are more clearly defined, and _interpretative_, which provides categorizations and justifications for possible solutions.
-A similar dichotomy is commonly used to distinguish the _performance_ \hl{criterion }of predictive models from the _interpretability_ criterion\hl{.} In these terms, \hl{SCBM}s are highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
+A similar dichotomy is commonly used to distinguish the _performance_ \hl{criterion }of predictive models from the _interpretability_ criterion\hl{.}
+@Rudin2022\hl{ distinguish \emph{prototype-based} CBR, based on the selection or creation of prototype cases for more efficient matching, from classical \emph{nearest neighbor--based} CBR.}
+In these terms, \hl{SCBM}s are \hl{nearest neighbor--based and }highly procedural: The step of fitting a predictive model to a retrieved cohort is an automated adaptive strategy [@Begum2011], and the parameters that govern cohort retrieval can be tuned alongside model hyperparameters in a machine learning (ML) workflow.
 Accordingly, \hl{SCBM}s have been primarily used for and evaluated on their ability to predict classes or outcomes.
@@
 
-We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018].
-While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; \hl{CDSS} and software tools).
-For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.}
+We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016]\hl{, of CDSS more broadly }[@Wasylewicz2019]\hl{ and of AI-driven systems }[@Elhaddad2024], \hl{of patient similarity measures} [@Dai2020]\hl{, of kNN algorithms }[@Halder2024], \hl{of theory and uses of CBR for} predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018; @Badra2023]\hl{, and of recent advances in CBR with emphasis on interpretability and explainability }[@Schoenborn2021; @Rudin2022].
+While the reviews of CBR \hl{and CDSS }focus on applications using health data, the \hl{others consider }additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction\hl{; }software tools).
+
+For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.
+Moreover, we found no reviews or surveys that focus specifically on the kinds of models we call SCBMs or include detailed comparisons between them.}
 
@@
 Of these, 360 entries met the screening criteria and were assessed through full text review.
-Of these, 51 fit the original criteria, and 21 were included as distinct from NN prediction.
+Of these, 51 fit the original criteria, and 21 were included as distinct from \hl{kNN} prediction.
 
@@
 We then reviewed \hl{a seed set comprising 43 studies} that inspired this review.
-After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using NN prediction.
+After removing duplicates and screening for eligibility, we were left with 6 additional studies [@Park2006; @Lowsky2013; @Lee2015; @Ng2015; @Lee2017; @Wang2019], 1 of which was excluded from the synthesis for using \hl{kNN} prediction.
 Reference tracking from the 43 + 360 = 403 studies assessed for eligibility led us to identify 1 additional study that met our criteria [@Kasabov2010].
@@
 We note that CBR in health and medicine originated as early as 1990, and \hl{SCBM} emerged soon after CBR had established itself; the idea has been "in the air" for as long as CBR has been in use.
+\hl{ Based on the motivations and methodologies observed in our sample (discussed in later sections), we speculate that the higher frequency of studies since 2013 is attributable in large part to the availability of vaster clinical databases with improved pre-processing tools, following the rapid uptake of EHRs; and of more efficient and powerful tools for retrieval (similarity-matching), the traditional bottleneck of CBR.
+Another possibility is an increased interest in interpretable machine learning }[@Rudin2022]\hl{.
+Each of these factors should apply with no less force to other CDSS and clinical applications of ML, and the rise we see may simply be commensurate with an overall increase in the volume of articles in these fields.}
 
@@
 For example, a 2006 study specifically evaluated the usefulness of CBR-based explanations for the purpose of decision support [@Doyle2006].
-This study was part of a much larger literature on CBR systems and was included here despite relying on NN prediction because it used an unconventional voting scheme to generate recommendations.
+This study was part of a much larger literature on CBR systems and was included here despite relying on \hl{kNN} prediction because it used an unconventional voting scheme to generate recommendations.
 A more recent study took essentially the same focus with respect to a proposed clinical risk prediction model, which amounted to CBR with a novel weighting scheme on predictors informed by expert consensus [@Fang2021].
@@
 
+Figure S1 **TODO**.
+
 ## Identified needs
@@
 These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as \hl{similarity cohort--based modeling}; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
-We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
-\hl{While ``local'' naturally contrasts with ``global'', the overloaded term ``localized models'' stands to cause more confusion than it resolves, so we use ``similarity cohort--based models'' to refer to this specific approach.}
+\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)} We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
+\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves, so we use \emph{similarity cohort--based models} to refer to this specific approach.}
 
@@
 Each used its similarity measure to retrieve training cases relevant to each testing case.
-\hl{Most} took a standard NN approach to generating predictions; \hl{two} [@Vilhena2016; @Tang2021] fit predictive models to the retrieved cohorts.
+\hl{Most} took a standard \hl{kNN} approach to generating predictions; \hl{two} [@Vilhena2016; @Tang2021] fit predictive models to the retrieved cohorts.
 
@@
 This was an adaptive step taken by @Park2006 to optimize the predictive performance of models fitted to cohorts retrieved using a fixed similarity rather than count threshold, a counterpart to conventional CBR they called "statistical CBR".[^combinatorial] The step used a heuristic procedure to locate a similarity threshold that (locally) maximizes predictive accuracy on the training set.
-This is analogous to optimizing the neighborhood size parameter in NN predictive modeling, so we will not discuss it further except to mention that the authors found statistical CBR to outperform conventional CBR on several data sets.
-@CampilloGimenez2013 combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the NN model using the previously optimized weights.
+This is analogous to optimizing the neighborhood size parameter in \hl{kNN} predictive modeling, so we will not discuss it further except to mention that the authors found statistical CBR to outperform conventional CBR on several data sets.
+@CampilloGimenez2013 combined these learning techniques: They used logistic regression on the training set to obtain weights for the predictors (by predicting the binary outcome of kidney transplant waitlist registration) and for the cases (by predicting agreement of outcome between an index case and other training cases), and they used exhaustion to optimize the size of the retrieved cohort for the accuracy of the \hl{kNN} model using the previously optimized weights.
 
@@
 Despite the early recommendation by @Park2006 to threshold cohorts by similarity rather than by cardinality, only the one previously published study [@Mariuzzi1997] and the most recent study [@Tang2021; @Ng2021] in this corpus took this approach, though @Liu2022 took a third option by retrieving fixed proportions of training data.
-Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in NN \hl{prediction}, so we will not discuss it further.
+Viewed as a hyperparameter of the individualized modeling approach, cohort size can be treated in the same way as neighborhood size in \hl{kNN} \hl{prediction}, so we will not discuss it further.
 
@@
 * AI: artificial intelligence
-* NN: nearest neighbors
+* \hl{kNN}: nearest neighbors
 * GLM: generalized linear model
@@
 
-**TODO: Place vertical separators between major columns.**
+**TODO: Place vertical separators (` |`) between major columns.**
 

### 2025-03-09 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
write additional paragraphs + visualize performance dependencies

@@
 retrieved cases \hl{ to inform human decision-making}\citep{Aamodt1994};
-the related nearest neighbors (\hl{kNN}) technique generates predictions from
-cohorts automatically via averaging (regression) or voting
+the related nearest neighbors (\hl{kNN}) technique generates predictions
+from cohorts automatically via averaging (regression) or voting
 (classification) of their outcomes. In contrast, \hl{SCBM} fits families
@@
 
-This approach harkens to the aspirational ``green button'' that would,
-in response to a query, automatically retrieve patient data from an
+\hl{Interpretability is often coupled and contrasted with \emph{explainability}.
+We follow }\citet{Lisboa2023}\hl{ in distinguishing these properties in terms of \emph{ante-hoc} versus \emph{post-hoc} model design and analysis choices, though we note that this usage is not universal }\citep{Molnar2023}\hl{.
+Work on explainable ML (or AI) is driven by many of the same needs as interpretable ML, including SCBM, in particular extracting knowledge from routinely collected data and balancing improvements in efficiency and accuracy with user or client comprehension and trust }\citep{Lisboa2023}\hl{.
+However, we find compelling the case that interpretable models are preferable to explainable models when their performance is commensurate, especially for high-stakes decisions, due to the often shallow and fragile insights offered by explainers in comparison to models with transparent and comprehensible structure }\citep{Rudin2022}\hl{.
+This review is driven by our interest specifically in interpretable ML, and by our suspicion that cohort retrieval has more potential for exploratory and predictive modeling of EHR and other routinely collected health data than is yet being exploited.}
+
+\hl{SCBM} harkens to the aspirational ``green button'' that would, in
+response to a query, automatically retrieve patient data from an
 institution's records with which to conduct on-demand retrospective
@@
 (solved problems) from a corpus, \emph{reuses} these to generate an
-understanding of (solution to) the new case, \emph{revises} this
-understanding (solution) to better fit the new case (sometimes called
-``adaptation''), and \emph{retains} the new case and its eventual
+understanding of (solution to) the new case, \emph{revises}
+\hl{or \emph{adapts}}this understanding (solution) to better fit the new
+\hl{case,} and \emph{retains} the new case and its eventual
 understanding (solution) in the corpus to be retrieved and reused in
@@
 \hl{criterion }of predictive models from the \emph{interpretability}
-criterion\hl{.} In these terms, \hl{SCBM}s are highly procedural: The
-step of fitting a predictive model to a retrieved cohort is an automated
-adaptive strategy \citep{Begum2011}, and the parameters that govern
-cohort retrieval can be tuned alongside model hyperparameters in a
-machine learning (ML) workflow. Accordingly, \hl{SCBM}s have been
-primarily used for and evaluated on their ability to predict classes or
-outcomes. Nevertheless, as \hl{we find in this review}, they can be used
-to draw inferences about the importance of different risk factors to
-specific individuals.
+criterion\hl{.}
+\citet{Rudin2022}\hl{ distinguish \emph{prototype-based} CBR, based on the selection or creation of prototype cases for more efficient matching, from classical \emph{nearest neighbor--based} CBR.}
+In these terms, \hl{SCBM}s are \hl{nearest neighbor--based and }highly
+procedural: The step of fitting a predictive model to a retrieved cohort
+is an automated adaptive strategy \citep{Begum2011}, and the parameters
+that govern cohort retrieval can be tuned alongside model
+hyperparameters in a machine learning (ML) workflow. Accordingly,
+\hl{SCBM}s have been primarily used for and evaluated on their ability
+to predict classes or outcomes. Nevertheless, as
+\hl{we find in this review}, they can be used to draw inferences about
+the importance of different risk factors to specific individuals.
 
@@
 We refer the interested reader to several previous reviews of CBR in
-medicine \citep{Gierl1998, Begum2011, Choudhury2016}, of measures of
-patient similarity \citep{Dai2020}, and of uses of patient similarity in
-predictive models \citep{Welch2013, Sharafoddini2017, Parimbelli2018}.
-While the reviews of CBR focus on applications using health data, the
-patient similarity reviews encompass many additional types of data
-(various molecular -omics, genetic tests, medical images, laboratory
-tests, patient preferences, patient-reported outcomes, tracking devices,
-social media) and survey a much broader scope of models (exploratory
-analysis via dimension reduction and cluster analysis; risk evaluation
-and outcome prediction; \hl{CDSS} and software tools). For the present
-review, we are interested in how similarity matching on patient-level
-health data is used to construct local cohorts for predictive or
-inferential
-modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.}
+medicine
+\citep{Gierl1998, Begum2011, Choudhury2016}\hl{, of CDSS more broadly }\citep{Wasylewicz2019}\hl{ and of AI-driven systems }\citep{Elhaddad2024},
+\hl{of patient similarity measures}
+\citep{Dai2020}\hl{, of kNN algorithms }\citep{Halder2024},
+\hl{of theory and uses of CBR for} predictive models
+\citep{Welch2013, Sharafoddini2017, Parimbelli2018, Badra2023}\hl{, and of recent advances in CBR with emphasis on interpretability and explainability }\citep{Schoenborn2021, Rudin2022}.
+While the reviews of CBR \hl{and CDSS }focus on applications using
+health data, the \hl{others consider }additional types of data (various
+molecular -omics, genetic tests, medical images, laboratory tests,
+patient preferences, patient-reported outcomes, tracking devices, social
+media) and survey a much broader scope of models (exploratory analysis
+via dimension reduction and cluster analysis; risk evaluation and
+outcome prediction\hl{; }software tools).
+
+For the present review, we are interested in how similarity matching on
+patient-level health data is used to construct local cohorts for
+predictive or inferential
+modeling.\hl{ While the similarity measure used in the retrieval step need not be changed as the corpus grows, proposed measures have been diverse, contested, and rarely systematically compared.
+Moreover, we found no reviews or surveys that focus specifically on the kinds of models we call SCBMs or include detailed comparisons between them.}
 
@@
 \citep{Park2006, Lowsky2013, Lee2015, Ng2015, Lee2017, Wang2019}, 1 of
-which was excluded from the synthesis for using \hl{kNN} prediction. Reference
-tracking from the 43 + 360 = 403 studies assessed for eligibility led us
-to identify 1 additional study that met our criteria
+which was excluded from the synthesis for using \hl{kNN} prediction.
+Reference tracking from the 43 + 360 = 403 studies assessed for
+eligibility led us to identify 1 additional study that met our criteria
 \citep{Kasabov2010}. This left us with 21 + 1 + 5 = 27 studies included
@@
 soon after CBR had established itself; the idea has been ``in the air''
-for as long as CBR has been in use.\hl{ Based on the motivations and methodologies observed in our sample (discussed in later sections), we speculate that the higher frequency of studies since 2013 is attributable in large part to the availability of vaster clinical databases with improved pre-processing tools, following the rapid uptake of EHRs; and of more efficient and powerful tools for retrieval (similarity-matching), the traditional bottleneck of CBR.
+for as long as CBR has been in use.
+\hl{ Based on the motivations and methodologies observed in our sample (discussed in later sections), we speculate that the higher frequency of studies since 2013 is attributable in large part to the availability of vaster clinical databases with improved pre-processing tools, following the rapid uptake of EHRs; and of more efficient and powerful tools for retrieval (similarity-matching), the traditional bottleneck of CBR.
 Another possibility is an increased interest in interpretable machine learning }\citep{Rudin2022}\hl{.
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1154}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1209}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1978}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1044}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1538}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1123}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1497}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1176}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1925}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1497}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1283}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1497}}@{}}
 \caption{\label{tab:synthesis}\hl{Studies included in the method synthesis, arranged by the earliest the study is known to have been public.}}\tabularnewline
@@
 Clinical, Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 366; 270;
-560; 760 & 35; 14; 31; 8 \\
+560; 760; 340 & 35; 14; 31; 8; 7 \\
 \citet{Song2006} & Prognosis\hspace{6em} & Practice\hspace{6em} &
-Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000; 447 & 6;
-6 \\
+Clinical\hspace{6em} & Cross-sectional\hspace{6em} & 1,000 & 6 \\
 \citet{Elter2007} & Diagnosis\hspace{6em} & Practice\hspace{6em} &
@@
 \citet{Zhang2018} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
-Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 810; 302 & 8 \\
+Imaging\hspace{6em} & Cross-sectional\hspace{6em} & 302 & 8 \\
 \citet{Malykh2018} & Recommend-ation\hspace{6em} & Practice\hspace{6em}
@@
 decision support \citep{Doyle2006}. This study was part of a much larger
-literature on CBR systems and was included here despite relying on \hl{kNN}
-prediction because it used an unconventional voting scheme to generate
-recommendations. A more recent study took essentially the same focus
-with respect to a proposed clinical risk prediction model, which
+literature on CBR systems and was included here despite relying on
+\hl{kNN} prediction because it used an unconventional voting scheme to
+generate recommendations. A more recent study took essentially the same
+focus with respect to a proposed clinical risk prediction model, which
 amounted to CBR with a novel weighting scheme on predictors informed by
@@
 
+Figure S1 \textbf{TODO}.
+
 \subsection{Identified needs}\label{identified-needs}
@@
 models'']{Liang2015}; \citealp{Ng2015}; \citealp{Wang2019}; \citealp{Ma2020}; \citealp{Liu2022}; \citealp{Doborjeh2022}},
-and ``precision cohort'' analysis \citep{Wang2019, Tang2021, Ng2021}. We
-find uses of ``targeted'', ``personalized'', and ``precision'' generic
-and imprecise, while transductive inference is an established term for a
-broader set of methods in ML.
-\hl{While ``local'' naturally contrasts with ``global'', the overloaded term ``localized models'' stands to cause more confusion than it resolves, so we use ``similarity cohort--based models'' to refer to this specific approach.}
+and ``precision cohort'' analysis \citep{Wang2019, Tang2021, Ng2021}.
+\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)}
+We find uses of ``targeted'', ``personalized'', and ``precision''
+generic and imprecise, while transductive inference is an established
+term for a broader set of methods in ML.
+\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves, so we use \emph{similarity cohort--based models} to refer to this specific approach.}
 
@@
 inequality. Each used its similarity measure to retrieve training cases
-relevant to each testing case. \hl{Most} took a standard \hl{kNN} approach to
-generating predictions; \hl{two} \citep{Vilhena2016, Tang2021} fit
-predictive models to the retrieved cohorts.
+relevant to each testing case. \hl{Most} took a standard \hl{kNN}
+approach to generating predictions; \hl{two}
+\citep{Vilhena2016, Tang2021} fit predictive models to the retrieved
+cohorts.
 
@@
 predictive accuracy on the training set. This is analogous to optimizing
-the neighborhood size parameter in \hl{kNN} predictive modeling, so we will
-not discuss it further except to mention that the authors found
+the neighborhood size parameter in \hl{kNN} predictive modeling, so we
+will not discuss it further except to mention that the authors found
 statistical CBR to outperform conventional CBR on several data sets.
@@
 hyperparameter of the individualized modeling approach, cohort size can
-be treated in the same way as neighborhood size in \hl{kNN} \hl{prediction},
-so we will not discuss it further.
+be treated in the same way as neighborhood size in \hl{kNN}
+\hl{prediction}, so we will not discuss it further.
 
@@
 
-%\textbf{TODO: Place vertical separators between major columns.}
+\textbf{TODO: Place vertical separators (\texttt{\textbar{}}) between
+major columns.}
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1088}} |
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1088}}
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0725}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1658}} |
+  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1658}}
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2435}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}} |
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0984}}
   >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2124}}
@@
 & & & & \hspace{6em} & LR & 91.14\%\hspace{6em} \\
-& & & & \hspace{6em} & \hl{kNN} & 88.29\%\hspace{6em} \\
+& & & & \hspace{6em} & kNN & 88.29\%\hspace{6em} \\
 & & Heart & Statistical CBR & 83.70\%\hspace{6em} & LR &
 84.44\%\hspace{6em} \\
-& & & & \hspace{6em} & \hl{kNN} & 82.22\%\hspace{6em} \\
+& & & & \hspace{6em} & kNN & 82.22\%\hspace{6em} \\
 & & & & \hspace{6em} & CBR & 81.85\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & C5.0 & 76.30\%\hspace{6em} \\
-& & Breast & Statistical CBR & 96.96\%\hspace{6em} & \hl{kNN} &
+& & Breast & Statistical CBR & 96.96\%\hspace{6em} & kNN &
 97.50\%\hspace{6em} \\
@@
 & & & & \hspace{6em} & CART & 72.89\%\hspace{6em} \\
-& & & & \hspace{6em} & \hl{kNN} & 65.39\%\hspace{6em} \\
+& & & & \hspace{6em} & kNN & 65.39\%\hspace{6em} \\
 & & Liver & Statistical CBR & 66.76\%\hspace{6em} & CART &
@@
 & & & & \hspace{6em} & CBR & 60.88\%\hspace{6em} \\
-& & & & \hspace{6em} & \hl{kNN} & 55.59\%\hspace{6em} \\
+& & & & \hspace{6em} & kNN & 55.59\%\hspace{6em} \\
 \citet{Song2006} & RMSE & GFR & TWNFI & 7.08\hspace{6em} & MDRD &
@@
 & & & & \hspace{6em} & WKNN & 90.3\%\hspace{6em} \\
-& & Crohn's disease risk & TWNFI & 82.45\%\hspace{6em} & &
-\hspace{6em} \\
 \citet{Liang2015} & Accuracy & Colon cancer & knnGSA &
@@
 & & & Individualized CSRF & 0.832\hspace{6em} & & \hspace{6em} \\
-\citet{Malykh2018} & Accuracy & J13/pneumonia & Case-based \hl{ANN} &
-81.6\%\hspace{6em} & & \hspace{6em} \\
-& & K80.1/gallbladder & Case-based \hl{ANN} & 76.7\%\hspace{6em} & &
-\hspace{6em} \\
-& & H25.1/age-related cataract & Case-based \hl{ANN} & 94.9\%\hspace{6em} & &
-\hspace{6em} \\
-& & H26.2/complicated cataract & Case-based \hl{ANN} & 91.4\%\hspace{6em} & &
-\hspace{6em} \\
-& & 167.4/encephalopathy & Case-based \hl{ANN} & 72.4\%\hspace{6em} & &
-\hspace{6em} \\
-& & 167.9/CBD & Case-based \hl{ANN} & 75.4\%\hspace{6em} & & \hspace{6em} \\
-& & N20.1/ureter & Case-based \hl{ANN} & 58.7\%\hspace{6em} & &
-\hspace{6em} \\
+\citet{Malykh2018} & Accuracy & J13/pneumonia & & \hspace{6em} &
+Case-based ANN & 81.6\%\hspace{6em} \\
+& & K80.1/gallbladder & & \hspace{6em} & Case-based ANN &
+76.7\%\hspace{6em} \\
+& & H25.1/age-related cataract & & \hspace{6em} & Case-based ANN &
+94.9\%\hspace{6em} \\
+& & H26.2/complicated cataract & & \hspace{6em} & Case-based ANN &
+91.4\%\hspace{6em} \\
+& & 167.4/encephalopathy & & \hspace{6em} & Case-based ANN &
+72.4\%\hspace{6em} \\
+& & 167.9/CBD & & \hspace{6em} & Case-based ANN & 75.4\%\hspace{6em} \\
+& & N20.1/ureter & & \hspace{6em} & Case-based ANN &
+58.7\%\hspace{6em} \\
 \citet{Zhang2018} & AUROC & Cognitive impairment & Gaussian process

### 2025-03-09 scr/synthesis.r

@@ -58,8 +58,8 @@ properties_inclusion %>%
     Aim = ifelse(generalizable_knowledge == "Science", "Knowledge", "Practice"),
     Source = type_of_data,
     Type = range_of_data,
-    `Cases` = number_of_cases_incidents,
-    `Features` = number_of_predictors_features
+    Cases = number_of_cases_incidents,
+    Features = number_of_predictors_features
   ) %>%
   mutate(across(
     c(Cases, Features),
@@ -125,12 +125,17 @@ properties_inclusion %>%
   transmute(
     zotero_key, date_of_publication,
     proposal = proposed_methods,
+    cases = number_of_cases_incidents,
+    features = number_of_predictors_features,
     evaluation = evaluation_comparison_approach,
     measure = performance_measure,
     performance = performance_values
   ) %>%
+  # break `cases`, `features`, and `performance` together
+  mutate(across(c(cases, features), \(l) str_split(l, "; "))) %>%
   # break `performance` into nested data with task, method, value
-  mutate(performance = str_split(performance, "\n")) %>% unnest(performance) %>%
+  mutate(performance = str_split(performance, "\n")) %>% 
+  unnest(c(cases, features, performance)) %>%
   separate(col = performance, into = c("task", "results"), sep = ": ") %>%
   mutate(results = str_split(results, "; ")) %>% unnest(results) %>%
   mutate(results = str_remove(results, " \\(.*\\)$")) %>%
@@ -381,3 +386,66 @@ composite_approaches %>%
 print(method_freq)
 ggsave(here::here("fig/fig-methods.png"), method_freq, width = 8, height = 3)
 ggsave(here::here("fig/fig-methods.jpg"), method_freq, width = 8, height = 3)
+
+# extract evaluations of proposed methods
+tab_eval_comp |> 
+  pivot_wider(
+    id_cols = c(zotero_key, date_of_publication, measure, task, cases, features),
+    names_from = role, values_from = results
+  ) |> 
+  transmute(
+    zotero_key, Date = as_date(date_of_publication),
+    Measure = measure, Problem = task, Cases = cases, Features = features,
+    proposal
+  ) |> 
+  arrange(Date) |> 
+  unnest(proposal) |> 
+  separate_wider_delim(
+    proposal, delim = ": ",
+    names = c("Proposal", "Performance")
+  ) |> 
[…]

### 2025-03-10 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
complete update per reviewer comments + track bbl file

@@
 that small distances between cases yield small changes in prediction; and kernel-based methods that rely directly on pairwise distances computed via inner product, rather than high-dimensional coordinates, extend far beyond their use in CBR.
-The key difference in SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
+Additionally, }@Schoenborn2021\hl{ survey CBR-powered explanations of opaque models and taxonomize them as \emph{model-agnostic} versus \emph{model-based}, though this coupling of CBR with predictive models is inverted from that of SCBM.
+What distinguishes SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
 Indeed, retrieval procedures were the most significant algorithmic requirements and accomplishments of CBR }[@Kolodner1992; @Gierl1998]\hl{ and the retrieved cohorts may be reused for numerous modeling tasks.}
-Thus, \hl{SCBM}s may embody both styles of CBR.
 
@@
 Based on these criteria, we formulated search strings for five platforms: PubMed, Web of Science, Academic Search Premier, Google Scholar, and MathSciNet.
-We finalized the PubMed search first, then adapted it to the other platforms\hl{.}
+\hl{These platforms were chosen to cover a range of relevant disciplines (biomedicine, mathematics, broader sciences) and types of reports (journal articles, conference proceedings, grey literature.} We finalized the PubMed search first, then adapted it to the other platforms\hl{.}
 
@@
 
-Figure S1 **TODO**.
+\hl{Figure }\ref{fig:performance}\hl{ visualizes the relationship between performance and case count, feature count, data source, and predictive task for those experiments that evaluated models using accuracy or AUROC (other performance measures were too rare to allow for useful comparisons).
+While not represented in the plot, results from different models applied to the same data can be inferred from their common case counts, data sources, and tasks.
+The data are too heterogeneous to meaningfully test for trends.
+Performance varies much more between experiments than between models within experiments, and there is no clear dependency on the source(s) of data or on the type of prediction.
+We also note that the plot is no less consistent with the alarming negative association between sample size and reported accuracy observed by }@Berisha2021\hl{ than with the inverse association expected in the absence of selective reporting.}
 
@@
 
-The study authors focused their recommendations and their own plans for future work mostly on technical improvements and evaluations [@Mariuzzi1997; @Yearwood1997].
+\hl{S}tudy authors focused their recommendations and \hl{future plans} mostly on technical improvements and evaluations\hl{.}
 Urged improvements included full or partial automation of predictor selection [@Mariuzzi1997; @Yearwood1997], similarity learning [@Mariuzzi1997; @Wang2019], and parameter optimization [@Song2006; @Lee2017]; extensions to new data structures [@Lopez2011], data types [@Liang2015; @Verma2015; @Malykh2018], and reasoning systems [@Nicolas2014]; and the use of more advanced model components to improve accuracy or efficiency [@Lowsky2013; @Lee2015; @Liang2015; @Zhang2018].
 Authors also urged validations and independent evaluations using larger or more comprehensive data sets [@Elter2007; @Xu2008; @Verma2015; @Ng2015], using data aggregated from multiple health systems [@Lee2015; @Lee2017; @Tang2021; @Ng2021], and in other care settings or disease contexts [@Song2006; @Zhang2018; @Tang2021; @Ng2021].
-
-Less common were calls to strengthen the connection between the methods and the users.
-Some authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997; @Wang2019].
-Others suggested new uses of their modeling approaches: to measure feature importance [@Wyns2004], to make models more expressive [@Lee2015], and to combine information obtained both from global and from \hl{SCBM}s [@Ng2015].
-Most of this incremental work was indeed carried out in later investigations in our sample.
-That said, we did not find reports of successful implementations of these tools in clinical practice. (Several excluded studies did showcase uses of CBR in practice.)
+\hl{That said, s}ome authors urged the incorporation of (human-derived) domain knowledge into the data or models [@Yearwood1997; @Wang2019]\hl{, and o}thers suggested new uses of their \hl{methods} to measure feature importance [@Wyns2004], to make models more expressive [@Lee2015], and to combine information obtained both from global and from \hl{SCBM}s [@Ng2015].
+\hl{(}Most of this incremental work was indeed carried out \hl{later} in our sample.\hl{)}
 
@@
 Some earlier studies tuned generalized regression models, but more recent studies did not.
+
+\hl{A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
+It would also make like-kind comparisons of performance easier on different tasks using different data sets and expedite validation studies.
+This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)} as well as the limitations of our own comparison of reported performances (Section }\ref{evaluations}\hl{).
+We also believe that this framework could be used to make the diversity of possible approaches more comprehensible to users, enabling them to most appropriately tailor these decisions to their tasks.}
 
@@
 The table contains no original results.
+\hl{Figure }\ref{fig:performance}\hl{ visualizes the relationship between the performances of the proposed methods (not of their comparators) and several aspects of the experiments used to evaluate them.}
 
@@
 
-**TODO: Place vertical separators (` |`) between major columns.**
+**TODO: Place vertical separators between major columns.**
 
@@
 
+```{r performance, fig.align='center', out.width="100%", echo=FALSE, fig.pos='H'}
+#| fig.cap =
+#| "\\hl{Scatterplot of predictive performance versus case count;
+#|  marker size, shape, and color
+#|  encode feature count, data source, and predictive task, respectively.
+#|  (Image included as a separate file.)}"
+# knitr::include_graphics("FigS1.png", error = FALSE)
+knitr::include_graphics("Fig6.png")
+```
+
 # References {-}

### 2025-03-10 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
complete update per reviewer comments + track bbl file

@@
 that small distances between cases yield small changes in prediction; and kernel-based methods that rely directly on pairwise distances computed via inner product, rather than high-dimensional coordinates, extend far beyond their use in CBR.
-The key difference in SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
+Additionally, }\citet{Schoenborn2021}\hl{ survey CBR-powered explanations of opaque models and taxonomize them as \emph{model-agnostic} versus \emph{model-based}, though this coupling of CBR with predictive models is inverted from that of SCBM.
+What distinguishes SCBM is that the retrieved cohort is treated as a component of the model and made available to the user.
 Indeed, retrieval procedures were the most significant algorithmic requirements and accomplishments of CBR }\citep{Kolodner1992, Gierl1998}\hl{ and the retrieved cohorts may be reused for numerous modeling tasks.}
-Thus, \hl{SCBM}s may embody both styles of CBR.
 
@@
 five platforms: PubMed, Web of Science, Academic Search Premier, Google
-Scholar, and MathSciNet. We finalized the PubMed search first, then
-adapted it to the other platforms\hl{.}
+Scholar, and MathSciNet.
+\hl{These platforms were chosen to cover a range of relevant disciplines (biomedicine, mathematics, broader sciences) and types of reports (journal articles, conference proceedings, grey literature.}
+We finalized the PubMed search first, then adapted it to the other
+platforms\hl{.}
 
@@
 
-Figure S1 \textbf{TODO}.
+\hl{Figure }\ref{fig:performance}\hl{ visualizes the relationship between performance and case count, feature count, data source, and predictive task for those experiments that evaluated models using accuracy or AUROC (other performance measures were too rare to allow for useful comparisons).
+While not represented in the plot, results from different models applied to the same data can be inferred from their common case counts, data sources, and tasks.
+The data are too heterogeneous to meaningfully test for trends.
+Performance varies much more between experiments than between models within experiments, and there is no clear dependency on the source(s) of data or on the type of prediction.
+We also note that the plot is no less consistent with the alarming negative association between sample size and reported accuracy observed by }\citet{Berisha2021}\hl{ than with the inverse association expected in the absence of selective reporting.}
 
@@
 
-The study authors focused their recommendations and their own plans for
-future work mostly on technical improvements and evaluations
-\citep{Mariuzzi1997, Yearwood1997}. Urged improvements included full or
-partial automation of predictor selection
+\hl{S}tudy authors focused their recommendations and \hl{future plans}
+mostly on technical improvements and evaluations\hl{.} Urged
+improvements included full or partial automation of predictor selection
 \citep{Mariuzzi1997, Yearwood1997}, similarity learning
@@
 or disease contexts \citep{Song2006, Zhang2018, Tang2021, Ng2021}.
-
-Less common were calls to strengthen the connection between the methods
-and the users. Some authors urged the incorporation of (human-derived)
-domain knowledge into the data or models \citep{Yearwood1997, Wang2019}.
-Others suggested new uses of their modeling approaches: to measure
-feature importance \citep{Wyns2004}, to make models more expressive
-\citep{Lee2015}, and to combine information obtained both from global
-and from \hl{SCBM}s \citep{Ng2015}. Most of this incremental work was
-indeed carried out in later investigations in our sample. That said, we
-did not find reports of successful implementations of these tools in
-clinical practice. (Several excluded studies did showcase uses of CBR in
-practice.)
+\hl{That said, s}ome authors urged the incorporation of (human-derived)
+domain knowledge into the data or models
+\citep{Yearwood1997, Wang2019}\hl{, and o}thers suggested new uses of
+their \hl{methods} to measure feature importance \citep{Wyns2004}, to
+make models more expressive \citep{Lee2015}, and to combine information
+obtained both from global and from \hl{SCBM}s \citep{Ng2015}. \hl{(}Most
+of this incremental work was indeed carried out \hl{later} in our
+sample.\hl{)}
 
@@
 regression models, but more recent studies did not.
+
+\hl{A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
+It would also make like-kind comparisons of performance easier on different tasks using different data sets and expedite validation studies.
+This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)}
+as well as the limitations of our own comparison of reported
+performances (Section \}\ref{evaluations}\hl{).
+We also believe that this framework could be used to make the diversity of possible approaches more comprehensible to users, enabling them to most appropriately tailor these decisions to their tasks.}
 
@@
 metrics, are reproduced here. The table contains no original results.
+\hl{Figure }\ref{fig:performance}\hl{ visualizes the relationship between the performances of the proposed methods (not of their comparators) and several aspects of the experiments used to evaluate them.}
 
@@
 
-\textbf{TODO: Place vertical separators (\texttt{\textbar{}}) between
-major columns.}
+%\textbf{TODO: Place vertical separators between major columns.}
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{Fig6}
+
+}
+
+\caption{\hl{Scatterplot of predictive performance versus case count;
+ marker size, shape, and color
+ encode feature count, data source, and predictive task, respectively.
+ (Image included as a separate file.)}}\label{fig:performance}
+\end{figure}
+
 \renewcommand\refname{References}

### 2025-03-10 scr/synthesis.r

@@ -441,6 +441,7 @@ tab_eval_prop |>
   theme(legend.position = "bottom") ->
   performance_determinants
 print(performance_determinants)
+# for consistent zoom with transpose orientation, set height here to width above
 ggsave(
   here::here("fig/fig-performance.png"), performance_determinants,
   width = 13, height = 8

### 2025-03-14 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
make final changes before resubmission

@@
 
-\usepackage{soul}
+%\usepackage{soul}
+\newcommand{\hl}[1]{#1}
 
@@
 later agreed to discard these results due to the lack of reproducibility
-of the search. We organized the remaining results into a public Zotero
-collection with one subcollection for each database. We then imported
-the pooled results to Covidence, which merged some duplicate entries.
-
+of the search.
 We repeated the search on 2024 January 24 on the PubMed and Web of
@@
 
-\subsection{\texorpdfstring{Rationales\hl{ for methods}}{Rationales}}\label{rationales}
-
-The included studies hypothesized, asserted, or assumed several benefits
-of \hl{SCBM} specific to clinical and medical settings and advantages
-over other modeling approaches. The most common was that the restriction
-to similar or relevant past cases would improve predictive performance
-for the index case \citep{Mariuzzi1997, Liang2015, Ng2015, Lee2017}. In
-particular, \citet{Lee2015} hypothesized and confirmed that the value of
-each past case would be positively related to its similarity to the
-index case, an assumption built in to the weighting schemes of other
-approaches. \citet{Lowsky2013} made a different case, that the fewer
-parametric assumptions and complications of a CBR-style model would
-allow for greater accuracy. Additionally, several investigators asserted
-that the use of localized cohorts befit the clinical focus on the
-individual patient rather than the population, without reference to
-performance \citep{Song2006, Xu2008}.
-
-In an interesting contrast, \citet{Tang2021} and \citet{Ng2021} argued
-that their localized approach using the larger and more heterogeneous
-populations covered by EHR-derived data could better capture the messy
-and diverse lessons of everyday practice, as a counterpart to guidelines
-based on randomized controlled trials. Their emphasis on the enabling
-role of EHRs to power methods well-suited to large, structured data
-repositories was shared by several others
-\citep{CampilloGimenez2013, Nicolas2014, Verma2015, Lee2017}.
-\citet{Verma2015} and \citet{Zhang2018} additionally pointed out that
-\hl{SCBM}s are adaptable to noise in the data, variation in patterns of
-missingness, and (harkening to the last step of the R4 cycle) addition
-of new cases with known outcomes to the corpus. These properties, they
-said, tend to be more difficult for whole-population models to handle.
-
-The other frequent advantage attributed to \hl{SCBM} was
-interpretability. \citet{Elter2007} and \citet{Nicolas2014} emphasized
-the value of the detailed past cases, available to the user, on which
-predictions are based. This ``self-explanation capability'' made outputs
-more intelligible to physicians in the \hl{CDSS} setting. \citet{Ng2015}
-additionally pointed out that localized GLMs yield localized effect
-estimates, which may help investigators identify individually relevant
-risk factors. \citet{Liu2022} took this idea further, systematically
-comparing regression coefficients for specific risk factors across
-dozens of diagnostic subpopulations. \citet{Doborjeh2022} used localized
-time series models to identify environmental changes associated with
-increased stroke risk.
-
-The remaining coded rationales were for augmentations or hybridizations
-of then-conventional CBR, most of which defended the use of other tools
-to improve performance via cohort selection
-\citep{CampilloGimenez2013, Nicolas2014, Vilhena2016} or to make models
-and outputs more interpretable \citep{Lopez2011, Wang2019}.
-\citet{Liang2015} argued for simultaneous optimization of cohort and
-feature selection with model parameterization; their TWNFI approach
-combines localization with regularization.
-
-\subsection{\texorpdfstring{Challenges\hl{ with methods}}{Challenges}}\label{challenges}
+\subsection{\texorpdfstring{Challenges\hl{ faced}}{Challenges}}\label{challenges}
 
@@
 discussions between the reviewing authors during full-text review to
-better specify an agreed-upon meaning. We will discuss the different
+better specify an agreed-upon meaning. \hl{We }discuss the different
 domains, terminologies, and reporting issues of the sample in the
@@
 disease \citep{Ng2015, Tang2021, Ng2021}.
-
-These programs used varying terminology for common concepts, and no
-common term is in use for what we here refer to as
-\hl{similarity cohort--based modeling}; authors described their
-approaches as ``targeted prognosis'' \citep{Mariuzzi1997},
-``transductive inference'' \citep{Song2006}, ``personalized decision
-support'' \citep{Lee2015}, ``personalized (predictive) models''
-\citetext{\citealp[in contrast to ``local
-models'']{Liang2015}; \citealp{Ng2015}; \citealp{Wang2019}; \citealp{Ma2020}; \citealp{Liu2022}; \citealp{Doborjeh2022}},
-and ``precision cohort'' analysis \citep{Wang2019, Tang2021, Ng2021}.
-\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)}
-We find uses of ``targeted'', ``personalized'', and ``precision''
-generic and imprecise, while transductive inference is an established
-term for a broader set of methods in ML.
-\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves, so we use \emph{similarity cohort--based models} to refer to this specific approach.}
 
@@
 three largely independent choices
-\hl{(Figure }\ref{fig:framework}\hl{})\}: (1) how to measure the
+\hl{(Figure }\ref{fig:framework}\hl{)}: (1) how to measure the
 similarity between cases, (2) how to retrieve a cohort of cases similar
@@
 
-\hl{A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
+\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work (Ravi et al.).
+A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
 It would also make like-kind comparisons of performance easier on different tasks using different data sets and expedite validation studies.
-This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)}
+This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)
 as well as the limitations of our own comparison of reported
-performances (Section \}\ref{evaluations}\hl{).
+performances (Section }\ref{evaluations}\hl{).
 We also believe that this framework could be used to make the diversity of possible approaches more comprehensible to users, enabling them to most appropriately tailor these decisions to their tasks.}
@@
 
-\hl{Most studies were premised on the potential predictive value of localized models (Section }\ref{rationales}\hl{), but not all experiments affirmed this premise (Section }\ref{evaluations}\hl{).
+\hl{Most studies were premised on the potential predictive value of localized models (Appendix), but not all experiments affirmed this premise (Section }\ref{evaluations}\hl{).
 Those that did mostly used DML to improve retrieval and sometimes iteratively optimized the metric and the model.
@@
 Future work should compare competitive alternatives to assess their performance and other trade-offs.
-This work would be facilitated by standardization and modularization using a framework like that outlined above.
+This work would be aided by standardization and modularization using a framework like that outlined above.
 For example, }\citet{CampilloGimenez2013}\hl{ and }\citet{Liu2022}\hl{ use a modular design to toggle multiple choices and obtain neatly nested comparisons.}
@@
 
-Two anonymous reviewers provided valuable feedback on a previous version
+\hl{Four} anonymous reviewers provided valuable feedback on \hl{previous versions}
 of this manuscript.
@@
 
-The search will have been designed to recover studies of the kind
+The search \hl{was} designed to recover studies of the kind
 reviewed by the review papers from which we obtained our ``seed set''.
-To validate the final search design, we will determine how many of the
+To validate the final search design, we \hl{determined} how many of the
 papers in this seed set that are indexed by PubMed are actually
@@
 Lee (2017), which focuses on patient similarity--based prediction models
-on health data. The proportion of each PubMed-indexed seed set that is
-recovered from our PubMed search provides a rough and optimistic yet
+on health data. The proportion of each PubMed-indexed seed set \hl{recovered} from our PubMed search provides a rough and optimistic yet
 useful estimate of the proportion of the relevant literature that our
@@
 
-Once we have finalized the search as a logical pattern, we will take the
+Once we \hl{finalized} the search as a logical pattern, we \hl{took} the
 following steps to generate the sample/corpus of literature that will be
@@
 \item
-  The logical pattern will be converted to a search string using the
+  The logical pattern \hl{was} converted to a search string using the
   syntax appropriate to each database in our search strategy. These
-  include PubMed (already done as part of the search design), Web of
+  include\hl{d} PubMed (already done as part of the search design), Web of
   Science, Academic Search Premier/Elite, and Mathematical Reviews.
 \item
-  The search will be conducted on each database and the results
+  The search \hl{was} conducted on each database and the results
   organized into a Zotero collection, with one subcollection for each
@@
 \item
-  Duplicate results will be identified and merged. (A result obtained
+  Duplicate results \hl{were} identified and merged. (A result obtained
   from multiple databases should have only one Zotero entry but should
@@
 
-\subsection*{Terminology}\label{terminology}
+\subsection*{\hl{Terminology}}\label{terminology}
 \addcontentsline{toc}{subsection}{Terminology}
+
+These programs used varying terminology for common concepts, and no
+common term is in use for what we here refer to as
+\hl{similarity cohort--based modeling}; authors described their
+approaches as ``targeted prognosis'' \citep{Mariuzzi1997},
+``transductive inference'' \citep{Song2006}, ``personalized decision
+support'' \citep{Lee2015}, ``personalized (predictive) models''
+\citetext{\citealp[in contrast to ``local
+models'']{Liang2015}; \citealp{Ng2015}; \citealp{Wang2019}; \citealp{Ma2020}; \citealp{Liu2022}; \citealp{Doborjeh2022}},
+and ``precision cohort'' analysis \citep{Wang2019, Tang2021, Ng2021}.
+\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)}
+We find uses of ``targeted'', ``personalized'', and ``precision''
+generic and imprecise, while transductive inference is an established
+term for a broader set of methods in ML.
+\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves.}
 
@@
 \end{figure}
+
+\subsection*{\texorpdfstring{Rationales\hl{ for proposed methods}}{Rationales}}\label{rationales}
+
+\hl{Our review was driven by an interest in the potential of SCBM toward interpretable ML.
+While not essential to the main text, we were therefore interested in understanding what needs motivated the various proposed methods we found.}
+
+The included studies hypothesized, asserted, or assumed several benefits
+of \hl{SCBM} specific to clinical and medical settings and advantages
+over other modeling approaches. The most common was that the restriction
+to similar or relevant past cases would improve predictive performance
+for the index case \citep{Mariuzzi1997, Liang2015, Ng2015, Lee2017}. In
+particular, \citet{Lee2015} hypothesized and confirmed that the value of
+each past case would be positively related to its similarity to the
+index case, an assumption built in to the weighting schemes of other
+approaches. \citet{Lowsky2013} made a different case, that the fewer
+parametric assumptions and complications of a CBR-style model would
+allow for greater accuracy. Additionally, several investigators asserted
+that the use of localized cohorts befit the clinical focus on the
+individual patient rather than the population, without reference to
+performance \citep{Song2006, Xu2008}.
+
+In an interesting contrast, \citet{Tang2021} and \citet{Ng2021} argued
+that their localized approach using the larger and more heterogeneous
+populations covered by EHR-derived data could better capture the messy
+and diverse lessons of everyday practice, as a counterpart to guidelines
+based on randomized controlled trials. Their emphasis on the enabling
+role of EHRs to power methods well-suited to large, structured data
+repositories was shared by several others
+\citep{CampilloGimenez2013, Nicolas2014, Verma2015, Lee2017}.
+\citet{Verma2015} and \citet{Zhang2018} additionally pointed out that
+\hl{SCBM}s are adaptable to noise in the data, variation in patterns of
+missingness, and (harkening to the last step of the R4 cycle) addition
+of new cases with known outcomes to the corpus. These properties, they
+said, tend to be more difficult for whole-population models to handle.
+
+The other frequent advantage attributed to \hl{SCBM} was
+interpretability. \citet{Elter2007} and \citet{Nicolas2014} emphasized
+the value of the detailed past cases, available to the user, on which
+predictions are based. This ``self-explanation capability'' made outputs
+more intelligible to physicians in the \hl{CDSS} setting. \citet{Ng2015}
+additionally pointed out that localized GLMs yield localized effect
+estimates, which may help investigators identify individually relevant
+risk factors. \citet{Liu2022} took this idea further, systematically
+comparing regression coefficients for specific risk factors across
+dozens of diagnostic subpopulations. \citet{Doborjeh2022} used localized
+time series models to identify environmental changes associated with
+increased stroke risk.
+
+The remaining coded rationales were for augmentations or hybridizations
+of then-conventional CBR, most of which defended the use of other tools
+to improve performance via cohort selection
+\citep{CampilloGimenez2013, Nicolas2014, Vilhena2016} or to make models
+and outputs more interpretable \citep{Lopez2011, Wang2019}.
+\citet{Liang2015} argued for simultaneous optimization of cohort and
+feature selection with model parameterization; their TWNFI approach
+combines localization with regularization.
 

### 2025-03-14 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
put latex revisions into rmarkdown

@@
 We tentatively excluded results from Google Scholar because they were missing abstracts, and later agreed to discard these results due to the lack of reproducibility of the search.
-We organized the remaining results into a public Zotero collection with one subcollection for each database.
-We then imported the pooled results to Covidence, which merged some duplicate entries.
 
@@
 
-## Rationales\hl{ for methods}
-
-The included studies hypothesized, asserted, or assumed several benefits of \hl{SCBM} specific to clinical and medical settings and advantages over other modeling approaches.
-The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015; @Ng2015; @Lee2017].
-In particular, @Lee2015 hypothesized and confirmed that the value of each past case would be positively related to its similarity to the index case, an assumption built in to the weighting schemes of other approaches.
-@Lowsky2013 made a different case, that the fewer parametric assumptions and complications of a CBR-style model would allow for greater accuracy.
-Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to performance [@Song2006; @Xu2008].
-
-In an interesting contrast, @Tang2021 and @Ng2021 argued that their localized approach using the larger and more heterogeneous populations covered by EHR-derived data could better capture the messy and diverse lessons of everyday practice, as a counterpart to guidelines based on randomized controlled trials.
-Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2013; @Nicolas2014; @Verma2015; @Lee2017].
-@Verma2015 and @Zhang2018 additionally pointed out that \hl{SCBM}s are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
-These properties, they said, tend to be more difficult for whole-population models to handle.
-
-The other frequent advantage attributed to \hl{SCBM} was interpretability.
-@Elter2007 and @Nicolas2014 emphasized the value of the detailed past cases, available to the user, on which predictions are based.
-This "self-explanation capability" made outputs more intelligible to physicians in the \hl{CDSS} setting.
-@Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
-@Liu2022 took this idea further, systematically comparing regression coefficients for specific risk factors across dozens of diagnostic subpopulations.
-@Doborjeh2022 used localized time series models to identify environmental changes associated with increased stroke risk.
-
-The remaining coded rationales were for augmentations or hybridizations of then-conventional CBR, most of which defended the use of other tools to improve performance via cohort selection [@CampilloGimenez2013; @Nicolas2014; @Vilhena2016] or to make models and outputs more interpretable [@Lopez2011; @Wang2019].
-@Liang2015 argued for simultaneous optimization of cohort and feature selection with model parameterization; their TWNFI approach combines localization with regularization.
-
 ## Challenges\hl{ with methods}
@@
 Regarding interpretation of criteria, some wording of the criteria was adjusted following discussions between the reviewing authors during full-text review to better specify an agreed-upon meaning.
-We will discuss the different domains, terminologies, and reporting issues of the sample in the remainder of this section.
+\hl{We }discuss the different domains, terminologies, and reporting issues of the sample in the remainder of this section.
 
@@
 
-These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as \hl{similarity cohort--based modeling}; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
-\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)} We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
-\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves, so we use \emph{similarity cohort--based models} to refer to this specific approach.}
-
 ## Synthesis
@@
 In this section, we focus on the diversity of assemblies.
-These depend on} three largely independent choices \hl{(Figure }\ref{fig:framework}\hl{})}: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
+These depend on} three largely independent choices \hl{(Figure }\ref{fig:framework}\hl{)}: (1) how to measure the similarity between cases, (2) how to retrieve a cohort of cases similar to an index case, and (3) how to generate a prediction or other statistical insight for the index case from the similarity cohort.
 
@@
 
-\hl{A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
+\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work [@Ravi2022].
+A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
 It would also make like-kind comparisons of performance easier on different tasks using different data sets and expedite validation studies.
-This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)} as well as the limitations of our own comparison of reported performances (Section }\ref{evaluations}\hl{).
+This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{) as well as the limitations of our own comparison of reported performances (Section }\ref{evaluations}\hl{).
 We also believe that this framework could be used to make the diversity of possible approaches more comprehensible to users, enabling them to most appropriately tailor these decisions to their tasks.}
@@
 Future work should compare competitive alternatives to assess their performance and other trade-offs.
-This work would be facilitated by standardization and modularization using a framework like that outlined above.
+This work would be aided by standardization and modularization using a framework like that outlined above.
 For example, }@CampilloGimenez2013\hl{ and }@Liu2022\hl{ use a modular design to toggle multiple choices and obtain neatly nested comparisons.}
@@
 
-Two anonymous reviewers provided valuable feedback on a previous version of this manuscript.
+\hl{Four} anonymous reviewers provided valuable feedback on \hl{previous versions} of this manuscript.
 
@@
 
-The search will have been designed to recover studies of the kind reviewed by the review papers from which we obtained our “seed set”.
-To validate the final search design, we will determine how many of the papers in this seed set that are indexed by PubMed are actually recovered by our search.
+The search \hl{was} designed to recover studies of the kind reviewed by the review papers from which we obtained our “seed set”.
+To validate the final search design, we \hl{determined} how many of the papers in this seed set that are indexed by PubMed are actually recovered by our search.
 In most cases, the focus of a review paper is different from ours, so we will only perform this validation test on the seed set obtained from two review papers that are (a) closest in focus to ours and (b) use terminology associated with the two distinct sub-literatures relevant to our focus: Choudhury & Begum (2016), which focuses on case-based reasoning in medicine, and Sharafoddini, Dubin, & Lee (2017), which focuses on patient similarity–based prediction models on health data.
-The proportion of each PubMed-indexed seed set that is recovered from our PubMed search provides a rough and optimistic yet useful estimate of the proportion of the relevant literature that our full search strategy will recover.
-
-Once we have finalized the search as a logical pattern, we will take the following steps to generate the sample/corpus of literature that will be the starting point for our selection process.
+The proportion of each PubMed-indexed seed set \hl{recovered} from our PubMed search provides a rough and optimistic yet useful estimate of the proportion of the relevant literature that our full search strategy will recover.
+
+Once we \hl{finalized} the search as a logical pattern, we \hl{took} the following steps to generate the sample/corpus of literature that \hl{was} the starting point for our selection process.
 
@@
 The logical pattern will be converted to a search string using the syntax appropriate to each database in our search strategy.
-These include PubMed (already done as part of the search design), Web of Science, Academic Search Premier/Elite, and Mathematical Reviews.
-2. The search will be conducted on each database and the results organized into a Zotero collection, with one subcollection for each database.
+These include\hl{d} PubMed (already done as part of the search design), Web of Science, Academic Search Premier/Elite, and Mathematical Reviews.
+2. The search \hl{was} conducted on each database and the results organized into a Zotero collection, with one subcollection for each database.
 3.
-Duplicate results will be identified and merged.
+Duplicate results \hl{were} identified and merged.
 (A result obtained from multiple databases should have only one Zotero entry but should be filed under the subcollection for each database in which it was found.)
@@
 ## Terminology {-}
+
+These programs used varying terminology for common concepts, and no common term is in use for what we here refer to as \hl{similarity cohort--based modeling}; authors described their approaches as "targeted prognosis" [@Mariuzzi1997], "transductive inference" [@Song2006], "personalized decision support" [@Lee2015], "personalized (predictive) models" [@Liang2015, in contrast to "local models"; @Ng2015; @Wang2019; @Ma2020; @Liu2022; @Doborjeh2022], and "precision cohort" analysis [@Wang2019; @Tang2021; @Ng2021].
+\hl{(Interestingly, none of the articles described their methods as ``explainable'' ML.)} We find uses of "targeted", "personalized", and "precision" generic and imprecise, while transductive inference is an established term for a broader set of methods in ML.
+\hl{While \emph{local} naturally contrasts with \emph{global}, the overloaded term \emph{localized models} stands to cause more confusion than it resolves.}
 
@@
 ```
+
+## Rationales\hl{ for methods}
+
+\hl{Our review was driven by an interest in the potential of SCBM toward interpretable ML.
+While not essential to the main text, we were therefore interested in understanding what needs motivated the various proposed methods we found.}
+
+The included studies hypothesized, asserted, or assumed several benefits of \hl{SCBM} specific to clinical and medical settings and advantages over other modeling approaches.
+The most common was that the restriction to similar or relevant past cases would improve predictive performance for the index case [@Mariuzzi1997; @Liang2015; @Ng2015; @Lee2017].
+In particular, @Lee2015 hypothesized and confirmed that the value of each past case would be positively related to its similarity to the index case, an assumption built in to the weighting schemes of other approaches.
+@Lowsky2013 made a different case, that the fewer parametric assumptions and complications of a CBR-style model would allow for greater accuracy.
+Additionally, several investigators asserted that the use of localized cohorts befit the clinical focus on the individual patient rather than the population, without reference to performance [@Song2006; @Xu2008].
+
+In an interesting contrast, @Tang2021 and @Ng2021 argued that their localized approach using the larger and more heterogeneous populations covered by EHR-derived data could better capture the messy and diverse lessons of everyday practice, as a counterpart to guidelines based on randomized controlled trials.
+Their emphasis on the enabling role of EHRs to power methods well-suited to large, structured data repositories was shared by several others [@CampilloGimenez2013; @Nicolas2014; @Verma2015; @Lee2017].
+@Verma2015 and @Zhang2018 additionally pointed out that \hl{SCBM}s are adaptable to noise in the data, variation in patterns of missingness, and (harkening to the last step of the R4 cycle) addition of new cases with known outcomes to the corpus.
+These properties, they said, tend to be more difficult for whole-population models to handle.
+
+The other frequent advantage attributed to \hl{SCBM} was interpretability.
+@Elter2007 and @Nicolas2014 emphasized the value of the detailed past cases, available to the user, on which predictions are based.
+This "self-explanation capability" made outputs more intelligible to physicians in the \hl{CDSS} setting.
+@Ng2015 additionally pointed out that localized GLMs yield localized effect estimates, which may help investigators identify individually relevant risk factors.
+@Liu2022 took this idea further, systematically comparing regression coefficients for specific risk factors across dozens of diagnostic subpopulations.
+@Doborjeh2022 used localized time series models to identify environmental changes associated with increased stroke risk.
+
+The remaining coded rationales were for augmentations or hybridizations of then-conventional CBR, most of which defended the use of other tools to improve performance via cohort selection [@CampilloGimenez2013; @Nicolas2014; @Vilhena2016] or to make models and outputs more interpretable [@Lopez2011; @Wang2019].
+@Liang2015 argued for simultaneous optimization of cohort and feature selection with model parameterization; their TWNFI approach combines localization with regularization.
 

### 2025-03-14 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
put latex revisions into rmarkdown

@@
 
-%\usepackage{soul}
-\newcommand{\hl}[1]{#1}
+\usepackage{soul}
 
@@
 of the search.
+
 We repeated the search on 2024 January 24 on the PubMed and Web of
@@
 
-\subsection{\texorpdfstring{Challenges\hl{ faced}}{Challenges}}\label{challenges}
+\subsection{\texorpdfstring{Challenges\hl{ with methods}}{Challenges}}\label{challenges}
 
@@
 
-\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work (Ravi et al.).
+\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work [@Ravi2022].
 A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
 It would also make like-kind comparisons of performance easier on different tasks using different data sets and expedite validation studies.
-This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{)
-as well as the limitations of our own comparison of reported
-performances (Section }\ref{evaluations}\hl{).
+This would address the core technical needs expressed in our sample (Section }\ref{identified-needs}\hl{) as well as the limitations of our own comparison of reported performances (Section }\ref{evaluations}\hl{).
 We also believe that this framework could be used to make the diversity of possible approaches more comprehensible to users, enabling them to most appropriately tailor these decisions to their tasks.}
@@
 
-\hl{Four} anonymous reviewers provided valuable feedback on \hl{previous versions}
-of this manuscript.
+\hl{Four} anonymous reviewers provided valuable feedback on
+\hl{previous versions} of this manuscript.
 
@@
 
-The search \hl{was} designed to recover studies of the kind
-reviewed by the review papers from which we obtained our ``seed set''.
-To validate the final search design, we \hl{determined} how many of the
-papers in this seed set that are indexed by PubMed are actually
-recovered by our search. In most cases, the focus of a review paper is
-different from ours, so we will only perform this validation test on the
-seed set obtained from two review papers that are (a) closest in focus
-to ours and (b) use terminology associated with the two distinct
-sub-literatures relevant to our focus: Choudhury \& Begum (2016), which
-focuses on case-based reasoning in medicine, and Sharafoddini, Dubin, \&
-Lee (2017), which focuses on patient similarity--based prediction models
-on health data. The proportion of each PubMed-indexed seed set \hl{recovered} from our PubMed search provides a rough and optimistic yet
-useful estimate of the proportion of the relevant literature that our
-full search strategy will recover.
+The search \hl{was} designed to recover studies of the kind reviewed by
+the review papers from which we obtained our ``seed set''. To validate
+the final search design, we \hl{determined} how many of the papers in
+this seed set that are indexed by PubMed are actually recovered by our
+search. In most cases, the focus of a review paper is different from
+ours, so we will only perform this validation test on the seed set
+obtained from two review papers that are (a) closest in focus to ours
+and (b) use terminology associated with the two distinct sub-literatures
+relevant to our focus: Choudhury \& Begum (2016), which focuses on
+case-based reasoning in medicine, and Sharafoddini, Dubin, \& Lee
+(2017), which focuses on patient similarity--based prediction models on
+health data. The proportion of each PubMed-indexed seed set
+\hl{recovered} from our PubMed search provides a rough and optimistic
+yet useful estimate of the proportion of the relevant literature that
+our full search strategy will recover.
 
 Once we \hl{finalized} the search as a logical pattern, we \hl{took} the
-following steps to generate the sample/corpus of literature that will be
-the starting point for our selection process.
+following steps to generate the sample/corpus of literature that
+\hl{was} the starting point for our selection process.
 
@@
 \item
-  The logical pattern \hl{was} converted to a search string using the
+  The logical pattern will be converted to a search string using the
   syntax appropriate to each database in our search strategy. These
-  include\hl{d} PubMed (already done as part of the search design), Web of
-  Science, Academic Search Premier/Elite, and Mathematical Reviews.
+  include\hl{d} PubMed (already done as part of the search design), Web
+  of Science, Academic Search Premier/Elite, and Mathematical Reviews.
 \item
@@
 
-\subsection*{\hl{Terminology}}\label{terminology}
+\subsection*{Terminology}\label{terminology}
 \addcontentsline{toc}{subsection}{Terminology}
@@
 
-\subsection*{\texorpdfstring{Rationales\hl{ for proposed methods}}{Rationales}}\label{rationales}
+\subsection{\texorpdfstring{Rationales\hl{ for methods}}{Rationales}}\label{rationales}
 
@@
 
-%\textbf{TODO: Place vertical separators between major columns.}
+\textbf{TODO: Place vertical separators between major columns.}
 

### 2025-03-14 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.Rmd
fix reference formatting

@@
 
-## Challenges\hl{ with methods}
+## Challenges\hl{ faced}
 
@@
 
-\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work [@Ravi2022].
+\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work }[@Ravi2022]\hl{.
 A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
@@
 
-## Rationales\hl{ for methods}
+## Rationales\hl{ for proposed methods}
 

### 2025-03-14 docs/Localized-Modeling-BMCAI/localized-modeling-dai-revision.tex
fix reference formatting

@@
 
-\subsection{\texorpdfstring{Challenges\hl{ with methods}}{Challenges}}\label{challenges}
+\subsection{\texorpdfstring{Challenges\hl{ faced}}{Challenges}}\label{challenges}
 
@@
 
-\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work [@Ravi2022].
+\hl{As in the case of data and, recently, of ML models, having a standardized general framework facilitates the recombination of ideas and prevents duplicate effort as well as supporting more comprehensive and efficient work }\citep{Ravi2022}\hl{.
 A SCBM implementation based on this framework would help investigators systematically compare predictions and other output obtained using different tools for cohort retrieval and predictive modeling, as well as tune the parameters governing these tools.
@@
 
-\subsection{\texorpdfstring{Rationales\hl{ for methods}}{Rationales}}\label{rationales}
+\subsection{\texorpdfstring{Rationales\hl{ for proposed methods}}{Rationales}}\label{rationales}
 
